### 2015-11-19 data_cleaning.tex
added files

\documentclass{article}
\usepackage{color}
\usepackage{listings}	
\usepackage{enumerate}
\usepackage{layouts}
\usepackage{times}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{url}
\usepackage{xspace}
\usepackage{tikz}

%\author{Authors}

\title{On tabular constraint learning}
\begin{document}
\maketitle
\vspace{-15pt}
\section{Headers}
Can be in columns or in row, see examples. Needs to be inferred using type information.
\vspace{-10pt}
\section{General Algorithm: Infer-Learn-Suggest -- ILS }
\paragraph{Infer} Datatypes, headers, i.e. variable names, and table layouts, borders of tables, the number of tables etc
\vspace{-10pt}
\paragraph{Learn} On each level, i.e., table-wise or column-row-wise learn applicable constraints
\vspace{-10pt}
\paragraph{Suggest} If there is a row or a column with missing data (not-completed cells), then suggest values satisfying the constraints
\vspace{-10pt}
\section*{Types of Constraints}
Most of the constraints I have seen are numeric (finance, accounting) and occasionally strings, also emails and dates are present but their quantity is negligible.

Types of constraints exact or noisy(way harder)
\begin{itemize}
  \item Ascending/Descending (all types)
  \item Ordered by (all types)
  \item All different (all types, similar to the key detection)
  \item Arithmetics (only for number type + possible ranges over two, N, all columns-rows or their consecutive subset)
  \item Concatenation (strings + any other type)
  \item Common substring? 
  \item Column row-inclusion, sort of FD, all values of a row or a column are present in the other row or a column
  \item Aggregates? Counting 
  \item Key -- foreign or primary
\end{itemize}
\end{document}

### 2015-11-19 z.sh

@@ -1,0 +1,2 @@
+pdflatex data_cleaning.tex
+evince data_cleaning.pdf&

### 2015-12-08 Notes/Description.tex
Added demo example and description

\documentclass[a4paper]{article}

% Essentials
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage[table]{xcolor}
%\usepackage{latexsym,graphicx,graphics,epsfig}

% Hyperlinks
\hypersetup{
    colorlinks,
    linkcolor={red},
    citecolor={blue},
    urlcolor={blue}
}

% Optional but very useful
\usepackage[tmargin=3.5cm, bmargin=3.5cm, lmargin=3cm, rmargin=3cm]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{float}
\usepackage{enumitem}
\usepackage{tabularx,booktabs,multirow}
\usepackage{longtable}
\usepackage{url}
\usepackage{enumitem}
\usepackage{todonotes}
\usepackage{caption}

% Title
\title{Excel Constraints}
\date{\today}

% Indentation
\setlength\parindent{0pt}

\begin{document}

\maketitle

% Content

\section{Demo example} % (fold)
\label{sec:demo_example}
The annotated example demonstrates the most used constraints in Excel which we aim to identify automatically.

% section demo_example (end)
\begin{description}
	\item[Aggregates] Aggregate functions like sum, max, average, ...
	\item[Conditional aggregates] Aggregates that use a filter on their input
	\item[Series] Ranges of integer numbers, ascending or descending
	\item[Lookups] Exact or fuzzy lookups that use a key to find a corresponding value
	\item[Ranks] Uses values to determine an order
	\item[Structural constraints] Foreign keys to test value consistency between ranges
	\item[Previous] Uses row values and the previous value to compute the current value
\end{description}

\begin{figure}
  \centering
    \includegraphics[width=1\linewidth]{../Demo.png}
  \label{fig:demo}
\end{figure}

\end{document}

### 2015-12-09 Notes/Description.tex
Updated description

@@
 
-\section{Demo example} % (fold)
+\section{Use cases} % (fold)
+\label{sec:use_cases}
+There are different ways that learned constraints could be used in practice:
+
+\begin{itemize}
+	\item Find mistakes
+	\item Suggest a formula for a specific field
+	\item Suggest a next value
+	\item Find structure and functions in a plain text sheet (such as a csv)
+\end{itemize}
+
+% section use_cases (end)
+
+\section{Desired output} % (fold)
 \label{sec:demo_example}
-The annotated example demonstrates the most used constraints in Excel which we aim to identify automatically.
+The annotated example (fig.~\ref{fig:demo}) demonstrates the most used constraints in Excel which we aim to identify automatically.
 
@@
 \begin{description}
-	\item[Aggregates] Aggregate functions like sum, max, average, ...
+	\item[Aggregates] Aggregate functions like sum, max, average, ... that reduce a range to a value
 	\item[Conditional aggregates] Aggregates that use a filter on their input
-	\item[Series] Ranges of integer numbers, ascending or descending
+	\item[Series] Ranges of integer numbers, ascending or descending (a special case of a permutation)
 	\item[Lookups] Exact or fuzzy lookups that use a key to find a corresponding value
-	\item[Ranks] Uses values to determine an order
+	\item[Ranks] Uses a range and a value to determine an order over elements
 	\item[Structural constraints] Foreign keys to test value consistency between ranges
@@
 
+Generally these constraints can be subdivided into:
+
+\begin{itemize}
+	\item Row constraints
+	\item Column constraints
+	\item Inter-table constraints
+	\item Nested constraints
+\end{itemize}
+
+% section types_of_constraints (end)
+
+\section{Approach} % (fold)
+\label{sec:approach}
+
+Currently we are considering a ModelSeeker like approach were ranges and values are generated and constraints tested upon them.
+One option would be to use constraint-specific generators to avoid the explosion of the search space.
+Meta-information would be used to find the most specific constraints.
+A heuristic could use various information to determine which learned constraints are useful.
+
+% section approach (end)
+
 \begin{figure}
@@
     \includegraphics[width=1\linewidth]{../Demo.png}
+  \caption{Demo example}
   \label{fig:demo}

### 2016-01-14 Notes/Description.tex
Added explanation, integrating notes and meetings + adding table ids on demo

@@
 \usepackage[table]{xcolor}
-%\usepackage{latexsym,graphicx,graphics,epsfig}
 
@@
 
+% \sam comment command
+\newcommand{\sam}[1]{{\color{red}(#1)}}
+
+% \range* commands
+\newcommand{\range}[3]{#1[#2,#3]}
+\newcommand{\rangeto}[2]{#1{:}#2}
+\newcommand{\rangeall}{:}
+
 % Title
@@
 	\item Suggest a formula for a specific field
-	\item Suggest a next value
+	\item Suggest a next value / autosuggest
 	\item Find structure and functions in a plain text sheet (such as a csv)
@@
 
+Notable omission so far is the \textit{if-then-else} constraint. Whether it will be included and if how this will be done exactly is still to be determined.
+
 Generally these constraints can be subdivided into:
@@
 \end{itemize}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=1\linewidth]{../Demo.png}
+  \caption{Demo example}
+  \label{fig:demo}
+\end{figure}
 
@@
 
-% section approach (end)
+\subsection{Group discovery}
+Vectors consist of columns or rows that are consistent in their type (e.g. the Salesperson column or the Average row).
+Groups are sets of vectors that are consistent in their type and size.
+If only neighboring vectors are allowed to be added to a group, the group consists of a range.
+The groups partition the data and for every group subgroups may be defined that can overlap.
 
-\begin{figure}
-  \centering
-    \includegraphics[width=1\linewidth]{../Demo.png}
-  \caption{Demo example}
-  \label{fig:demo}
-\end{figure}
+Groups describe subsets of data and can be used to form an $m \times n$ matrix with a specific type.
+Constraints can then be defined in terms of these matrices.
+
+Our current approach would be to precompute all groups and subgroups and store metadata about their content (e.g. type, size) and hierarchy (e.g. child and parent groups).
+
+An important consideration is that some data could be considered input while other data might have to be computed.
+This influences the search algorithm.
+
+\subsection{Finding constraints}
+Currently we see three approaches to finding constraints:
+
+\begin{description}
+	\item[Constraint view] In this approach the search for constraints is structured foremost from the point of view of the constraints.
+	This means that for every type of constraint groups and transformations  are searched such that they fulfill the constraint.
+	\item[Vector view] From an Excel point of view constraints are typically functional.
+	Constraint are either a property of a vector (e.g. series) or describe how a specific vector can be calculated in terms of other vectors (e.g. sum).
+	This lead us to the vector view in which for every vector constraints are searched that hold on the vector or describe how to calculate it.
+	\item[Set view] The last approach is to generate sets of groups and test various constraints on the set.
+	This approach seems simpler but also more naive than the previous approaches.
+\end{description}
+
+In order to choose what groups may occur in what constraints and what combinations are valid, metadata about the constraints is also required (e.g. consistency in sizes, types).
+Additionally the system needs information about the relation between different constraints to remove redundancies.
+
+\subsection{Transformations}
+So far the concept of transformations is only defined losely.
+The idea is that some constraints cannot act on groups without applying a transformation on one or multiple groups first (e.g. the \textit{previous} transformation).
+Therefore, there should be a catalog of transformations which might be applied to groups.
+As with constraints, metadata about the transformations should be given.
+It is yet a bit unclear what the difference is between transformations and constraints.
+On one hand, one might see some constraints as transformations and vice versa.
+One could generalize to \textit{functions} that map input to output.
+On the other hand, however, a property of transformations could be that they preserve their input size and distantiate them from constraint / functions in that way.
+Considering transformations separately might avoid some of the combinatorial explosion that arbitrary function chaining might provide.
+
+One notable decision about transformations in general will be if the system supports types of backwards reasoning.
+An easy example of this would be adding a constant to the input or multiplying it with a constant.
+
+\section{Notation}
+\subsection{Ranges}
+Every table $T$ is considered like a matrix.
+Indexing allows to specify a subset of the table, for example, $\range{T}{r}{c}$ represents the cell at row~$r$ and column~$c$.
+The colon $:$ is used to represent ranges.
+For example, $\range{T}{\rangeall}{c}$ selects all cells in column~$c$ and $\range{T}{\rangeto{r_1}{r_2}}{\rangeall}$ selects all cells in the rows $r_1$ to $r_2$.
+Disjoint columns can be specified by using $\range{T}{\rangeall}{\{c_1, ..., c_n\}}$.
+
+\subsection{Example}
+\paragraph{Aggregates}
+There are two types of aggregates in the example: The \textit{Total} column consists of the sum of the four previous columns and the \textit{Total} row consist of the sum of the column corresponding to each cell.
+These constraints can be written as:
+\begin{align}
+	\range{T_1}{\rangeall}{7} = \mathit{SUM}(\range{T_1}{\rangeall}{\rangeto{3}{6}}, row)\\
+	\range{T_2}{\rangeall}{1} = \mathit{SUM}(\range{T_1}{\rangeall}{\rangeto{3}{7}}, column).
+\end{align}
+
+\paragraph{Conditional aggregates}
+Conditional aggregates form more challenging constraints.
+There is an arbitrary reduction in size of the input data based on a condition to be matched.
+The complexity of this constraint is also dependent on the restrictions imposed on the constraint part.
+The conditional sum that calculates the \textit{Items sold total} column uses exact matching the salesperson name. Sticking close to the Excel notation this could be written as:
+\begin{align}
+	\range{T_1}{\rangeall}{10} = \mathit{SUMIF}(\range{T_5}{\rangeall}{1}, \range{T_1}{\rangeall}{2}, \range{T_5}{\rangeall}{2}).
+\end{align}
+Here, the condition will be applied to the \textit{Salesperson} column of table~5 (the test vector).
+Generally speaking the test vector has to match the sum column in size and the condition vector has to match the output vector.
+However, the condition can also consist of a boolean test such as $>5$ or a single value.
+To represent these constraints the Excel style using strings or exact values could be used but this might change yet to accomodate possibly more expressive conditions.
+
+\paragraph{Unary}
+The \textit{ID} column consists of a series:
+\begin{align}
+	\mathit{SERIES}(\range{T_1}{\rangeall}{1}).
+\end{align}
+Series are general cases of permutations.
+If a vector of size $n$ is a permutation of the numbers $[1, n]$, this can be an interesting observation, but often it is an interesting property of the vector which might unlock other constraints.
+See, for example, the column \textit{Rank} which is a permutation ($\mathit{PERMUTATION}(\range{T_1}{\rangeall}{8})$).
+This property allows the rank constraint to be applied to it (See below).
+Similarly, all-different constraints can be uselful as constraint or precondition.
+For numberic data all-different is not very useful but it might be for textual data such as colum \textit{Salesperson} ($\mathit{ALLDIFFERENT}(\range{T_1}{\rangeall}{2})$).
+Here, the all-different property allows the column to be refered to by a foreign key.
+
+\paragraph{Lookups}
+Consider the exact lookup in the example.
+This constraint can be expressed as:
+\begin{align}
+	\range{T_4}{\rangeall}{3} = \mathit{LOOKUP}(\range{T_1}{\rangeall}{1}, \range{T_1}{\rangeall}{2}, \range{T_4}{\rangeall}{2}).
+\end{align}
+Lookups, however, also support more complex constraints.
+For example, a fuzzy lookup assigns values to intervals to lookup the output value.
+Moreover, the input values can be transformed.
+The fuzzy lookup for the \textit{Label} column demonstrates this:
+\begin{align}
+	\range{T_1}{\rangeall}{9} = \mathit{LOOKUP}(\range{T_1}{\rangeall}{7} / \range{T_2}{2}{5}, \range{T_3}{\rangeall}{1}, \range{T_3}{\rangeall}{2}).
+\end{align}
+Here, the \textit{Total} column is divided by the calculated total average before being used to lookup what interval the performance falls into.
+
+\paragraph{Ranks}
+The column \textit{Rank} ranks based on the the column \textit{Total}, which can be expressed as:
+\begin{align}
+	\range{T_1}{\rangeall}{8} = RANK(\range{T_1}{\rangeall}{7}).
+\end{align}
+
+\paragraph{Structural constraints}
+The \textit{Salesperson} column in table~5 can be seen as a foreign key refering to the corresponding column in table~1.
+There are different options of representing this constraint:
+\begin{align}
+	\mathit{UNIQUE}(\range{T_5}{\rangeall}{1}) \subseteq \range{T_1}{\rangeall}{2} \label{eq:unique}\\
+	\mathit{FOREIGNKEY}(\range{T_1}{\rangeall}{2}, \range{T_5}{\rangeall}{1}) \label{eq:foreign_key}
+\end{align}
+Equation~\ref{eq:unique} uses a size-changing transformation or non-reversible function to express the underlying property.
+Instead one could also introduce a binary constraint expressing the foreign-key property directly (as in equation~\ref{eq:foreign_key}).
+
+\paragraph{Previous} Using the previous value in the constraint for a field can be solved easily by using a transformation.
+The calculation of the \textit{Total} column of table $T_6$ can be expressed as:
+\begin{align}
+	\range{T_6}{\rangeall}{4} = \range{T_6}{\rangeall}{2} - \range{T_6}{\rangeall}{3} + \mathit{PREVIOUS}(0, \range{T_6}{\rangeall}{4}, column).
+\end{align}
+The previous transformation drops the last value and adds an initial value, maintaining the size of the transformation input.
+In this specific case it is worth noting that this constraint can also be expressed differently:
+\begin{align}
+	\range{T_6}{\rangeall}{2} = \mathit{SUM}(\range{T_6}{\rangeall}{\rangeto{3}{4}}, row).
+\end{align}
+This constraint makes a different assumption about which columns are considered input and which are considered output.
+If no additional information is given, the algorithm could decide to provide both (redundant) constraints, or make a decision based on generality / specificity or a heuristic.
 

### 2016-03-08 Notes/Description.tex
Work in progress on algorithm

@@
 
+\section{Constraint-view Algorithm}
+\subsection{Input}
+The input to the algorithm consists of a set of version spaces~$S$ describing a hierarchical set of vectors.
+
+Additionally the user specifies if he wants the most general, specific or all solutions.
+This bias is captured is captured in the refinement operator~$\rho_2$ which traverses version spaces based on the bias.
+
+Implicitly this algorithm also uses a catalogue of constraints~$C$, an oracle~$O$ to calculate constraints and background knowledg~$B$ about 1) constraints and 2) properties.
+
+\subsection{Output}
+This algorithm~$\rho_1$ returns a set of constraints where the variables are instantiated by (sub-)version spaces contained in~$S$.
+
+\subsection{Method}
+In this case we will assume that a constraint~$c \in C$ has be selected to be learned.
+
+\begin{itemize}
+ 	\item For every free variable~$x$ in~$c$, prune~$S$ to obtain candidate values~$S_x$ using the properties that have been calculated so far: repeat:
+ 	\item For every version space~$s$ in~$S_x$, choose~$s$ as a value for $x$ and proceed
+ 	\item Use~$\rho_2$ to return the 
+ \end{itemize}
+
 \end{document}

### 2016-03-08 Progress.tex
Reorganized folder
Described the new algorithm

--- 2016-03-08 Notes/Description.tex
@@
 \setlength\parindent{0pt}
+\graphicspath{{Graphics/}}
 
@@
   \centering
-    \includegraphics[width=1\linewidth]{../Demo.png}
+    \includegraphics[width=1\linewidth]{Demo.png}
   \caption{Demo example}
@@
 
-\section{Constraint-view Algorithm}
+{\color{gray}
+\section{Constraint-view Algorithm (see sec~\ref{sec:algorithm2.0} for newer version)}
+
 \subsection{Input}
@@
  \end{itemize}
+}
+
+\section{Algorithm 2.0}
+\label{sec:algorithm2.0}
+
+\subsection{Concepts \& Issues}
+In order to abstract from the details of an excel sheet or CSV file, we use the notation introduced previously and reason about vectors and groups (powersets of vectors).
+Based on the bias of the algorithm the powerset maybe be restricted to, for example, continuous ranges.
+\\\\
+Currently, transformations and dependencies between constraints are not really included in the algorithm.
+They would, however, be useful for expressivity and eliminating redundancy, respectively.
+
+\subsection*{Step 1: Group detection}
+$\textit{CSV / Excel} \rightarrow \textit{Set of groups}$\\\\
+In this step the input CSV / Excel file is analyzed and partitioned into a set of groups.
+Preliminarily this step will be performed manually but it should be automized eventually.
+
+\subsection*{Step 2: Constraint detection}
+$\textit{Set of groups} \rightarrow \textit{Set of constraints}$\\\\
+In this step all constraints that satisfy the data and the bias are detected and returned.
+The main bias that will have to be accounted for is whether to return all, all most specific or all most general constraints.
+
+\subsubsection*{Step 2.1: Constraint selection}
+This step repeatedly selects a constraint template.
+Constraints are found for every template and the constraints that have been found already are used to prune the results for the current template.
+
+\subsubsection*{Step 2.2: Group selection}
+The group selection step repeatedly assigns groups to the different free variables present in the constraint template.
+It uses properties of the groups and preconditions on the variables imposed by the constraint template to avoid impossible assignments.
+This step is either implemented directly or as a CSP.
+
+\subsubsection*{Step 2.3: Subset selection}
+The last step within constraint detection finds assignments of subgroups to variables such that they satisfy the constraint template.
+It uses properties and the actually data to perform this step efficiently.
+This step is realised by translating the assignment problem to a constraint-specific CSP that includes the actual data.
+
+\subsection*{Step 3: Heuristic filtering}
+$\textit{Set of constraints} \rightarrow \textit{Set of relevant constraints}$\\\\
+In this last step the set of constraints is filtered and optionally ranked in order such that only relevant constraints are returned.
+So far we have identified three (increasingly difficult) steps to find a heuristic.
+First, rule of thumb heuristics can be used to prune some unlikely constraints.
+In a second step, dependencies in the data can be used to identify useful constraints.
+Finally, we can also attempt to learn what constraints are used frequently and how from existing data.
 

### 2016-03-31 Code/src/aspengine.py
added min,max,avg in asp

@@ -3,7 +3,7 @@ from os import system
 
 import numpy as np
 
-from core.constraint import SumColumn, SumRow, MaxRow, MaxColumn
+from core.constraint import SumColumn, SumRow, MaxRow, MaxColumn, MinColumn, MinRow, AvgColumn, AvgRow
 from core.group import GType
 from core.strategy import DictSolvingStrategy
 from functools import partial
@@ -23,18 +23,19 @@ class AspSolvingStrategy(DictSolvingStrategy):
             for i,xy_dict in enumerate(assignments):
               X = xy_dict["X"]
               Y = xy_dict["Y"]
+        #     print(X,Y,i)
               if X.row == False:
                 SAT = self.handle_aggregate_column_data_in_column(X,Y,i,aggregate)
                 if SAT:
                   selected_y, x_positions = SAT
-                  print("X COLUMN GROUP","SAT","X",X,"X Positions: ",x_positions,"Y",Y,"selected y vector",selected_y, sep="\n")
+            #     print("X COLUMN GROUP","SAT","X",X,"X Positions: ",x_positions,"Y",Y,"selected y vector",selected_y, sep="\n")
                   solution = {"X":X.vector_subset(min(x_positions),max(x_positions)),"Y":Y.vector_subset(selected_y,selected_y)}
                   solutions.append(solution)
               else: #X.row == True
                 SAT = self.handle_aggregate_column_data_in_rows(X,Y,i,aggregate)
                 if SAT:
                   start,end,selected_y = SAT
-                  print("SAT",start,end,selected_y, X,Y)
+            #     print("SAT",start,end,selected_y, X,Y)
                   solution = {"X":X.vector_subset(start,end),"Y":Y.vector_subset(selected_y,selected_y)}
                   solutions.append(solution)
             return solutions
@@ -59,13 +60,24 @@ class AspSolvingStrategy(DictSolvingStrategy):
                         solutions.append(solution)
             return solutions
         
+        #sum
         self.add_strategy(SumColumn(), partial(aggregate_columns,"sum"))
         self.add_strategy(SumRow(), partial(aggregate_rows,"sum"))
+        #max
         self.add_strategy(MaxColumn(), partial(aggregate_columns,"max"))
         self.add_strategy(MaxRow(), partial(aggregate_rows,"max"))
+        #min
+        self.add_strategy(MinColumn(), partial(aggregate_columns,"min"))
[…]

### 2016-03-31 Code/src/core/constraint.py

@@ -66,6 +66,34 @@ class MaxRow(Constraint):
         filters = []
         super().__init__("row-max", "{Y} = MAX({X}, row)", Source(variables), filters)
 
+class MinColumn(Constraint):
+    def __init__(self):
+        variables = [Variable("X", numeric=True), Variable("Y", vector=True, numeric=True)]
+        filters = []
+        super().__init__("column-min", "{Y} = MIN({X}, col)", Source(variables), filters)
+
+
+class MinRow(Constraint):
+    def __init__(self):
+        variables = [Variable("X", numeric=True), Variable("Y", vector=True, numeric=True)]
+        filters = []
+        super().__init__("row-min", "{Y} = MIN({X}, row)", Source(variables), filters)
+
+
+class AvgColumn(Constraint):
+    def __init__(self):
+        variables = [Variable("X", numeric=True), Variable("Y", vector=True, numeric=True)]
+        filters = []
+        super().__init__("column-avg", "{Y} = AVG({X}, col)", Source(variables), filters)
+
+
+class AvgRow(Constraint):
+    def __init__(self):
+        variables = [Variable("X", numeric=True), Variable("Y", vector=True, numeric=True)]
+        filters = []
+        super().__init__("row-avg", "{Y} = AVG({X}, row)", Source(variables), filters)
+
+
 
 class Permutation(Constraint):
     def __init__(self):

### 2016-03-31 Code/src/minizinc.py

@@ -126,11 +126,17 @@ class MinizincAssignmentStrategy(DictAssignmentStrategy):
             # TODO Implementation not finished
             return self._get_assignments(groups, constraint, local("minizinc/group/rank.mzn"))
 
-        self.add_strategy(SumColumn(), sum_columns)
+        self.add_strategy(SumColumn(), sum_columns) # probably, sum_columns should be renamed into binary assignments?
         self.add_strategy(SumRow(), sum_rows)
 
-        self.add_strategy(MaxColumn(), sum_columns)
+        self.add_strategy(MaxColumn(), sum_columns) 
         self.add_strategy(MaxRow(), sum_rows)
+
+        self.add_strategy(MinColumn(), sum_columns)
+        self.add_strategy(MinRow(), sum_rows)
+
+        self.add_strategy(AvgColumn(), sum_columns)
+        self.add_strategy(AvgRow(), sum_rows)
 
     def applies_to(self, constraint):
         return constraint in self.strategies

### 2016-03-31 Code/src/workflow.py

@@ -19,8 +19,8 @@ def main(csv_file, groups_file):
     solutions = Solutions()
     t_origin = time.time()
 
-#   constraints = [Permutation(), Series(), AllDifferent(), SumColumn(), SumRow(), Rank(), ForeignKey(), Lookup(), SumIf(), RunningTotal(), MaxColumn(), MaxRow()]
-    constraints = [MaxColumn(), MaxRow()]
+    constraints = [Permutation(), Series(), AllDifferent(), SumColumn(), SumRow(), Rank(), ForeignKey(), Lookup(), SumIf(), RunningTotal(), MaxColumn(), MaxRow(), MinColumn(), MinRow(), AvgColumn(), AvgRow()]
+#   constraints = [AvgColumn(), AvgRow()]
     for constraint in constraints:
         if not manager.supports_assignments_for(constraint):
             print("No assignment strategy for {}\n".format(constraint))

### 2016-03-31 Paper/ECAI/tabular.tex
added tex files for ECAI

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}

%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
                  % Do NOT use for camera-ready version of paper.

\begin{document}

\title{Tabular Constraint Learning}

\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3\institute{KU Leuven,
  Belgium, email: firstname.lastname@kuleuven} }

\maketitle
\bibliographystyle{ecai}

\begin{abstract}
  abstract
  \end{abstract}
\section{Introduction}
test

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-03-31 Paper/ECAI/z.sh

@@ -1,0 +1,9 @@
+file=tabular
+pdflatex $file
+bibtex $file
+pdflatex $file
+pdflatex $file 
+
+rm *log *aux *bbl *blg
+
+evince ${file}.pdf &

### 2016-03-31 Paper/ECAI/tabular.tex
added files for the paper

@@
 \usepackage{latexsym}
+\usepackage{xspace}
+\usepackage[dvipsnames]{xcolor}
+\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey} #1}\xspace}
 
@@
 
-\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3\institute{KU Leuven,
-  Belgium, email: firstname.lastname@kuleuven} }
+\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{KU Leuven,Belgium, email: firstname.lastname@kuleuven} }
 
@@
 \section{Introduction}
-test
+\sergey{we need structure} test
 
+\section{Related Work}
+\sergey{what else should be in the related work?}
+flashfill
+
+sketch
+
+tabular?
+
+\bibliography{references}
 \end{document}

### 2016-03-31 Paper/ECAI/tabular.tex
added references
updating the paper
added short related work annotation

@@
 \usepackage{xspace}
+\usepackage[numbers]{natbib}
+\usepackage{notoccite}
+
 \usepackage[dvipsnames]{xcolor}
@@
 
-\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{KU Leuven,Belgium, email: firstname.lastname@kuleuven} }
+\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{KU Leuven, Belgium, email: firstname.lastname@kuleuven} }
 
 \maketitle
-\bibliographystyle{ecai}
 
@@
 \section{Related Work}
-\sergey{what else should be in the related work?}
-flashfill
+\sergey{ECAI reference style file ignores their guideline and their guideline ignores what is written in the guidelines!}
+flashfill, flashextract, flashmeta \cite{flashfill,flashextract,flashmeta}
+\begin{itemize}
+  \item their supervised vs our unsupervised approach
+  \item they look for a single ``smallest'' solution, we enumerate them all
+  \item they are looking for a function, we solve constraint satisfaction problems
+  \item we do not assume classic row based data layout, we work in the tabular setting
+\end{itemize}
 
-sketch
+sketch \cite{sketch}
+\begin{itemize}
+  \item look for a constant that would fill in the gap in a program
+  \item tailored for programming languages
+  \item similar to model checking
+  \item looks for a single solution
+  \item similar to constraint satisfaction and sat, where one is interested in a single assignment that works for any potential input
+\end{itemize}
 
-tabular?
+tabular \cite{tabular}
+\begin{itemize}
+  \item language based on the excel tables that specify probabilistic models
+  \item a system for probabilistic inference and similarity mostly in the usage of excel
+  \item probabilistic constraint satisfaction (?) and graphical models
+  \item single solution again
+\end{itemize}
 
+modelseeker \cite{modelseeker} \sergey{Samuel, Luc, probably you would need elaborate here more in details}
+
+\begin{itemize}
+  \item not designed for excel-like data representation (type consistency, groups, etc)
+  \item not designed for excel-like constraints (lookups, conditional ifs, etc)
+  \item does not support user extensions (?)
+\end{itemize
+
+claudien \cite{claudien} \sergey{Samuel, Luc, you would need to help with this one}
+
+\bibliographystyle{ecai}
 \bibliography{references}

### 2016-04-01 Paper/ECAI/tabular.tex
added short related work annotation
text

@@
 \section{Introduction}
-\sergey{we need structure} test
+\sergey{bullet points for luc to start introduction}\\
+\textbf{Key question}:\\
+Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\
+
+\textbf{Motivation}:
+\begin{itemize}
+  \item File generated from model, model got lost - > reconstruct
+  \item Constraint programming is hard - is Excel hard?
+  \item Avoid manual analysis, provide selection of constraints
+  \item Error checking
+  \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
+\end{itemize}
+
+\textbf{Novelty:}
+\begin{itemize}
+  \item Unsupervised setting (contrary to flashfill, etc)
+  \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
+  \item Data format (2D) -- data is no longer in rows like a classic ML or DM settings
+  \item Declarative, general / modular, stacking of constraint problems 
+\end{itemize}
+
+\sergey{we need structure here}
+
+\textbf{Approach}
+\begin{itemize}
+  \item Notation
+  \item Algorithm (select constraints, find assignments, find solutions)
+\end{itemize}
+
+{\bfseries 
+  Experimental questions
+}
+\begin{itemize}
+  \item  How accurate are we? (Accuracy / recall)
+  \item  How fast are we and which factors affect the runtime (how)?
+  \item  How general is our approach, what limitations are there?
+\end{itemize}
 
 \section{Related Work}
+\sergey{key bullet points for Luc and possibly Samuel and me to make related work section}
+
 \sergey{ECAI reference style file ignores their guideline and their guideline ignores what is written in the guidelines!}
@@
   \item does not support user extensions (?)
-\end{itemize
+\end{itemize}
 

### 2016-04-01 Paper/ECAI/tabular.tex
refactoring the tex file

@@
 \usepackage[dvipsnames]{xcolor}
-\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey} #1}\xspace}
+\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
 
@@
 \textbf{Key question}:\\
-Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\
+Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\\
 
@@
 \begin{itemize}
-  \item File generated from model, model got lost - > reconstruct
+  \item File generated from model, model got lost, need to reconstruct
   \item Constraint programming is hard - is Excel hard?
@@
 
-\sergey{we need structure here}
+\sergey{to himself we need structure here}
 

### 2016-04-04 Code/src/core/constraint.py
Generalized sum col / row

@@ -2,7 +2,7 @@ from typing import List, Dict
 
 from core.assignment import Source, Filter, Variable, SameLength, ConstraintSource, NotSubgroup, SameTable, \
     SameOrientation, SameType, SizeFilter, Not
-from core.group import GType, Group
+from core.group import GType, Group, Orientation
 
 
 class Constraint:
@@ -43,6 +43,45 @@ integer = {GType.int}
 numeric = {GType.int, GType.float}
 textual = {GType.string}
 discrete = {GType.string, GType.int}
+
+
+class Sum(Constraint):
+    x = Variable("X", types=numeric)
+    y = Variable("Y", vector=True, types=numeric)
+
+    def __init__(self, orientation: Orientation):
+        self._orientation = orientation
+        size = Group.columns if orientation == Orientation.VERTICAL else Group.rows
+        o_string = "col" if orientation == Orientation.VERTICAL else "row"
+        variables = [self.x, self.y]
+
+        def test(_, a: Dict[str, Group]):
+            x_group, y_group = [a[v.name] for v in variables]
+            o_match = x_group.row == (orientation == Orientation.HORIZONTAL)
+            return y_group.length() <= size(x_group) if o_match else y_group.length() == size(x_group)
+
+        filter_class = type("Sum{}Length".format(o_string.capitalize()), (Filter,), {"test": test})
+        size_filter = SizeFilter([self.x], rows=2) if Orientation.column(orientation) else SizeFilter([self.x], cols=2)
+        filters = [size_filter, filter_class(variables)]
+        super().__init__("sum ({})".format(o_string), "{Y} = SUM({X}, " + o_string + ")", Source(variables), filters)
+
+    @property
+    def orientation(self):
+        return self._orientation
+
+
+class ColumnSum(Sum):
+    def __init__(self):
+        super().__init__(Orientation.VERTICAL)
+
+# constraint not g_row_orientation[assign[1]] -> g_length[assign[2]] = g_rows[assign[1]];
+# constraint g_row_orientation[assign[1]] -> g_length[assign[2]] <= g_rows[assign[1]];
+# constraint g_columns[assign[1]] > 1;
+# constraint forall(v in 1..nV)(v_numeric[v] -> g_numeric[assign[v]]);
+
+# TODO Same table, different orientation, overlapping bounds => prune assignment already
+
+# TODO Subset -> Fuzzy lookup
 
 
 class SumColumn(Constraint):
[…]

### 2016-04-04 Code/src/internal.py

@@ -12,7 +12,7 @@ class InternalCSPStrategy(AssignmentStrategy):
     def __init__(self):
         super().__init__()
         self.constraints = {Series(), AllDifferent(), Permutation(), Rank(), ForeignKey(), Lookup(), FuzzyLookup(),
-                            SumIf(), MaxIf(), RunningTotal(), ForeignProduct(), SumColumn()}
+                            SumIf(), MaxIf(), RunningTotal(), ForeignProduct(), SumColumn(), ColumnSum()}
 
     def applies_to(self, constraint):
         return constraint in self.constraints
@@ -184,6 +184,36 @@ class InternalSolvingStrategy(DictSolvingStrategy):
 
             return solutions
 
+        def sum__(c: Sum, assignments: List[Dict[str, Group]], solutions: Solutions):
+            solutions = []
+            o_column = Orientation.column(c.orientation)
+
+            for assignment in assignments:
+                x_group, y_group = (assignment[k.name] for k in [c.x, c.y])
+                x_data = x_group.data
+                assert isinstance(x_group, Group)
+                assert isinstance(y_group, Group)
+                y_length = y_group.length()
+
+                o_match = x_group.row == Orientation.row(c.orientation)
+                if o_match:
+                    sums = numpy.sum(x_group.data, 0 if o_column else 1)
+                    for y_vector_group in y_group:
+                        match = pattern_finder(sums, y_vector_group.get_vector(1))
+                        x_match = [x_group.vector_subset(m + 1, m + y_length) for m in match]
+                        solutions += [{c.x.name: x, c.y.name: y_vector_group} for x in x_match]
+                else:
+                    x_length = x_group.rows() if o_column else x_group.columns()
+                    for x1 in range(x_length - 1):
+                        for x2 in range(x1 + 2, x_length + 1):
+                            sums = numpy.sum(x_data[x1:x2, :], 0) if o_column else numpy.sum(x_data[:, x1:x2], 1)
+                            for y_vector_group in y_group:
+                                if numpy.vectorize(equal)(sums, y_vector_group.get_vector(1)).all():
+                                    x_subgroup = x_group.vector_subset(x1 + 1, x2)
+                                    solutions.append({c.x.name: x_subgroup, c.y.name: y_vector_group})
+
+            return solutions
+
         self.add_strategy(Series(), series)
         self.add_strategy(AllDifferent(), all_different)
[…]

### 2016-04-04 Code/src/parser.py

@@ -15,8 +15,8 @@ def parse(filename):
 
 def get_groups_tables(csv_file, groups_file=None):
     data = parse(csv_file)
+    type_data = np.vectorize(detect_type)(data)
     if groups_file is None:
-        type_data = np.vectorize(detect_type)(data)
         t = list(detect_tables(type_data))
         print("PARSE: Detected table areas: {}".format(", ".join(["[{}:{}, {}:{}]".format(*r) for r, o in t])))
         t = [(b, Table("T{}".format(i + 1), Bounds(b).subset(data), o)) for i, (b, o) in enumerate(t)]
@@ -25,7 +25,8 @@ def get_groups_tables(csv_file, groups_file=None):
         print()
         return groups
     else:
-        tables = {}
+        table_dict = {}
+        t = []
         groups = []
         with open(groups_file, "r") as group_file:
             json_data = json.load(group_file)
@@ -33,10 +34,14 @@ def get_groups_tables(csv_file, groups_file=None):
                 bounds = Bounds(table_description["Bounds"])
                 data_subset = bounds.subset(data)
                 name = table_description["Name"]
-                tables[name] = Table(name, data_subset)
-            for group_description in json_data["Groups"]:
-                table = tables[group_description["Table"]]
-                groups.append(create_group(group_description["Bounds"], table))
+                table_dict[name] = Table(name, data_subset)
+                t.append((bounds.bounds, table_dict[name]))
+            if "Groups" in json_data:
+                for group_description in json_data["Groups"]:
+                    table = table_dict[group_description["Table"]]
+                    groups.append(create_group(group_description["Bounds"], table))
+            else:
+                groups = detect_groups(data, type_data, t)
         return groups
 
 
@@ -83,7 +88,8 @@ def detect_tables(type_data):
         current = new_current
     tables = [((r1 + 1, r2, c1 + 1, c2), o) for (r1, r2, c1, c2), o in [remove_header(rec, type_data) for rec in saved]]
     tables = [(rec, o) for rec, o in tables if rec[0] <= rec[1] and rec[2] <= rec[3]]
-    return sorted(tables, key=lambda t: (t[0][0], t[0][2], t[0][1], t[0][3], t[1].value))
+    # TODO Error BMI (no T1)
+    return sorted(tables, key=lambda t: (t[0][0], t[0][2], t[0][1], t[0][3]))
 
 
 def detect_groups(data, type_data, tables):
@@ -125,7 +131,7 @@ def detect_groups(data, type_data, tables):
[…]

### 2016-04-04 Code/src/workflow.py

@@ -24,6 +24,7 @@ def main(csv_file, groups_file, verbose):
         Series(),
         AllDifferent(),
         SumColumn(),
+        ColumnSum(),
         SumRow(),
         MaxColumn(),
         MaxRow(),

### 2016-04-04 Paper/ECAI/tabular.tex
ADDED THE FIGURE

@@
 \usepackage{xspace}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 \usepackage[numbers]{natbib}
@@
 \sergey{to himself we need structure here} 
+\begin{figure}[thpb]
+  \begin{center}
+    \includegraphics[width=0.55\textwidth]{figures/Demo.png}
+  \end{center}
+  \vspace{-10pt}
+  \caption{An example of constraint reconstruction (in blue) with indicated groups (in green)}
+  \label{fig:main_example}
+\end{figure}
+
+\section{Formalization}
+
+\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
+\newcommand{\generategroups}{\textit{generateGroups}\xspace}
+\newcommand{\extractgroups}{\textit{extractGroups}\xspace}
+\newcommand{\findassignment}{\textit{findAssignment}\xspace}
+
+\begin{algorithm}[thb]
+  \begin{algorithmic}
+    \footnotesize
+    \State \textbf{Input:} $D$ -- dataset, (optional groups $G$)
+    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
+    \If{$G$ is not defined}
+       \State $G \gets \extractgroups(D)$
+    \EndIf
+  \For{$c \in \constraints$} \Comment{\constraints -- the set of predefined Excel constraints}
+    \For{$G_1, \dots, G_n \in \generategroups(c,G)$}
+      \State $S \gets S \cup \findassignment(c, G1, \dots, G_n)$
+    \EndFor
+  \EndFor\\
+\Return $S$
+\end{algorithmic}
+\caption{First Order Model: Iterative Query Enumeration}
+\label{algo:top_cycle}
+\end{algorithm}
+
+\section{Case Study aka Experiments}
 

### 2016-04-04 Code/src/parser.py
Allow user to provide table orientation

@@ -34,7 +34,14 @@ def get_groups_tables(csv_file, groups_file=None):
                 bounds = Bounds(table_description["Bounds"])
                 data_subset = bounds.subset(data)
                 name = table_description["Name"]
-                table_dict[name] = Table(name, data_subset)
+                orientation = None
+                if "Orientation" in table_description:
+                    o_string = table_description["Orientation"].lower()
+                    if o_string == "row" or o_string == "horizontal":
+                        orientation = Orientation.HORIZONTAL
+                    elif o_string == "column" or o_string == "col" or o_string == "vertical":
+                        orientation = Orientation.VERTICAL
+                table_dict[name] = Table(name, data_subset, orientation)
                 t.append((bounds.bounds, table_dict[name]))
             if "Groups" in json_data:
                 for group_description in json_data["Groups"]:

### 2016-04-06 Code/src/core/group.py
Added projection

@@ -125,7 +125,8 @@ class Group:
         self._row = row
         self._dtype = dtype
         self._data = data
-        self._is_partial = numpy.any(numpy.vectorize(lambda x: x is None)(self._data))
+        from core.constraint import Operation
+        self._is_partial = numpy.any(numpy.vectorize(Operation.blank_filter(self._data)[1])(self._data))
 
     @property
     def is_partial(self):

### 2016-04-06 Code/src/engine/internal.py

@@ -6,6 +6,27 @@ from core.solutions import Solutions
 from core.strategy import AssignmentStrategy, DictSolvingStrategy
 
 
+class MaxRange:
+    def __init__(self, test: callable):
+        self._test = test
+
+    def find(self, start, end, size, limit=None, last=None):
+        limit = start + size - 1 if limit is None else limit
+        last = end if last is None else last
+        self._find(start, end, size, limit, last)
+
+    def _find(self, start, end, size, limit, last):
+        if last - start < size:
+            pass
+        elif end - start < size or end <= limit:
+            self._find(start + 1, last, size, max(start + size, limit), last)
+        else:
+            if self._test(start, end):
+                self._find(start + 1, last, size, end, last)
+            else:
+                self._find(start, end - 1, size, limit, last)
+
+
 # TODO Sum IF with Nones
 class InternalCSPStrategy(AssignmentStrategy):
     def __init__(self):
@@ -230,20 +251,36 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             return self._generate_test_vectors(assignments, keys, is_sum_product, lambda r, o1, o2: ordered(o1, o2))
 
         def project(c: Projection, assignments, _):
+            solutions = []
             masks = {}
             for assignment in assignments:
                 r_group, p_group = [assignment[v.name] for v in [c.result, c.projected]]
                 assert isinstance(r_group, Group)
                 assert isinstance(p_group, Group)
                 if p_group not in masks:
-                    masks[p_group] = numpy.vectorize(lambda x: x is not None)(p_group.data)
+                    bool_mask = numpy.vectorize(Operation.blank_filter(p_group.data)[1])(p_group.data)
+                    masks[p_group] = numpy.vectorize(lambda e: 1 if e else 0)(bool_mask)
                 p_masked = masks[p_group] if p_group.row else masks[p_group].T
-                print(p_masked)
                 size = 2
-                for s in range(p_group.vectors() - 1):
-                    for e in range(s + size - 1, p_group.vectors() - 1):
-                        pass
-            return []
+
+                def check(start, end):
+                    result = numpy.sum(p_masked[start:end, :], 0)
+                    if numpy.vectorize(lambda e: e == 1)(result).all():
+                        p_subgroup = p_group.vector_subset(start + 1, end)
+                        p_data = p_subgroup.data if p_subgroup.row else p_subgroup.data.T
[…]

### 2016-04-06 Code/src/engine/internal.py
Added projection (fix)

@@ -253,6 +253,8 @@ class InternalSolvingStrategy(DictSolvingStrategy):
         def project(c: Projection, assignments, _):
             solutions = []
             masks = {}
+            size = 2
+
             for assignment in assignments:
                 r_group, p_group = [assignment[v.name] for v in [c.result, c.projected]]
                 assert isinstance(r_group, Group)
@@ -261,7 +263,6 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     bool_mask = numpy.vectorize(Operation.blank_filter(p_group.data)[1])(p_group.data)
                     masks[p_group] = numpy.vectorize(lambda e: 1 if e else 0)(bool_mask)
                 p_masked = masks[p_group] if p_group.row else masks[p_group].T
-                size = 2
 
                 def check(start, end):
                     result = numpy.sum(p_masked[start:end, :], 0)
@@ -275,11 +276,12 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     return False
 
                 max_range = MaxRange(check)
-                if r_group == p_group:
-                    for r_i in range(r_group.vectors()):
+                for r_i in range(r_group.vectors()):
+                    if r_group == p_group:
                         max_range.find(0, r_i, size)
                         max_range.find(r_i + 1, p_group.vectors(), size)
-
+                    else:
+                        max_range.find(0, p_group.vectors(), size)
             return solutions
 
         self.add_strategy(Series(), series)

### 2016-04-06 Code/src/core/constraint.py
Partial test fix

@@ -83,10 +83,10 @@ class Operation(Enum):
 
     @staticmethod
     def blank_filter(data):
-        if data.dtype == numpy.object:
-            return None, lambda e: e is not None
-        else:
+        if numpy.issubdtype(data.dtype, numpy.float):
             return numpy.nan, lambda e: not numpy.isnan(e)
+        else:
+            return None, lambda e: e is not None
 
 
 class Aggregate(Constraint):

### 2016-04-06 Code/src/core/group.py

@@ -126,7 +126,7 @@ class Group:
         self._dtype = dtype
         self._data = data
         from core.constraint import Operation
-        self._is_partial = numpy.any(numpy.vectorize(Operation.blank_filter(self._data)[1])(self._data))
+        self._is_partial = not numpy.all(numpy.vectorize(Operation.blank_filter(self._data)[1])(self._data))
 
     @property
     def is_partial(self):

### 2016-04-06 Code/src/engine/internal.py

@@ -257,8 +257,6 @@ class InternalSolvingStrategy(DictSolvingStrategy):
 
             for assignment in assignments:
                 r_group, p_group = [assignment[v.name] for v in [c.result, c.projected]]
-                assert isinstance(r_group, Group)
-                assert isinstance(p_group, Group)
                 if p_group not in masks:
                     bool_mask = numpy.vectorize(Operation.blank_filter(p_group.data)[1])(p_group.data)
                     masks[p_group] = numpy.vectorize(lambda e: 1 if e else 0)(bool_mask)

### 2016-04-06 Code/src/engine/internal.py
Changed aggregates to maximizing range

@@ -222,15 +222,23 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                         solutions += [{c.x.name: x, c.y.name: y_vector_group} for x in x_match
                                       if not x.overlaps_with(y_vector_group)]
                 else:
-                    x_length = x_group.rows() if o_column else x_group.columns()
-                    for x1 in range(x_length - 1):
-                        for x2 in range(x1 + 2, x_length + 1):
-                            sums = operation_f(x_data[x1:x2, :], 0) if o_column else operation_f(x_data[:, x1:x2], 1)
-                            for y_vector_group in y_group:
-                                if numpy.vectorize(equal)(sums, y_vector_group.get_vector(1)).all():
-                                    x_subgroup = x_group.vector_subset(x1 + 1, x2)
-                                    if not x_subgroup.overlaps_with(y_vector_group):
-                                        solutions.append({c.x.name: x_subgroup, c.y.name: y_vector_group})
+                    if not o_column:
+                        x_data = x_data.T
+
+                    def check(start, end):
+                        result = operation_f(x_data[start:end, :], 0)
+                        if equal_v(result, y_group.get_vector(y_i + 1)).all():
+                            x_subgroup = x_group.vector_subset(start + 1, end)
+                            y_subgroup = y_group.vector_subset(y_i + 1, y_i + 1)
+                            solutions.append({c.x.name: x_subgroup, c.y.name: y_subgroup})
+
+                    max_range = MaxRange(check)
+                    for y_i in range(y_group.vectors()):
+                        if y_group == x_group:
+                            max_range.find(0, y_i, 2)
+                            max_range.find(y_i + 1, x_group.vectors(), 2)
+                        else:
+                            max_range.find(0, x_group.vectors(), 2)
 
             return solutions
 

### 2016-04-06 Code/src/core/solutions.py
Added projected pruning

@@ -17,4 +17,4 @@ class Solutions:
 
     @staticmethod
     def _to_tuple(constraint, solution):
-        return (solution[v.name] for v in constraint.variables)
+        return tuple(solution[v.name] for v in constraint.variables)

### 2016-04-06 Code/src/engine/internal.py

@@ -200,11 +200,16 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             return self._generate_test_vectors(assignments, keys, is_foreign_product)
 
         def aggregate(c: Aggregate, assignments: List[Dict[str, Group]], solutions: Solutions):
-            solutions = []
+            results = []
             o_column = Orientation.column(c.orientation)
             operation_f = c.operation.aggregate
+            projection = Projection()
 
-            # TODO Largest Area?
+            def add(solution):
+                mapping = {c.x: projection.projected, c.y: projection.result}
+                mapped = {mapping[v].name: solution[v.name] for v in c.variables}
+                if not solutions.has_solution(projection, mapped):
+                    results.append(solution)
 
             for assignment in assignments:
                 x_group, y_group = (assignment[k.name] for k in [c.x, c.y])
@@ -219,8 +224,9 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     for y_vector_group in y_group:
                         match = pattern_finder(sums, y_vector_group.get_vector(1))
                         x_match = [x_group.vector_subset(m + 1, m + y_length) for m in match]
-                        solutions += [{c.x.name: x, c.y.name: y_vector_group} for x in x_match
-                                      if not x.overlaps_with(y_vector_group)]
+                        for x in x_match:
+                            if not x.overlaps_with(y_vector_group):
+                                add({c.x.name: x, c.y.name: y_vector_group})
                 else:
                     if not o_column:
                         x_data = x_data.T
@@ -230,7 +236,7 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                         if equal_v(result, y_group.get_vector(y_i + 1)).all():
                             x_subgroup = x_group.vector_subset(start + 1, end)
                             y_subgroup = y_group.vector_subset(y_i + 1, y_i + 1)
-                            solutions.append({c.x.name: x_subgroup, c.y.name: y_subgroup})
+                            add({c.x.name: x_subgroup, c.y.name: y_subgroup})
 
                     max_range = MaxRange(check)
                     for y_i in range(y_group.vectors()):
@@ -240,7 +246,7 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                         else:
                             max_range.find(0, x_group.vectors(), 2)
 
-            return solutions
+            return results
 
         def product(c: Product, assignments, solutions):
             keys = [c.result, c.first, c.second]

### 2016-04-07 Paper/ECAI/tabular.tex
Expanded pseudo code
new examples
formalization

@@
 \usepackage{xspace}
+\usepackage{hyperref}
 \usepackage{algorithm}
@@
 \newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
+\newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
 
@@
   \begin{center}
-    \includegraphics[width=0.55\textwidth]{figures/Demo.png}
+    \includegraphics[width=0.51\textwidth]{figures/Demo.png}
   \end{center}
@@
 \newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
-\newcommand{\generategroups}{\textit{generateGroups}\xspace}
-\newcommand{\extractgroups}{\textit{extractGroups}\xspace}
-\newcommand{\findassignment}{\textit{findAssignment}\xspace}
+\newcommand{\format}[1]{\textit{#1}\xspace}
+\newcommand{\generategroups}{\format{generateAssignments}}
+\newcommand{\extractgroups}{\format{extractGroups}}
+\newcommand{\extracttables}{\format{extractTables}}
+\newcommand{\learnconstraints}{\format{learnConstraints}}
+\newcommand{\findassignment}{\format{findSolutions}}
+\newcommand{\postprocess}{\format{pruneRedundant}}
 
@@
     \footnotesize
-    \State \textbf{Input:} $D$ -- dataset, (optional groups $G$)
-    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
-    \If{$G$ is not defined}
-       \State $G \gets \extractgroups(D)$
-    \EndIf
+    \State \textbf{Input:} $G$ -- set of groups
+    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
+    \State $S \gets \emptyset$ \Comment{The set of solutions}
   \For{$c \in \constraints$} \Comment{\constraints -- the set of predefined Excel constraints}
-    \For{$G_1, \dots, G_n \in \generategroups(c,G)$}
-      \State $S \gets S \cup \findassignment(c, G1, \dots, G_n)$
+  \State $v_1, ..., v_n =$ variables of $c$ 
+    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S)$}
+      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
     \EndFor
@@
 \end{algorithmic}
-\caption{First Order Model: Iterative Query Enumeration}
-\label{algo:top_cycle}
+\caption{Tabular constraint learning}
+\label{algo:tcl}
+\end{algorithm}
+
+\sergey{we assume an order on the constraints (it doesn't have to be complete though)}
+
+\subsection{Groups}
+In this work we distinguish the following types of data: numeric and textual.
+The numeric type has two subtypes: integers and floats.
+We also consider the special element called \textit{None}, which has two types: numeric and textual.
+A set is called type-consistent iff all elements are either numeric or textual.
+Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.
+
+A \textit{vector} is a subrange of a column or of a row that is type-consistent.
+If a vector is a subrange of a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
+A \textit{group} is a subrange of vectors with the same orientation in a table.
+We use the following notation to refer to a row group $G$ in the $N$-th table with rows ranging from $a$ to $b$: $G = T_N[a{:}b,:]$, where $N,a,b$ are natural numbers.
+Similarly for a column group $G$ ranging from $a$ to $b$ columns in the $N$-th table we write $G = T_N[{:},a{:}b]$.
+
+\subsection{Algorithm description}
+
+Let us describe the key functions in Algorithm \ref{algo:tcl}.
+\paragraph{$\generategroups(\textit{Constraint,GroupSet,Solutions})$} is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}.
+Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint.
+However, many constraints have the same candidate generation procedures, e.g., min, max, avg, count, etc.
+
+\paragraph{$\findassignment(\textit{Constraint,Candidates,Solutions})$} is the function looking for the subsets of vectors in the candidates satisfying the constraint.
+If multiple subsets satisfy the constraint, a maximal is selected.
+If $G_1,G_2$, associated with the variables $X,Y$ in \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$.
+For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}7]$ can be used for both $X$ and $Y$ in the row-wise sum constraint, then as $x$ would be selected $T_1[{:},3{:}6]$ and as $y$ would be $T_1[{:},7]$.
+
+\paragraph{}
+
+\begin{algorithm}[thb]
+  \begin{algorithmic}
+    \footnotesize
+    \State \textbf{Input:} $D$ -- dataset, (optional: tables $T$, groups $G$)
+    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
+    \If{$T$ is \textbf{not} provided}
+      \State $T \gets \extracttables(D)$
+    \EndIf
+    \If{$G$ is \textbf{not} provided}
+      \State $G \gets \extractgroups(D, T)$
+    \EndIf
+    \State $S \gets \learnconstraints(G)$
+    \State $S \gets \postprocess(S)$
+    \State \Return $S$
+\end{algorithmic}
+\caption{Workflow}
+\label{algo:workflow}
 \end{algorithm}

### 2016-04-07 Paper/ECAI/z.sh

@@ -6,4 +6,4 @@ pdflatex $file
 
 rm *log *aux *bbl *blg
 
-evince ${file}.pdf &
+evince ${file}.pdf &> /dev/null &

### 2016-04-07 Paper/ECAI/tabular.tex
writing the formalization

@@
 
+\newcommand{\CName}{Name\xspace}
+\newcommand{\CSignature}{Signature\xspace}
+\newcommand{\CFunction}{Function\xspace}
+
 \begin{algorithm}[thb]
@@
     \State $S \gets \emptyset$ \Comment{The set of solutions}
-  \For{$c \in \constraints$} \Comment{\constraints -- the set of predefined Excel constraints}
+  \For{$c \in \constraints$} \Comment{\constraints -- partially ordered set of Excel constraints; Table \ref{table:constraints}}
   \State $v_1, ..., v_n =$ variables of $c$ 
@@
 \end{algorithm}
-
-\sergey{we assume an order on the constraints (it doesn't have to be complete though)}
-
-\subsection{Groups}
+\subsection{Groups, Type-consistency and Constraints}
 In this work we distinguish the following types of data: numeric and textual.
@@
 We also consider the special element called \textit{None}, which has two types: numeric and textual.
-A set is called type-consistent iff all elements are either numeric or textual.
+A set is called \textit{type-consistent} iff all elements are either numeric or textual.
 Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.
@@
 
-\subsection{Algorithm description}
+A \textit{constraint} is a triple of \textit{(\CName,\CSignature,\CFunction)}, where \CName is a textual name of the constraint; \CSignature is the signature of the constraint consisting of its arity and properties of the arguments, such as their types, sizes (properties might constrain sizes of two arguments as well, by for example, require them to be equal), etc; \CFunction is the function over arguments specified in \CSignature that represents constraint satisfaction.
+For example, a column-wise sum constraint has \CName equal to \textit{Y = SUM(X, col)}, its \CFunction sets arity to 2 and both arguments must be numeric and \sergey{Samuel, could you put the actual properties and \CFunction here?}
 
-Let us describe the key functions in Algorithm \ref{algo:tcl}.
-\paragraph{$\generategroups(\textit{Constraint,GroupSet,Solutions})$} is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}.
+\subsection{Constraint learning algorithm description}
+Let us describe the key steps and functions in Algorithm \ref{algo:tcl}.
+Essentially, the algorithm has two steps: candidate group generation and subgroup satisfaction search, which is in line with the well-known in AI paradigm of ``generate-and-test'' \cite{whaisasp}.
+Let us elaborate on each step in detail.
+
+\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}.
 Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint.
@@
 
-\paragraph{$\findassignment(\textit{Constraint,Candidates,Solutions})$} is the function looking for the subsets of vectors in the candidates satisfying the constraint.
+\paragraph{Subgroup satisfaction search} $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function looking for the subsets of vectors in the candidates satisfying the constraint.
 If multiple subsets satisfy the constraint, a maximal is selected.
-If $G_1,G_2$, associated with the variables $X,Y$ in \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$.
+If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$.
 For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}7]$ can be used for both $X$ and $Y$ in the row-wise sum constraint, then as $x$ would be selected $T_1[{:},3{:}6]$ and as $y$ would be $T_1[{:},7]$.
 
-\paragraph{}
+\subsection{Constraints}
+\sergey{we need to explain that constraints order is a partial order, which is a DAG}
+\begin{table}
+  \centering
+  \begin{tabular}{lcc}
+    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ 
+  \end{tabular}
+  \caption{Tabular Constraints}
+  \label{table:constraints}
+\end{table}
 

### 2016-04-07 Code/src/accuracy.py
Removed partiality test from conditional aggregate -> solver

@@ -15,7 +15,6 @@ files = ["bmi", "age_department_sumif", "average_ablebits", "columnwise-sum-rows
          "shares", "sum_if_double_condition", "sum_table_from_guide", "sumif_blanks", "sumif_example",
          "sumif_games_toys", "sumif_region", "sumif_uk", "sumproduct", "week_2_busn_store_2"]
 
-# TODO sumif_example
 
 def main():
     categories = {"essential constraints": "Essential"}

### 2016-04-07 Code/src/core/constraint.py

@@ -200,7 +200,7 @@ class Rank(Constraint):
 
     def __init__(self):
         variables = [self.x, self.y]
-        source = Source(variables)
+        source = Source(variables)  # Not from Permutation because of possible ties
         filters = [SameLength(variables), NotPartial(variables)]
         super().__init__("rank", "{Y} = RANK({X})", source, filters)
 
@@ -249,9 +249,9 @@ class FuzzyLookup(Constraint):
 
 
 class ConditionalAggregate(Constraint):
-    o_key = Variable("OK", vector=True, types=textual)
+    o_key = Variable("OK", vector=True, types=discrete)
     result = Variable("R", vector=True, types=numeric)
-    f_key = Variable("FK", vector=True, types=textual)
+    f_key = Variable("FK", vector=True, types=discrete)
     values = Variable("V", vector=True, types=numeric)
 
     def __init__(self, name: str, operator, default=0):
@@ -260,7 +260,7 @@ class ConditionalAggregate(Constraint):
         variables = [self.o_key, self.result, self.f_key, self.values]
         foreign_key = ForeignKey()
         source = ConstraintSource(variables, foreign_key, {foreign_key.pk.name: "OK", foreign_key.fk.name: "FK"})
-        filters = [SameLength([self.o_key, self.result]), SameLength([self.f_key, self.values]), NotPartial(variables),
+        filters = [SameLength([self.o_key, self.result]), SameLength([self.f_key, self.values]),
                    SameTable([self.o_key, self.result]), SameTable([self.f_key, self.values]),
                    SameOrientation([self.o_key, self.result]), SameOrientation([self.f_key, self.values])]
         super().__init__("{}-if".format(name.lower()), "{R} = " + name.upper() + "IF({FK}={OK}, {V})", source, filters)

### 2016-04-07 Paper/ECAI/tabular.tex
fixing the paper

@@
 
-A \textit{vector} is a subrange of a column or of a row that is type-consistent.
-If a vector is a subrange of a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
+A \textit{vector} is either a column or a row that is type-consistent.
+If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
 A \textit{group} is a subrange of vectors with the same orientation in a table.
-We use the following notation to refer to a row group $G$ in the $N$-th table with rows ranging from $a$ to $b$: $G = T_N[a{:}b,:]$, where $N,a,b$ are natural numbers.
-Similarly for a column group $G$ ranging from $a$ to $b$ columns in the $N$-th table we write $G = T_N[{:},a{:}b]$.
+We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
 
-A \textit{constraint} is a triple of \textit{(\CName,\CSignature,\CFunction)}, where \CName is a textual name of the constraint; \CSignature is the signature of the constraint consisting of its arity and properties of the arguments, such as their types, sizes (properties might constrain sizes of two arguments as well, by for example, require them to be equal), etc; \CFunction is the function over arguments specified in \CSignature that represents constraint satisfaction.
-For example, a column-wise sum constraint has \CName equal to \textit{Y = SUM(X, col)}, its \CFunction sets arity to 2 and both arguments must be numeric and \sergey{Samuel, could you put the actual properties and \CFunction here?}
+An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}, where \textit{\CName} is a textual name of the constraint; \textit{\CSignature} is a set of constraints specifying the properties of the arguments, such as their types, e.g., to require them to be integers or sizes, e.g.
+the length of vectors in the arguments must be equal, i.e., these are constraints on the group meta-information not on the actual group content; \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function.
+For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}; its \CSignature is: $X$ is numeric, $Y$ is integer and the length of vectors in $X$ is the same as in $Y$; its \CFunction is the following constraint: a pair of vectors $x,y$ is a solution iff $x \in X, y \in Y$ and each value in $x$ has the rank (possibly with ties) specified in $y$.
 
@@
 Let us describe the key steps and functions in Algorithm \ref{algo:tcl}.
-Essentially, the algorithm has two steps: candidate group generation and subgroup satisfaction search, which is in line with the well-known in AI paradigm of ``generate-and-test'' \cite{whaisasp}.
+Essentially, the algorithm has two steps: candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}.
 Let us elaborate on each step in detail.
 
-\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}.
+\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = RANK(X)} and $X,Y$ satisfy the constraints from \CSignature above.
 Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint.
-However, many constraints have the same candidate generation procedures, e.g., min, max, avg, count, etc.
+However, many constraints have the same candidate generation procedures, e.g., sum, min, max, avg, count, etc.
 
@@
 If multiple subsets satisfy the constraint, a maximal is selected.
-If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$.
-For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}7]$ can be used for both $X$ and $Y$ in the row-wise sum constraint, then as $x$ would be selected $T_1[{:},3{:}6]$ and as $y$ would be $T_1[{:},7]$.
+If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = RANK(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and a single vector $x$ in $G_1$ such that $x$ is ranked by $y$.
+For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}8]$ serves as $X$ and $Y$ and $x$ can be picked as $T_1[{:},7]$ and $y$ would be $T_1[{:},8]$.
 
 \subsection{Constraints}
-\sergey{we need to explain that constraints order is a partial order, which is a DAG}
+The set of Excel constraints has a partial order in which they should be learned.
+For certain constraints this order does not matter, such as rank or product.
+For many others they have to be learned in the order, which is specified as a DAG in Figure \ref{fig:learning_order} \sergey{add the figure here}.
+This figure specifies a potential parallelization of the computation, since each connected component is independent of the others.
+
 \begin{table}
@@
 
+\subsection{Workflow}
 \begin{algorithm}[thb]

### 2016-04-08 Paper/ECAI/tabular.tex
added the table

@@
 We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
+We denote as the \textit{length} of a row (column) group $G$, written as \textit{length(G)}, the number of its columns (rows).
+We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if its vectors contain numeric (textual, etc) elements.
 
@@
 the length of vectors in the arguments must be equal, i.e., these are constraints on the group meta-information not on the actual group content; \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function.
-For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}; its \CSignature is: $X$ is numeric, $Y$ is integer and the length of vectors in $X$ is the same as in $Y$; its \CFunction is the following constraint: a pair of vectors $x,y$ is a solution iff $x \in X, y \in Y$ and each value in $x$ has the rank (possibly with ties) specified in $y$.
+For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
 
@@
 For certain constraints this order does not matter, such as rank or product.
-For many others they have to be learned in the order, which is specified as a DAG in Figure \ref{fig:learning_order} \sergey{add the figure here}.
+For many others they have to be learned in the order, which is specified as a DAG in Figure \ref{fig:learning_order} (if a constraint is not in the graph, it is independent).
 This figure specifies a potential parallelization of the computation, since each connected component is independent of the others.
 
-\begin{table}
+\begin{figure}[htb]
   \centering
-  \begin{tabular}{lcc}
+  \includegraphics[width=0.20\textwidth]{figures/constraint_dependency.png}
+  \caption{Constraint Learning Order.
+The \textcolor{green}{solid} arrows indicate a construction order, i.e., learning one constraint is necessary to build the other.
+The \textcolor{red}{dashed} arrows indicate a pruning order, the former constraint is used to prune the latter.}
+  \label{fig:learning_order}
+\end{figure}
+
+\newcommand{\numeric}{\format{numeric}}
+\newcommand{\textual}{\format{textual}}
+\newcommand{\integer}{\format{integer}}
+\newcommand{\length}{\format{length}}
+\newcommand{\nat}{\mathcal{N}}
+
+\begin{table*}
+  \centering
+  \begin{tabular}{ccc}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ 
+    Y = RANK(X)     & $\numeric(G_X) \wedge \integer(G_Y) \wedge \length(G_X) = \length(G_Y)$ & $\begin{array}{r} X \in G_X, Y \in G_Y{:}~\forall i,j \in \nat{:}~(Y_i \leq Y_j \rightarrow X_{Y_i} \leq X_{Y_j}) \\ \wedge~ 1 \leq Y_i \leq \length(G_Y) \end{array}$ \\
+    ALLDIFFERENT(X) & $\textual(G_X) \vee \integer(G_X) $ & $X \in G_X{:}~\forall i,j \in \nat{:}~i \neq j \rightarrow X[i] \neq X[j]$ \\
   \end{tabular}
-  \caption{Tabular Constraints}
+  \caption{Excel Tabular Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}
   \label{table:constraints}
-\end{table}
+\end{table*}
 

### 2016-04-08 Paper/ECAI/tabular.tex
refactoring the paper

@@
 
-\sergey{to himself we need structure here} 
-\begin{figure}[thpb]
+\sergey{need to elaborate the example here: like in a story}
+
+\begin{figure*}[tbh]
   \begin{center}
-    \includegraphics[width=0.51\textwidth]{figures/Demo.png}
+    \includegraphics[width=0.75\textwidth]{figures/Demo.png}
   \end{center}
@@
   \label{fig:main_example}
-\end{figure}
+\end{figure*}
 
@@
 
-An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}, where \textit{\CName} is a textual name of the constraint; \textit{\CSignature} is a set of constraints specifying the properties of the arguments, such as their types, e.g., to require them to be integers or sizes, e.g.
-the length of vectors in the arguments must be equal, i.e., these are constraints on the group meta-information not on the actual group content; \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function.
-For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
+An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
+Let us elaborate on each of them.
+\textit{\CName} is the textual name of the constraint together with its variable names.
+\textit{\CSignature} is a set of constraints specifying the properties of the group assignments, corresponding to the variables in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g.
+the length of vectors in the arguments must be equal.
+These are constraints on the group meta-information not on the actual group content.
+\textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function.
+For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
 
-\subsection{Constraint learning algorithm description}
+\section{Problem Statement}
+
+\section{Approach to Tabular Constraint Learning}
 Let us describe the key steps and functions in Algorithm \ref{algo:tcl}.

### 2016-04-08 Code/src/workflow.py
Added equal and redundancy testing

@@ -10,6 +10,8 @@ from engine.minizinc import MinizincAssignmentStrategy, MinizincSolvingStrategy
 from parse.parser import get_groups_tables
 
 constraint_list = [
+    Equal(),
+    # EqualGroup(),
     Permutation(),
     Series(),
     AllDifferent(),

### 2016-04-08 Paper/ECAI/tabular.tex

@@
 \usepackage{notoccite}
+\usepackage{framed}
 
@@
 \newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
+
+\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
+\newcommand{\format}[1]{\textit{#1}\xspace}
+\newcommand{\generategroups}{\format{generateAssignments}}
+\newcommand{\extractgroups}{\format{extractGroups}}
+\newcommand{\extracttables}{\format{extractTables}}
+\newcommand{\learnconstraints}{\format{learnConstraints}}
+\newcommand{\findassignment}{\format{findSolutions}}
+\newcommand{\postprocess}{\format{pruneRedundant}}
+
+\newcommand{\CName}{Name\xspace}
+\newcommand{\CSignature}{Signature\xspace}
+\newcommand{\CFunction}{Function\xspace}
+\newcommand{\dependecies}{\ensuremath{\mathcal{D}}\xspace}
+\newcommand{\groups}{\ensuremath{\mathcal{G}}\xspace}
 
@@
 
-\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
-\newcommand{\format}[1]{\textit{#1}\xspace}
-\newcommand{\generategroups}{\format{generateAssignments}}
-\newcommand{\extractgroups}{\format{extractGroups}}
-\newcommand{\extracttables}{\format{extractTables}}
-\newcommand{\learnconstraints}{\format{learnConstraints}}
-\newcommand{\findassignment}{\format{findSolutions}}
-\newcommand{\postprocess}{\format{pruneRedundant}}
-
-\newcommand{\CName}{Name\xspace}
-\newcommand{\CSignature}{Signature\xspace}
-\newcommand{\CFunction}{Function\xspace}
-
 \begin{algorithm}[thb]
@@
 \section{Problem Statement}
+In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
+Here we formalize the statement in terms of Excel constraints and group assignments.
+
+\begin{framed}
+  \begin{tabular}{ll}
+
+    \multicolumn{2}{c}{{\textbf{Tabular Constraint Learning Problem}}}\\
+    &\\
+    \textbf{Given:} & The set of all groups $\groups$ and the set of Excel  \\
+    \phantom{Given:} & constraints $\constraints$ with its dependencies DAG $\dependecies$\\
+    \textbf{Find: } & text
+  \end{tabular}
+\end{framed}
 

### 2016-04-08 Code/src/core/constraint.py
Added order

@@ -9,11 +9,12 @@ from core.group import GType, Group, Orientation
 
 
 class Constraint:
-    def __init__(self, name, print_format, source: Source, filters: List[Filter]):
+    def __init__(self, name, print_format, source: Source, filters: List[Filter], depends_on=set()):
         self.name = name
         self.print_format = print_format
         self.source = source
         self._filters = filters
+        self._depends_on = set.union(depends_on, source.depends_on())
 
     @property
     def filters(self):
@@ -22,6 +23,9 @@ class Constraint:
     @property
     def variables(self):
         return self.source.variables
+
+    def depends_on(self):
+        return self._depends_on
 
     def get_variables(self):
         return self.variables
@@ -119,7 +123,8 @@ class Aggregate(Constraint):
         size_filter = SizeFilter([self.x], rows=2) if Orientation.column(orientation) else SizeFilter([self.x], cols=2)
         filters = [size_filter, filter_class(variables)]
         format_s = "{Y} = " + op_string.upper() + "({X}, " + or_string + ")"
-        super().__init__("{} ({})".format(op_string.lower(), or_string), format_s, Source(variables), filters)
+        name = "{} ({})".format(op_string.lower(), or_string)
+        super().__init__(name, format_s, Source(variables), filters, {Equal(), Projection()})
 
     @property
     def orientation(self):

### 2016-04-08 Code/src/workflow.py

@@ -10,7 +10,6 @@ from engine.minizinc import MinizincAssignmentStrategy, MinizincSolvingStrategy
 from parse.parser import get_groups_tables
 
 constraint_list = [
-    Equal(),
     # EqualGroup(),
     Permutation(),
     Series(),
@@ -30,7 +29,26 @@ constraint_list += [
     ForeignProduct(),
     Product(),
     SumProduct(),
+    Equal(),
 ]
+
+
+def order_constraints(constraints: List[Constraint]):
+    ordered = []
+    found = set()
+    spill = constraints
+    while len(spill) > 0:
+        new_spill = []
+        for constraint in constraints:
+            if constraint.depends_on().issubset(found):
+                ordered.append(constraint)
+                found.add(constraint)
+            else:
+                spill.append(constraint)
+        if len(new_spill) == len(spill):
+            raise Exception("Dependency order is not a DAG")
+        spill = new_spill
+    return ordered
 
 
 def main(csv_file, groups_file, verbose, silent=False, parse_silent=False, constraints=constraint_list):
@@ -58,7 +76,9 @@ def main(csv_file, groups_file, verbose, silent=False, parse_silent=False, const
     if (len(unsupported_assignment) > 0 or len(unsupported_solving) > 0) and not silent:
         print()
 
-    for constraint in supported:
+    ordered = order_constraints(supported)
+
+    for constraint in ordered:
         t_start = time.time()
         assignments = manager.find_assignments(constraint, groups, solutions)
         t_assign = time.time()

### 2016-04-08 Code/src/core/constraint.py
Comment

@@ -124,6 +124,7 @@ class Aggregate(Constraint):
         filters = [size_filter, filter_class(variables)]
         format_s = "{Y} = " + op_string.upper() + "({X}, " + or_string + ")"
         name = "{} ({})".format(op_string.lower(), or_string)
+        # TODO Dependency only min max average
         super().__init__(name, format_s, Source(variables), filters, {Equal(), Projection()})
 
     @property

### 2016-04-08 Paper/ECAI/tabular.tex
writing the problem statement
Constraint table expanded

@@
     \phantom{Given:} & constraints $\constraints$ with its dependencies DAG $\dependecies$\\
-    \textbf{Find: } & text
+    \textbf{Find:} & For each constraint $c$ in \constraints all the subgroups of $\groups$ satisfying the \CSignature and \CFunction of $c$
   \end{tabular}
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
+    ALLDIFFERENT(X) & $\textual(G_X) \vee \integer(G_X) $ & $X \in G_X{:}~\forall i,j \in \nat{:}~i \neq j \rightarrow X[i] \neq X[j]$ \\
+    X = Y & & \\
+    FK -> PK & & \\
+    R = PRODUCT(FV, FK=PK | PV) & & \\
+    FV = FUZZY-LOOKUP(FK, PK, PV) & & \\
+    FV = LOOKUP(FK, PK, PV) & & \\
+    PERMUTATION(X) & & \\
+    R = O1 * O2 & & \\
+    R = PROJECT(P) & & \\
      Y = RANK(X)     & $\numeric(G_X) \wedge \integer(G_Y) \wedge \length(G_X) = \length(G_Y)$ & $\begin{array}{r} X \in G_X, Y \in G_Y{:}~\forall i,j \in \nat{:}~(Y_i \leq Y_j \rightarrow X_{Y_i} \leq X_{Y_j}) \\ \wedge~ 1 \leq Y_i \leq \length(G_Y) \end{array}$ \\
-    ALLDIFFERENT(X) & $\textual(G_X) \vee \integer(G_X) $ & $X \in G_X{:}~\forall i,j \in \nat{:}~i \neq j \rightarrow X[i] \neq X[j]$ \\
+    A = PREV(A) + P - N & & \\
+    SERIES(X) & & \\
+    Y = SUM(X, col) & & \\
+    Y = SUM(X, row) & & \\
+    R = SUMIF(FK=PK, V) & & \\
+    R = SUMPRODUCT(O1, O2) & & \\
+
   \end{tabular}

### 2016-04-08 Paper/ECAI/tabular.tex
writing the paper

--- 2016-04-08 Paper/ECAI/tabular.tex
@@
 \newcommand{\postprocess}{\format{pruneRedundant}}
+\newcommand{\constrainttorder}{\format{orderedConstraints}}
 
@@
 \newcommand{\CFunction}{Function\xspace}
-\newcommand{\dependecies}{\ensuremath{\mathcal{D}}\xspace}
+\newcommand{\dependencies}{\ensuremath{\mathcal{D}}\xspace}
 \newcommand{\groups}{\ensuremath{\mathcal{G}}\xspace}
@@
 
-\begin{algorithm}[thb]
-  \begin{algorithmic}
-    \footnotesize
-    \State \textbf{Input:} $G$ -- set of groups
-    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
-    \State $S \gets \emptyset$ \Comment{The set of solutions}
-  \For{$c \in \constraints$} \Comment{\constraints -- partially ordered set of Excel constraints; Table \ref{table:constraints}}
-  \State $v_1, ..., v_n =$ variables of $c$ 
-    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S)$}
-      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
-    \EndFor
-  \EndFor\\
-\Return $S$
-\end{algorithmic}
-\caption{Tabular constraint learning}
-\label{algo:tcl}
-\end{algorithm}
 \subsection{Groups, Type-consistency and Constraints}
@@
 
-\begin{framed}
+\begin{minipage}[c]{14em}
+  \vspace{5pt}
   \begin{tabular}{ll}
-
-    \multicolumn{2}{c}{{\textbf{Tabular Constraint Learning Problem}}}\\
+    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
+    \vspace{-4pt}
     &\\
-    \textbf{Given:} & The set of all groups $\groups$ and the set of Excel  \\
-    \phantom{Given:} & constraints $\constraints$ with its dependencies DAG $\dependecies$\\
-    \textbf{Find: } & text
+    \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$\\
+    \textbf{Find:}& for each constraint  $c$ in \constraints all the subgroups of $\groups$\\ 
+    & satisfying the \CSignature and \CFunction of $c$ 
   \end{tabular}
-\end{framed}
+\end{minipage}
 
 \section{Approach to Tabular Constraint Learning}
+\begin{algorithm}[thb]
+  \begin{algorithmic}
+    \footnotesize
+    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Excel constraints, \dependencies -- dependency graph 
+    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
+    \State $S \gets \emptyset$ \Comment{The set of solutions}
+  \For{$c \in \constrainttorder(\constraints,\dependencies)$} 
+  \State $v_1, ..., v_n =$ variables of $c$ 
+    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S)$}
+      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
+    \EndFor
+  \EndFor\\
+\Return $S$
+\end{algorithmic}
+\caption{Tabular constraint learning}
+\label{algo:tcl}
+\end{algorithm}
+
 Let us describe the key steps and functions in Algorithm \ref{algo:tcl}.
-Essentially, the algorithm has two steps: candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}.
+Essentially, the algorithm has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}.
 Let us elaborate on each step in detail.
+
+\paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ \sergey{explaination that solutions to certain constraints allows us to prune solutions to other constraints -- subsumption, condensed representation}
 
@@
 The \textcolor{green}{solid} arrows indicate a construction order, i.e., learning one constraint is necessary to build the other.
-The \textcolor{red}{dashed} arrows indicate a pruning order, the former constraint is used to prune the latter.}
+The \textcolor{red}{dashed} arrows indicate a subsumption order, the former constraint is used to prune the latter.}
   \label{fig:learning_order}
@@
     \footnotesize
-    \State \textbf{Input:} $D$ -- dataset, (optional: tables $T$, groups $G$)
+    \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
@@
     \EndIf
-    \State $S \gets \learnconstraints(G)$
+    \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
     \State $S \gets \postprocess(S)$

### 2016-04-10 Paper/ECAI/tabular.tex
Corrections and case study evaluation

@@
 \usepackage{framed}
+\usepackage{amsmath}
 
@@
 \newcommand{\groups}{\ensuremath{\mathcal{G}}\xspace}
+
+\newcommand{\range}[3]{\ensuremath{#1[#2,#3]}}
+\newcommand{\rangeto}[2]{#1{:}#2}
+\newcommand{\rangeall}{:}
+
+\newcommand{\ecrank}[2]{\ensuremath{#1 = \mathit{RANK}(#2)}}
 
@@
   \centering
-  \begin{tabular}{ccc}
+  \begin{tabular}{lcc}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
@@
 
-\section{Case Study aka Experiments}
-
 \textbf{Approach}
@@
 \end{itemize}
+
+\section{Case Study aka Experiments}
+In this section we experimentally validate our approach.
+The implementation is first illustrated using a case study on a specific Excel sheet.
+In order to quantify the results and generalize our findings we created a benchmark of 30 (\samuel{check number}) Excel sheets from various sources.
+
+\subsection{Benchmark}
+There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online, publicly available spreadsheets such as crime statistics or financial reports\footnote{\samuel{link to links.txt}}.
+
+All spreadsheets have been converted to CSV with no manual intervention unless noted.
+Definitions of tables where added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).
+
+For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expressible in Excel and expected to be discovered.
+\footnote{Constraints currently outside of the scope of this system (e.g. generic nested mathematical formulas) are in a separate category (non-trivial).}
+
+\subsection{Case study}
+Let us illustrate our implementation using the example presented in the introduction.
+We obtained this example by combining several Excel sheets from exercise sessions into one.
+Figure~\ref{fig:sol_example} shows the constraints that we expect to find.
+
+\paragraph{Accuracy}
+Our current implementation finds 18 constraints, containing all of the 12 solutions described in figure~\ref{fig:sol_example}.
+Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}
+  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5})} \\
+  & \ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{4}}
+\end{align*}
+The other redundant constraint is a additional $\mathit{LOOKUP}$ that holds because the vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).
+
+We conclude that for this example our primary goal of finding all constraints is achieved.
+A number of redundant constraints ($33.33\%$ of the total) are also found, which, as we will show, is rather high compared to other spreadsheets.
+However, this example also contains many short vectors which increases the chance for constraints to be true by accident.
+
+\begin{figure}
+  \label{fig:sol_example}
+  {\small
+    \begin{align*}
+      & SERIES(\range{T_{1}}{\rangeall}{1}) \\
+      & \range{T_{1}}{\rangeall}{8} = RANK(\range{T_{1}}{\rangeall}{7}) \\
+      & \range{T_{2}}{1}{\rangeall} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
+      & \range{T_{6}}{\rangeall}{2} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{6}}, col) \\
+      & \range{T_{1}}{\rangeall}{7} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{6}}, row) \\
+      & \range{T_{2}}{2}{\rangeall} = AVERAGE(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
+      & \range{T_{2}}{3}{\rangeall} = MAX(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
+      & \range{T_{2}}{4}{\rangeall} = MIN(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
+      & \range{T_{1}}{\rangeall}{10} = SUMIF(\range{T_{5}}{\rangeall}{1}=\range{T_{1}}{\rangeall}{2}, \range{T_{5}}{\rangeall}{2}) \\
+      & \range{T_{1}}{\rangeall}{11} = MAXIF(\range{T_{5}}{\rangeall}{1}=\range{T_{1}}{\rangeall}{2}, \range{T_{5}}{\rangeall}{2}) \\
+      & \range{T_{4}}{\rangeall}{3} = LOOKUP(\range{T_{4}}{\rangeall}{2}, \range{T_{1}}{\rangeall}{1}, \range{T_{1}}{\rangeall}{2}) \\
+      & \range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
+    \end{align*}
+  }%
+  \caption{Expected constraints case study}
+\end{figure}
 

### 2016-04-10 Code/src/core/solutions.py

@@ -15,6 +15,9 @@ class Solutions:
     def has_solution(self, constraint, solution):
         return self._to_tuple(constraint, solution) in self.properties[constraint]
 
+    def has(self, constraint, keys, values):
+        return self.has_solution(constraint, {k.name: v for k, v in zip(keys, values)})
+
     @staticmethod
     def _to_tuple(constraint, solution):
         return tuple(solution[v.name] for v in constraint.variables)

### 2016-04-10 Code/src/engine/internal.py

@@ -119,6 +119,7 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             return solutions
 
         def lookups(c: Lookup, assignments, solutions):
+            # TODO redundant lookups
             results = []
 
             def is_lookup(reference_dictionary, keys, values):

### 2016-04-10 Paper/ECAI/tabular.tex
filling in the structure and writing the story

@@
 In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
-Here we formalize the statement in terms of Excel constraints and group assignments.
+Here we formalize the statement in terms of Excel constraints and group assignmentsd as follows:
 
@@
     \textbf{Find:}& for each constraint  $c$ in \constraints all the subgroups of $\groups$\\ 
-    & satisfying the \CSignature and \CFunction of $c$ 
+    & satisfying the \CSignature and \CFunction of $c$ \\
   \end{tabular}
+  \vspace{6pt}
 \end{minipage}
+
+  The key observation here is that essentially the problem is a constraint enumeration problem, where each constraint is independent.
+This property comes from fact that in Excel each formula is applied based on the existing data in the cells.
+This allows learning constraints independently of each other by examining constraint satisfaction on the groups.
 
@@
   \State $v_1, ..., v_n =$ variables of $c$ 
-    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S)$}
+    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
       \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
@@
 \end{algorithmic}
-\caption{Tabular constraint learning}
+\caption{Tabular constraint learning \sergey{Samuel, I think \generategroups takes \dependencies as an argument}}
 \label{algo:tcl}
 \end{algorithm}
+In theory each constraint is independent and equally useful.
+In practice, however, things are different.
+Assume that for some $X$ and $Y$ it holds that \textit{SUM(X)=Y}, and also we discovered that the vector $Y$ is equal to a vector $Z$.
+Then, of course, \textit{SUM(X)=Z} holds and it is a valid solution to the sum constraint problem.
+From a user's perspective the last solution is useless, since he or she already knows that $Y=Z$ and the latter constraint satisfaction follows.
+To capture this condensed representation \cite{condensed} of solutions, we have introduced a subsumption (generality) order between constraints, which is depicted using \textcolor{red}{dashed} lines in Figure \ref{fig:learning_order}.
+Besides generality, we know that certain constraints have as a subproblem other constraints, e.g.
+if $X$ is a \textit{series}, then it is also a \textit{permutation}.
+If discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all.
+To capture this solving strategy we introduced a constraint construction order depicted using \textcolor{green}{solid} lines in Figure \ref{fig:learning_order}.
 
-Let us describe the key steps and functions in Algorithm \ref{algo:tcl}.
-Essentially, the algorithm has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}.
+Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}.
 Let us elaborate on each step in detail.
 
-\paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ \sergey{explaination that solutions to certain constraints allows us to prune solutions to other constraints -- subsumption, condensed representation}
+\paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
+Also, if there is a construction edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
+If there is a sumsumption edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
 
@@
 \subsection{Constraints}
-The set of Excel constraints has a partial order in which they should be learned.
-For certain constraints this order does not matter, such as rank or product.
+As emphasized before, the set of Excel constraints has a partial order in which they should be learned.
+For certain constraints this order does not matter, such as \textit{rank} or \textit{product}.
 For many others they have to be learned in the order, which is specified as a DAG in Figure \ref{fig:learning_order} (if a constraint is not in the graph, it is independent).
-This figure specifies a potential parallelization of the computation, since each connected component is independent of the others.
+This figure also specifies a potential parallelization of the computation, since each connected component is independent of the others.
+
+\sergey{need to elaborate here especially on the structure and how the constraints we picked}
 
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
-    ALLDIFFERENT(X) & $\textual(G_X) \vee \integer(G_X) $ & $X \in G_X{:}~\forall i,j \in \nat{:}~i \neq j \rightarrow X[i] \neq X[j]$ \\
+    ALLDIFFERENT(X) & & \\
     X = Y & & \\
-    FK -> PK & & \\
+    FK $\rightarrow$ PK & & \\
     R = PRODUCT(FV, FK=PK | PV) & & \\
@@
     R = PROJECT(P) & & \\
-     Y = RANK(X)     & $\numeric(G_X) \wedge \integer(G_Y) \wedge \length(G_X) = \length(G_Y)$ & $\begin{array}{r} X \in G_X, Y \in G_Y{:}~\forall i,j \in \nat{:}~(Y_i \leq Y_j \rightarrow X_{Y_i} \leq X_{Y_j}) \\ \wedge~ 1 \leq Y_i \leq \length(G_Y) \end{array}$ \\
+     Y = RANK(X)   & & \\
     A = PREV(A) + P - N & & \\
@@
 
+\section{Declarative modeling}
+\sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
+
 \section{Case Study aka Experiments}
@@
 
+\sergey{I think there should be three things: key example evaluated, benchmark of 30 spreadsheets and maybe the FBI example, how we can detect interesting dependecies and correct potential mistakes}
+
 \paragraph{Accuracy}
@@
 We conclude that for this example our primary goal of finding all constraints is achieved.
-A number of redundant constraints ($33.33\%$ of the total) are also found, which, as we will show, is rather high compared to other spreadsheets.
+A number of redundant constraints ($\frac{1}{3}$ of the total) are also found, which, as we will show, is rather high compared to other spreadsheets.
 However, this example also contains many short vectors which increases the chance for constraints to be true by accident.

### 2016-04-11 Code/src/accuracy.py
Fixed conditional aggregate

@@ -18,7 +18,8 @@ def is_excel_constraint(c: Constraint):
     return isinstance(c, Aggregate) or isinstance(c, ConditionalAggregate) or isinstance(c, Series) \
         or isinstance(c, Rank) or isinstance(c, Lookup) or isinstance(c, FuzzyLookup) or isinstance(c, RunningTotal) \
         or isinstance(c, Product) or isinstance(c, Diff) or isinstance(c, SumProduct) \
-        or isinstance(c, ForeignProduct) or isinstance(c, Equal) or isinstance(c, PercentualDiff)
+        or isinstance(c, ForeignProduct) or isinstance(c, Equal) or isinstance(c, PercentualDiff) \
+        or isinstance(c, Projection)
 
 
 constraint_map = {c.name: c for c in workflow.constraint_list}

### 2016-04-11 Code/src/core/constraint.py
Fixed conditional aggregate

@@ -83,16 +83,19 @@ class Operation(Enum):
                 data = data.T
             if not data.shape:
                 data = numpy.array([[data]])
-            rows, cols = data.shape
-            blank, blank_test = blank_filter(data)
-            results = numpy.array([blank] * cols)
-            mask = numpy.vectorize(blank_test)
-            for i in range(cols):
-                vec = data[:, i][mask(data[:, i])]
-                if len(vec) != 0:
-                    vec = numpy.array(vec, dtype=numpy.float64)
-                    results[i] = self._aggregate(vec)
-            return results
+            blank, blank_test = blank_filter(data, vectorized=True)
+            if len(data.shape) > 1:
+                rows, cols = data.shape
+                results = numpy.array([blank] * cols)
+                for i in range(cols):
+                    vec = data[:, i][blank_test(data[:, i])]
+                    if len(vec) != 0:
+                        vec = numpy.array(vec, dtype=numpy.float64)
+                        results[i] = self._aggregate(vec)
+                return results
+            else:
+                array = numpy.array(data[blank_test(data)], dtype=numpy.float64)
+                return self._aggregate(array) if len(array) > 0 else blank
 
         return apply
 

### 2016-04-11 Code/src/engine/internal.py

@@ -176,18 +176,12 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             def is_aggregate(ok_v, r_v, fk_v, v_v):
                 if not all(all(numpy.vectorize(blank_filter(v)[1])(v)) for v in [ok_v, r_v, v_v]):
                     return False
-                blank_f = blank_filter(fk_v)[1]
-                if not blank_f(fk_v[0]):
-                    return False
-                m = dict(zip(ok_v, range(len(ok_v))))
-                acc = [None] * len(r_v)
-                for i in range(len(fk_v)):
-                    if blank_f(fk_v[i]):
-                        key = m[fk_v[i]]
-                        aggregated = c.operation.aggregate(v_v[i])
-                        acc[key] = aggregated if acc[key] is None else c.operation.func(acc[key], aggregated)
-                acc = [c.default if acc[i] is None else acc[i] for i in range(len(acc))]
-                return all(equal(r_v[i], acc[i]) for i in range(len(r_v)))
+                for i in range(len(ok_v)):
+                    data = v_v[fk_v == ok_v[i]]
+                    res = c.operation.aggregate(data) if len(data) > 0 else c.default
+                    if not equal(res, r_v[i], True):
+                        return False
+                return True
 
             return list(self._generate_test_vectors(assignments, keys, is_aggregate))
 

### 2016-04-11 Code/src/core/constraint.py
Broader conditional aggregate

@@ -244,10 +244,11 @@ class ConditionalAggregate(Constraint):
         self._operation = operation
         name = operation.name
         variables = [self.o_key, self.result, self.f_key, self.values]
-        foreign_key = ForeignKey()
-        source = ConstraintSource(variables, foreign_key, {foreign_key.pk.name: "OK", foreign_key.fk.name: "FK"})
+        all_diff = AllDifferent()
+        source = ConstraintSource(variables, all_diff, {all_diff.x.name: "OK"})
         filters = [SameLength([self.o_key, self.result]), SameLength([self.f_key, self.values]),
-                   SameTable([self.f_key, self.values]),
+                   SameTable([self.f_key, self.values]), Not(SameTable([self.f_key, self.o_key])),
+                   NotPartial([self.o_key]), SameType([self.f_key, self.o_key]),
                    SameOrientation([self.o_key, self.result]), SameOrientation([self.f_key, self.values])]
         super().__init__("{}-if".format(name.lower()), "{R} = " + name.upper() + "IF({FK}={OK}, {V})", source, filters)
 

### 2016-04-11 Code/src/engine/internal.py

@@ -178,7 +178,10 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     return False
                 for i in range(len(ok_v)):
                     data = v_v[fk_v == ok_v[i]]
-                    res = c.operation.aggregate(data) if len(data) > 0 else c.default
+                    try:
+                        res = c.operation.aggregate(data) if len(data) > 0 else c.default
+                    except TypeError:
+                        print(v_v, v_v[fk_v == "Management"], fk_v, ok_v[i])
                     if not equal(res, r_v[i], True):
                         return False
                 return True

### 2016-04-11 Paper/ECAI/tabular.tex

@@
 Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}
-  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5})} \\
+  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}} \\
   & \ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{4}}

### 2016-04-11 Code/src/engine/internal.py
Removed debugging message

@@ -178,10 +178,7 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     return False
                 for i in range(len(ok_v)):
                     data = v_v[fk_v == ok_v[i]]
-                    try:
-                        res = c.operation.aggregate(data) if len(data) > 0 else c.default
-                    except TypeError:
-                        print(v_v, v_v[fk_v == "Management"], fk_v, ok_v[i])
+                    res = c.operation.aggregate(data) if len(data) > 0 else c.default
                     if not equal(res, r_v[i], True):
                         return False
                 return True

### 2016-04-11 Code/src/core/constraint.py
Removed artifacts

@@ -250,7 +250,8 @@ class ConditionalAggregate(Constraint):
                    SameTable([self.f_key, self.values]), Not(SameTable([self.f_key, self.o_key])),
                    NotPartial([self.o_key]), SameType([self.f_key, self.o_key]),
                    SameOrientation([self.o_key, self.result]), SameOrientation([self.f_key, self.values])]
-        super().__init__("{}-if".format(name.lower()), "{R} = " + name.upper() + "IF({FK}={OK}, {V})", source, filters)
+        p_format = "{R} = " + name.upper() + "IF({FK}={OK}, {V})"
+        super().__init__("{}-if".format(name.lower()), p_format, source, filters, depends_on={ForeignKey()})
 
     @property
     def operation(self) -> Operation:

### 2016-04-11 Code/src/engine/internal.py

@@ -170,7 +170,7 @@ class InternalSolvingStrategy(DictSolvingStrategy):
 
             return self._generate_test_vectors(assignments, keys, is_lookup)
 
-        def conditional_aggregate(c: ConditionalAggregate, assignments, solutions):
+        def conditional_aggregate(c: ConditionalAggregate, assignments, solutions: Solutions):
             keys = [c.o_key, c.result, c.f_key, c.values]
 
             def is_aggregate(ok_v, r_v, fk_v, v_v):
@@ -183,7 +183,12 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                         return False
                 return True
 
-            return list(self._generate_test_vectors(assignments, keys, is_aggregate))
+            def not_inverted(ok, r, fk, v):
+                foreign_key = ForeignKey()
+                return not solutions.has(foreign_key, [foreign_key.fk, foreign_key.pk], [ok, fk]) \
+                    and not solutions.has(foreign_key, [foreign_key.fk, foreign_key.pk], [r, v])
+
+            return list(self._generate_test_vectors(assignments, keys, is_aggregate, not_inverted))
 
         def running_total(c: RunningTotal, assignments, solutions):
             def is_running_diff(acc, pos, neg):

### 2016-04-11 Code/src/core/assignment.py
Allowed product aggregate

@@ -1,5 +1,6 @@
 from typing import List, Dict, Set
 
+import itertools
 from constraint import Problem
 from core.group import Group
 from parse.parser import GType

### 2016-04-11 Code/src/core/constraint.py

@@ -111,6 +111,7 @@ class Aggregate(Constraint):
     def __init__(self, orientation: Orientation, operation: Operation):
         self._orientation = orientation
         self._operation = operation
+        self.min_size = 3 if operation == Operation.PRODUCT else 2
         size = Group.columns if orientation == Orientation.VERTICAL else Group.rows
         or_string = "col" if orientation == Orientation.VERTICAL else "row"
         op_string = operation.name
@@ -123,8 +124,9 @@ class Aggregate(Constraint):
 
         filter_class = type("{}{}Length".format(op_string.lower().capitalize(), or_string.capitalize()),
                             (Filter,), {"test": test})
-        size_filter = SizeFilter([self.x], rows=2) if Orientation.column(orientation) else SizeFilter([self.x], cols=2)
-        filters = [size_filter, filter_class(variables)]
+        x_size_filter = SizeFilter([self.x], rows=self.min_size) if Orientation.column(orientation)\
+            else SizeFilter([self.x], cols=self.min_size)
+        filters = [x_size_filter, filter_class(variables)]
         format_s = "{Y} = " + op_string.upper() + "({X}, " + or_string + ")"
         name = "{} ({})".format(op_string.lower(), or_string)
         # TODO Dependency only min max average
@@ -144,7 +146,7 @@ class Aggregate(Constraint):
 
     @classmethod
     def instances(cls):
-        return list(cls.instance(o, op) for o in Orientation for op in Operation if not op == Operation.PRODUCT)
+        return list(cls.instance(o, op) for o in Orientation for op in Operation)  # if not op == Operation.PRODUCT)
 
 
 # TODO Same table, different orientation, overlapping bounds => prune assignment already

### 2016-04-11 Code/src/engine/internal.py

@@ -257,10 +257,10 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     max_range = MaxRange(check)
                     for y_i in range(y_group.vectors()):
                         if y_group == x_group:
-                            max_range.find(0, y_i, 2)
-                            max_range.find(y_i + 1, x_group.vectors(), 2)
+                            max_range.find(0, y_i, c.min_size)
+                            max_range.find(y_i + 1, x_group.vectors(), c.min_size)
                         else:
-                            max_range.find(0, x_group.vectors(), 2)
+                            max_range.find(0, x_group.vectors(), c.min_size)
 
             return results
 

### 2016-04-11 Code/src/core/assignment.py
Allowed product aggregate

@@ -2,7 +2,7 @@ from typing import List, Dict, Set
 
 import itertools
 from constraint import Problem
-from core.group import Group
+from core.group import Group, Orientation
 from parse.parser import GType
 
 
@@ -115,6 +115,10 @@ class Filter:
         groups = list([assignment[v.name] for v in self.variables])
         return all(f(groups[i]) == f(groups[j]) for i in range(len(groups)) for j in range(i + 1, len(groups)))
 
+    def test_all(self, assignment, f):
+        groups = list([assignment[v.name] for v in self.variables])
+        return all(f(g) for g in groups)
+
 
 class Not(Filter):
     def __init__(self, original_filter: Filter):
@@ -123,6 +127,35 @@ class Not(Filter):
 
     def test(self, assignment):
         return not self._original_filter.test(assignment)
+
+
+class If(Filter):
+    def __init__(self, if_filter: Filter, then_filter: Filter, else_filter: Filter=None):
+        if else_filter is None:
+            else_filter = NoFilter([])
+        variables = set(if_filter.variables).union(set(then_filter.variables)).union(set(else_filter.variables))
+        super().__init__(variables)
+        self._if_filter = if_filter
+        self._then_filter = then_filter
+        self._else_filter = else_filter
+
+    @property
+    def if_filter(self):
+        return self._if_filter
+
+    @property
+    def then_filter(self):
+        return self._then_filter
+
+    @property
+    def else_filter(self):
+        return self._else_filter
+
+    def test(self, assignment: Dict[str, Group]):
+        if self.if_filter.test(assignment):
+            return self.then_filter.test(assignment)
+        else:
+            return self.else_filter.test(assignment)
 
 
 class NoFilter(Filter):
@@ -166,6 +199,19 @@ class SizeFilter(Filter):
             and all(self._cols is None or op(g.columns(), self._cols) for g in groups) \
             and all(self._length is None or op(g.length(), self._length) for g in groups) \
             and all(self._vectors is None or op(g.vectors(), self._vectors) for g in groups)
+
+
+class OrientationFilter(Filter):
+    def __init__(self, variables, orientation):
+        self._orientation = orientation
+        super().__init__(variables)
+
+    @property
+    def orientation(self):
+        return self._orientation
+
+    def test(self, assignment: Dict[str, Group]):
+        return self.test_all(assignment, lambda g: g.row == (self.orientation == Orientation.HORIZONTAL))
 
 
 class NotPartial(Filter):

### 2016-04-11 Code/src/core/constraint.py

@@ -4,7 +4,7 @@ from typing import List, Dict
 import numpy
 
 from core.assignment import Source, Filter, Variable, SameLength, ConstraintSource, SameTable, \
-    SameOrientation, SameType, SizeFilter, Not, NotPartial, Partial
+    SameOrientation, SameType, SizeFilter, Not, NotPartial, Partial, If, OrientationFilter
 from core.group import GType, Group, Orientation
 
 
@@ -111,7 +111,8 @@ class Aggregate(Constraint):
     def __init__(self, orientation: Orientation, operation: Operation):
         self._orientation = orientation
         self._operation = operation
-        self.min_size = 3 if operation == Operation.PRODUCT else 2
+        self.min_size = 2
+        self.min_vectors = 3 if operation == Operation.PRODUCT else 2
         size = Group.columns if orientation == Orientation.VERTICAL else Group.rows
         or_string = "col" if orientation == Orientation.VERTICAL else "row"
         op_string = operation.name
@@ -120,12 +121,14 @@ class Aggregate(Constraint):
         def test(_, a: Dict[str, Group]):
             x_group, y_group = [a[v.name] for v in variables]
             o_match = x_group.row == (orientation == Orientation.HORIZONTAL)
+            if not o_match and x_group.vectors() < self.min_vectors:
+                return False
             return y_group.length() <= size(x_group) if o_match else y_group.length() == size(x_group)
 
         filter_class = type("{}{}Length".format(op_string.lower().capitalize(), or_string.capitalize()),
                             (Filter,), {"test": test})
-        x_size_filter = SizeFilter([self.x], rows=self.min_size) if Orientation.column(orientation)\
-            else SizeFilter([self.x], cols=self.min_size)
+        x_size_filter = SizeFilter([self.x], rows=self.min_size)\
+            if Orientation.column(orientation) else SizeFilter([self.x], cols=self.min_size)
         filters = [x_size_filter, filter_class(variables)]
         format_s = "{Y} = " + op_string.upper() + "({X}, " + or_string + ")"
         name = "{} ({})".format(op_string.lower(), or_string)

### 2016-04-11 Code/src/engine/internal.py

@@ -257,10 +257,10 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     max_range = MaxRange(check)
                     for y_i in range(y_group.vectors()):
                         if y_group == x_group:
-                            max_range.find(0, y_i, c.min_size)
-                            max_range.find(y_i + 1, x_group.vectors(), c.min_size)
+                            max_range.find(0, y_i, c.min_vectors)
+                            max_range.find(y_i + 1, x_group.vectors(), c.min_vectors)
                         else:
-                            max_range.find(0, x_group.vectors(), c.min_size)
+                            max_range.find(0, x_group.vectors(), c.min_vectors)
 
             return results
 

### 2016-04-11 Code/src/core/constraint.py
Display time

@@ -4,7 +4,7 @@ from typing import List, Dict
 import numpy
 
 from core.assignment import Source, Filter, Variable, SameLength, ConstraintSource, SameTable, \
-    SameOrientation, SameType, SizeFilter, Not, NotPartial, Partial, If, OrientationFilter
+    SameOrientation, SameType, SizeFilter, Not, NotPartial, Partial
 from core.group import GType, Group, Orientation
 
 

### 2016-04-11 Paper/ECAI/tabular.tex
added dependencies into findassignment
updatint the problem statement

@@
 \usepackage{hyperref}
+\usepackage{amssymb}
 \usepackage{algorithm}
@@
 
+A \textit{tabular solution} $S$ to a constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+
 \section{Problem Statement}
 In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
-Here we formalize the statement in terms of Excel constraints and group assignmentsd as follows:
+Here we formalize the statement in terms of Excel constraints and group assignments as follows:
 
@@
     \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$\\
-    \textbf{Find:}& for each constraint  $c$ in \constraints all the subgroups of $\groups$\\ 
-    & satisfying the \CSignature and \CFunction of $c$ \\
+    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ \constraints \\ 
   \end{tabular}
@@
 This allows learning constraints independently of each other by examining constraint satisfaction on the groups.
+
+In theory each constraint is independent and equally useful.
+In practice, however, things are different.
+Assume that for some $X$ both \textit{series(X)} and \textit{alldifferent(X)} hold.
+From a user's perspective the last solution is useless, since he or she already knows that series constraint implies alldifferent constraint.
+To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity (generality) order between constraints, which is depicted using solid (dashed) arrows in Figure \ref{fig:learning_order}.
+This matters from implementation perspective as well, if we discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all.
+A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is a more specific constraint than $C_2$ and therefore if $X$ is a solution for $C_1$ it should not be tested for $C_2$.
+
+We call a tabular solution $S$ to a constraint $C$ \textit{the most specific} wrt \dependencies, iff there is no other constraint $C'$ for which $S$ is a solution and $C'$ is more specific in \dependencies.
+
+Let us introduce the version of tabular constraint learning that takes into account the specificity relation between constraints.
+
+\begin{minipage}[c]{14em}
+  \vspace{5pt}
+  \begin{tabular}{ll}
+    \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
+    \vspace{-4pt}
+    &\\
+    \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$,\\ 
+    & a constraint specificity DAG \dependencies \\
+    \textbf{Find:}& all the most specific solution $S$ in \groups\\
+    & for each $c$ in \constraints \\
+  \end{tabular}
+  \vspace{6pt}
+\end{minipage}
 
@@
     \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
-      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S)$
+      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S, \dependencies)$
     \EndFor
@@
 \end{algorithmic}
-\caption{Tabular constraint learning \sergey{Samuel, I think \generategroups takes \dependencies as an argument}}
+\caption{Tabular constraint learning}
 \label{algo:tcl}
 \end{algorithm}
-In theory each constraint is independent and equally useful.
-In practice, however, things are different.
-Assume that for some $X$ and $Y$ it holds that \textit{SUM(X)=Y}, and also we discovered that the vector $Y$ is equal to a vector $Z$.
-Then, of course, \textit{SUM(X)=Z} holds and it is a valid solution to the sum constraint problem.
-From a user's perspective the last solution is useless, since he or she already knows that $Y=Z$ and the latter constraint satisfaction follows.
-To capture this condensed representation \cite{condensed} of solutions, we have introduced a subsumption (generality) order between constraints, which is depicted using \textcolor{red}{dashed} lines in Figure \ref{fig:learning_order}.
-Besides generality, we know that certain constraints have as a subproblem other constraints, e.g.
-if $X$ is a \textit{series}, then it is also a \textit{permutation}.
-If discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all.
-To capture this solving strategy we introduced a constraint construction order depicted using \textcolor{green}{solid} lines in Figure \ref{fig:learning_order}.
-
 Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}.
@@
   \caption{Constraint Learning Order.
-The \textcolor{green}{solid} arrows indicate a construction order, i.e., learning one constraint is necessary to build the other.
-The \textcolor{red}{dashed} arrows indicate a subsumption order, the former constraint is used to prune the latter.}
+The solid arrows indicate a specificity order, i.e., \textit{permutation} $\rightarrow$ \textit{series} means that the series constraint is more specific than permutation.
+The dashed arrows indicate an inverse construction order, \textit{project} $\dashrightarrow$ \textit{aggregates} means the project constraint is more specific than aggregates.
+\sergey{Samuel,  can you give an explanation why we keep both arrow and how it is internally motivated?}}
   \label{fig:learning_order}
@@
 \end{itemize}
-
-\section{Declarative modeling}
-\sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
 

### 2016-04-11 Paper/ECAI/tabular.tex
updatint the problem statement
writing the paper

@@
 
-A \textit{tabular solution} $S$ to a constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+A \textit{tabular solution} $S$ to an Excel constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
 
@@
 This matters from implementation perspective as well, if we discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all.
-A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is a more specific constraint than $C_2$ and therefore if $X$ is a solution for $C_1$ it should not be tested for $C_2$.
+A solid lines from $C_1$ to $C_2$ indicates that $C_2$ is \textit{more specific} than $C_1$ and if $X$ is not a solution to $C_1$, then it is not a solution to $C_2$.
+A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is \textit{a more specific} constraint than $C_2$ and therefore if $X$ is a solution for $C_1$ it should not be tested for $C_2$.
 
@@
     & a constraint specificity DAG \dependencies \\
-    \textbf{Find:}& all the most specific solution $S$ in \groups\\
-    & for each $c$ in \constraints \\
+    \textbf{Find:}& all the most specific solution $S$ in \groups for each $c$ in \constraints \\
   \end{tabular}
@@
 
+This version explicitly takes into account the specificity of the constraints and invalidates some of the solutions that are entailed by the tabular other solutions.
+
 \section{Approach to Tabular Constraint Learning}
+Here we describe how the problem of Condensed Tabular Constraint Learning can be solved.
+The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to find possible candidate groups,\sergey{continue here..}
+
 \begin{algorithm}[thb]
@@
 Also, if there is a construction edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
-If there is a sumsumption edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
+If there is a reversed specificity edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
 

### 2016-04-11 Code/src/core/constraint.py
improvements

@@ -63,13 +63,25 @@ def blank_filter(data, vectorized=False):
     return blank, (blank_f if not vectorized else numpy.vectorize(blank_f))
 
 
+def count_agg(data, axis=None):
+    if axis is None:
+        res = [len(data.flatten())]
+    elif axis == 0:
+        res = [data.shape[0]] * data.shape[1]
+    elif axis == 1:
+        res = [data.shape[1]] * data.shape[0]
+    else:
+        raise Exception("Invalid axis: {}".format(axis))
+    return numpy.array(res)
+
+
 class Operation(Enum):
     SUM = (numpy.sum, lambda x, y: x + y)
     MAX = (numpy.max, lambda x, y: max(x, y))
     MIN = (numpy.min, lambda x, y: min(x, y))
     AVERAGE = (numpy.average, lambda x, y: (x + y) / 2)
     PRODUCT = (numpy.product, lambda x, y: x * y)
-    COUNT = (lambda a: len(blank_filter(a.flatten(), True)[1](a.flatten())), lambda x, y: x + y)
+    COUNT = (count_agg, lambda x, y: x + y)
 
     # noinspection PyInitNewSignature
     def __init__(self, aggregate, func):
@@ -78,7 +90,7 @@ class Operation(Enum):
 
     @property
     def aggregate(self):
-        def apply(data, axis=None):
+        def apply(data, axis=None, partial=True):
             if axis == 1:
                 data = data.T
             if not data.shape:
@@ -86,16 +98,22 @@ class Operation(Enum):
             blank, blank_test = blank_filter(data, vectorized=True)
             if len(data.shape) > 1:
                 rows, cols = data.shape
-                results = numpy.array([blank] * cols)
-                for i in range(cols):
-                    vec = data[:, i][blank_test(data[:, i])]
-                    if len(vec) != 0:
-                        vec = numpy.array(vec, dtype=numpy.float64)
-                        results[i] = self._aggregate(vec)
-                return results
+                if partial:
+                    results = numpy.array([blank] * cols)
+                    for i in range(cols):
+                        vec = data[:, i][blank_test(data[:, i])]
+                        if len(vec) != 0:
+                            vec = numpy.array(vec, dtype=numpy.float64)
+                            results[i] = self._aggregate(vec)
+                    return results
+                else:
+                    return self._aggregate(data, 0)
             else:
-                array = numpy.array(data[blank_test(data)], dtype=numpy.float64)
-                return self._aggregate(array) if len(array) > 0 else blank
[…]

### 2016-04-11 Paper/ECAI/tabular.tex

--- 2016-04-11 Paper/ECAI/tabular.tex
@@
 
+\samuel{Explain concepts on example}
 \subsection{Groups, Type-consistency and Constraints}
@@
 For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
-
+\samuel{2nd problem statement (from approach), remove EQUAL}
 \section{Problem Statement}
@@
 
+\samuel{Running example (RANK?)}
 \paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
@@
     FK $\rightarrow$ PK & & \\
-    R = PRODUCT(FV, FK=PK | PV) & & \\
+    % TODO check out
+    R = LOOKUP-PRODUCT(FV, FK=PK, PV) & & \\
     FV = FUZZY-LOOKUP(FK, PK, PV) & & \\
@@
 
+\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}
+
   \end{tabular}
@@
 
+\samuel{Move to beginning without algo}
 \subsection{Workflow}
@@
     \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
-    \State $S \gets \postprocess(S)$
     \State \Return $S$
@@
 
+\samuel{Move next to subgroup satisfaction search}
 \section{Declarative modeling}
@@
 }
+
 \begin{itemize}

### 2016-04-11 Paper/ECAI/tabular.tex

--- 2016-04-11 Paper/ECAI/tabular.tex
@@
 For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
+
+\samuel{2nd problem statement (from approach), remove EQUAL}
 
@@
 
+\samuel{Running example (RANK?)}
 \paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
 Also, if there is a construction edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
-If there is a reversed specificity edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
+If there is a sumsumption edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
 
@@
 \end{itemize}
+
+\samuel{Move next to subgroup satisfaction search}
+\section{Declarative modeling}
+\sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
 

### 2016-04-11 Paper/ECAI/tabular.tex
Tias suggestion
removed algo, added an explation in text :

@@
 \textbf{Motivation}:
+\samuel{Suggest better table layouts / formula translation}
 \begin{itemize}
@@
 \section{Approach to Tabular Constraint Learning}
-Here we describe how the problem of Condensed Tabular Constraint Learning can be solved.
-The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to find possible candidate groups,\sergey{continue here..}
+As we have seen the statement before assumes the set of groups to be given.
+It is often, however, not the case and a possible approach is to extract the groups from the CSV file directly.
+In our system, we have implement the group construction step using the type-consistency check.
+\sergey{Samuel, we need details here, I think} In general, separating data from a header is a task on its own  \cite{header} and goes beyond the scope of this paper.
+
+Another solution to the group extraction problem is to make it interactive, when a user can select the groups in the spreadsheet or correct automatically generated groups.
+We regard this approach to be a possible future direction in the system development.
+Later in the paper we assume the groups to be given.
+
+Now we describe how the problem of Condensed Tabular Constraint Learning can be solved.
+The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to generate possible candidate groups, and in the next step we find the subgroups satisfying the constrain.
+After an iteration we accumulate all solutions and proceed to the next constraint.
 
@@
 \paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
-Also, if there is a construction edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
-If there is a sumsumption edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
+Also, if there is a solid edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
+If there is a dashed edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
 
@@
 \samuel{Move to beginning without algo}
-\subsection{Workflow}
-\begin{algorithm}[thb]
-  \begin{algorithmic}
-    \footnotesize
-    \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
-    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
-    \If{$T$ is \textbf{not} provided}
-      \State $T \gets \extracttables(D)$
-    \EndIf
-    \If{$G$ is \textbf{not} provided}
-      \State $G \gets \extractgroups(D, T)$
-    \EndIf
-    \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
-    \State \Return $S$
-\end{algorithmic}
-\caption{Workflow}
-\label{algo:workflow}
-\end{algorithm}
+% \subsection{Workflow}
+% \begin{algorithm}[thb]
+%   \begin{algorithmic}
+%     \footnotesize
+%     \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
+%     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
+%     \If{$T$ is \textbf{not} provided}
+%       \State $T \gets \extracttables(D)$
+%     \EndIf
+%     \If{$G$ is \textbf{not} provided}
+%       \State $G \gets \extractgroups(D, T)$
+%     \EndIf
+%     \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
+%     \State \Return $S$
+% \end{algorithmic}
+% \caption{Workflow}
+% \label{algo:workflow}
+% \end{algorithm}
 
-\textbf{Approach}
-\begin{itemize}
-  \item Notation
-  \item Algorithm (select constraints, find assignments, find solutions)
-\end{itemize}
+% \textbf{Approach}
+% \begin{itemize}
+%   \item Notation
+%   \item Algorithm (select constraints, find assignments, find solutions)
+% \end{itemize}
 
-\samuel{Move next to subgroup satisfaction search}
-\section{Declarative modeling}
-\sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
+% \samuel{Move next to subgroup satisfaction search}
+% \section{Declarative modeling}
+% \sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
 

### 2016-04-12 Code/src/accuracy.py
Added sumif diskettes

@@ -13,8 +13,7 @@ files = ["bmi", "age_department_sumif", "average_ablebits", "columnwise-sum-rows
          "financial_result", "fruits", "help_lookup", "household", "inventory", "multsum",
          "paper_supply", "price_weight", "repair", "rides", "sales_blanks", "school", "score",
          "shares", "sum_if_double_condition", "sum_table_from_guide", "sumif_blanks", "sumif_example",
-         "sumif_games_toys", "sumif_region", "sumif_uk", "sumproduct", "week_2_busn_store_2"]
-
+         "sumif_games_toys", "sumif_region", "sumif_uk", "sumproduct", "week_2_busn_store_2", "sumif_diskettes"]
 
 def is_excel_constraint(c: Constraint):
     return isinstance(c, Aggregate) or isinstance(c, ConditionalAggregate) or isinstance(c, Series) \

### 2016-04-12 Paper/ECAI/tabular.tex
removed algo, added an explation in text :
Extended evaluation

@@
     & a constraint specificity DAG \dependencies \\
-    \textbf{Find:}& all the most specific solution $S$ in \groups for each $c$ in \constraints \\
+    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\ & for each $c$ in \constraints \\
   \end{tabular}
@@
 For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}8]$ serves as $X$ and $Y$ and $x$ can be picked as $T_1[{:},7]$ and $y$ would be $T_1[{:},8]$.
+
+\samuel{Expand aggregates to max within group, tweak groups if necessary}
 
@@
 
-\section{Case Study aka Experiments}
+\section{Evaluation}
 In this section we experimentally validate our approach.
-The implementation is first illustrated using a case study on a specific Excel sheet.
-In order to quantify the results and generalize our findings we created a benchmark of 30 (\samuel{check number}) Excel sheets from various sources.
+We studied various questions, most notably with what accuracy our algorithm can find essential constraints.
 
-\subsection{Benchmark}
-There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online, publicly available spreadsheets such as crime statistics or financial reports\footnote{\samuel{link to links.txt}}.
+The implementation is illustrated using a case study on the spreadsheet corresponding with the previously introduced example (figure~\ref{fig:main_example}).
+In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 30 (\samuel{check number}) spreadsheets that we assembled from various sources.
 
-All spreadsheets have been converted to CSV with no manual intervention unless noted.
-Definitions of tables where added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).
-
-For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expressible in Excel and expected to be discovered.
-\footnote{Constraints currently outside of the scope of this system (e.g. generic nested mathematical formulas) are in a separate category (non-trivial).}
+In this section we focus on \textit{functional} constraints that could be used in Excel, ignoring constraints such as all-different or foreign-key.
 
 \subsection{Case study}
-Let us illustrate our implementation using the example presented in the introduction.
-We obtained this example by combining several Excel sheets from exercise sessions into one.
+Let us illustrate our implementation using the example presented in figure~\ref{fig:main_example}.
+This example combines several smaller examples that were used in an exercise session to teach Excel into one spreadsheet.
 Figure~\ref{fig:sol_example} shows the constraints that we expect to find.
-
-\sergey{I think there should be three things: key example evaluated, benchmark of 30 spreadsheets and maybe the FBI example, how we can detect interesting dependecies and correct potential mistakes}
-
-\paragraph{Accuracy}
-Our current implementation finds 18 constraints, containing all of the 12 solutions described in figure~\ref{fig:sol_example}.
-Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}
-  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}} \\
-  & \ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{4}}
-\end{align*}
-The other redundant constraint is a additional $\mathit{LOOKUP}$ that holds because the vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).
-
-We conclude that for this example our primary goal of finding all constraints is achieved.
-A number of redundant constraints ($\frac{1}{3}$ of the total) are also found, which, as we will show, is rather high compared to other spreadsheets.
-However, this example also contains many short vectors which increases the chance for constraints to be true by accident.
 
@@
     \end{align*}
-  }%
+  }
   \caption{Expected constraints case study}
 \end{figure}
+
+\subsubsection{Results}
+Our current implementation takes a few seconds\footnote{3 seconds on our Macbook Pro} to find 18 constraints, including all of the 12 solutions described in figure~\ref{fig:sol_example}.
+Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}
+  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}} \\
+  & \ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{4}}
+\end{align*}
+The other redundant constraint is an additional $\mathit{LOOKUP}$ that holds because the vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).
+
+For this example our primary goal of finding all constraints is achieved.
+The implementation also returns a number of redundant constraints ($33.33\%$ of the total).
+This ratio is, as we will show, rather high compared to other spreadsheets.
+However, this example contains many short vectors which increases the chance for constraints to be true by accident.
+
+\subsection{Benchmark}
+There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{\samuel{link to links.txt}}.
+The case study is also included.
+
+All spreadsheets have been converted to CSV with no manual intervention unless noted.
+Definitions of tables where added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).
+
+For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
+We consider three types of constraints:
+\begin{description}
+  \item[Expected] Constraint that have been implemented and are expected to be found
+  \item[Essential] Constraint that have or have not been implemented but could be found using our algorithm
+  \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)
+\end{description}
+
+\subsubsection*{Q1. How accurate is the approach?}
+Fortunately, our system is currently able to find $100\%$ of the expected constraints (90) on the benchmark suite.
+There are only three essential constraints that have not been implemented yet, therefore our system currently identifies $96.77\%$ of the constraints that we hope to discover.
+
+\subsubsection*{Q2. How many redundant constraints are discovered?}
+Our primary focus was accuracy, therefore we sometimes traded off less redundancy for more accuracy.
+The motivation is that solutions can still be pruned using entailment or heuristics afterwards.
+The constraints we considered to be redundant are either duplications (results that can be calculated in different ways, one of which seems superior) or constraints that hold by accident.
+Duplications should be detected in a post-processing step while accidental constraints should primarily be removed by adding data.
+
+Across all spreadsheets our implementation finds 121 constraints, 28 ($23.14\%$) of which are redundant.
+However, the average redundancy per spreadsheet is only $8.33\%$.
+Examining the redundant constraints reveals that many (12) of them are duplications occurring in a single spreadsheet.
+Many of the remaining constraints are duplications stemming from the overlapping role of difference and sum.
+Accidental constraints are limited to the 5 rank constraints that were discussed in the case study.
+
+\subsubsection*{Q3. How fast is the algorithm?}
+Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
+For the 32 spreadsheets in the benchmark our implementation ran in $17.86s \pm 0.87s$\footnote{Macbook Pro, Intel Core i7 2.3 GHz, 16GB RAM}.
+The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
+Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
+The search for aggregates will is slow on spreadsheets containing larger groups of numeric data.
+For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time.
+
+\subsection{Insights}
+- Running time -> dependency
+
+\sergey{I think there should be three things: key example evaluated, benchmark of 30 spreadsheets and maybe the FBI example, how we can detect interesting dependencies and correct potential mistakes}
 

### 2016-04-12 Paper/ECAI/tabular.tex
implemenation note

--- 2016-04-11 Paper/ECAI/tabular.tex
@@
     & a constraint specificity DAG \dependencies \\
-    \textbf{Find:}& all the most specific solution $S$ in \groups for each $c$ in \constraints \\
+    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\ 
+    & for each $c$ in \constraints \\
   \end{tabular}
@@
 For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}8]$ serves as $X$ and $Y$ and $x$ can be picked as $T_1[{:},7]$ and $y$ would be $T_1[{:},8]$.
+
+\paragraph{Implementation note} Essentially for each constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
+Hence, it is natural to solve them using declarative solvers.
+Even though each is a separate constraint problem, they are variations of each other and can be modeled using similar techniques.
+It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem.
+Furthermore, the declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
+
+Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native python CSP solver \cite{python_constraint}.
+All the constraints mentored in the paper can be run purely in the internal python engine without any external software.
+The future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages.
 

### 2016-04-12 Paper/ECAI/tabular.tex
Added paragraph on dependencies

--- 2016-04-12 Paper/ECAI/tabular.tex
@@
 Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
-The search for aggregates will is slow on spreadsheets containing larger groups of numeric data.
-For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time.
+The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
+For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).
+
+\paragraph{Dependencies}
+In order for the algorithm to run efficiently it is crucial to use dependencies and find constraint incrementally whenever possible.
+This avoids some of the explosion of combinations for constraints that have variables.
+For example, using foreign keys as a base constraint to find conditional aggregates reduces the running time for the case study from about 3 seconds to under 1 second.
+Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys present.
 
 \subsection{Insights}
-- Running time -> dependency
 

### 2016-04-12 Paper/ECAI/tabular.tex

--- 2016-04-12 Paper/ECAI/tabular.tex
@@
     & a constraint specificity DAG \dependencies \\
-    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\ & for each $c$ in \constraints \\
+    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\ 
+    & for each $c$ in \constraints \\
   \end{tabular}
@@
 \samuel{Expand aggregates to max within group, tweak groups if necessary}
+\sergey{[23:32:21] Samuel Kolb: And the motivation is that you can make the group smaller if you want to exclude a column from the aggregate that does not belong in it but accidentally gets included}
+
+\paragraph{Implementation note} Essentially for each constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
+Hence, it is natural to solve them using declarative solvers.
+Even though each is a separate constraint problem, they are variations of each other and can be modeled using similar techniques.
+It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem.
+Furthermore, the declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
+
+Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native python CSP solver \cite{python_constraint}.
+All the constraints mentored in the paper can be run purely in the internal python engine without any external software.
+The future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages.
 
@@
 This figure also specifies a potential parallelization of the computation, since each connected component is independent of the others.
-
-\sergey{need to elaborate here especially on the structure and how the constraints we picked}
 

### 2016-04-12 Paper/ECAI/tabular.tex
intro sketch

@@
 \section{Introduction}
-\sergey{bullet points for luc to start introduction}\\
-\textbf{Key question}:\\
+Imagine you have received a CSV file.
+Upon examining the content it becomes fairly clear it was made in Excel.
+Even the names of the fields such as Total or Avg indicated the usage of the standard functions.
+Having a closer look, you could deduce the functional dependencies in the table that look as in Figure \ref{fig:main_example}.
+Provided that a spreadsheet and the numbers in it are small, and dependencies are not complicated, one can examine the spreadsheet and find all connections.
+But wouldn't it be good to do it automatically?
+
+This brings us to the \textbf{key question}:\\
 Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\\
 
+Of course, it is stated semi-informally here, but it already allows us to provide an intuition on the task.
+We know that often people work in Excel with ranges, e.g.
+if we look at the Table 1  and at the first row in the Table 2 in Figure \ref{fig:main_example}, it is clear that one could have selected the first columns, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
+Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
+I.e., all operations are vectorized and make use of the standard Excel functions.
+
+It is not hard to see that computation goes not only in columns but in rows as well: if we sum the values in Table 1 in the first row in all four quarters, we would obtain the seventh column in Table 1.
+That gives us already a hint how this problem is different from the standard data mining settings, when the data is in rows and variables are in columns.
+Here everything is mixed, some values are missing in the results, some missing the arguments, even a header might have missing values.
+The data is relational on one hand, since we have multiple table with relationships between them.
+And on the other hand, the data is numeric, since people tend to use spreadsheets for financial and accounting computations, which is heavily numeric.
+
+\sergey{we need to incorporate the motivation part like error checking and autocompletion in the intro}
+
+\sergey{bullet points for Luc and Tias to rework introduction}\\
 \textbf{Motivation}:

### 2016-04-12 Paper/ECAI/tabular.tex

@@
 \newcommand{\ecrank}[2]{\ensuremath{#1 = \mathit{RANK}(#2)}}
+\newcommand{\ecfkey}[2]{\ensuremath{#1 \rightarrow #2}}
+\newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
 
@@
 \section{Approach to Tabular Constraint Learning}
-As we have seen the statement before assumes the set of groups to be given.
+As we have seen the statement before we assume the set of groups to be given.
 It is often, however, not the case and a possible approach is to extract the groups from the CSV file directly.
@@
 Now we describe how the problem of Condensed Tabular Constraint Learning can be solved.
-The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to generate possible candidate groups, and in the next step we find the subgroups satisfying the constrain.
+The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to generate possible candidate groups, and in the next step we find the subgroups satisfying the constraint.
 After an iteration we accumulate all solutions and proceed to the next constraint.
@@
 \samuel{Running example (RANK?)}
-\paragraph{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
-Also, if there is a solid edge from $C_1$ to $C_2$, and $X$ is not a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
-If there is a dashed edge from $C_1$ to $C_2$ and $X$ is a solution to $C_1$, then $X$ is not generated as a candidate group for $C_2$.
+\subsection{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
+Solid edges point from more general to more specific constraints.
+Therefore, if there is a solid edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ must contain the subgroups satisfying $C_1$.
 
-\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = RANK(X)} and $X,Y$ satisfy the constraints from \CSignature above.
-Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint.
-However, many constraints have the same candidate generation procedures, e.g., sum, min, max, avg, count, etc.
+Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
+If there is a dashed edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ may not contain a subset that also satisfies $C_1$.
 
-\paragraph{Subgroup satisfaction search} $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function looking for the subsets of vectors in the candidates satisfying the constraint.
-If multiple subsets satisfy the constraint, a maximal is selected.
-If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = RANK(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and a single vector $x$ in $G_1$ such that $x$ is ranked by $y$.
-For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}8]$ serves as $X$ and $Y$ and $x$ can be picked as $T_1[{:},7]$ and $y$ would be $T_1[{:},8]$.
+\paragraph{Example}
+Consider the \ecfkey{FK}{PK} (foreign-key) constraint: for every solution the subgroup (vector) $G_{PK}$ assigned to $PK$ must fulfill \ecalldiff{G_{PK}}.
+Therefore, we say indicate \textit{foreign-key} as being dependent on \textit{all-different} and the latter will be treated first.
+The \ecrank{Y}{X} constraint does not have any dependencies, however in a restricted version without ties the vector $Y$ would always be a permutation of numbers $1, 2, ..., length(Y)$.
 
-\samuel{Expand aggregates to max within group, tweak groups if necessary}
-\sergey{[23:32:21] Samuel Kolb: And the motivation is that you can make the group smaller if you want to exclude a column from the aggregate that does not belong in it but accidentally gets included}
+\subsection{Candidate group generation}
+$\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that may contains subgroups satisfying the constraint.
+The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
 
-\paragraph{Implementation note} Essentially for each constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
+Finding candidate groups for an excel constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
+Several constraints (e.g. sum, max, count), however, have the same candidate generation procedure as they share the same \CSignature.
+
+\paragraph{Example}
+The \ecrank{Y}{X} constraint requires $Y$ to be integer or textual and $X$ to be numeric.
+This allows to filter candidate groups for these variables.
+Additionally, $Y$ and~$X$ must be vectors of the same length, therefore, only groups with the same length are considered.
+Groups that does not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{Y}{X}.
+
+\subsection{Subgroup satisfaction search}
+$\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
+
+\paragraph{Example}
+Consider a candidate assignment $G_X, G_Y$ associated with the variables $X,Y$ of \ecrank{Y}{X}, then \findassignment selects single vectors $G_X^{i}, G_Y^{j}$ such that \ecrank{G_Y^{j}}{G_X^{i}} is true.
+For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_X$ and $G_Y$.
+\findassignment could then select $G_X^{i}$ to be $\range{T_1}{\rangeall}{7}$ and $G_Y^{j}$ to be $\range{T_1}{\rangeall}{8}$ which forms a valid solution.
+\\\\
+For some constraints (most notably aggregates) there may be multiple overlapping subsets that satisfy the constraint.
+We chose to omit subsets that are fully contained within a larger subset since this can lead to overfitting.
+If the returned subsets are too large and accidentally contains unrelated vectors, the user can split the group to reflect this structure.
+
+\subsection{Implementation note} Essentially for each constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
 Hence, it is natural to solve them using declarative solvers.

### 2016-04-12 Paper/ECAI/tabular.tex
Table WIP

@@
   \centering
-  \begin{tabular}{lcc}
+  \begin{tabular}{ccc}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
-    ALLDIFFERENT(X) & & \\
-    X = Y & & \\
-    FK $\rightarrow$ PK & & \\
+    ALLDIFFERENT(X) & X is integer or textual & All elements in X are different \\
+    %X = Y & & \\
+    FK $\rightarrow$ PK & FK and PK have the same type; they originate from different tables and PK is all-different & all values in FK exist in PK \\
     % TODO check out
@@
      Y = RANK(X) & & \\
-    A = PREV(A) + P - N & & \\
+    A = PREV(A) + P - N & A, P and N have the same length and their length is at least 2 & *\\
     SERIES(X) & & \\
-    Y = SUM(X, col) & & \\
-    Y = SUM(X, row) & & \\
+    Y = SUM(X, col) & & Y is the result of summing each column in X \\
+    Y = SUM(X, row) & & Y is the result of summing each row in X \\
     R = SUMIF(FK=PK, V) & & \\

### 2016-04-12 Paper/ECAI/tabular.tex
Fixed table
rank

@@
 \usepackage{amsmath}
+\usepackage{tabularx}
 
@@
   \centering
-  \begin{tabular}{ccc}
+  \begin{tabularx}{\textwidth}{XXX}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
@@
     R = PROJECT(P) & & \\
-     Y = RANK(X) & & \\
+    Y = RANK(X) & Y and X have the same length & Y is the rank of X (including ties)\\
     A = PREV(A) + P - N & A, P and N have the same length and their length is at least 2 & *\\
@@
 
-  \end{tabular}
+  \end{tabularx}
   \caption{Excel Tabular Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}

### 2016-04-13 Paper/ECAI/tabular.tex
Put Tias comments in
editing the paper

@@
 \newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
+\newcommand{\tias}[1]{\textcolor{blue}{{\sc Tias:} #1}\xspace}
 
@@
 \end{abstract}
+
+\tias{Use spreadsheet instead of Excel}
+\tias{Spreadsheet is an instance of tabular data, but not the only one?}
+
+\tias{Rewrite}
 \section{Introduction}
-Imagine you have received a CSV file.
-Upon examining the content it becomes fairly clear it was made in Excel.
-Even the names of the fields such as Total or Avg indicated the usage of the standard functions.
-Having a closer look, you could deduce the functional dependencies in the table that look as in Figure \ref{fig:main_example}.
-Provided that a spreadsheet and the numbers in it are small, and dependencies are not complicated, one can examine the spreadsheet and find all connections.
-But wouldn't it be good to do it automatically?
+Millions of people across use spreadsheets everyday.
+Yet, many people lack a complete understanding of the structures and existing dependencies in the data.
+Indeed, the most useful part of spreadsheets is often the formulas and often these formula disappear, when a file is exported and shared as a CSV file.
+This brings us to the question: can we discover or reconstruct structural relations in flat tabular spreadsheet data?
+[in a general way that allows declarative specification of constraints to discover]
 
-This brings us to the \textbf{key question}:\\
-Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\\
-
-Of course, it is stated semi-informally here, but it already allows us to provide an intuition on the task.
-We know that often people work in Excel with ranges, e.g.
-if we look at the Table 1  and at the first row in the Table 2 in Figure \ref{fig:main_example}, it is clear that one could have selected the first columns, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
+Let us look at an example in Figure \ref{fig:main_example}, the name already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
+If we look at the Table 1 and at the first row in the Table 2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
 Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
-I.e., all operations are vectorized and make use of the standard Excel functions.
+I.e., all operations are vectorized and make use of the standard spreadsheet operations.
 
@@
 
-\sergey{we need to incorporate the motivation part like error checking and autocompletion in the intro}
+This makes tabular data understanding and constraint learning new and challenging, and it is not only challenging for a machine but for a human as well.
+Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry.
+Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
 
@@
 \end{itemize}
+
+\samuel{Spreadsheet risks: financial losses due to errors in spreadsheets}
 
@@
 \section{Formalization}
-
+\tias{More gentle}
 \samuel{Explain concepts on example}
@@
 
+\tias{Use generic constraint}
 An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
@@
   \end{tabularx}
-  \caption{Excel Tabular Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}
+  \caption{Excel Tabular Constraints}
   \label{table:constraints}
@@
 
-\samuel{Move to beginning without algo}
 % \subsection{Workflow}

### 2016-04-13 Paper/ECAI/tabular.tex
Fixes

--- 2016-04-12 Paper/ECAI/tabular.tex
@@
 
-\newcommand{\ecrank}[2]{\ensuremath{#1 = \mathit{RANK}(#2)}}
+\newcommand{\eccalc}[2]{\ensuremath{#1 = #2}}
+\newcommand{\ecrank}[2]{\eccalc{#1}{\mathit{RANK}(#2)}}
 \newcommand{\ecfkey}[2]{\ensuremath{#1 \rightarrow #2}}
 \newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
+\newcommand{\eclookupf}[4]{\ensuremath{\mathit{LOOKUP}_{#4}(#1, #2, #3)}}
+\newcommand{\eclookup}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{}}}
+\newcommand{\eclookupprod}[5]{\eccalc{#1}{#2 \times \eclookupf{#3}{#4}{#5}{}}}
+\newcommand{\eclookupfuzzy}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{fuzzy}}}
+\newcommand{\ecperm}[1]{\ensuremath{\mathit{PERMUTATION}(#1)}}
+\newcommand{\ecseries}[1]{\ensuremath{\mathit{SERIES}(#1)}}
+\newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
+\newcommand{\ectotal}[3]{\eccalc{#1}{\mathit{PREV}(#1) + #2 - #3}}
+\newcommand{\ecproj}[2]{\eccalc{#1}{\mathit{PROJECT}(#2)}}
+\newcommand{\ecaggc}[3]{\eccalc{#2}{\mathit{#1}(#3, col)}}
+\newcommand{\ecaggr}[3]{\eccalc{#2}{\mathit{#1}(#3, row)}}
+\newcommand{\ecsumc}[2]{\eccalc{#1}{\mathit{SUM}(#2, col)}}
+\newcommand{\ecsumr}[2]{\eccalc{#1}{\mathit{SUM}(#2, row)}}
+\newcommand{\ecaggif}[5]{\eccalc{#2}{\mathit{#1IF}(#3, #4, #5)}}
+\newcommand{\ecsumif}[4]{\eccalc{#1}{\mathit{SUMIF}(#2, #3, #4)}}
+\newcommand{\ecsumprod}[3]{\eccalc{#1}{\mathit{SUMPRODUCT}(#2, #3)}}
 
@@
 \end{abstract}
+
+\tias{Use spreadsheet instead of Excel}
+\tias{Spreadsheet is an instance of tabular data, but not the only one?}
+
+\tias{Rewrite}
 \section{Introduction}
@@
 Of course, it is stated semi-informally here, but it already allows us to provide an intuition on the task.
-We know that often people work in Excel with ranges, e.g.
+We (know) that often people work in spreadsheets with ranges, e.g.
 if we look at the Table 1  and at the first row in the Table 2 in Figure \ref{fig:main_example}, it is clear that one could have selected the first columns, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
 Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
-I.e., all operations are vectorized and make use of the standard Excel functions.
+I.e., all operations are vectorized and make use of the standard spreadsheet functions.
 
@@
 
-An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
+\tias{Use generic constraint}
+A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
 Let us elaborate on each of them.
@@
 
-A \textit{tabular solution} $S$ to an Excel constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+A \textit{tabular solution} $S$ to an spreadsheet constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
 
@@
 In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
-Here we formalize the statement in terms of Excel constraints and group assignments as follows:
+Here we formalize the statement in terms of spreadsheet constraints and group assignments as follows:
 
@@
     &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$\\
+    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$\\
     \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ \constraints \\ 
@@
   The key observation here is that essentially the problem is a constraint enumeration problem, where each constraint is independent.
-This property comes from fact that in Excel each formula is applied based on the existing data in the cells.
+This property comes from fact that in spreadsheets each formula is applied based on the existing data in the cells.
 This allows learning constraints independently of each other by examining constraint satisfaction on the groups.
@@
     &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of Excel constraints $\constraints$,\\ 
+    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$,\\ 
     & a constraint specificity DAG \dependencies \\
@@
     \footnotesize
-    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Excel constraints, \dependencies -- dependency graph 
+    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Spreadsheet constraints, \dependencies -- dependency graph 
     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
@@
 \samuel{Running example (RANK?)}
+
 \subsection{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
@@
 If there is a dashed edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ may not contain a subset that also satisfies $C_1$.
+
+\dependencies thus describes a partial order in which constraints should be learned.
+Some constraints have no dependencies (ommitted from the figure) and can be learned independently.
+Generally, a parallel version of the algorithm could learn each connected component independently, only taking into account the partial ordering within components.
 
@@
 \subsection{Constraints}
-As emphasized before, the set of Excel constraints has a partial order in which they should be learned.
+% As emphasized before, the set of Excel constraints has a partial order in which they should be learned.
 For certain constraints this order does not matter, such as \textit{rank} or \textit{product}.
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
-    ALLDIFFERENT(X) & X is integer or textual & All elements in X are different \\
+    \ecalldiff{V_x} & X is integer or textual & All elements in X are different \\
     %X = Y & & \\
-    FK $\rightarrow$ PK & FK and PK have the same type; they originate from different tables and PK is all-different & all values in FK exist in PK \\
+    \ecfkey{V_{fk}}{V_{pk}} & FK and PK have the same type; they originate from different tables and PK is all-different & all values in FK exist in PK \\
     % TODO check out
-    R = LOOKUP-PRODUCT(FV, FK=PK, PV) & & \\
-    FV = FUZZY-LOOKUP(FK, PK, PV) & & \\
-    FV = LOOKUP(FK, PK, PV) & & \\
-    PERMUTATION(X) & & \\
-    R = O1 * O2 & & \\
-    R = PROJECT(P) & & \\
-    Y = RANK(X) & Y and X have the same length & Y is the rank of X (including ties)\\
-    A = PREV(A) + P - N & A, P and N have the same length and their length is at least 2 & *\\
-    SERIES(X) & & \\
-    Y = SUM(X, col) & & Y is the result of summing each column in X \\
-    Y = SUM(X, row) & & Y is the result of summing each row in X \\
-    R = SUMIF(FK=PK, V) & & \\
-    R = SUMPRODUCT(O1, O2) & & \\
+    \eclookupprod{V_r}{V_1}{V_{fk}}{V_{pk}}{V_{val}} & & \\
+    \eclookup{V_r}{V_{fk}}{V_{pk}}{V_{val}} & & \\
+    \eclookupfuzzy{V_r}{V_{fk}}{V_{pk}}{V_{val}} & & \\
+    \ecperm{V_x} & & \\
+    \ecprod{V_r}{V_1}{V_2} & & \\
+    \ecproj{V_r}{G_x} & & \\
+    \ecrank{V_r}{V_1} & Y and X have the same length & Y is the rank of X (including ties)\\
+    \ectotal{V_r}{V_{pos}}{V_{neg}} & A, P and N have the same length and their length is at least 2 & *\\
+    \ecseries{V_1} & & \\
+    \ecsumc{V_r}{G_1} & & Y is the result of summing each column in X \\
+    \ecsumr{V_r}{G_1} & & Y is the result of summing each row in X \\
+    \ecsumif{V_r}{V_{fk}}{V_{pk}}{V_{val}} & & \\
+    \ecsumprod{V_r}{V_1}{V_2} & & \\
 
@@
   \end{tabularx}
-  \caption{Excel Tabular Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}
+  \caption{Spreadsheet Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}
   \label{table:constraints}
@@
 
-In this section we focus on \textit{functional} constraints that could be used in Excel, ignoring constraints such as all-different or foreign-key.
+In this section we focus on \textit{functional} constraints that could be used in spreadsheets, ignoring constraints such as all-different or foreign-key.
 
@@
       & SERIES(\range{T_{1}}{\rangeall}{1}) \\
-      & \range{T_{1}}{\rangeall}{8} = RANK(\range{T_{1}}{\rangeall}{7}) \\
-      & \range{T_{2}}{1}{\rangeall} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
-      & \range{T_{6}}{\rangeall}{2} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{6}}, col) \\
-      & \range{T_{1}}{\rangeall}{7} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{6}}, row) \\
-      & \range{T_{2}}{2}{\rangeall} = AVERAGE(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
-      & \range{T_{2}}{3}{\rangeall} = MAX(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
-      & \range{T_{2}}{4}{\rangeall} = MIN(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
-      & \range{T_{1}}{\rangeall}{10} = SUMIF(\range{T_{5}}{\rangeall}{1}=\range{T_{1}}{\rangeall}{2}, \range{T_{5}}{\rangeall}{2}) \\
-      & \range{T_{1}}{\rangeall}{11} = MAXIF(\range{T_{5}}{\rangeall}{1}=\range{T_{1}}{\rangeall}{2}, \range{T_{5}}{\rangeall}{2}) \\
-      & \range{T_{4}}{\rangeall}{3} = LOOKUP(\range{T_{4}}{\rangeall}{2}, \range{T_{1}}{\rangeall}{1}, \range{T_{1}}{\rangeall}{2}) \\
+%
+      & \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
+%
+      & \ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      & \ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+      & \ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+      & \ecaggc{AVERAGE}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      & \ecaggc{MAX}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      & \ecaggc{MIN}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      & \ecaggif{SUM}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+%
+      & \ecaggif{MAX}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+%
+      & \eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
+%
       & \range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}

### 2016-04-13 Paper/ECAI/tabular.tex
added the figure with insight

--- 2016-04-13 Paper/ECAI/tabular.tex
@@
 \subsection{Insights}
+We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.
+
+\begin{figure*}[thb]
+  \begin{center}
+    \includegraphics[width=0.85\textwidth]{figures/fbi_figure_highlighted.png}
+  \end{center}
+  \caption{Real world tabular constraint reconstruction: FBI crime statistics}
+  \label{fig:fbi}
+\end{figure*}
 

### 2016-04-13 Paper/ECAI/tabular.tex

--- 2016-04-13 Paper/ECAI/tabular.tex
@@
 \section{Introduction}
-Imagine you have received a CSV file.
-Upon examining the content it becomes fairly clear it was made in Excel.
-Even the names of the fields such as Total or Avg indicated the usage of the standard functions.
-Having a closer look, you could deduce the functional dependencies in the table that look as in Figure \ref{fig:main_example}.
-Provided that a spreadsheet and the numbers in it are small, and dependencies are not complicated, one can examine the spreadsheet and find all connections.
-But wouldn't it be good to do it automatically?
+Millions of people across use spreadsheets everyday.
+Yet, many people lack a complete understanding of the structures and existing dependencies in the data.
+Indeed, the most useful part of spreadsheets is often the formulas and often these formula disappear, when a file is exported and shared as a CSV file.
+This brings us to the question: can we discover or reconstruct structural relations in flat tabular spreadsheet data?
+[in a general way that allows declarative specification of constraints to discover]
 
-This brings us to the \textbf{key question}:\\
-Can we discover or reconstruct structural relations in flat tabular spreadsheet data? [in a general way that allows declarative specification of constraints to discover]\\
-
-Of course, it is stated semi-informally here, but it already allows us to provide an intuition on the task.
-We (know) that often people work in spreadsheets with ranges, e.g.
-if we look at the Table 1  and at the first row in the Table 2 in Figure \ref{fig:main_example}, it is clear that one could have selected the first columns, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
+Let us look at an example in Figure \ref{fig:main_example}, the name already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
+If we look at the Table 1 and at the first row in the Table 2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
 Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
-I.e., all operations are vectorized and make use of the standard spreadsheet functions.
+I.e., all operations are vectorized and make use of the standard spreadsheet operations.
 

### 2016-04-13 Paper/ECAI/tabular.tex
fixed typo

@@
 
-A \textit{tabular solution} $S$ to an spreadsheet constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+A \textit{tabular solution} $S$ to a spreadsheet constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
 
@@
     \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$\\
-    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ \constraints \\ 
+    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ in \constraints \\ 
   \end{tabular}

### 2016-04-13 Paper/ECAI/tabular.tex
put my comments

@@
 \samuel{Explain concepts on example}
+\sergey{subgroup notation, rework the solution definition}
 \subsection{Groups, Type-consistency and Constraints}
@@
 \section{Problem Statement}
+\sergey{should reflect all the changes, Luc wants to introduce constraint satisfation into problem statement, maybe we should move the definition of a solution here?
+Like a solution is actually constraint satisfaction?}
 In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
@@
 \section{Approach to Tabular Constraint Learning}
+\sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
 As we have seen the statement before we assume the set of groups to be given.
@@
 
+\sergey{Table should take into account our new notation and balance between formality and intuition, heh}
 \begin{table*}

### 2016-04-14 Paper/ECAI/tabular.tex
updating the formalization

@@
 \section{Formalization}
-\tias{More gentle}
 \samuel{Explain concepts on example}
@@
 \subsection{Groups, Type-consistency and Constraints}
-In this work we distinguish the following types of data: numeric and textual.
+In this work we focus on the spreadsheet data, which is physically one big table, but conceptually consists of multiple tables, see Fig \ref{fig:main_example} for an example.
+Formally, a table is $n \times m$ matrix, each entry is called a \textit{cell}.
+A cell has a type, e.g.
+namely numeric.
+In this work we distinguish the two types of data: numeric and textual.
 The numeric type has two subtypes: integers and floats.
 We also consider the special element called \textit{None}, which has two types: numeric and textual.
+
+Based on these types we introduce a notation of type consistency for a set of elements.
 A set is called \textit{type-consistent} iff all elements are either numeric or textual.
 Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.
+
+We use the notation $T[a,b]$ to refer to a particular cell in the table $T$ at the row with index $a$ and the column with index $b$.
+To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
 

### 2016-04-14 Paper/ECAI/tabular.tex
updating the formalization

@@
 
+We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If the subgroup $g$ must constain only one vector, we write $g \in G$.
+
 \tias{Use generic constraint}
@@
 \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function.
-For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
+For example, a spreadsheet constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
 
@@
 
-Finding candidate groups for an excel constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
+Finding candidate groups for a spreadsheet constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
 Several constraints (e.g. sum, max, count), however, have the same candidate generation procedure as they share the same \CSignature.

### 2016-04-14 Paper/ECAI/tabular.tex
updating the formalization

@@
 \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function.
-For example, a spreadsheet constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
+For example, a spreadsheet constraint \textit{rank} has \CName \textit{$g_Y$ = RANK($g_X$)}, where $g_X$ and $g_Y$ are the variables; its \CSignature is: the subgroup $g_X$ (associated with X) is numeric, $g_Y$ is integer and the length of vectors in $g_X$ is the same as in $g_Y$; its \CFunction is the following constraint: a pair of vectors $g_X,g_Y$ is a solution iff $g_X \in G_X, g_Y \in G_Y$ and each value in $g_X$ has the rank (possibly with ties) specified in $g_Y$.
 
@@
 
-A \textit{tabular solution} $S$ to a spreadsheet constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+A \textit{tabular solution} $S$ to a spreadsheet constraint $C$ (with $n$ variables) is a subgroup assignment $(g_1, \dots, g_n)$ where $g_i$, such that $(g_1,\dots,g_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+\sergey{to finish this definition later, need to say that $g_i$ is the subgroup assigned to the $i$-th variable in $C$}
 

### 2016-04-14 Paper/ECAI/tabular.tex
rewriting the paper

--- 2016-04-14 Paper/ECAI/tabular.tex
@@
 
-\tias{Use spreadsheet instead of Excel}
-\tias{Spreadsheet is an instance of tabular data, but not the only one?}
-
 \tias{Rewrite}
@@
 This makes tabular data understanding and constraint learning new and challenging, and it is not only challenging for a machine but for a human as well.
-Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry.
+Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the \textit{Spreadsheet risk}\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaws in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
 Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
@@
 \subsection{Groups, Type-consistency and Constraints}
-In this work we focus on the spreadsheet data, which is physically one big table, but conceptually consists of multiple tables, see Fig \ref{fig:main_example} for an example.
+In this work we focus on the spreadsheet data, which is physically one big table, but conceptually consists of multiple tables, see Figure \ref{fig:main_example} for an example.
 Formally, a table is $n \times m$ matrix, each entry is called a \textit{cell}.
@@
 
-We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If the subgroup $g$ must constain only one vector, we write $g \in G$.
+We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must constain only one vector, we write $g \in G$.
 
-\tias{Use generic constraint}
 A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
 Let us elaborate on each of them.
-\textit{\CName} is the textual name of the constraint together with its variable names.
-\textit{\CSignature} is a set of constraints specifying the properties of the group assignments, corresponding to the variables in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g.
+\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.
+\textit{\CSignature} is a constraint specifying the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g.
 the length of vectors in the arguments must be equal.
 These are constraints on the group meta-information not on the actual group content.
-\textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function.
-For example, a spreadsheet constraint \textit{rank} has \CName \textit{Y = RANK(X)}, where $X$ and $Y$ are the variables; its \CSignature is: the group $G_X$ (associated with X) is numeric, $G_Y$ is integer and the length of vectors in $G_X$ is the same as in $G_Y$; its \CFunction is the following constraint: a pair of vectors $X,Y$ is a solution iff $X \in G_X, Y \in G_Y$ and each value in $X$ has the rank (possibly with ties) specified in $Y$.
+\textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
+For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_y$ = RANK($v_x$)}, where $v_x$ and $v_y$ are the variables; its \CSignature is: the subgroup assigned to $v_x$ must be numeric, $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_x$ has the rank (possibly with ties) specified in $g_y$.
 
-\samuel{2nd problem statement (from approach), remove EQUAL}
-
-A \textit{tabular solution} $S$ to a spreadsheet constraint $C$ is a tuple $(X_1, \dots, X_n)$ where $n$ is the number of variables in $C$ and each $X_i$ is a subgroup in \groups, such that $(X_1,\dots,X_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+Let us now define what is a tabular solution for a spreadsheet constraint in general.
+Let $C$ denote a spreadsheet constraint, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
 
@@
 In practice, however, things are different.
-Assume that for some $X$ both \textit{series(X)} and \textit{alldifferent(X)} hold.
-From a user's perspective the last solution is useless, since he or she already knows that series constraint implies alldifferent constraint.
+Assume that for some $g_x$ both \textit{series($g_x$)} and \textit{alldifferent($g_x$)} hold.
+From a user's perspective the last solution is useless, since he or she already knows that the series constraint implies the alldifferent constraint.
 To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity (generality) order between constraints, which is depicted using solid (dashed) arrows in Figure \ref{fig:learning_order}.
-This matters from implementation perspective as well, if we discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all.
-A solid lines from $C_1$ to $C_2$ indicates that $C_2$ is \textit{more specific} than $C_1$ and if $X$ is not a solution to $C_1$, then it is not a solution to $C_2$.
-A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is \textit{a more specific} constraint than $C_2$ and therefore if $X$ is a solution for $C_1$ it should not be tested for $C_2$.
+This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all.
+A solid lines from $C_1$ to $C_2$ indicates that $C_2$ is \textit{more specific} than $C_1$ and if $S$ is not a solution to $C_1$, then it is not a solution to $C_2$.
+A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is \textit{a more specific} constraint than $C_2$ and therefore if $S$ is a solution for $C_1$ it should not be tested for $C_2$.
 
@@
 \paragraph{Example}
-Consider the \ecfkey{FK}{PK} (foreign-key) constraint: for every solution the subgroup (vector) $G_{PK}$ assigned to $PK$ must fulfill \ecalldiff{G_{PK}}.
+Consider the \ecfkey{FK}{PK} (foreign-key) constraint: for every solution the subgroup (vector) $v_{PK}$ assigned to \textit{PK} must fulfill \ecalldiff{v_{PK}}.
 Therefore, we say indicate \textit{foreign-key} as being dependent on \textit{all-different} and the latter will be treated first.
-The \ecrank{Y}{X} constraint does not have any dependencies, however in a restricted version without ties the vector $Y$ would always be a permutation of numbers $1, 2, ..., length(Y)$.
+The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties the vector $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
 
@@
 \paragraph{Example}
-The \ecrank{Y}{X} constraint requires $Y$ to be integer or textual and $X$ to be numeric.
+The \ecrank{v_y}{v_x} constraint requires $v_y$ to be integer or textual and $v_x$ to be numeric.
 This allows to filter candidate groups for these variables.
-Additionally, $Y$ and~$X$ must be vectors of the same length, therefore, only groups with the same length are considered.
-Groups that does not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{Y}{X}.
+Additionally, $v_y$ and~$v_x$ must be vectors of the same length, therefore, only groups with the same length are considered.
+Groups that does not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{v_y}{v_x}.
 
@@
 \paragraph{Example}
-Consider a candidate assignment $G_X, G_Y$ associated with the variables $X,Y$ of \ecrank{Y}{X}, then \findassignment selects single vectors $G_X^{i}, G_Y^{j}$ such that \ecrank{G_Y^{j}}{G_X^{i}} is true.
-For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_X$ and $G_Y$.
-\findassignment could then select $G_X^{i}$ to be $\range{T_1}{\rangeall}{7}$ and $G_Y^{j}$ to be $\range{T_1}{\rangeall}{8}$ which forms a valid solution.
+Consider a candidate assignment $G_x, G_y$ associated with the variables $v_x,v_y$ of \ecrank{v_y}{v_x}, then \findassignment selects single vectors $g_y \in G_y, g_x \in G_x$ such that \ecrank{g_y}{g_x} is true, i.e.
+it searches for a mapping $S$ where $v_x$ ($v_y$) must be mapped to a subgroup of $G_x$ ($G_y$).
+For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_x$ and $G_y$.
+Then \findassignment could select $g_x$ to be $\range{T_1}{\rangeall}{7}$ and $g_y$ to be $\range{T_1}{\rangeall}{8}$ which forms a valid solution.
 \\\\
@@
 
-Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native python CSP solver \cite{python_constraint}.
-All the constraints mentored in the paper can be run purely in the internal python engine without any external software.
-The future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages.
+Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native Python CSP solver \cite{python_constraint}.
+All the constraints mentored in the paper can be run purely in the internal Python engine without any external software.
+The future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages to allow users to add new constraints in a declarative way.
 
 \subsection{Constraints}
-% As emphasized before, the set of Excel constraints has a partial order in which they should be learned.
-For certain constraints this order does not matter, such as \textit{rank} or \textit{product}.
-For many others they have to be learned in the order, which is specified as a DAG in Figure \ref{fig:learning_order} (if a constraint is not in the graph, it is independent).
-This figure also specifies a potential parallelization of the computation, since each connected component is independent of the others.
 
@@
   \end{tabularx}
-  \caption{Spreadsheet Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}
+  \caption{Spreadsheet Constraints \sergey{Here a detailed explanation on what is going on, what is essential, what is not? etc should be self-explanatory}}
   \label{table:constraints}
@@
 \subsection{Case study}
-Let us illustrate our implementation using the example presented in figure~\ref{fig:main_example}.
+Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
 This example combines several smaller examples that were used in an exercise session to teach Excel into one spreadsheet.
@@
   }
-  \caption{Expected constraints case study}
+  \caption{Expected constraints in the case study \sergey{in Figure \ref{fig:main_example}; some more explanation needed, not self explanatory}}
 \end{figure}
@@
 \subsubsection*{Q1. How accurate is the approach?}
+\sergey{that's confusing, it can do 100\% but then all of a sudden 96\%, short explanation needed}
 Fortunately, our system is currently able to find $100\%$ of the expected constraints (90) on the benchmark suite.
@@
 \subsection{Insights}
-We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.
+\sergey{We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.}
 

### 2016-04-14 Paper/ECAI/tabular.tex
rewriting the paper

@@
 
-Let us look at an example in Figure \ref{fig:main_example}, the name already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
+Let us have a look at the example in Figure \ref{fig:main_example}, certainly there are several constraint present in the data and even the names in the headers already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
 If we look at the Table 1 and at the first row in the Table 2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.

### 2016-04-14 Paper/ECAI/tabular.tex
fixed minor, added comments

--- 2016-04-14 Paper/ECAI/tabular.tex
@@
 
-\tias{Rewrite}
 \section{Introduction}
@@
 
-Let us look at an example in Figure \ref{fig:main_example}, the name already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
-If we look at the Table 1 and at the first row in the Table 2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
+Let us have a look at the example in Figure \ref{fig:main_example}.
+Certainly there are several constraint present in the data and even the names in the headers already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
+If we look at the Table 1 and at the first row in the Table 2, it is clear that one could have selected the column called \textbf{1st Quarter} in the Table  in the Table 11, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
 Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
@@
 
-It is not hard to see that computation goes not only in columns but in rows as well: if we sum the values in Table 1 in the first row in all four quarters, we would obtain the seventh column in Table 1.
-That gives us already a hint how this problem is different from the standard data mining settings, when the data is in rows and variables are in columns.
-Here everything is mixed, some values are missing in the results, some missing the arguments, even a header might have missing values.
-The data is relational on one hand, since we have multiple table with relationships between them.
+It is not hard to see that computation goes not only in columns but in rows as well: if we sum the values in Table 1 in the first row in all four quarters, we obtain the seventh column in Table 1.
+That gives us already a hint how this problem is different from the standard data mining setting, when the data is in rows and variables are in columns.
+Here everything is mixed, also some values are missing in the results, some missing the arguments, even a header might have missing values.
+The data is relational on the one hand, since we have multiple tables with relationships between them.
 And on the other hand, the data is numeric, since people tend to use spreadsheets for financial and accounting computations, which is heavily numeric.
@@
 Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the \textit{Spreadsheet risk}\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaws in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
-Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
+Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggesting possible fixes.
 
@@
 
-We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must constain only one vector, we write $g \in G$.
+We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
 
@@
 \section{Problem Statement}
-\sergey{should reflect all the changes, Luc wants to introduce constraint satisfation into problem statement, maybe we should move the definition of a solution here?
+\sergey{should reflect all the changes, Luc wants to introduce constraint satisfaction into problem statement, maybe we should move the definition of a solution here?
 Like a solution is actually constraint satisfaction?}
@@
 
-\dependencies thus describes a partial order in which constraints should be learned.
-Some constraints have no dependencies (ommitted from the figure) and can be learned independently.
+Thus, \dependencies describes a partial order in which constraints should be learned.
+Some constraints have no dependencies (omitted from the figure) and can be learned independently.
 Generally, a parallel version of the algorithm could learn each connected component independently, only taking into account the partial ordering within components.
@@
 \begin{figure}
-  \label{fig:sol_example}
   {\small
@@
   \caption{Expected constraints in the case study \sergey{in Figure \ref{fig:main_example}; some more explanation needed, not self explanatory}}
+  \label{fig:sol_example}
 \end{figure}
+\sergey{Samuel, hint, don't put label before caption, it can't resolve it sometimes}
 
 \subsubsection{Results}
-Our current implementation takes a few seconds\footnote{3 seconds on our Macbook Pro} to find 18 constraints, including all of the 12 solutions described in figure~\ref{fig:sol_example}.
+\sergey{Samuel, you should just say at the beginning that all experiments has been run on macbook la-la system and that's it, I guess, I don't think we need to mention it multiple times}
+Our current implementation takes a few seconds\footnote{3 seconds on our Macbook Pro} to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
 Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}

### 2016-04-14 Paper/ECAI/luc1.tex
Added luc's changes

--- 2016-04-14 Paper/ECAI/tabular.tex
@@
 
+\newcommand{\luc}[1]{{\textcolor{red}{#1}}}
+
 %%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
@@
 \end{abstract}
+
+\tias{Use spreadsheet instead of Excel}
+\tias{Spreadsheet is an instance of tabular data, but not the only one?}
 
@@
 This makes tabular data understanding and constraint learning new and challenging, and it is not only challenging for a machine but for a human as well.
-Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the \textit{Spreadsheet risk}\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaws in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
+Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry.
 Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
@@
 \subsection{Groups, Type-consistency and Constraints}
-In this work we focus on the spreadsheet data, which is physically one big table, but conceptually consists of multiple tables, see Figure \ref{fig:main_example} for an example.
-Formally, a table is $n \times m$ matrix, each entry is called a \textit{cell}.
-A cell has a type, e.g.
-namely numeric.
-In this work we distinguish the two types of data: numeric and textual.
-The numeric type has two subtypes: integers and floats.
+In this work we focus on the spreadsheet data, which is physically one big table, but conceptually consists of multiple tables, see Fig \ref{fig:main_example} for an example.
+Formally, a table is $n \times m$ matrix, each entry table is called  a \textit{cell}.
+
+Each cell has a {\em type} which can be numeric or textual. The numeric type has two subtypes: integers and floats. 
 We also consider the special element called \textit{None}, which has two types: numeric and textual.
+A set of cells is called \textit{type-consistent} iff all cells in the set are either numeric or textual. 
+%Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric subtype by requiring its arguments to be integers.
 
-Based on these types we introduce a notation of type consistency for a set of elements.
-A set is called \textit{type-consistent} iff all elements are either numeric or textual.
-Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.
-
+We consider various sets of cells in tables: 
+\item {\bf Cells}: Each entry in the matrix/table is called  a \textit{cell}.
 We use the notation $T[a,b]$ to refer to a particular cell in the table $T$ at the row with index $a$ and the column with index $b$. 
-To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
-
+\item {\bf Columns and rows}: To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
+\item 
 A \textit{vector} is either a column or a row that is type-consistent. 
 If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation. 
+\item
 A \textit{group} is a subrange of vectors with the same orientation in a table.
@@
 We denote as the \textit{length} of a row (column) group $G$, written as \textit{length(G)}, the number of its columns (rows).
-We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if its vectors contain numeric (textual, etc) elements.
+We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if \luc{all?????} its vectors contain numeric (textual, etc) elements.
 
-We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must constain only one vector, we write $g \in G$.
+\item
+A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
+\end{itemize}
+For notational convenience, we will refer to cells with $C$,  to vectors with $V$, and to groups and subgroups with $G$. 
+When we do not want to distinguish between cells, columns, rows, vectors or groups, we shall talk about a {\em set of cells}.
+\luc{add an example of group/ subgroup}
 
-A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
-Let us elaborate on each of them.
-\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.
-\textit{\CSignature} is a constraint specifying the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g.
+\tias{Use generic constraint}
+A \textit{constraint} definition is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. 
+\begin{itemize}
+\item
+\textit{\CName} is the name of the constraint together with the name of the groups, 
+\item \textit{\CSignature} is specifying a set of restrictions on the types and variables of the constraint, 
+% the properties of the group assignments, corresponding to the variables in \CName, such as their types, 
+e.g., requiring the variables to be integers or constraining the sizes, e.g.
 the length of vectors in the arguments must be equal.
-These are constraints on the group meta-information not on the actual group content.
-\textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
-For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_y$ = RANK($v_x$)}, where $v_x$ and $v_y$ are the variables; its \CSignature is: the subgroup assigned to $v_x$ must be numeric, $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_x$ has the rank (possibly with ties) specified in $g_y$.
+These are restrictions on the group meta-information not on the actual group content.
+\item \textit{\CFunction} defines the actual constraint that needs to be satisfied in the data.
+\end{itemize}
 
-Let us now define what is a tabular solution for a spreadsheet constraint in general.
-Let $C$ denote a spreadsheet constraint, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
+\begin{example}
+The constraint \textit{rank} has 
+\begin{itemize}
+\item \CName = \textit{$g_Y$ = RANK($g_X$)}, where $g_X$ and $g_Y$ are the group variables; 
+\item \CSignature = subgroup $g_X$ (associated with X) is numeric, $g_Y$ is integer and the length of vectors in $g_X$ is the same as in $g_Y$; and
+\item
+\CFunction = \textit{$g_Y$ = RANK($g_X$)} iff $g_X \in G_X, g_Y \in G_Y$ and each value in $g_X$ has the rank (possibly with ties) specified in $g_Y$.
+\end{itemize}
+\end{example}
+
+\samuel{2nd problem statement (from approach), remove EQUAL}
+
+We can now define a classical constraint satisfaction problem on tables. 
+A {\em constraint satisfaction problem} (CSP)  consists of 1) a set of cells, and 2) a set of constraints over those cells. 
+Notice that the constraints specify the types (or the domains) that one encounters in traditional CSPs.  
+Notice also, that as in traditional CSPs, the CSP problem defines the variables (the cells) of the problem
+as well as the constraint that should hold amongst them. The task of CSP-solvers is then to find an assignment 
+to the variables that satisfies all the constraints. The distinction between variables (cells) and their assignments (their values)
+is important here. So, when we speak about cells, vectors and groups, we refer to the variables;
+when we talk about values, assignments or instances, we refer to the values these cells take in a particular "instantiated" table.
+
+While the traditional CSP problem is given a set of cells and a set of constraints,
+and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem}.
+the can be defined as follows.\\
+\\
+{\bf Given }
+\begin{itemize}
+\item
+a set of instantiated tables (so all cells have values) ${\cal T}$; 
+\item
+a set of possible constraints ${\cal C}$; 
+\ite
+a set of possible groups for the tables ${\cal G}$;
+\end{itemize}
+\noindent
+{\bf Find}  the set $C\subset {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
+for which
+all $G_i \in {\cal G}$ and  3)  $c(G_1, ... , G_N)$ satisfies the signature of the constraint $c$.
+
+\luc{ADD/REFER TO EXAMPLE}
+
+%A \textit{tabular solution} $S$ to a spreadsheet constraint $C$ (with $n$ variables) is a subgroup assignment $(g_1, \dots, g_n)$ where $g_i$, such that $(g_1,\dots,g_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+\sergey{to finish this definition later, need to say that $g_i$ is the subgroup assigned to the $i$-th variable in $C$}
 
@@
 In practice, however, things are different.
-Assume that for some $g_x$ both \textit{series($g_x$)} and \textit{alldifferent($g_x$)} hold.
-From a user's perspective the last solution is useless, since he or she already knows that the series constraint implies the alldifferent constraint.
+Assume that for some $X$ both \textit{series(X)} and \textit{alldifferent(X)} hold.
+From a user's perspective the last solution is useless, since he or she already knows that series constraint implies alldifferent constraint.
 To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity (generality) order between constraints, which is depicted using solid (dashed) arrows in Figure \ref{fig:learning_order}.
-This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all.
-A solid lines from $C_1$ to $C_2$ indicates that $C_2$ is \textit{more specific} than $C_1$ and if $S$ is not a solution to $C_1$, then it is not a solution to $C_2$.
-A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is \textit{a more specific} constraint than $C_2$ and therefore if $S$ is a solution for $C_1$ it should not be tested for $C_2$.
+This matters from implementation perspective as well, if we discovered that $X$ is a not permutation, then $X$ should not be tested for the series constraint at all.
+A solid lines from $C_1$ to $C_2$ indicates that $C_2$ is \textit{more specific} than $C_1$ and if $X$ is not a solution to $C_1$, then it is not a solution to $C_2$.
+A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is \textit{a more specific} constraint than $C_2$ and therefore if $X$ is a solution for $C_1$ it should not be tested for $C_2$.
 
@@
 \paragraph{Example}
-Consider the \ecfkey{FK}{PK} (foreign-key) constraint: for every solution the subgroup (vector) $v_{PK}$ assigned to \textit{PK} must fulfill \ecalldiff{v_{PK}}.
+Consider the \ecfkey{FK}{PK} (foreign-key) constraint: for every solution the subgroup (vector) $G_{PK}$ assigned to $PK$ must fulfill \ecalldiff{G_{PK}}.
 Therefore, we say indicate \textit{foreign-key} as being dependent on \textit{all-different} and the latter will be treated first.
-The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties the vector $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+The \ecrank{Y}{X} constraint does not have any dependencies, however in a restricted version without ties the vector $Y$ would always be a permutation of numbers $1, 2, ..., length(Y)$.
 
@@
 \paragraph{Example}
-The \ecrank{v_y}{v_x} constraint requires $v_y$ to be integer or textual and $v_x$ to be numeric.
+The \ecrank{Y}{X} constraint requires $Y$ to be integer or textual and $X$ to be numeric.
 This allows to filter candidate groups for these variables.
-Additionally, $v_y$ and~$v_x$ must be vectors of the same length, therefore, only groups with the same length are considered.
-Groups that does not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{v_y}{v_x}.
+Additionally, $Y$ and~$X$ must be vectors of the same length, therefore, only groups with the same length are considered.
+Groups that does not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{Y}{X}.
 
@@
 \paragraph{Example}
-Consider a candidate assignment $G_x, G_y$ associated with the variables $v_x,v_y$ of \ecrank{v_y}{v_x}, then \findassignment selects single vectors $g_y \in G_y, g_x \in G_x$ such that \ecrank{g_y}{g_x} is true, i.e.
-it searches for a mapping $S$ where $v_x$ ($v_y$) must be mapped to a subgroup of $G_x$ ($G_y$).
-For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_x$ and $G_y$.
-Then \findassignment could select $g_x$ to be $\range{T_1}{\rangeall}{7}$ and $g_y$ to be $\range{T_1}{\rangeall}{8}$ which forms a valid solution.
+Consider a candidate assignment $G_X, G_Y$ associated with the variables $X,Y$ of \ecrank{Y}{X}, then \findassignment selects single vectors $G_X^{i}, G_Y^{j}$ such that \ecrank{G_Y^{j}}{G_X^{i}} is true.
+For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_X$ and $G_Y$.
+\findassignment could then select $G_X^{i}$ to be $\range{T_1}{\rangeall}{7}$ and $G_Y^{j}$ to be $\range{T_1}{\rangeall}{8}$ which forms a valid solution.
 \\\\
@@
 
-Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native Python CSP solver \cite{python_constraint}.
-All the constraints mentored in the paper can be run purely in the internal Python engine without any external software.
-The future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages to allow users to add new constraints in a declarative way.
+Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native python CSP solver \cite{python_constraint}.
+All the constraints mentored in the paper can be run purely in the internal python engine without any external software.
+The future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages.
 
 \subsection{Constraints}
+% As emphasized before, the set of Excel constraints has a partial order in which they should be learned.
+For certain constraints this order does not matter, such as \textit{rank} or \textit{product}.
+For many others they have to be learned in the order, which is specified as a DAG in Figure \ref{fig:learning_order} (if a constraint is not in the graph, it is independent).
+This figure also specifies a potential parallelization of the computation, since each connected component is independent of the others.
 
@@
   \end{tabularx}
-  \caption{Spreadsheet Constraints \sergey{Here a detailed explanation on what is going on, what is essential, what is not? etc should be self-explanatory}}
+  \caption{Spreadsheet Constraints \sergey{Luc, we need your comments here on the notation of $X,Y$ and $G_X,G_Y$}}
   \label{table:constraints}
@@
 \subsection{Case study}
-Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
+Let us illustrate our implementation using the example presented in figure~\ref{fig:main_example}.
 This example combines several smaller examples that were used in an exercise session to teach Excel into one spreadsheet.
@@
   }
-  \caption{Expected constraints in the case study \sergey{in Figure \ref{fig:main_example}; some more explanation needed, not self explanatory}}
+  \caption{Expected constraints case study}
 \end{figure}
@@
 \subsubsection*{Q1. How accurate is the approach?}
-\sergey{that's confusing, it can do 100\% but then all of a sudden 96\%, short explanation needed}
 Fortunately, our system is currently able to find $100\%$ of the expected constraints (90) on the benchmark suite.
@@
 \subsection{Insights}
-\sergey{We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.}
+We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.
 

### 2016-04-14 Paper/ECAI/tabular.tex
added the discussion section with comments
Small corrections

--- 2016-04-14 Paper/ECAI/tabular.tex
@@
 
+\tias{Spreadsheet is an instance of tabular data, but not the only one?}
+
 \tias{Rewrite}
 \section{Introduction}
-Millions of people across use spreadsheets everyday.
+Millions of people across the world use spreadsheets everyday.
 Yet, many people lack a complete understanding of the structures and existing dependencies in the data.
-Indeed, the most useful part of spreadsheets is often the formulas and often these formula disappear, when a file is exported and shared as a CSV file.
+Indeed, the most useful part of spreadsheets is often the formulas and often these formula disappear when a file is exported and shared, for example as a CSV file.
 This brings us to the question: can we discover or reconstruct structural relations in flat tabular spreadsheet data?
-[in a general way that allows declarative specification of constraints to discover]
+Moreover, can we accomplish this task in a general way that allows declarative specification of constraints to discover?
 
-Let us look at an example in Figure \ref{fig:main_example}, the name already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
-If we look at the Table 1 and at the first row in the Table 2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
+\samuel{Maybe slightly more formal:}
+Let us look at an example in Figure~\ref{fig:main_example}, the header names already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
+If we look at Table~1 and at the first row in Table~2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
 Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
@@
 
-It is not hard to see that computation goes not only in columns but in rows as well: if we sum the values in Table 1 in the first row in all four quarters, we would obtain the seventh column in Table 1.
+Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
+\samuel{Maybe slightly more formal:}
 That gives us already a hint how this problem is different from the standard data mining settings, when the data is in rows and variables are in columns.
@@
 
-This makes tabular data understanding and constraint learning new and challenging, and it is not only challenging for a machine but for a human as well.
-Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the \textit{Spreadsheet risk}\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaws in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
+Given these novel aspects, we conclude that understanding and learning constraints in tabular data is a new and challenging problem setting.
+Futhermore, it is not only challenging for a machine but for a human as well.
+\samuel{Maybe slightly more formal:} Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry.
 Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
@@
 \end{itemize}
-
-\samuel{Spreadsheet risks: financial losses due to errors in spreadsheets}
 
@@
 \subsection{Groups, Type-consistency and Constraints}
-In this work we focus on the spreadsheet data, which is physically one big table, but conceptually consists of multiple tables, see Figure \ref{fig:main_example} for an example.
-Formally, a table is $n \times m$ matrix, each entry is called a \textit{cell}.
-A cell has a type, e.g.
-namely numeric.
-In this work we distinguish the two types of data: numeric and textual.
+In this work we focus on spreadsheet data, which is physically one big table, but conceptually often consists of multiple tables, see Figure~\ref{fig:main_example} for an example.
+Formally, a table is an $n \times m$ matrix, each table entry is called  a \textit{cell}.
+
+Each cell has a {\em type} which can be numeric or textual.
 The numeric type has two subtypes: integers and floats. 
-We also consider the special element called \textit{None}, which has two types: numeric and textual.
+We also consider the special element called \textit{None}, which can be either type: numeric or textual.
+A set of cells is called \textit{type-consistent} iff all cells in the set are either numeric or textual. 
+%Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric subtype by requiring its arguments to be integers.
 
-Based on these types we introduce a notation of type consistency for a set of elements.
-A set is called \textit{type-consistent} iff all elements are either numeric or textual.
-Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric sub-types by requiring its arguments to be integers.
-
-We use the notation $T[a,b]$ to refer to a particular cell in the table $T$ at the row with index $a$ and the column with index $b$.
-To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
-
+We consider various sets of cells in tables: 
+\begin{itemize}
+  \item {\bf Cells}: Each entry in the matrix/table is called  a \textit{cell}.
+  We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$. 
+  \item {\bf Columns and rows}: To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
+  \item 
   A \textit{vector} is either a column or a row that is type-consistent. 
   If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
+  \item
   A \textit{group} is a subrange of vectors with the same orientation in a table.
@@
   We denote as the \textit{length} of a row (column) group $G$, written as \textit{length(G)}, the number of its columns (rows).
-We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if its vectors contain numeric (textual, etc) elements.
+We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if \luc{all?????} its vectors contain numeric (textual, etc) elements.
+  \item
+  A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
+\end{itemize}
+For notational convenience, we will refer to cells with $C$,  to vectors with $V$, and to groups and subgroups with $G$. 
+When we do not want to distinguish between cells, columns, rows, vectors or groups, we shall talk about a {\em set of cells}.
+\luc{add an example of group/ subgroup}
 
-We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must constain only one vector, we write $g \in G$.
+\tias{Use generic constraint}
+A \textit{constraint} definition is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. 
+\begin{itemize}
+\item
+\textit{\CName} is the name of the constraint together with the name of the groups, 
+\item \textit{\CSignature} is specifying a set of restrictions on the types and variables of the constraint, 
+% the properties of the group assignments, corresponding to the variables in \CName, such as their types, 
+e.g., requiring the variables to be integers or constraining the sizes, e.g.
+the length of vectors in the arguments must be equal.
+These are restrictions on the group meta-information not on the actual group content.
+\item \textit{\CFunction} defines the actual constraint that needs to be satisfied in the data.
+\end{itemize}
 
-A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
-Let us elaborate on each of them.
-\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.
-\textit{\CSignature} is a constraint specifying the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g.
-the length of vectors in the arguments must be equal.
-These are constraints on the group meta-information not on the actual group content.
-\textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
-For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_y$ = RANK($v_x$)}, where $v_x$ and $v_y$ are the variables; its \CSignature is: the subgroup assigned to $v_x$ must be numeric, $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_x$ has the rank (possibly with ties) specified in $g_y$.
+\paragraph{Example}
+The constraint \textit{rank} has 
+\begin{itemize}
+\item \CName = \textit{$g_Y$ = RANK($g_X$)}, where $g_X$ and $g_Y$ are the group variables; 
+\item \CSignature = subgroup $g_X$ (associated with X) is numeric, $g_Y$ is integer and the length of vectors in $g_X$ is the same as in $g_Y$; and
+\item
+\CFunction = \textit{$g_Y$ = RANK($g_X$)} iff $g_X \in G_X, g_Y \in G_Y$ and each value in $g_X$ has the rank (possibly with ties) specified in $g_Y$.
+\end{itemize}
 
-Let us now define what is a tabular solution for a spreadsheet constraint in general.
-Let $C$ denote a spreadsheet constraint, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
+\samuel{2nd problem statement (from approach), remove EQUAL}
+
+We can now define a classical constraint satisfaction problem on tables. 
+A {\em constraint satisfaction problem} (CSP)  consists of 1) a set of cells, and 2) a set of constraints over those cells. 
+Notice that the constraints specify the types (or the domains) that one encounters in traditional CSPs.  
+Notice also, that as in traditional CSPs, the CSP problem defines the variables (the cells) of the problem
+as well as the constraint that should hold amongst them. The task of CSP-solvers is then to find an assignment 
+to the variables that satisfies all the constraints. The distinction between variables (cells) and their assignments (their values)
+is important here. So, when we speak about cells, vectors and groups, we refer to the variables;
+when we talk about values, assignments or instances, we refer to the values these cells take in a particular "instantiated" table.
+
+While the traditional CSP problem is given a set of cells and a set of constraints,
+and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem}.
+the can be defined as follows.\\
+\\
+{\bf Given }
+\begin{itemize}
+\item
+a set of instantiated tables (so all cells have values) ${\cal T}$; 
+\item
+a set of possible constraints ${\cal C}$; 
+\item
+a set of possible groups for the tables ${\cal G}$;
+\end{itemize}
+\noindent
+{\bf Find}  the set $C\subset {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
+for which
+all $G_i \in {\cal G}$ and  3)  $c(G_1, ... , G_N)$ satisfies the signature of the constraint $c$.
+
+\luc{ADD/REFER TO EXAMPLE}
+
+%A \textit{tabular solution} $S$ to a spreadsheet constraint $C$ (with $n$ variables) is a subgroup assignment $(g_1, \dots, g_n)$ where $g_i$, such that $(g_1,\dots,g_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
+\sergey{to finish this definition later, need to say that $g_i$ is the subgroup assigned to the $i$-th variable in $C$}
 

### 2016-04-14 Paper/ECAI/tabular.tex
Sergey's version
Merged

@@
   }
-  \caption{Expected constraints case study}
+  \caption{Expected constraints in the case study \sergey{in Figure \ref{fig:main_example}; some more explanation needed, not self explanatory}}
+  \label{fig:sol_example}
 \end{figure}
+\sergey{Samuel, hint, don't put label before caption, it can't resolve it sometimes}
 
 \subsubsection{Results}
-Our current implementation takes a few seconds\footnote{3 seconds on our Macbook Pro} to find 18 constraints, including all of the 12 solutions described in figure~\ref{fig:sol_example}.
+\sergey{Samuel, you should just say at the beginning that all experiments has been run on macbook la-la system and that's it, I guess, I don't think we need to mention it multiple times}
+Our current implementation takes a few seconds\footnote{3 seconds on our Macbook Pro} to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
 Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}
@@
 
+\section{Discussion and Conclusions}
+\sergey{link to the github with implementations and the dataset}
+
+\sergey{Future version, extensions, declarative stuff, bla-bla here}
+
 \bibliographystyle{ecai}

### 2016-04-14 Paper/ECAI/tabular.tex
merged mostly

@@
 
-\tias{Spreadsheet is an instance of tabular data, but not the only one?}
-
-\tias{Rewrite}
 \section{Introduction}
@@
 
-\samuel{Maybe slightly more formal:}
 Let us look at an example in Figure~\ref{fig:main_example}, the header names already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
@@
 Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
-\samuel{Maybe slightly more formal:}
+
 That gives us already a hint how this problem is different from the standard data mining settings, when the data is in rows and variables are in columns.
@@
 Futhermore, it is not only challenging for a machine but for a human as well.
-\samuel{Maybe slightly more formal:} Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry.
+
+Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
 Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
@@
   \item
-  A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
+ A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$. If the subgroup $g$ must contain only one vector, we write $g \in G$.
+%We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
+% A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
 \end{itemize}
-For notational convenience, we will refer to cells with $C$,  to vectors with $V$, and to groups and subgroups with $G$.
+For notational convenience, we will refer to cells with $C$ \sergey{maybe better name? confusing with constaraints},  to vectors with $V$, and to groups and subgroups with $G$.
+
 When we do not want to distinguish between cells, columns, rows, vectors or groups, we shall talk about a {\em set of cells}.
@@
 
-\tias{Use generic constraint}
-A \textit{constraint} definition is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them.
+A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. 
 \begin{itemize}
 \item
-\textit{\CName} is the name of the constraint together with the name of the groups,
-\item \textit{\CSignature} is specifying a set of restrictions on the types and variables of the constraint,
-% the properties of the group assignments, corresponding to the variables in \CName, such as their types,
-e.g., requiring the variables to be integers or constraining the sizes, e.g.
+\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.
+\item
+  \textit{\CSignature} is a constraint specifying the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g.
 the length of vectors in the arguments must be equal.
-These are restrictions on the group meta-information not on the actual group content.
-\item \textit{\CFunction} defines the actual constraint that needs to be satisfied in the data.
-\end{itemize}
+These are constraints on the group meta-information not on the actual group content.
+\item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
 
 \paragraph{Example}
+\sergey{need to use this below:}
+
+  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_y$ = RANK($v_x$)}, where $v_x$ and $v_y$ are the variables; its \CSignature is: the subgroup assigned to $v_x$ must be numeric, $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_x$ has the rank (possibly with ties) specified in $g_y$.
+
 The constraint \textit{rank} has
@@
 \end{itemize}
-
-\samuel{2nd problem statement (from approach), remove EQUAL}
 
@@
 
-%A \textit{tabular solution} $S$ to a spreadsheet constraint $C$ (with $n$ variables) is a subgroup assignment $(g_1, \dots, g_n)$ where $g_i$, such that $(g_1,\dots,g_n)$ satisfies both the \CSignature and \CFunction of the constraint $C$.
-\sergey{to finish this definition later, need to say that $g_i$ is the subgroup assigned to the $i$-th variable in $C$}
+Let us now define what is a tabular solution for a spreadsheet constraint in general.
+Let $C$ denote a spreadsheet constraint, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
 

### 2016-04-14 Paper/ECAI/tabular.tex
fixed typo
Merge

@@
   We denote as the \textit{length} of a row (column) group $G$, written as \textit{length(G)}, the number of its columns (rows).
-We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if \luc{all?????} its vectors contain numeric (textual, etc) elements.
+We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
   \item
@@
 \end{itemize}
-For notational convenience, we will refer to cells with $C$ \sergey{maybe better name? confusing with constaraints},  to vectors with $V$, and to groups and subgroups with $G$.
+For notational convenience, we will refer to vectors with $V$ and to groups and subgroups with $G$.
 
@@
 \luc{add an example of group/ subgroup}
+
+\paragraph{Example}
+Consider Table~1 in the Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
+The (column) vectors of Table~1 are subdivided into five groups (\range{T_1}{\rangeall}{1}, \range{T_1}{\rangeall}{2}, \range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{9}, \range{T_1}{\rangeall}{\rangeto{10}{11}}).
 
@@
 \item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
+  \end{itemize}
 

### 2016-04-14 Paper/ECAI/tabular.tex
General improvements

@@
 When we do not want to distinguish between cells, columns, rows, vectors or groups, we shall talk about a {\em set of cells}.
-\luc{add an example of group/ subgroup}
 
@@
 Consider Table~1 in the Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
-The (column) vectors of Table~1 are subdivided into five groups (\range{T_1}{\rangeall}{1}, \range{T_1}{\rangeall}{2}, \range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{9}, \range{T_1}{\rangeall}{\rangeto{10}{11}}).
+The (column) vectors of Table~1 are subdivided into five groups:
+\begin{align*}
+&G_1 = \range{T_1}{\rangeall}{1},
+&G_2 = \range{T_1}{\rangeall}{2},\\
+&G_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
+&G_4 = \range{T_1}{\rangeall}{9},\\
+&G_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}
+\end{align*}.
+In order to sum the results of a salesperson across quarters we can consider for example the subgroup $\range{T_1}{\rangeall}{\rangeto{3}{6}} \subset G_3$
 
-A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}. Let us elaborate on each of them. 
+\subsection{Constraints}
+A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
+Let us elaborate on each of them. 
 \begin{itemize}
@@
 \item
-  \textit{\CSignature} is a constraint specifying the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName, such as their types, e.g., requiring them to be integers or constraining the sizes, e.g.
-the length of vectors in the arguments must be equal.
-These are constraints on the group meta-information not on the actual group content.
+  \textit{\CSignature} is a set of constraints on the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName.
+  For example, the signature can restrict the types of the groups, e.g., requiring them to be integers or constrain their sizes, e.g. the length of vectors in the arguments must be equal.
+  The constraints act on porperties or group meta-information, not on the actual group content. 
 \item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
@@
 \sergey{need to use this below:}
-
   For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_y$ = RANK($v_x$)}, where $v_x$ and $v_y$ are the variables; its \CSignature is: the subgroup assigned to $v_x$ must be numeric, $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_x$ has the rank (possibly with ties) specified in $g_y$.
@@
 
+\samuel{Should this section maybe be moved to problem statement?}
 We can now define a classical constraint satisfaction problem on tables.
@@
 Notice that the constraints specify the types (or the domains) that one encounters in traditional CSPs.
-Notice also, that as in traditional CSPs, the CSP problem defines the variables (the cells) of the problem
-as well as the constraint that should hold amongst them. The task of CSP-solvers is then to find an assignment
-to the variables that satisfies all the constraints. The distinction between variables (cells) and their assignments (their values)
-is important here. So, when we speak about cells, vectors and groups, we refer to the variables;
-when we talk about values, assignments or instances, we refer to the values these cells take in a particular "instantiated" table.
+Notice also, that as in traditional CSPs, the CSP problem defines the variables (the cells) of the problem as well as the constraint that should hold amongst them.
+The task of CSP-solvers is then to find an assignment to the variables that satisfies all the constraints.
+The distinction between variables (cells) and their assignments (their values) is important here.
+So, when we speak about cells, vectors and groups, we refer to the variables; when we talk about values, assignments or instances, we refer to the values these cells take in a particular "instantiated" table.
 
 While the traditional CSP problem is given a set of cells and a set of constraints,
-and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem}.
-the can be defined as follows.\\
+and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
 \\
@@
 \item
-a set of possible constraints ${\cal C}$;
+a set of spreadsheet constraints ${\cal C}$;
 \item
-a set of possible groups for the tables ${\cal G}$;
+a set of groups ${\cal G}$ for the tables ${\cal T}$;
 \end{itemize}
 \noindent
+\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a spreadsheet constraint}
 {\bf Find}  the set $C\subset {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
@@
 \sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
-As we have seen the statement before we assume the set of groups to be given.
+As we have seen in the statement before we assume the set of groups to be given.
 It is often, however, not the case and a possible approach is to extract the groups from the CSV file directly.
-In our system, we have implement the group construction step using the type-consistency check.
+In our system, we have implemented the group construction step using the type-consistency check.
 \sergey{Samuel, we need details here, I think} In general, separating data from a header is a task on its own  \cite{header} and goes beyond the scope of this paper.
@@
 
-\subsection{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG \dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of \dependencies.
+\subsection{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of~\dependencies.
 Solid edges point from more general to more specific constraints.
-Therefore, if there is a solid edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ must contain the subgroups satisfying $C_1$.
+Therefore, if there is a solid edge from constraint~$C_1$ to constraint~$C_2$, then a set of subgroups satisfying~$C_2$ must contain the subgroups satisfying~$C_1$.
 
@@
 \paragraph{Example}
-Consider the \ecfkey{FK}{PK} (foreign-key) constraint: for every solution the subgroup (vector) $v_{PK}$ assigned to \textit{PK} must fulfill \ecalldiff{v_{PK}}.
+\samuel{Fix notation variable vs. group}
+Consider the \ecfkey{v_{fk}}{v_{pk}} (foreign-key) constraint: for every solution the subgroup (vector) $V$ assigned to $\mathit{v_{pk}}$ must fulfill \ecalldiff{V}.
 Therefore, we say indicate \textit{foreign-key} as being dependent on \textit{all-different} and the latter will be treated first.
@@
 Finding candidate groups for a spreadsheet constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
-Several constraints (e.g. sum, max, count), however, have the same candidate generation procedure as they share the same \CSignature.
+Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.
 
 \paragraph{Example}
-The \ecrank{v_y}{v_x} constraint requires $v_y$ to be integer or textual and $v_x$ to be numeric.
+The \ecrank{v_y}{v_x} constraint requires $v_y$ to consist of only integer or only textual data and $v_x$ to be numeric.
 This allows to filter candidate groups for these variables.
 Additionally, $v_y$ and~$v_x$ must be vectors of the same length, therefore, only groups with the same length are considered.
-Groups that does not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{v_y}{v_x}.
+Groups that do not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{v_y}{v_x}.
 
@@
 $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
+Every spreadsheet constraint~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.
 
@@
 \\\\
-For some constraints (most notably aggregates) there may be multiple overlapping subsets that satisfy the constraint.
+For some constraints (most notably aggregates) there may be multiple overlapping subsets of a given group~$G$ that satisfy the constraint.
 We chose to omit subsets that are fully contained within a larger subset since this can lead to overfitting.
-If the returned subsets are too large and accidentally contains unrelated vectors, the user can split the group to reflect this structure.
+If the returned subsets are too large and accidentally contain unrelated vectors, the user can split such a group, disallowing.
 
-\subsection{Implementation note} Essentially for each constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
+\subsection{Implementation note} Essentially for each spreadsheet constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
 Hence, it is natural to solve them using declarative solvers.
@@
 It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem.
-Furthermore, the declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
+Furthermore, declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
 
 Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native Python CSP solver \cite{python_constraint}.
-All the constraints mentored in the paper can be run purely in the internal Python engine without any external software.
-The future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages to allow users to add new constraints in a declarative way.
+All the constraints mentioned in the paper can be run purely using the internal Python engine without any external software.
+Future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages to allow users to add new constraints in a declarative way.
 
@@
 
-\newcommand{\numeric}{\format{numeric}}
-\newcommand{\textual}{\format{textual}}
-\newcommand{\integer}{\format{integer}}
-\newcommand{\length}{\format{length}}
+\newcommand{\numeric}[1]{\format{numeric}(#1)}
+\newcommand{\textual}[1]{\format{textual}(#1)}
+\newcommand{\integer}[1]{\format{integer}(#1)}
+\newcommand{\length}[1]{\format{length}(#1)}
 \newcommand{\nat}{\mathcal{N}}
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
-    \ecalldiff{V_x} & X is integer or textual & All elements in X are different \\
+    \ecalldiff{v_x} & $\integer{v_x}$ or $\textual{v_x}$ & All elements in $v_x$ are different \\
     %X = Y & & \\
-    \ecfkey{V_{fk}}{V_{pk}} & FK and PK have the same type; they originate from different tables and PK is all-different & all values in FK exist in PK \\
+    \ecfkey{v_{fk}}{v_{pk}} & FK and PK have the same type; they originate from different tables and PK is all-different & all values in FK exist in PK \\
     % TODO check out
-    \eclookupprod{V_r}{V_1}{V_{fk}}{V_{pk}}{V_{val}} & & \\
-    \eclookup{V_r}{V_{fk}}{V_{pk}}{V_{val}} & & \\
-    \eclookupfuzzy{V_r}{V_{fk}}{V_{pk}}{V_{val}} & & \\
-    \ecperm{V_x} & & \\
-    \ecprod{V_r}{V_1}{V_2} & & \\
-    \ecproj{V_r}{G_x} & & \\
-    \ecrank{V_r}{V_1} & Y and X have the same length & Y is the rank of X (including ties)\\
-    \ectotal{V_r}{V_{pos}}{V_{neg}} & A, P and N have the same length and their length is at least 2 & *\\
-    \ecseries{V_1} & & \\
-    \ecsumc{V_r}{G_1} & & Y is the result of summing each column in X \\
-    \ecsumr{V_r}{G_1} & & Y is the result of summing each row in X \\
-    \ecsumif{V_r}{V_{fk}}{V_{pk}}{V_{val}} & & \\
-    \ecsumprod{V_r}{V_1}{V_2} & & \\
+    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}} & & \\
+    \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
+    \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
+    \ecperm{v_x} & & \\
+    \ecprod{v_r}{v_1}{v_2} & & \\
+    \ecproj{v_r}{\mathbf{v_x}} & & \\
+    \ecrank{v_r}{v_1} & Y and X have the same length & Y is the rank of X (including ties)\\
+    \ectotal{v_r}{v_{pos}}{v_{neg}} & A, P and N have the same length and their length is at least 2 & *\\
+    \ecseries{v_x} & & \\
+    \ecsumc{v_r}{\mathbf{v_x}} & & Y is the result of summing each column in X \\
+    \ecsumr{v_r}{\mathbf{v_x}} & & Y is the result of summing each row in X \\
+    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
+    \ecsumprod{v_r}{v_1}{v_2} & & \\
 
@@
 
+All experiments were run on a Macbook Pro, Intel Core i7 2.3 GHz with 16GB RAM.
+
 \subsection{Case study}
@@
 \sergey{Samuel, you should just say at the beginning that all experiments has been run on macbook la-la system and that's it, I guess, I don't think we need to mention it multiple times}
-Our current implementation takes a few seconds\footnote{3 seconds on our Macbook Pro} to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
-Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}
-  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}} \\
-  & \ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{4}}
+Our current implementation takes a few seconds to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
+Of the 6 remaining (redundant) constraints, 5 are $\mathit{rank}$ constraints that are true by accident, such as: \begin{align*}
+  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}
 \end{align*}
-The other redundant constraint is an additional $\mathit{LOOKUP}$ that holds because the vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).
+The other redundant constraint is an additional $\mathit{lookup}$ that holds because the two vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).
 
@@
 All spreadsheets have been converted to CSV with no manual intervention unless noted.
-Definitions of tables where added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).
+Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).
 
@@
 \begin{description}
-  \item[Expected] Constraint that have been implemented and are expected to be found
+  \item[Implemented] Constraint that have been implemented and are expected to be found
   \item[Essential] Constraint that have or have not been implemented but could be found using our algorithm
@@
 \sergey{that's confusing, it can do 100\% but then all of a sudden 96\%, short explanation needed}
-Fortunately, our system is currently able to find $100\%$ of the expected constraints (90) on the benchmark suite.
-There are only three essential constraints that have not been implemented yet, therefore our system currently identifies $96.77\%$ of the constraints that we hope to discover.
+Fortunately, our system is currently able to find $100\%$ of the implemented constraints (90) on the benchmark suite.
+There are only three essential constraints that have not been implemented yet, therefore our system currently identifies $96.77\%$ of the essential constraints.
 
@@
 Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
-For the 32 spreadsheets in the benchmark our implementation ran in $17.86s \pm 0.87s$\footnote{Macbook Pro, Intel Core i7 2.3 GHz, 16GB RAM}.
+For the 32 spreadsheets in the benchmark our implementation ran in $17.86s \pm 0.87s$.
 The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
@@
 This avoids some of the explosion of combinations for constraints that have variables.
-For example, using foreign keys as a base constraint to find conditional aggregates reduces the running time for the case study from about 3 seconds to under 1 second.
-Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys present.
+For example, using foreign keys as a base constraint to find conditional aggregates reduces the running time for the case study from about 3 seconds to below 1 second.
+Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.
 
-\subsection{Insights}
+\section{Discussion}
 \sergey{We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.}

### 2016-04-14 Paper/ECAI/tabular.tex
Signature changes

@@
 \sergey{need to use this below:}
-  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_y$ = RANK($v_x$)}, where $v_x$ and $v_y$ are the variables; its \CSignature is: the subgroup assigned to $v_x$ must be numeric, $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_x$ has the rank (possibly with ties) specified in $g_y$.
+  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subset G_x, G_y' \subset G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
 
-The constraint \textit{rank} has
-\begin{itemize}
-\item \CName = \textit{$g_Y$ = RANK($g_X$)}, where $g_X$ and $g_Y$ are the group variables;
-\item \CSignature = subgroup $g_X$ (associated with X) is numeric, $g_Y$ is integer and the length of vectors in $g_X$ is the same as in $g_Y$; and
-\item
-\CFunction = \textit{$g_Y$ = RANK($g_X$)} iff $g_X \in G_X, g_Y \in G_Y$ and each value in $g_X$ has the rank (possibly with ties) specified in $g_Y$.
-\end{itemize}
+%The constraint \textit{rank} has
+%\begin{itemize}
+%\item \CName = \textit{$v_r$ = RANK($v_x$)}, where $v_r$ and $v_x$ are %the variables;
+%\item \CSignature = subgroup $g_X$ (associated with X) is numeric, $g_Y$ %is integer and the length of vectors in $g_X$ is the same as in $g_Y$; and
+%\item
+%\CFunction = \textit{$g_Y$ = RANK($g_X$)} iff $g_X \in G_X, g_Y \in G_Y$ %and each value in $g_X$ has the rank (possibly with ties) specified in $%g_Y$.
+%\end{itemize}
 
@@
     \ecproj{v_r}{\mathbf{v_x}} & & \\
-    \ecrank{v_r}{v_1} & Y and X have the same length & Y is the rank of X (including ties)\\
+    \ecrank{v_r}{v_x} & Y and X have the same length & Y is the rank of X (including ties)\\
     \ectotal{v_r}{v_{pos}}{v_{neg}} & A, P and N have the same length and their length is at least 2 & *\\
@@
 \begin{description}
-  \item[Implemented] Constraint that have been implemented and are expected to be found
-  \item[Essential] Constraint that have or have not been implemented but could be found using our algorithm
+  \item[Implemented] Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
+  \item[Essential] Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
   \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)

### 2016-04-14 Paper/ECAI/tabular.tex
adding comments

@@
 &G_4 = \range{T_1}{\rangeall}{9},\\
-&G_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}
-\end{align*}.
+&G_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}.
+\end{align*}
 In order to sum the results of a salesperson across quarters we can consider for example the subgroup $\range{T_1}{\rangeall}{\rangeto{3}{6}} \subset G_3$
@@
 \subsection{Constraints}
+\sergey{write correspondence with ILP terminology}
 A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
@@
 \sergey{need to use this below:}
-  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subset G_x, G_y' \subset G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
+  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
 
@@
 \samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a spreadsheet constraint}
-{\bf Find}  the set $C\subset {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
+{\bf Find}  the set $C\subseteq {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
 for which

### 2016-04-14 Paper/ECAI/tabular.tex
adding comments

@@
 
-\subsection{Constraint ordering} $\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of~\dependencies.
+\subsection{Constraint ordering} 
+\sergey{refinement operator, we need emphasize it}
+$\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of~\dependencies.
 Solid edges point from more general to more specific constraints.
@@
 \section{Evaluation}
+\sergey{we need to add a summary of the dataset, avg number of constraints, cells, rows, columns}
 In this section we experimentally validate our approach.

### 2016-04-14 Paper/ECAI/tabular.tex
fixing the ordering issue

@@
 \newcommand{\postprocess}{\format{pruneRedundant}}
-\newcommand{\constrainttorder}{\format{orderedConstraints}}
+\newcommand{\constrainttorder}{\format{generalityOrder}}
 
@@
 \subsection{Constraint ordering} 
-\sergey{refinement operator, we need emphasize it}
+Since some constraints are more general than the other, we generate them in the generality order similar to the ILP techniques known as refinement operators \cite{luc_book}.
+
 $\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of~\dependencies.

### 2016-04-14 Paper/ECAI/tabular.tex
added ILP lines

@@
 \item
-\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.
+\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
 \item
@@
   For example, the signature can restrict the types of the groups, e.g., requiring them to be integers or constrain their sizes, e.g. the length of vectors in the arguments must be equal.
-  The constraints act on porperties or group meta-information, not on the actual group content. 
+  The constraints act on properties or group meta-information, not on the actual group content.
+In ILP, it is known as mode declarations, when one needs to declare types of variables in the vocabulary.
+Also, it is analogous to standard linear algebra well-defined requirements, e.g., in order to multiply matrices $A$ and $B$, $A$ must have the size of $n \times k$ and $B$ the size $k \times m$.
 \item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
+In ILP, it is known as bias, where we constrain ourselves to only particular subset of language in which we perform the search.
 \end{itemize}

### 2016-04-14 Paper/ECAI/tabular.tex
Table

@@
 
-\newcommand{\numeric}[1]{\format{numeric}(#1)}
-\newcommand{\textual}[1]{\format{textual}(#1)}
-\newcommand{\integer}[1]{\format{integer}(#1)}
-\newcommand{\length}[1]{\format{length}(#1)}
+\newcommand{\numeric}{\format{numeric}}
+\newcommand{\textual}{\format{textual}}
+\newcommand{\integer}{\format{integer}}
+\newcommand{\plength}{\format{length}}
+\newcommand{\ptype}{\format{type}}
+\newcommand{\ptable}{\format{table}}
+\newcommand{\por}{\format{orientation}}
 \newcommand{\nat}{\mathcal{N}}
@@
   \centering
-  \begin{tabularx}{\textwidth}{XXX}
+  \begin{tabularx}{\textwidth}{l X X}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
-    \ecalldiff{v_x} & $\integer{v_x}$ or $\textual{v_x}$ & All elements in $v_x$ are different \\
+    \ecalldiff{v_x} & $\integer(v_x)$ or $\textual(v_x)$ & All elements in $v_x$ are different \\
     %X = Y & & \\
-    \ecfkey{v_{fk}}{v_{pk}} & FK and PK have the same type; they originate from different tables and PK is all-different & all values in FK exist in PK \\
+    \ecfkey{v_{fk}}{v_{pk}} & $\ptype(v_{fk})$ = $ptype(v_{pk})$,$\ptable(v_{fk}) \neq \ptable(v_{pk})$, $\ecalldiff{v_{pk}}$ & Every value in @fk also exist in @pk \\
     % TODO check out
-    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}} & & \\
-    \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
+    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
+      & the variables in $\{@r, @1, @fk\}$ and in $\{@pk, @val\}$ all have the same \plength, \ptable and \por.
+      & $@r$ is the obtained by multiplying $@1$ with $\eclookup{@fk}{@pk}{@val}$  
+      \\
+    \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
+      & the variables in $\{@fk, @val\}$ and in $\{@pk, @r\}$ all have the same \plength, \ptable and \por; $@r$ and~$@val$ have the same type; and \ecfkey{@fk}{@pk}.
+      & Lookup $@fk$ in $@pk$ and return the corresponding value in $@val$.
+      \\
     \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
@@
 
-\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}
-
   \end{tabularx}
-  \caption{Spreadsheet Constraints \sergey{Here a detailed explanation on what is going on, what is essential, what is not? etc should be self-explanatory}}
+  \caption{Spreadsheet Constraints \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
+etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}
+}
   \label{table:constraints}

### 2016-04-14 Paper/ECAI/tabular.tex
Removed comments and corrected lookup

@@
     \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
-      & the variables in $\{@fk, @val\}$ and in $\{@pk, @r\}$ all have the same \plength, \ptable and \por; $@r$ and~$@val$ have the same type; and \ecfkey{@fk}{@pk}.
+      & the variables in $\{@fk, @r\}$ and in $\{@pk, @val\}$ all have the same \plength, \ptable and \por; $@r$ and~$@val$ have the same type; and \ecfkey{@fk}{@pk}.
       & Lookup $@fk$ in $@pk$ and return the corresponding value in $@val$.
@@
 \subsubsection{Results}
-\sergey{Samuel, you should just say at the beginning that all experiments has been run on macbook la-la system and that's it, I guess, I don't think we need to mention it multiple times}
 Our current implementation takes a few seconds to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
@@
 \subsubsection*{Q1. How accurate is the approach?}
-\sergey{that's confusing, it can do 100\% but then all of a sudden 96\%, short explanation needed}
 Fortunately, our system is currently able to find $100\%$ of the implemented constraints (90) on the benchmark suite.

### 2016-04-15 Paper/ECAI/tabular.tex
Small fix
fixed ILP terminology

@@
   \item
- A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$. If the subgroup $g$ must contain only one vector, we write $g \in G$.
+ A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$.
+If the subgroup $g$ must contain only one vector, we write $g \in G$.it defines what it means for a constraint be well-formed
 %We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
@@
   For example, the signature can restrict the types of the groups, e.g., requiring them to be integers or constrain their sizes, e.g. the length of vectors in the arguments must be equal.
-  The constraints act on properties or group meta-information, not on the actual group content.
-In ILP, it is known as mode declarations, when one needs to declare types of variables in the vocabulary.
-Also, it is analogous to standard linear algebra well-defined requirements, e.g., in order to multiply matrices $A$ and $B$, $A$ must have the size of $n \times k$ and $B$ the size $k \times m$.
-\item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
-In ILP, it is known as bias, where we constrain ourselves to only particular subset of language in which we perform the search.
+  The constraints act on properties or group meta-information, not on the actual group content. In ILP, it is known as bias, it defines what it means for a constraint be well-formed.
+  \item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function. In ILP, it is known as background knowledge, which provides a definition to a predicate.
 \end{itemize}
@@
 The \ecrank{v_y}{v_x} constraint requires $v_y$ to consist of only integer or only textual data and $v_x$ to be numeric.
-This allows to filter candidate groups for these variables.
+This allows us to filter candidate groups for these variables.
 Additionally, $v_y$ and~$v_x$ must be vectors of the same length, therefore, only groups with the same length are considered.

### 2016-04-15 Paper/ECAI/tabular.tex
Table WIP

--- 2016-04-14 Paper/ECAI/tabular.tex
@@
       & the variables in $\{@r, @1, @fk\}$ and in $\{@pk, @val\}$ all have the same \plength, \ptable and \por.
-      & $@r$ is the obtained by multiplying $@1$ with $\eclookup{@fk}{@pk}{@val}$  
+      & $@r$ is the obtained by multiplying $@1$ with $\eclookup{@fk}{@pk}{@val}$.
       \\
@@
       \\
-    \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
-    \ecperm{v_x} & & \\
-    \ecprod{v_r}{v_1}{v_2} & & \\
-    \ecproj{v_r}{\mathbf{v_x}} & & \\
-    \ecrank{v_r}{v_x} & Y and X have the same length & Y is the rank of X (including ties)\\
-    \ectotal{v_r}{v_{pos}}{v_{neg}} & A, P and N have the same length and their length is at least 2 & *\\
-    \ecseries{v_x} & & \\
-    \ecsumc{v_r}{\mathbf{v_x}} & & Y is the result of summing each column in X \\
+    \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}}
+      & \textit{Same as lookup}
+      & Lookup the first item in $@pk$ smaller than $@fk$ and return the corresponding value in $@val$.
+      \\
+    \ecperm{v_x}
+      & $\numeric(@x)$, $\ecalldiff{@x}$
+      & The values in $@x$ are a permutation of the numbers $1$ through $\plength(@x)$.
+      \\
+    \ecprod{v_r}{v_1}{v_2}
+      & variables $\{@r, @1, @2\}$ are all $\numeric$ and have the same $\plength$
+      & Every element in $@r$ is the product of the corresponding elements in~$@1$ and~$@2$.
+      \\
+    \ecproj{v_r}{\mathbf{v_x}}
+      & $\{@r, @x\}$ all have the same $\plength$, $\por$, $\ptable$ and type; @x contains at least 2~vectors
+      & At every position in $1$ through $\plength(@x)$ there is exactly one vector in $@x$ with a non-blank value at that position which is the same as the corresponding value in $@r$.
+      \\
+    \ecrank{v_r}{v_x}
+      & $\integer(@r)$; $\numeric(@x)$; and $\plength(@r) = \plength(@x)$
+      & The values in $@r$ represent the rank (from largest to smallest) of the values in ($@x$) - including ties
+      \\
+    \ectotal{v_r}{v_{pos}}{v_{neg}}
+      & $\{@r, @pos, @neg\}$ all have the same $\plength$, which is at least $2$
+      & The values in $@r$ are a running total, each value $@r[i] = @r[i - 1] + @pos[i] - @neg[i]$.
+      \\
+    \ecseries{v_x}
+      & $\numeric(@x)$ and $\ecperm{@x}$
+      & $@x[1] = 1$ and $@x[i] = @x[i - 1] + 1$.
+      \\
+    \ecsumc{v_r}{\mathbf{v_x}}
+      &
+      & Y is the result of summing each column in X \\
     \ecsumr{v_r}{\mathbf{v_x}} & & Y is the result of summing each row in X \\

### 2016-04-15 Code/src/core/constraint.py
Code fix

@@ -399,7 +399,8 @@ class SumProduct(Constraint):
         variables = [self.result, self.first, self.second]
         source = Source(variables)
         filters = [SameLength([self.first, self.second]), NotPartial(variables),
-                   SizeFilter([self.first, self.second], length=2), SizeFilter([self.result], rows=1, cols=1)]
+                   SizeFilter([self.first, self.second], length=2), SizeFilter([self.result], rows=1, cols=1),
+                    SizeFilter([self.result], rows=1, cols=1, max_size=True)]
         super().__init__("sum-product", "{R} = SUMPRODUCT({O1}, {O2})", source, filters)
 
 

### 2016-04-15 Code/src/core/constraint.py
-

@@ -400,7 +400,7 @@ class SumProduct(Constraint):
         source = Source(variables)
         filters = [SameLength([self.first, self.second]), NotPartial(variables),
                    SizeFilter([self.first, self.second], length=2), SizeFilter([self.result], rows=1, cols=1),
-                    SizeFilter([self.result], rows=1, cols=1, max_size=True)]
+                   SizeFilter([self.result], rows=1, cols=1, max_size=True)]
         super().__init__("sum-product", "{R} = SUMPRODUCT({O1}, {O2})", source, filters)
 
 

### 2016-04-15 Paper/ECAI/tabular.tex
typo

@@
 \newcommand{\constrainttorder}{\format{generalityOrder}}
+\newcommand{\template}{\format{Constraint template}}
 
@@
  A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$.
-If the subgroup $g$ must contain only one vector, we write $g \in G$.it defines what it means for a constraint be well-formed
+If the subgroup $g$ must contain only one vector, we write $g \in G$.it defines what it means for a constraint to be well-formed
 %We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
@@
 \subsection{Constraints}
-\sergey{write correspondence with ILP terminology}
-A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
+A \template is a triple \textit{(\CName, \CSignature, \CFunction)}.
 Let us elaborate on each of them. 
@@
 \sergey{need to use this below:}
-  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
+  For example, a \template \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
 
@@
 \item
-a set of spreadsheet constraints ${\cal C}$;
+a set of \template ${\cal C}$;
 \item
@@
 \noindent
-\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a spreadsheet constraint}
+\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a \template}
 {\bf Find}  the set $C\subseteq {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
@@
 
-Let us now define what is a tabular solution for a spreadsheet constraint in general.
-Let $C$ denote a spreadsheet constraint, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
+Let us now define what is a tabular solution for a \template in general.
+Let $C$ denote a \template, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
 
@@
 In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
-Here we formalize the statement in terms of spreadsheet constraints and group assignments as follows:
+Here we formalize the statement in terms of \template and group assignments as follows:
 
@@
     &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$\\
+    \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
     \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ in \constraints \\
@@
     &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$,\\
+    \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
     & a constraint specificity DAG \dependencies \\
@@
     \footnotesize
-    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Spreadsheet constraints, \dependencies -- dependency graph
+    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- \template's, \dependencies -- dependency graph
     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
@@
 
-Finding candidate groups for a spreadsheet constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
+Finding candidate groups for a \template~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
 Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.
@@
 $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
-Every spreadsheet constraint~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.
+Every \template~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.
 
@@
 
-\subsection{Implementation note} Essentially for each spreadsheet constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
+\subsection{Implementation note} Essentially for each \template both \generategroups and \findassignment are separate constraint satisfaction problems.
 Hence, it is natural to solve them using declarative solvers.
@@
   \end{tabularx}
-  \caption{Spreadsheet Constraints \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
+  \caption{\template's \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
 etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}

### 2016-04-15 Paper/ECAI/tabular.tex
reworking the statement and formalization

@@
 
-\newcommand{\constraints}{\ensuremath{\mathcal{C}}\xspace}
+\newcommand{\constraints}{\ensuremath{\mathcal{T}}\xspace}
 \newcommand{\format}[1]{\textit{#1}\xspace}
@@
 
-Let us now define what is a tabular solution for a \template in general.
-Let $C$ denote a \template, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
+Let us introduce \template instantiation syntax.
+Let $S$ be a mapping from variables to the subgroups of \groups and $T$ be a \template, then $T^S$ is the template, in which all variables in all constrainst in \CSignature and \CFunction are instantiated with the corresponding subgroups from $S$.
+We refer to $T^S$ as a \textit{constraint} and to the instantiated \CSignature (\CFunction) as $\text{\CSignature}^S$ ($\text{\CFunction}^S$).
+
+Let us now define what it means for a constraint from \template to hold in general.
+Let $T$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $T^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
 
@@
     \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
-    \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ in \constraints \\
+    \textbf{Find:}&  all constraints $C$ over \groups for each template $t$ in \constraints \\
   \end{tabular}
@@
 This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all.
-A solid lines from $C_1$ to $C_2$ indicates that $C_2$ is \textit{more specific} than $C_1$ and if $S$ is not a solution to $C_1$, then it is not a solution to $C_2$.
-A dashed arrow from $C_1$ to $C_2$ indicates that $C_1$ is \textit{a more specific} constraint than $C_2$ and therefore if $S$ is a solution for $C_1$ it should not be tested for $C_2$.
+A solid lines from $t_1$ to $t_2$ indicates that $t_2$ is \textit{more specific} than $t_1$ and if $C_1=t_1^S$ from $t_1$ does not hold, then the constraint $C_2=t_2^S$ from $t_2$ does not hold neither.
+A dashed arrow from $t_1$ to $t_2$ indicates that $t_1$ is \textit{a more specific} constraint than $t_2$ and therefore if $C=t_1^S$ from $t_1$ holds it should not be tested for $t_2$.
 
-We call a tabular solution $S$ to a constraint $C$ \textit{the most specific} wrt \dependencies, iff there is no other constraint $C'$ for which $S$ is a solution and $C'$ is more specific in \dependencies.
+Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, \dependencies be the DAG of constraint specificity, then the constraint $C$ equal to $t^S$ is called  \textit{the most specific} wrt \dependencies iff $C$ holds and there is no other template $t_{*}$ for which the constraint $C_{*}=t_{*}^S$ holds and $t_{*}$ is more specific in \dependencies.
 
@@
     & a constraint specificity DAG \dependencies \\
-    \textbf{Find:}& all the most specific wrt \dependencies solutions $S$ in \groups\\
-    & for each $c$ in \constraints \\
+    \textbf{Find:}& all the most specific wrt \dependencies constraints $C$ over \groups\\
+    & for each $t$ in \constraints \\
   \end{tabular}
@@
     \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- \template's, \dependencies -- dependency graph
-    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
-    \State $S \gets \emptyset$ \Comment{The set of solutions}
-  \For{$c \in \constrainttorder(\constraints,\dependencies)$}
-  \State $v_1, ..., v_n =$ variables of $c$
-    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(c, G, S, \dependencies)$}
-      \State $S \gets S \cup \findassignment(c, v_1{:}~G_1, \dots, v_n{:}~G_n, S, \dependencies)$
+    \State \textbf{Output:} $C$ -- learned constraints 
+    \State $C \gets \emptyset$ \Comment{The set of constraints}
+  \For{$t \in \constrainttorder(\constraints,\dependencies)$}
+  \State $v_1, ..., v_n =$ variables of $T$
+    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(t, \groups, C, \dependencies)$}
+      \State $C \gets C \cup \findassignment(t, v_1{:}~G_1, \dots, v_n{:}~G_n, C, \dependencies)$
     \EndFor
   \EndFor\\
-\Return $S$
+\Return $C$
 \end{algorithmic}
@@
 Solid edges point from more general to more specific constraints.
-Therefore, if there is a solid edge from constraint~$C_1$ to constraint~$C_2$, then a set of subgroups satisfying~$C_2$ must contain the subgroups satisfying~$C_1$.
+Therefore, if there is a solid edge from constraint~$t_1$ to constraint~$t_2$, then a set of subgroups satisfying~$t_2$ must contain the subgroups satisfying~$t_1$.
 
 Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
-If there is a dashed edge from $C_1$ to $C_2$, then a set of subgroups satisfying $C_2$ may not contain a subset that also satisfies $C_1$.
+If there is a dashed edge from $t_1$ to $t_2$, then a set of subgroups satisfying $t_2$ may not contain a subset that also satisfies $t_1$.
 
@@
 
-Finding candidate groups for a \template~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
+Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $t$.
 Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.

### 2016-04-15 Paper/ECAI/tabular.tex
fixing notation, formalization

@@
 \paragraph{Example}
-\sergey{need to use this below:}
-  For example, a \template \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
-
-%The constraint \textit{rank} has
-%\begin{itemize}
-%\item \CName = \textit{$v_r$ = RANK($v_x$)}, where $v_r$ and $v_x$ are %the variables;
-%\item \CSignature = subgroup $g_X$ (associated with X) is numeric, $g_Y$ %is integer and the length of vectors in $g_X$ is the same as in $g_Y$; and
-%\item
-%\CFunction = \textit{$g_Y$ = RANK($g_X$)} iff $g_X \in G_X, g_Y \in G_Y$ %and each value in $g_X$ has the rank (possibly with ties) specified in $%g_Y$.
-%\end{itemize}
+  For example, the \template \textit{rank} has
+\begin{itemize}
+\item \CName: \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables;
+\item \CSignature: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$;
+\item \CFunction: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
+\end{itemize}
 
@@
 We refer to $T^S$ as a \textit{constraint} and to the instantiated \CSignature (\CFunction) as $\text{\CSignature}^S$ ($\text{\CFunction}^S$).
+Throughout the work we refer as \textit{find constraints} to finding all constraints that hold.
 
@@
 $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
-Every \template~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.
+Every \template~$t$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$t$.
 

### 2016-04-15 Paper/ECAI/tabular.tex
Completed table

--- 2016-04-15 Paper/ECAI/tabular.tex
@@
 \newcommand{\constrainttorder}{\format{generalityOrder}}
-\newcommand{\template}{\format{Constraint template}}
 
@@
   \item
- A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$.
-If the subgroup $g$ must contain only one vector, we write $g \in G$.it defines what it means for a constraint to be well-formed
+ A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$. If the subgroup $g$ must contain only one vector, we write $g \in G$.
 %We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
@@
 \subsection{Constraints}
-A \template is a triple \textit{(\CName, \CSignature, \CFunction)}.
+\sergey{write correspondence with ILP terminology}
+A \textit{Spreadsheet constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}.
 Let us elaborate on each of them. 
@@
   For example, the signature can restrict the types of the groups, e.g., requiring them to be integers or constrain their sizes, e.g. the length of vectors in the arguments must be equal.
-  The constraints act on properties or group meta-information, not on the actual group content. In ILP, it is known as bias, it defines what it means for a constraint be well-formed.
-  \item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function. In ILP, it is known as background knowledge, which provides a definition to a predicate.
+  The constraints act on properties or group meta-information, not on the actual group content.
+In ILP, it is known as mode declarations, when one needs to declare types of variables in the vocabulary.
+Also, it is analogous to standard linear algebra well-defined requirements, e.g., in order to multiply matrices $A$ and $B$, $A$ must have the size of $n \times k$ and $B$ the size $k \times m$.
+\item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function.
+In ILP, it is known as bias, where we constrain ourselves to only particular subset of language in which we perform the search.
 \end{itemize}
@@
 \sergey{need to use this below:}
-  For example, a \template \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
+  For example, a spreadsheet constraint \textit{rank} has \CName \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables; its \CSignature is: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$; its \CFunction is the following constraint: the mapping $\{ v_x \mapsto G_x', v_y \mapsto G_y' \}$ is a solution iff $G_x' \subseteq G_x, G_y' \subseteq G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $G_y'$ is the rank of the corresponding value in $G_x'$ (possibly with ties).
 
@@
 \item
-a set of \template ${\cal C}$;
+a set of spreadsheet constraints ${\cal C}$;
 \item
@@
 \noindent
-\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a \template}
+\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a spreadsheet constraint}
 {\bf Find}  the set $C\subseteq {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
@@
 
-Let us now define what is a tabular solution for a \template in general.
-Let $C$ denote a \template, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
+Let us now define what is a tabular solution for a spreadsheet constraint in general.
+Let $C$ denote a spreadsheet constraint, $S$ a mapping from variables to the subgroups of \groups, and $C^S$ the constraint $C$, in which all variables are substituted with the corresponding subgroups from $S$, then $S$ is a \textit{tabular solution} to $C$ iff both $\CSignature^S$ and $\CFunction^S$ hold.
 
@@
 In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
-Here we formalize the statement in terms of \template and group assignments as follows:
+Here we formalize the statement in terms of spreadsheet constraints and group assignments as follows:
 
@@
     &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
+    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$\\
     \textbf{Find:}&  all tabular solution $S$ in \groups for each constraint $c$ in \constraints \\
@@
     &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
+    \textbf{Given:}& the set of all groups $\groups$ and of spreadsheet constraints $\constraints$,\\
     & a constraint specificity DAG \dependencies \\
@@
     \footnotesize
-    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- \template's, \dependencies -- dependency graph
+    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- Spreadsheet constraints, \dependencies -- dependency graph
     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignments
@@
 
-Finding candidate groups for a \template~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
+Finding candidate groups for a spreadsheet constraint~$c$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $c$.
 Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.
@@
 $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
-Every \template~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.
+Every spreadsheet constraint~$C$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$C$.
 
@@
 
-\subsection{Implementation note} Essentially for each \template both \generategroups and \findassignment are separate constraint satisfaction problems.
+\subsection{Implementation note} Essentially for each spreadsheet constraint both \generategroups and \findassignment are separate constraint satisfaction problems.
 Hence, it is natural to solve them using declarative solvers.
@@
 \newcommand{\integer}{\format{integer}}
+\newcommand{\discrete}{\format{discrete}}
 \newcommand{\plength}{\format{length}}
@@
 \newcommand{\por}{\format{orientation}}
+\newcommand{\prows}{\format{rows}}
+\newcommand{\pcols}{\format{columns}}
 \newcommand{\nat}{\mathcal{N}}
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
-    \ecalldiff{v_x} & $\integer(v_x)$ or $\textual(v_x)$ & All elements in $v_x$ are different \\
+    \ecalldiff{v_x} & $\discrete(v_x)$ & All values in $v_x$ are different: $v_x[i] \neq v_x[j]$ if $i \neq j$ \\
     %X = Y & & \\
-    \ecfkey{v_{fk}}{v_{pk}} & $\ptype(v_{fk})$ = $ptype(v_{pk})$,$\ptable(v_{fk}) \neq \ptable(v_{pk})$, $\ecalldiff{v_{pk}}$ & Every value in @fk also exist in @pk \\
+    \ecperm{v_x}
+      & $\numeric(v_{x})$, $\ecalldiff{v_{x}}$
+      & The values in $v_{x}$ are a permutation of the numbers $1$ through $\plength(v_{x})$.
+      \\
+    \ecseries{v_x}
+      & $\numeric(v_{x})$ and $\ecperm{v_{x}}$
+      & $v_{x}[1] = 1$ and $v_{x}[i] = v_{x}[i - 1] + 1$.
+      \\
+    \ecfkey{v_{fk}}{v_{pk}} & $v_{fk}$ and~$v_{pk}$ are both $\discrete$; $\ptype(v_{fk}) = \ptype(v_{pk})$; $\ptable(v_{fk}) \neq \ptable(v_{pk})$; and $\ecalldiff{v_{pk}}$ & Every value in~$v_{fk}$ also exist in~$v_{pk}$ \\
     % TODO check out
-    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
-      & the variables in $\{@r, @1, @fk\}$ and in $\{@pk, @val\}$ all have the same \plength, \ptable and \por.
-      & $@r$ is the obtained by multiplying $@1$ with $\eclookup{@fk}{@pk}{@val}$.
-      \\
     \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
-      & the variables in $\{@fk, @r\}$ and in $\{@pk, @val\}$ all have the same \plength, \ptable and \por; $@r$ and~$@val$ have the same type; and \ecfkey{@fk}{@pk}.
-      & Lookup $@fk$ in $@pk$ and return the corresponding value in $@val$.
+      & $v_{fk}$ and $v_{pk}$ are both $\discrete$; variables $\{v_{fk}, v_{r}\}$ and $\{v_{pk}, v_{val}\}$ within the same set have the same \plength, \ptable and \por; $v_{r}$ and~$v_{val}$ have the same type; and \ecfkey{v_{fk}}{v_{pk}}.
+      & $v_r[i]$ is the same value as looking up~$v_{fk}[i]$ in~$v_{pk}$  and returning the corresponding value in~$v_{val}$.
       \\
@@
       & \textit{Same as lookup}
-      & Lookup the first item in $@pk$ smaller than $@fk$ and return the corresponding value in $@val$.
+      & $v_r[i]$ is the same value as looking up the last item in~$v_{pk}$ smaller than~$v_{fk}[i]$ and returning the corresponding value in~$v_{val}$.
       \\
-    \ecperm{v_x}
-      & $\numeric(@x)$, $\ecalldiff{@x}$
-      & The values in $@x$ are a permutation of the numbers $1$ through $\plength(@x)$.
+    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
+      & Variables $\{v_{r}, v_{1}, v_{fk}\}$ are $\numeric$, variables $\{v_{pk}, v_{val}\}$ are $\discrete$ and within both sets all variables have the same \plength, \ptable and \por; also \ecfkey{v_{fk}}{v_{pk}}.
+      & $v_{r}[i]$ is the obtained by multiplying $v_{1}[i]$ with $\eclookupf{v_{fk}}{v_{pk}}{v_{val}}{}[i]$.
       \\
     \ecprod{v_r}{v_1}{v_2}
-      & variables $\{@r, @1, @2\}$ are all $\numeric$ and have the same $\plength$
-      & Every element in $@r$ is the product of the corresponding elements in~$@1$ and~$@2$.
+      & Variables $\{v_{r}, v_{1}, v_{2}\}$ are all $\numeric$ and have the same $\plength$
+      & $v_{r}[i] = v_{1}[i] \times v_{2}[i]$.
       \\
     \ecproj{v_r}{\mathbf{v_x}}
-      & $\{@r, @x\}$ all have the same $\plength$, $\por$, $\ptable$ and type; @x contains at least 2~vectors
-      & At every position in $1$ through $\plength(@x)$ there is exactly one vector in $@x$ with a non-blank value at that position which is the same as the corresponding value in $@r$.
+      & Variables $\{v_{r}, \mathbf{v_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{v_x}$ contains at least~2 vectors
+      & At every position~$i$ in $1$ through $\plength(v_{r})$ there is exactly one vector~$v$ in $\mathbf{v_x}$ such that $v[i]$ is a non-blank value, then $v[i] = v_{r}[i]$.
       \\
     \ecrank{v_r}{v_x}
-      & $\integer(@r)$; $\numeric(@x)$; and $\plength(@r) = \plength(@x)$
-      & The values in $@r$ represent the rank (from largest to smallest) of the values in ($@x$) - including ties
+      & $\integer(v_{r})$; $\numeric(v_{x})$; and $\plength(v_{r}) = \plength(v_{x})$
+      & The values in $v_{r}$ represent the rank (from largest to smallest) of the values in $v_{x}$ (including ties)
       \\
     \ectotal{v_r}{v_{pos}}{v_{neg}}
-      & $\{@r, @pos, @neg\}$ all have the same $\plength$, which is at least $2$
-      & The values in $@r$ are a running total, each value $@r[i] = @r[i - 1] + @pos[i] - @neg[i]$.
-      \\
-    \ecseries{v_x}
-      & $\numeric(@x)$ and $\ecperm{@x}$
-      & $@x[1] = 1$ and $@x[i] = @x[i - 1] + 1$.
+      & Variables $\{v_{r}, v_{pos}, v_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+      & The values in $v_{r}$ are a running total, each value $v_{r}[i] = v_{r}[i - 1] + v_{pos}[i] - v_{neg}[i]$.
       \\
     \ecsumc{v_r}{\mathbf{v_x}}
-      &
-      & Y is the result of summing each column in X \\
-    \ecsumr{v_r}{\mathbf{v_x}} & & Y is the result of summing each row in X \\
-    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
-    \ecsumprod{v_r}{v_1}{v_2} & & \\
+      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\prows(\mathbf{v_x}) \geq 2$; and $\pcols(\mathbf{v_x}) = \plength(v_r)$
+      & Each value in $v_{r}$ is obtained by summing over the corresponding column in $\mathbf{v_x}$.
+      \\
+    \ecsumr{v_r}{\mathbf{v_x}}
+      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\pcols(\mathbf{v_x}) \geq 2$; and $\prows(\mathbf{v_x}) = \plength(v_r)$
+      & Each value in $v_{r}$ is obtained by summing over the corresponding row in $\mathbf{v_x}$.
+      \\
+    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}
+      & $v_{fk}, v_{pk}$ are $\discrete$; $v_{r}, v_{val}$ are $\numeric$; within the sets $\{v_{val}, v_{fk}\}$ and $\{v_{pk}, v_{r}\}$ variables have the same $\plength$ and $\por$; $v_{fk}$ and $v_{val}$ have the same $\ptable$; $v_{fk}$ and $v_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{v_{pk}}
+      & The value for $v_{r}[i]$ is obtained by summing all values $v_{val}[j]$ where $v_{fk}[j] = v_{pk}[i]$
+      \\
+    \ecsumprod{v_r}{v_1}{v_2}
+      & Variables $\{v_r, v_1, v_2\}$ are all $\numeric$; $\plength(v_{1}) = \plength(v_{2}) \geq 2$; and $\prows(v_{r}) = \pcols(v_{r}) = 1$
+      & $v_{r}[i] = \sum_{i = 1}^{\plength(v_{1})} v_{1}[i] \times v_{2}[i]$.
+      \\
 
   \end{tabularx}
-  \caption{\template's \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
+  \caption{Spreadsheet Constraints \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
 etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}
@@
 \end{table*}
+
+\samuel{Satisfy signature for group assignment: find groups such that there exists a subgroup which satisfies the signature}
 

### 2016-04-15 Paper/ECAI/tabular.tex

--- 2016-04-15 Paper/ECAI/tabular.tex
@@
       & \textit{Same as lookup}
-      & Lookup the first item in $@pk$ smaller than $@fk$ and return the corresponding value in $@val$.
+      & $v_r[i]$ is the same value as looking up the last item in~$v_{pk}$ smaller than~$v_{fk}[i]$ and returning the corresponding value in~$v_{val}$.
       \\
-    \ecperm{v_x}
-      & $\numeric(@x)$, $\ecalldiff{@x}$
-      & The values in $@x$ are a permutation of the numbers $1$ through $\plength(@x)$.
+    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
+      & Variables $\{v_{r}, v_{1}, v_{fk}\}$ are $\numeric$, variables $\{v_{pk}, v_{val}\}$ are $\discrete$ and within both sets all variables have the same \plength, \ptable and \por; also \ecfkey{v_{fk}}{v_{pk}}.
+      & $v_{r}[i]$ is the obtained by multiplying $v_{1}[i]$ with $\eclookupf{v_{fk}}{v_{pk}}{v_{val}}{}[i]$.
       \\
     \ecprod{v_r}{v_1}{v_2}
-      & variables $\{@r, @1, @2\}$ are all $\numeric$ and have the same $\plength$
-      & Every element in $@r$ is the product of the corresponding elements in~$@1$ and~$@2$.
+      & Variables $\{v_{r}, v_{1}, v_{2}\}$ are all $\numeric$ and have the same $\plength$
+      & $v_{r}[i] = v_{1}[i] \times v_{2}[i]$.
       \\
     \ecproj{v_r}{\mathbf{v_x}}
-      & $\{@r, @x\}$ all have the same $\plength$, $\por$, $\ptable$ and type; @x contains at least 2~vectors
-      & At every position in $1$ through $\plength(@x)$ there is exactly one vector in $@x$ with a non-blank value at that position which is the same as the corresponding value in $@r$.
+      & Variables $\{v_{r}, \mathbf{v_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{v_x}$ contains at least~2 vectors
+      & At every position~$i$ in $1$ through $\plength(v_{r})$ there is exactly one vector~$v$ in $\mathbf{v_x}$ such that $v[i]$ is a non-blank value, then $v[i] = v_{r}[i]$.
       \\
     \ecrank{v_r}{v_x}
-      & $\integer(@r)$; $\numeric(@x)$; and $\plength(@r) = \plength(@x)$
-      & The values in $@r$ represent the rank (from largest to smallest) of the values in ($@x$) - including ties
+      & $\integer(v_{r})$; $\numeric(v_{x})$; and $\plength(v_{r}) = \plength(v_{x})$
+      & The values in $v_{r}$ represent the rank (from largest to smallest) of the values in $v_{x}$ (including ties)
       \\
     \ectotal{v_r}{v_{pos}}{v_{neg}}
-      & $\{@r, @pos, @neg\}$ all have the same $\plength$, which is at least $2$
-      & The values in $@r$ are a running total, each value $@r[i] = @r[i - 1] + @pos[i] - @neg[i]$.
-      \\
-    \ecseries{v_x}
-      & $\numeric(@x)$ and $\ecperm{@x}$
-      & $@x[1] = 1$ and $@x[i] = @x[i - 1] + 1$.
+      & Variables $\{v_{r}, v_{pos}, v_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+      & The values in $v_{r}$ are a running total, each value $v_{r}[i] = v_{r}[i - 1] + v_{pos}[i] - v_{neg}[i]$.
       \\
     \ecsumc{v_r}{\mathbf{v_x}}
-      &
-      & Y is the result of summing each column in X \\
-    \ecsumr{v_r}{\mathbf{v_x}} & & Y is the result of summing each row in X \\
-    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}} & & \\
-    \ecsumprod{v_r}{v_1}{v_2} & & \\
+      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\prows(\mathbf{v_x}) \geq 2$; and $\pcols(\mathbf{v_x}) = \plength(v_r)$
+      & Each value in $v_{r}$ is obtained by summing over the corresponding column in $\mathbf{v_x}$.
+      \\
+    \ecsumr{v_r}{\mathbf{v_x}}
+      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\pcols(\mathbf{v_x}) \geq 2$; and $\prows(\mathbf{v_x}) = \plength(v_r)$
+      & Each value in $v_{r}$ is obtained by summing over the corresponding row in $\mathbf{v_x}$.
+      \\
+    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}
+      & $v_{fk}, v_{pk}$ are $\discrete$; $v_{r}, v_{val}$ are $\numeric$; within the sets $\{v_{val}, v_{fk}\}$ and $\{v_{pk}, v_{r}\}$ variables have the same $\plength$ and $\por$; $v_{fk}$ and $v_{val}$ have the same $\ptable$; $v_{fk}$ and $v_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{v_{pk}}
+      & The value for $v_{r}[i]$ is obtained by summing all values $v_{val}[j]$ where $v_{fk}[j] = v_{pk}[i]$
+      \\
+    \ecsumprod{v_r}{v_1}{v_2}
+      & Variables $\{v_r, v_1, v_2\}$ are all $\numeric$; $\plength(v_{1}) = \plength(v_{2}) \geq 2$; and $\prows(v_{r}) = \pcols(v_{r}) = 1$
+      & $v_{r}[i] = \sum_{i = 1}^{\plength(v_{1})} v_{1}[i] \times v_{2}[i]$.
+      \\
 

### 2016-04-15 Paper/ECAI/tabular.tex
moved from formalization to statement section

@@
 \item \CSignature: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$;
-\item \CFunction: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ is a solution iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
+\item \CFunction: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
 \end{itemize}
 
-\samuel{Should this section maybe be moved to problem statement?}
+\section{Problem Statement}
+\sergey{moved from the formalization BEGIN}
 We can now define a classical constraint satisfaction problem on tables.
@@
 Let us introduce \template instantiation syntax.
-Let $S$ be a mapping from variables to the subgroups of \groups and $T$ be a \template, then $T^S$ is the template, in which all variables in all constrainst in \CSignature and \CFunction are instantiated with the corresponding subgroups from $S$.
-We refer to $T^S$ as a \textit{constraint} and to the instantiated \CSignature (\CFunction) as $\text{\CSignature}^S$ ($\text{\CFunction}^S$).
+Let $S$ be a mapping from variables to the subgroups of \groups and $t$ be a \template, then $t^S$ is the template, in which all variables in all constrainst in \CSignature and \CFunction are instantiated with the corresponding subgroups from $S$.
+We refer to $t^S$ as a \textit{constraint} and to the instantiated \CSignature (\CFunction) as $\text{\CSignature}^S$ ($\text{\CFunction}^S$).
 Throughout the work we refer as \textit{find constraints} to finding all constraints that hold.
@@
 Let us now define what it means for a constraint from \template to hold in general.
-Let $T$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $T^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
+Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
+\sergey{moved from formalization END}
 
-\section{Problem Statement}
-\sergey{should reflect all the changes, Luc wants to introduce constraint satisfaction into problem statement, maybe we should move the definition of a solution here?
-Like a solution is actually constraint satisfaction?}
 In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.

### 2016-04-15 Paper/ECAI/tabular.tex
lines

@@
   \begin{tabularx}{\textwidth}{l X X}
-    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
-    \ecalldiff{v_x} & $\discrete(v_x)$ & All values in $v_x$ are different: $v_x[i] \neq v_x[j]$ if $i \neq j$ \\
+    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
+    \ecalldiff{v_x}
+      & $\discrete(v_x)$
+      & All values in $v_x$ are different: $v_x[i] \neq v_x[j]$ if $i \neq j$
+      \\ \hline
     %X = Y & & \\
@@
       & The values in $v_{x}$ are a permutation of the numbers $1$ through $\plength(v_{x})$.
-      \\
+      \\ \hline
     \ecseries{v_x}
@@
       & $v_{x}[1] = 1$ and $v_{x}[i] = v_{x}[i - 1] + 1$.
-      \\
-    \ecfkey{v_{fk}}{v_{pk}} & $v_{fk}$ and~$v_{pk}$ are both $\discrete$; $\ptype(v_{fk}) = \ptype(v_{pk})$; $\ptable(v_{fk}) \neq \ptable(v_{pk})$; and $\ecalldiff{v_{pk}}$ & Every value in~$v_{fk}$ also exist in~$v_{pk}$ \\
+      \\ \hline
+    \ecfkey{v_{fk}}{v_{pk}} & $v_{fk}$ and~$v_{pk}$ are both $\discrete$; $\ptype(v_{fk}) = \ptype(v_{pk})$; $\ptable(v_{fk}) \neq \ptable(v_{pk})$; and $\ecalldiff{v_{pk}}$ & Every value in~$v_{fk}$ also exist in~$v_{pk}$ \\ \hline
     % TODO check out
@@
       & $v_r[i]$ is the same value as looking up~$v_{fk}[i]$ in~$v_{pk}$  and returning the corresponding value in~$v_{val}$.
-      \\
+      \\ \hline
     \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}}
@@
       & $v_r[i]$ is the same value as looking up the last item in~$v_{pk}$ smaller than~$v_{fk}[i]$ and returning the corresponding value in~$v_{val}$.
-      \\
+      \\ \hline
     \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
@@
       & $v_{r}[i]$ is the obtained by multiplying $v_{1}[i]$ with $\eclookupf{v_{fk}}{v_{pk}}{v_{val}}{}[i]$.
-      \\
+      \\ \hline
     \ecprod{v_r}{v_1}{v_2}
@@
       & $v_{r}[i] = v_{1}[i] \times v_{2}[i]$.
-      \\
+      \\ \hline
     \ecproj{v_r}{\mathbf{v_x}}
@@
       & At every position~$i$ in $1$ through $\plength(v_{r})$ there is exactly one vector~$v$ in $\mathbf{v_x}$ such that $v[i]$ is a non-blank value, then $v[i] = v_{r}[i]$.
-      \\
+      \\ \hline
     \ecrank{v_r}{v_x}
@@
       & The values in $v_{r}$ represent the rank (from largest to smallest) of the values in $v_{x}$ (including ties)
-      \\
+      \\ \hline
     \ectotal{v_r}{v_{pos}}{v_{neg}}
@@
       & The values in $v_{r}$ are a running total, each value $v_{r}[i] = v_{r}[i - 1] + v_{pos}[i] - v_{neg}[i]$.
-      \\
+      \\ \hline
     \ecsumc{v_r}{\mathbf{v_x}}
@@
       & Each value in $v_{r}$ is obtained by summing over the corresponding column in $\mathbf{v_x}$.
-      \\
+      \\ \hline
     \ecsumr{v_r}{\mathbf{v_x}}
@@
       & Each value in $v_{r}$ is obtained by summing over the corresponding row in $\mathbf{v_x}$.
-      \\
+      \\ \hline
     \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}
@@
       & The value for $v_{r}[i]$ is obtained by summing all values $v_{val}[j]$ where $v_{fk}[j] = v_{pk}[i]$
-      \\
+      \\ \hline
     \ecsumprod{v_r}{v_1}{v_2}

### 2016-04-15 Paper/ECAI/tabular.tex
comments and typos

--- 2016-04-15 Paper/ECAI/tabular.tex
@@
  A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$.
-If the subgroup $g$ must contain only one vector, we write $g \in G$.it defines what it means for a constraint to be well-formed
+If the subgroup $g$ must contain only one vector, we write $g \in G$.
+It defines what it means for a constraint to be well-formed
 %We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
@@
   \caption{\template's \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
-etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display}
+etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display} \sergey{1) should be readable alone without lookin in the text of the paper 2) that's the constraints we found popular according the spreadsheets in the web 3) that's what is actually implemented in the system 4) corresponds to the actual Excel functions}
 } 

### 2016-04-15 Paper/ECAI/tabular.tex
update the figure and comments

@@
 \sergey{Future version, extensions, declarative stuff, bla-bla here}
+discussion: what would be a language to extend the system, what would be the user interface? can it be used by community?  can it be complitely automated? interaction with user and highlighing?
+
+conclusions 
+points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
 

### 2016-04-15 Paper/ECAI/tabular.tex
Added comment

@@
 \subsection{Groups, Type-consistency and Constraints}
+\samuel{Instantiated table... references}
 In this work we focus on spreadsheet data, which is physically one big table, but conceptually often consists of multiple tables, see Figure~\ref{fig:main_example} for an example.

### 2016-04-15 Paper/ECAI/tabular.tex
My changes (approach) + luc

@@
 \newcommand{\constrainttorder}{\format{generalityOrder}}
-\newcommand{\template}{\format{Constraint template}}
+\newcommand{\template}{\format{constraint template}}
 
-\newcommand{\CName}{Name\xspace}
+\newcommand{\CName}{Syntax\xspace}
 \newcommand{\CSignature}{Signature\xspace}
-\newcommand{\CFunction}{Function\xspace}
+\newcommand{\CFunction}{Definition\xspace}
 \newcommand{\dependencies}{\ensuremath{\mathcal{D}}\xspace}
@@
 \newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
-\newcommand{\eclookupf}[4]{\ensuremath{\mathit{LOOKUP}_{#4}(#1, #2, #3)}}
+\newcommand{\eclookupf}[4]{\ensuremath{\mathit{LOOKUP}_{\mathit{#4}}(#1, #2, #3)}}
 \newcommand{\eclookup}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{}}}
@@
 
+\newcommand{\numeric}{\format{numeric}}
+\newcommand{\textual}{\format{textual}}
+\newcommand{\integer}{\format{integer}}
+\newcommand{\discrete}{\format{discrete}}
+\newcommand{\plength}{\format{length}}
+\newcommand{\ptype}{\format{type}}
+\newcommand{\ptable}{\format{table}}
+\newcommand{\por}{\format{orientation}}
+\newcommand{\prows}{\format{rows}}
+\newcommand{\pcols}{\format{columns}}
+\newcommand{\nat}{\mathcal{N}}
+
 \newcommand{\luc}[1]{{\textcolor{red}{#1}}}
+
+\renewcommand{\arraystretch}{1.5}
 
@@
   \item
-  A \textit{vector} is either a column or a row that is type-consistent.
+  A \textbf{vector} is either a column or a row that is type-consistent.
   If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
   \item
-  A \textit{group} is a subrange of vectors with the same orientation in a table.
+  A \textbf{group} is a subrange of vectors with the same orientation in a table.
   We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
-  We denote as the \textit{length} of a row (column) group $G$, written as \textit{length(G)}, the number of its columns (rows).
-We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
+  The \textit{length} of a row (column) group $G$, written as \textit{length(G)}, is the number of its columns (rows); We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
   \item
@@
 If the subgroup $g$ must contain only one vector, we write $g \in G$.
-It defines what it means for a constraint to be well-formed
+%It defines what it means for a constraint to be well-formed
 %We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
@@
 \end{itemize}
-For notational convenience, we will refer to vectors with $V$ and to groups and subgroups with $G$.
+%For notational convenience, we will refer to vectors in {\rm roman} and to groups in {\bf bold}. {\bf CHECK}
 
@@
 A \template is a triple \textit{(\CName, \CSignature, \CFunction)}.
-Let us elaborate on each of them. 
+Let us elaborate on this:
 \begin{itemize}
 \item
-\textit{\CName} is the textual name of the constraint together with the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
-\item
-  \textit{\CSignature} is a set of constraints on the properties of the group assignments, corresponding to the variables $v_1,\dots,v_n$ in \CName.
-  For example, the signature can restrict the types of the groups, e.g., requiring them to be integers or constrain their sizes, e.g. the length of vectors in the arguments must be equal.
-  The constraints act on properties or group meta-information, not on the actual group content. In ILP, it is known as bias, it defines what it means for a constraint be well-formed.
-  \item \textit{\CFunction} is a constraint specifying that the data in the subgroups satisfies the function. In ILP, it is known as background knowledge, which provides a definition to a predicate.
+\textit{\CName}  specifies the syntactic form of the constraint $c(A_1, ...,A_n)$, that is, the name of the constraint $c$ together
+with $n$ arguments $A_i$. Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
+% the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
+\item the \textit{\CSignature} constrains the arguments $A_i$ of the constraint, that is, the types of the arguments, together
+with possible further restrictions using properties of the arguments $A_i$ such as their length.  For example, the signature can restrict the types of the arguments
+to be vectors of integers and require that  the length of all arguments be equal.  In terms of logical and relational learning, this is known as the {\em bias} of the learner,
+it specifies when the constraint is well-formed.
+  %The constraints act on properties or group meta-information, not on the actual group content. In ILP, it is known as bias,
+  \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds, in practice this will be a function that can be called when
+the arguments are full instantiated to decide whether the constraint is satisfied or not.
+In logical and relational learning this is known as the background knowledge, which provides the definition of the predicate.
 \end{itemize}
@@
 \paragraph{Example}
-  For example, the \template \textit{rank} has
+Several example constraints are illustrated in Table 1.  For example, the \template \textit{rank} has
 \begin{itemize}
-\item \CName: \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are its variables;
-\item \CSignature: groups assigned to $v_x$ must be numeric, groups assigned to $v_y$ must be integer and the length of vectors in $v_x$ is the same as in $v_y$;
-\item \CFunction: the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
+\item \CName: \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are the arguments;
+\item \CSignature: the two arguments $v_r$ and $v_x$ must be numeric; furthermore, the first argument $v_r$ must be integer and the length of the vectors
+of $v_x$ and $v_$ be identical, i.e., $length(v_x) = length(v_r)$;
+\item \CFunction: \textit{$v_r$ = RANK($v_x$)} whenever each value of $v_r$ is the rank of the corresponding value in $v_x$.
+% is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
 \end{itemize}
 
-\section{Problem Statement}
-\sergey{moved from the formalization BEGIN}
-We can now define a classical constraint satisfaction problem on tables.
-A {\em constraint satisfaction problem} (CSP)  consists of 1) a set of cells, and 2) a set of constraints over those cells.
+When looking into constraints like \textit{$v_r$ = RANK($v_x$)}, it is important to see the analogy with both first order logic and with constraint satisfaction problems.
+From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $v_r$ and $v_x$ are the terms, which can be either variables or values.
+This also holds in our setting: when we write
+      $\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}$,
+we can interpret the argument ${\range{T_{1}}{\rangeall}{8}}$ as group variables that would apply to any table,
+or we can interpret the argument as the value it takes in the particular table called $T_1$ that is shown in Figure 1, that is, $\[1,2,3,4\]$. {\bf CHECK}
+This also means that we can talk about constraints being satisfied. Basically, a constraint $c(t_1, ..., t_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
+if the constraint holds on the values that the $t_i$ take in the set of tables ${\cal T}$. Figure 3 provides a list of constraints that are satisfied by the tables
+in Figure 1.
+
+%\sergey{moved from the formalization BEGIN}
+%l 1) a set of cells, and 2) a set of constraints over those cells.
 Notice that the constraints specify the types (or the domains) that one encounters in traditional CSPs.
@@
 
-While the traditional CSP problem is given a set of cells and a set of constraints,
+\section{Problem Statement}
+A {\em constraint satisfaction problem} (CSP) consists of a set of constraints and the
+task is to find assignments of values to the variables in the CSP so that all constraints are satisfied.  As the variables typically refer to
+cells in tables, solving a CSP problem often means completing the tables so that all constraints are satisfied.
+
+While the traditional CSP problem is given a set of variables and a set of constraints,
 and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
@@
 \noindent
-\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a \template}
-{\bf Find}  the set $C\subseteq {\cal C}$ that contains all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
-for which
-all $G_i \in {\cal G}$ and  3)  $c(G_1, ... , G_N)$ satisfies the signature of the constraint $c$.
+%\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a \template}
+{\bf Find}  the set of all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
+for which all $G_i \in {\cal G}$ and  3)  $c(G_1, ... , G_N)$ satisfies the signature of the constraint $c$.
 
+For instance, Figure 3 contains the solution to the tabular constraint learning problem
+that uses the tables of Figure 1, the signatures of Table 1 and the set of groups mentioned in {\bf CHECK / COMPLETE}
+
+\luc{Delete from here}
 \luc{ADD/REFER TO EXAMPLE}
@@
 \end{minipage}
+\luc{To here}
 
@@
 
+\newcommand{\tcl}{Tabular Constraint Learning}
+\newcommand{\ctl}{Condensed Tabular Constraint Learning}
 \section{Approach to Tabular Constraint Learning}
 \sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
-As we have seen in the statement before we assume the set of groups to be given.
-It is often, however, not the case and a possible approach is to extract the groups from the CSV file directly.
-In our system, we have implemented the group construction step using the type-consistency check.
-\sergey{Samuel, we need details here, I think} In general, separating data from a header is a task on its own  \cite{header} and goes beyond the scope of this paper.
+This section will describe our approach to \tcl.
+Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
+Often, however, this not the case and a possible approach is to extract the groups from the CSV file directly.
+This requires discovering \textit{tables} and \textit{groups}.
+Our implementation offers automated procedures for both steps.
 
-Another solution to the group extraction problem is to make it interactive, when a user can select the groups in the spreadsheet or correct automatically generated groups.
-We regard this approach to be a possible future direction in the system development.
-Later in the paper we assume the groups to be given.
+Tables are detected by detecting rectangular collections of data and attempting to detect and separate headers from the data.
+Especially this last step is a complex task on its own  \cite{header} and a sophisticated implementation goes beyond the scope of this paper.
+On well formated data, however, the user will not have to specify tables manually.
 
-Now we describe how the problem of Condensed Tabular Constraint Learning can be solved.
-The algorithm mimics the structure of the problem in the following way: first we need to order constraints in the order of specificity, then for each constraint we need to generate possible candidate groups, and in the next step we find the subgroups satisfying the constraint.
-After an iteration we accumulate all solutions and proceed to the next constraint.
+Our algorithm also attempts to detect groups within tables by partitioning the data into consecutive, type consistent groups, essentially separating textual and numerical data.
+We support some extensions such as currencies and percentual data and our approach allows us to omit manual specifications of groups.
+
+A more elaborate implementation could allow group (and table) detection could be done interactively / visually such that a user can select the groups in a spreadsheet or correct automatically generated groups.
+From here on we will assume that groups and tables are given.
+
+\newcommand{\temps}{\ensuremath{S}}
+
+\subsection{Algorithm}
+Our algorithm is illustrated in Algorithm~\ref{algo:tcl} and contains three key steps.
+First, constraint templates are ordered according to the generality ordering such that they can be treated sequentially.
+Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
+Every group assignment assigns a group~$G \in \groups$ to every argument of the template.
+In the third step, for every assignments $(G_1, ..., G_n) \in A$ we find all constraints $s(G_1', ...
+G_n')$ that hold on the data (satisfy the \CFunction), such that every subgroup $G_i' \subseteq G_i$.
 
@@
 \end{algorithm}
-Essentially, our Algorithm \ref{algo:tcl} has three steps: constraint ordering, candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}.
-Let us elaborate on each step in detail.
 
-\samuel{Running example (RANK?)}
+Our Algorithm \ref{algo:tcl} is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}, however, we apply the approach on two levels, generating and testing groups according to their properties and generating and testing subgroups on the actual data.
+We will explain and illustrate the three key steps in detail.
 
-\subsection{Constraint ordering} 
-Since some constraints are more general than the other, we generate them in the generality order similar to the ILP techniques known as refinement operators \cite{luc_book}.
-
-$\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} in the following way: all constraints are generated in the topological order of~\dependencies.
-Solid edges point from more general to more specific constraints.
-Therefore, if there is a solid edge from constraint~$t_1$ to constraint~$t_2$, then a set of subgroups satisfying~$t_2$ must contain the subgroups satisfying~$t_1$.
-
-Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
-If there is a dashed edge from $t_1$ to $t_2$, then a set of subgroups satisfying $t_2$ may not contain a subset that also satisfies $t_1$.
-
-Thus, \dependencies describes a partial order in which constraints should be learned.
-Some constraints have no dependencies (omitted from the figure) and can be learned independently.
-Generally, a parallel version of the algorithm could learn each connected component independently, only taking into account the partial ordering within components.
+\subsection{Generality ordering}
+A constraint~$c_1$ is a special case of another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
+We consider $c_1$ to be more specific than $c_2$ or $c_2$ to be more general than $c_1$.
+A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.
 
 \paragraph{Example}
-\samuel{Fix notation variable vs. group}
-Consider the \ecfkey{v_{fk}}{v_{pk}} (foreign-key) constraint: for every solution the subgroup (vector) $V$ assigned to $\mathit{v_{pk}}$ must fulfill \ecalldiff{V}.
-Therefore, we say indicate \textit{foreign-key} as being dependent on \textit{all-different} and the latter will be treated first.
-The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties the vector $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+Consider templates \textit{all-different} and \textit{foreign-key}.
+If \ecfkey{G_{pk}}{G_{fk}}, then \ecalldiff{G_{pk}} must also hold, therefore, \textit{all-different} is considered to be more general than \textit{foreign-key}.
+\\\\
+In ILP one often exploits these type of generality relations to structure the search for constraints using refinement operators \cite{luc_book}.
+Using such techniques leads to a more efficient traversal of the search space (of constraints).
+Aside from the efficiency gained by exploiting constraints learned earlier, the generality also allows us to build constraint templates on top of more general templates.
+
+$\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} to generate all constraint templates according to the topological order of~\dependencies.
+Constraint templates that have no dependencies are omitted.
+In the figure solid edges point from more general to more specific templates.
+Therefore, if there is a solid edge from constraint template~$s_1$ to template~$s_2$, then a set of subgroups satisfying~$s_2$ must contain the subgroups satisfying~$t_1$.
+
+% Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
+% If there is a dashed edge from $t_1$ to $t_2$, then a set of subgroups satisfying $t_2$ may not contain a subset that also satisfies $t_1$.
+
+\paragraph{Example}
+As mentioned above \textit{foreign-key} is more specific than \textit{all-different}.
+Therefore, we can first generate all subgroups $G_i'$ for which \ecalldiff{G_i'} is satisfied and use these subgroups as candidates for $v_{pk}$ in \ecfkey{v_{fk}}{v_{pk}}.
+
+The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+\\\\
+Most parts of the algorithm could be easily parallelized.
+However, \dependencies would impose a partial order in which constraint templates would have to be treated.
 
 \subsection{Candidate group generation}
-$\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that may contains subgroups satisfying the constraint.
+$\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups (group-assignments~$A$) that may contains subgroups satisfying the constraint.
 The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
 
-Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of $t$.
-Several constraints (e.g. sum, max, count) have the same candidate generation procedure as they share the same \CSignature.
+Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
+For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that those groups or some of their subgroups satisfy the \CSignature.
+
+For some constraints (e.g. sum, max, count) that share the same \CSignature the same set of assignments~$A$ will be found.
 
 \paragraph{Example}
-The \ecrank{v_y}{v_x} constraint requires $v_y$ to consist of only integer or only textual data and $v_x$ to be numeric.
-This allows us to filter candidate groups for these variables.
-Additionally, $v_y$ and~$v_x$ must be vectors of the same length, therefore, only groups with the same length are considered.
-Groups that do not satisfy these requirements cannot contain subgroups (vectors) that satisfy \ecrank{v_y}{v_x}.
+\ecrank{v_r}{v_x} will require groups assigned to $v_r$ to be $\integer$, groups assigned to $v_x$ to be $\discrete$ (i.e.
+$\integer$ or $\textual$) and only allows assignments in which groups assigned to both variables have the same length.
 
 \subsection{Subgroup satisfaction search}
-$\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function that searches for subsets of vectors in the candidate groups that satisfy the constraint.
-Every \template~$t$ needs to have an implementation that finds subgroup assignments, given set of tuples each containing a candidate group for every variable in~$t$.
+$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that $t(G_1', ..., G_n')$ is satisfied.
+For every \template~$t$ an implementation (declarative or imperative) need to be provided that finds valid subgroups given the assignments.
+ assignments, given set of tuples each containing a candidate group for every variable in~$t$.
 
@@
   \centering
-  \includegraphics[width=0.20\textwidth]{figures/constraint_dependency.png}
+  \includegraphics[width=0.9\linewidth]{figures/constraint_dependency.png}
   \caption{Constraint Learning Order.
@@
 
-\newcommand{\numeric}{\format{numeric}}
-\newcommand{\textual}{\format{textual}}
-\newcommand{\integer}{\format{integer}}
-\newcommand{\discrete}{\format{discrete}}
-\newcommand{\plength}{\format{length}}
-\newcommand{\ptype}{\format{type}}
-\newcommand{\ptable}{\format{table}}
-\newcommand{\por}{\format{orientation}}
-\newcommand{\prows}{\format{rows}}
-\newcommand{\pcols}{\format{columns}}
-\newcommand{\nat}{\mathcal{N}}
-
 \sergey{Table should take into account our new notation and balance between formality and intuition, heh}
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    \ecalldiff{v_x}
+    $\ecalldiff{v_x}^*$
       & $\discrete(v_x)$
@@
     %X = Y & & \\
-    \ecperm{v_x}
+    $\ecperm{v_x}^*$
       & $\numeric(v_{x})$, $\ecalldiff{v_{x}}$
@@
   \caption{\template's \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
-etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display} \sergey{1) should be readable alone without lookin in the text of the paper 2) that's the constraints we found popular according the spreadsheets in the web 3) that's what is actually implemented in the system 4) corresponds to the actual Excel functions}
-} 
+etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display} \sergey{1) should be readable alone without lookin in the text of the paper 2) that's the constraints we found popular according the spreadsheets in the web 3) that's what is actually implemented in the system 4) corresponds to the actual Excel functions} These constraints are \textit{structural}, they do not compute a result.
+Our algorithm can and does discover these constraints, however, they cannot be directly used in spreadsheet software }
+
   \label{table:constraints}

### 2016-04-15 Paper/ECAI/tabular.tex
Fixes

@@
 \item \CSignature: the two arguments $v_r$ and $v_x$ must be numeric; furthermore, the first argument $v_r$ must be integer and the length of the vectors
-of $v_x$ and $v_$ be identical, i.e., $length(v_x) = length(v_r)$;
+of $v_x$ and $v_r$ be identical, i.e., $length(v_x) = length(v_r)$;
 \item \CFunction: \textit{$v_r$ = RANK($v_x$)} whenever each value of $v_r$ is the rank of the corresponding value in $v_x$.
@@
 we can interpret the argument ${\range{T_{1}}{\rangeall}{8}}$ as group variables that would apply to any table,
-or we can interpret the argument as the value it takes in the particular table called $T_1$ that is shown in Figure 1, that is, $\[1,2,3,4\]$. {\bf CHECK}
+or we can interpret the argument as the value it takes in the particular table called $T_1$ that is shown in Figure 1, that is, $[1,2,3,4]$. {\bf CHECK}
 This also means that we can talk about constraints being satisfied. Basically, a constraint $c(t_1, ..., t_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only

### 2016-04-26 Paper/ECAI/tabular.tex
Cleanup and approach

@@
 Given these novel aspects, we conclude that understanding and learning constraints in tabular data is a new and challenging problem setting.
-Futhermore, it is not only challenging for a machine but for a human as well.
+Furthermore, it is not only challenging for a machine but for a human as well.
 
@@
 with $n$ arguments $A_i$. Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
+\samuel{Or a function}
 % the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
@@
 
-\luc{Delete from here}
-\luc{ADD/REFER TO EXAMPLE}
+% \luc{Delete from here}
+% \luc{ADD/REFER TO EXAMPLE}
 
-Let us introduce \template instantiation syntax.
+% Let us introduce \template instantiation syntax.
 Let $S$ be a mapping from variables to the subgroups of \groups and $t$ be a \template, then $t^S$ is the template, in which all variables in all constrainst in \CSignature and \CFunction are instantiated with the corresponding subgroups from $S$.
@@
 
-Let us now define what it means for a constraint from \template to hold in general.
+% Let us now define what it means for a constraint from \template to hold in general.
 Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
-\sergey{moved from formalization END}
+% \sergey{moved from formalization END}
 
-In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
+% In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
 Here we formalize the statement in terms of \template and group assignments as follows:
 
-\begin{minipage}[c]{14em}
-  \vspace{5pt}
-  \begin{tabular}{ll}
-    \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
-    \vspace{-4pt}
-    &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
-    \textbf{Find:}&  all constraints $C$ over \groups for each template $t$ in \constraints \\
-  \end{tabular}
-  \vspace{6pt}
-\end{minipage}
-\luc{To here}
+% \begin{minipage}[c]{14em}
+%   \vspace{5pt}
+%   \begin{tabular}{ll}
+%     \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
+%     \vspace{-4pt}
+%     &\\
+%     \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
+%     \textbf{Find:}&  all constraints $C$ over \groups for each template $t$ in \constraints \\
+%   \end{tabular}
+%   \vspace{6pt}
+% \end{minipage}
+% \luc{To here}
+
+\samuel{Not reworked yet from here ---}
 
@@
 
-\subsection{Generality ordering}
+\subsubsection{Generality ordering}
+
+\begin{figure}[tbh]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figures/constraint_dependency.png}
+  \caption{Constraint Learning Order. The solid arrows points from more general to more specific constraints templates (e.g. \textit{all-different} is more general than \textit{permutation}).}
+  \label{fig:learning_order}
+\end{figure}
+
 A constraint~$c_1$ is a special case of another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
@@
 
-\subsection{Candidate group generation}
+\subsubsection{Candidate group generation}
 $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups (group-assignments~$A$) that may contains subgroups satisfying the constraint.
@@
 \ecrank{v_r}{v_x} will require groups assigned to $v_r$ to be $\integer$, groups assigned to $v_x$ to be $\discrete$ (i.e.
-$\integer$ or $\textual$) and only allows assignments in which groups assigned to both variables have the same length.
+$\integer$ or $\textual$) and only allows assignments in which groups assigned to both arguments have the same length.
 
-\subsection{Subgroup satisfaction search}
-$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that $t(G_1', ..., G_n')$ is satisfied.
-For every \template~$t$ an implementation (declarative or imperative) need to be provided that finds valid subgroups given the assignments.
- assignments, given set of tuples each containing a candidate group for every variable in~$t$.
+\subsubsection{Subgroup satisfaction search}
+$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied.
+For every \template~$t$ an implementation (declarative or imperative) needs to be provided that finds valid subgroups given the assignments.
 
 \paragraph{Example}
-Consider a candidate assignment $G_x, G_y$ associated with the variables $v_x,v_y$ of \ecrank{v_y}{v_x}, then \findassignment selects single vectors $g_y \in G_y, g_x \in G_x$ such that \ecrank{g_y}{g_x} is true, i.e.
-it searches for a mapping $S$ where $v_x$ ($v_y$) must be mapped to a subgroup of $G_x$ ($G_y$).
+Consider a group assignment $G_x, G_y$ associated with the arguments $v_x,v_y$ of \ecrank{v_y}{v_x}, then \findassignment selects single vectors $V_y \in G_y, V_x \in G_x$ such that \ecrank{V_y}{V_x} is true.
 For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_x$ and $G_y$.
-Then \findassignment could select $g_x$ to be $\range{T_1}{\rangeall}{7}$ and $g_y$ to be $\range{T_1}{\rangeall}{8}$ which forms a valid solution.
+Then \findassignment could select $V_x$ to be $\range{T_1}{\rangeall}{7}$ and $V_y$ to be $\range{T_1}{\rangeall}{8}$ since \ecrank{\range{T_1}{\rangeall}{7}}{\range{T_1}{\rangeall}{8}} is satisfied in the data.
 \\\\
-For some constraints (most notably aggregates) there may be multiple overlapping subsets of a given group~$G$ that satisfy the constraint.
-We chose to omit subsets that are fully contained within a larger subset since this can lead to overfitting.
-If the returned subsets are too large and accidentally contain unrelated vectors, the user can split such a group, disallowing.
+For some constraint templates (most notably aggregates) there may be multiple (overlapping) subgroups of a given group~$G$ for which the template satisfied.
+In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
+If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.
 
-\subsection{Implementation note} Essentially for each \template both \generategroups and \findassignment are separate constraint satisfaction problems.
+\subsection{Implementation note}
+For each \template, both \generategroups and \findassignment are can be seen as separate constraint satisfaction problems.
 Hence, it is natural to solve them using declarative solvers.
-Even though each is a separate constraint problem, they are variations of each other and can be modeled using similar techniques.
-It also makes an extension of the system easy, we can solve new variations by adding, removing or modifying constraints from a similar constraint satisfaction problem.
-Furthermore, declarative languages such as ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
+Even though each of these problems differ from each other, they are variations that can be modeled using similar techniques.
+This also makes it easy to extend the implementation as templates can be easily added (or removed).
+Furthermore, declarative languages such as, for example, ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
 
-Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and a native Python CSP solver \cite{python_constraint}.
+Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and uses a native Python CSP solver \cite{python_constraint} for \generategroups.
 All the constraints mentioned in the paper can be run purely using the internal Python engine without any external software.
-Future versions of the system will provide user interfaces to the ASP, Minizinc and Python-Constraint languages to allow users to add new constraints in a declarative way.
+Future versions of the system will provide user interfaces to some of these constraint languages (e.g.
+ASP and Minizinc) to allow users to specify solving procedures for new constraint templates in a declarative way.
 
@@
 
-\begin{figure}[htb]
-  \centering
-  \includegraphics[width=0.9\linewidth]{figures/constraint_dependency.png}
-  \caption{Constraint Learning Order.
-The solid arrows indicate a specificity order, i.e., \textit{permutation} $\rightarrow$ \textit{series} means that the series constraint is more specific than permutation.
-The dashed arrows indicate an inverse construction order, \textit{project} $\dashrightarrow$ \textit{aggregates} means the project constraint is more specific than aggregates.
-\sergey{Samuel,  can you give an explanation why we keep both arrow and how it is internally motivated?}}
-  \label{fig:learning_order}
-\end{figure}
-
-\sergey{Table should take into account our new notation and balance between formality and intuition, heh}
 \begin{table*}
@@
     \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
-      & $v_{fk}$ and $v_{pk}$ are both $\discrete$; variables $\{v_{fk}, v_{r}\}$ and $\{v_{pk}, v_{val}\}$ within the same set have the same \plength, \ptable and \por; $v_{r}$ and~$v_{val}$ have the same type; and \ecfkey{v_{fk}}{v_{pk}}.
+      & $v_{fk}$ and $v_{pk}$ are both $\discrete$; arguments $\{v_{fk}, v_{r}\}$ and $\{v_{pk}, v_{val}\}$ within the same set have the same \plength, \ptable and \por; $v_{r}$ and~$v_{val}$ have the same type; and \ecfkey{v_{fk}}{v_{pk}}.
       & $v_r[i]$ is the same value as looking up~$v_{fk}[i]$ in~$v_{pk}$  and returning the corresponding value in~$v_{val}$.
@@
     \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
-      & Variables $\{v_{r}, v_{1}, v_{fk}\}$ are $\numeric$, variables $\{v_{pk}, v_{val}\}$ are $\discrete$ and within both sets all variables have the same \plength, \ptable and \por; also \ecfkey{v_{fk}}{v_{pk}}.
+      & Arguments $\{v_{r}, v_{1}, v_{fk}\}$ are $\numeric$, arguments $\{v_{pk}, v_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{v_{fk}}{v_{pk}}.
       & $v_{r}[i]$ is the obtained by multiplying $v_{1}[i]$ with $\eclookupf{v_{fk}}{v_{pk}}{v_{val}}{}[i]$.
@@
     \ecprod{v_r}{v_1}{v_2}
-      & Variables $\{v_{r}, v_{1}, v_{2}\}$ are all $\numeric$ and have the same $\plength$
+      & Arguments $\{v_{r}, v_{1}, v_{2}\}$ are all $\numeric$ and have the same $\plength$
       & $v_{r}[i] = v_{1}[i] \times v_{2}[i]$.
@@
     \ecproj{v_r}{\mathbf{v_x}}
-      & Variables $\{v_{r}, \mathbf{v_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{v_x}$ contains at least~2 vectors
+      & Arguments $\{v_{r}, \mathbf{v_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{v_x}$ contains at least~2 vectors; and $v_r = \mathit{SUM}(\mathbf{v_x}, \por(\mathbf{v_x}))$
       & At every position~$i$ in $1$ through $\plength(v_{r})$ there is exactly one vector~$v$ in $\mathbf{v_x}$ such that $v[i]$ is a non-blank value, then $v[i] = v_{r}[i]$.
@@
     \ectotal{v_r}{v_{pos}}{v_{neg}}
-      & Variables $\{v_{r}, v_{pos}, v_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+      & Arguments $\{v_{r}, v_{pos}, v_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
       & The values in $v_{r}$ are a running total, each value $v_{r}[i] = v_{r}[i - 1] + v_{pos}[i] - v_{neg}[i]$.
       \\ \hline
-    \ecsumc{v_r}{\mathbf{v_x}}
+    $\ecsumc{v_r}{\mathbf{v_x}}^\dagger$
       & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\prows(\mathbf{v_x}) \geq 2$; and $\pcols(\mathbf{v_x}) = \plength(v_r)$
@@
       \\ \hline
-    \ecsumr{v_r}{\mathbf{v_x}}
+    $\ecsumr{v_r}{\mathbf{v_x}}^\dagger$
       & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\pcols(\mathbf{v_x}) \geq 2$; and $\prows(\mathbf{v_x}) = \plength(v_r)$
@@
       \\ \hline
-    \ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}
-      & $v_{fk}, v_{pk}$ are $\discrete$; $v_{r}, v_{val}$ are $\numeric$; within the sets $\{v_{val}, v_{fk}\}$ and $\{v_{pk}, v_{r}\}$ variables have the same $\plength$ and $\por$; $v_{fk}$ and $v_{val}$ have the same $\ptable$; $v_{fk}$ and $v_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{v_{pk}}
+    $\ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}^\dagger$
+      & $v_{fk}, v_{pk}$ are $\discrete$; $v_{r}, v_{val}$ are $\numeric$; within the sets $\{v_{val}, v_{fk}\}$ and $\{v_{pk}, v_{r}\}$ arguments have the same $\plength$ and $\por$; $v_{fk}$ and $v_{val}$ have the same $\ptable$; $v_{fk}$ and $v_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{v_{pk}}
       & The value for $v_{r}[i]$ is obtained by summing all values $v_{val}[j]$ where $v_{fk}[j] = v_{pk}[i]$
@@
     \ecsumprod{v_r}{v_1}{v_2}
-      & Variables $\{v_r, v_1, v_2\}$ are all $\numeric$; $\plength(v_{1}) = \plength(v_{2}) \geq 2$; and $\prows(v_{r}) = \pcols(v_{r}) = 1$
+      & Arguments $\{v_r, v_1, v_2\}$ are all $\numeric$; $\plength(v_{1}) = \plength(v_{2}) \geq 2$; and $\prows(v_{r}) = \pcols(v_{r}) = 1$
       & $v_{r}[i] = \sum_{i = 1}^{\plength(v_{1})} v_{1}[i] \times v_{2}[i]$.
@@
 
-\samuel{Satisfy signature for group assignment: find groups such that there exists a subgroup which satisfies the signature}
+Table~\ref{table:constraints} shows all constraint templates currently supported by our implementation.
+Their dependencies are reflected in their \CSignature and shown in Figure~\ref{fig:learning_order}.
+These constraints have been selected based on the most-used Excel functions as well as their occurrence in various example spreadsheets we observed.
 
@@
   }
-  \caption{Expected constraints in the case study \sergey{in Figure \ref{fig:main_example}; some more explanation needed, not self explanatory}}
+  \caption{Constraints that are satisfied in the case study (Figure~\ref{fig:main_example}) and expected to be found by our implementation. Functional constraint are omitted.}
   \label{fig:sol_example}
 \end{figure}
-\sergey{Samuel, hint, don't put label before caption, it can't resolve it sometimes}
 
@@
 In order for the algorithm to run efficiently it is crucial to use dependencies and find constraint incrementally whenever possible.
-This avoids some of the explosion of combinations for constraints that have variables.
+This avoids some of the explosion of combinations for constraints that have many arguments.
 For example, using foreign keys as a base constraint to find conditional aggregates reduces the running time for the case study from about 3 seconds to below 1 second.
@@
 conclusions 
-points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
+% points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
+Our goal was to automatically identify constraints that hold in a spreadsheet.
+We have presented and evaluated an approach that is able to learn a large set constraints from CSV data.
+
+fututre work
 

### 2016-04-27 Paper/ECAI/tabular.tex
fixing introduction

@@
 
-\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{KU Leuven, Belgium, email: firstname.lastname@kuleuven} }
+\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
 
@@
 
-Let us look at an example in Figure~\ref{fig:main_example}, the header names already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
+Let us look at an example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
 If we look at Table~1 and at the first row in Table~2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
 Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
-I.e., all operations are vectorized and make use of the standard spreadsheet operations.
+I.e., all operations here are vectorized and make use of the standard spreadsheet operations.
 
@@
 Here everything is mixed, some values are missing in the results, some missing the arguments, even a header might have missing values.
-The data is relational on one hand, since we have multiple table with relationships between them.
+The data is relational on the one hand, since we have multiple table with relationships between them.
 And on the other hand, the data is numeric, since people tend to use spreadsheets for financial and accounting computations, which is heavily numeric.
@@
 
-Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to so called the Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and caused billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
+Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
 Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.

### 2016-04-27 Paper/ECAI/tabular.tex
fixing introduction

@@
 \section{Introduction}
+\sergey{Needs to be checked, bullet points for Luc and Tias to rework introduction}\\
 Millions of people across the world use spreadsheets everyday.
@@
 
-\sergey{bullet points for Luc and Tias to rework introduction}\\
 \textbf{Motivation}:
@@
   A \textbf{vector} is either a column or a row that is type-consistent.
-  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
+  If a vector is a row (column), we say that it has the \textit{row} (\textit{column}) orientation.
   \item

### 2016-04-28 Paper/ECAI/tabular.tex
Merge

@@
   A \textbf{vector} is either a column or a row that is type-consistent.
-  If a vector is a row (column), we say that it has the \textit{row} (\textit{column}) orientation.
+  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
+  \samuel{Vector also has a length; Group has a size}
   \item
@@
 \item \CSignature: the two arguments $v_r$ and $v_x$ must be numeric; furthermore, the first argument $v_r$ must be integer and the length of the vectors
+\samuel{Both must be vectors}
 of $v_x$ and $v_r$ be identical, i.e., $length(v_x) = length(v_r)$;
@@
 we can interpret the argument ${\range{T_{1}}{\rangeall}{8}}$ as group variables that would apply to any table,
+\samuel{clarify, e.g. what is group variable}
 or we can interpret the argument as the value it takes in the particular table called $T_1$ that is shown in Figure 1, that is, $[1,2,3,4]$. {\bf CHECK}
@@
 
+\tias{Redundancy instead of dependency}
+
   In theory each constraint is independent and equally useful.
@@
 
+\tias{Reformulate: describe there's an naive approach...}
 Tables are detected by detecting rectangular collections of data and attempting to detect and separate headers from the data.
@@
 
+\tias{Split extensions and group detection, clarify group detection, example (e.g. currency for extensions)}
 Our algorithm also attempts to detect groups within tables by partitioning the data into consecutive, type consistent groups, essentially separating textual and numerical data.
@@
 
+\tias{Move this to problem statement second part}
 A constraint~$c_1$ is a special case of another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
@@
 \\\\
+\sergey{Can be parallelized by looking at connected components and respecting order}
 Most parts of the algorithm could be easily parallelized.
@@
 Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
+\tias{...such that at least one subgroup satisfies the signature}
 For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that those groups or some of their subgroups satisfy the \CSignature.
+
+\tias{Add paragraph on enforcement of group generation}
 
@@
 \subsection{Case study}
+\tias{Move explanation of origin to first time it is introduced}
 Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
@@
 The constraints we considered to be redundant are either duplications (results that can be calculated in different ways, one of which seems superior) or constraints that hold by accident.
+\tias{Advice, move to discussion}
 Duplications should be detected in a post-processing step while accidental constraints should primarily be removed by adding data.
@@
 \subsubsection*{Q3. How fast is the algorithm?}
+\tias{Mention trade-off in the approach}
 Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
@@
 Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.
+
+\sergey{Add summary table of results}
 

### 2016-04-29 Paper/ECAI/tabular.tex
Edits part 1

@@
 \\\\
-\sergey{Can be parallelized by looking at connected components and respecting order}
 Most parts of the algorithm could be easily parallelized.
-However, \dependencies would impose a partial order in which constraint templates would have to be treated.
+Every connected component in \dependencies can be learned independently.
+Within every component, constraint templates have to be treated respecting the order, i.e. learning the most general ones first.
+The generality ordering thus imposes a partial ordering according to which constraint templates must be learned.
 
@@
 Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
-\tias{...such that at least one subgroup satisfies the signature}
-For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that those groups or some of their subgroups satisfy the \CSignature.
+For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that a group~$G \in \groups$ is selected if there is at least one subgroup~$G' \subseteq G$ that satisfies the \CSignature.
 

### 2016-04-29 Paper/ECAI/tabular.tex
Edits part 2

@@
 
+Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty. \samuel{More on this / example?}
+
 % \subsection{Workflow}
@@
 
+We examine three main questions concerning the accuracy, redundancy and efficiency of our approach, our main focus being accuracy.
+
 \subsubsection*{Q1. How accurate is the approach?}
@@
 \subsubsection*{Q2. How many redundant constraints are discovered?}
-Our primary focus was accuracy, therefore we sometimes traded off less redundancy for more accuracy.
+Our primary focus was, as mentioned before, accuracy, therefore we sometimes traded off less redundancy for more accuracy.
 The motivation is that solutions can still be pruned using entailment or heuristics afterwards.
 The constraints we considered to be redundant are either duplications (results that can be calculated in different ways, one of which seems superior) or constraints that hold by accident.
-\tias{Advice, move to discussion}
-Duplications should be detected in a post-processing step while accidental constraints should primarily be removed by adding data.
 
@@
 \subsubsection*{Q3. How fast is the algorithm?}
-\tias{Mention trade-off in the approach}
 Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
@@
 
-\sergey{I think there should be three things: key example evaluated, benchmark of 30 spreadsheets and maybe the FBI example, how we can detect interesting dependencies and correct potential mistakes}
+\paragraph{Redundancy}
+As mentioned before, there are two types of redundant constraints: duplications and accidental constraints.
+Duplications should be detected throught in a post-processing step that uses entailment or heursitics to filter constraints.
+Accidental constraints can sometimes be detected through heuristics, as well, however they can also be removed by adding additional data.
+
+(\sergey{I think there should be three things: key example evaluated, benchmark of 30 spreadsheets and maybe the FBI example, how we can detect interesting dependencies and correct potential mistakes})
 
@@
 \sergey{key bullet points for Luc and possibly Samuel and me to make related work section}
+
+\samuel{Two possible takes here: 1) recent interest in topics such as flashfill, etc -> sparked the idea for this project.
+This actually relates to constraint learning which has seen an uptick recently and is a subfield of long-established ILP.
+2) ILP established field trying to learn structure / .., has been brought to the less traditional context (constraint learning), learning constraints / programs has been picked up by specific approaches such as flashfill etc which inspired this research}
 
@@
 Our goal was to automatically identify constraints that hold in a spreadsheet.
-We have presented and evaluated an approach that is able to learn a large set constraints from CSV data.
+We have presented and evaluated an approach that is able to learn a large set of constraints from CSV data.
 
-fututre work
+\paragraph{Future work}
+...
 

### 2016-05-02 Paper/ECAI/tabular.tex
Expanded table caption

@@
   \end{tabularx}
-  \caption{\template's \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
-etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display} \sergey{1) should be readable alone without lookin in the text of the paper 2) that's the constraints we found popular according the spreadsheets in the web 3) that's what is actually implemented in the system 4) corresponds to the actual Excel functions} These constraints are \textit{structural}, they do not compute a result.
-Our algorithm can and does discover these constraints, however, they cannot be directly used in spreadsheet software }
+  \caption{
+An overview of the different constraint templates that we have implemented.
+Most of the constraint correspond directly to spreadsheet functions and have been selected based on the most popular constraints in Excel and their occurence in spreadsheets that we have collected.
+The constraints marked with $*$ are \textit{structural}, they are discovered by our algorithm, however, they cannot be readily used in spreadsheet software.
+The constraints marked with $\dagger$ are aggregate constraints, the table shows the constraints for sum but the implementation also supports versions for max, min, average, product and count.
+  \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?} 
+  }
+
+  % \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
+etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display} \sergey{1) should be readable alone without lookin in the text of the paper 2) that's the constraints we found popular according the spreadsheets in the web 3) that's what is actually implemented in the system 4) corresponds to the actual Excel functions}}
 

### 2016-05-02 Paper/ECAI/tabular.tex
Rewrote table and group detection

@@
 
-\tias{Reformulate: describe there's an naive approach...}
-Tables are detected by detecting rectangular collections of data and attempting to detect and separate headers from the data.
-Especially this last step is a complex task on its own  \cite{header} and a sophisticated implementation goes beyond the scope of this paper.
-On well formated data, however, the user will not have to specify tables manually.
+\subsection{Table and group detection}
+We chose to implement a naive approach to extract tables as the detection of tables and headers from data is a complex task on its own (\cite{header}) and a sophisticated implementation goes beyond the scope of this paper.
+Our approach first detects rectangular collections of data.
+It then attempts to identify the header by looking at the type of the first row.
+If this row consists only of textual entries, the row is expected to be the header and is removed, if not, the same procedure is applied on the first column.
+We suggest to provide tables in general, for well formated data, however, the user will not have to specify tables manually.
 
-\tias{Split extensions and group detection, clarify group detection, example (e.g. currency for extensions)}
-Our algorithm also attempts to detect groups within tables by partitioning the data into consecutive, type consistent groups, essentially separating textual and numerical data.
-We support some extensions such as currencies and percentual data and our approach allows us to omit manual specifications of groups.
+Our approach for detecting groups works by partitioning the data of a table into consecutive, type consistent groups, essentially separating textual and numerical data.
+The assumption is that often numerical and textual data are not mixed and this approach works well in practice as often groups do not have to be specified manually\footnote{\samuel{Check on benchmark}}.
+Unless an explicit orientation (rows or columns) is provided for the table, this approach will attempt to create both row- and column-groups.
 
-A more elaborate implementation could allow group (and table) detection could be done interactively / visually such that a user can select the groups in a spreadsheet or correct automatically generated groups.
+So far we have considered textual and numerical (floating point or integer) data.
+Our implementation also supports some extensions such as currencies and percentual data, which is considered to be numerical data.
+Moreover, percentual data is corrected to the floating point representation (i.e. divided by 100).
+
+A more elaborate implementation could allow group (and table) detection to be done interactively / visually such that a user can select the groups in a spreadsheet or correct automatically generated groups.
 From here on we will assume that groups and tables are given.

### 2016-05-02 Paper/ECAI/tabular.tex
Comments

@@
 \section{Approach to Tabular Constraint Learning}
-\sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
+% \sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
 This section will describe our approach to \tcl.
@@
 \subsection{Table and group detection}
+\samuel{Is an example needed? I find no because it is not our main concern but open for discussion}
 We chose to implement a naive approach to extract tables as the detection of tables and headers from data is a complex task on its own (\cite{header}) and a sophisticated implementation goes beyond the scope of this paper.

### 2016-05-02 Paper/ECAI/tabular.tex
Added paragraph on group generation and provided some more details on subgroup assignment finder

@@
 
-\tias{Add paragraph on enforcement of group generation}
+\samuel{TODO: correct code integer groups}\\
+Consider the \ecrank{v_r}{v_x} constraint template, then both arguments must be of the same length.
+Therefore, assignments containing groups of different length do not have to be considered.
+Constraints of the \CSignature cannot always be enforced directly on a group but have to consider whether subgroups could satisfy them.
+For example, the \CSignature requires $v_r$ to be integer.
+If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this this group cannot be discarted.
 
@@
 \subsubsection{Subgroup satisfaction search}
-$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied.
-For every \template~$t$ an implementation (declarative or imperative) needs to be provided that finds valid subgroups given the assignments.
+$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
+the \CSignature and \CFunction of~$t$ are satisfied).
+For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.
 

### 2016-05-02 Paper/ECAI/tabular.tex
rework intro, first pass

@@
 \section{Introduction}
-\sergey{Needs to be checked, bullet points for Luc and Tias to rework introduction}\\
 Millions of people across the world use spreadsheets everyday.
-Yet, many people lack a complete understanding of the structures and existing dependencies in the data.
-Indeed, the most useful part of spreadsheets is often the formulas and often these formula disappear when a file is exported and shared, for example as a CSV file.
-This brings us to the question: can we discover or reconstruct structural relations in flat tabular spreadsheet data?
-Moreover, can we accomplish this task in a general way that allows declarative specification of constraints to discover?
+The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
+However, large sheets (possible with multiple tables) and relations between different tables and sheets can be tedious and error-prone to handle.
+As a result, many people lack a complete understanding of the structures and existing dependencies in the data.
 
-Let us look at an example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of standard spreadsheet operations such as \textit{average} or \textit{sum}.
-If we look at Table~1 and at the first row in Table~2, it is clear that one could have selected the column called \textbf{1st Quarter}, clicked on the \textit{SUM} function and then just used the drag fill handle to apply the formula to the other columns.
-Similarly, that would work for the second (\textit{Average}), third (\textit{Min})  and forth (\textit{Max}) rows in Table 2.
-I.e., all operations here are vectorized and make use of the standard spreadsheet operations.
+This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages.
+In this case, often a comma-separated values (CSV) format is used, meaning that all formula's are lost, including inter-sheet formulas and relations.
 
+In this work, we investigate in how far machine learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
+However, this is an unconventional machine learning problem.
+Consider the example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of spreadsheet operations such as \textit{average} or \textit{sum}.
+Looking at the first row in Table~2 and the data in Table~1, it is clear that the values in this row were computed by summing the values in the corresponding column of Table~1.
 Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
+This already provides a flavor of how this problem is different from standard data mining settings, when the data is just in rows and variables are in columns.
+Here everything is mixed.
+The data is relational on the one hand, since we have multiple table with relationships between them.
+And on the other hand the data is mixed textual and numeric, since people tend to use spreadsheets for financial and accounting computations.
+One can see that understanding and learning constraints in tabular data is hence a new and challenging problem setting.
+%Furthermore, it is not only challenging for a machine but for a human as well.
 
-That gives us already a hint how this problem is different from the standard data mining settings, when the data is in rows and variables are in columns.
-Here everything is mixed, some values are missing in the results, some missing the arguments, even a header might have missing values.
-The data is relational on the one hand, since we have multiple table with relationships between them.
-And on the other hand, the data is numeric, since people tend to use spreadsheets for financial and accounting computations, which is heavily numeric.
+Relatedly, the field of Inductive Logic Programming has long...
+\tias{TODO for Sergey or Luc}.
+Additionally, the field of Constraint Programming has recently started investigating ways in which Constraint Satisfaction Problems can automatically be inferred from known solutions.
+Here, methods ranging from generate-and-test~\cite{todo:modelseeker}, concept learning~\cite{todo:conacq} and Inductive Logic Programming~\cite{todo:lallouet} have been used with encouraging results.
 
-Given these novel aspects, we conclude that understanding and learning constraints in tabular data is a new and challenging problem setting.
-Furthermore, it is not only challenging for a machine but for a human as well.
+Another line of work from which we draw inspiration is that of \tias{TODO for ?, flashfill et al}.
+We will consider the case where all data is given at once.
+However, as in the use case of flashfill, one could also use this technique to derive formula's interactively while a user is entering values in a cell \tias{TODO SAMUEL: proofread/refine last statement}.
 
-Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
+The question that we hence try to answer in this work is: can we discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
+Moreover, we investigate general-purpose ways for doing so, that do not require hard coding a number of constraints.
+Rather, we use techniques from constraint solving to derive whether and which constraints are satisfied.
+This approach can be extended to any constraint in a principled way.
+
+%Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
 Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
@@
 \textbf{Motivation}:
-\samuel{Suggest better table layouts / formula translation}
 \begin{itemize}
-  \item File generated from model, model got lost, need to reconstruct
+  \item USED -- File generated from model, model got lost, need to reconstruct
   \item Constraint programming is hard - is Excel hard?
   \item Avoid manual analysis, provide selection of constraints
-  \item Error checking
+  \item SOMEWHAT USED --Error checking
   \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
@@
 \begin{itemize}
-  \item Unsupervised setting (contrary to flashfill, etc)
+  \item USED -- Unsupervised setting (contrary to flashfill, etc)
   \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
-  \item Data format (2D) -- data is no longer in rows like a classic ML or DM settings
-  \item Declarative, general / modular, stacking of constraint problems
+  \item USED -- Data format (2D) -- data is no longer in rows like a classic ML or DM settings
+  \item USED -- Declarative, general / modular, stacking of constraint problems
 \end{itemize}
-
-\sergey{need to elaborate the example here: like in a story}
 

### 2016-05-02 Paper/ECAI/tabular.tex
pass over section2

@@
 
+{\it
 \textbf{Motivation}:
@@
   \vspace{-10pt}
-  \caption{An example of constraint reconstruction (in blue) with indicated groups (in green)}
+  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations.}
   \label{fig:main_example}
 \end{figure*}
+}
 
 \section{Formalization}
-\samuel{Explain concepts on example}
-\sergey{subgroup notation, rework the solution definition}
-\subsection{Groups, Type-consistency and Constraints}
-\samuel{Instantiated table... references}
-In this work we focus on spreadsheet data, which is physically one big table, but conceptually often consists of multiple tables, see Figure~\ref{fig:main_example} for an example.
-Formally, a table is an $n \times m$ matrix, each table entry is called  a \textit{cell}.
+In this work we focus on spreadsheet data.
+We will first introduce the general concepts of cells, group and constraint templates, using the (multi-table) sheet in Figure~\ref{fig:main_example} as running example.
 
-Each cell has a {\em type} which can be numeric or textual.
-The numeric type has two subtypes: integers and floats.
-We also consider the special element called \textit{None}, which can be either type: numeric or textual.
-A set of cells is called \textit{type-consistent} iff all cells in the set are either numeric or textual.
+\subsection{Cells, vectors and groups}
+A spreadsheet is physically one big table, but conceptually may consist of multiple tables, such as in Figure~\ref{fig:main_example}.
+We will describe a method of extracting tables from a big spreadsheet in Section~\ref{todo}, however this is not the core of our learning method.
+Our formalization reasons over tables directly.
+
+Formally, a table is an $n \times m$ matrix. Each entry is called a \textit{cell}.
+A cell has a {\em type}, which can be numeric or textual.
+We further distinguish numeric types in subtypes integer and float.
+We also consider \textit{None} as a special type when a cell is empty; \textit{None} is a subtype of all other types.
+A set of cells is called \textit{type-consistent} iff all cells in the set are of the same type.
 %Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric subtype by requiring its arguments to be integers.
 
-We consider various sets of cells in tables:
+We consider various sets of cells with the following notation:
 \begin{itemize}
-  \item {\bf Cells}: Each entry in the matrix/table is called  a \textit{cell}.
-  We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$.
-  \item {\bf Columns and rows}: To refer to the $a$-th row (column) we use the following syntax $T[a,{:}]$ ($T[{:},a]$), to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
+  \item {\bf Cell}: We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$.
+
+  \item {\bf Columns and rows}: For a table $T$, we use notation $T[a,{:}]$ to refer to the $a$-th row, and similarly $T[{:},a]$ for the $a$-th column.
+% to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
+
   \item
-  A \textbf{vector} is either a column or a row that is type-consistent.
-  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation.
-  \samuel{Vector also has a length; Group has a size}
+  A \textbf{vector} is a type-consistent row or column.
+  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation. The \textit{length} of a vector is the number of cells in that vector.
+
   \item
-  A \textbf{group} is a subrange of vectors with the same orientation in a table.
-  We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.
-  The \textit{length} of a row (column) group $G$, written as \textit{length(G)}, is the number of its columns (rows); We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
+  A \textbf{group} is a projection of a matrix onto a set of consecutive rows or columns. %In this work we only consider consecutive rows and columns.
+  We use the following notation to refer to a row-oriented group with rows ranging from $a$ to $b$: $G = T[a{:}b,:]$, where $a,b$ are natural numbers.
+Similarly, we have $G = T[{:},a{:}b]$ for a column-oriented group.
+  The vectors in a group have to be type-consistent, consecutive in the original table and of equal length.
+  The \textit{size} of a group is the number of vectors in the group, while the \textit{length} of a group is the length of its vectors. 
+  %The \textit{length} of a row (column) group $G$, written as \textit{length(G)}, is the number of its columns (rows); We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
+
   \item
- A \textit{subgroup} $g$ is a subrange of vectors belonging to a group $G$.
+  A \textbf{subgroup} $G'$ of $G$ is simply a projection of $G$ onto a consecutive subset its vectors.
+% is a subrange of vectors belonging to a group $G$.
 If the subgroup $g$ must contain only one vector, we write $g \in G$.
@@
 
-When we do not want to distinguish between cells, columns, rows, vectors or groups, we shall talk about a {\em set of cells}.
+%When we do not want to distinguish between cells, columns, rows, vectors or groups, we shall talk about a {\em set of cells}.
 
@@
 Consider Table~1 in the Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
-The (column) vectors of Table~1 are subdivided into five groups:
+Ignoring the (all-textual) header, Table~1 can be divided into the following five column-oriented groups:
+{\small
 \begin{align*}
@@
 \end{align*}
-In order to sum the results of a salesperson across quarters we can consider for example the subgroup $\range{T_1}{\rangeall}{\rangeto{3}{6}} \subset G_3$
+}
+An example subgroup is $\range{T_1}{\rangeall}{\rangeto{3}{6}} \subset G_3$, which contains the sales numbers of all employees for the four quarters.
 
-\subsection{Constraints}
-A \template is a triple \textit{(\CName, \CSignature, \CFunction)}.
-Let us elaborate on this:
+\subsection{Constraint templates}
+Our method will extract the applicable constraints (including functions) in a set of tables, based on \textit{constraint templates}.
+%
+A \template is a triple \textit{(\CName, \CSignature, \CFunction)}:
+%Let us elaborate on this:
 \begin{itemize}
@@
 \textit{\CName}  specifies the syntactic form of the constraint $c(A_1, ...,A_n)$, that is, the name of the constraint $c$ together
-with $n$ arguments $A_i$. Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
-\samuel{Or a function}
+with $n$ arguments $A_i$.
+Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
+Note that a function $B=f(A_1,...,A_n)$ can be represented with the $n+1$-ary predicate $cf(B,A_1,...,A_n)$.
 % the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
-\item the \textit{\CSignature} constrains the arguments $A_i$ of the constraint, that is, the types of the arguments, together
-with possible further restrictions using properties of the arguments $A_i$ such as their length.  For example, the signature can restrict the types of the arguments
-to be vectors of integers and require that  the length of all arguments be equal.  In terms of logical and relational learning, this is known as the {\em bias} of the learner,
-it specifies when the constraint is well-formed.
-  %The constraints act on properties or group meta-information, not on the actual group content. In ILP, it is known as bias,
-  \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds, in practice this will be a function that can be called when
-the arguments are full instantiated to decide whether the constraint is satisfied or not.
+
+\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy. That is, the required types of the arguments, together
+with possible further restrictions such as on their length.
+For example, the signature can restrict the types of the arguments to be groups integers and require that the length of all arguments be equal.
+\\
+In terms of logical and relational learning, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
+
+\item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
+Given an assignments to all its arguments, it can verify whether the constraint is satisfied or not.
+%holds, in practice this will be a function that can be called when
+%the arguments are full instantiated to decide whether the constraint is satisfied or not.
 In logical and relational learning this is known as the background knowledge, which provides the definition of the predicate.
@@
 \paragraph{Example}
-Several example constraints are illustrated in Table 1.  For example, the \template \textit{rank} has
+Several example constraints are illustrated in Table 1.  For example, the constraint template \textit{rank} has
 \begin{itemize}
-\item \CName: \textit{$v_r$ = RANK($v_x$)}, where $v_x$ and $v_r$ are the arguments;
-\item \CSignature: the two arguments $v_r$ and $v_x$ must be numeric; furthermore, the first argument $v_r$ must be integer and the length of the vectors
-\samuel{Both must be vectors}
-of $v_x$ and $v_r$ be identical, i.e., $length(v_x) = length(v_r)$;
-\item \CFunction: \textit{$v_r$ = RANK($v_x$)} whenever each value of $v_r$ is the rank of the corresponding value in $v_x$.
+\item \CName: \textit{$B$ = RANK($A$)}, where $B$ and $A$ are the arguments;
+\item \CSignature: the two arguments $B$ and $A$ must be numeric groups and of size 1; furthermore, the group $B$ must be integer (the rank) and the length of groups $A$ and $B$ must be identical, i.e., $length(B) = length(A)$;
+\item \CFunction: \textit{$B$ = RANK($A$)} whenever each value of $B$ is the rank of the corresponding value in $A$.
 % is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
@@
 
-When looking into constraints like \textit{$v_r$ = RANK($v_x$)}, it is important to see the analogy with both first order logic and with constraint satisfaction problems.
-From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $v_r$ and $v_x$ are the terms, which can be either variables or values.
+When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
 This also holds in our setting: when we write
-      $\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}$,
-we can interpret the argument ${\range{T_{1}}{\rangeall}{8}}$ as group variables that would apply to any table,
-\samuel{clarify, e.g. what is group variable}
-or we can interpret the argument as the value it takes in the particular table called $T_1$ that is shown in Figure 1, that is, $[1,2,3,4]$. {\bf CHECK}
-This also means that we can talk about constraints being satisfied. Basically, a constraint $c(t_1, ..., t_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
-if the constraint holds on the values that the $t_i$ take in the set of tables ${\cal T}$. Figure 3 provides a list of constraints that are satisfied by the tables
-in Figure 1.
+      $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
+we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table \tias{Of the right (minimum) dimension??
+Note that this is also sloppy because the ':8' is not considered part of the variable...},
+or we can interpret the argument as the value it takes in a particular table $T$ such as $T_1$ in Figure 1, that is, $[1,2,3,4]$.
+This also means that we can talk about constraints being satisfied. Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
+if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
+\tias{I find this confusing: a group as we defined it is defined on one table, not on multiple tables.
+we should separate the table variable from the 'range' variables then...}
+%Figure 3 provides a list of constraints that are satisfied by the tables in Figure 1.
 

### 2016-05-02 Code/src/draw.py
Added more evaluation data

@@ -1,0 +1,22 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def main():
+    m = [[0.05, 0.05, 0.05, 0.06, 0.06, 0.06, 0.05, 0.05, 0.05, 0.05], [0.07, 0.08, 0.08, 0.09, 0.10, 0.09, 0.07, 0.07, 0.08, 0.07], [0.02, 0.02, 0.02, 0.02, 0.02, 0.01, 0.01, 0.01, 0.02, 0.02], [0.08, 0.08, 0.07, 0.08, 0.08, 0.08, 0.07, 0.08, 0.07, 0.07], [2.67, 2.68, 2.66, 2.97, 2.77, 2.67, 2.69, 2.69, 2.68, 2.66], [0.12, 0.13, 0.12, 0.13, 0.13, 0.13, 0.12, 0.12, 0.12, 0.12], [0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02], [8.39, 8.73, 8.48, 9.72, 9.12, 8.56, 8.54, 8.24, 8.28, 8.41], [0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02], [0.06, 0.06, 0.06, 0.06, 0.06, 0.06, 0.06, 0.06, 0.06, 0.06], [0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02], [0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02, 0.02], [0.05, 0.05, 0.06, 0.06, 0.06, 0.05, 0.05, 0.05, 0.05, 0.05], [0.03, 0.03, 0.03, 0.03, 0.03, 0.03, 0.03, 0.03, 0.03, 0.03], [0.09, 0.11, 0.10, 0.08, 0.09, 0.09, 0.09, 0.10, 0.09, 0.09], [0.09, 0.09, 0.09, 0.08, 0.09, 0.08, 0.09, 0.09, 0.09, 0.10], [0.04, 0.04, 0.04, 0.04, 0.04, 0.04, 0.04, 0.04, 0.04, 0.04], [0.07, 0.07, 0.08, 0.07, 0.08, 0.07, 0.07, 0.07, 0.07, 0.08], [0.02, 0.02, 0.02
[…]

### 2016-05-04 Paper/ECAI/tabular.tex
Added more evaluation data
first stab at abstract

@@
 \begin{abstract}
-  abstract
+Spreadsheets, comma separated value files and other tabular data representations are in wide use today.
+However, the correct use and maintenance of functions in and between the tabular data can be error-prone and overwhelming.
+In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
+Our method takes inspiration form inductive logic programming and constraint satisfiability.
+We represent common spreadsheet functions as predicates with arguments that must satisfy number of constraints over the arguments.
+This allows generic constraint satisfaction techniques to be used to find all such satisfying predicates/functions within the rows, columns and submatrices of the tables.
+We show the effectiveness and accuracy of this approach on a number of spreadsheets from varying sources.
 \end{abstract}
@@
 
+\newcommand{\runtotal}{16.12}
+\newcommand{\runtotalstd}{0.62}
+
+\newcommand{\runfile}{0.50}
+\newcommand{\runfilestd}{0.02}
+
 \section{Evaluation}
@@
 Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
-For the 32 spreadsheets in the benchmark our implementation ran in $17.86s \pm 0.87s$.
+For the 32 spreadsheets in the benchmark our implementation ran in ${\runtotal}s \pm {\runtotalstd}s$.
 The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
@@
 For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).
+
+\begin{table}
+  \centering
+  \begin{tabular}{lll}
+    & \textbf{Total} & \textbf{Average per spreadsheet} \\
+    \textbf{Essential constraints} & $93$ & $2.91$ \\
+    \textbf{Accuracy} & $96.77\%$ & $94.27\%$ \\
+    \textbf{Redundant} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
+    \textbf{Speed (s)} & ${\runtotal}s \pm {\runtotalstd}s$ & ${\runfile}s \pm {\runfilestd}s$
+  \end{tabular}
+  \caption{Overview evaluation on essential constraints}
+\end{table}
 
@@
 Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.
-
 \sergey{Add summary table of results}

### 2016-05-04 Paper/ECAI/tabular.tex
Abstract v2

@@
 \begin{abstract}
-Spreadsheets, comma separated value files and other tabular data representations are in wide use today.
+%Spreadsheets, comma separated value files and other tabular data representations are in wide use today.
 However, the correct use and maintenance of functions in and between the tabular data can be error-prone and overwhelming.
@@
 We show the effectiveness and accuracy of this approach on a number of spreadsheets from varying sources.
+Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
+However, modeling, maintaining and discovering formulas in tabular data and between spreadsheets can be time consuming and error-prone.
+In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
+We see multiple promising applications for this technique, e.g.
+in rediscovering constraints, auto-completion and error checking.
+Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfiability.
+Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
+Constraint satisfaction techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
+We show that our approach is able to accurately discover constraints in a number of spreadsheets from varying sources.
 \end{abstract}

### 2016-05-09 Paper/ECAI/tabular.tex
Updated abstract
Updates (problem statement amongst others)

@@
 in rediscovering constraints, auto-completion and error checking.
-Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfiability.
+Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
 Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
 Constraint satisfaction techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
+Therefore, new types of constraints can be easily added to the system by specifying them using declarative methods such as Minizinc or ASP.
 We show that our approach is able to accurately discover constraints in a number of spreadsheets from varying sources.
@@
 A {\em constraint satisfaction problem} (CSP) consists of a set of constraints and the
-task is to find assignments of values to the variables in the CSP so that all constraints are satisfied.  As the variables typically refer to
-cells in tables, solving a CSP problem often means completing the tables so that all constraints are satisfied.
+task is to find assignments of values to the variables in the CSP so that all constraints are satisfied.
+In the context of spreadsheets the variables typically refer to cells in tables, and solving the corresponding CSP problem means completing the tables so that all constraints in the spreadsheet are satisfied.
 
-While the traditional CSP problem is given a set of variables and a set of constraints,
-and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
+While the traditional CSP problem is given a set of variables and a set of constraints, and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
 \\
@@
 \item
-a set of \template ${\cal C}$;
+a set of {\template}s ${\cal C}$;
 \item
@@
 \noindent
-%\samuel{This section is incorrect, yet, we find assignments over subgroups, associated with a \template}
-{\bf Find}  the set of all constraints $c(G_1, ... , G_N)$ that 1) are satisfied  in ${\cal T}$, 2)
-for which all $G_i \in {\cal G}$ and  3)  $c(G_1, ... , G_N)$ satisfies the signature of the constraint $c$.
+{\bf Find}  the set of all constraints $s(G_1', ..., G_n')$ that 1) are satisfied  in ${\cal T}$, 2)
+for which all $G_i' \subseteq G_i \in {\cal G}$ and  3) the groups $G_1', ... , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
 
 For instance, Figure 3 contains the solution to the tabular constraint learning problem
-that uses the tables of Figure 1, the signatures of Table 1 and the set of groups mentioned in {\bf CHECK / COMPLETE}
+that uses the tables of Figure 1, the signatures of Table 1 and the set of groups mentioned in (\samuel{show groups somewhere}).
 
@@
 A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.
+\samuel{Formalize: $\forall x1..x_k s_1(x_1...x_n) => s_2(x_1...x_k)$ :: looking for most specific}
 
@@
 
+Redundant constraints can be reduced in different ways.
+Duplications should be detected in a post-processing step that uses entailment or heursitics to filter constraints.
+Accidental constraints may be detected through heuristics, however, they can be also be removed by adding additional data.
+
 \subsubsection*{Q3. How fast is the algorithm?}
@@
 
-\section{Discussion}
+\section{Applications}
 \sergey{We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.}
@@
 
-\paragraph{Redundancy}
-As mentioned before, there are two types of redundant constraints: duplications and accidental constraints.
-Duplications should be detected throught in a post-processing step that uses entailment or heursitics to filter constraints.
-Accidental constraints can sometimes be detected through heuristics, as well, however they can also be removed by adding additional data.
+\samuel{application autocomplete}
+
+\samuel{FBI: work out error detection algorithm}
 

### 2016-05-09 Paper/ECAI/tabular.tex
reworked the problem statement

@@
 \renewcommand{\arraystretch}{1.5}
+
+\usepackage{amsthm}
+
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
 
@@
 While the traditional CSP problem is given a set of variables and a set of constraints, and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
-\\
-{\bf Given }
-\begin{itemize}
-\item
-a set of instantiated tables (so all cells have values) ${\cal T}$;
-\item
-a set of {\template}s ${\cal C}$;
-\item
-a set of groups ${\cal G}$ for the tables ${\cal T}$;
-\end{itemize}
-\noindent
-{\bf Find}  the set of all constraints $s(G_1', ..., G_n')$ that 1) are satisfied  in ${\cal T}$, 2)
-for which all $G_i' \subseteq G_i \in {\cal G}$ and  3) the groups $G_1', ... , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
-
+\begin{definition}[Tabular Constraint Learning]\label{def:problem_statement}
+  Let ${\cal T}$ a set of instantiated tables (all cells have values),  ${\cal C}$ be a set of {\template}s, ${\cal G}$ be a set of groups  for the tables ${\cal T}$, then find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, and the groups $G_1', ...
+, G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
+\end{definition}
+% {\bf Given }
+% \begin{enumerate}
+% \item
+% a set of instantiated tables (so all cells have values) ${\cal T}$;
+% \item
+% a set of {\template}s ${\cal C}$;
+% \item
+% a set of groups ${\cal G}$ for the tables ${\cal T}$;
+% \end{enumerate}
+% \noindent
+% {\bf Find}  the set of all constraints $s(G_1', ..., G_n')$ that 
+% \begin{enumerate}
+% \item are satisfied  in ${\cal T}$;
+% \item for which all $G_i' \subseteq G_i \in {\cal G}$ ;
+% \item the groups $G_1', ... , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
+% \end{enumerate}
 For instance, Figure 3 contains the solution to the tabular constraint learning problem
@@
 
-\samuel{Not reworked yet from here ---}
-
 The key observation here is that essentially the problem is a constraint enumeration problem, where each constraint is independent.
@@
 
-\tias{Redundancy instead of dependency}
+In theory each constraint is independent and equally useful.
+In practice, however, things are different and some constraints are considered to be \textit{redundant}.
+Assume that for some $g_x$ both \textit{series($g_x$)} and \textit{alldifferent($g_x$)} hold.
+From a user's perspective the last solution is useless, since he or she already knows that the series constraint implies the alldifferent constraint, or formally speaking $\forall g_x{:}~\textit{series}(g_x) \rightarrow \textit{alldifferent}(g_x)$  holds.
+To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity order between constraints, which is depicted using solid arrows in Figure \ref{fig:learning_order}.
+This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all.
 
-  In theory each constraint is independent and equally useful.
-In practice, however, things are different.
-Assume that for some $g_x$ both \textit{series($g_x$)} and \textit{alldifferent($g_x$)} hold.
-From a user's perspective the last solution is useless, since he or she already knows that the series constraint implies the alldifferent constraint.
-To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity (generality) order between constraints, which is depicted using solid (dashed) arrows in Figure \ref{fig:learning_order}.
-This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all.
-A solid lines from $t_1$ to $t_2$ indicates that $t_2$ is \textit{more specific} than $t_1$ and if $C_1=t_1^S$ from $t_1$ does not hold, then the constraint $C_2=t_2^S$ from $t_2$ does not hold neither.
-A dashed arrow from $t_1$ to $t_2$ indicates that $t_1$ is \textit{a more specific} constraint than $t_2$ and therefore if $C=t_1^S$ from $t_1$ holds it should not be tested for $t_2$.
+Formally, we say that a constraint template $t_1$ is \textit{more specific} than a constraint template $t_2$ iff for $t_1$ and $t_2$  holds the statements of the form: $\forall g_1, \dots, g_n{:}~ t_1(g_1,\dots,g_n) \rightarrow t_2(g_h, \dots, g_k)$ where $1 \leq h \leq k \leq n$.
+The intuition behind the formula is the following, if a list of groups is a solution for $t_1$ then its subset is also a solution for $t_2$.
 
-Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, \dependencies be the DAG of constraint specificity, then the constraint $C$ equal to $t^S$ is called \textit{the most specific} wrt \dependencies iff $C$ holds and there is no other template $t_{*}$ for which the constraint $C_{*}=t_{*}^S$ holds and $t_{*}$ is more specific in \dependencies.
+Let us introduce the version of tabular constraint learning by modifying Definition \ref{def:problem_statement} that takes into account the specificity relation between constraint templates.
 
-Let us introduce the version of tabular constraint learning that takes into account the specificity relation between constraints.
+\begin{definition}[Condensed Tabular Constraint Learning]\label{def:condensed_problem_statement}
+  Let ${\cal T}$ a set of instantiated tables (all cells have values),  ${\cal C}$ be a set of {\template}s,  ${\cal G}$ be a set of groups for the tables ${\cal T}$, \dependencies be the set of all specifity rules, then find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, the groups $G_1', ...
+, G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$ and there is no $s'$ in ${\cal C}$ such that $s'(G_h',\dots,G_k')$ holds (for $1 \leq h \leq k \leq n$) and  $s'$ is more specific than $s$ in \dependencies.
+\end{definition}
 
-\begin{minipage}[c]{14em}
-  \vspace{5pt}
-  \begin{tabular}{ll}
-    \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
-    \vspace{-4pt}
-    &\\
-    \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
-    & a constraint specificity DAG \dependencies \\
-    \textbf{Find:}& all the most specific wrt \dependencies constraints $C$ over \groups\\
-    & for each $t$ in \constraints \\
-  \end{tabular}
-  \vspace{6pt}
-\end{minipage}
+% \begin{minipage}[c]{14em}
+%   \vspace{5pt}
+%   \begin{tabular}{ll}
+%     \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
+%     \vspace{-4pt}
+%     &\\
+%     \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
+%     & a constraint specificity DAG \dependencies \\
+%     \textbf{Find:}& all the most specific wrt \dependencies constraints $C$ over \groups\\
+%     & for each $t$ in \constraints \\
+%   \end{tabular}
+%   \vspace{6pt}
+% \end{minipage}
 
 This version explicitly takes into account the specificity of the constraints and invalidates some of the solutions that are entailed by the tabular other solutions.
+
+\sergey{still need to remove some repetitions later on about specisifity}
 
@@
 \subsubsection{Generality ordering}
+\sergey{should we show all the formulas here? or some maybe and say that the specificity DAG can be extracted?}
 

### 2016-05-10 Paper/ECAI/tabular.tex
Conclusion, removed discussion

--- 2016-05-09 Paper/ECAI/tabular.tex
@@
 
+\samuel{discussion: what would be a language to extend the system, what would be the user interface?
+can it be used by community?
+can it be complitely automated?
+interaction with user and highlighing?
+}
+
 {\bfseries
@@
 
-\section{Discussion and Conclusions}
+\section{Conclusions}
 \sergey{link to the github with implementations and the dataset}
@@
 \sergey{Future version, extensions, declarative stuff, bla-bla here}
-discussion: what would be a language to extend the system, what would be the user interface? can it be used by community?  can it be complitely automated? interaction with user and highlighing?
 
@@
 We have presented and evaluated an approach that is able to learn a large set of constraints from CSV data.
+Moreover, the amount of redundant constraints found by the algorithm is limited despite the lack of more sophisticated post-processing steps.
 

### 2016-05-10 Paper/ECAI/tabular.tex
quick and dirty update to CIKM/ACM style
some reworking of problem statement, not finished at all

@@
-\documentclass{ecai}
+\documentclass{sig-alternate-05-2015}
+
 \usepackage{times}
@@
 
-\usepackage{amsthm}
-
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
+%\usepackage{amsthm} % incompatible with ACM
+%\theoremstyle{definition}
+%\newtheorem{definition}{Definition}
+\newdef{definition}{Definition} % ACM specific
 
@@
 
-\title{Tabular Constraint Learning}
+\title{Learning constraints in tabular data}
+% see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
 
-\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
+%\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
 
@@
 \section{Problem Statement}
-A {\em constraint satisfaction problem} (CSP) consists of a set of constraints and the
+The problem of learning constraints from tabular data can be seen as an inverse {\em constraint satisfaction problem} (CSP). A CSP consists of a set of constraints and the
 task is to find assignments of values to the variables in the CSP so that all constraints are satisfied.
-In the context of spreadsheets the variables typically refer to cells in tables, and solving the corresponding CSP problem means completing the tables so that all constraints in the spreadsheet are satisfied.
+In the context of spreadsheets where variables could represent cells in tables, solving a corresponding CSP problem would mean completing the tables so that all constraints of the spreadsheet are satisfied.
 
-While the traditional CSP problem is given a set of variables and a set of constraints, and the task is to find an assignment, the {\bf inverse CSP} or the {\bf Tabular Constraint Learning Problem} can be defined as follows.\\
-\begin{definition}[Tabular Constraint Learning]\label{def:problem_statement}
-  Let ${\cal T}$ a set of instantiated tables (all cells have values),  ${\cal C}$ be a set of {\template}s, ${\cal G}$ be a set of groups  for the tables ${\cal T}$, then find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, and the groups $G_1', ...
-, G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
+The inverse problem is, given an assignment to all cells, to find the constraints that are present in the spreadsheet.
+We define the {\bf Tabular Constraint Learning Problem} as follows:
+%
+\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}
+  Given ${\cal G}$ a set of groups derived from fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal C}$ ; find the set of all constraints $s(G_1', ..., G_n')$ that match the syntax of a \template in ${\cal C}$ and where the groups $G_1', ...
+, G_n'$ with $\forall i, G_i' \subseteq G_i, G_i \in {\cal G}$ satisfy both the signature and the definition of that \template.
 \end{definition}
@@
 % \end{enumerate}
-For instance, Figure 3 contains the solution to the tabular constraint learning problem
-that uses the tables of Figure 1, the signatures of Table 1 and the set of groups mentioned in (\samuel{show groups somewhere}).
-
-% \luc{Delete from here}
-% \luc{ADD/REFER TO EXAMPLE}
+An example solution for the type-consistent groups of the tables in Figure~\ref{fig:main_example} and {\template}'s in Table~\ref{table:constraints} is shown in Figure~\ref{fig:sol_example}.
 
@@
 
-The key observation here is that essentially the problem is a constraint enumeration problem, where each constraint is independent.
-This property comes from fact that in spreadsheets each formula is applied based on the existing data in the cells.
-This allows learning constraints independently of each other by examining constraint satisfaction on the groups.
+As is well-studied in Inductive Logic Programming, some constraints can \textit{entail} other constraints, meaning that some constraints are \textit{implied} by others.
+More formally, a constraint $c_1$ implies another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
+As a result, a solution to the tabular constraint learning problem may contain constraints that are implied by others.
+We can hence say that there is \textit{redundancy} in the solution set of constraints.
+
+Two types of impliciation:
+
+1) on constraint definitions for same signature (example: permutation and alldiff or foreign-key and alldiff)
+
+2) on constraint signature for same definition (example: ?) \textit{Formally, we say that a constraint template $t_1$ is \textit{more specific} than a constraint template $t_2$ iff for $t_1$ and $t_2$  holds the statements of the form: $\forall g_1, \dots, g_n{:}~ t_1(g_1,\dots,g_n) \rightarrow t_2(g_h, \dots, g_k)$ where $1 \leq h \leq k \leq n$.
+The intuition behind the formula is the following, if a list of groups is a solution for $t_1$ then its subset is also a solution for $t_2$.}
+
+3) combination
+
+\tias{OLD TEXT:}
+
+A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.
+\samuel{Formalize: $\forall x1..x_k s_1(x_1...x_n) => s_2(x_1...x_k)$ :: looking for most specific}
+
+\paragraph{Example}
+Consider templates \textit{all-different} and \textit{foreign-key}.
+If \ecfkey{G_{pk}}{G_{fk}}, then \ecalldiff{G_{pk}} must also hold, therefore, \textit{all-different} is considered to be more general than \textit{foreign-key}.
+\\\\
+In ILP one often exploits these type of generality relations to structure the search for constraints using refinement operators \cite{luc_book}.
+Using such techniques leads to a more efficient traversal of the search space (of constraints).
+Aside from the efficiency gained by exploiting constraints learned earlier, the generality also allows us to build constraint templates on top of more general templates.
 
@@
 This section will describe our approach to \tcl.
+
+The key observation is that essentially the problem is a constraint enumeration problem, where each constraint is independent.
+This property comes from fact that in spreadsheets each formula is applied based on the existing data in the cells.
+This allows learning constraints independently of each other by examining constraint satisfaction on the groups.
+
 Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
@@
 \subsubsection{Generality ordering}
-\sergey{should we show all the formulas here? or some maybe and say that the specificity DAG can be extracted?}
+\tias{ordering is because some constraint templates use other templates in their signature, doesn't have to be more complicated?}
 
@@
 \end{figure}
-
-\tias{Move this to problem statement second part}
-A constraint~$c_1$ is a special case of another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
-We consider $c_1$ to be more specific than $c_2$ or $c_2$ to be more general than $c_1$.
-A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.
-\samuel{Formalize: $\forall x1..x_k s_1(x_1...x_n) => s_2(x_1...x_k)$ :: looking for most specific}
-
-\paragraph{Example}
-Consider templates \textit{all-different} and \textit{foreign-key}.
-If \ecfkey{G_{pk}}{G_{fk}}, then \ecalldiff{G_{pk}} must also hold, therefore, \textit{all-different} is considered to be more general than \textit{foreign-key}.
-\\\\
-In ILP one often exploits these type of generality relations to structure the search for constraints using refinement operators \cite{luc_book}.
-Using such techniques leads to a more efficient traversal of the search space (of constraints).
-Aside from the efficiency gained by exploiting constraints learned earlier, the generality also allows us to build constraint templates on top of more general templates.
 

### 2016-05-10 Paper/ECAI/tabular.tex
Added future work

@@
 
-conclusions 
 % points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
 Our goal was to automatically identify constraints that hold in a spreadsheet.
-We have presented and evaluated an approach that is able to learn a large set of constraints from CSV data.
+We presented and evaluated an approach that is able to learn a large set of constraints from CSV data.
 Moreover, the amount of redundant constraints found by the algorithm is limited despite the lack of more sophisticated post-processing steps.
@@
 \paragraph{Future work}
-...
+We see multiple directions for future work on this topic.
+The system we presented would become more useful in an interactive setting where users can easily receive and provide feedback.
+Moreover, the system could be adapted to produce less redundant constraints through the use of heuristic filtering.
+
+Currently, the system only learns single constraints, however, extending this approach to nested constraints would allow more expressive concepts to be learned.
+This shift would bring the approach more in line with programming by example.
+
+Aside from finding errors, the system could be extended to deal with noise and the ability to also learn soft constraints.
+Soft constraints are (potentially weighted) constraints that hold only on some of the data.
+This could extend the approach to new application domains as well as provide more native error detection.
 

### 2016-05-10 Paper/ECAI/tabular.tex
Use S for templates

@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}
-  Given ${\cal G}$ a set of groups derived from fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal C}$ ; find the set of all constraints $s(G_1', ..., G_n')$ that match the syntax of a \template in ${\cal C}$ and where the groups $G_1', ...
+  Given ${\cal G}$ a set of groups derived from fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal S}$ ; find the set of all constraints $s(G_1', ..., G_n')$ that match the syntax of a \template~$s \in {\cal S}$ and where the groups $G_1', ...
 , G_n'$ with $\forall i, G_i' \subseteq G_i, G_i \in {\cal G}$ satisfy both the signature and the definition of that \template.

### 2016-05-10 Paper/ECAI/tabular.tex
Reworked definitions a bit

--- 2016-05-10 Paper/ECAI/tabular.tex
@@
 %
-\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}
-  Given ${\cal G}$ a set of groups derived from fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal C}$ ; find the set of all constraints $s(G_1', ..., G_n')$ that match the syntax of a \template in ${\cal C}$ and where the groups $G_1', ...
-, G_n'$ with $\forall i, G_i' \subseteq G_i, G_i \in {\cal G}$ satisfy both the signature and the definition of that \template.
+\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
+  Given ${\cal G}$ a set of groups over fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal S}$: find the set of all constraints $s(G_1', ..., G_n')$ that match the syntax of a \template~$s \in {\cal S}$ and where the (sub)groups $G_1', ...
+, G_n'$ with $\forall i: G_i' \subseteq G_i, G_i \in {\cal G}$ satisfy both the signature and the definition of that \template.
 \end{definition}
@@
 
-\begin{definition}[Condensed Tabular Constraint Learning]\label{def:condensed_problem_statement}
-  Let ${\cal T}$ a set of instantiated tables (all cells have values),  ${\cal C}$ be a set of {\template}s,  ${\cal G}$ be a set of groups for the tables ${\cal T}$, \dependencies be the set of all specifity rules, then find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, the groups $G_1', ...
+\begin{definition} \textit{Condensed Tabular Constraint Learning}\label{def:condensed_problem_statement}\\
+  Given a set of instantiated tables~${\cal T}$ (all cells have values), a set of {\template}s~${\cal S}$, a set of groups~${\cal G}$ over the tables~${\cal T}$, and the set of all specifity rules~\dependencies: find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, the (sub)groups $G_1', ...
 , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$ and there is no $s'$ in ${\cal C}$ such that $s'(G_h',\dots,G_k')$ holds (for $1 \leq h \leq k \leq n$) and  $s'$ is more specific than $s$ in \dependencies.

### 2016-05-10 Paper/ECAI/tabular.tex
Added end marker old text
Updated, added applications

@@
 
+\tias{End old text}
+
 In theory each constraint is independent and equally useful.
@@
 \tias{ordering is because some constraint templates use other templates in their signature, doesn't have to be more complicated?}
+\samuel{That is true - is it clear now?}
 
@@
   \includegraphics[width=0.9\linewidth]{figures/constraint_dependency.png}
-  \caption{Constraint Learning Order. The solid arrows points from more general to more specific constraints templates (e.g. \textit{all-different} is more general than \textit{permutation}).}
+  \caption{Constraint Learning Order.
+  An arrow from~$s_1$ to~$s_2$ indicates that the \CSignature of~$s_2$ requires some arguments of~$s_2$ to fulfill~$s_1$.
+  Therefore, constraints for the more general \template~$s_1$ are learned before constraints for the more specific \template~$s_2$.
+  }
   \label{fig:learning_order}
@@
 Constraint templates that have no dependencies are omitted.
-In the figure solid edges point from more general to more specific templates.
-Therefore, if there is a solid edge from constraint template~$s_1$ to template~$s_2$, then a set of subgroups satisfying~$s_2$ must contain the subgroups satisfying~$t_1$.
-
-% Dashed edges represent the inverse relationship, pointing from specific to more general constraints.
-% If there is a dashed edge from $t_1$ to $t_2$, then a set of subgroups satisfying $t_2$ may not contain a subset that also satisfies $t_1$.
 
 \paragraph{Example}
-As mentioned above \textit{foreign-key} is more specific than \textit{all-different}.
+As mentioned earlier, \textit{foreign-key} is more specific than \textit{all-different} as $\ecfkey{v_{fk}}{v_{pk}} \implies \ecalldiff{v_{pk}}$.
 Therefore, we can first generate all subgroups $G_i'$ for which \ecalldiff{G_i'} is satisfied and use these subgroups as candidates for $v_{pk}$ in \ecfkey{v_{fk}}{v_{pk}}.
 
-The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
 \\\\
@@
     & \textbf{Total} & \textbf{Average per spreadsheet} \\
-    \textbf{Essential constraints} & $93$ & $2.91$ \\
+    \textbf{Constraints} & $93$ & $2.91$ \\
     \textbf{Accuracy} & $96.77\%$ & $94.27\%$ \\
@@
 \end{table}
+\samuel{Expand summary}
 
@@
 Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.
-\sergey{Add summary table of results}
 
 \section{Applications}
+In this section we will illustrate how various motivating applications could be realized using our system.
+
+\subsection{Autocompletion}
+Autocompletion can be seen as using knowledge derived from data at time~$t_0$ and new input data at time~$t_1 > t_0$ to predict data that the user has not yet written.
+Therefore, constraints are learned at time~$t_0$ and combined with data from a new row~$i$ (column~$j$) that the user has added to a table~$T$ by time~$t_1$.
+Constraints that do not conflict with the added data are considered viable.
+Viable constraints that have enough input data to compute the result for a blank cell~$T[i,j]$ in the new row (column) are \textit{active}, which means they can be used to predict the outcome for that cell.
+Active constraints can be used to autocomplete blank cells such that the user does not have to fill in the values manually.
+Once the user has stopped editing the new row~$i$ (column~$j$), the table~$T$ is updated to include~$i$ ($j$) and constraints are learned once again.
+
+\subsection{}
 \sergey{We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.}
@@
 \end{figure*}
-
-\samuel{application autocomplete}
 

### 2016-05-10 Paper/ECAI/tabular.tex
Added skeleton error detection
rework problem statement

@@
 \newcommand{\postprocess}{\format{pruneRedundant}}
-\newcommand{\constrainttorder}{\format{generalityOrder}}
+\newcommand{\constrainttorder}{\format{templateOrder}}
 \newcommand{\template}{\format{constraint template}}
@@
 \newcommand{\ecrank}[2]{\eccalc{#1}{\mathit{RANK}(#2)}}
-\newcommand{\ecfkey}[2]{\ensuremath{#1 \rightarrow #2}}
+\newcommand{\ecfkey}[2]{\ensuremath{FOREIGNKEY(#1,#2)}}
 \newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-  Given ${\cal G}$ a set of groups over fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal S}$: find the set of all constraints $s(G_1', ..., G_n')$ that match the syntax of a \template~$s \in {\cal S}$ and where the (sub)groups $G_1', ...
-, G_n'$ with $\forall i: G_i' \subseteq G_i, G_i \in {\cal G}$ satisfy both the signature and the definition of that \template.
+  Given ${\cal G}$ a set of groups over fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal S}$: find all constraints $c_s(G'_1, ..., G'_n)$ that match the syntax of a \template~$s \in {\cal S}$ and where all subgroups $G'_i \subseteq G_i$ with $G_i \in {\cal G}$, satisfy both the signature and the definition of that \template.
 \end{definition}
-% {\bf Given }
-% \begin{enumerate}
-% \item
-% a set of instantiated tables (so all cells have values) ${\cal T}$;
-% \item
-% a set of {\template}s ${\cal C}$;
-% \item
-% a set of groups ${\cal G}$ for the tables ${\cal T}$;
-% \end{enumerate}
-% \noindent
-% {\bf Find}  the set of all constraints $s(G_1', ..., G_n')$ that 
-% \begin{enumerate}
-% \item are satisfied  in ${\cal T}$;
-% \item for which all $G_i' \subseteq G_i \in {\cal G}$ ;
-% \item the groups $G_1', ... , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$.
-% \end{enumerate}
-An example solution for the type-consistent groups of the tables in Figure~\ref{fig:main_example} and {\template}'s in Table~\ref{table:constraints} is shown in Figure~\ref{fig:sol_example}.
-
-% Let us introduce \template instantiation syntax.
-Let $S$ be a mapping from variables to the subgroups of \groups and $t$ be a \template, then $t^S$ is the template, in which all variables in all constrainst in \CSignature and \CFunction are instantiated with the corresponding subgroups from $S$.
-We refer to $t^S$ as a \textit{constraint} and to the instantiated \CSignature (\CFunction) as $\text{\CSignature}^S$ ($\text{\CFunction}^S$).
-Throughout the work we refer as \textit{find constraints} to finding all constraints that hold.
-
 % Let us now define what it means for a constraint from \template to hold in general.
@@
 
-% In the previous section we introduced the problem of tabular constraint learning informally using the example in Figure \ref{fig:main_example}.
-Here we formalize the statement in terms of \template and group assignments as follows:
-
-% \begin{minipage}[c]{14em}
-%   \vspace{5pt}
+%Here we formalize the statement in terms of \template and group assignments as follows:
 %   \begin{tabular}{ll}
 %     \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
-%     \vspace{-4pt}
-%     &\\
 %     \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
@@
 %   \end{tabular}
-%   \vspace{6pt}
-% \end{minipage}
-% \luc{To here}
 
-As is well-studied in Inductive Logic Programming, some constraints can \textit{entail} other constraints, meaning that some constraints are \textit{implied} by others.
-More formally, a constraint $c_1$ implies another constraint~$c_2$ iff whenever $c_1$ is true, $c_2$ must also be true.
-As a result, a solution to the tabular constraint learning problem may contain constraints that are implied by others.
-We can hence say that there is \textit{redundancy} in the solution set of constraints.
+An example solution for the type-consistent groups of the tables in Figure~\ref{fig:main_example} and {\template}'s in Table~\ref{table:constraints} is shown in Figure~\ref{fig:sol_example}.
 
-Two types of impliciation:
+Depending on the application, a solution to the tabular constraint learning problem can be considered to contain \textit{redundant} constraints.
+This is because constraints may be logically equivalent and constraints may also be implied by other constraints.
 
-1) on constraint definitions for same signature (example: permutation and alldiff or foreign-key and alldiff)
-
-2) on constraint signature for same definition (example: ?) \textit{Formally, we say that a constraint template $t_1$ is \textit{more specific} than a constraint template $t_2$ iff for $t_1$ and $t_2$  holds the statements of the form: $\forall g_1, \dots, g_n{:}~ t_1(g_1,\dots,g_n) \rightarrow t_2(g_h, \dots, g_k)$ where $1 \leq h \leq k \leq n$.
-The intuition behind the formula is the following, if a list of groups is a solution for $t_1$ then its subset is also a solution for $t_2$.}
-
-3) combination
-
-\tias{OLD TEXT:}
-
-A constraint template~$s_1$ is more specific than a template~$s_2$ iff a constraint applying $s_1$ to a set of subgroups is always more specific than a constraint applying $s_2$ to (part of) the same set.
-\samuel{Formalize: $\forall x1..x_k s_1(x_1...x_n) => s_2(x_1...x_k)$ :: looking for most specific}
+A constraint $c_1(G'_1, ...
+G'_n)$ \textit{implies} another constraint~$c_2(G''_1, ..., G''_m)$ where the second one ranges over a subsets of the arguments of the first, iff whenever the first constraint is true, the second must also be true.
+More formally, $\forall G'_1, ..., G'_n$ s.t.
+$\{G''_1, ..., G''_m\} \subseteq \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
+G'_n) \rightarrow c_2(G''_1, ..., G''_m)$.
+We say that the first constraint is more specific than the second.
+\tias{Some reference?}
 
 \paragraph{Example}
-Consider templates \textit{all-different} and \textit{foreign-key}.
-If \ecfkey{G_{pk}}{G_{fk}}, then \ecalldiff{G_{pk}} must also hold, therefore, \textit{all-different} is considered to be more general than \textit{foreign-key}.
-\\\\
-In ILP one often exploits these type of generality relations to structure the search for constraints using refinement operators \cite{luc_book}.
-Using such techniques leads to a more efficient traversal of the search space (of constraints).
-Aside from the efficiency gained by exploiting constraints learned earlier, the generality also allows us to build constraint templates on top of more general templates.
+A first example is $permutation(G') \rightarrow alldifferent(G')$ (it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G'$ is a vector).
+In this case the arguments are the same in both constraints.
+For most applications, it will be sufficient to output only the most specific one.
+A more involved example is that a look-up constraint implies a foreign-key constraint.
+The syntax of the look-up constraint is $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$, where the values in the vector $G'_2$ are determined by looking up the corresponding value in vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
+In Figure~\ref{fig:main_example} we have $\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}$.
+This implies a foreign-key constraint $\ecfkey{G'_1}{H'_1}$ where each of the values in vector $G'_1$ must appear in $H'_1$ for the look-up to work.
 
-\tias{End old text}
+Two constraints are \textit{equivalent} if they imply each other.
+More formally, $\forall G'_1, ..., G'_n$ s.t.
+$\{G''_1, ..., G''_n\} = \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
+G'_n) \leftrightarrow c_2(G''_1, ..., G''_n)$.
+An example within the same \template is for the sum over vectors constraint: $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
+A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = SUM(H)$ with $size(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
 
-In theory each constraint is independent and equally useful.
-In practice, however, things are different and some constraints are considered to be \textit{redundant}.
-Assume that for some $g_x$ both \textit{series($g_x$)} and \textit{alldifferent($g_x$)} hold.
-From a user's perspective the last solution is useless, since he or she already knows that the series constraint implies the alldifferent constraint, or formally speaking $\forall g_x{:}~\textit{series}(g_x) \rightarrow \textit{alldifferent}(g_x)$  holds.
-To capture this condensed representation \cite{condensed} of solutions, we have introduced a specificity order between constraints, which is depicted using solid arrows in Figure \ref{fig:learning_order}.
-This matters from implementation perspective as well, if we discovered that $g_x$ is not a permutation, then $g_x$ should not be tested for the series constraint at all.
+A good way to deal with equivalent constraints is to identify in a generic way a \textit{canonical} constraint among the equivalent ones.
+For example, for equivalent constraints of the same \template, the constraint where the groups are ordered by table, orientation and row/column identifier.
+For equivalent constraints of different templates one could define heuristics such as that some templates are preferred over others, e.g.
+binary sum over n-ary sum, sum over difference, etc.
 
-Formally, we say that a constraint template $t_1$ is \textit{more specific} than a constraint template $t_2$ iff for $t_1$ and $t_2$  holds the statements of the form: $\forall g_1, \dots, g_n{:}~ t_1(g_1,\dots,g_n) \rightarrow t_2(g_h, \dots, g_k)$ where $1 \leq h \leq k \leq n$.
-The intuition behind the formula is the following, if a list of groups is a solution for $t_1$ then its subset is also a solution for $t_2$.
+\tias{I am not defining a new problem for 'most specific canonical tabular constraint learning' because we loosely showed how implications and equivalences would be expressed, but not how to express canonicity and we did not claim that most specific is always best.}
+\tias{Note also that our method finds ALL constraints; it uses specificity to faster traverse the search space, but the resulting set of constraints found would be the same, whether or not using the specificity (see 4.2.1).}
 
-Let us introduce the version of tabular constraint learning by modifying Definition \ref{def:problem_statement} that takes into account the specificity relation between constraint templates.
-
-\begin{definition} \textit{Condensed Tabular Constraint Learning}\label{def:condensed_problem_statement}\\
-  Given a set of instantiated tables~${\cal T}$ (all cells have values), a set of {\template}s~${\cal S}$, a set of groups~${\cal G}$ over the tables~${\cal T}$, and the set of all specifity rules~\dependencies: find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, the (sub)groups $G_1', ...
+%Let us introduce the version of tabular constraint learning by modifying Definition \ref{def:problem_statement} that takes into account the specificity relation between constraint templates.
+%
+%\begin{definition} \textit{Condensed Tabular Constraint Learning}\label{def:condensed_problem_statement}\\
+%  Given a set of instantiated tables~${\cal T}$ (all cells have values), a set of {\template}s~${\cal S}$, a set of groups~${\cal G}$ over the tables~${\cal T}$, and the set of all specifity rules~\dependencies: find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, the (sub)groups $G_1', ...
 , G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$ and there is no $s'$ in ${\cal C}$ such that $s'(G_h',\dots,G_k')$ holds (for $1 \leq h \leq k \leq n$) and  $s'$ is more specific than $s$ in \dependencies.
-\end{definition}
+%\end{definition}
 
@@
 
-This version explicitly takes into account the specificity of the constraints and invalidates some of the solutions that are entailed by the tabular other solutions.
-
-\sergey{still need to remove some repetitions later on about specisifity}
+%This version explicitly takes into account the specificity of the constraints and invalidates some of the solutions that are entailed by the tabular other solutions.
 
 \newcommand{\tcl}{Tabular Constraint Learning}
-\newcommand{\ctl}{Condensed Tabular Constraint Learning}
+%\newcommand{\ctl}{Condensed Tabular Constraint Learning}
 \section{Approach to Tabular Constraint Learning}
-% \sergey{Luc wants this to be more intuitive, should we just add some examples here to make it easier to follow and refer to the running example?}
-This section will describe our approach to \tcl.
-
-The key observation is that essentially the problem is a constraint enumeration problem, where each constraint is independent.
-This property comes from fact that in spreadsheets each formula is applied based on the existing data in the cells.
-This allows learning constraints independently of each other by examining constraint satisfaction on the groups.
+This section describes our approach to \tcl.
 
@@
 Our algorithm is illustrated in Algorithm~\ref{algo:tcl} and contains three key steps.
-First, constraint templates are ordered according to the generality ordering such that they can be treated sequentially.
+First, constraint templates are ordered such that they can be treated sequentially and each only once.
 Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
@@
 
-\subsubsection{Generality ordering}
-\tias{ordering is because some constraint templates use other templates in their signature, doesn't have to be more complicated?}
-\samuel{That is true - is it clear now?}
+\subsubsection{Template ordering}
+Looking at the template signatures in Table~\ref{table:constraints} one can see that some signatures depend on other constraints.
+This dependency also translates into the template implying the other templace.
+In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
+Similarly, we can use these implications to define an ordering over the templates, such that the more general constraints (the requirements) are verified before the more specific constraints.
+Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}.
+Constraint templates that have no dependencies are omitted. Because there are no equivalences or loops in the signatures, the resulting graph is a directed acyclic graph (DAG).
 
@@
 
-$\constrainttorder(\constraints,\dependencies)$ uses the DAG~\dependencies in Figure \ref{fig:learning_order} to generate all constraint templates according to the topological order of~\dependencies.
-Constraint templates that have no dependencies are omitted.
+$\constrainttorder(\constraints,\dependencies)$ uses such a DAG over a subset of the constraint templates to derive a topological order of~\dependencies.
 
@@
 \begin{table*}
+\tias{Notation inconsistent (no $G'$?
+also the 'bold' notation not explained.
+Also, in many places missing that must be vector...
+switched $\rightarrow$ to FOREIGNKEY to avoid confusion with implied constraints.}
   \centering
@@
 
-\subsection{}
+\subsection{Error detection}
+For this application we consider a use case concerning crime statistics provided by the FBI.
+
 \sergey{We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.}

### 2016-05-11 Paper/ECAI/tabular.tex
some tweaks to problem statement

@@
 A more involved example is that a look-up constraint implies a foreign-key constraint.
-The syntax of the look-up constraint is $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$, where the values in the vector $G'_2$ are determined by looking up the corresponding value in vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
+The syntax of the look-up constraint is $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$, where the values in the vector $G'_2$ are determined by looking up the corresponding value of vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
 In Figure~\ref{fig:main_example} we have $\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}$.
@@
 
-Two constraints are \textit{equivalent} if they imply each other.
+Another form of redundancy is because of \textit{equivalent} constraints.
+Two constraints are equivalent if they imply each other.
 More formally, $\forall G'_1, ..., G'_n$ s.t.
 $\{G''_1, ..., G''_n\} = \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
-G'_n) \leftrightarrow c_2(G''_1, ..., G''_n)$.
+G'_n)$ $\leftrightarrow c_2(G''_1, ..., G''_n)$.
 An example within the same \template is for the sum over vectors constraint: $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
@@
 binary sum over n-ary sum, sum over difference, etc.
+
+Because dealing with redundancy is often application-dependent, we explain in the next section our generic method for finding all constraints.
 

### 2016-05-11 Paper/ECAI/tabular.tex
inbetween commit with personal comments

--- 2016-05-10 Paper/ECAI/tabular.tex
@@
 \newcommand{\ecrank}[2]{\eccalc{#1}{\mathit{RANK}(#2)}}
-\newcommand{\ecfkey}[2]{\ensuremath{FOREIGNKEY(#1,#2)}}
+\newcommand{\ecfkey}[2]{\ensuremath{\mathit{FOREIGNKEY}(#1,#2)}}
 \newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
@@
 Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
-Constraint satisfaction techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
-Therefore, new types of constraints can be easily added to the system by specifying them using declarative methods such as Minizinc or ASP.
+Constraint learning techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
+New types of constraints can be easily added to the system by specifying them using in a declarative language.
 We show that our approach is able to accurately discover constraints in a number of spreadsheets from varying sources.
@@
 \section{Introduction}
-Millions of people across the world use spreadsheets everyday.
+Millions of people across the world use spreadsheets every day.
 The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
-However, large sheets (possible with multiple tables) and relations between different tables and sheets can be tedious and error-prone to handle.
-As a result, many people lack a complete understanding of the structures and existing dependencies in the data.
+However, large and complex sheets (possibly with multiple tables and relations between different them) can be hard to handle.
+Many end-users lack the understanding of the underlying structures and dependencies in such sheets and the data they contain.
+Especially when spreadsheets have been exported from
+other software such as ERP packages, it is hard to understand the data as in this case,
+often a comma-separated values (CSV) format is used, meaning that all formula's are lost, including inter-sheet formulas and relations.
+%This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
 
-This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages.
-In this case, often a comma-separated values (CSV) format is used, meaning that all formula's are lost, including inter-sheet formulas and relations.
-
-In this work, we investigate in how far machine learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
-However, this is an unconventional machine learning problem.
+In this paper,  we investigate whether machine learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
+This is a new and unconventional machine learning problem.
 Consider the example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of spreadsheet operations such as \textit{average} or \textit{sum}.
@@
 Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
-This already provides a flavor of how this problem is different from standard data mining settings, when the data is just in rows and variables are in columns.
+This already provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns.
 Here everything is mixed.
 The data is relational on the one hand, since we have multiple table with relationships between them.
-And on the other hand the data is mixed textual and numeric, since people tend to use spreadsheets for financial and accounting computations.
-One can see that understanding and learning constraints in tabular data is hence a new and challenging problem setting.
-%Furthermore, it is not only challenging for a machine but for a human as well.
+And on the other hand, the data is mixed textual and numeric, since people tend to use spreadsheets for financial and accounting computations.
+One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
+Furthermore, it is not only challenging for a machine but for humans as well.
 
-Relatedly, the field of Inductive Logic Programming has long...
+The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
+\cite{claudien,lallouet} have studied the discovery of clausal constraints in an inductive logic programming setting, from constraint learning,  where \cite{Quacq,Conacq,modelseeker} have developed several approaches to learn conjunctive constraints in a constraint programming setting,
+and from mining in databases,  where e.g.
+\cite{Shavlik} has applied inductive techniques to discover constraints (such as functional and multi-valued dependencies) in relational databases.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from
+very few examples.  Our approach contrasts with all of the mentioned approaches (except Modelseeker) in that it focusses
+on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
+A more detailed discussion on related work is contained in Section \ref{sec:related_work}.
+
+%the field of Inductive Logic Programming has long...
 \tias{TODO for Sergey or Luc}.
@@
 
-Another line of work from which we draw inspiration is that of \tias{TODO for ?, flashfill et al}.
+%Another line of work from which we draw inspiration is that of \tias{TODO for ?, flashfill et al}.
 We will consider the case where all data is given at once.
@@
 
-The question that we hence try to answer in this work is: can we discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
-Moreover, we investigate general-purpose ways for doing so, that do not require hard coding a number of constraints.
-Rather, we use techniques from constraint solving to derive whether and which constraints are satisfied.
-This approach can be extended to any constraint in a principled way.
+The question that we answer in this paper is: whether it is possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose algorithm and system for discovering such constraints.
+Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias") and covers both textual and numeric fields.
+It works in an unsupervised setting, as only the tabular data is provided and no examples need to be specified, and discovers column- and row-wise constraints.
+We demonstrate the utility of our approach in an experimental evaluation and we also sketch several application scenario's.
+\luc{todo:mention these here too}
+
+This paper is organised as follows: \luc{todo}
 
@@
 \section{Formalization}
-In this work we focus on spreadsheet data.
-We will first introduce the general concepts of cells, group and constraint templates, using the (multi-table) sheet in Figure~\ref{fig:main_example} as running example.
+As we focus on spreadsheet data, we first introduce the general concepts of cells, group and constraint templates, using the (multi-table) sheet in Figure~\ref{fig:main_example} as running example.
 
 \subsection{Cells, vectors and groups}
-A spreadsheet is physically one big table, but conceptually may consist of multiple tables, such as in Figure~\ref{fig:main_example}.
+A spreadsheet may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
 We will describe a method of extracting tables from a big spreadsheet in Section~\ref{todo}, however this is not the core of our learning method.
@@
   \item
-  A \textbf{subgroup} $G'$ of $G$ is simply a projection of $G$ onto a consecutive subset its vectors.
+  A \textbf{subgroup} $G'$ of a group $G$ is simply a projection of $G$ onto a consecutive subset its vectors.
 % is a subrange of vectors belonging to a group $G$.
@@
 \paragraph{Example}
-Consider Table~1 in the Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
+Consider Table~1 in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
 Ignoring the (all-textual) header, Table~1 can be divided into the following five column-oriented groups:
@@
 \item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy. That is, the required types of the arguments, together
-with possible further restrictions such as on their length.
-For example, the signature can restrict the types of the arguments to be groups integers and require that the length of all arguments be equal.
+with possible further restrictions, e.g., on their length.
+For example, the signature can restrict the types of the arguments to be groups of integers and require that the length of all arguments be equal.
 \\
-In terms of logical and relational learning, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
+In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
 
 \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
-Given an assignments to all its arguments, it can verify whether the constraint is satisfied or not.
+Given an assignments to its arguments, it can be used to verify whether the constraint is satisfied or not.
 %holds, in practice this will be a function that can be called when
@@
       $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
-we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table \tias{Of the right (minimum) dimension??
-Note that this is also sloppy because the ':8' is not considered part of the variable...},
-or we can interpret the argument as the value it takes in a particular table $T$ such as $T_1$ in Figure 1, that is, $[1,2,3,4]$.
-This also means that we can talk about constraints being satisfied. Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
+we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ of the right dimensions.
+So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
+
+\tias{Of the right (minimum) dimension?? Note that this is also sloppy because the ':8' is not considered part of the variable...},
+Alternatively, and perhaps more naturally, we can interpret the argument as the value it takes in a particular table $T$ such as $T_1$ in Figure 1, that is,
+${\range{T_1}{\rangeall}{8}} =  [1,2,3,4]$.
+Then we can also talk about constraints being satisfied. Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
 if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
@@
 
-The inverse problem is, given an assignment to all cells, to find the constraints that are present in the spreadsheet.
+The inverse problem is, given an assignment to all cells, to find the constraints that are satisfied in the spreadsheet.
 We define the {\bf Tabular Constraint Learning Problem} as follows:
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-  Given ${\cal G}$ a set of groups over fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal S}$: find all constraints $c_s(G'_1, ..., G'_n)$ that match the syntax of a \template~$s \in {\cal S}$ and where all subgroups $G'_i \subseteq G_i$ with $G_i \in {\cal G}$, satisfy both the signature and the definition of that \template.
+{\bf   Given} ${\cal G}$ a set of groups over fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal S}$: {\bf find } all constraints $c_s(G'_1, ..., G'_n)$ that match the syntax of a \template~$s \in {\cal S}$ and where all subgroups $G'_i \subseteq G_i$ with $G_i \in {\cal G}$, satisfy both the signature and the definition of that \template.
 \end{definition}
@@
 
-Depending on the application, a solution to the tabular constraint learning problem can be considered to contain \textit{redundant} constraints.
-This is because constraints may be logically equivalent and constraints may also be implied by other constraints.
+Depending on the application, a solution to the tabular constraint learning problem may contain \textit{redundant} constraints, cf.
+\cite{luc_book}.
+This is because constraints may be logically equivalent or may be implied by other constraints.
 
-A constraint $c_1(G'_1, ...
-G'_n)$ \textit{implies} another constraint~$c_2(G''_1, ..., G''_m)$ where the second one ranges over a subsets of the arguments of the first, iff whenever the first constraint is true, the second must also be true.
+More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
+the  following statement holds: $$\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m).$$
+
+We shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
+To simplify notation, and in line with logic programming, we shall often make the quantification implicit and simply write this as rules $ c_1(G_1, ...
+G_n) \rightarrow c_2(G'_1, ...
+G'_m)$.
+We say that the first constraint is more specific than the second, and the second is more general.
+
+% where the second one ranges over a subsets of the arguments of the first, iff whenever the first constraint is true, the second must also be true.
 More formally, $\forall G'_1, ..., G'_n$ s.t.
 $\{G''_1, ..., G''_m\} \subseteq \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
-G'_n) \rightarrow c_2(G''_1, ..., G''_m)$.
-We say that the first constraint is more specific than the second.
-\tias{Some reference?}
+G'_n) \rightarrow c_2(G''_1, ..., G''_m)$ll.
 
 \paragraph{Example}
-A first example is $permutation(G') \rightarrow alldifferent(G')$ (it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G'$ is a vector).
-In this case the arguments are the same in both constraints.
-For most applications, it will be sufficient to output only the most specific one.
-A more involved example is that a look-up constraint implies a foreign-key constraint.
-The syntax of the look-up constraint is $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$, where the values in the vector $G'_2$ are determined by looking up the corresponding value in vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
+Consider the following implication:
+$$\ecperm{G} \rightarrow \ecalldiff{G}$$
+(it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G$ is a vector).
+In this case the arguments are identical in both constraints.
+Given that the problem is to find the set of {\em all} constraints, and that the more general constraint is entailed
+by the more specific one, the $ALLDDIFFERENT(G)$ constraint will be {\em redundant}.
+So, for most applications, it will be sufficient to output only the most specific one.
+
+A more involved example is
+$$\eclookup{G'_2}{G'_1}{H'_1}{H'_2} \rightarrow \ecfkey{G'_1}{H'_1}$$
+which states
+that a look-up constraint implies a foreign-key constraint.
+In the look-up constraint $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ the values in the vector $G'_2$ are determined by looking up the corresponding value in vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
 In Figure~\ref{fig:main_example} we have $\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}$.
@@
 
-Two constraints are \textit{equivalent} if they imply each other.
-More formally, $\forall G'_1, ..., G'_n$ s.t.
+Two constraints are \textit{equivalent} if they imply one another.
+%More formally, $\forall G'_1, ..., G'_n$ s.t.
 $\{G''_1, ..., G''_n\} = \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
@@
 An example within the same \template is for the sum over vectors constraint: $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = SUM(H)$ with $size(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
+A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = \textit{SUM}(H)$ with $\textit{size}(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
+\luc{Do we really need this, or can we write it in a real equivalence --- conjunctive formula?}
 
-A good way to deal with equivalent constraints is to identify in a generic way a \textit{canonical} constraint among the equivalent ones.
-For example, for equivalent constraints of the same \template, the constraint where the groups are ordered by table, orientation and row/column identifier.
+A good way to deal with equivalent constraints is to work with a \textit{canonical} constraint among the equivalent ones.
+For example, for equivalent constraints of the same \template, this could be the constraint where the groups are ordered by table, orientation and row/column identifier.
 For equivalent constraints of different templates one could define heuristics such as that some templates are preferred over others, e.g.
 binary sum over n-ary sum, sum over difference, etc.
+
+Because dealing with redundancy is often application-dependent, we explain in the next section our generic method for finding all constraints.
 
@@
 
+\tias{Make general intro of subsections}
+
 Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
@@
 \subsection{Table and group detection}
+\tias{0. preprocess: currencies and precentual to numeric, 1. tables from spreadsheets, 2. groups from tables. Ext: could be done interactively.}
+\tias{Mention spreadsheet-specific hacks in the exps, not here}
 \samuel{Is an example needed? I find no because it is not our main concern but open for discussion}
@@
 We suggest to provide tables in general, for well formated data, however, the user will not have to specify tables manually.
+\tias{Samuel, what does this mean?}
 
@@
 \subsection{Algorithm}
+\tias{1) all templates, order so only once; 2) generate groups that satisfy signature; 3) generate subgroups that satisfy definition}
 Our algorithm is illustrated in Algorithm~\ref{algo:tcl} and contains three key steps.
@@
   \For{$t \in \constrainttorder(\constraints,\dependencies)$}
-  \State $v_1, ..., v_n =$ variables of $T$
+  \State $v_1, ..., v_n =$ variables of $T$ \tias{variables not defined, only nr of groups important?}
     \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(t, \groups, C, \dependencies)$}
@@
 We will explain and illustrate the three key steps in detail.
+
+We now describe each of the three steps in more detail.
 
@@
 \\\\
-Most parts of the algorithm could be easily parallelized.
-Every connected component in \dependencies can be learned independently.
-Within every component, constraint templates have to be treated respecting the order, i.e. learning the most general ones first.
+Constraint templates have to be treated respecting the order, i.e. learning the most general ones first.
 The generality ordering thus imposes a partial ordering according to which constraint templates must be learned.
@@
 \subsubsection{Candidate group generation}
+\tias{find groups, each $n$ groups and $m$ arguments than m-choose-n candidates.
+instead of generate-and-test we will search for all valid groups.
+To do so, CSP with signature in logical constraints and properties of each group as data.
+Example should use constraint syntax}
 $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups (group-assignments~$A$) that may contains subgroups satisfying the constraint.
@@
 For example, the \CSignature requires $v_r$ to be integer.
-If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this this group cannot be discarted.
+If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this this group cannot be discarded.
 
-For some constraints (e.g. sum, max, count) that share the same \CSignature the same set of assignments~$A$ will be found.
+For some constraints (e.g. sum, max, count) that share the same \CSignature the same set of assignments~$A$ will be found. \tias{so?}
 
@@
 
-\subsubsection{Subgroup satisfaction search}
-$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
+\subsubsection{Satisfying subgroups}
+\tias{not set of assignment but single assignment $a$?}
+\tias{requires checking for which of the subgroups of the groups the constraint is satisfied.
+many subgroups of the groups will not satisfy signature, hence no generate-and-test.
+Instead, we use again satisfiability to both search for the subgroups and verify satisfiab of the sign and definition.}
+$\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
 the \CSignature and \CFunction of~$t$ are satisfied).
@@
 For some constraint templates (most notably aggregates) there may be multiple (overlapping) subgroups of a given group~$G$ for which the template satisfied.
+\tias{like min/max? should mention this example in implied constraints. Not sure we should mention this here, how do we remove overlapping, in post-processing? if so, move to later...}
 In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
@@
 \subsection{Implementation note}
+
+\paragraph{Constraint solvers}
+\tias{stand-alone (with text-based modeling language) versus embedded.}
 For each \template, both \generategroups and \findassignment are can be seen as separate constraint satisfaction problems.
-Hence, it is natural to solve them using declarative solvers.
-Even though each of these problems differ from each other, they are variations that can be modeled using similar techniques.
-This also makes it easy to extend the implementation as templates can be easily added (or removed).
 Furthermore, declarative languages such as, for example, ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
@@
 
-\subsection{Constraints}
+\paragraph{Parallelization}
+\tias{Do we do this? If not, cut...}
+The two for loops in the algorithm can be easily parallelized.
+Each generated assignment can be checked independently.
+Templates may not be independent as expressed by the DAG \dependencies.
+However, each independent component of the DAG (including one component per template not in the DAG) can be verified independently.
+
+\paragraph{Redundancy}
+Mention 'largest overlapping subgroup' here?
+
+\paragraph{Constraints}
 
@@
 
-Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty. \samuel{More on this / example?}
+Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty.
+\samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
+
+\section{Examples}
+\subsection{Group generation CSP}
+variables from syntax, domain from input groups, filtered using type
+
+Group generation CSP (Sum Product)\\
+variables: $R, O_1, O_2$\\
+domains: $D(R) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_1) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_2) = \{ G | G \in {\cal G} \land \numeric(G)\}$\\
+constraints: $\plength(O_1) = \plength(O_2) \geq 2, \pcols(R) = 1, \prows(R) = 1$ % excluded not partial because we don't focus on it much
+
+Translation:
+\begin{description}
+  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
+  \item[Length] Require group to have exact length (subgroups have the same length)
+  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
+  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
+  \item[Same length / table / orientation] Require groups to gave the same length table ...
+\end{description}
+
+\subsection{Subgroup generation}
+Sum Product
+
+Implementation: finds subgroups based on assignments\\
+Method $Test(G_R, G_{O_1}, G_{O_2})$\\
+$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
+$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
+\\\\
+Method $FindSubgroups(Assignments)$\\
+$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$
+
+Register implementation with the system for constraint X
 
@@
 Redundant constraints can be reduced in different ways.
-Duplications should be detected in a post-processing step that uses entailment or heursitics to filter constraints.
+Duplications should be detected in a post-processing step that uses entailment or heuristics to filter constraints.
 Accidental constraints may be detected through heuristics, however, they can be also be removed by adding additional data.
@@
 
-\section{Related Work}
+\section{Related Work}\label{sec:related_work}
 \sergey{key bullet points for Luc and possibly Samuel and me to make related work section}

### 2016-05-11 Paper/ECAI/tabular.tex
Added Luc's changes

@@
 \newcommand{\ecrank}[2]{\eccalc{#1}{\mathit{RANK}(#2)}}
-\newcommand{\ecfkey}[2]{\ensuremath{\mathit{FOREIGNKEY}(#1,#2)}}
+\newcommand{\ecfkey}[2]{\ensuremath{FOREIGNKEY(#1,#2)}}
 \newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
@@
 
-The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
-\cite{claudien,lallouet} have studied the discovery of clausal constraints in an inductive logic programming setting, from constraint learning,  where \cite{Quacq,Conacq,modelseeker} have developed several approaches to learn conjunctive constraints in a constraint programming setting,
+The approach that we will introduce will borrow techniques from logical and relational learning \cite{LucBook}, where
+\cite{Claudien,todo:lallouet} have studied the discovery of clausal constraints in an inductive logic programming setting, from constraint learning,  where \cite{Quacq,Conacq,Modelseeker} have developed several approaches to learn conjunctive constraints in a constraint programming setting,
 and from mining in databases,  where e.g.
-\cite{Shavlik} has applied inductive techniques to discover constraints (such as functional and multi-valued dependencies) in relational databases.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from
+\cite{Savlik} has applied inductive techniques to discover constraints (such as functional and multi-valued dependencies) in relational databases.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{}, where the definition of function is learned in spreadsheet data from
 very few examples.  Our approach contrasts with all of the mentioned approaches (except Modelseeker) in that it focusses
 on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
-A more detailed discussion on related work is contained in Section \ref{sec:related_work}.
+A more detailed discussion on related work is contained in Section \luc{TODO}.
 
@@
 \\
-In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
+In terms of logical and relational learning \cite{bookLuc}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
 
@@
 Depending on the application, a solution to the tabular constraint learning problem may contain \textit{redundant} constraints, cf.
-\cite{luc_book}.
+\cite{bookLuc}.
 This is because constraints may be logically equivalent or may be implied by other constraints.
@@
 More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
-the  following statement holds: $$\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m).$$
-
-We shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
+the statement $\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m)$ holds. To simplify testing,
+we shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
 To simplify notation, and in line with logic programming, we shall often make the quantification implicit and simply write this as rules $ c_1(G_1, ...
@@
 Consider the following implication:
-$$\ecperm{G} \rightarrow \ecalldiff{G}$$
+$$PERMUTATION(G) \rightarrow ALLDIFFERENT(G)$$
 (it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G$ is a vector).
@@
 An example within the same \template is for the sum over vectors constraint: $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = \textit{SUM}(H)$ with $\textit{size}(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
+A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = SUM(H)$ with $size(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
 \luc{Do we really need this, or can we write it in a real equivalence --- conjunctive formula?}
@@
 
-\tias{Make general intro of subsections}
-
 Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
@@
 \subsection{Table and group detection}
-\tias{0. preprocess: currencies and precentual to numeric, 1. tables from spreadsheets, 2. groups from tables. Ext: could be done interactively.}
-\tias{Mention spreadsheet-specific hacks in the exps, not here}
 \samuel{Is an example needed? I find no because it is not our main concern but open for discussion}
@@
 We suggest to provide tables in general, for well formated data, however, the user will not have to specify tables manually.
-\tias{Samuel, what does this mean?}
 
@@
 \subsection{Algorithm}
-\tias{1) all templates, order so only once; 2) generate groups that satisfy signature; 3) generate subgroups that satisfy definition}
 Our algorithm is illustrated in Algorithm~\ref{algo:tcl} and contains three key steps.
@@
   \For{$t \in \constrainttorder(\constraints,\dependencies)$}
-  \State $v_1, ..., v_n =$ variables of $T$ \tias{variables not defined, only nr of groups important?}
+  \State $v_1, ..., v_n =$ variables of $T$
     \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(t, \groups, C, \dependencies)$}
@@
 We will explain and illustrate the three key steps in detail.
-
-We now describe each of the three steps in more detail.
 
@@
 \\\\
-Constraint templates have to be treated respecting the order, i.e. learning the most general ones first.
+Most parts of the algorithm could be easily parallelized.
+Every connected component in \dependencies can be learned independently.
+Within every component, constraint templates have to be treated respecting the order, i.e. learning the most general ones first.
 The generality ordering thus imposes a partial ordering according to which constraint templates must be learned.
@@
 \subsubsection{Candidate group generation}
-\tias{find groups, each $n$ groups and $m$ arguments than m-choose-n candidates.
-instead of generate-and-test we will search for all valid groups.
-To do so, CSP with signature in logical constraints and properties of each group as data.
-Example should use constraint syntax}
 $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups (group-assignments~$A$) that may contains subgroups satisfying the constraint.
@@
 For example, the \CSignature requires $v_r$ to be integer.
-If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this this group cannot be discarded.
+If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this this group cannot be discarted.
 
-For some constraints (e.g. sum, max, count) that share the same \CSignature the same set of assignments~$A$ will be found. \tias{so?}
+For some constraints (e.g. sum, max, count) that share the same \CSignature the same set of assignments~$A$ will be found.
 
@@
 
-\subsubsection{Satisfying subgroups}
-\tias{not set of assignment but single assignment $a$?}
-\tias{requires checking for which of the subgroups of the groups the constraint is satisfied.
-many subgroups of the groups will not satisfy signature, hence no generate-and-test.
-Instead, we use again satisfiability to both search for the subgroups and verify satisfiab of the sign and definition.}
-$\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
+\subsubsection{Subgroup satisfaction search}
+$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
 the \CSignature and \CFunction of~$t$ are satisfied).
@@
 For some constraint templates (most notably aggregates) there may be multiple (overlapping) subgroups of a given group~$G$ for which the template satisfied.
-\tias{like min/max? should mention this example in implied constraints. Not sure we should mention this here, how do we remove overlapping, in post-processing? if so, move to later...}
 In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
@@
 \subsection{Implementation note}
-
-\paragraph{Constraint solvers}
-\tias{stand-alone (with text-based modeling language) versus embedded.}
 For each \template, both \generategroups and \findassignment are can be seen as separate constraint satisfaction problems.
+Hence, it is natural to solve them using declarative solvers.
+Even though each of these problems differ from each other, they are variations that can be modeled using similar techniques.
+This also makes it easy to extend the implementation as templates can be easily added (or removed).
 Furthermore, declarative languages such as, for example, ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
@@
 
-\paragraph{Parallelization}
-\tias{Do we do this? If not, cut...}
-The two for loops in the algorithm can be easily parallelized.
-Each generated assignment can be checked independently.
-Templates may not be independent as expressed by the DAG \dependencies.
-However, each independent component of the DAG (including one component per template not in the DAG) can be verified independently.
-
-\paragraph{Redundancy}
-Mention 'largest overlapping subgroup' here?
-
-\paragraph{Constraints}
+\subsection{Constraints}
 
@@
 
-Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty.
-\samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
-
-\section{Examples}
-\subsection{Group generation CSP}
-variables from syntax, domain from input groups, filtered using type
-
-Group generation CSP (Sum Product)\\
-variables: $R, O_1, O_2$\\
-domains: $D(R) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_1) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_2) = \{ G | G \in {\cal G} \land \numeric(G)\}$\\
-constraints: $\plength(O_1) = \plength(O_2) \geq 2, \pcols(R) = 1, \prows(R) = 1$ % excluded not partial because we don't focus on it much
-
-Translation:
-\begin{description}
-  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
-  \item[Length] Require group to have exact length (subgroups have the same length)
-  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
-  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
-  \item[Same length / table / orientation] Require groups to gave the same length table ...
-\end{description}
-
-\subsection{Subgroup generation}
-Sum Product
-
-Implementation: finds subgroups based on assignments\\
-Method $Test(G_R, G_{O_1}, G_{O_2})$\\
-$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
-$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
-\\\\
-Method $FindSubgroups(Assignments)$\\
-$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$
-
-Register implementation with the system for constraint X
+Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty. \samuel{More on this / example?}
 
@@
 Redundant constraints can be reduced in different ways.
-Duplications should be detected in a post-processing step that uses entailment or heuristics to filter constraints.
+Duplications should be detected in a post-processing step that uses entailment or heursitics to filter constraints.
 Accidental constraints may be detected through heuristics, however, they can be also be removed by adding additional data.
@@
 
-\section{Related Work}\label{sec:related_work}
+\section{Related Work}
 \sergey{key bullet points for Luc and possibly Samuel and me to make related work section}

### 2016-05-11 Paper/ECAI/tabular.tex
4.1 table and group detection

--- 2016-05-11 Paper/ECAI/tabular.tex
@@
 This section describes our approach to \tcl.
+The main method takes \textit{groups} and {\template}s as input, and is described in Section~\ref{sec:algo}.
+We first explain how tables and groups can be extracted from raw spreadsheet data.
 
-\tias{Make general intro of subsections}
+\subsection{Table and group detection} \label{sec:make_groups}
+%A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) groups.
+Many spreadsheets contain headers that provide hints at where the table(s) in the sheet are and what the meaning of the rows and columns is.
+However, detecting tables and headers is a complex problem on its own~\cite{header}.
+Furthermore, headers may be missing and their use often requires incorporating language-specific vocabulary, e.g.
+in English.
 
-Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
-Often, however, this not the case and a possible approach is to extract the groups from the CSV file directly.
-This requires discovering \textit{tables} and \textit{groups}.
-Our implementation offers automated procedures for both steps.
+%\tias{0. preprocess: currencies and precentual to numeric, 1. tables from spreadsheets, 2. groups from tables. Ext: could be done interactively.}
+Fully automatic table and header detection is not the core of our work. We hence built a visual tool that uses a simple method to detect tables in the data and allows a user to correct these.
 
-\subsection{Table and group detection}
-\tias{0. preprocess: currencies and precentual to numeric, 1. tables from spreadsheets, 2. groups from tables. Ext: could be done interactively.}
-\tias{Mention spreadsheet-specific hacks in the exps, not here}
-\samuel{Is an example needed? I find no because it is not our main concern but open for discussion}
-We chose to implement a naive approach to extract tables as the detection of tables and headers from data is a complex task on its own (\cite{header}) and a sophisticated implementation goes beyond the scope of this paper.
-Our approach first detects rectangular collections of data.
-It then attempts to identify the header by looking at the type of the first row.
-If this row consists only of textual entries, the row is expected to be the header and is removed, if not, the same procedure is applied on the first column.
-We suggest to provide tables in general, for well formated data, however, the user will not have to specify tables manually.
-\tias{Samuel, what does this mean?}
+The way the table detection in the tool works is as follows:
+First, we preprocess the spreadsheets so that currency values are transformed into their integer or float value and percentual data is transformed into the corresponding floating point representation (e.g.
+$85\%$ as $0.85$).
+Next, every row in the spreadsheet is partitioned into blocks of columns that are separate by at least one $None$ value.
+Then, the blocks in each row are greedily merged to obtain rectangles.
+Finally headers are removed by checking whether the first row or column is all textual while the next row or column is not; in that case the assumed header is removed.
 
-Our approach for detecting groups works by partitioning the data of a table into consecutive, type consistent groups, essentially separating textual and numerical data.
-The assumption is that often numerical and textual data are not mixed and this approach works well in practice as often groups do not have to be specified manually\footnote{\samuel{Check on benchmark}}.
-Unless an explicit orientation (rows or columns) is provided for the table, this approach will attempt to create both row- and column-groups.
+It should be clear this simple method only works if all tables are clearly separated by $None$ values and that headers of fully textual tables are not detected, among other things.
+Hence, in the visual tool the user can correct the detected tables if needed.
+\tias{I think it would be good if we could show a screenshot of this tool...}
 
-So far we have considered textual and numerical (floating point or integer) data.
-Our implementation also supports some extensions such as currencies and percentual data, which is considered to be numerical data.
-Moreover, percentual data is corrected to the floating point representation (i.e. divided by 100).
+From each table, row-oriented and column-oriented groups are then extracted by partitioning the rows (columns) into type-consistent parts. This step can also be verified by a user in the visual tool.
 
-A more elaborate implementation could allow group (and table) detection to be done interactively / visually such that a user can select the groups in a spreadsheet or correct automatically generated groups.
 From here on we will assume that groups and tables are given.
@@
 \newcommand{\temps}{\ensuremath{S}}
-
-\subsection{Algorithm}
+\subsection{Algorithm} \label{sec:algo}
 \tias{1) all templates, order so only once; 2) generate groups that satisfy signature; 3) generate subgroups that satisfy definition}
@@
 Instead, we use again satisfiability to both search for the subgroups and verify satisfiab of the sign and definition.}
+\tias{can implement any way}
 $\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.

### 2016-05-11 Paper/ECAI/tabular.tex
minor corrections

--- 2016-05-11 Paper/ECAI/tabular.tex
@@
 More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
-the statement $\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m)$ holds. To simplify testing,
-we shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
+the  following statement holds: $$\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m).$$
+
+We shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
 To simplify notation, and in line with logic programming, we shall often make the quantification implicit and simply write this as rules $ c_1(G_1, ...
@@
 Consider the following implication:
-$$PERMUTATION(G) \rightarrow ALLDIFFERENT(G)$$
+$$\ecperm{G} \rightarrow \ecalldiff{G}$$
 (it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G$ is a vector).
@@
 An example within the same \template is for the sum over vectors constraint: $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = SUM(H)$ with $size(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
+A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = \textit{SUM}(H)$ with $\textit{size}(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
 \luc{Do we really need this, or can we write it in a real equivalence --- conjunctive formula?}

### 2016-05-11 Paper/ECAI/tabular.tex
Added examples

--- 2016-05-11 Paper/ECAI/tabular.tex
@@
 
+\section{Examples}
+\subsection{Group generation CSP}
+variables from syntax, domain from input groups, filtered using type
+
+Group generation CSP (Sum Product)\\
+variables: $R, O_1, O_2$\\
+domains: $D(R) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_1) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_2) = \{ G | G \in {\cal G} \land \numeric(G)\}$\\
+constraints: $\plength(O_1) = \plength(O_2) \geq 2, \pcols(R) = 1, \prows(R) = 1$ % excluded not partial because we don't focus on it much
+
+Translation:
+\begin{description}
+  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
+  \item[Length] Require group to have exact length (subgroups have the same length)
+  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
+  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
+  \item[Same length / table / orientation] Require groups to gave the same length table ...
+\end{description}
+
+\subsection{Subgroup generation}
+Sum Product
+
+Implementation: finds subgroups based on assignments\\
+Method $Test(G_R, G_{O_1}, G_{O_2})$\\
+$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
+$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
+\\\\
+Method $FindSubgroups(Assignments)$\\
+$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$
+
+Register implementation with the system for constraint X
+
 % \subsection{Workflow}

### 2016-05-11 Paper/ECAI/tabular.tex
fixed all links in the intro

--- 2016-05-11 Paper/ECAI/tabular.tex
@@
 
-The approach that we will introduce will borrow techniques from logical and relational learning \cite{LucBook}, where
-\cite{Claudien,todo:lallouet} have studied the discovery of clausal constraints in an inductive logic programming setting, from constraint learning,  where \cite{Quacq,Conacq,Modelseeker} have developed several approaches to learn conjunctive constraints in a constraint programming setting,
+The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
+\cite{claudien,lallouet} have studied the discovery of clausal constraints in an inductive logic programming setting, from constraint learning,  where \cite{Quacq,Conacq,modelseeker} have developed several approaches to learn conjunctive constraints in a constraint programming setting,
 and from mining in databases,  where e.g.
-\cite{Savlik} has applied inductive techniques to discover constraints (such as functional and multi-valued dependencies) in relational databases.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{}, where the definition of function is learned in spreadsheet data from
+\cite{Shavlik} has applied inductive techniques to discover constraints (such as functional and multi-valued dependencies) in relational databases.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from
 very few examples.  Our approach contrasts with all of the mentioned approaches (except Modelseeker) in that it focusses
 on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
-A more detailed discussion on related work is contained in Section \luc{TODO}.
+A more detailed discussion on related work is contained in Section \ref{sec:related_work}.
 
@@
 \\
-In terms of logical and relational learning \cite{bookLuc}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
+In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
 
@@
 Depending on the application, a solution to the tabular constraint learning problem may contain \textit{redundant} constraints, cf.
-\cite{bookLuc}.
+\cite{luc_book}.
 This is because constraints may be logically equivalent or may be implied by other constraints.

### 2016-05-11 Paper/ECAI/tabular.tex
4.2.1, template ordering

@@
 \newcommand{\format}[1]{\textit{#1}\xspace}
-\newcommand{\generategroups}{\format{generateAssignments}}
+\newcommand{\generategroups}{\format{generateGroups}}
 \newcommand{\extractgroups}{\format{extractGroups}}
@@
 \newcommand{\learnconstraints}{\format{learnConstraints}}
-\newcommand{\findassignment}{\format{findSolutions}}
+\newcommand{\findassignment}{\format{findConstraints}}
 \newcommand{\postprocess}{\format{pruneRedundant}}
@@
 This section describes our approach to \tcl.
+The main method takes \textit{groups} and {\template}s as input, and is described in Section~\ref{sec:algo}.
+We first explain how tables and groups can be extracted from raw spreadsheet data.
 
-Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
-Often, however, this not the case and a possible approach is to extract the groups from the CSV file directly.
-This requires discovering \textit{tables} and \textit{groups}.
-Our implementation offers automated procedures for both steps.
+\subsection{Table and group detection} \label{sec:make_groups}
+%A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) groups.
+Many spreadsheets contain headers that provide hints at where the table(s) in the sheet are and what the meaning of the rows and columns is.
+However, detecting tables and headers is a complex problem on its own~\cite{header}.
+Furthermore, headers may be missing and their use often requires incorporating language-specific vocabulary, e.g.
+in English.
 
-\subsection{Table and group detection}
-\samuel{Is an example needed? I find no because it is not our main concern but open for discussion}
-We chose to implement a naive approach to extract tables as the detection of tables and headers from data is a complex task on its own (\cite{header}) and a sophisticated implementation goes beyond the scope of this paper.
-Our approach first detects rectangular collections of data.
-It then attempts to identify the header by looking at the type of the first row.
-If this row consists only of textual entries, the row is expected to be the header and is removed, if not, the same procedure is applied on the first column.
-We suggest to provide tables in general, for well formated data, however, the user will not have to specify tables manually.
+%\tias{0. preprocess: currencies and precentual to numeric, 1. tables from spreadsheets, 2. groups from tables. Ext: could be done interactively.}
+Fully automatic table and header detection is not the core of our work. We hence built a visual tool that uses a simple method to detect tables in the data and allows a user to correct these.
 
-Our approach for detecting groups works by partitioning the data of a table into consecutive, type consistent groups, essentially separating textual and numerical data.
-The assumption is that often numerical and textual data are not mixed and this approach works well in practice as often groups do not have to be specified manually\footnote{\samuel{Check on benchmark}}.
-Unless an explicit orientation (rows or columns) is provided for the table, this approach will attempt to create both row- and column-groups.
+The way the table detection in the tool works is as follows:
+First, we preprocess the spreadsheets so that currency values are transformed into their integer or float value and percentual data is transformed into the corresponding floating point representation (e.g.
+$85\%$ as $0.85$).
+Next, every row in the spreadsheet is partitioned into blocks of columns that are separate by at least one $None$ value.
+Then, the blocks in each row are greedily merged to obtain rectangles.
+Finally headers are removed by checking whether the first row or column is all textual while the next row or column is not; in that case the assumed header is removed.
 
-So far we have considered textual and numerical (floating point or integer) data.
-Our implementation also supports some extensions such as currencies and percentual data, which is considered to be numerical data.
-Moreover, percentual data is corrected to the floating point representation (i.e. divided by 100).
+It should be clear this simple method only works if all tables are clearly separated by $None$ values and that headers of fully textual tables are not detected, among other things.
+Hence, in the visual tool the user can correct the detected tables if needed.
+\tias{I think it would be good if we could show a screenshot of this tool...}
 
-A more elaborate implementation could allow group (and table) detection to be done interactively / visually such that a user can select the groups in a spreadsheet or correct automatically generated groups.
+From each table, row-oriented and column-oriented groups are then extracted by partitioning the rows (columns) into type-consistent parts. This step can also be verified by a user in the visual tool.
+
 From here on we will assume that groups and tables are given.
@@
 \newcommand{\temps}{\ensuremath{S}}
+\subsection{Algorithm} \label{sec:algo}
+Our method solves the \tcl problem by checking for each template what combination of groups can satisfy the signature, and which specific subgroups satisfy the definition.
 
-\subsection{Algorithm}
-Our algorithm is illustrated in Algorithm~\ref{algo:tcl} and contains three key steps.
-First, constraint templates are ordered such that they can be treated sequentially and each only once.
-Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
-Every group assignment assigns a group~$G \in \groups$ to every argument of the template.
-In the third step, for every assignments $(G_1, ..., G_n) \in A$ we find all constraints $s(G_1', ...
-G_n')$ that hold on the data (satisfy the \CFunction), such that every subgroup $G_i' \subseteq G_i$.
+The main challenge is to avoid having to check each possible combination of group and subgroup on each constraint template, as well as developing an \textit{extensible} method in which new constraint templates can be plugged in with relative ease.
+%We will use constraint satisfaction technology to partly answer this challenge.
+To answer this challenge we separate checking the signature of a constraint template from checking the definition of a constraint template.
+Furthermore, we use constraint satisfaction technology to jointly search for and test the signature over all group combinations.
+
+The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}; it consists of three key components: 1) templates can depend on each other, so they are ordered to avoid having to verify the same template multiple times; 2) we generate all combinations of groups that can satisfy the signature using constraint satisfaction techniques; 3) from this, all satisfying subgroups are generated.
 
@@
     \State \textbf{Output:} $C$ -- learned constraints
-    \State $C \gets \emptyset$ \Comment{The set of constraints}
-  \For{$t \in \constrainttorder(\constraints,\dependencies)$}
-  \State $v_1, ..., v_n =$ variables of $T$
-    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(t, \groups, C, \dependencies)$}
-      \State $C \gets C \cup \findassignment(t, v_1{:}~G_1, \dots, v_n{:}~G_n, C, \dependencies)$
+    \State $C \gets \emptyset$ %\Comment{The set of constraints}
+  \ForAll{$t \in \constrainttorder(\constraints,\dependencies)$}
+  \State $n \gets$ number of arguments of constraint $c_t$
+    \ForAll{$(G_1, \dots, G_n) \in \generategroups(t, \groups, C)$}
+      \State $C \gets C \cup \findassignment(t, (G_1, \dots, G_n))$
     \EndFor
@@
 
-Our Algorithm \ref{algo:tcl} is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}, however, we apply the approach on two levels, generating and testing groups according to their properties and generating and testing subgroups on the actual data.
-We will explain and illustrate the three key steps in detail.
+%First, constraint templates are ordered such that they can be treated sequentially and each only once.
+%Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
+%Every group assignment assigns a group~$G \in \groups$ to every argument of the template.
+%In the third step, for every assignments $(G_1, ..., G_n) \in A$ we find all constraints $s(G_1', ...
+G_n')$ that hold on the data (satisfy the \CFunction), such that every subgroup $G_i' \subseteq G_i$.
+
+%Our Algorithm \ref{algo:tcl} is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}, however, we apply the approach on two levels, generating and testing groups according to their properties and generating and testing subgroups on the actual data.
+%We will explain and illustrate the three key steps in detail.
+
+We now describe each of the three components in more detail.
 
 \subsubsection{Template ordering}
-Looking at the template signatures in Table~\ref{table:constraints} one can see that some signatures depend on other constraints.
-This dependency also translates into the template implying the other templace.
+As discussed before, a constraint may imply another constraint.
+This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as $LOOKUP$ depending on $FOREIGNKEY$.
+
 In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
-Similarly, we can use these implications to define an ordering over the templates, such that the more general constraints (the requirements) are verified before the more specific constraints.
+In our method, we use these dependencies/implications to define a dependency graph between the templates.
+We assume the signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
 Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}. 
-Constraint templates that have no dependencies are omitted. Because there are no equivalences or loops in the signatures, the resulting graph is a directed acyclic graph (DAG).
+Constraint templates that have no dependencies are omitted.
 
-\begin{figure}[tbh]
+\begin{figure}[t]
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/constraint_dependency.png}
-  \caption{Constraint Learning Order.
-  An arrow from~$s_1$ to~$s_2$ indicates that the \CSignature of~$s_2$ requires some arguments of~$s_2$ to fulfill~$s_1$.
-  Therefore, constraints for the more general \template~$s_1$ are learned before constraints for the more specific \template~$s_2$.
+  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency.png}
+  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
   }
@@
 
-$\constrainttorder(\constraints,\dependencies)$ uses such a DAG over a subset of the constraint templates to derive a topological order of~\dependencies.
+The $\constrainttorder(\constraints,\dependencies)$ module in our method takes a set of {\template}s and a DAG as input and produces a linear ordering of the templates such that the more general, implied constraint templates (the requirements) are ordered before the templates that depend on them.
 
 \paragraph{Example}
-As mentioned earlier, \textit{foreign-key} is more specific than \textit{all-different} as $\ecfkey{v_{fk}}{v_{pk}} \implies \ecalldiff{v_{pk}}$.
-Therefore, we can first generate all subgroups $G_i'$ for which \ecalldiff{G_i'} is satisfied and use these subgroups as candidates for $v_{pk}$ in \ecfkey{v_{fk}}{v_{pk}}.
+As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence $LOOKUP$ depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
 
-The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
-\\\\
-Most parts of the algorithm could be easily parallelized.
-Every connected component in \dependencies can be learned independently.
-Within every component, constraint templates have to be treated respecting the order, i.e. learning the most general ones first.
-The generality ordering thus imposes a partial ordering according to which constraint templates must be learned.
+% not helpful:
+%The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
 
 \subsubsection{Candidate group generation}
+\tias{find groups, each $n$ groups and $m$ arguments than m-choose-n candidates.
+instead of generate-and-test we will search for all valid groups.
+To do so, CSP with signature in logical constraints and properties of each group as data.
+Example should use constraint syntax}
 $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups (group-assignments~$A$) that may contains subgroups satisfying the constraint.
@@
 
-\subsubsection{Subgroup satisfaction search}
-$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
+\subsubsection{Constraints over satisfying subgroups}
+\tias{not set of assignment but single assignment $a$?}
+\tias{requires checking for which of the subgroups of the groups the constraint is satisfied.
+many subgroups of the groups will not satisfy signature, hence no generate-and-test.
+Instead, we use again satisfiability to both search for the subgroups and verify satisfiab of the sign and definition.}
+\tias{can implement any way}
+$\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
 the \CSignature and \CFunction of~$t$ are satisfied).
@@
 
-Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty. \samuel{More on this / example?}
+Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty.
+\samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
 

### 2016-05-11 Paper/ECAI/tabular.tex
4.2.2 candidate group generation

@@
 
-\begin{algorithm}[thb]
+\begin{algorithm}[t]
   \begin{algorithmic}
@@
 \paragraph{Example}
-As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence $LOOKUP$ depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
+As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $LOOKUP$ constraint template depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
 
@@
 \subsubsection{Candidate group generation}
-\tias{find groups, each $n$ groups and $m$ arguments than m-choose-n candidates.
-instead of generate-and-test we will search for all valid groups.
-To do so, CSP with signature in logical constraints and properties of each group as data.
-Example should use constraint syntax}
-$\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups (group-assignments~$A$) that may contains subgroups satisfying the constraint.
-The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
+%The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
+Given a constraint template and the set of all groups ${\cal G}$, the goal is to find all combinations of groups that satisfy the constraint signature.
+Let there be $n$ groups and let the constraint have $k$ arguments, then a simple generate-and-test method would have to test $\tbinom{n+k-1}{k}$ combinations (given that groups can be repeated), which grows very rapidly with $n$.
 
-Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
-For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that a group~$G \in \groups$ is selected if there is at least one subgroup~$G' \subseteq G$ that satisfies the \CSignature.
+However, the properties defined in the signature, such as related to type or size, can be used to avoid considering certain groups for certain arguments.
+Furthermore, the choice of one argument can influence the possible candidates for the arguments, for example if they must have the same length.
+Instead of writing specialized code for each of these cases, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.
 
-\samuel{TODO: correct code integer groups}\\
-Consider the \ecrank{v_r}{v_x} constraint template, then both arguments must be of the same length.
-Therefore, assignments containing groups of different length do not have to be considered.
-Constraints of the \CSignature cannot always be enforced directly on a group but have to consider whether subgroups could satisfy them.
-For example, the \CSignature requires $v_r$ to be integer.
-If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this this group cannot be discarded.
+%Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
+%For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that a group~$G \in \groups$ is selected if there is at least one subgroup~$G' \subseteq G$ that satisfies the \CSignature.
 
-For some constraints (e.g. sum, max, count) that share the same \CSignature the same set of assignments~$A$ will be found. \tias{so?}
+A Constraint Satisfaction Problem (CSP) is a triple $(V,D,C)$ where $V$ is a set of \textit{variables}, $D$ the domain of possible values each variable can take and $C$ the set of constraints over $V$ that must be satisfied.
+In our case, we define one variable $V_i$ for each argument of the constraint.
+Each variable can be assigned to each of the groups in ${\cal G}$, so the domain consists of $|\{\cal G\}|$ identifiers.
+
+To reason over the groups, we automatically add to the constraint program a set of background facts that contain the properties of each group, namely its type, size and length.
+We also add all previously found constraints of constraint templates that are implied by this template.
+The actual constraints then express the properties defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
+\tias{More detailed description of the properties commented out, feel free to put in}
+%Translation:
+%\begin{description}
+%  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
+%  \item[Length] Require group to have exact length (subgroups have the same length)
+%  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
+%  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
+%  \item[Same length / table / orientation] Require groups to gave the same length table ...
+%\end{description}
+
+% TODO: old text, rephrase a bit
+Constraints derived from the signature can usually but not always be enforced exactly on a group, as we have to consider whether subgroups could satisfy them.
+For example, the \CSignature of ?? requires $v_r$ to be integer.
+If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this group cannot be discarded.
 
 \paragraph{Example}
-\ecrank{v_r}{v_x} will require groups assigned to $v_r$ to be $\integer$, groups assigned to $v_x$ to be $\discrete$ (i.e.
-$\integer$ or $\textual$) and only allows assignments in which groups assigned to both arguments have the same length.
+We show this construction of a CSP for the constraint template $\ecsumprod{G'}{H'}{I'}$ in Table~\ref{table:constraints}.
+It defines that $G'$ is a cell whose values is the result of the sum of products of rows in $H'$ and columns in $I'$.
+%
+There are three \textit{variables} $X, Y, Z$ corresponding to group $G'$, $H'$ and $I'$ respectively. The domains are $D(X) = D(Y) = D(Z) = \{1, \ldots, |{\cal G}|\}$. Finally, the constraints are:
+\begin{align*}
+&numeric(X) \land numeric(Y) \land numeric(Z) \\
+&\pcols(X) = 1 \land \prows(X) = 1 \\
+&\plength(Y) = \plength(Z) \land \plength(Y) \geq 2
+\end{align*}
+\tias{Sumproduct is a rather peculiar constraint for spreadsheets... perhaps switch to one with a dependency, e.g. look-up?}
+\\\\
+
+The $\generategroups(\textit{t,{\cal G},C})$ method will construct this CSP and query a solver for all satisfying solutions. This is the required result set.
 
 \subsubsection{Constraints over satisfying subgroups}
-\tias{not set of assignment but single assignment $a$?}
-\tias{requires checking for which of the subgroups of the groups the constraint is satisfied.
-many subgroups of the groups will not satisfy signature, hence no generate-and-test.
-Instead, we use again satisfiability to both search for the subgroups and verify satisfiab of the sign and definition.}
-\tias{can implement any way}
+\tias{I have not yet reworked this part}
+
+\tias{can implement in any way, CSP or imperative (we: imperative)}
 $\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
@@
 If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.
+
+\paragraph{Example 2}
+Sum Product
+
+Implementation: finds subgroups based on assignments\\
+Method $Test(G_R, G_{O_1}, G_{O_2})$\\
+$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
+$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
+\\\\
+Method $FindSubgroups(Assignments)$\\
+$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$
+
+Register implementation with the system for constraint X
 
@@
 \samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
-
-\section{Examples}
-\subsection{Group generation CSP}
-variables from syntax, domain from input groups, filtered using type
-
-Group generation CSP (Sum Product)\\
-variables: $R, O_1, O_2$\\
-domains: $D(R) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_1) = \{ G | G \in {\cal G} \land \numeric(G)\}$, $D(O_2) = \{ G | G \in {\cal G} \land \numeric(G)\}$\\
-constraints: $\plength(O_1) = \plength(O_2) \geq 2, \pcols(R) = 1, \prows(R) = 1$ % excluded not partial because we don't focus on it much
-
-Translation:
-\begin{description}
-  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
-  \item[Length] Require group to have exact length (subgroups have the same length)
-  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
-  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
-  \item[Same length / table / orientation] Require groups to gave the same length table ...
-\end{description}
-
-\subsection{Subgroup generation}
-Sum Product
-
-Implementation: finds subgroups based on assignments\\
-Method $Test(G_R, G_{O_1}, G_{O_2})$\\
-$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
-$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
-\\\\
-Method $FindSubgroups(Assignments)$\\
-$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$
-
-Register implementation with the system for constraint X
 

### 2016-05-11 Paper/ECAI/tabular.tex
started conclusions
editing the conclusions

@@
 \newcommand{\template}{\format{constraint template}}
+\newcommand{\sname}{\format{TaCKLe}}
 
@@
 \section{Conclusions}
-\sergey{link to the github with implementations and the dataset}
-
-\sergey{Future version, extensions, declarative stuff, bla-bla here}
 
 % points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
-Our goal was to automatically identify constraints that hold in a spreadsheet.
-We presented and evaluated an approach that is able to learn a large set of constraints from CSV data.
-Moreover, the amount of redundant constraints found by the algorithm is limited despite the lack of more sophisticated post-processing steps.
+Our goal is to automatically identify constraints in a spreadsheet.
+We have presented and evaluated our approach, implemented as the system \sname, that is able to learn a large set of constraints from CSV data.
+Moreover, the amount of redundant constraints found by the system is limited despite the lack of more sophisticated post-processing steps.
 
 \paragraph{Future work}
-We see multiple directions for future work on this topic.
+There are multiple directions for future work on this topic.
 The system we presented would become more useful in an interactive setting where users can easily receive and provide feedback.
-Moreover, the system could be adapted to produce less redundant constraints through the use of heuristic filtering.
+Moreover, the system can be adapted to produce less redundant constraints through the use of heuristic filtering.
 
 Currently, the system only learns single constraints, however, extending this approach to nested constraints would allow more expressive concepts to be learned.
-This shift would bring the approach more in line with programming by example.
+This shift would bring the approach in line with programming by example.
 
-Aside from finding errors, the system could be extended to deal with noise and the ability to also learn soft constraints.
-Soft constraints are (potentially weighted) constraints that hold only on some of the data.
-This could extend the approach to new application domains as well as provide more native error detection.
+Aside from finding errors, the system can be extended to deal with noise and the ability to learn soft constraints.
+Soft constraints are the (potentially weighted) constraints that hold only on some of the data.
+This would extend the approach to new application domains as well as provide more native error detection.
 
-\bibliographystyle{ecai}
+%\bibliographystyle{ecai}
 \bibliography{references}

### 2016-05-11 Paper/ECAI/tabular.tex
WIP error detection

--- 2016-05-11 Paper/ECAI/tabular.tex
@@
 \newcommand{\format}[1]{\textit{#1}\xspace}
-\newcommand{\generategroups}{\format{generateGroups}}
+\newcommand{\generategroups}{\format{generateAssignments}}
 \newcommand{\extractgroups}{\format{extractGroups}}
@@
 \newcommand{\learnconstraints}{\format{learnConstraints}}
-\newcommand{\findassignment}{\format{findConstraints}}
+\newcommand{\findassignment}{\format{findSolutions}}
 \newcommand{\postprocess}{\format{pruneRedundant}}
@@
 This section describes our approach to \tcl.
-The main method takes \textit{groups} and {\template}s as input, and is described in Section~\ref{sec:algo}.
-We first explain how tables and groups can be extracted from raw spreadsheet data.
 
-\subsection{Table and group detection} \label{sec:make_groups}
-%A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) groups.
-Many spreadsheets contain headers that provide hints at where the table(s) in the sheet are and what the meaning of the rows and columns is.
-However, detecting tables and headers is a complex problem on its own~\cite{header}.
-Furthermore, headers may be missing and their use often requires incorporating language-specific vocabulary, e.g.
-in English.
+Our algorithm reasons on the level of groups and vectors on the data and assumes that these are provided.
+Often, however, this not the case and a possible approach is to extract the groups from the CSV file directly.
+This requires discovering \textit{tables} and \textit{groups}.
+Our implementation offers automated procedures for both steps.
 
-%\tias{0. preprocess: currencies and precentual to numeric, 1. tables from spreadsheets, 2. groups from tables. Ext: could be done interactively.}
-Fully automatic table and header detection is not the core of our work. We hence built a visual tool that uses a simple method to detect tables in the data and allows a user to correct these.
+\subsection{Table and group detection}
+\samuel{Is an example needed? I find no because it is not our main concern but open for discussion}
+We chose to implement a naive approach to extract tables as the detection of tables and headers from data is a complex task on its own (\cite{header}) and a sophisticated implementation goes beyond the scope of this paper.
+Our approach first detects rectangular collections of data.
+It then attempts to identify the header by looking at the type of the first row.
+If this row consists only of textual entries, the row is expected to be the header and is removed, if not, the same procedure is applied on the first column.
+We suggest to provide tables in general, for well formated data, however, the user will not have to specify tables manually.
 
-The way the table detection in the tool works is as follows:
-First, we preprocess the spreadsheets so that currency values are transformed into their integer or float value and percentual data is transformed into the corresponding floating point representation (e.g.
-$85\%$ as $0.85$).
-Next, every row in the spreadsheet is partitioned into blocks of columns that are separate by at least one $None$ value.
-Then, the blocks in each row are greedily merged to obtain rectangles.
-Finally headers are removed by checking whether the first row or column is all textual while the next row or column is not; in that case the assumed header is removed.
+Our approach for detecting groups works by partitioning the data of a table into consecutive, type consistent groups, essentially separating textual and numerical data.
+The assumption is that often numerical and textual data are not mixed and this approach works well in practice as often groups do not have to be specified manually\footnote{\samuel{Check on benchmark}}.
+Unless an explicit orientation (rows or columns) is provided for the table, this approach will attempt to create both row- and column-groups.
 
-It should be clear this simple method only works if all tables are clearly separated by $None$ values and that headers of fully textual tables are not detected, among other things.
-Hence, in the visual tool the user can correct the detected tables if needed.
-\tias{I think it would be good if we could show a screenshot of this tool...}
+So far we have considered textual and numerical (floating point or integer) data.
+Our implementation also supports some extensions such as currencies and percentual data, which is considered to be numerical data.
+Moreover, percentual data is corrected to the floating point representation (i.e. divided by 100).
 
-From each table, row-oriented and column-oriented groups are then extracted by partitioning the rows (columns) into type-consistent parts. This step can also be verified by a user in the visual tool.
-
+A more elaborate implementation could allow group (and table) detection to be done interactively / visually such that a user can select the groups in a spreadsheet or correct automatically generated groups.
 From here on we will assume that groups and tables are given.
@@
 \newcommand{\temps}{\ensuremath{S}}
-\subsection{Algorithm} \label{sec:algo}
-Our method solves the \tcl problem by checking for each template what combination of groups can satisfy the signature, and which specific subgroups satisfy the definition.
 
-The main challenge is to avoid having to check each possible combination of group and subgroup on each constraint template, as well as developing an \textit{extensible} method in which new constraint templates can be plugged in with relative ease.
-%We will use constraint satisfaction technology to partly answer this challenge.
-To answer this challenge we separate checking the signature of a constraint template from checking the definition of a constraint template.
-Furthermore, we use constraint satisfaction technology to jointly search for and test the signature over all group combinations.
-
-The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}; it consists of three key components: 1) templates can depend on each other, so they are ordered to avoid having to verify the same template multiple times; 2) we generate all combinations of groups that can satisfy the signature using constraint satisfaction techniques; 3) from this, all satisfying subgroups are generated.
+\subsection{Algorithm}
+Our algorithm is illustrated in Algorithm~\ref{algo:tcl} and contains three key steps.
+First, constraint templates are ordered such that they can be treated sequentially and each only once.
+Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
+Every group assignment assigns a group~$G \in \groups$ to every argument of the template.
+In the third step, for every assignments $(G_1, ..., G_n) \in A$ we find all constraints $s(G_1', ...
+G_n')$ that hold on the data (satisfy the \CFunction), such that every subgroup $G_i' \subseteq G_i$.
 
@@
     \State \textbf{Output:} $C$ -- learned constraints
-    \State $C \gets \emptyset$ %\Comment{The set of constraints}
-  \ForAll{$t \in \constrainttorder(\constraints,\dependencies)$}
-  \State $n \gets$ number of arguments of constraint $c_t$
-    \ForAll{$(G_1, \dots, G_n) \in \generategroups(t, \groups, C)$}
-      \State $C \gets C \cup \findassignment(t, (G_1, \dots, G_n))$
+    \State $C \gets \emptyset$ \Comment{The set of constraints}
+  \For{$t \in \constrainttorder(\constraints,\dependencies)$}
+  \State $v_1, ..., v_n =$ variables of $T$
+    \For{$v_1{:}~G_1, \dots, v_n{:}~G_n \in \generategroups(t, \groups, C, \dependencies)$}
+      \State $C \gets C \cup \findassignment(t, v_1{:}~G_1, \dots, v_n{:}~G_n, C, \dependencies)$
     \EndFor
@@
 
-%First, constraint templates are ordered such that they can be treated sequentially and each only once.
-%Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
-%Every group assignment assigns a group~$G \in \groups$ to every argument of the template.
-%In the third step, for every assignments $(G_1, ..., G_n) \in A$ we find all constraints $s(G_1', ...
-G_n')$ that hold on the data (satisfy the \CFunction), such that every subgroup $G_i' \subseteq G_i$.
-
-%Our Algorithm \ref{algo:tcl} is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}, however, we apply the approach on two levels, generating and testing groups according to their properties and generating and testing subgroups on the actual data.
-%We will explain and illustrate the three key steps in detail.
-
-We now describe each of the three components in more detail.
+Our Algorithm \ref{algo:tcl} is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}, however, we apply the approach on two levels, generating and testing groups according to their properties and generating and testing subgroups on the actual data.
+We will explain and illustrate the three key steps in detail.
 
 \subsubsection{Template ordering}
-As discussed before, a constraint may imply another constraint.
-This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as $LOOKUP$ depending on $FOREIGNKEY$.
+Looking at the template signatures in Table~\ref{table:constraints} one can see that some signatures depend on other constraints.
+This dependency also translates into the template implying the other templace.
+In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
+Similarly, we can use these implications to define an ordering over the templates, such that the more general constraints (the requirements) are verified before the more specific constraints.
+Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}.
+Constraint templates that have no dependencies are omitted. Because there are no equivalences or loops in the signatures, the resulting graph is a directed acyclic graph (DAG).
 
-In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
-In our method, we use these dependencies/implications to define a dependency graph between the templates.
-We assume the signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
-Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}. 
-Constraint templates that have no dependencies are omitted.
-
-\begin{figure}[t]
+\begin{figure}[tbh]
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency.png}
-  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
+  \includegraphics[width=0.9\linewidth]{figures/constraint_dependency.png}
+  \caption{Constraint Learning Order.
+  An arrow from~$s_1$ to~$s_2$ indicates that the \CSignature of~$s_2$ requires some arguments of~$s_2$ to fulfill~$s_1$.
+  Therefore, constraints for the more general \template~$s_1$ are learned before constraints for the more specific \template~$s_2$.
   }
@@
 
-The $\constrainttorder(\constraints,\dependencies)$ module in our method takes a set of {\template}s and a DAG as input and produces a linear ordering of the templates such that the more general, implied constraint templates (the requirements) are ordered before the templates that depend on them.
+$\constrainttorder(\constraints,\dependencies)$ uses such a DAG over a subset of the constraint templates to derive a topological order of~\dependencies.
 
 \paragraph{Example}
-As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence $LOOKUP$ depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
+As mentioned earlier, \textit{foreign-key} is more specific than \textit{all-different} as $\ecfkey{v_{fk}}{v_{pk}} \implies \ecalldiff{v_{pk}}$.
+Therefore, we can first generate all subgroups $G_i'$ for which \ecalldiff{G_i'} is satisfied and use these subgroups as candidates for $v_{pk}$ in \ecfkey{v_{fk}}{v_{pk}}.
 
-% not helpful:
-%The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+\\\\
+Most parts of the algorithm could be easily parallelized.
+Every connected component in \dependencies can be learned independently.
+Within every component, constraint templates have to be treated respecting the order, i.e. learning the most general ones first.
+The generality ordering thus imposes a partial ordering according to which constraint templates must be learned.
 
 \subsubsection{Candidate group generation}
-\tias{find groups, each $n$ groups and $m$ arguments than m-choose-n candidates.
-instead of generate-and-test we will search for all valid groups.
-To do so, CSP with signature in logical constraints and properties of each group as data.
-Example should use constraint syntax}
 $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups (group-assignments~$A$) that may contains subgroups satisfying the constraint.
@@
 
-\subsubsection{Constraints over satisfying subgroups}
-\tias{not set of assignment but single assignment $a$?}
-\tias{requires checking for which of the subgroups of the groups the constraint is satisfied.
-many subgroups of the groups will not satisfy signature, hence no generate-and-test.
-Instead, we use again satisfiability to both search for the subgroups and verify satisfiab of the sign and definition.}
-\tias{can implement any way}
-$\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
+\subsubsection{Subgroup satisfaction search}
+$\findassignment(t,A,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
 the \CSignature and \CFunction of~$t$ are satisfied).
@@
 
-Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty.
-\samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
+Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty. \samuel{More on this / example?}
 
@@
 Therefore, constraints are learned at time~$t_0$ and combined with data from a new row~$i$ (column~$j$) that the user has added to a table~$T$ by time~$t_1$.
-Constraints that do not conflict with the added data are considered viable.
+Constraints that do not conflict with the added data are considered \textit{viable}.
 Viable constraints that have enough input data to compute the result for a blank cell~$T[i,j]$ in the new row (column) are \textit{active}, which means they can be used to predict the outcome for that cell.
-Active constraints can be used to autocomplete blank cells such that the user does not have to fill in the values manually.
+\textit{Active} constraints can be used to autocomplete blank cells such that the user does not have to fill in the values manually.
 Once the user has stopped editing the new row~$i$ (column~$j$), the table~$T$ is updated to include~$i$ ($j$) and constraints are learned once again.
@@
 \subsection{Error detection}
-For this application we consider a use case concerning crime statistics provided by the FBI.
+We consider two cases of error detection.
+The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
+However, instead of considering viable constraints, we look at \textit{conflict} constraints, i.e. constraints that were learned at time~$t_0$ but do not agree with the newly added data.
+Such conflict constraints can indicate an error in the input, however, they could also be accidental constraints that are invalidated by the new data.
+
+The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
+Consider, for example, a use case concerning crime statistics provided by the FBI.
+An extract of the spreadsheet is depicted in Figure~\ref{fig:fbi}.
+When run on this sheet, our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
 

### 2016-05-11 Paper/ECAI/tabular.tex
style correction

--- 2016-05-11 Paper/ECAI/tabular.tex
@@
 
-{\it
-\textbf{Motivation}:
-\begin{itemize}
-  \item USED -- File generated from model, model got lost, need to reconstruct
-  \item Constraint programming is hard - is Excel hard?
-  \item Avoid manual analysis, provide selection of constraints
-  \item SOMEWHAT USED --Error checking
-  \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
-\end{itemize}
+% {\it
+% \textbf{Motivation}:
+% \begin{itemize}
+%   \item USED -- File generated from model, model got lost, need to reconstruct
+%   \item Constraint programming is hard - is Excel hard?
+%   \item Avoid manual analysis, provide selection of constraints
+%   \item SOMEWHAT USED --Error checking
+%   \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
+% \end{itemize}
 
-\textbf{Novelty:}
-\begin{itemize}
-  \item USED -- Unsupervised setting (contrary to flashfill, etc)
-  \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
-  \item USED -- Data format (2D) -- data is no longer in rows like a classic ML or DM settings
-  \item USED -- Declarative, general / modular, stacking of constraint problems
-\end{itemize}
+% \textbf{Novelty:}
+% \begin{itemize}
+%   \item USED -- Unsupervised setting (contrary to flashfill, etc)
+%   \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
+%   \item USED -- Data format (2D) -- data is no longer in rows like a classic ML or DM settings
+%   \item USED -- Declarative, general / modular, stacking of constraint problems
+% \end{itemize}
+%}
 
@@
 \end{figure*}
-}
 
@@
 
-%\bibliographystyle{ecai}
+\bibliographystyle{plain}
 \bibliography{references}

### 2016-05-11 Paper/ECAI/tabular.tex
conclusion style correction

@@
 There are multiple directions for future work on this topic.
-The system we presented would become more useful in an interactive setting where users can easily receive and provide feedback.
+The system would be easier to use in an interactive setting where users can easily receive and provide feedback.
 Moreover, the system can be adapted to produce less redundant constraints through the use of heuristic filtering.
 
-Currently, the system only learns single constraints, however, extending this approach to nested constraints would allow more expressive concepts to be learned.
+Currently, the system only learns single constraints, however, extending it to nested constraints would allow more expressive concepts to be learned.
 This shift would bring the approach in line with programming by example.

### 2016-05-11 Paper/ECAI/tabular.tex
added a stab for the extension
Expanded error detection, small fixes conclusion

@@
 \newcommand{\template}{\format{constraint template}}
-\newcommand{\sname}{\format{TaCKLe}}
 
@@
 
-% {\it
-% \textbf{Motivation}:
-% \begin{itemize}
-%   \item USED -- File generated from model, model got lost, need to reconstruct
-%   \item Constraint programming is hard - is Excel hard?
-%   \item Avoid manual analysis, provide selection of constraints
-%   \item SOMEWHAT USED --Error checking
-%   \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
-% \end{itemize}
+{\it
+\textbf{Motivation}:
+\begin{itemize}
+  \item USED -- File generated from model, model got lost, need to reconstruct
+  \item Constraint programming is hard - is Excel hard?
+  \item Avoid manual analysis, provide selection of constraints
+  \item SOMEWHAT USED --Error checking
+  \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
+\end{itemize}
 
-% \textbf{Novelty:}
-% \begin{itemize}
-%   \item USED -- Unsupervised setting (contrary to flashfill, etc)
-%   \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
-%   \item USED -- Data format (2D) -- data is no longer in rows like a classic ML or DM settings
-%   \item USED -- Declarative, general / modular, stacking of constraint problems
-% \end{itemize}
-%}
+\textbf{Novelty:}
+\begin{itemize}
+  \item USED -- Unsupervised setting (contrary to flashfill, etc)
+  \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
+  \item USED -- Data format (2D) -- data is no longer in rows like a classic ML or DM settings
+  \item USED -- Declarative, general / modular, stacking of constraint problems
+\end{itemize}
 
@@
 \end{figure*}
+}
 
@@
 
-\subsection{Error detection}
-We consider two cases of error detection.
-The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
-However, instead of considering viable constraints, we look at \textit{conflict} constraints, i.e. constraints that were learned at time~$t_0$ but do not agree with the newly added data.
-Such conflict constraints can indicate an error in the input, however, they could also be accidental constraints that are invalidated by the new data.
-
-The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
-Consider, for example, a use case concerning crime statistics provided by the FBI.
-An extract of the spreadsheet is depicted in Figure~\ref{fig:fbi}.
-When run on this sheet, our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
-
-\sergey{We should explain Figure \ref{fig:fbi}, e.g. if we aggregate by the crime now, the missing value would not get into statistics, which seems faulty, since we can directly compute it.}
-
 \begin{figure*}[thb]
@@
 \end{figure*}
+
+\subsection{Error detection}
+We consider two cases of error detection.
+The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
+However, instead of considering viable constraints, we look at \textit{conflict} constraints, i.e. constraints that were learned at time~$t_0$ but do not agree with the newly added data at time~$t_1$.
+Such conflict constraints can indicate errors in the input.
+Since conflict constraints could also be accidental constraints that are correctly invalidated by the new data, the size of the data at timet~$t_0$ should be large enough.
+
+The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
+Consider, for example, a use case concerning crime statistics provided by the FBI.
+An extract of the spreadsheet is depicted in Figure~\ref{fig:fbi}.
+When run on this sheet, our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
+A possible approach to detect such missing or wrong values is to look at a sample of the table that excludes the erroneous row and compare the constraints learned on this sample with the constraints learned by adding the row containing the mistake.
+The sample needs to large enough such that we have enough confidence that the constraints are correct.
+If we define the sample to be our data at time~$t_0$ and the added row to be the additional input at~$t_1$, this task can be formulated as online error detection and dealt with in the same way.
+Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
+In the limited case of detecting single errors one can leave out every row (column), consider the remaining data to be the sample and
+and the left out row (column) to be new input.
+\samuel{Should this be more specific?}
+
+Error detection could be integrated more closely with the algorithm, reducing the computational complexity by looking for errors during the constraint learning phase.
+This would, however, make the modeling constraints (considerably) more difficult.
 
@@
 \section{Conclusions}
+\sergey{link to the github with implementations and the dataset}
+
+\sergey{Future version, extensions, declarative stuff, bla-bla here}
 
 % points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
-Our goal is to automatically identify constraints in a spreadsheet.
-We have presented and evaluated our approach, implemented as the system \sname, that is able to learn a large set of constraints from CSV data.
-Moreover, the amount of redundant constraints found by the system is limited despite the lack of more sophisticated post-processing steps.
+Our goal was to automatically identify constraints that hold in a spreadsheet.
+We presented and evaluated an approach that is able to learn a large set of constraints from CSV data.
+Moreover, the amount of redundant constraints found by the algorithm is limited despite the lack of more sophisticated post-processing steps.
 
 \paragraph{Future work}
-There are multiple directions for future work on this topic.
-The system would be easier to use in an interactive setting where users can easily receive and provide feedback.
-Moreover, the system can be adapted to produce less redundant constraints through the use of heuristic filtering.
+We see multiple directions for future work on this topic.
+The system we presented would become more useful in an interactive setting where users can easily receive and provide feedback.
+Moreover, the system could be adapted to deal with additional types and produce less redundant constraints through the use of (heuristic) filtering or post-processing.
 
-Currently, the system only learns single constraints, however, extending it to nested constraints would allow more expressive concepts to be learned.
-This shift would bring the approach in line with programming by example.
+Currently, the system only learns single constraints, however, extending this approach to nested constraints would allow more expressive concepts to be learned.
+This shift would bring the approach closer towards programming by example.
 
-Aside from finding errors, the system can be extended to deal with noise and the ability to learn soft constraints.
-Soft constraints are the (potentially weighted) constraints that hold only on some of the data.
-This would extend the approach to new application domains as well as provide more native error detection.
+Aside from finding errors, the system could be extended to deal with noise and the ability to also learn soft constraints.
+Soft constraints are (potentially weighted) constraints that hold only on some of the data.
+This could extend the approach to new application domains as well as provide more native error detection.
 
-\bibliographystyle{plain}
+\bibliographystyle{ecai}
 \bibliography{references}

### 2016-05-11 Paper/ECAI/tabular.tex

--- 2016-05-11 Paper/ECAI/tabular.tex
@@
 
+\paragraph{Application specific system extension}
+\sergey{Explain here, how the system can be extended with a new constraints like vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors}
+
+\begin{equation}
+  \forall i{:}~ v_r[i] \leq v[i] < v_r[i] + 0.5
+  \label{eq:floor_extension}
+\end{equation}
+
 % \subsection{Workflow}
@@
 There are multiple directions for future work on this topic.
-The system would be easier to use in an interactive setting where users can easily receive and provide feedback.
-Moreover, the system can be adapted to produce less redundant constraints through the use of heuristic filtering.
+Integrating the system in an interactive setting would offer users the possibility to easily receive and provide feedback.
+Moreover, the system can be adapted to deal with additional types produce less redundant constraints through the use of (heuristic) filtering or post-processing.
 
-Currently, the system only learns single constraints, however, extending it to nested constraints would allow more expressive concepts to be learned.
+Currently, the system only learns single constraints, however, extending our approach to nested constraints would allow more expressive concepts to be learned.
 This shift would bring the approach in line with programming by example.

### 2016-05-12 Paper/ECAI/tabular.tex
Did TODO's in intro

@@
 on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
-A more detailed discussion on related work is contained in Section \ref{sec:related_work}.
+A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.
 
@@
 It works in an unsupervised setting, as only the tabular data is provided and no examples need to be specified, and discovers column- and row-wise constraints.
-We demonstrate the utility of our approach in an experimental evaluation and we also sketch several application scenario's.
-\luc{todo:mention these here too}
-
-This paper is organised as follows: \luc{todo}
+We demonstrate the utility of our approach in an experimental evaluation.
+Moreover, we sketch several application scenario's, such as autocompletion and error detection.
+\\\\
+This paper is organised as follows:
+Section~2 introduces concepts relevant to our approach, Section~3 presents our problem statement, Section~4 describes our approach, the approach is evaluated in Section~5, in Section~6 we show how our system can be used for autocompletion and error detection, and we conclude in Section~7.
 

### 2016-05-12 Paper/ECAI/tabular.tex
minor rewriting in the introduction

@@
 
-In this paper,  we investigate whether machine learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
+In this paper,  we investigate whether learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
 This is a new and unconventional machine learning problem.
@@
 The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
-\cite{claudien,lallouet} have studied the discovery of clausal constraints in an inductive logic programming setting, from constraint learning,  where \cite{Quacq,Conacq,modelseeker} have developed several approaches to learn conjunctive constraints in a constraint programming setting,
-and from mining in databases,  where e.g.
-\cite{Shavlik} has applied inductive techniques to discover constraints (such as functional and multi-valued dependencies) in relational databases.
+the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to conjunctive constraint learning have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
+and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{Shavlik}.
 It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from

### 2016-05-12 Paper/ECAI/tabular.tex
rewriting paper strucutre

@@
 \\\\
-This paper is organised as follows:
-Section~2 introduces concepts relevant to our approach, Section~3 presents our problem statement, Section~4 describes our approach, the approach is evaluated in Section~5, in Section~6 we show how our system can be used for autocompletion and error detection, and we conclude in Section~7.
+This paper is organized as follows.
+Section \ref{sec:formalization} introduces concepts relevant to our approach.
+Section \ref{sec:problem_statement} presents our problem statement.
+Section \ref{sec:approach} describes our approach.
+Section \ref{sec:evaluation} presents the evaluation of the approach.
+Section \ref{sec:applications} shows how our system can be used for autocompletion and error detection.
+Section \ref{sec:related_work} present an overview of the related work.
+And Section \ref{sec:conclusions} provides conclusions.
 
@@
 
-\section{Formalization}
+\section{Formalization}\label{sec:formalization}
 As we focus on spreadsheet data, we first introduce the general concepts of cells, group and constraint templates, using the (multi-table) sheet in Figure~\ref{fig:main_example} as running example.
@@
 
-\section{Problem Statement}
+\section{Problem Statement}\label{sec:problem_statement}
 The problem of learning constraints from tabular data can be seen as an inverse {\em constraint satisfaction problem} (CSP). A CSP consists of a set of constraints and the
@@
 %\newcommand{\ctl}{Condensed Tabular Constraint Learning}
-\section{Approach to Tabular Constraint Learning}
+\section{Approach to Tabular Constraint Learning}\label{sec:approach}
 This section describes our approach to \tcl.
@@
 
-\section{Evaluation}
+\section{Evaluation}\label{sec:evaluation}
 \sergey{we need to add a summary of the dataset, avg number of constraints, cells, rows, columns}
@@
 
-\section{Applications}
+\section{Applications}\label{sec:applications}
 In this section we will illustrate how various motivating applications could be realized using our system.
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 

### 2016-05-12 Paper/ECAI/tabular.tex
rewriting paper strucutre

@@
 Section \ref{sec:formalization} introduces concepts relevant to our approach.
-Section \ref{sec:problem_statement} presents our problem statement.
-Section \ref{sec:approach} describes our approach.
+Section \ref{sec:problem_statement} presents the problem statement.
+Section \ref{sec:approach} describes the approach.
 Section \ref{sec:evaluation} presents the evaluation of the approach.
-Section \ref{sec:applications} shows how our system can be used for autocompletion and error detection.
+Section \ref{sec:applications} shows how our system can be used for applications such as autocompletion and error detection.
 Section \ref{sec:related_work} present an overview of the related work.

### 2016-05-12 Paper/ECAI/tabular.tex
minor changes

@@
 \newcommand{\tias}[1]{\textcolor{blue}{{\sc Tias:} #1}\xspace}
+\newcommand{\luc}[1]{\textcolor{red}{{\sc Luc:} #1}\xspace}
 
@@
 \newcommand{\template}{\format{constraint template}}
-\newcommand{\sname}{\format{TaCKLe}}
+\newcommand{\sname}{\format{TaCLe}}
 
@@
 
-\newcommand{\luc}[1]{{\textcolor{red}{#1}}}
-
 \renewcommand{\arraystretch}{1.5}
@@
 the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to conjunctive constraint learning have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
-and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{Shavlik}.
+and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
 It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from

### 2016-05-12 Paper/ECAI/tabular.tex
Updated table

--- 2016-05-12 Paper/ECAI/tabular.tex
@@
 \newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
+\newcommand{\ecdiff}[3]{\eccalc{#1}{#2 - #3}}
 \newcommand{\ectotal}[3]{\eccalc{#1}{\mathit{PREV}(#1) + #2 - #3}}
@@
 \newcommand{\nat}{\mathcal{N}}
+
+\newcommand{\sg}{G'}
 
@@
 \begin{table*}
-\tias{Notation inconsistent (no $G'$?
-also the 'bold' notation not explained.
-Also, in many places missing that must be vector...
-switched $\rightarrow$ to FOREIGNKEY to avoid confusion with implied constraints.}
-  \centering
+\medskip
+Constraint templates marked with $*$ are \textit{structural}, they are discovered by our algorithm, however, they cannot be readily used in spreadsheet software.
+The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
+Only bold subgroups may contain more than one vector.
+\medskip
+
+  {\centering
   \begin{tabularx}{\textwidth}{l X X}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    $\ecalldiff{v_x}^*$
-      & $\discrete(v_x)$
-      & All values in $v_x$ are different: $v_x[i] \neq v_x[j]$ if $i \neq j$
+    $\ecalldiff{\sg_x}^*$
+      & $\discrete(\sg_x)$
+      & All values in $\sg_x$ are different: $\sg_x[i] \neq \sg_x[j]$ if $i \neq j$
       \\ \hline
     %X = Y & & \\
-    $\ecperm{v_x}^*$
-      & $\numeric(v_{x})$, $\ecalldiff{v_{x}}$
-      & The values in $v_{x}$ are a permutation of the numbers $1$ through $\plength(v_{x})$.
-      \\ \hline
-    \ecseries{v_x}
-      & $\numeric(v_{x})$ and $\ecperm{v_{x}}$
-      & $v_{x}[1] = 1$ and $v_{x}[i] = v_{x}[i - 1] + 1$.
-      \\ \hline
-    \ecfkey{v_{fk}}{v_{pk}} & $v_{fk}$ and~$v_{pk}$ are both $\discrete$; $\ptype(v_{fk}) = \ptype(v_{pk})$; $\ptable(v_{fk}) \neq \ptable(v_{pk})$; and $\ecalldiff{v_{pk}}$ & Every value in~$v_{fk}$ also exist in~$v_{pk}$ \\ \hline
+    $\ecperm{\sg_x}^*$
+      & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
+      & The values in $\sg_{x}$ are a permutation of the numbers $1$ through $\plength(\sg_{x})$.
+      \\ \hline
+    \ecseries{\sg_x}
+      & $\numeric(\sg_{x})$ and $\ecperm{\sg_{x}}$
+      & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
+      \\ \hline
+    \ecfkey{\sg_{fk}}{\sg_{pk}} & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$ & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
     % TODO check out
-    \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
-      & $v_{fk}$ and $v_{pk}$ are both $\discrete$; arguments $\{v_{fk}, v_{r}\}$ and $\{v_{pk}, v_{val}\}$ within the same set have the same \plength, \ptable and \por; $v_{r}$ and~$v_{val}$ have the same type; and \ecfkey{v_{fk}}{v_{pk}}.
-      & $v_r[i]$ is the same value as looking up~$v_{fk}[i]$ in~$v_{pk}$  and returning the corresponding value in~$v_{val}$.
-      \\ \hline
-    \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}}
+    \eclookup{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+      & $\sg_{fk}$ and $\sg_{pk}$ are both $\discrete$; arguments $\{\sg_{fk}, \sg_{r}\}$ and $\{\sg_{pk}, \sg_{val}\}$ within the same set have the same \plength, \ptable and \por; $\sg_{r}$ and~$\sg_{val}$ have the same type; and \ecfkey{\sg_{fk}}{\sg_{pk}}.
+      & $\sg_r[i]$ is the same value as looking up~$\sg_{fk}[i]$ in~$\sg_{pk}$  and returning the corresponding value in~$\sg_{val}$.
+      \\ \hline
+    \eclookupfuzzy{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
       & \textit{Same as lookup}
-      & $v_r[i]$ is the same value as looking up the last item in~$v_{pk}$ smaller than~$v_{fk}[i]$ and returning the corresponding value in~$v_{val}$.
-      \\ \hline
-    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
-      & Arguments $\{v_{r}, v_{1}, v_{fk}\}$ are $\numeric$, arguments $\{v_{pk}, v_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{v_{fk}}{v_{pk}}.
-      & $v_{r}[i]$ is the obtained by multiplying $v_{1}[i]$ with $\eclookupf{v_{fk}}{v_{pk}}{v_{val}}{}[i]$.
-      \\ \hline
-    \ecprod{v_r}{v_1}{v_2}
-      & Arguments $\{v_{r}, v_{1}, v_{2}\}$ are all $\numeric$ and have the same $\plength$
-      & $v_{r}[i] = v_{1}[i] \times v_{2}[i]$.
-      \\ \hline
-    \ecproj{v_r}{\mathbf{v_x}}
-      & Arguments $\{v_{r}, \mathbf{v_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{v_x}$ contains at least~2 vectors; and $v_r = \mathit{SUM}(\mathbf{v_x}, \por(\mathbf{v_x}))$
-      & At every position~$i$ in $1$ through $\plength(v_{r})$ there is exactly one vector~$v$ in $\mathbf{v_x}$ such that $v[i]$ is a non-blank value, then $v[i] = v_{r}[i]$.
-      \\ \hline
-    \ecrank{v_r}{v_x}
-      & $\integer(v_{r})$; $\numeric(v_{x})$; and $\plength(v_{r}) = \plength(v_{x})$
-      & The values in $v_{r}$ represent the rank (from largest to smallest) of the values in $v_{x}$ (including ties)
-      \\ \hline
-    \ectotal{v_r}{v_{pos}}{v_{neg}}
-      & Arguments $\{v_{r}, v_{pos}, v_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
-      & The values in $v_{r}$ are a running total, each value $v_{r}[i] = v_{r}[i - 1] + v_{pos}[i] - v_{neg}[i]$.
-      \\ \hline
-    $\ecsumc{v_r}{\mathbf{v_x}}^\dagger$
-      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\prows(\mathbf{v_x}) \geq 2$; and $\pcols(\mathbf{v_x}) = \plength(v_r)$
-      & Each value in $v_{r}$ is obtained by summing over the corresponding column in $\mathbf{v_x}$.
-      \\ \hline
-    $\ecsumr{v_r}{\mathbf{v_x}}^\dagger$
-      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\pcols(\mathbf{v_x}) \geq 2$; and $\prows(\mathbf{v_x}) = \plength(v_r)$
-      & Each value in $v_{r}$ is obtained by summing over the corresponding row in $\mathbf{v_x}$.
-      \\ \hline
-    $\ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}^\dagger$
-      & $v_{fk}, v_{pk}$ are $\discrete$; $v_{r}, v_{val}$ are $\numeric$; within the sets $\{v_{val}, v_{fk}\}$ and $\{v_{pk}, v_{r}\}$ arguments have the same $\plength$ and $\por$; $v_{fk}$ and $v_{val}$ have the same $\ptable$; $v_{fk}$ and $v_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{v_{pk}}
-      & The value for $v_{r}[i]$ is obtained by summing all values $v_{val}[j]$ where $v_{fk}[j] = v_{pk}[i]$
-      \\ \hline
-    \ecsumprod{v_r}{v_1}{v_2}
-      & Arguments $\{v_r, v_1, v_2\}$ are all $\numeric$; $\plength(v_{1}) = \plength(v_{2}) \geq 2$; and $\prows(v_{r}) = \pcols(v_{r}) = 1$
-      & $v_{r}[i] = \sum_{i = 1}^{\plength(v_{1})} v_{1}[i] \times v_{2}[i]$.
+      & $\sg_r[i]$ is the same value as looking up the last item in~$\sg_{pk}$ smaller than~$\sg_{fk}[i]$ and returning the corresponding value in~$\sg_{val}$.
+      \\ \hline
+    \eclookupprod{\sg_r}{\sg_1}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{fk}\}$ are $\numeric$, arguments $\{\sg_{pk}, \sg_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sg_{fk}}{\sg_{pk}}.
+      & $\sg_{r}[i]$ is the obtained by multiplying $\sg_{1}[i]$ with $\eclookupf{\sg_{fk}}{\sg_{pk}}{\sg_{val}}{}[i]$.
+      \\ \hline
+    \ecprod{\sg_r}{\sg_1}{\sg_2}
+      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$
+      & $\sg_{r}[i] = \sg_{1}[i] \times \sg_{2}[i]$.
+      \\ \hline
+    \ecdiff{\sg_r}{\sg_1}{\sg_2}
+      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
+      & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
+      \\ \hline
+    \ecproj{\sg_r}{\mathbf{\sg_x}}
+      & Arguments $\{\sg_{r}, \mathbf{\sg_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{\sg_x}$ contains at least~2 vectors; and $\sg_r = \mathit{SUM}(\mathbf{\sg_x}, \por(\mathbf{\sg_x}))$
+      & At every position~$i$ in $1$ through $\plength(\sg_{r})$ there is exactly one vector~$v$ in $\mathbf{\sg_x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sg_{r}[i]$.
+      \\ \hline
+    \ecrank{\sg_r}{\sg_x}
+      & $\integer(\sg_{r})$; $\numeric(\sg_{x})$; and $\plength(\sg_{r}) = \plength(\sg_{x})$
+      & The values in $\sg_{r}$ represent the rank (from largest to smallest) of the values in $\sg_{x}$ (including ties)
+      \\ \hline
+    \ectotal{\sg_r}{\sg_{pos}}{\sg_{neg}}
+      & Arguments $\{\sg_{r}, \sg_{pos}, \sg_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+      & The values in $\sg_{r}$ are a running total, each value $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
+      \\ \hline
+    $\ecsumc{\sg_r}{\mathbf{\sg_x}}^\dagger$
+      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\prows(\mathbf{\sg_x}) \geq 2$; and $\pcols(\mathbf{\sg_x}) = \plength(\sg_r)$
+      & Each value in $\sg_{r}$ is obtained by summing over the corresponding column in $\mathbf{\sg_x}$.
+      \\ \hline
+    $\ecsumr{\sg_r}{\mathbf{\sg_x}}^\dagger$
+      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\pcols(\mathbf{\sg_x}) \geq 2$; and $\prows(\mathbf{\sg_x}) = \plength(\sg_r)$
+      & Each value in $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
+      \\ \hline
+    $\ecsumif{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}^\dagger$
+      & $\sg_{fk}, \sg_{pk}$ are $\discrete$; $\sg_{r}, \sg_{val}$ are $\numeric$; within the sets $\{\sg_{val}, \sg_{fk}\}$ and $\{\sg_{pk}, \sg_{r}\}$ arguments have the same $\plength$ and $\por$; $\sg_{fk}$ and $\sg_{val}$ have the same $\ptable$; $\sg_{fk}$ and $\sg_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sg_{pk}}
+      & The value for $\sg_{r}[i]$ is obtained by summing all values $\sg_{val}[j]$ where $\sg_{fk}[j] = \sg_{pk}[i]$
+      \\ \hline
+    \ecsumprod{\sg_r}{\sg_1}{\sg_2}
+      & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are all $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
+      & $\sg_{r}[i] = \sum_{i = 1}^{\plength(\sg_{1})} \sg_{1}[i] \times \sg_{2}[i]$.
       \\
 
-  \end{tabularx}
+  \end{tabularx}}
   \caption{
@@
 Most of the constraint correspond directly to spreadsheet functions and have been selected based on the most popular constraints in Excel and their occurence in spreadsheets that we have collected.
-The constraints marked with $*$ are \textit{structural}, they are discovered by our algorithm, however, they cannot be readily used in spreadsheet software.
-The constraints marked with $\dagger$ are aggregate constraints, the table shows the constraints for sum but the implementation also supports versions for max, min, average, product and count.
   \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?} 
   }
-
     % \sergey{Here a detailed explanation on what is going on, what is essential, what is not?

### 2016-05-12 Paper/ECAI/tabular.tex
Updated table

@@
 Such conflict constraints can indicate errors in the input.
-Since conflict constraints could also be accidental constraints that are correctly invalidated by the new data, the size of the data at timet~$t_0$ should be large enough.
+Since conflict constraints could also be accidental constraints that are correctly invalidated by the new data, the size of the data at time ~$t_0$ should be large enough.
 
@@
 The sample needs to large enough such that we have enough confidence that the constraints are correct.
+For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
 If we define the sample to be our data at time~$t_0$ and the added row to be the additional input at~$t_1$, this task can be formulated as online error detection and dealt with in the same way.
-Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
-In the limited case of detecting single errors one can leave out every row (column), consider the remaining data to be the sample and
-and the left out row (column) to be new input.
-\samuel{Should this be more specific?}
-
-Error detection could be integrated more closely with the algorithm, reducing the computational complexity by looking for errors during the constraint learning phase.
-This would, however, make the modeling constraints (considerably) more difficult.
-
-\samuel{FBI: work out error detection algorithm}
-
-(\sergey{I think there should be three things: key example evaluated, benchmark of 30 spreadsheets and maybe the FBI example, how we can detect interesting dependencies and correct potential mistakes})
-
-\samuel{discussion: what would be a language to extend the system, what would be the user interface?
-can it be used by community?
-can it be complitely automated?
-interaction with user and highlighing?
-}
-
-{\bfseries
-  Experimental questions
-}
-
-\begin{itemize}
-  \item  How accurate are we? (Accuracy / recall)
-  \item  How fast are we and which factors affect the runtime (how)?
-  \item  How general is our approach, what limitations are there?
-\end{itemize}
+%Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
+%In the limited case of detecting single errors one can leave out every row (column), consider the remaining data to be the sample and
+%and the left out row (column) to be new input.
 

### 2016-05-12 Paper/ECAI/tabular.tex
polishing the implementation and application sections

--- 2016-05-12 Paper/ECAI/tabular.tex
@@
 \newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
-\newcommand{\ecdiff}[3]{\eccalc{#1}{#2 - #3}}
 \newcommand{\ectotal}[3]{\eccalc{#1}{\mathit{PREV}(#1) + #2 - #3}}
@@
 \newcommand{\nat}{\mathcal{N}}
-
-\newcommand{\sg}{G'}
 
@@
 \begin{table*}
-\medskip
-Constraint templates marked with $*$ are \textit{structural}, they are discovered by our algorithm, however, they cannot be readily used in spreadsheet software.
-The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
-Only bold subgroups may contain more than one vector.
-\medskip
-
-  {\centering
+\tias{Notation inconsistent (no $G'$?
+also the 'bold' notation not explained.
+Also, in many places missing that must be vector...
+switched $\rightarrow$ to FOREIGNKEY to avoid confusion with implied constraints.}
+  \centering
   \begin{tabularx}{\textwidth}{l X X}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    $\ecalldiff{\sg_x}^*$
-      & $\discrete(\sg_x)$
-      & All values in $\sg_x$ are different: $\sg_x[i] \neq \sg_x[j]$ if $i \neq j$
+    $\ecalldiff{v_x}^*$
+      & $\discrete(v_x)$
+      & All values in $v_x$ are different: $v_x[i] \neq v_x[j]$ if $i \neq j$
       \\ \hline
     %X = Y & & \\
-    $\ecperm{\sg_x}^*$
-      & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
-      & The values in $\sg_{x}$ are a permutation of the numbers $1$ through $\plength(\sg_{x})$.
+    $\ecperm{v_x}^*$
+      & $\numeric(v_{x})$, $\ecalldiff{v_{x}}$
+      & The values in $v_{x}$ are a permutation of the numbers $1$ through $\plength(v_{x})$.
       \\ \hline
-    \ecseries{\sg_x}
-      & $\numeric(\sg_{x})$ and $\ecperm{\sg_{x}}$
-      & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
+    \ecseries{v_x}
+      & $\numeric(v_{x})$ and $\ecperm{v_{x}}$
+      & $v_{x}[1] = 1$ and $v_{x}[i] = v_{x}[i - 1] + 1$.
       \\ \hline
-    \ecfkey{\sg_{fk}}{\sg_{pk}} & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$ & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
+    \ecfkey{v_{fk}}{v_{pk}} & $v_{fk}$ and~$v_{pk}$ are both $\discrete$; $\ptype(v_{fk}) = \ptype(v_{pk})$; $\ptable(v_{fk}) \neq \ptable(v_{pk})$; and $\ecalldiff{v_{pk}}$ & Every value in~$v_{fk}$ also exist in~$v_{pk}$ \\ \hline
     % TODO check out
-    \eclookup{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
-      & $\sg_{fk}$ and $\sg_{pk}$ are both $\discrete$; arguments $\{\sg_{fk}, \sg_{r}\}$ and $\{\sg_{pk}, \sg_{val}\}$ within the same set have the same \plength, \ptable and \por; $\sg_{r}$ and~$\sg_{val}$ have the same type; and \ecfkey{\sg_{fk}}{\sg_{pk}}.
-      & $\sg_r[i]$ is the same value as looking up~$\sg_{fk}[i]$ in~$\sg_{pk}$  and returning the corresponding value in~$\sg_{val}$.
+    \eclookup{v_r}{v_{fk}}{v_{pk}}{v_{val}}
+      & $v_{fk}$ and $v_{pk}$ are both $\discrete$; arguments $\{v_{fk}, v_{r}\}$ and $\{v_{pk}, v_{val}\}$ within the same set have the same \plength, \ptable and \por; $v_{r}$ and~$v_{val}$ have the same type; and \ecfkey{v_{fk}}{v_{pk}}.
+      & $v_r[i]$ is the same value as looking up~$v_{fk}[i]$ in~$v_{pk}$  and returning the corresponding value in~$v_{val}$.
       \\ \hline
-    \eclookupfuzzy{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+    \eclookupfuzzy{v_r}{v_{fk}}{v_{pk}}{v_{val}}
       & \textit{Same as lookup}
-      & $\sg_r[i]$ is the same value as looking up the last item in~$\sg_{pk}$ smaller than~$\sg_{fk}[i]$ and returning the corresponding value in~$\sg_{val}$.
+      & $v_r[i]$ is the same value as looking up the last item in~$v_{pk}$ smaller than~$v_{fk}[i]$ and returning the corresponding value in~$v_{val}$.
       \\ \hline
-    \eclookupprod{\sg_r}{\sg_1}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{fk}\}$ are $\numeric$, arguments $\{\sg_{pk}, \sg_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sg_{fk}}{\sg_{pk}}.
-      & $\sg_{r}[i]$ is the obtained by multiplying $\sg_{1}[i]$ with $\eclookupf{\sg_{fk}}{\sg_{pk}}{\sg_{val}}{}[i]$.
+    \eclookupprod{v_r}{v_1}{v_{fk}}{v_{pk}}{v_{val}}
+      & Arguments $\{v_{r}, v_{1}, v_{fk}\}$ are $\numeric$, arguments $\{v_{pk}, v_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{v_{fk}}{v_{pk}}.
+      & $v_{r}[i]$ is the obtained by multiplying $v_{1}[i]$ with $\eclookupf{v_{fk}}{v_{pk}}{v_{val}}{}[i]$.
       \\ \hline
-    \ecprod{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$
-      & $\sg_{r}[i] = \sg_{1}[i] \times \sg_{2}[i]$.
+    \ecprod{v_r}{v_1}{v_2}
+      & Arguments $\{v_{r}, v_{1}, v_{2}\}$ are all $\numeric$ and have the same $\plength$
+      & $v_{r}[i] = v_{1}[i] \times v_{2}[i]$.
       \\ \hline
-    \ecdiff{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
-      & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
+    \ecproj{v_r}{\mathbf{v_x}}
+      & Arguments $\{v_{r}, \mathbf{v_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{v_x}$ contains at least~2 vectors; and $v_r = \mathit{SUM}(\mathbf{v_x}, \por(\mathbf{v_x}))$
+      & At every position~$i$ in $1$ through $\plength(v_{r})$ there is exactly one vector~$v$ in $\mathbf{v_x}$ such that $v[i]$ is a non-blank value, then $v[i] = v_{r}[i]$.
       \\ \hline
-    \ecproj{\sg_r}{\mathbf{\sg_x}}
-      & Arguments $\{\sg_{r}, \mathbf{\sg_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{\sg_x}$ contains at least~2 vectors; and $\sg_r = \mathit{SUM}(\mathbf{\sg_x}, \por(\mathbf{\sg_x}))$
-      & At every position~$i$ in $1$ through $\plength(\sg_{r})$ there is exactly one vector~$v$ in $\mathbf{\sg_x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sg_{r}[i]$.
+    \ecrank{v_r}{v_x}
+      & $\integer(v_{r})$; $\numeric(v_{x})$; and $\plength(v_{r}) = \plength(v_{x})$
+      & The values in $v_{r}$ represent the rank (from largest to smallest) of the values in $v_{x}$ (including ties)
       \\ \hline
-    \ecrank{\sg_r}{\sg_x}
-      & $\integer(\sg_{r})$; $\numeric(\sg_{x})$; and $\plength(\sg_{r}) = \plength(\sg_{x})$
-      & The values in $\sg_{r}$ represent the rank (from largest to smallest) of the values in $\sg_{x}$ (including ties)
+    \ectotal{v_r}{v_{pos}}{v_{neg}}
+      & Arguments $\{v_{r}, v_{pos}, v_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+      & The values in $v_{r}$ are a running total, each value $v_{r}[i] = v_{r}[i - 1] + v_{pos}[i] - v_{neg}[i]$.
       \\ \hline
-    \ectotal{\sg_r}{\sg_{pos}}{\sg_{neg}}
-      & Arguments $\{\sg_{r}, \sg_{pos}, \sg_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
-      & The values in $\sg_{r}$ are a running total, each value $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
+    $\ecsumc{v_r}{\mathbf{v_x}}^\dagger$
+      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\prows(\mathbf{v_x}) \geq 2$; and $\pcols(\mathbf{v_x}) = \plength(v_r)$
+      & Each value in $v_{r}$ is obtained by summing over the corresponding column in $\mathbf{v_x}$.
       \\ \hline
-    $\ecsumc{\sg_r}{\mathbf{\sg_x}}^\dagger$
-      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\prows(\mathbf{\sg_x}) \geq 2$; and $\pcols(\mathbf{\sg_x}) = \plength(\sg_r)$
-      & Each value in $\sg_{r}$ is obtained by summing over the corresponding column in $\mathbf{\sg_x}$.
+    $\ecsumr{v_r}{\mathbf{v_x}}^\dagger$
+      & $v_r$ and $\mathbf{v_x}$ are $\numeric$; $\pcols(\mathbf{v_x}) \geq 2$; and $\prows(\mathbf{v_x}) = \plength(v_r)$
+      & Each value in $v_{r}$ is obtained by summing over the corresponding row in $\mathbf{v_x}$.
       \\ \hline
-    $\ecsumr{\sg_r}{\mathbf{\sg_x}}^\dagger$
-      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\pcols(\mathbf{\sg_x}) \geq 2$; and $\prows(\mathbf{\sg_x}) = \plength(\sg_r)$
-      & Each value in $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
+    $\ecsumif{v_r}{v_{fk}}{v_{pk}}{v_{val}}^\dagger$
+      & $v_{fk}, v_{pk}$ are $\discrete$; $v_{r}, v_{val}$ are $\numeric$; within the sets $\{v_{val}, v_{fk}\}$ and $\{v_{pk}, v_{r}\}$ arguments have the same $\plength$ and $\por$; $v_{fk}$ and $v_{val}$ have the same $\ptable$; $v_{fk}$ and $v_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{v_{pk}}
+      & The value for $v_{r}[i]$ is obtained by summing all values $v_{val}[j]$ where $v_{fk}[j] = v_{pk}[i]$
       \\ \hline
-    $\ecsumif{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}^\dagger$
-      & $\sg_{fk}, \sg_{pk}$ are $\discrete$; $\sg_{r}, \sg_{val}$ are $\numeric$; within the sets $\{\sg_{val}, \sg_{fk}\}$ and $\{\sg_{pk}, \sg_{r}\}$ arguments have the same $\plength$ and $\por$; $\sg_{fk}$ and $\sg_{val}$ have the same $\ptable$; $\sg_{fk}$ and $\sg_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sg_{pk}}
-      & The value for $\sg_{r}[i]$ is obtained by summing all values $\sg_{val}[j]$ where $\sg_{fk}[j] = \sg_{pk}[i]$
-      \\ \hline
-    \ecsumprod{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are all $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
-      & $\sg_{r}[i] = \sum_{i = 1}^{\plength(\sg_{1})} \sg_{1}[i] \times \sg_{2}[i]$.
+    \ecsumprod{v_r}{v_1}{v_2}
+      & Arguments $\{v_r, v_1, v_2\}$ are all $\numeric$; $\plength(v_{1}) = \plength(v_{2}) \geq 2$; and $\prows(v_{r}) = \pcols(v_{r}) = 1$
+      & $v_{r}[i] = \sum_{i = 1}^{\plength(v_{1})} v_{1}[i] \times v_{2}[i]$.
       \\
 
-  \end{tabularx}}
+  \end{tabularx}
   \caption{
@@
 Most of the constraint correspond directly to spreadsheet functions and have been selected based on the most popular constraints in Excel and their occurence in spreadsheets that we have collected.
+The constraints marked with $*$ are \textit{structural}, they are discovered by our algorithm, however, they cannot be readily used in spreadsheet software.
+The constraints marked with $\dagger$ are aggregate constraints, the table shows the constraints for sum but the implementation also supports versions for max, min, average, product and count.
   \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?} 
   }
+
   % \sergey{Here a detailed explanation on what is going on, what is essential, what is not?

### 2016-05-12 Paper/ECAI/tabular.tex
removed ||

--- 2016-05-12 Paper/ECAI/tabular.tex
@@
 
-\paragraph{Parallelization}
-\tias{Do we do this? If not, cut...}
-The two for loops in the algorithm can be easily parallelized.
-Each generated assignment can be checked independently.
-Templates may not be independent as expressed by the DAG \dependencies.
-However, each independent component of the DAG (including one component per template not in the DAG) can be verified independently.
-
 \paragraph{Redundancy}

### 2016-05-12 Paper/ECAI/tabular.tex
title of the table

@@
   \end{tabular}
-  \caption{Overview evaluation on essential constraints}
+  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
 \end{table}
@@
 \section{Applications}\label{sec:applications}
-In this section we will illustrate how various motivating applications could be realized using our system.
+In this section, we illustrate how various motivating applications could be realized using our system.
 

### 2016-05-12 Paper/ECAI/tabular.tex
comments

@@
   \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
+    The graph is extracted from the Signature column in Table \ref{table:constraints}.
   }
@@
 \tias{Sumproduct is a rather peculiar constraint for spreadsheets... perhaps switch to one with a dependency, e.g. look-up?}
+\sergey{ok, we need to see if look-up would fit here}
 \\\\
@@
 \paragraph{Redundancy}
+\sergey{Samuel, probably it is better if you look here}
 Mention 'largest overlapping subgroup' here?
@@
 \subsection{Benchmark}
-There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{\samuel{link to links.txt}}.
+There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paperthe submitted paper)}.
 The case study is also included.
@@
 \samuel{Expand summary}
+\sergey{would it make sense to specify the size of a spreadsheet in the number of non-zero cells?}
 

### 2016-05-13 Paper/ECAI/tabular.tex
surprisingly difficult reworking of 4.2.3, satisf subgroups

@@
 
+\tias{Check that it is clear enough throughout the paper that: our method does not reason at the level of individual cells, but rather about \textit{vectors} of cells, as is common in spreadsheets.}
 %Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
@@
 \subsubsection{Constraints over satisfying subgroups}
-\tias{I have not yet reworked this part}
-
-\tias{can implement in any way, CSP or imperative (we: imperative)}
-$\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
-the \CSignature and \CFunction of~$t$ are satisfied).
-For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.
+In the previous step, $\generategroups(\textit{t,{\cal G},C})$ generates tuples of candidate groups $(G_1, ..., G_n)$ for the constraint template $t$ with $n$ arguments.
+In this step, $\findassignment(t, (G_1, \dots, G_n))$, the goal is to enumerate the subgroups $(G'_1, ..., G'_n), \forall i, G'_i \subseteq G_i$ that satisfy both the signature and the definition of the constraint template $t$.
+The main difference with the previous step is that the group generation reasons about the \textit{properties} of the groups (type, size, length, ...), while this step reasons about the actual \textit{content} of the vectors in the groups, that is, the data.
 
 \paragraph{Example}
-Consider a group assignment $G_x, G_y$ associated with the arguments $v_x,v_y$ of \ecrank{v_y}{v_x}, then \findassignment selects single vectors $V_y \in G_y, V_x \in G_x$ such that \ecrank{V_y}{V_x} is true.
-For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_x$ and $G_y$.
-Then \findassignment could select $V_x$ to be $\range{T_1}{\rangeall}{7}$ and $V_y$ to be $\range{T_1}{\rangeall}{8}$ since \ecrank{\range{T_1}{\rangeall}{7}}{\range{T_1}{\rangeall}{8}} is satisfied in the data.
+Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
+In the group generation, all numeric groups where the length of $\sg_r$ vector is equal to the number of rows in $\mathbf{\sg_x}$ has been enumerated.
+An example from Table~\ref{table:constraints} is $(\sg_r, \mathbf{\sg_x}) = (T_1[:,3:8],T_1[:,3:8])$.
+Note that this does not match the signature exactly yet: $\sg_r$ contains more than one vector.
+We can now generate all vectors in $T_1[:,3:8]$ for $\sg_r$ and all subgroups of $T_1[:,3:8]$ for $\mathbf{\sg_x}$ and test whether $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$ where $G[j][i]$ indicates the $i$the value of the $j$th vector in $G$.
+An example solution is $(T_1[:7], T_1[:,3:6])$.
 \\\\
-For some constraint templates (most notably aggregates) there may be multiple (overlapping) subgroups of a given group~$G$ for which the template satisfied.
-\tias{like min/max? should mention this example in implied constraints. Not sure we should mention this here, how do we remove overlapping, in post-processing? if so, move to later...}
-In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
-If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.
-
-\paragraph{Example 2}
+
+One can formulate this as a constraint satisfaction problem, similar to what we did in the previous section, or use a simple generate-and-test approach.
+In contrast to the previous step, the number of subgroups to generate here are typically manageable: $O(n*m^2)$ with $m$ the number of vectors in the largest group; we only consider continuous ranges in subgroups, so a subgroup of size $m$ has $m*(m+1)/2$ possible subgroups to consider.
+Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheet.
+
+For this reason, our current implementation uses a simple generate-and-test method that takes the required size/length of the groups into account, and sequentially verifies whether the elements of the vectors satisfy the definition.
+The pseudo-code for this approach is shown in ...\tias{Add pseudocode!!}
+
+\begin{verbatim}
+VERY PSEUDOcode:
+something recursive that handles vectors/subgroups differently like this?
+check(Groups, X, Test):
+  if |Groups| == 0:
+    forall i in ?: vector elements...
+      if not Test(X):
+        return []
+    return X
+  // else
+  all = []
+  Group = Groups.pop(0)
+  if vector(Group):
+    forall Vec in Group:
+      all += check(Groups, X + Vec, Test)
+  else:
+    forall Subgroup in Group:
+      all += check(Groups, X + Subgroup, Test)
+  return all
+\end{verbatim}
+
+Note that for some constraint definitions, there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
+\tias{How are they stopped in the code?
+should be part of the pseud-code because not part of $Test(X)$?}
+
+\paragraph{Example 2, not used (should be replaced by pseudo-code above?)}
 Sum Product
@@
 Method $Test(G_R, G_{O_1}, G_{O_2})$\\
-$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
-$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
+$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
+$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
 \\\\
 Method $FindSubgroups(Assignments)$\\
-$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$
-
-Register implementation with the system for constraint X
+$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g.
+vectors)?}
 
@@
 Mention 'largest overlapping subgroup' here?
+
+ Also, there may be overlapping subgroups that satisfy the definition, such as \tias{There is no example in the example figure? I was thinking of max/min, but it is over the entire vector always...}.
+In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
+If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.
+
+For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.
 
@@
 \samuel{Expand summary}
-\sergey{would it make sense to specify the size of a spreadsheet in the number of non-zero cells?}
 

### 2016-05-13 Paper/ECAI/tabular.tex
4.3 implementation reworked

@@
 
-\subsubsection{Constraints over satisfying subgroups}
+\subsubsection{Constraints over satisfying subgroups} \label{sec:algo:subgr}
 In the previous step, $\generategroups(\textit{t,{\cal G},C})$ generates tuples of candidate groups $(G_1, ..., G_n)$ for the constraint template $t$ with $n$ arguments.
@@
 
-\subsection{Implementation note}
+\subsection{Implementation}
+We provide some more details of our implementation of the above algorithmic approach, to ease reproducibility.
 
 \paragraph{Constraint solvers}
-\tias{stand-alone (with text-based modeling language) versus embedded.}
-For each \template, both \generategroups and \findassignment are can be seen as separate constraint satisfaction problems.
-Furthermore, declarative languages such as, for example, ASP \cite{whaisasp} and Minizinc \cite{minizinc} provide primitives that make modeling transparent and generic.
-
-Our system internally supports ASP (using Clingo 4.5.4 \cite{clingo}), Minizinc (Gecode-backend 2.0.2 \cite{minizinc}) and uses a native Python CSP solver \cite{python_constraint} for \generategroups.
-All the constraints mentioned in the paper can be run purely using the internal Python engine without any external software.
-Future versions of the system will provide user interfaces to some of these constraint languages (e.g.
-ASP and Minizinc) to allow users to specify solving procedures for new constraint templates in a declarative way.
+Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
+Constraint solvers can be categorized as being stand-alone solvers, meaning that they have a text-based modeling language in which the CSP must be formulated, and \textit{embedded}, meaning that they have a programmatic API to formulate the CSP.
+
+Our implementation supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the stand-alone Constraint Programming language MiniZinc \cite{minizinc} with the Gecode 4.4.0 solver which supports floats, as well as an embedded Python-based CP solver \cite{python_constraint}.
+The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints.
+The embedded solver is fastest as it does not have the overhead of file-writing and calling an external program, and is used in the experiments below.
 
 \paragraph{Redundancy}
-\sergey{Samuel, probably it is better if you look here}
-Mention 'largest overlapping subgroup' here?
-
- Also, there may be overlapping subgroups that satisfy the definition, such as \tias{There is no example in the example figure? I was thinking of max/min, but it is over the entire vector always...}.
-In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
-If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.
-
-For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.
+There are only types of redundancy that we eliminate during search. The first is the removal of symmetric attribute assignments as explained in Section~\ref{sec:algo:subgr}.
+
+The second is that there be \textit{overlapping} subgroups that satisfy the definition, such as \tias{There is no example in the example figure?
+I was thinking of max/min, but it is over the entire vector always...}.
+In this case we chose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
 
 \paragraph{Constraints}
+The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
+All these constraints are available in popular spreadsheet software, except the first two constraints (alldifferent and permutation).
+These two constraints are \textit{structural} constraints, which we added so they can be used in the signature of other constraint templates; they are also popular constraints in constraint satisfaction~\cite{modelseeker}.
+
+%Their dependencies are reflected in their \CSignature and shown in Figure~\ref{fig:learning_order}.
+%These constraints have been selected based on the most-used Excel functions as well as their occurrence in various example spreadsheets we observed.
+
+Note that \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?}
+
+\tias{Not used, should be cut!?:}
+Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty.
+\samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
 
 \begin{table*}
-\medskip
-Constraint templates marked with $*$ are \textit{structural}, they are discovered by our algorithm, however, they cannot be readily used in spreadsheet software.
+\caption{Overview of constraint templates implemented in our system, which covers the most popular spreadsheet constraints we encountered.
 The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
-Only bold subgroups may contain more than one vector.
-\medskip
-
+Subgroups in normal font have to be vectors, while subgroups in bold may contain more than one vector.\label{table:constraints}}
   {\centering
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    $\ecalldiff{\sg_x}^*$
+    $\ecalldiff{\sg_x}$
       & $\discrete(\sg_x)$
@@
     %X = Y & & \\
-    $\ecperm{\sg_x}^*$
+    $\ecperm{\sg_x}$
       & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
@@
   \end{tabularx}}
-  \caption{
-An overview of the different constraint templates that we have implemented.
-Most of the constraint correspond directly to spreadsheet functions and have been selected based on the most popular constraints in Excel and their occurence in spreadsheets that we have collected.
-  \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?} 
-  }
-    % \sergey{Here a detailed explanation on what is going on, what is essential, what is not?
-etc should be self-explanatory}\samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display} \sergey{1) should be readable alone without lookin in the text of the paper 2) that's the constraints we found popular according the spreadsheets in the web 3) that's what is actually implemented in the system 4) corresponds to the actual Excel functions}}
-
-  \label{table:constraints}
 \end{table*}
 
-Table~\ref{table:constraints} shows all constraint templates currently supported by our implementation.
-Their dependencies are reflected in their \CSignature and shown in Figure~\ref{fig:learning_order}.
-These constraints have been selected based on the most-used Excel functions as well as their occurrence in various example spreadsheets we observed.
-
-Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty.
-\samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
-
-\paragraph{Application specific system extension}
-\sergey{Explain here, how the system can be extended with a new constraints like vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors}
+Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}
+
+\sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}
 

### 2016-05-13 Paper/ECAI/tabular.tex
stab at making the formalization closer to ILP

@@
 \subsection{Cells, vectors and groups}
+\tias{perhaps we should not introduce cells, we reason at the level of vectors!}
 A spreadsheet may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
@@
 \end{itemize}
+\tias{Requested by Luc: formulate signature and definition as formula?}
 
@@
 
-When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
-From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
-This also holds in our setting: when we write
-      $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
-we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ of the right dimensions.
-So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
-
-\tias{Of the right (minimum) dimension?? Note that this is also sloppy because the ':8' is not considered part of the variable...},
-Alternatively, and perhaps more naturally, we can interpret the argument as the value it takes in a particular table $T$ such as $T_1$ in Figure 1, that is,
-${\range{T_1}{\rangeall}{8}} =  [1,2,3,4]$.
-Then we can also talk about constraints being satisfied. Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
-if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
-\tias{I find this confusing: a group as we defined it is defined on one table, not on multiple tables.
-we should separate the table variable from the 'range' variables then...}
-%Figure 3 provides a list of constraints that are satisfied by the tables in Figure 1.
+When looking into constraints like $\ecrank{B}{A}$, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
+
+This also holds in our setting, where an instantiation of a variable corresponds to a concrete set of vectors (a subgroup).
+For example for $\ecrank{B}{A}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B$ is the $8$th vector in $T_1$: $B = \range{T_1}{\rangeall}{8} = \{[2,1,3,4]\}$ and the value of $A$ is the $7$th vector: $A = \range{T_1}{\rangeall}{7} = \{[1514, 1551, 691, 392]\}$.
+\tias{not OK to use set because of column-wise sum? we must properly define it then...}
+
+With this interpretation, we can speak about the signature and definition of a constraint template being satisfied.
+We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by an assignment $(A_1, ..., A_n)$ if $Sig_c(A_1, ..., A_n)$ (respectively $Def_c(A_1, ..., A_n)$) is satisfied.
+Likewise, an assignment $(A_1, ..., A_n)$ satisfies a constraint template $s$ iff $|(A_1, ..., A_n)| = n_s \wedge Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$.
+
+\tias{Luc will probably prefer: $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is abused to be both the constraint name and the template}.
 

### 2016-05-13 Paper/ECAI/tabular.tex
minor changes in first two sections

--- 2016-05-13 Paper/ECAI/tabular.tex
@@
 Constraint learning techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
-New types of constraints can be easily added to the system by specifying them using in a declarative language.
+New types of constraints can be easily added to the system by specifying them in a declarative language.
 We show that our approach is able to accurately discover constraints in a number of spreadsheets from varying sources.
@@
 One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
-Furthermore, it is not only challenging for a machine but for humans as well.
+Furthermore, it is not only challenging for a machine but for humans as well, e.g.
+the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' turned to be bogus due an excel formula mistake \cite{flaw_excel}.
 
@@
 
-The question that we answer in this paper is: whether it is possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose algorithm and system for discovering such constraints.
+The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose algorithm and the system, named \sname (pronounced as the word ``tackle''), for discovering such constraints.
 Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias") and covers both textual and numeric fields.
@@
 We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch several application scenario's, such as autocompletion and error detection.
+Moreover, we sketch several application scenario's such as autocompletion and error detection.
 \\\\
@@
 Section \ref{sec:evaluation} presents the evaluation of the approach.
-Section \ref{sec:applications} shows how our system can be used for applications such as autocompletion and error detection.
-Section \ref{sec:related_work} present an overview of the related work.
-And Section \ref{sec:conclusions} provides conclusions.
-
-\tias{Check that it is clear enough throughout the paper that: our method does not reason at the level of individual cells, but rather about \textit{vectors} of cells, as is common in spreadsheets.}
+Section \ref{sec:applications} shows how our system can be used for applications.
+Section \ref{sec:related_work} presents an overview of the related work.
+Section \ref{sec:conclusions} provides conclusions.
+
 %Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
@@
 A spreadsheet may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
-We will describe a method of extracting tables from a big spreadsheet in Section~\ref{todo}, however this is not the core of our learning method.
+We will describe a method of extracting tables from a big spreadsheet in Section~\ref{sec:approach}, however this is not the core of our learning method.
 Our formalization reasons over tables directly.
@@
 A cell has a {\em type}, which can be numeric or textual.
-We further distinguish numeric types in subtypes integer and float.
+We further distinguish numeric types in subtypes: integer and float.
 We also consider \textit{None} as a special type when a cell is empty; \textit{None} is a subtype of all other types.
-A set of cells is called \textit{type-consistent} iff all cells in the set are of the same type.
+A set of cells is called \textit{type-consistent} iff all cells in the set are of the same type (\textit{None} is consistent with everything).
 %Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric subtype by requiring its arguments to be integers.
@@
 For example, the signature can restrict the types of the arguments to be groups of integers and require that the length of all arguments be equal.
-\\
 In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
@@
 \begin{itemize}
-\item \CName: \textit{$B$ = RANK($A$)}, where $B$ and $A$ are the arguments;
-\item \CSignature: the two arguments $B$ and $A$ must be numeric groups and of size 1; furthermore, the group $B$ must be integer (the rank) and the length of groups $A$ and $B$ must be identical, i.e., $length(B) = length(A)$;
-\item \CFunction: \textit{$B$ = RANK($A$)} whenever each value of $B$ is the rank of the corresponding value in $A$.
+  \item \CName: \ecrank{G_r}{G_x}, where $G_r$ and $G_x$ are the arguments;
+\item \CSignature: the two arguments $G_r$ and $G_x$ must be numeric groups and of size one; furthermore, the group $G_r$ must be integer (the rank) and the length of groups $G_x$ and $G_r$ must be identical, i.e., $length(G_r) = length(G_x)$;
+\item \CFunction: \ecrank{G_r}{G_x}, whenever each value of $G_r$ is the rank of the corresponding value in $G_x$.
 % is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
@@
 
-When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
-From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
+When looking into constraints like \ecrank{G_r}{G_x}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+From a FOL perspective, the name of the constraint (\textit{RANK}) is just the predicate, and the arguments $G_r$ and $G_x$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
 This also holds in our setting: when we write
@@
 ${\range{T_1}{\rangeall}{8}} =  [1,2,3,4]$.
-Then we can also talk about constraints being satisfied. Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
+Then we can also talk about constraints being satisfied. Basically, a constraint $c(G_1, ..., G_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
 if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
@@
 \subsubsection{Constraints over satisfying subgroups}
-In the previous step, $\generategroups(\textit{t,{\cal G},C})$ generates tuples of candidate groups $(G_1, ..., G_n)$ for the constraint template $t$ with $n$ arguments.
-In this step, $\findassignment(t, (G_1, \dots, G_n))$, the goal is to enumerate the subgroups $(G'_1, ..., G'_n), \forall i, G'_i \subseteq G_i$ that satisfy both the signature and the definition of the constraint template $t$.
-The main difference with the previous step is that the group generation reasons about the \textit{properties} of the groups (type, size, length, ...), while this step reasons about the actual \textit{content} of the vectors in the groups, that is, the data.
+\tias{I have not yet reworked this part}
+
+\tias{can implement in any way, CSP or imperative (we: imperative)}
+$\findassignment(t,a,C_{solutions})$ finds, for all assignments~$a = (G_1, ..., G_n)$ in $A$, the subgroups $G_i' \subseteq G_i$ such that the corresponding constraint $t(G_1', ..., G_n')$ is satisfied (i.e.
+the \CSignature and \CFunction of~$t$ are satisfied).
+For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.
 
 \paragraph{Example}
-Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
-In the group generation, all numeric groups where the length of $\sg_r$ vector is equal to the number of rows in $\mathbf{\sg_x}$ has been enumerated.
-An example from Table~\ref{table:constraints} is $(\sg_r, \mathbf{\sg_x}) = (T_1[:,3:8],T_1[:,3:8])$.
-Note that this does not match the signature exactly yet: $\sg_r$ contains more than one vector.
-We can now generate all vectors in $T_1[:,3:8]$ for $\sg_r$ and all subgroups of $T_1[:,3:8]$ for $\mathbf{\sg_x}$ and test whether $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$ where $G[j][i]$ indicates the $i$the value of the $j$th vector in $G$.
-An example solution is $(T_1[:7], T_1[:,3:6])$.
+Consider a group assignment $G_x, G_y$ associated with the arguments $v_x,v_y$ of \ecrank{v_y}{v_x}, then \findassignment selects single vectors $V_y \in G_y, V_x \in G_x$ such that \ecrank{V_y}{V_x} is true.
+For example, in Figure~\ref{fig:main_example} the group $G = \range{T_1}{\rangeall}{\rangeto{3}{8}}$ can serve as both $G_x$ and $G_y$.
+Then \findassignment could select $V_x$ to be $\range{T_1}{\rangeall}{7}$ and $V_y$ to be $\range{T_1}{\rangeall}{8}$ since \ecrank{\range{T_1}{\rangeall}{7}}{\range{T_1}{\rangeall}{8}} is satisfied in the data.
 \\\\
-
-One can formulate this as a constraint satisfaction problem, similar to what we did in the previous section, or use a simple generate-and-test approach.
-In contrast to the previous step, the number of subgroups to generate here are typically manageable: $O(n*m^2)$ with $m$ the number of vectors in the largest group; we only consider continuous ranges in subgroups, so a subgroup of size $m$ has $m*(m+1)/2$ possible subgroups to consider.
-Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheet.
-
-For this reason, our current implementation uses a simple generate-and-test method that takes the required size/length of the groups into account, and sequentially verifies whether the elements of the vectors satisfy the definition.
-The pseudo-code for this approach is shown in ...\tias{Add pseudocode!!}
-
-\begin{verbatim}
-VERY PSEUDOcode:
-something recursive that handles vectors/subgroups differently like this?
-check(Groups, X, Test):
-  if |Groups| == 0:
-    forall i in ?: vector elements...
-      if not Test(X):
-        return []
-    return X
-  // else
-  all = []
-  Group = Groups.pop(0)
-  if vector(Group):
-    forall Vec in Group:
-      all += check(Groups, X + Vec, Test)
-  else:
-    forall Subgroup in Group:
-      all += check(Groups, X + Subgroup, Test)
-  return all
-\end{verbatim}
-
-Note that for some constraint definitions, there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-\tias{How are they stopped in the code?
-should be part of the pseud-code because not part of $Test(X)$?}
-
-\paragraph{Example 2, not used (should be replaced by pseudo-code above?)}
+For some constraint templates (most notably aggregates) there may be multiple (overlapping) subgroups of a given group~$G$ for which the template satisfied.
+\tias{like min/max? should mention this example in implied constraints. Not sure we should mention this here, how do we remove overlapping, in post-processing? if so, move to later...}
+In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
+If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.
+
+\paragraph{Example 2}
 Sum Product
@@
 Method $Test(G_R, G_{O_1}, G_{O_2})$\\
-$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
-$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
+$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\\
+$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$
 \\\\
 Method $FindSubgroups(Assignments)$\\
-$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g.
-vectors)?}
+$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$
+
+Register implementation with the system for constraint X
 
@@
 Mention 'largest overlapping subgroup' here?
-
- Also, there may be overlapping subgroups that satisfy the definition, such as \tias{There is no example in the example figure? I was thinking of max/min, but it is over the entire vector always...}.
-In this case we choose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
-If the returned subsets are too large and accidentally include unrelated vectors the user can split such a group.
-
-For every \template~$t$ an implementation (declarative or imperative) needs to be provided that, given an assignment, finds valid subgroups.
 
@@
 \paragraph{Application specific system extension}
+\sergey{under question now, might worth deleting} \\
 \sergey{Explain here, how the system can be extended with a new constraints like vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors}

### 2016-05-13 Paper/ECAI/tabular.tex

--- 2016-05-13 Paper/ECAI/tabular.tex
@@
 Section \ref{sec:evaluation} presents the evaluation of the approach.
-Section \ref{sec:applications} shows how our system can be used for applications such as autocompletion and error detection.
-Section \ref{sec:related_work} present an overview of the related work.
-And Section \ref{sec:conclusions} provides conclusions.
+Section \ref{sec:applications} shows how our system can be used for applications.
+Section \ref{sec:related_work} presents an overview of the related work.
+Section \ref{sec:conclusions} provides conclusions.
 
@@
 \item
-\textit{\CName}  specifies the syntactic form of the constraint $c(A_1, ...,A_n)$, that is, the name of the constraint $c$ together
-with $n$ arguments $A_i$.
+\textit{\CName}  specifies the syntactic form of the constraint $c(G_1, ...,G_n)$, that is, the name of the constraint $c$ together
+with $n$ arguments $G_i$.
 Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
-Note that a function $B=f(A_1,...,A_n)$ can be represented with the $n+1$-ary predicate $cf(B,A_1,...,A_n)$.
+Note that a function $G_r=f(G_1,...,G_n)$ can be represented with the $n+1$-ary predicate $cf(G_r,G_1,...,G_n)$.
 % the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
@@
 \paragraph{Example}
-Several example constraints are illustrated in Table 1.  For example, the constraint template \textit{rank} has
+Several example constraints are illustrated in Table \ref{table:constraints}.  For example, the constraint template \textit{rank} has
 \begin{itemize}
-\item \CName: \textit{$B$ = RANK($A$)}, where $B$ and $A$ are the arguments;
-\item \CSignature: the two arguments $B$ and $A$ must be numeric groups and of size 1; furthermore, the group $B$ must be integer (the rank) and the length of groups $A$ and $B$ must be identical, i.e., $length(B) = length(A)$;
-\item \CFunction: \textit{$B$ = RANK($A$)} whenever each value of $B$ is the rank of the corresponding value in $A$.
+  \item \CName: \ecrank{G_r}{G_x}, where $G_r$ and $G_x$ are the arguments;
+\item \CSignature: the two arguments $G_r$ and $G_x$ must be numeric groups and of size one; furthermore, the group $G_r$ must be integer (the rank) and the length of groups $G_x$ and $G_r$ must be identical, i.e., $\plength(G_r) = \plength(G_x)$;
+\item \CFunction: \ecrank{G_r}{G_x}, whenever each value of $G_r$ is the rank of the corresponding value in $G_x$.
 % is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
 \end{itemize}
-
-When looking into constraints like $\ecrank{B}{A}$, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
-From a FOL perspective, the name of the constraint ($RANK$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
+When looking into constraints like $\ecrank{G_r}{G_x}$, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+From a FOL perspective, the name of the constraint (\textit{RANK}) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
 
 This also holds in our setting, where an instantiation of a variable corresponds to a concrete set of vectors (a subgroup).
-For example for $\ecrank{B}{A}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B$ is the $8$th vector in $T_1$: $B = \range{T_1}{\rangeall}{8} = \{[2,1,3,4]\}$ and the value of $A$ is the $7$th vector: $A = \range{T_1}{\rangeall}{7} = \{[1514, 1551, 691, 392]\}$.
+For example for $\ecrank{G_r}{G_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $G_r$ is the $8$th vector in $T_1$: $G_r = \range{T_1}{\rangeall}{8} = \{[2,1,3,4]\}$ and the value of $G_x$ is the $7$th vector: $G_x = \range{T_1}{\rangeall}{7} = \{[1514, 1551, 691, 392]\}$.
 \tias{not OK to use set because of column-wise sum? we must properly define it then...}
 
+\newcommand{\sigc}{\ensuremath{\format{Sig}_c}}
+\newcommand{\defc}{\ensuremath{\format{Def}_c}}
+
 With this interpretation, we can speak about the signature and definition of a constraint template being satisfied.
-We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by an assignment $(A_1, ..., A_n)$ if $Sig_c(A_1, ..., A_n)$ (respectively $Def_c(A_1, ..., A_n)$) is satisfied.
-Likewise, an assignment $(A_1, ..., A_n)$ satisfies a constraint template $s$ iff $|(A_1, ..., A_n)| = n_s \wedge Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$.
+We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by an assignment $(G_1, ..., G_n)$ if $\sigc(G_1, ..., G_n)$ (respectively $\defc(G_1, ..., G_n)$) is satisfied.
+Likewise, an assignment $(G_1, ..., G_n)$ satisfies a constraint template $s$ iff $|(G_1, ..., G_n)| = n_s \wedge \sigc(G_1, ..., G_n) \wedge \defc(G_1, ..., G_n)$.
 
@@
 
+\sergey{under question now, might worth deleting} \\
 \sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}

### 2016-05-13 Paper/ECAI/tabular.tex
moved on of the figures

@@
 \usepackage{tabularx}
+\usepackage{subcaption}
 
@@
 
-\begin{figure*}[tbh]
+\begin{figure*}[thb]
+
+  \begin{subfigure}{.70\textwidth}
   \begin{center}
-    \includegraphics[width=0.75\textwidth]{figures/Demo.png}
+    \includegraphics[width=1\textwidth]{figures/Demo.png}
   \end{center}
@@
   \label{fig:main_example}
+\end{subfigure}
+\hfill
+\begin{subfigure}{.30\textwidth}
+  {\small
+    \begin{align*}
+      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
+%
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
+%
+      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+      &~\ecaggc{AVERAGE}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecaggc{MAX}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecaggc{MIN}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecaggif{SUM}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+%
+      &~\ecaggif{MAX}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+%
+      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
+%
+      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
+    \end{align*}}
+  \vspace{-10pt}
+  \caption{Constraints that are satisfied in the case study (Figure on the left) and expected to be found by our implementation. Functional constraint are omitted.}
+  \label{fig:sol_example}
+\end{subfigure}
 \end{figure*}
@@
 
-Depending on the application, a solution to the tabular constraint learning problem may contain \textit{redundant} constraints, cf.
-\cite{luc_book}.
+Depending on the application, a solution to the tabular constraint learning problem may contain \textit{redundant} constraints, for details we refer to relational learning literature \cite{luc_book}.
 This is because constraints may be logically equivalent or may be implied by other constraints.
@@
 Given that the problem is to find the set of {\em all} constraints, and that the more general constraint is entailed
-by the more specific one, the $ALLDDIFFERENT(G)$ constraint will be {\em redundant}.
+by the more specific one, the $\ecdiff{G}$ constraint will be {\em redundant}.
 So, for most applications, it will be sufficient to output only the most specific one.
@@
 
-\begin{figure}
-  {\small
-    \begin{align*}
-      & SERIES(\range{T_{1}}{\rangeall}{1}) \\
-%
-      & \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
-%
-      & \ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      & \ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-      & \ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-      & \ecaggc{AVERAGE}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      & \ecaggc{MAX}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      & \ecaggc{MIN}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      & \ecaggif{SUM}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
-%
-      & \ecaggif{MAX}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
-%
-      & \eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
-%
-      & \range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
-    \end{align*}
-  }
-  \caption{Constraints that are satisfied in the case study (Figure~\ref{fig:main_example}) and expected to be found by our implementation. Functional constraint are omitted.}
-  \label{fig:sol_example}
-\end{figure}
-
 \subsubsection{Results}

### 2016-05-13 Paper/ECAI/tabular.tex
chaning minor
Clarifications

--- 2016-05-13 Paper/ECAI/tabular.tex
@@
 Constraint learning techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
-New types of constraints can be easily added to the system by specifying them in a declarative language.
+New types of constraints can be easily added to the system by specifying them using in a declarative language.
 We show that our approach is able to accurately discover constraints in a number of spreadsheets from varying sources.
@@
 One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
-Furthermore, it is not only challenging for a machine but for humans as well, e.g.
-the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' turned to be bogus due an excel formula mistake \cite{flaw_excel}.
+Furthermore, it is not only challenging for a machine but for humans as well.
 
@@
 
-The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose algorithm and the system, named \sname (pronounced as the word ``tackle''), for discovering such constraints.
+The question that we answer in this paper is: whether it is possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose algorithm and system for discovering such constraints.
 Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias") and covers both textual and numeric fields.
@@
 We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch several application scenario's such as autocompletion and error detection.
+Moreover, we sketch several application scenario's, such as autocompletion and error detection.
 \\\\
@@
 Section \ref{sec:evaluation} presents the evaluation of the approach.
-Section \ref{sec:applications} shows how our system can be used for applications.
-Section \ref{sec:related_work} presents an overview of the related work.
-Section \ref{sec:conclusions} provides conclusions.
+Section \ref{sec:applications} shows how our system can be used for applications such as autocompletion and error detection.
+Section \ref{sec:related_work} present an overview of the related work.
+And Section \ref{sec:conclusions} provides conclusions.
 
@@
 A spreadsheet may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
-We will describe a method of extracting tables from a big spreadsheet in Section~\ref{sec:approach}, however this is not the core of our learning method.
+We will describe a method of extracting tables from a big spreadsheet in Section~\ref{todo}, however this is not the core of our learning method.
 Our formalization reasons over tables directly.
@@
 A cell has a {\em type}, which can be numeric or textual.
-We further distinguish numeric types in subtypes: integer and float.
+We further distinguish numeric types in subtypes integer and float.
 We also consider \textit{None} as a special type when a cell is empty; \textit{None} is a subtype of all other types.
-A set of cells is called \textit{type-consistent} iff all cells in the set are of the same type (\textit{None} is consistent with everything).
+A set of cells is called \textit{type-consistent} iff all cells in the set are of the same type.
 %Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric subtype by requiring its arguments to be integers.
@@
   \item
-  A \textbf{subgroup} $G'$ of a group $G$ is simply a projection of $G$ onto a consecutive subset its vectors.
+  A \textbf{subgroup} $G'$ of a group $G$ is simply a projection of $G$ onto a consecutive subset of its vectors.
 % is a subrange of vectors belonging to a group $G$.
@@
 For example, the signature can restrict the types of the arguments to be groups of integers and require that the length of all arguments be equal.
+\\
 In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
@@
 \begin{itemize}
-  \item \CName: \ecrank{G_r}{G_x}, where $G_r$ and $G_x$ are the arguments;
-\item \CSignature: the two arguments $G_r$ and $G_x$ must be numeric groups and of size one; furthermore, the group $G_r$ must be integer (the rank) and the length of groups $G_x$ and $G_r$ must be identical, i.e., $length(G_r) = length(G_x)$;
-\item \CFunction: \ecrank{G_r}{G_x}, whenever each value of $G_r$ is the rank of the corresponding value in $G_x$.
+\item \CName: \textit{$B$ = RANK($A$)}, where $B$ and $A$ are the arguments;
+\item \CSignature: the two arguments $B$ and $A$ must be numeric groups and of size 1; furthermore, the group $B$ must be integer (the rank) and the length of groups $A$ and $B$ must be identical, i.e., $length(B) = length(A)$;
+\item \CFunction: \textit{$B$ = RANK($A$)} whenever each value of $B$ is the rank of the corresponding value in $A$.
 % is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
@@
 
-When looking into constraints like \ecrank{G_r}{G_x}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
-From a FOL perspective, the name of the constraint (\textit{RANK}) is just the predicate, and the arguments $G_r$ and $G_x$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
+When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+From a FOL perspective, the name of the constraint ($\mathit{RANK}$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
 This also holds in our setting: when we write
       $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
-we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ of the right dimensions.
+we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ (provided~$T$ has an 8th column).
 So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
 
-\tias{Of the right (minimum) dimension?? Note that this is also sloppy because the ':8' is not considered part of the variable...},
+\samuel{\tias{Of the right (minimum) dimension??
+Note that this is also sloppy because the ':8' is not considered part of the variable...} True, I changed the wording a bit but personally I would leave it at this now, you can always say you look at $T[:,i:j]$ iff $cols(T) >= max(i,j)$}.
 Alternatively, and perhaps more naturally, we can interpret the argument as the value it takes in a particular table $T$ such as $T_1$ in Figure 1, that is,
-${\range{T_1}{\rangeall}{8}} =  [1,2,3,4]$.
-Then we can also talk about constraints being satisfied. Basically, a constraint $c(G_1, ..., G_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only
-if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
-\tias{I find this confusing: a group as we defined it is defined on one table, not on multiple tables.
-we should separate the table variable from the 'range' variables then...}
+${\range{T_1}{\rangeall}{8}} =  [2,1,3,4]$.
+Then we can also talk about constraints being satisfied.
+Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
+\samuel{\tias{I find this confusing: a group as we defined it is defined on one table, not on multiple tables.
+we should separate the table variable from the 'range' variables then...} waiting for Luc's opinion here}
 %Figure 3 provides a list of constraints that are satisfied by the tables in Figure 1.
@@
 \luc{Do we really need this, or can we write it in a real equivalence --- conjunctive formula?}
+\samuel{Tias, Sergey what do you think? Also I find this notation confusing: +(G2, G3), why not just use G1 = G2 + G3?}
 
@@
 
-\tias{I am not defining a new problem for 'most specific canonical tabular constraint learning' because we loosely showed how implications and equivalences would be expressed, but not how to express canonicity and we did not claim that most specific is always best.}
-\tias{Note also that our method finds ALL constraints; it uses specificity to faster traverse the search space, but the resulting set of constraints found would be the same, whether or not using the specificity (see 4.2.1).}
+%\tias{I am not defining a new problem for 'most specific canonical tabular constraint learning' because we loosely showed how implications and equivalences would be expressed, but not how to express canonicity and we did not claim that most specific is always best.}
+%\tias{Note also that our method finds ALL constraints; it uses specificity to faster traverse the search space, but the resulting set of constraints found would be the same, whether or not using the specificity (see 4.2.1).}
 
@@
 
-From each table, row-oriented and column-oriented groups are then extracted by partitioning the rows (columns) into type-consistent parts. This step can also be verified by a user in the visual tool.
+Row-oriented and column-oriented groups are then extracted from each table by partitioning the rows (columns) into type-consistent parts.
 
@@
 \paragraph{Application specific system extension}
-\sergey{under question now, might worth deleting} \\
 \sergey{Explain here, how the system can be extended with a new constraints like vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors}

### 2016-05-13 Paper/ECAI/tabular.tex

--- 2016-05-13 Paper/ECAI/tabular.tex
@@
 \newcommand{\eccalc}[2]{\ensuremath{#1 = #2}}
-\newcommand{\ecrank}[2]{\eccalc{#1}{\mathit{RANK}(#2)}}
-\newcommand{\ecfkey}[2]{\ensuremath{\mathit{FOREIGNKEY}(#1,#2)}}
-\newcommand{\ecalldiff}[1]{\ensuremath{\mathit{ALLDIFFERENT}(#1)}}
-\newcommand{\eclookupf}[4]{\ensuremath{\mathit{LOOKUP}_{\mathit{#4}}(#1, #2, #3)}}
+\newcommand{\ecrank}[2]{\eccalc{#1}{\textit{RANK}(#2)}}
+\newcommand{\ecfkey}[2]{\ensuremath{\textit{FOREIGNKEY}(#1,#2)}}
+\newcommand{\ecalldiff}[1]{\ensuremath{\textit{ALLDIFFERENT}(#1)}}
+\newcommand{\eclookupf}[4]{\ensuremath{\textit{LOOKUP}_{\textit{#4}}(#1, #2, #3)}}
 \newcommand{\eclookup}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{}}}
@@
 \newcommand{\eclookupfuzzy}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{fuzzy}}}
-\newcommand{\ecperm}[1]{\ensuremath{\mathit{PERMUTATION}(#1)}}
-\newcommand{\ecseries}[1]{\ensuremath{\mathit{SERIES}(#1)}}
+\newcommand{\ecperm}[1]{\ensuremath{\textit{PERMUTATION}(#1)}}
+\newcommand{\ecseries}[1]{\ensuremath{\textit{SERIES}(#1)}}
 \newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
 \newcommand{\ecdiff}[3]{\eccalc{#1}{#2 - #3}}
-\newcommand{\ectotal}[3]{\eccalc{#1}{\mathit{PREV}(#1) + #2 - #3}}
-\newcommand{\ecproj}[2]{\eccalc{#1}{\mathit{PROJECT}(#2)}}
-\newcommand{\ecaggc}[3]{\eccalc{#2}{\mathit{#1}(#3, col)}}
-\newcommand{\ecaggr}[3]{\eccalc{#2}{\mathit{#1}(#3, row)}}
-\newcommand{\ecsumc}[2]{\eccalc{#1}{\mathit{SUM}(#2, col)}}
-\newcommand{\ecsumr}[2]{\eccalc{#1}{\mathit{SUM}(#2, row)}}
-\newcommand{\ecaggif}[5]{\eccalc{#2}{\mathit{#1IF}(#3, #4, #5)}}
-\newcommand{\ecsumif}[4]{\eccalc{#1}{\mathit{SUMIF}(#2, #3, #4)}}
-\newcommand{\ecsumprod}[3]{\eccalc{#1}{\mathit{SUMPRODUCT}(#2, #3)}}
+\newcommand{\ectotal}[3]{\eccalc{#1}{\textit{PREV}(#1) + #2 - #3}}
+\newcommand{\ecproj}[2]{\eccalc{#1}{\textit{PROJECT}(#2)}}
+\newcommand{\ecaggc}[3]{\eccalc{#2}{\textit{#1}(#3, col)}}
+\newcommand{\ecaggr}[3]{\eccalc{#2}{\textit{#1}(#3, row)}}
+\newcommand{\ecsumc}[2]{\eccalc{#1}{\textit{SUM}(#2, col)}}
+\newcommand{\ecsumr}[2]{\eccalc{#1}{\textit{SUM}(#2, row)}}
+\newcommand{\ecaggif}[5]{\eccalc{#2}{\textit{#1IF}(#3, #4, #5)}}
+\newcommand{\ecsumif}[4]{\eccalc{#1}{\textit{SUMIF}(#2, #3, #4)}}
+\newcommand{\ecsumprod}[3]{\eccalc{#1}{\textit{SUMPRODUCT}(#2, #3)}}
 
@@
 %
-      &~\ecaggc{AVERAGE}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggc{MAX}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggc{MIN}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggif{SUM}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
-%
-      &~\ecaggif{MAX}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+%
+      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
 %
@@
 
+%When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+%From a FOL perspective, the name of the constraint ($\mathit{RANK}$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
+%This also holds in our setting: when we write
+%      $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
+%we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ (provided~$T$ has an 8th column).
+%So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
+
 \newcommand{\sigc}{\ensuremath{\format{Sig}_c}}
@@
 \tias{Luc will probably prefer: $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is abused to be both the constraint name and the template}.
+%\samuel{\tias{Of the right (minimum) dimension??
+Note that this is also sloppy because the ':8' is not considered part of the variable...} True, I changed the wording a bit but personally I would leave it at this now, you can always say you look at $T[:,i:j]$ iff $cols(T) >= max(i,j)$}.
+%Alternatively, and perhaps more naturally, we can interpret the argument as the value it takes in a particular table $T$ such as $T_1$ in Figure 1, that is,
+%${\range{T_1}{\rangeall}{8}} =  [2,1,3,4]$.
+%Then we can also talk about constraints being satisfied.
+Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
+\samuel{\tias{I find this confusing: a group as we defined it is defined on one table, not on multiple tables.
+we should separate the table variable from the 'range' variables then...} waiting for Luc's opinion here}
+%Figure 3 provides a list of constraints that are satisfied by the tables in Figure 1.
 

### 2016-05-13 Paper/ECAI/tabular.tex
added preliminary figure

--- 2016-05-13 Paper/ECAI/tabular.tex
@@
 \subsection{Autocompletion}
+\begin{figure}[thb]
+  \begin{center}
+    \includegraphics[width=0.33\textwidth]{figures/autocompletion_example.png}
+  \end{center}
+  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
+  \label{fig:autocompletion_example}
+\end{figure}
+
 Autocompletion can be seen as using knowledge derived from data at time~$t_0$ and new input data at time~$t_1 > t_0$ to predict data that the user has not yet written.
@@
 
+Let us illustrate this on one of the spreadsheets from the gathered dataset. In Figure \ref{fig:autocompletion_example}, \sergey{to be finished}
+
+\subsection{Error detection}
 \begin{figure*}[thb]
@@
 
-\subsection{Error detection}
 We consider two cases of error detection.

### 2016-05-13 Paper/ECAI/tabular.tex
Added figure, comment

@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{figures/tabletool.png}
+  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
+    The graph is extracted from the Signature column in Table \ref{table:constraints}.
+  }
+  \label{fig:learning_order}
+\end{figure}
+
+\samuel{This does not fully correspond with the real tool, the real tool does not allow detected tables to be corrected: Either use detection tool or use visual tool to input tables manually}
 The way the table detection in the tool works is as follows:
@@
 Hence, in the visual tool the user can correct the detected tables if needed.
-\tias{I think it would be good if we could show a screenshot of this tool...}
+\samuel{\tias{I think it would be good if we could show a screenshot of this tool...} Added screenshot, needs to be referred to}
 

### 2016-05-13 Paper/ECAI/tabular.tex
Corrected algorithm
Cosmetic

@@
     \State $C \gets \emptyset$ %\Comment{The set of constraints}
-  \ForAll{$t \in \constrainttorder(\constraints,\dependencies)$}
-  \State $n \gets$ number of arguments of constraint $c_t$
-    \ForAll{$(G_1, \dots, G_n) \in \generategroups(t, \groups, C)$}
-      \State $C \gets C \cup \findassignment(t, (G_1, \dots, G_n))$
+  \ForAll{$s \in \constrainttorder(\constraints,\dependencies)$}
+  \State $n \gets$ number of arguments of template $s$
+    \ForAll{$(G_1, \dots, G_n) \in \generategroups(s, \groups, C)$}
+      \State $C \gets C \cup \findassignment(s, (G_1, \dots, G_n))$
     \EndFor
@@
 
-The $\constrainttorder(\constraints,\dependencies)$ module in our method takes a set of {\template}s and a DAG as input and produces a linear ordering of the templates such that the more general, implied constraint templates (the requirements) are ordered before the templates that depend on them.
+The $\constrainttorder(\constraints,\dependencies)$ module in our method takes a set of {\template}s~(\constraints) and a DAG~(\dependencies) as input and produces a linear ordering of the templates such that the more general, implied constraint templates (the requirements) are ordered before the templates that depend on them.
 

### 2016-05-13 Paper/ECAI/tabular.tex
Cosmetic
Showed bold

@@
 VERY PSEUDOcode:
-something recursive that handles vectors/subgroups differently like this?
+something recursive that handles vectors/
+subgroups differently like this?
 check(Groups, X, Test):
@@
 The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
-Subgroups in normal font have to be vectors, while subgroups in bold may contain more than one vector.\label{table:constraints}}
+Subgroups in normal font have to be vectors, while subgroups in \textbf{bold} may contain more than one vector.\label{table:constraints}}
   {\centering

### 2016-05-13 Paper/ECAI/tabular.tex
Comment
Corrections group generation

@@
 
-%Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
-%For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that a group~$G \in \groups$ is selected if there is at least one subgroup~$G' \subseteq G$ that satisfies the \CSignature.
-
 A Constraint Satisfaction Problem (CSP) is a triple $(V,D,C)$ where $V$ is a set of \textit{variables}, $D$ the domain of possible values each variable can take and $C$ the set of constraints over $V$ that must be satisfied.
@@
 
-To reason over the groups, we automatically add to the constraint program a set of background facts that contain the properties of each group, namely its type, size and length.
+%Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
+%For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that a group~$G \in \groups$ is selected if there is at least one subgroup~$G' \subseteq G$ that satisfies the \CSignature.
+
+To reason over the groups, we automatically add to the constraint program a set of background facts that contain the properties of each group, namely its type, size, number of columns and rows, length, orientation, and table.
 We also add all previously found constraints of constraint templates that are implied by this template.
 The actual constraints then express the properties defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
-\tias{More detailed description of the properties commented out, feel free to put in}
+% \tias{More detailed description of the properties commented out, feel free to put in}
 %Translation:
@@
 
-% TODO: old text, rephrase a bit
 Constraints derived from the signature can usually but not always be enforced exactly on a group, as we have to consider whether subgroups could satisfy them.
+Requirements on the lengths, orientations and tables of groups can be directly enforced since they stay the same for subgroups.
+Typing requirements need to be relaxed to consider only non-subtypes, since, for example, numeric groups might contain integer subgroups.
+Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since subgroups might contains fewer vectors.
+Finally, exact restrictions on the number of rows or columns may have to be relaxed depending on the orientation of the group.
+Subgroups of row-oriented (column-oriented) group will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
+Therefore, we can only require row-oriented (column-oriented) groups to contain \textit{at least} the required amount of rows (columns).
+
 For example, the \CSignature of ?? requires $v_r$ to be integer.
@@
 Note that for some constraint definitions, there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-\tias{How are they stopped in the code?
-should be part of the pseud-code because not part of $Test(X)$?}
+\samuel{\tias{How are they stopped in the code?
+should be part of the pseud-code because not part of $Test(X)$?} will make it part of pseudo-code}
 

### 2016-05-13 Code/src/core/assignment.py
Added comment (CODE)

@@ -51,6 +51,8 @@ class Source:
     def _complete(self, assignments, groups, filters):
         result = []
 
+        # TODO type as constraint
+
         def try_assignment():
             problem = Problem()
 

### 2016-05-17 Paper/ECAI/tabular.tex
revise abstract

@@
 Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
-Constraint learning techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
-New types of constraints can be easily added to the system by specifying them in a declarative language.
-We show that our approach is able to accurately discover constraints in a number of spreadsheets from varying sources.
+Constraint learning techniques are used to identify predicates (constraints) that hold between blocks of rows and columns of the tables.
+%New types of constraints can be easily added to the system by specifying them in a declarative language.
+We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
 \end{abstract}
@@
 Especially when spreadsheets have been exported from
-other software such as ERP packages, it is hard to understand the data as in this case,
-often a comma-separated values (CSV) format is used, meaning that all formula's are lost, including inter-sheet formulas and relations.
+other software such as Enterprise Resource Planning (ERP) systems; in this case, %, it is hard to understand the data as in this case,
+often a comma-separated values (CSV) format is used meaning that all formula's are lost, including inter-sheet formulas and relations.
 %This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
+Even in manually created spreadsheets, it can be challenging to be consistent and correct with formula's across big spreadsheets.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain some mistakes in formulae.
 
@@
 Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
-This already provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns.
+This provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns.
 Here everything is mixed.
-The data is relational on the one hand, since we have multiple table with relationships between them.
-And on the other hand, the data is mixed textual and numeric, since people tend to use spreadsheets for financial and accounting computations.
+The data is relational on the one hand, since we have multiple tables with relationships between them.
+And on the other hand, the data is most often mixed textual and numeric.
+%, since people tend to use spreadsheets for financial and accounting computations.
 One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
-Furthermore, it is not only challenging for a machine but for humans as well, e.g.
-the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' turned to be bogus due an excel formula mistake \cite{flaw_excel}.
-
-The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
-the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to conjunctive constraint learning have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
+
+The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where
+the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
 and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from
-very few examples.  Our approach contrasts with all of the mentioned approaches (except Modelseeker) in that it focusses
-on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
+very few examples.  Our approach contrasts with these in that it focusses
+on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional CSPs.
+It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
 A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.
@@
 The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose algorithm and the system, named \sname (pronounced as the word ``tackle''), for discovering such constraints.
-Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias") and covers both textual and numeric fields.
-It works in an unsupervised setting, as only the tabular data is provided and no examples need to be specified, and discovers column- and row-wise constraints.
+To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
+Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias").
+\tias{revise declarative claim?} It operates directly on the tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given.
+%, and discovers column- and row-wise constraints.
 We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch several application scenario's such as autocompletion and error detection.
+Moreover, we sketch additional application scenario's such as autocompletion and error detection.
 \\\\
@@
 
-\tias{Check that it is clear enough throughout the paper that: our method does not reason at the level of individual cells, but rather about \textit{vectors} of cells, as is common in spreadsheets.}
 %Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
@@
   \vspace{-10pt}
-  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations.}
+  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations. \tias{TODO: headers not in green boxes}}
   \label{fig:main_example}

### 2016-05-17 Paper/ECAI/tabular.tex
refactor first half of section 2, formalization

@@
   \vspace{-10pt}
-  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations. \tias{TODO: headers not in green boxes}}
+  \caption{Example spreadsheet (black words and numbers only).
+Green borders indicate headerless tables; the blue elements are added to clarify some constraints and relations.
+\tias{TODO: headers not in green boxes} \sergey{show blocks?
+orientation only over blocks, not over tables?}\tias{Use specific constraint names, e.g.
+'SUM-row' above Total in T1}}
   \label{fig:main_example}
@@
   \vspace{-10pt}
-  \caption{Constraints that are satisfied in the case study (Figure on the left) and expected to be found by our implementation. Functional constraint are omitted.}
+  \caption{Constraints that are satisfied in the case study (Figure on the left) and expected to be found by our implementation. Non-functional constraint are omitted \tias{Add them too}.}
   \label{fig:sol_example}
@@
 \section{Formalization}\label{sec:formalization}
-As we focus on spreadsheet data, we first introduce the general concepts of cells, group and constraint templates, using the (multi-table) sheet in Figure~\ref{fig:main_example} as running example.
-
-\subsection{Cells, vectors and groups}
-\tias{perhaps we should not introduce cells, we reason at the level of vectors!}
-A spreadsheet may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
-We will describe a method of extracting tables from a big spreadsheet in Section~\ref{sec:approach}, however this is not the core of our learning method.
-Our formalization reasons over tables directly.
-
-Formally, a table is an $n \times m$ matrix. Each entry is called a \textit{cell}.
-A cell has a {\em type}, which can be numeric or textual.
+Our goal is to automatically discover constraints (functions and relations) between the rows and the columns of tables in a spreadsheet.
+This is applicable not just to data from spreadsheets, but any data in tabular form, hence the name.
+
+%An example spreadsheet with tables is given in Figure~\ref{fig:main_example}, we will use this as running example throughout the paper.
+
+\subsection{Terminology}
+Spreadsheets and tabular data may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
+Note that a table can contain a header; however, we wish to reason over entire rows and columns of data, and hence we will consider \textbf{headerless tables} only.
+
+Formally, a (headerless) table is an $n \times m$ matrix. Each entry is called a \textit{cell}.
+A cell has a {\bf type}, which can be numeric or textual.
 We further distinguish numeric types in subtypes: integer and float.
 We also consider \textit{None} as a special type when a cell is empty; \textit{None} is a subtype of all other types.
-A set of cells is called \textit{type-consistent} iff all cells in the set are of the same type (\textit{None} is consistent with everything).
-%Certain constraints, such as \textit{rank} or \textit{series}, make use of the numeric subtype by requiring its arguments to be integers.
-
-We consider various sets of cells with the following notation:
-\begin{itemize}
-  \item {\bf Cell}: We use the notation $T[a,b]$ to refer to a particular cell in the table~$T$ at the row with index~$a$ and the column with index~$b$.
-
-  \item {\bf Columns and rows}: For a table $T$, we use notation $T[a,{:}]$ to refer to the $a$-th row, and similarly $T[{:},a]$ for the $a$-th column.
-% to refer to a subrange from $a$ to $b$ we use $T[a{:}b,]$ ($T[,a{:}b]$).
-
-  \item
-  A \textbf{vector} is a type-consistent row or column.
-  If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation. The \textit{length} of a vector is the number of cells in that vector.
-
-  \item
-  A \textbf{group} is a projection of a matrix onto a set of consecutive rows or columns. %In this work we only consider consecutive rows and columns.
-  We use the following notation to refer to a row-oriented group with rows ranging from $a$ to $b$: $G = T[a{:}b,:]$, where $a,b$ are natural numbers.
-Similarly, we have $G = T[{:},a{:}b]$ for a column-oriented group.
-  The vectors in a group have to be type-consistent, consecutive in the original table and of equal length.
-  The \textit{size} of a group is the number of vectors in the group, while the \textit{length} of a group is the length of its vectors. 
-  %The \textit{length} of a row (column) group $G$, written as \textit{length(G)}, is the number of its columns (rows); We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
-
-  \item
-  A \textbf{subgroup} $G'$ of a group $G$ is simply a projection of $G$ onto a consecutive subset of its vectors.
-% is a subrange of vectors belonging to a group $G$.
-If the subgroup $g$ must contain only one vector, we write $g \in G$.
-%It defines what it means for a constraint to be well-formed
-%We refer to a subgroup of a group $G$ in \groups as $g$. A \textit{subgroup} $g$ is a subrange of vectors in a group $G$. If a subgroup $g$ must contain only one vector, we write $g \in G$.
-% A \textit{subgroup} $G'$ is a subrange of vectors belonging to a group $G$. If the subgroup $\{v\}$ must contain only one vector, we write $v \in G$.
-\end{itemize}
+
+A row or columns is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual.
+We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$, and similarly $T[{:},a]$ for the $a$-th column.
+For example in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_5[:,1] = ['$Diana Coolen$', 5]$.
+The latter is not type-consistent while the former is.
+We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.
+
+The most important concept is that of a \textbf{block}.
+A block has three properties: 1) contains only entire rows or entire columns of a single headerless table 2) it is contiguous and 3) it is type-consistent.
+If it contains only rows we say it has \textit{row-orientation}, if only columns, \textit{column-orientation}.
+The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type.
+
+Because of these properties, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
+%
+The \textit{size} of a block is the number of vectors in the block, e.g.
+$b-a+1$.
+The \textit{length} of a block is the length of its vectors; as all vectors are from the same table, they always have the same length.
+%We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
+
 %For notational convenience, we will refer to vectors in {\rm roman} and to groups in {\bf bold}. {\bf CHECK}
 
-%When we do not want to distinguish between cells, columns, rows, vectors or groups, we shall talk about a {\em set of cells}.
-
 \paragraph{Example}
-Consider Table~1 in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
-Ignoring the (all-textual) header, Table~1 can be divided into the following five column-oriented groups:
+Consider headerless Table~1 in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e.
+they contain both numeric and textual data).
+The table can be partitioned into the following five column-oriented blocks:
 {\small
 \begin{align*}
-&G_1 = \range{T_1}{\rangeall}{1},
-&G_2 = \range{T_1}{\rangeall}{2},\\
-&G_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
-&G_4 = \range{T_1}{\rangeall}{9},\\
-&G_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}.
+&B_1 = \range{T_1}{\rangeall}{1},
+&B_2 = \range{T_1}{\rangeall}{2},\\
+&B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
+&B_4 = \range{T_1}{\rangeall}{9},\\
+&B_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}.
 \end{align*}
+\tias{Indicate in figure? see also Sergey's comment in figure}
 }
-An example subgroup is $\range{T_1}{\rangeall}{\rangeto{3}{6}} \subset G_3$, which contains the sales numbers of all employees for the four quarters.
+
+We write that a block $B'$ is \textbf{contained} in a block $B$, $B' \sqsubseteq B$ iff both are valid blocks (single orientation, contiguous, type consistent) and each of the vectors in $B'$ is also in $B$.
+For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=T[\rangeto{a}{b},:] \wedge B'=T[\rangeto{a'}{b'},:] \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
+We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
+
+An example is that $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.
+
+\tias{TODO: update everything to speak about blocks instead of groups}
 
@@
 \end{itemize}
-\tias{Requested by Luc: formulate signature and definition as formula?}
 
@@
 % is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
+
+\tias{Maybe better row-sum so that actual 'blocks' is clear instead of just vectors? (and then use same row-sum for all subsequent examples}
 \end{itemize}
+
 When looking into constraints like $\ecrank{G_r}{G_x}$, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
@@
 This also holds in our setting, where an instantiation of a variable corresponds to a concrete set of vectors (a subgroup).
-For example for $\ecrank{G_r}{G_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $G_r$ is the $8$th vector in $T_1$: $G_r = \range{T_1}{\rangeall}{8} = \{[2,1,3,4]\}$ and the value of $G_x$ is the $7$th vector: $G_x = \range{T_1}{\rangeall}{7} = \{[1514, 1551, 691, 392]\}$.
-\tias{not OK to use set because of column-wise sum? we must properly define it then...}
+For example for $\ecrank{G_r}{G_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $G_r$ is the $8$th vector in $T_1$: $G_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $G_x$ is the $7$th vector: $G_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
 
@@
 \newcommand{\defc}{\ensuremath{\format{Def}_c}}
-
+\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
 With this interpretation, we can speak about the signature and definition of a constraint template being satisfied.
@@
 
-\tias{Luc will probably prefer: $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is abused to be both the constraint name and the template}.
 %\samuel{\tias{Of the right (minimum) dimension??

### 2016-05-17 Paper/ECAI/tabular.tex
start of new intro section 4, have to go now

@@
 \section{Approach to Tabular Constraint Learning}\label{sec:approach}
+The aim of our method is to detect constraints between rows and columns of tabular data.
+Recall that a valid argument assignment for a constraint is an assignment of each argument of the constraint to a block, such that the constraint, that is, its signature and definition, is satisfied.
+\tias{I here use constraint and constraint template interchangeably}
+
+Our proposed methodology contains the following steps:
+\begin{enumerate}
+\item Extract headerless tables from tabular data
+\item Partition the tables into type-consistent contiguous \textit{superblocks}
+\item Generate for each constraint all valid argument assignments:
+\begin{enumerate}
+\item For each constraint, generate all candidate \textit{super-assignments} by selecting compatible superblocks for each argument
+\item For each candidate super-assignment, generate all valid argument assignments from the superblocks.
+\end{enumerate}
+\end{enumerate}
+
+The core of our method is step 3.
+In principle, one could use a generate-and-test approach by generating all possible blocks from the superblock and testing each combination of blocks for each of the arguments of the constraints.
+However, this ...
+\textit{blows up} ...
+
+Hence, we divide this step into two parts: in the first part, we will reason over entire superblocks by generating all possible combinations of superblocks that...
+
+\paragraph{Rest old text}
+
 This section describes our approach to \tcl.

### 2016-05-17 Paper/ECAI/tabular.tex
Add comment

--- 2016-05-17 Paper/ECAI/tabular.tex
@@
 Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
-Constraint learning techniques are used to identify predicates (constraints) that hold between blocks of rows and columns of the tables.
-%New types of constraints can be easily added to the system by specifying them in a declarative language.
-We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
+Constraint learning techniques are used to identify predicates (constraints) that hold between the rows, columns and submatrices of the tables.
+New types of constraints can be easily added to the system by specifying them in a declarative language.
+We show that our approach is able to accurately discover constraints in a number of spreadsheets from varying sources.
 \end{abstract}
@@
 Especially when spreadsheets have been exported from
-other software such as Enterprise Resource Planning (ERP) systems; in this case, %, it is hard to understand the data as in this case,
-often a comma-separated values (CSV) format is used meaning that all formula's are lost, including inter-sheet formulas and relations.
+other software such as ERP packages, it is hard to understand the data as in this case,
+often a comma-separated values (CSV) format is used, meaning that all formula's are lost, including inter-sheet formulas and relations.
 %This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
-Even in manually created spreadsheets, it can be challenging to be consistent and correct with formula's across big spreadsheets.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain some mistakes in formulae.
 
@@
 Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
-This provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns.
+This already provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns.
 Here everything is mixed.
-The data is relational on the one hand, since we have multiple tables with relationships between them.
-And on the other hand, the data is most often mixed textual and numeric.
-%, since people tend to use spreadsheets for financial and accounting computations.
+The data is relational on the one hand, since we have multiple table with relationships between them.
+And on the other hand, the data is mixed textual and numeric, since people tend to use spreadsheets for financial and accounting computations.
 One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
-
-The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where
-the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
+Furthermore, it is not only challenging for a machine but for humans as well, e.g.
+the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' turned to be bogus due an excel formula mistake \cite{flaw_excel}.
+
+The approach that we will introduce will borrow techniques from logical and relational learning \cite{luc_book}, where
+the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to conjunctive constraint learning have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
 and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
-very few examples.  Our approach contrasts with these in that it focusses
-on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional CSPs.
-It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of function is learned in spreadsheet data from
+very few examples.  Our approach contrasts with all of the mentioned approaches (except Modelseeker) in that it focusses
+on learning both column- and row-constraints, it contrasts with Flashfill in learning constraints rather than functions.
 A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.
@@
 The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
-Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias").
-\tias{revise declarative claim?} It operates directly on the tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given.
-%, and discovers column- and row-wise constraints.
+To answer this question we contribute a general-purpose algorithm and the system, named \sname (pronounced as the word ``tackle''), for discovering such constraints.
+Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias") and covers both textual and numeric fields.
+It works in an unsupervised setting, as only the tabular data is provided and no examples need to be specified, and discovers column- and row-wise constraints.
 We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch additional application scenario's such as autocompletion and error detection.
+Moreover, we sketch several application scenario's such as autocompletion and error detection.
 \\\\
@@
   \vspace{-10pt}
-  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations. \tias{TODO: headers not in green boxes}}
+  \caption{Example spreadsheet (black words and numbers only). Blue and green elements are added by us to clarify the relations.}
   \label{fig:main_example}
@@
 \section{Formalization}\label{sec:formalization}
+\samuel{Make explicit that we consider vectors as minimal unit (maybe in intro already?)}
 As we focus on spreadsheet data, we first introduce the general concepts of cells, group and constraint templates, using the (multi-table) sheet in Figure~\ref{fig:main_example} as running example.

### 2016-05-17 Paper/ECAI/tabular.tex
remove comment (and stray paragraph) from samuel again: should be fixed
form: small tweak vector -> row/column on first use in block
further rework of Section 4 intro

--- 2016-05-17 Paper/ECAI/tabular.tex
@@
 The \textit{size} of a block is the number of vectors in the block, e.g.
-$b-a+1$.
+$b-a+1$, the number of rows or columns.
 The \textit{length} of a block is the length of its vectors; as all vectors are from the same table, they always have the same length.
@@
 
-Hence, we divide this step into two parts: in the first part, we will reason over entire superblocks by generating all possible combinations of superblocks that...
+Hence, we divide this step into two parts: in the first part, we will not reason over individual blocks, but rather over all possible combinations of superblocks as candidates for the arguments.
+\tias{Example To develop in more detail and clarity:}
+Consider the sum constraints, the row-blocks of T1 are ...
+and column-blocks are ...
+instead of considering all possible blocks in each super-block, we reason at level of superblocks, for example block (with ID) can not be in a row-sum and neither can (block with Items sold total, Max items sold) because its size (1 and 2) is smaller than length of blocks (4).
+The only possible combination of superblocks is (1st Quarter ..
+Rank) for both arguments.
+
+In the second part, we start from such a superblock combination and generate and test all possible block assignments to arguments.
+As we only have to consider the blocks contained in each superblock, this is typically much easier.
+For the above example, \tias{to develop with more clarity} each of the vectors in the block will be considered as candidate for the left-hand side, and one could enumerate all subblocks for the right-hand side and verify for each of the rows in the vector.
+In practice, this can be optimized further.
+\\\\
+We now describe in more detail how the headerless tables are extracted and how the superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3).
+We then describe a number of optimizations that we perform to speed up step 3 and 4 (Section 4.4).
 

### 2016-05-19 Paper/ECAI/tabular.tex
switch to IEEEtran for DSAA
more on the formalisation

@@
-\documentclass{sig-alternate-05-2015}
+\documentclass{IEEEtran}
 
@@
 
-%\usepackage{amsthm} % incompatible with ACM
-%\theoremstyle{definition}
-%\newtheorem{definition}{Definition}
-\newdef{definition}{Definition} % ACM specific
+\usepackage{amsthm} % incompatible with ACM
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}
+\newtheorem{example}{Example}
+%\newdef{definition}{Definition} % ACM specific
 
@@
   \vspace{-10pt}
-  \caption{Constraints that are satisfied in the case study (Figure on the left) and expected to be found by our implementation. Non-functional constraint are omitted \tias{Add them too}.}
+  \caption{Constraints present in the running example (left) \tias{Add ALL, including non-functional and duplicates}.}
   \label{fig:sol_example}
@@
 %An example spreadsheet with tables is given in Figure~\ref{fig:main_example}, we will use this as running example throughout the paper.
+
+\tias{section description (add prob def here?)}
 
@@
 The latter is not type-consistent while the former is.
-We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.
+%We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.
 
 The most important concept is that of a \textbf{block}.
-A block has three properties: 1) contains only entire rows or entire columns of a single headerless table 2) it is contiguous and 3) it is type-consistent.
+A block has to satisfy three conditions: 1) it contains only entire rows or entire columns of a single headerless table 2) it is contiguous and 3) it is type-consistent.
 If it contains only rows we say it has \textit{row-orientation}, if only columns, \textit{column-orientation}.
 The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type.
+We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independent of their orientation.
 
 Because of these properties, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
-%
-The \textit{size} of a block is the number of vectors in the block, e.g.
-$b-a+1$, the number of rows or columns.
-The \textit{length} of a block is the length of its vectors; as all vectors are from the same table, they always have the same length.
+
+A block has the following properties:
+\begin{itemize}
+\item \textit{type}: a block has to be type-consistent, so it has a single type
+\item \textit{table}: the table that this block belongs to
+\item \textit{orientation}: either row-oriented or column-oriented
+\item \textit{size}: the size of a block is the number of vectors in it, e.g. $b-a+1$, the number of rows or columns.
+\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
+\item \textit{rows}: this is orientation specific, for row-based blocks this is the size, for column-based blocks the length
+\item \textit{columns}: also orientation specific, for row-based blocks this is the length, for column-based blocks the size
+\item \tias{What is missing is 'discrete'}
+\end{itemize}
+
 %We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
@@
 
-\tias{TODO: update everything to speak about blocks instead of groups}
-
 \subsection{Constraint templates}
-Our method will extract the applicable constraints (including functions) in a set of tables, based on \textit{constraint templates}.
-%
-A \template is a triple \textit{(\CName, \CSignature, \CFunction)}:
+The goal is to learn constraints over blocks in the data. The knowledge needed to learn a constraint is expressed through {\template}s.
+%
+A \template $s$ is a triple $s=\textit{(\CName, \CSignature, \CFunction)}$:
 %Let us elaborate on this:
@@
 \textit{\CName}  specifies the syntactic form of the constraint $c(G_1, ...,G_n)$, that is, the name of the constraint $c$ together
-with $n$ arguments $G_i$.
+with $n$ abstract arguments $G_i$.
 Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
-Note that a function $G_r=f(G_1,...,G_n)$ can be represented with the $n+1$-ary predicate $cf(G_r,G_1,...,G_n)$.
+Note that a function $G_r=f(G_1,...,G_n)$ can be represented with the $(n{+}1)$-ary predicate $c_f(G_r,G_1,...,G_n)$.
+Each argument will have to be instantiated with a block.
 % the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
 
-\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy. That is, the required types of the arguments, together
-with possible further restrictions, e.g., on their length.
-For example, the signature can restrict the types of the arguments to be groups of integers and require that the length of all arguments be equal.
-In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint is well-formed.
+\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy, that is, the blocks.
+This can be any property of individual blocks, such as type, length, size and orientation, as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table, or have equal type or length.
+In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
 
 \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
-Given an assignments to its arguments, it can be used to verify whether the constraint is satisfied or not.
+Given an assignments of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not.
 %holds, in practice this will be a function that can be called when
 %the arguments are full instantiated to decide whether the constraint is satisfied or not.
-In logical and relational learning this is known as the background knowledge, which provides the definition of the predicate.
+In logical and relational learning this is known as the background knowledge. %, which provides the definition of the predicate.
 \end{itemize}
 
-\paragraph{Example}
-Several example constraints are illustrated in Table \ref{table:constraints}.  For example, the constraint template \textit{rank} has
+\begin{example}
+Several example constraint templates are illustrated in Table \ref{table:constraints}. An non-trivial example is the constraint template for the row-based sum:
+
 \begin{itemize}
-  \item \CName: \ecrank{G_r}{G_x}, where $G_r$ and $G_x$ are the arguments;
-\item \CSignature: the two arguments $G_r$ and $G_x$ must be numeric groups and of size one; furthermore, the group $G_r$ must be integer (the rank) and the length of groups $G_x$ and $G_r$ must be identical, i.e., $\plength(G_r) = \plength(G_x)$;
-\item \CFunction: \ecrank{G_r}{G_x}, whenever each value of $G_r$ is the rank of the corresponding value in $G_x$.
-% is satisfied in the mapping $\{ v_x \mapsto g_x, v_y \mapsto g_y \}$ satisfies the constraint iff $g_x \in G_x, g_y \in G_y$ ($G_x \in \groups, G_y \in \groups$) and each value in $g_y$ is the rank of the corresponding value in $g_x$ (possibly with ties).
-
-\tias{Maybe better row-sum so that actual 'blocks' is clear instead of just vectors? (and then use same row-sum for all subsequent examples}
+  \item \CName: $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, where $\sg_r$ and $\mathbf{\sg_x}$ are the arguments;
+
+  \item \CSignature: $\sg_r$ has to be a single vector ($size=1$) while $\mathbf{\sg_x}$ can be a block ($size>=1$), which can be derived from the use of a normal or \textbf{bold} font.
+The two blocks have to be numeric.
+This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$.
+While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint;
+
+  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
 \end{itemize}
+\end{example}
 
@@
 
-%\sergey{moved from the formalization BEGIN}
-%l 1) a set of cells, and 2) a set of constraints over those cells.
-Notice that the constraints specify the types (or the domains) that one encounters in traditional CSPs.
-Notice also, that as in traditional CSPs, the CSP problem defines the variables (the cells) of the problem as well as the constraint that should hold amongst them.
-The task of CSP-solvers is then to find an assignment to the variables that satisfies all the constraints.
-The distinction between variables (cells) and their assignments (their values) is important here.
-So, when we speak about cells, vectors and groups, we refer to the variables; when we talk about values, assignments or instances, we refer to the values these cells take in a particular "instantiated" table.
+\begin{table*}
+\caption{Overview of constraint templates implemented in our system, which covers the most popular spreadsheet constraints we encountered.
+The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
+Subgroups in normal font have to be vectors, while subgroups in \textbf{bold} may contain more than one vector.\label{table:constraints}}
+  {\centering
+  \begin{tabularx}{\textwidth}{l X X}
+    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
+    $\ecalldiff{\sg_x}$
+      & $\discrete(\sg_x)$
+      & All values in $\sg_x$ are different: $\sg_x[i] \neq \sg_x[j]$ if $i \neq j$
+      \\ \hline
+    %X = Y & & \\
+    $\ecperm{\sg_x}$
+      & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
+      & The values in $\sg_{x}$ are a permutation of the numbers $1$ through $\plength(\sg_{x})$.
+      \\ \hline
+    \ecseries{\sg_x}
+      & $\numeric(\sg_{x})$ and $\ecperm{\sg_{x}}$
+      & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
+      \\ \hline
+    \ecfkey{\sg_{fk}}{\sg_{pk}} & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$ & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
+    % TODO check out
+    \eclookup{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+      & $\sg_{fk}$ and $\sg_{pk}$ are both $\discrete$; arguments $\{\sg_{fk}, \sg_{r}\}$ and $\{\sg_{pk}, \sg_{val}\}$ within the same set have the same \plength, \ptable and \por; $\sg_{r}$ and~$\sg_{val}$ have the same type; and \ecfkey{\sg_{fk}}{\sg_{pk}}.
+      & $\sg_r[i]$ is the same value as looking up~$\sg_{fk}[i]$ in~$\sg_{pk}$  and returning the corresponding value in~$\sg_{val}$.
+      \\ \hline
+    \eclookupfuzzy{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+      & \textit{Same as lookup}
+      & $\sg_r[i]$ is the same value as looking up the last item in~$\sg_{pk}$ smaller than~$\sg_{fk}[i]$ and returning the corresponding value in~$\sg_{val}$.
+      \\ \hline
+    \eclookupprod{\sg_r}{\sg_1}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{fk}\}$ are $\numeric$, arguments $\{\sg_{pk}, \sg_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sg_{fk}}{\sg_{pk}}.
+      & $\sg_{r}[i]$ is the obtained by multiplying $\sg_{1}[i]$ with $\eclookupf{\sg_{fk}}{\sg_{pk}}{\sg_{val}}{}[i]$.
+      \\ \hline
+    \ecprod{\sg_r}{\sg_1}{\sg_2}
+      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$
+      & $\sg_{r}[i] = \sg_{1}[i] \times \sg_{2}[i]$.
+      \\ \hline
+    \ecdiff{\sg_r}{\sg_1}{\sg_2}
+      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
+      & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
+      \\ \hline
+    \ecproj{\sg_r}{\mathbf{\sg_x}}
+      & Arguments $\{\sg_{r}, \mathbf{\sg_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{\sg_x}$ contains at least~2 vectors; and $\sg_r = \mathit{SUM}(\mathbf{\sg_x}, \por(\mathbf{\sg_x}))$
+      & At every position~$i$ in $1$ through $\plength(\sg_{r})$ there is exactly one vector~$v$ in $\mathbf{\sg_x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sg_{r}[i]$.
+      \\ \hline
+    \ecrank{\sg_r}{\sg_x}
+      & $\integer(\sg_{r})$; $\numeric(\sg_{x})$; and $\plength(\sg_{r}) = \plength(\sg_{x})$
+      & The values in $\sg_{r}$ represent the rank (from largest to smallest) of the values in $\sg_{x}$ (including ties)
+      \\ \hline
+    \ectotal{\sg_r}{\sg_{pos}}{\sg_{neg}}
+      & Arguments $\{\sg_{r}, \sg_{pos}, \sg_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+      & The values in $\sg_{r}$ are a running total, each value $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
+      \\ \hline
+    $\ecsumr{\sg_r}{\mathbf{\sg_x}}^\dagger$
+      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\pcols(\mathbf{\sg_x}) \geq 2$; and $\prows(\mathbf{\sg_x}) = \plength(\sg_r)$
+      & Each value in $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
+      \\ \hline
+    $\ecsumc{\sg_r}{\mathbf{\sg_x}}^\dagger$
+      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\prows(\mathbf{\sg_x}) \geq 2$; and $\pcols(\mathbf{\sg_x}) = \plength(\sg_r)$
+      & Each value in $\sg_{r}$ is obtained by summing over the corresponding column in $\mathbf{\sg_x}$.
+      \\ \hline
+    $\ecsumif{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}^\dagger$
+      & $\sg_{fk}, \sg_{pk}$ are $\discrete$; $\sg_{r}, \sg_{val}$ are $\numeric$; within the sets $\{\sg_{val}, \sg_{fk}\}$ and $\{\sg_{pk}, \sg_{r}\}$ arguments have the same $\plength$ and $\por$; $\sg_{fk}$ and $\sg_{val}$ have the same $\ptable$; $\sg_{fk}$ and $\sg_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sg_{pk}}
+      & The value for $\sg_{r}[i]$ is obtained by summing all values $\sg_{val}[j]$ where $\sg_{fk}[j] = \sg_{pk}[i]$
+      \\ \hline
+    \ecsumprod{\sg_r}{\sg_1}{\sg_2}
+      & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are all $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
+      & $\sg_{r}[i] = \sum_{i = 1}^{\plength(\sg_{1})} \sg_{1}[i] \times \sg_{2}[i]$.
+      \\
+
+  \end{tabularx}}
+\end{table*}
 
@@
 
-\begin{table*}
-\caption{Overview of constraint templates implemented in our system, which covers the most popular spreadsheet constraints we encountered.
-The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
-Subgroups in normal font have to be vectors, while subgroups in \textbf{bold} may contain more than one vector.\label{table:constraints}}
-  {\centering
-  \begin{tabularx}{\textwidth}{l X X}
-    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    $\ecalldiff{\sg_x}$
-      & $\discrete(\sg_x)$
-      & All values in $\sg_x$ are different: $\sg_x[i] \neq \sg_x[j]$ if $i \neq j$
-      \\ \hline
-    %X = Y & & \\
-    $\ecperm{\sg_x}$
-      & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
-      & The values in $\sg_{x}$ are a permutation of the numbers $1$ through $\plength(\sg_{x})$.
-      \\ \hline
-    \ecseries{\sg_x}
-      & $\numeric(\sg_{x})$ and $\ecperm{\sg_{x}}$
-      & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
-      \\ \hline
-    \ecfkey{\sg_{fk}}{\sg_{pk}} & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$ & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
-    % TODO check out
-    \eclookup{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
-      & $\sg_{fk}$ and $\sg_{pk}$ are both $\discrete$; arguments $\{\sg_{fk}, \sg_{r}\}$ and $\{\sg_{pk}, \sg_{val}\}$ within the same set have the same \plength, \ptable and \por; $\sg_{r}$ and~$\sg_{val}$ have the same type; and \ecfkey{\sg_{fk}}{\sg_{pk}}.
-      & $\sg_r[i]$ is the same value as looking up~$\sg_{fk}[i]$ in~$\sg_{pk}$  and returning the corresponding value in~$\sg_{val}$.
-      \\ \hline
-    \eclookupfuzzy{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
-      & \textit{Same as lookup}
-      & $\sg_r[i]$ is the same value as looking up the last item in~$\sg_{pk}$ smaller than~$\sg_{fk}[i]$ and returning the corresponding value in~$\sg_{val}$.
-      \\ \hline
-    \eclookupprod{\sg_r}{\sg_1}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{fk}\}$ are $\numeric$, arguments $\{\sg_{pk}, \sg_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sg_{fk}}{\sg_{pk}}.
-      & $\sg_{r}[i]$ is the obtained by multiplying $\sg_{1}[i]$ with $\eclookupf{\sg_{fk}}{\sg_{pk}}{\sg_{val}}{}[i]$.
-      \\ \hline
-    \ecprod{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$
-      & $\sg_{r}[i] = \sg_{1}[i] \times \sg_{2}[i]$.
-      \\ \hline
-    \ecdiff{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
-      & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
-      \\ \hline
-    \ecproj{\sg_r}{\mathbf{\sg_x}}
-      & Arguments $\{\sg_{r}, \mathbf{\sg_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{\sg_x}$ contains at least~2 vectors; and $\sg_r = \mathit{SUM}(\mathbf{\sg_x}, \por(\mathbf{\sg_x}))$
-      & At every position~$i$ in $1$ through $\plength(\sg_{r})$ there is exactly one vector~$v$ in $\mathbf{\sg_x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sg_{r}[i]$.
-      \\ \hline
-    \ecrank{\sg_r}{\sg_x}
-      & $\integer(\sg_{r})$; $\numeric(\sg_{x})$; and $\plength(\sg_{r}) = \plength(\sg_{x})$
-      & The values in $\sg_{r}$ represent the rank (from largest to smallest) of the values in $\sg_{x}$ (including ties)
-      \\ \hline
-    \ectotal{\sg_r}{\sg_{pos}}{\sg_{neg}}
-      & Arguments $\{\sg_{r}, \sg_{pos}, \sg_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
-      & The values in $\sg_{r}$ are a running total, each value $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
-      \\ \hline
-    $\ecsumc{\sg_r}{\mathbf{\sg_x}}^\dagger$
-      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\prows(\mathbf{\sg_x}) \geq 2$; and $\pcols(\mathbf{\sg_x}) = \plength(\sg_r)$
-      & Each value in $\sg_{r}$ is obtained by summing over the corresponding column in $\mathbf{\sg_x}$.
-      \\ \hline
-    $\ecsumr{\sg_r}{\mathbf{\sg_x}}^\dagger$
-      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\pcols(\mathbf{\sg_x}) \geq 2$; and $\prows(\mathbf{\sg_x}) = \plength(\sg_r)$
-      & Each value in $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
-      \\ \hline
-    $\ecsumif{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}^\dagger$
-      & $\sg_{fk}, \sg_{pk}$ are $\discrete$; $\sg_{r}, \sg_{val}$ are $\numeric$; within the sets $\{\sg_{val}, \sg_{fk}\}$ and $\{\sg_{pk}, \sg_{r}\}$ arguments have the same $\plength$ and $\por$; $\sg_{fk}$ and $\sg_{val}$ have the same $\ptable$; $\sg_{fk}$ and $\sg_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sg_{pk}}
-      & The value for $\sg_{r}[i]$ is obtained by summing all values $\sg_{val}[j]$ where $\sg_{fk}[j] = \sg_{pk}[i]$
-      \\ \hline
-    \ecsumprod{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are all $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
-      & $\sg_{r}[i] = \sum_{i = 1}^{\plength(\sg_{1})} \sg_{1}[i] \times \sg_{2}[i]$.
-      \\
-
-  \end{tabularx}}
-\end{table*}
-
 Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}

### 2016-05-19 Paper/ECAI/tabular.tex
update steps based on feedback from Luc

@@
 \item Extract headerless tables from tabular data
-\item Partition the tables into type-consistent contiguous \textit{superblocks}
+\item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
 \item Generate for each constraint all valid argument assignments:
 \begin{enumerate}
-\item For each constraint, generate all candidate \textit{super-assignments} by selecting compatible superblocks for each argument
-\item For each candidate super-assignment, generate all valid argument assignments from the superblocks.
+\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal \textit{superblock} and the $B_i$ are compatible with the signature.
+\item For each generated atom $c(B_1, \ldots ,B_n)$, find all valid $c(B'_1, \ldots, B'_n)$ such that $\forall i, B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ is satisfied.
 \end{enumerate}

### 2016-05-19 Paper/ECAI/tabular.tex
reworked section 2

@@
 \newcommand{\ecproj}[2]{\eccalc{#1}{\textit{PROJECT}(#2)}}
-\newcommand{\ecaggc}[3]{\eccalc{#2}{\textit{#1}(#3, col)}}
-\newcommand{\ecaggr}[3]{\eccalc{#2}{\textit{#1}(#3, row)}}
-\newcommand{\ecsumc}[2]{\eccalc{#1}{\textit{SUM}(#2, col)}}
-\newcommand{\ecsumr}[2]{\eccalc{#1}{\textit{SUM}(#2, row)}}
+\newcommand{\ecaggc}[3]{\eccalc{#2}{\textit{#1\textsubscript{col}}(#3)}}
+\newcommand{\ecaggr}[3]{\eccalc{#2}{\textit{#1\textsubscript{row}}(#3)}}
+\newcommand{\ecsumc}[2]{\eccalc{#1}{\textit{SUM\textsubscript{col}}(#2)}}
+\newcommand{\ecsumr}[2]{\eccalc{#1}{\textit{SUM\textsubscript{row}}(#2)}}
 \newcommand{\ecaggif}[5]{\eccalc{#2}{\textit{#1IF}(#3, #4, #5)}}
@@
 To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
-Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias").
-\tias{revise declarative claim?} It operates directly on the tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given.
+%Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
+It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given.
 %, and discovers column- and row-wise constraints.
@@
 
-\tias{section description (add prob def here?)}
+We first introduce some terminology and the concept of \template, after which we define the problem and make some additional considerations.
 
@@
 The most important concept is that of a \textbf{block}. 
-A block has to satisfy three conditions: 1) it contains only entire rows or entire columns of a single headerless table 2) it is contiguous and 3) it is type-consistent.
+\begin{definition}
+A \textbf{block} has to satisfy three conditions: 1) it contains only entire rows or entire columns of a single headerless table 2) it is contiguous and 3) it is type-consistent.
+The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type. 
 If it contains only rows we say it has \textit{row-orientation}, if only columns, \textit{column-orientation}. 
-The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type.
+\end{definition}
+
+In line with this definition, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
 We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independent of their orientation.
-
-Because of these properties, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
 
@@
 \begin{itemize}
-\item \textit{type}: a block has to be type-consistent, so it has a single type
+\item \textit{type}: a block is type-consistent, so it has a type
 \item \textit{table}: the table that this block belongs to
 \item \textit{orientation}: either row-oriented or column-oriented
-\item \textit{size}: the size of a block is the number of vectors in it, e.g. $b-a+1$, the number of rows or columns.
+\item \textit{size}: the size of a block is the number of vectors in it, that is, the number of rows or columns.
 \item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
@@
 \item \textit{columns}: also orientation specific, for row-based blocks this is the length, for column-based blocks the size
-\item \tias{What is missing is 'discrete'}
+\item \tias{What is missing is 'discrete', not needed?}
 \end{itemize}
@@
 
-\paragraph{Example}
+\begin{example}
 Consider headerless Table~1 in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e.
@@
 }
-
-We write that a block $B'$ is \textbf{contained} in a block $B$, $B' \sqsubseteq B$ iff both are valid blocks (single orientation, contiguous, type consistent) and each of the vectors in $B'$ is also in $B$.
+\end{example}
+
+\begin{definition}
+\textbf{Block containement $\sqsubseteq$.} 
+A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (single orientation, contiguous, type consistent) and each of the vectors in $B'$ is also in $B$.
 For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=T[\rangeto{a}{b},:] \wedge B'=T[\rangeto{a'}{b'},:] \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
+\end{definition}
 We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
-
+%
 An example is that $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.
@@
 \item
-\textit{\CName}  specifies the syntactic form of the constraint $c(G_1, ...,G_n)$, that is, the name of the constraint $c$ together
-with $n$ abstract arguments $G_i$.
+\textit{\CName}  specifies the syntactic form of the constraint $c(B_1, ...,B_n)$, that is, the name of the constraint $c$ together
+with $n$ abstract arguments $B_i$.
 Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
-Note that a function $G_r=f(G_1,...,G_n)$ can be represented with the $(n{+}1)$-ary predicate $c_f(G_r,G_1,...,G_n)$.
+Note that a function $B_r=f(B_1,...,B_n)$ can be represented with the $(n{+}1)$-ary predicate $c_f(B_r,B_1,...,B_n)$.
 Each argument will have to be instantiated with a block.
@@
 
-\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy, that is, the blocks.
+\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy.
 This can be any property of individual blocks, such as type, length, size and orientation, as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table, or have equal type or length.
@@
 \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
-Given an assignments of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not.
+Given an assignments of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks.
 %holds, in practice this will be a function that can be called when
@@
 
-When looking into constraints like $\ecrank{G_r}{G_x}$, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
-From a FOL perspective, the name of the constraint (\textit{RANK}) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
-
-This also holds in our setting, where an instantiation of a variable corresponds to a concrete set of vectors (a subgroup).
-For example for $\ecrank{G_r}{G_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $G_r$ is the $8$th vector in $T_1$: $G_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $G_x$ is the $7$th vector: $G_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
+It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
+From a FOL perspective, the name of the constraint (e.g.
+\textit{SUM}) is just the predicate name, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
+This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
+For example for $\ecrank{B_r}{B_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
 
@@
 \newcommand{\defc}{\ensuremath{\format{Def}_c}}
-\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
-With this interpretation, we can speak about the signature and definition of a constraint template being satisfied.
-We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by an assignment $(G_1, ..., G_n)$ if $\sigc(G_1, ..., G_n)$ (respectively $\defc(G_1, ..., G_n)$) is satisfied.
-Likewise, an assignment $(G_1, ..., G_n)$ satisfies a constraint template $s$ iff $|(G_1, ..., G_n)| = n_s \wedge \sigc(G_1, ..., G_n) \wedge \defc(G_1, ..., G_n)$.
+%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
+With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
+We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
+Likewise, the template is satisfied is both the signature and definition is satisfed; in first order logic, we would write: $c(B_1, ..., B_n) :-\sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
+Under this interpretation, the term constraint and constraint template can be used interchangeably.
+
+\begin{definition}
+A \textbf{valid argument assignment} of a constraint $c$ is a tuple $(B_1, ..., B_n)$ such that $c(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
+\end{definition}
 
@@
 
-\section{Problem Statement}\label{sec:problem_statement}
-The problem of learning constraints from tabular data can be seen as an inverse {\em constraint satisfaction problem} (CSP). A CSP consists of a set of constraints and the
-task is to find assignments of values to the variables in the CSP so that all constraints are satisfied.
-In the context of spreadsheets where variables could represent cells in tables, solving a corresponding CSP problem would mean completing the tables so that all constraints of the spreadsheet are satisfied.
-
-The inverse problem is, given an assignment to all cells, to find the constraints that are satisfied in the spreadsheet.
+\subsection{Problem Definition}\label{sec:problem_statement}
+The problem of learning constraints from tabular data can be seen as an inverse {\em constraint satisfaction problem} (CSP).
+In a CSP one is given a set of constraints over variables that must all be satisfied, and the goal is to find an instantiation of all the variables that satisfies these constraints.
+In the context of spreadsheets, the variables would be (blocks of) cells, and one would be given the actual constraints and functions with the goal of finding the values in the cells.
+The inverse problem is, given only an instantiation of the cells, find the constraints that are satisfied in the spreadsheet.
+
 We define the {\bf Tabular Constraint Learning Problem} as follows:
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf   Given} ${\cal G}$ a set of groups over fully instantiated tables ${\cal T}$ (all cells have values) and a set of {\template}s ${\cal S}$: {\bf find } all constraints $c_s(G'_1, ..., G'_n)$ that match the syntax of a \template~$s \in {\cal S}$ and where all subgroups $G'_i \subseteq G_i$ with $G_i \in {\cal G}$, satisfy both the signature and the definition of that \template.
+{\bf   Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf Find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
 \end{definition}
@@
 
-An example solution for the type-consistent groups of the tables in Figure~\ref{fig:main_example} and {\template}'s in Table~\ref{table:constraints} is shown in Figure~\ref{fig:sol_example}.
-
-Depending on the application, a solution to the tabular constraint learning problem may contain \textit{redundant} constraints, for details we refer to relational learning literature \cite{luc_book}.
+Figure~\ref{fig:sol_example} shows the solution to the tabular constraint learning problem when applied on the blocks of Figure~\ref{fig:main_example} and constraint templates listed in Table~\ref{table:constraints}.
+
+\subsection{Other considerations}
+
+\paragraph{Dependencies}
+In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g.
+for \textit{PERMUTATION}.
+This expresses a dependency of the constraint on that other constraint.
+This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the depending constraint, and its definition of its own definition plus the definition of the depending constraint.
+In FOL, we can see that one constraint entails the other, for example $\ecperm{\sg_x} \models \ecalldiff{\sg_{x}}$.
+
+Apart from easing the specification of the signature and definition, in Section~\tias{TODO} we will see how such dependencies can be used to speed up the search for constraints.
+
+\paragraph{Redundancies}
+Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}.
 This is because constraints may be logically equivalent or may be implied by other constraints.
 
-More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
-the  following statement holds: $$\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m).$$
-
-We shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
-To simplify notation, and in line with logic programming, we shall often make the quantification implicit and simply write this as rules $ c_1(G_1, ...
+\tias{This should contain examples that recur later in the text}
+
+For example, for some constraints, the order of some of the arguments may not matter.
+For example for the product constraint, $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecprod{\sg_r}{\sg_2}{\sg_1}$ so one can be considered redundant to the other.
+
+One way to deal with such equivalences is to identify a \textit{canonical} constraint among the equivalent ones.
+For example, for the product constraint above, we could define a lexicographic ordering over the blocks and only consider the constraint with the smallest ordering.
+
+Because dealing with redundancy is often application-dependent, we explain in the next section our generic method for finding all constraints.
+
+%
+%More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
+%the  following statement holds: $$\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m).$$
+%
+%We shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
+%To simplify notation, and in line with logic programming, we shall often make the quantification implicit and simply write this as rules $ c_1(G_1, ...
 G_n) \rightarrow c_2(G'_1, ...
@@
 We say that the first constraint is more specific than the second, and the second is more general.
-
-% where the second one ranges over a subsets of the arguments of the first, iff whenever the first constraint is true, the second must also be true.
+%
+%
+%% where the second one ranges over a subsets of the arguments of the first, iff whenever the first constraint is true, the second must also be true.
 More formally, $\forall G'_1, ..., G'_n$ s.t.
@@
 G'_n) \rightarrow c_2(G''_1, ..., G''_m)$ll.
-
-\paragraph{Example}
-Consider the following implication:
-$$\ecperm{G} \rightarrow \ecalldiff{G}$$
-(it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G$ is a vector).
-In this case the arguments are identical in both constraints.
-Given that the problem is to find the set of {\em all} constraints, and that the more general constraint is entailed
-by the more specific one, the $\ecdiff{G}$ constraint will be {\em redundant}.
-So, for most applications, it will be sufficient to output only the most specific one.
-
-A more involved example is
-$$\eclookup{G'_2}{G'_1}{H'_1}{H'_2} \rightarrow \ecfkey{G'_1}{H'_1}$$
-which states
-that a look-up constraint implies a foreign-key constraint.
+%
+%\paragraph{Example}
+%Consider the following implication:
+%$$\ecperm{G} \rightarrow \ecalldiff{G}$$
+%(it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G$ is a vector).
+%In this case the arguments are identical in both constraints.
+%Given that the problem is to find the set of {\em all} constraints, and that the more general constraint is entailed
+%by the more specific one, the $\ecdiff{G}$ constraint will be {\em redundant}.
+%So, for most applications, it will be sufficient to output only the most specific one.
+%
+%A more involved example is
+%$$\eclookup{G'_2}{G'_1}{H'_1}{H'_2} \rightarrow \ecfkey{G'_1}{H'_1}$$
+%which states
+%that a look-up constraint implies a foreign-key constraint.
 In the look-up constraint $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ the values in the vector $G'_2$ are determined by looking up the corresponding value in vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
@@
 This implies a foreign-key constraint $\ecfkey{G'_1}{H'_1}$ where each of the values in vector $G'_1$ must appear in $H'_1$ for the look-up to work.
-
-Two constraints are \textit{equivalent} if they imply one another.
-%More formally, $\forall G'_1, ..., G'_n$ s.t.
-$\{G''_1, ..., G''_n\} = \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
-G'_n) \leftrightarrow c_2(G''_1, ..., G''_n)$.
-An example within the same \template is for the sum over vectors constraint: $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-A related example over multiple constraints is $G_1 = +(G_2, G_3) \leftrightarrow G_1 = \textit{SUM}(H)$ with $\textit{size}(H) = 2$ and $G_2 \subset H, G_3 \subset H$.
-\luc{Do we really need this, or can we write it in a real equivalence --- conjunctive formula?}
-\samuel{Tias, Sergey what do you think? Also I find this notation confusing: +(G2, G3), why not just use G1 = G2 + G3?}
-
-A good way to deal with equivalent constraints is to work with a \textit{canonical} constraint among the equivalent ones.
-For example, for equivalent constraints of the same \template, this could be the constraint where the groups are ordered by table, orientation and row/column identifier.
-For equivalent constraints of different templates one could define heuristics such as that some templates are preferred over others, e.g.
-binary sum over n-ary sum, sum over difference, etc.
-
-Because dealing with redundancy is often application-dependent, we explain in the next section our generic method for finding all constraints.
 

### 2016-05-19 Paper/ECAI/tabular.tex
Table + group detection

@@
 \usepackage{subcaption}
+\usepackage[inline]{enumitem}
 
@@
 Recall that a valid argument assignment for a constraint is an assignment of each argument of the constraint to a block, such that the constraint, that is, its signature and definition, is satisfied.
+\samuel{In general: should it not be: assign blocks to arguments instead of arguments to blocks?}
 \tias{I here use constraint and constraint template interchangeably}
@@
 
-\subsection{Table and group detection} \label{sec:make_groups}
-%A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) groups.
+\subsection{Table extraction}
+\label{sec:table_extraction}
 Many spreadsheets contain headers that provide hints at where the table(s) in the sheet are and what the meaning of the rows and columns is.
@@
 
-%\tias{0. preprocess: currencies and precentual to numeric, 1. tables from spreadsheets, 2. groups from tables. Ext: could be done interactively.}
-Fully automatic table and header detection is not the core of our work. We hence built a visual tool that uses a simple method to detect tables in the data and allows a user to correct these.
-
 \begin{figure}[t]
@@
   \includegraphics[width=1\linewidth]{figures/tabletool.png}
-  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
-    The graph is extracted from the Signature column in Table \ref{table:constraints}.
-  }
-  \label{fig:learning_order}
+  \caption{This figure shows our table selection tool.
+A user selects rectangular ranges of cells that correspond to (headerless) tables.
+If the data has an unambiguous orientation, this information can be provided as well.}
+  \label{fig:visual_tool}
 \end{figure}
 
-\samuel{This does not fully correspond with the real tool, the real tool does not allow detected tables to be corrected: Either use detection tool or use visual tool to input tables manually}
-The way the table detection in the tool works is as follows:
-First, we preprocess the spreadsheets so that currency values are transformed into their integer or float value and percentual data is transformed into the corresponding floating point representation (e.g.
-$85\%$ as $0.85$).
-Next, every row in the spreadsheet is partitioned into blocks of columns that are separate by at least one $None$ value.
-Then, the blocks in each row are greedily merged to obtain rectangles.
-Finally headers are removed by checking whether the first row or column is all textual while the next row or column is not; in that case the assumed header is removed.
-
-It should be clear this simple method only works if all tables are clearly separated by $None$ values and that headers of fully textual tables are not detected, among other things.
-Hence, in the visual tool the user can correct the detected tables if needed.
-\samuel{\tias{I think it would be good if we could show a screenshot of this tool...} Added screenshot, needs to be referred to}
-
-Row-oriented and column-oriented groups are then extracted from each table by partitioning the rows (columns) into type-consistent parts.
-
-From here on we will assume that groups and tables are given.
+Our algorithm requires tables to be specified, this can, however, be done in various ways.
+Tables are specified in a JSON configuration file which contains the table names, their coordinates within a CSV file and optionally the orientation of certain tables.
+The orientation can be used to indicate that a table should be interpreted as only rows or only columns.
+
+The generation of this JSON file can be done in many ways, ranging from manually specifying the tables to automatically detecting tables.
+Requiring the user to manually generate these files decreases the usability of our tool in a standalone setting, therefore, we offer two proof of concept approaches that aim to facilitate the specification of tables within a CSV.
+
+\paragraph{Automatic detection}
+As mentioned above, automatic detection of headers and tables is a complex task.
+Therefore, we define a restricted setting in which the task becomes easy enough to be automated.
+We assume that 1) tables are rectangular blocks of data not containing any empty cells; and 2) tables are separated by at least one empty cell from other tables.
+Additionally, if the first row contains only textual values it treated as a header, otherwise, if the first column consists of only textual data is will be the header.
+If a header row (column) is detected, it is removed from the table and the orientation of the table is fixed to columns (rows), otherwise, we assume there is no header and the orientation is not fixed.
+
+The tool will now detect ranges of non-empty cells in every row and greedily merge these ranges across columns to obtain rectangular tables.
+If the first row of a table contains 
+This tool removes the need for human intervention, however, it is rather limited and can only be used in simple settings.
+
+\paragraph{Visual}
+Since the specification of tables is usually easy for humans, we propose a second approach which aims to decrease the complexity of specifying tables by allowing users to indicate tables using a visual tool (Figure~\ref{fig:visual_tool}).
+Users select tables excluding the header and optionally specify an orientation.
+The tool then generates the table definitions automatically.
+\\\\
+Aside from extending the automatic detection to be more applicable, one could combine the visual and automatic detection to let the user approve or correct detected table.
+
+\subsection{Block detection} \label{sec:make_groups}
+%A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) blocks.
+The goal of the block detection step is to partition tables into type-consistent (all-numeric or all-textual) blocks.
+We generate blocks automatically, partitioning tables into maximal type-consistent blocks.
+First, the spreadsheet data is preprocessed so that currency values and percentual values are transformed into their corresponding numeric (i.e.
+integer or floating point) representation (e.g.
+$2.75 \$$ as $2.75$ or $85\%$ as~$0.85$).
+Afterwards, tables are partitioned into maximal type-consistent blocks.
+If a table has no fixed orientation, the algorithm will attempt to construct both row-blocks and column-blocks.
+In order to find row-blocks (column-blocks), the type of each row (column) is determined, i.e. whether it is numeric or textual.
+All adjacent rows (columns) that have the same type are merged to form a block. 
+If any row (column) of the table contains both numeric and textual values, it is type inconsistent and the algorithm will not produce any row-blocks (column-blocks) for that table.
+For tables that have a fixed orientation only blocks of that orientation are built.
+
+While this step can be done algorithmically, there are cases where specifying finer-grained blocks manually may be beneficial.
+Firstly, blocks may group data from different contexts.
+For example, consider the type consistent block $B_3=\range{T_1}{\rangeall}{\rangeto{3}{8}}$ from the example in Figure~\ref{fig:main_example}: $\range{T_1}{\rangeall}{8}$ is contextually different from $\range{T_1}{\rangeall}{\rangeto{3}{7}}$, therefore, we could split~$B_3$ into two blocks.
+\samuel{Secondly, accidental blocks}
+Finer grained groups can reduce the amount of redundant constraints that are found.
 

### 2016-05-19 Paper/ECAI/tabular.tex
Fig: add all constraints
Rewrite algorithm intro

@@
 \begin{subfigure}{.30\textwidth}
-  {\small
+  {\footnotesize
     \begin{align*}
+      % Table 1
       &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
 %
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
-%
+      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
+%
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}} \\
+%
+      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+%
+      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+%
+	  % Table 2
       &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
 %
+      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
+%
+      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+	  % Table 4
+      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}} \\
+%
+      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
+%
+	  % Table 6
       &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
-%
-      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
-%
-      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
 %
@@
   \vspace{-10pt}
-  \caption{Constraints present in the running example (left) \tias{Add ALL, including non-functional and duplicates}.}
+  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).}
   \label{fig:sol_example}
@@
 
-\paragraph{Automatic detection}
+\subsubsection{Automatic detection}
 As mentioned above, automatic detection of headers and tables is a complex task.
@@
 
-\paragraph{Visual}
+\subsubsection{Visual}
 Since the specification of tables is usually easy for humans, we propose a second approach which aims to decrease the complexity of specifying tables by allowing users to indicate tables using a visual tool (Figure~\ref{fig:visual_tool}).
@@
 \newcommand{\temps}{\ensuremath{S}}
-\subsection{Algorithm} \label{sec:algo}
-Our method solves the \tcl problem by checking for each template what combination of groups can satisfy the signature, and which specific subgroups satisfy the definition.
-
-The main challenge is to avoid having to check each possible combination of group and subgroup on each constraint template, as well as developing an \textit{extensible} method in which new constraint templates can be plugged in with relative ease.
+
+\subsection{Algorithm}
+\label{sec:algo}
+Our method assumes that tables and blocks are given and solves the \tcl~problem by checking for each template: what combination of given (super)blocks may satisfy the signature and which specific (sub)blocks satisfy both the signature and the definition.
+
+The main challenge we aim to address with our approach is to avoid having to check each possible combination of blocks for each constraint template as well as developing an \textit{extensible} method in which new constraint templates can be plugged in with relative ease.
 %We will use constraint satisfaction technology to partly answer this challenge.
-To answer this challenge we separate checking the signature of a constraint template from checking the definition of a constraint template.
-Furthermore, we use constraint satisfaction technology to jointly search for and test the signature over all group combinations.
-
-The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}; it consists of three key components: 1) templates can depend on each other, so they are ordered to avoid having to verify the same template multiple times; 2) we generate all combinations of groups that can satisfy the signature using constraint satisfaction techniques; 3) from this, all satisfying subgroups are generated.
+To answer this challenge we separate checking the signature (concerning properties of a block) of a constraint template from checking the definition (concerning the values in a block) of a constraint template.
+Furthermore, we use constraint satisfaction technology to jointly search for and test the signature over all combinations of superblocks.
 
@@
 
-%First, constraint templates are ordered such that they can be treated sequentially and each only once.
-%Then, for each constraint template~$s \in \temps$ we generate a set~$A$ of \textit{group-assignments} that satisfy the \CSignature of~$s$.
-%Every group assignment assigns a group~$G \in \groups$ to every argument of the template.
-%In the third step, for every assignments $(G_1, ..., G_n) \in A$ we find all constraints $s(G_1', ...
-G_n')$ that hold on the data (satisfy the \CFunction), such that every subgroup $G_i' \subseteq G_i$.
-
-%Our Algorithm \ref{algo:tcl} is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}, however, we apply the approach on two levels, generating and testing groups according to their properties and generating and testing subgroups on the actual data.
-%We will explain and illustrate the three key steps in detail.
-
+The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}; it consists of three key components: 1) templates can depend on each other, so they are ordered to avoid having to verify the same template multiple times; 2) we generate all combinations of superblocks that can satisfy the signature using constraint satisfaction techniques; 3) from this, all satisfying subblocks are generated.
 We now describe each of the three components in more detail.
 
+\samuel{MARKER-C}
 \subsubsection{Template ordering}

### 2016-05-20 Paper/ECAI/tabular.tex
Removed old text
minor changes

--- 2016-05-19 Paper/ECAI/tabular.tex
@@
 \usepackage{notoccite}
+\usepackage{fixltx2e}
 \usepackage{framed}
@@
 
-Let us illustrate this on one of the spreadsheets from the gathered dataset. In Figure \ref{fig:autocompletion_example}, \sergey{to be finished}
+Let us illustrate this on one of the spreadsheets from the gathered dataset.
+In Figure \ref{fig:autocompletion_example}, we see two tables the upper one is completed and in the lower one values in the last raw are not yet filled, then the system finds a satisfying constraint (\textit{SUMIF}) and based on it proposes filling values.
 

### 2016-05-20 Paper/ECAI/tabular.tex

--- 2016-05-19 Paper/ECAI/tabular.tex
@@
 
-\paragraph{Rest old text}
-
-This section describes our approach to \tcl.
-The main method takes \textit{groups} and {\template}s as input, and is described in Section~\ref{sec:algo}.
-We first explain how tables and groups can be extracted from raw spreadsheet data.
-
 \subsection{Table extraction}

### 2016-05-20 Paper/ECAI/tabular.tex
Reworked intro algorithm

--- 2016-05-19 Paper/ECAI/tabular.tex
@@
 \newcommand{\format}[1]{\textit{#1}\xspace}
-\newcommand{\generategroups}{\format{generateGroups}}
+\newcommand{\generategroups}{\format{SuperblockAssignments}}
 \newcommand{\extractgroups}{\format{extractGroups}}
 \newcommand{\extracttables}{\format{extractTables}}
-\newcommand{\learnconstraints}{\format{learnConstraints}}
-\newcommand{\findassignment}{\format{findConstraints}}
+\newcommand{\learnconstraints}{\format{LearnConstraints}}
+\newcommand{\findassignment}{\format{FindSubassignments}}
 \newcommand{\postprocess}{\format{pruneRedundant}}
@@
     \footnotesize
-    \State \textbf{Input:} $\groups$ -- groups, $\constraints$ -- \template's, \dependencies -- dependency graph
+    \State \textbf{Input:} $\groups$ -- groups, $s$ -- constraint template
     \State \textbf{Output:} $C$ -- learned constraints
+    \Procedure{LearnConstraints}{$\groups$, $s$}
     \State $C \gets \emptyset$ %\Comment{The set of constraints}
-  \ForAll{$s \in \constrainttorder(\constraints,\dependencies)$}
     \State $n \gets$ number of arguments of template $s$
-    \ForAll{$(G_1, \dots, G_n) \in \generategroups(s, \groups, C)$}
-      \State $C \gets C \cup \findassignment(s, (G_1, \dots, G_n))$
+    \State $A \gets \generategroups(s, \groups)$
+    \ForAll{$(B_1, \dots, B_n) \in A$}
+      \ForAll{$(B_1', \dots, B_n') \in \findassignment(s, (B_1, \dots, B_n))$}
+        \State $C \gets C \cup \{ s(B_1', \dots, B_n') \}$
       \EndFor
-  \EndFor\\
+    \EndFor
 \Return $C$
+\EndProcedure
 \end{algorithmic}
-\caption{Tabular constraint learning}
+\caption{Learn tabular constraints}
 \label{algo:tcl}
@@
 
-The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}; it consists of three key components: 1) templates can depend on each other, so they are ordered to avoid having to verify the same template multiple times; 2) we generate all combinations of superblocks that can satisfy the signature using constraint satisfaction techniques; 3) from this, all satisfying subblocks are generated.
-We now describe each of the three components in more detail.
-
-\samuel{MARKER-C}
-\subsubsection{Template ordering}
-As discussed before, a constraint may imply another constraint.
-This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as $LOOKUP$ depending on $FOREIGNKEY$.
-
-In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
-In our method, we use these dependencies/implications to define a dependency graph between the templates.
-We assume the signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
-Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}. 
-Constraint templates that have no dependencies are omitted.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency.png}
-  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
-    The graph is extracted from the Signature column in Table \ref{table:constraints}.
-  }
-  \label{fig:learning_order}
-\end{figure}
-
-The $\constrainttorder(\constraints,\dependencies)$ module in our method takes a set of {\template}s~(\constraints) and a DAG~(\dependencies) as input and produces a linear ordering of the templates such that the more general, implied constraint templates (the requirements) are ordered before the templates that depend on them.
-
-\paragraph{Example}
-As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $LOOKUP$ constraint template depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
-
-% not helpful:
-%The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}.
+Given a constraint template~$s$, the algorithm first generates all combinations of superblocks (i.e. superblock assignments) that are consistent with the signature.
+In a second step it looks at every superblock assignment and finds subassignments that satisfy the signature and definition of~$s$.
 
@@
 \subsection{Implementation}
+
+\samuel{Start: Moved template ordering}
+\subsubsection{Template ordering}
+As discussed before, a constraint may imply another constraint.
+This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as \textit{LOOKUP} depending on \textit{FOREIGNKEY}.
+
+In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
+In our method, we use these dependencies/implications to define a dependency graph between the templates.
+We assume the signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
+Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}. 
+Constraint templates that have no dependencies are omitted.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency.png}
+  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
+    The graph is extracted from the Signature column in Table \ref{table:constraints}.
+  }
+  \label{fig:learning_order}
+\end{figure}
+
+The $\constrainttorder(\constraints,\dependencies)$ module in our method takes a set of {\template}s~(\constraints) and a DAG~(\dependencies) as input and produces a linear ordering of the templates such that the more general, implied constraint templates (the requirements) are ordered before the templates that depend on them.
+
+\paragraph{Example}
+As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $LOOKUP$ constraint template depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
+
+% not helpful:
+%The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
+\samuel{End: Moved template ordering}
+
 We provide some more details of our implementation of the above algorithmic approach, to ease reproducibility.

### 2016-05-20 Paper/ECAI/tabular.tex
minor

@@
 
-\newcommand{\sg}{G'}
+\newcommand{\sg}{B}
 
@@
 Section \ref{sec:formalization} introduces concepts relevant to our approach.
-Section \ref{sec:problem_statement} presents the problem statement.
-Section \ref{sec:approach} describes the approach.
+Section \ref{sec:problem_statement} presents the problem statement and the approach.
 Section \ref{sec:evaluation} presents the evaluation of the approach.
@@
 
-A row or columns is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual.
+A row or a column is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual.
 We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$, and similarly $T[{:},a]$ for the $a$-th column.
@@
 \end{definition}
+\sergey{do we need to mention explicitly that blocks have the same orientation and are in the same table?}
 We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
@@
 This can be any property of individual blocks, such as type, length, size and orientation, as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table, or have equal type or length.
-In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
+In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
 
@@
 \begin{example}
-Several example constraint templates are illustrated in Table \ref{table:constraints}. An non-trivial example is the constraint template for the row-based sum:
+Several constraint templates are illustrated in Table \ref{table:constraints}. A non-trivial example is the constraint template for the row-based sum:
 
@@
 
-  \item \CSignature: $\sg_r$ has to be a single vector ($size=1$) while $\mathbf{\sg_x}$ can be a block ($size>=1$), which can be derived from the use of a normal or \textbf{bold} font.
+  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
 The two blocks have to be numeric.
@@
 From a FOL perspective, the name of the constraint (e.g.
-\textit{SUM}) is just the predicate name, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
+\textit{SUM}) is just the predicate name, and the arguments $B_r$ and $B_x$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
 This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
@@
 We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
-Likewise, the template is satisfied is both the signature and definition is satisfed; in first order logic, we would write: $c(B_1, ..., B_n) :-\sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
+Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would write a Prolog-like clause: $c(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
 Under this interpretation, the term constraint and constraint template can be used interchangeably.
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf   Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf Find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
+{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
 \end{definition}
@@
 \paragraph{Dependencies}
+\sergey{question: the guidlines say caption for tables above the tables, but dsaa papers from 2015 (I have checked two) put it below, that's confusing...}
 In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g.
@@
 This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the depending constraint, and its definition of its own definition plus the definition of the depending constraint.
-In FOL, we can see that one constraint entails the other, for example $\ecperm{\sg_x} \models \ecalldiff{\sg_{x}}$.
-
-Apart from easing the specification of the signature and definition, in Section~\tias{TODO} we will see how such dependencies can be used to speed up the search for constraints.
+In FOL, we can see that one constraint entails the other, for example if $\ecperm{\sg_x}$ holds for a block $\sg_x$, then $\ecalldiff{\sg_{x}}$ also holds.
+
+Apart from easing the specification of the signature and definition, in Section~\ref{sec:approach} we will see how such dependencies can be used to speed up the search for constraints.
 

### 2016-05-20 Paper/ECAI/tabular.tex
minor

@@
 \paragraph{Dependencies}
-\sergey{question: the guidlines say caption for tables above the tables, but dsaa papers from 2015 (I have checked two) put it below, that's confusing...}
+\sergey{question: the guidlines say caption for tables above the tables, but dsaa papers from 2014 and 2015 (I have checked three) put it below, that's confusing...}
 In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g.

### 2016-05-20 Paper/ECAI/tabular.tex
fixed the algorithm

@@
 \item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal \textit{superblock} and the $B_i$ are compatible with the signature.
-\item For each generated atom $c(B_1, \ldots ,B_n)$, find all valid $c(B'_1, \ldots, B'_n)$ such that $\forall i, B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ is satisfied.
+\item For each generated atom $c(B_1, \ldots ,B_n)$, find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
 \end{enumerate}
@@
     \ForAll{$(B_1, \dots, B_n) \in A$}
-      \ForAll{$(B_1', \dots, B_n') \in \findassignment(s, (B_1, \dots, B_n))$}
+    \ForAll{${(B_1',\dots,B_n')\in}~\findassignment(s, (B_1,\dots,B_n))$}
         \State $C \gets C \cup \{ s(B_1', \dots, B_n') \}$
@@
 In the previous step, $\generategroups(\textit{t,{\cal G},C})$ generates tuples of candidate groups $(G_1, ..., G_n)$ for the constraint template $t$ with $n$ arguments.
-In this step, $\findassignment(t, (G_1, \dots, G_n))$, the goal is to enumerate the subgroups $(G'_1, ..., G'_n), \forall i, G'_i \subseteq G_i$ that satisfy both the signature and the definition of the constraint template $t$.
+In this step, $\findassignment(t, (G_1, \dots, G_n))$, the goal is to enumerate the subgroups $(G'_1, ..., G'_n)$ (for all $i$, $G'_i \subseteq G_i$) that satisfy both the signature and the definition of the constraint template $t$.
 The main difference with the previous step is that the group generation reasons about the \textit{properties} of the groups (type, size, length, ...), while this step reasons about the actual \textit{content} of the vectors in the groups, that is, the data.
@@
 
+\sergey{there should be a discussion on what to do if multiple constraints satisfy the data and propose predictions, just a sketch here for now}
+If multiple constraints satisfy the data and allow autocompletion, then there is a number of ways to handle it.
+Firstly, autocompletion can be disabled.
+Secondly, the most specific constrain can be applied.
+If they are not compatible in the generality sense, one needs to introduce an autocompletion order, which would resolve the problem by picking only one constraint to suggest the values.
+
 \subsection{Error detection}

### 2016-05-20 Paper/ECAI/tabular.tex
fixing things catching my eye

@@
 Section \ref{sec:formalization} introduces concepts relevant to our approach.
-Section \ref{sec:problem_statement} presents the problem statement and the approach.
+Section \ref{sec:approach} presents the problem statement and the approach.
 Section \ref{sec:evaluation} presents the evaluation of the approach.
@@
 \item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
-\item Generate for each constraint all valid argument assignments:
+\item Generate for each constraint all valid argument assignments in two steps:
 \begin{enumerate}
@@
   \includegraphics[width=1\linewidth]{figures/tabletool.png}
-  \caption{This figure shows our table selection tool.
+  \caption{This figure shows the table selection tool of \sname.
 A user selects rectangular ranges of cells that correspond to (headerless) tables.
@@
 \paragraph{Example}
-As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $LOOKUP$ constraint template depends on $FOREIGNKEY$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
+As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $\textit{LOOKUP}$ constraint template depends on $\textit{FOREIGNKEY}$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
 
@@
 
-Our implementation supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the stand-alone Constraint Programming language MiniZinc \cite{minizinc} with the Gecode 4.4.0 solver which supports floats, as well as an embedded Python-based CP solver \cite{python_constraint}.
+\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the stand-alone Constraint Programming language MiniZinc \cite{minizinc} with the Gecode 4.4.0 solver which supports floats, as well as an embedded Python-based CP solver \cite{python_constraint}.
 The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints.
@@
 
-Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}
-
-\sergey{under question now, might worth deleting} \\
-\sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}
-
-\begin{equation}
-  \forall i{:}~ v_r[i] \leq v[i] < v_r[i] + 0.5
-  \label{eq:floor_extension}
-\end{equation}
+% Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}
+
+% \sergey{under question now, might worth deleting} \\
+% \sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}
+
+% \begin{equation}
+%   \forall i{:}~ v_r[i] \leq v[i] < v_r[i] + 0.5
+%   \label{eq:floor_extension}
+% \end{equation}
 
@@
 \subsubsection{Results}
-Our current implementation takes a few seconds to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
+\sname takes a few seconds to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
 Of the 6 remaining (redundant) constraints, 5 are $\mathit{rank}$ constraints that are true by accident, such as: \begin{align*}
@@
 \subsection{Benchmark}
-There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paperthe submitted paper)}.
+There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
 The case study is also included.
@@
 \end{figure}
+\sergey{I think the time discussion is irrelevant, should it be cut out completely?}
 
@@
 Secondly, the most specific constrain can be applied.
-If they are not compatible in the generality sense, one needs to introduce an autocompletion order, which would resolve the problem by picking only one constraint to suggest the values.
+Thirdly, if one is not more specific than the other, one needs to introduce an autocompletion order, which would resolve the problem by picking only one constraint to suggest the values.
 
@@
 \end{figure*}
+
+\sergey{Either we need to elaborate on the online detection or remove offline-online and simplify the rest, otherwise it is not clear what is going on; again I don't see any point in time here}
 
@@
 
-\sergey{ECAI reference style file ignores their guideline and their guideline ignores what is written in the guidelines!}
 flashfill, flashextract, flashmeta \cite{flashfill,flashextract,flashmeta}

### 2016-05-20 Paper/ECAI/tabular.tex
Reworked superblock assignment

--- 2016-05-20 Paper/ECAI/tabular.tex
@@
 \usepackage{notoccite}
-\usepackage{fixltx2e}
 \usepackage{framed}
@@
 \newcommand{\dependencies}{\ensuremath{\mathcal{D}}\xspace}
-\newcommand{\groups}{\ensuremath{\mathcal{G}}\xspace}
+\newcommand{\groups}{\ensuremath{\mathcal{B}}\xspace}
+\newcommand{\blocks}{\ensuremath{\mathcal{B}}\xspace}
 
@@
 \newcommand{\plength}{\format{length}}
+\newcommand{\psize}{\format{size}}
 \newcommand{\ptype}{\format{type}}
@@
 
-\newcommand{\sg}{B}
+\newcommand{\sg}{G'}
 
@@
 Section \ref{sec:formalization} introduces concepts relevant to our approach.
-Section \ref{sec:problem_statement} presents the problem statement and the approach.
+Section \ref{sec:problem_statement} presents the problem statement.
+Section \ref{sec:approach} describes the approach.
 Section \ref{sec:evaluation} presents the evaluation of the approach.
@@
 
-A row or a column is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual.
+A row or columns is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual.
 We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$, and similarly $T[{:},a]$ for the $a$-th column.
@@
 \begin{definition}
-\textbf{Block containement $\sqsubseteq$.} 
+\textbf{Block containment $\sqsubseteq$.} 
 A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (single orientation, contiguous, type consistent) and each of the vectors in $B'$ is also in $B$.
@@
 \end{definition}
-\sergey{do we need to mention explicitly that blocks have the same orientation and are in the same table?}
 We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
@@
 This can be any property of individual blocks, such as type, length, size and orientation, as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table, or have equal type or length.
-In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
+In terms of logical and relational learning \cite{luc_book}, such a \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
 
@@
 \begin{example}
-Several constraint templates are illustrated in Table \ref{table:constraints}. A non-trivial example is the constraint template for the row-based sum:
+Several example constraint templates are illustrated in Table \ref{table:constraints}. An non-trivial example is the constraint template for the row-based sum:
 
@@
 
-  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
+  \item \CSignature: $\sg_r$ has to be a single vector ($size=1$) while $\mathbf{\sg_x}$ can be a block ($size>=1$), which can be derived from the use of a normal or \textbf{bold} font.
 The two blocks have to be numeric.
@@
 From a FOL perspective, the name of the constraint (e.g.
-\textit{SUM}) is just the predicate name, and the arguments $B_r$ and $B_x$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
+\textit{SUM}) is just the predicate name, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
 This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
@@
 We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
-Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would write a Prolog-like clause: $c(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
+Likewise, the template is satisfied is both the signature and definition is satisfed; in first order logic, we would write: $c(B_1, ..., B_n) :-\sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
 Under this interpretation, the term constraint and constraint template can be used interchangeably.
@@
     \ecseries{\sg_x}
-      & $\numeric(\sg_{x})$ and $\ecperm{\sg_{x}}$
+      & $\integer(\sg_{x})$ and $\ecperm{\sg_{x}}$
       & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
+{\bf   Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf Find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
 \end{definition}
@@
 \paragraph{Dependencies}
-\sergey{question: the guidlines say caption for tables above the tables, but dsaa papers from 2014 and 2015 (I have checked three) put it below, that's confusing...}
 In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g.
@@
 This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the depending constraint, and its definition of its own definition plus the definition of the depending constraint.
-In FOL, we can see that one constraint entails the other, for example if $\ecperm{\sg_x}$ holds for a block $\sg_x$, then $\ecalldiff{\sg_{x}}$ also holds.
-
-Apart from easing the specification of the signature and definition, in Section~\ref{sec:approach} we will see how such dependencies can be used to speed up the search for constraints.
+In FOL, we can see that one constraint entails the other, for example $\ecperm{\sg_x} \models \ecalldiff{\sg_{x}}$.
+
+Apart from easing the specification of the signature and definition, in Section~\tias{TODO} we will see how such dependencies can be used to speed up the search for constraints.
 
@@
 
-\subsubsection{Candidate group generation}
+\subsubsection{Generating superblock assignments}
 %The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
-Given a constraint template and the set of all groups ${\cal G}$, the goal is to find all combinations of groups that satisfy the constraint signature.
-Let there be $n$ groups and let the constraint have $k$ arguments, then a simple generate-and-test method would have to test $\tbinom{n+k-1}{k}$ combinations (given that groups can be repeated), which grows very rapidly with $n$.
-
-However, the properties defined in the signature, such as related to type or size, can be used to avoid considering certain groups for certain arguments.
-Furthermore, the choice of one argument can influence the possible candidates for the arguments, for example if they must have the same length.
+Given a constraint template~$s$ and the set of all blocks~\blocks, the goal is to find all combinations of blocks that satisfy the constraint signature.
+Let there be $n$ blocks and let~$s$ have~$k$ arguments, then a simple generate-and-test method would have to test $n^k$ combinations (given that blocks can be repeated), which can grow very rapidly with $n$, depending on $k$.
+
+However, the properties defined in the signature, such as those related to the type or size of a block, can be used to avoid considering certain groups for certain arguments.
+Furthermore, the choice of one argument can influence the possible candidates for the other arguments, for example, if they must have the same length.
 Instead of writing specialized code for each of these cases, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.
@@
 A Constraint Satisfaction Problem (CSP) is a triple $(V,D,C)$ where $V$ is a set of \textit{variables}, $D$ the domain of possible values each variable can take and $C$ the set of constraints over $V$ that must be satisfied.
-In our case, we define one variable $V_i$ for each argument of the constraint.
-Each variable can be assigned to each of the groups in ${\cal G}$, so the domain consists of $|\{\cal G\}|$ identifiers.
+In our case, we define one variable $V_i$ for each argument of a constraint template.
+Each variable can be assigned to each of the blocks in \blocks, so the domain consists of $|\{\blocks\}|$ identifiers.
 
@@
 
-To reason over the groups, we automatically add to the constraint program a set of background facts that contain the properties of each group, namely its type, size, number of columns and rows, length, orientation, and table.
-We also add all previously found constraints of constraint templates that are implied by this template.
-The actual constraints then express the properties defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
+To reason over the blocks, we automatically add to the constraint program a set of background facts that contain the properties of each block, namely its type, size, number of columns and rows, length, orientation, and table.
+\samuel{Review: We also add all previously found constraints of constraint templates that are implied by this template.}
+The actual constraints correspond to the requirements defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
+
 % \tias{More detailed description of the properties commented out, feel free to put in}
@@
 
-Constraints derived from the signature can usually but not always be enforced exactly on a group, as we have to consider whether subgroups could satisfy them.
-Requirements on the lengths, orientations and tables of groups can be directly enforced since they stay the same for subgroups.
-Typing requirements need to be relaxed to consider only non-subtypes, since, for example, numeric groups might contain integer subgroups.
+\begin{table}[htb]
+  \caption{Translation of signature requirements to superblock constraints}
+  \label{tbl:translation}
+  \begin{tabularx}{\linewidth}{lX}
+    \textbf{Constraint} & \textbf{Superblock constraint} \\ \hline \hline
+    $\plength(B) = x$ & $\plength(B) = x$ \\ \hline
+    $\por(B) = x$ & $\por(B) = x$ \\\hline
+    $\ptable(B) = x$ & $\ptable(B) = x$ \\ \hline
+    $\ptype(B) = x$ & $\mathit{supertype}(\ptype(B)) = \mathit{supertype}(x)$ \\ \hline
+    $\psize(B) \geq x$ & $\psize(B) \geq x$ \\ \hline
+    $\psize(B) = x$ & $\psize(B) \geq x$ \\ \hline
+    $\pcols(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \pcols(B) \geq x$ \\ 
+    & $\mathit{if}~\por(B) = \mathit{row}: \pcols(B) = x$ \\ \hline
+    $\prows(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \prows(B) = x$ \\ 
+    & $\mathit{if}~\por(B) = \mathit{row}: \prows(B) \geq x$
+  \end{tabularx}
+\end{table}
+
+In order for our approach to be complete, the signature constraints can not always be enforced directly on a block.
+If a block matches the signature it is a valid candidate, however, if it does not there might exist subblocks that still satisfy the signature.
+\begin{example}
+  Consider, for example, the constraint template \ecseries{x} and a superblock candidate~$B$ for~$x$ containing both integer and floating point values.
+  While the signature of the template requires~$x$ to be integer we cannot discard~$B$ since there might be subblocks~$B' \sqsubseteq B$ that only contain integer values and satisfy the signature.
+\end{example}
+Table~\ref{tbl:translation} shows the how constraints from the signature are converted to constraints on superblocks.
+Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
+Typing requirements need to be relaxed to check only for supertypes (numeric and textual).
 Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since subgroups might contains fewer vectors.
-Finally, exact restrictions on the number of rows or columns may have to be relaxed depending on the orientation of the group.
+Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the group they are applied to.
 Subgroups of row-oriented (column-oriented) group will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
-Therefore, we can only require row-oriented (column-oriented) groups to contain \textit{at least} the required amount of rows (columns).
-
-For example, the \CSignature of ?? requires $v_r$ to be integer.
-If a numeric group is assigned to $v_r$ then some subgroups may have type float while other may be of type integer, therefore, this group cannot be discarded.
-
-\paragraph{Example}
-We show this construction of a CSP for the constraint template $\ecsumprod{G'}{H'}{I'}$ in Table~\ref{table:constraints}.
-It defines that $G'$ is a cell whose values is the result of the sum of products of rows in $H'$ and columns in $I'$.
-%
-There are three \textit{variables} $X, Y, Z$ corresponding to group $G'$, $H'$ and $I'$ respectively. The domains are $D(X) = D(Y) = D(Z) = \{1, \ldots, |{\cal G}|\}$. Finally, the constraints are:
+
+The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all satisfying solutions.
+These solutions will correspond with the valid superblock assignments for constraint template~$s$.
+
+\begin{example}
+  \samuel{Is this still necessary?}
+  Consider the template $\ecsumprod{x}{y}{z}$, then the generated CSP will contain three variables $V_x, V_y, V_z$ corresponding to arguments $x, y$ and $z$ respectively.
+  Given blocks~\blocks, the domain of these variables are $D(V_x) = D(V_y) = D(V_z) = \{ 1, \dots, |\blocks| \}$.
+  Finally, the signature will be translated into constraints:
   \begin{align*}
-&numeric(X) \land numeric(Y) \land numeric(Z) \\
-&\pcols(X) = 1 \land \prows(X) = 1 \\
-&\plength(Y) = \plength(Z) \land \plength(Y) \geq 2
+    & \numeric(V_x) \land \numeric(V_y) \land \numeric(V_z) \\
+    & \por(V_x) = \mathit{row} \implies \pcols(V_x) = 1 \\
+    & \por(V_x) = \mathit{row} \implies \prows(V_x) = 1 & \\
+    & \plength(V_y) \geq 2 \land \plength(V_z) \geq 2 \\
+    & \plength(V_y) = \plength(V_z)
   \end{align*}
-\tias{Sumproduct is a rather peculiar constraint for spreadsheets... perhaps switch to one with a dependency, e.g. look-up?}
-\sergey{ok, we need to see if look-up would fit here}
-\\\\
-
-The $\generategroups(\textit{t,{\cal G},C})$ method will construct this CSP and query a solver for all satisfying solutions. This is the required result set.
+  Every solution is a tuple of values for~$V_x$, $V_y$ and~$V_z$ corresponding directly with a superblock assignment over~\blocks.
+\end{example}
 
@@
 
-Let us illustrate this on one of the spreadsheets from the gathered dataset.
-In Figure \ref{fig:autocompletion_example}, we see two tables the upper one is completed and in the lower one values in the last raw are not yet filled, then the system finds a satisfying constraint (\textit{SUMIF}) and based on it proposes filling values.
+Let us illustrate this on one of the spreadsheets from the gathered dataset. In Figure \ref{fig:autocompletion_example}, \sergey{to be finished}
 

### 2016-05-20 Paper/ECAI/tabular.tex
Reworked subblock assignments

@@
 Given a constraint template~$s$ and the set of all blocks~\blocks, the goal is to find all combinations of blocks that satisfy the constraint signature.
-Let there be $n$ blocks and let~$s$ have~$k$ arguments, then a simple generate-and-test method would have to test $n^k$ combinations (given that blocks can be repeated), which can grow very rapidly with $n$, depending on $k$.
+Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.
 
@@
 
-\subsubsection{Constraints over satisfying subgroups} \label{sec:algo:subgr}
-In the previous step, $\generategroups(\textit{t,{\cal G},C})$ generates tuples of candidate groups $(G_1, ..., G_n)$ for the constraint template $t$ with $n$ arguments.
-In this step, $\findassignment(t, (G_1, \dots, G_n))$, the goal is to enumerate the subgroups $(G'_1, ..., G'_n)$ (for all $i$, $G'_i \subseteq G_i$) that satisfy both the signature and the definition of the constraint template $t$.
-The main difference with the previous step is that the group generation reasons about the \textit{properties} of the groups (type, size, length, ...), while this step reasons about the actual \textit{content} of the vectors in the groups, that is, the data.
-
-\paragraph{Example}
+\subsubsection{Constraints over satisfying subgroups}
+\label{sec:algo:subgr}
+In the previous step superblock assignments are generated for a constraint template $s$ with $n$ arguments.
+In this step, given such a superblock assignment $(B_1, \dots, B_n)$, the goal is to discover valid \textit{subassignments}, i.e.
+assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.
+While the previous step reasons about the \textit{properties} of blocks (type, size, length, ...), this step reasons about the actual \textit{content} of the vectors in the blocks, that is, the data.
+
+\begin{example}
   Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
-In the group generation, all numeric groups where the length of $\sg_r$ vector is equal to the number of rows in $\mathbf{\sg_x}$ has been enumerated.
-An example from Table~\ref{table:constraints} is $(\sg_r, \mathbf{\sg_x}) = (T_1[:,3:8],T_1[:,3:8])$.
-Note that this does not match the signature exactly yet: $\sg_r$ contains more than one vector.
-We can now generate all vectors in $T_1[:,3:8]$ for $\sg_r$ and all subgroups of $T_1[:,3:8]$ for $\mathbf{\sg_x}$ and test whether $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$ where $G[j][i]$ indicates the $i$the value of the $j$th vector in $G$.
-An example solution is $(T_1[:7], T_1[:,3:6])$.
-\\\\
-
-One can formulate this as a constraint satisfaction problem, similar to what we did in the previous section, or use a simple generate-and-test approach.
-In contrast to the previous step, the number of subgroups to generate here are typically manageable: $O(n*m^2)$ with $m$ the number of vectors in the largest group; we only consider continuous ranges in subgroups, so a subgroup of size $m$ has $m*(m+1)/2$ possible subgroups to consider.
+  An example superblock assignment from Figure~\ref{fig:main_example} is $(B_r, \mathbf{\sg_x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
+  Note that this assignment does not satisfy the signature yet, $\sg_r$ contains more than one vector and $\psize(\mathbf{\sg_x}) \neq \plength(B_r)$.
+  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition: $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$, where $B[i][j]$ indicates the $i$th value of the $j$th vector in $B$.
+  In Figure~\ref{fig:main_example} there is exactly one such satisfying subassignment: $(T_1[:7], T_1[:,3:6])$.
+\end{example}
+
+The search for subassignments is represented as the generic procedure $\findassignment$.
+However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
+One possible method is to reformulate the problem as a CSP, similar to the approach we employed in the last section.
+Often, however, a simple generate-and-test approach will suffice.
+In contrast to the previous step, the number of subblocks to generate here are typically manageable: $O(n*m^2)$ with $m$ the size of the largest block and $n$ the number of arguments of the template.
+We only consider continuous ranges as subblocks, so for a block of size $m$ there are $m*(m+1)/2$ possible subblocks to consider.
 Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheet.
 
-For this reason, our current implementation uses a simple generate-and-test method that takes the required size/length of the groups into account, and sequentially verifies whether the elements of the vectors satisfy the definition.
-The pseudo-code for this approach is shown in ...\tias{Add pseudocode!!}
-
-\begin{verbatim}
-VERY PSEUDOcode:
-something recursive that handles vectors/
-subgroups differently like this?
-check(Groups, X, Test):
-  if |Groups| == 0:
-    forall i in ?: vector elements...
-      if not Test(X):
-        return []
-    return X
-  // else
-  all = []
-  Group = Groups.pop(0)
-  if vector(Group):
-    forall Vec in Group:
-      all += check(Groups, X + Vec, Test)
-  else:
-    forall Subgroup in Group:
-      all += check(Groups, X + Subgroup, Test)
-  return all
-\end{verbatim}
-
-Note that for some constraint definitions, there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example $G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
+For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
+Given a superblock assignment, all subassignments are generated taking into account the required size, columns or rows.
+The subassignments are sequentially tested using a template-specific procedure \textit{Test} and valid subassignments are returned.
+The \textit{Test} procedure verifies if the definition is satisfied and, if necessary, checks signature constraints that have not yet been enforced (e.g. subtype constraints).
+
+\begin{algorithm}[tbh]
+  \begin{algorithmic}
+    \footnotesize
+    \Procedure{FindSubassignments}{$s$, $(B_1, \dots, B_n)$}
+    \State $n \gets$ number of arguments of template $s$
+    \State $A_{sub} \gets \emptyset$
+    \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i \}$}
+      \If{$disjoint(\{B_1', \dots, B_n'\}) \land Test((B_1',\dots,B_n'))$}
+        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
+      \EndIf
+    \EndFor
+\Return $A_{sub}$
+\EndProcedure
+\end{algorithmic}
+\caption{Generate-and-test for $\findassignment$}
+\label{algo:gen-test-subblocks}
+\end{algorithm}
+
+% \begin{verbatim}
+% VERY PSEUDOcode:
+% something recursive that handles vectors/
+% subgroups differently like this?
+% check(Groups, X, Test):
+%   if |Groups| == 0:
+%     forall i in ?: vector elements...
+%       if not Test(X):
+%         return []
+%     return X
+%   // else
+%   all = []
+%   Group = Groups.pop(0)
+%   if vector(Group):
+%     forall Vec in Group:
+%       all += check(Groups, X + Vec, Test)
+%   else:
+%     forall Subgroup in Group:
+%       all += check(Groups, X + Subgroup, Test)
+%   return all
+% \end{verbatim}
+
+Note that for some constraint definitions, there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
+Such duplicates are avoided by imposing an arbitrary but fixed group ordering on exchangeable blocks and discarding assignments not adhering to the ordering.
+
+ %$G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
 \samuel{\tias{How are they stopped in the code?
@@
 
-\paragraph{Example 2, not used (should be replaced by pseudo-code above?)}
-Sum Product
-
-Implementation: finds subgroups based on assignments\\
-Method $Test(G_R, G_{O_1}, G_{O_2})$\\
-$If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
-$If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
-\\\\
-Method $FindSubgroups(Assignments)$\\
-$Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g.
+% \paragraph{Example 2, not used (should be replaced by pseudo-code above?)}
+% Sum Product
+
+% Implementation: finds subgroups based on assignments\\
+% Method $Test(G_R, G_{O_1}, G_{O_2})$\\
+% $If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
+% $If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
+% \\\\
+% Method $FindSubgroups(Assignments)$\\
+% $Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g.
 vectors)?}
 
+\samuel{MARKER-C}
 \subsection{Implementation}

### 2016-05-23 Paper/ECAI/tabular.tex
added and fixed lucs related work

@@
 \section{Related Work}\label{sec:related_work}
-\sergey{key bullet points for Luc and possibly Samuel and me to make related work section}
-
-\samuel{Two possible takes here: 1) recent interest in topics such as flashfill, etc -> sparked the idea for this project.
-This actually relates to constraint learning which has seen an uptick recently and is a subfield of long-established ILP.
-2) ILP established field trying to learn structure / .., has been brought to the less traditional context (constraint learning), learning constraints / programs has been picked up by specific approaches such as flashfill etc which inspired this research}
-
-flashfill, flashextract, flashmeta \cite{flashfill,flashextract,flashmeta}
-\begin{itemize}
-  \item their supervised vs our unsupervised approach
-  \item they look for a single ``smallest'' solution, we enumerate them all
-  \item they are looking for a function, we solve constraint satisfaction problems
-  \item we do not assume classic row based data layout, we work in the tabular setting
-\end{itemize}
-
-sketch \cite{sketch}
-\begin{itemize}
-  \item look for a constant that would fill in the gap in a program
-  \item tailored for programming languages
-  \item similar to model checking
-  \item looks for a single solution
-  \item similar to constraint satisfaction and sat, where one is interested in a single assignment that works for any potential input
-\end{itemize}
-
-tabular \cite{tabular}
-\begin{itemize}
-  \item language based on the excel tables that specify probabilistic models
-  \item a system for probabilistic inference and similarity mostly in the usage of excel
-  \item probabilistic constraint satisfaction (?) and graphical models
-  \item single solution again
-\end{itemize}
-
-modelseeker \cite{modelseeker} \sergey{Samuel, Luc, probably you would need elaborate here more in details}
-
-\begin{itemize}
-  \item not designed for excel-like data representation (type consistency, groups, etc)
-  \item not designed for excel-like constraints (lookups, conditional ifs, etc)
-  \item does not support user extensions (?)
-\end{itemize}
-
-claudien \cite{claudien} \sergey{Samuel, Luc, you would need to help with this one}
+The system that we have introduced combines ideas from several existing
+approaches from multiple different fields of research.
+First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
+constraints in multiple relations (or tables). However, unlike typical logical and relational learning
+approaches, it focuses on data in spreadsheets and on constraints that can be column or row-wise.
+Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien} or \cite{Llalouet}.
+Still several ideas from logical and relational learning
+proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
+Also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha}
+although this line of research has focused on more specialized techniques for specific types of constraints. Furthermore, the constraints used
+in spreadsheets are often quite different than those in databases.
+Second, there exist several algorithms in the constraint programming community that induce constraint (programs)
+from one or more examples and questions. Two well-known lines of research include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}
+Modelseeker starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints
+from a large constraint library that hold in the vector. To this aim, it cleverly enumerates different tables from the vector. For instance, if the initial vector
+was of dimension 1 x 20, it would consider rearrangements of size 2 x 10, 4 x 5, 5 x 4, … and Modelseeker would then look for both column, row  and even diagonal constraints.
+Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet),
+and also that ModelSeeker’s constraints are the typical ones from the constraint programming community, which are aimed at combinatorial optimization
+and scheduling rather than what is typically used in spreadsheets. Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints
+using techniques inspired on Mitchell’s version spaces to process examples, some variants also asking informative questions to the user. As the conjunction
+for typical constraint programming problems (such as SUDOKU) can be quite large, a key problem is to cope with redundant constraints.   As Modelseeker,
+these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they 
+do not assume the variables are organized in tabular form.
+Third, our work borrows from the seminal work of Gulwani et al.
+\cite{flashfill} on program synthesis for tabular data that is incorporated in Microsoft’s Excel.
+In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians.
+The first row might contain the values A1 = Blair and B1 = Tony, and the user might then enter in cell C1 the value T.B.
+At that point FlashFill would generate a program that produces the output T.B.
+for the inputs Blair and Tony, would also apply it to automatically generate values for the other politicians (rows) in the table.
+While it might need an extra example to deal with names such as Van Rompuy, Herman and determine whether it should output H.V.or H.V.R.,FlashFill learns correct programs from very few examples.
+Flashfill has been extended in e.g.
+FlashExtract \cite{flashextract} to produce a set of normalized database tables starting from naively filled out spreadsheets.
+However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.
+In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.
+Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
+Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
+However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify “hard” constraints.
 

### 2016-05-23 Paper/ECAI/tabular.tex
Introduce optimizations

--- 2016-05-20 Paper/ECAI/tabular.tex
@@
     & \numeric(V_x) \land \numeric(V_y) \land \numeric(V_z) \\
-    & \por(V_x) = \mathit{row} \implies \pcols(V_x) = 1 \\
-    & \por(V_x) = \mathit{row} \implies \prows(V_x) = 1 & \\
+    & (\por(V_x) = \mathit{row}) \implies (\pcols(V_x) = 1) \\
+    & (\por(V_x) = \mathit{column}) \implies (\prows(V_x) = 1) & \\
     & \plength(V_y) \geq 2 \land \plength(V_z) \geq 2 \\
@@
 \samuel{MARKER-C}
-\subsection{Implementation}
+\subsection{Optimizations}
+This section discusses various essential design decisions and optimizations.
+The aim is to improve the algorithm in three areas: 1) speed; 2) redundant constraints; and 3) extensibility.
 
@@
 
-\paragraph{Constraint solvers}
+\subsubsection{Constraint solvers}
 Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
@@
 
-\paragraph{Redundancy}
+\subsubsection{Redundancy}
 There are only types of redundancy that we eliminate during search. The first is the removal of symmetric attribute assignments as explained in Section~\ref{sec:algo:subgr}.
@@
 
-\paragraph{Constraints}
+\subsubsection{Constraints}
 The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.

### 2016-05-23 Paper/ECAI/tabular.tex
figures, related work fixes

@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency.png}
+  \includegraphics[width=0.65\linewidth]{figures/constraint_dependency.png}
   \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
@@
 \subsection{Autocompletion}
-\begin{figure}[thb]
-  \begin{center}
-    \includegraphics[width=0.33\textwidth]{figures/autocompletion_example.png}
-  \end{center}
-  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
-  \label{fig:autocompletion_example}
-\end{figure}
 \sergey{I think the time discussion is irrelevant, should it be cut out completely?}
@@
 \begin{figure*}[thb]
+\begin{subfigure}{.70\textwidth}
   \begin{center}
-    \includegraphics[width=0.85\textwidth]{figures/fbi_figure_highlighted.png}
+    \includegraphics[width=\textwidth]{figures/fbi_figure_highlighted.png}
   \end{center}
@@
   \label{fig:fbi}
+\end{subfigure}
+\hfill
+\begin{subfigure}{.29\textwidth}
+\begin{center}
+    \includegraphics[width=\textwidth]{figures/autocompletion_example.png}
+  \end{center}
+  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
+  \label{fig:autocompletion_example}
+
+\end{subfigure}
+
 \end{figure*}
@@
 \section{Related Work}\label{sec:related_work}
-The system that we have introduced combines ideas from several existing
+\sname combines ideas from several existing
 approaches from multiple different fields of research.
+
 First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
@@
 approaches, it focuses on data in spreadsheets and on constraints that can be column or row-wise.
-Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien} or \cite{Llalouet}.
+Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,Llalouet}.
 Still several ideas from logical and relational learning
@@
 in spreadsheets are often quite different than those in databases.
+
 Second, there exist several algorithms in the constraint programming community that induce constraint (programs)
-from one or more examples and questions. Two well-known lines of research include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}
-Modelseeker starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints
-from a large constraint library that hold in the vector. To this aim, it cleverly enumerates different tables from the vector. For instance, if the initial vector
-was of dimension 1 x 20, it would consider rearrangements of size 2 x 10, 4 x 5, 5 x 4, … and Modelseeker would then look for both column, row  and even diagonal constraints.
-Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet),
-and also that ModelSeeker’s constraints are the typical ones from the constraint programming community, which are aimed at combinatorial optimization
-and scheduling rather than what is typically used in spreadsheets. Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints
-using techniques inspired on Mitchell’s version spaces to process examples, some variants also asking informative questions to the user. As the conjunction
-for typical constraint programming problems (such as SUDOKU) can be quite large, a key problem is to cope with redundant constraints.   As Modelseeker,
-these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they 
-do not assume the variables are organized in tabular form.
+from one or more examples and questions. Two well-known lines of research include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
+The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
+To this aim, it cleverly enumerates different tables from the vector.
+For instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
+Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and also that ModelSeeker’s constraints are the typical ones from the constraint programming community, which are aimed at combinatorial optimization and scheduling rather than what is typically used in spreadsheets.
+Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints using techniques inspired on Mitchell’s version spaces to process examples, some variants also asking informative questions to the user.
+As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
+As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.
+
 Third, our work borrows from the seminal work of Gulwani et al.
@@
 In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians.
-The first row might contain the values A1 = Blair and B1 = Tony, and the user might then enter in cell C1 the value T.B.
-At that point FlashFill would generate a program that produces the output T.B.
-for the inputs Blair and Tony, would also apply it to automatically generate values for the other politicians (rows) in the table.
-While it might need an extra example to deal with names such as Van Rompuy, Herman and determine whether it should output H.V.or H.V.R.,FlashFill learns correct programs from very few examples.
-Flashfill has been extended in e.g.
+The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''.
+At that point FlashFill would generate a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', would also apply it to automatically generate values for the other politicians (rows) in the table.
+While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
+Flashfill has been extended to e.g.
 FlashExtract \cite{flashextract} to produce a set of normalized database tables starting from naively filled out spreadsheets.
@@
 Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
+
 Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.

### 2016-05-24 Paper/ECAI/tabular.tex
Small edits table

--- 2016-05-23 Paper/ECAI/tabular.tex
@@
 \begin{table*}
-\caption{Overview of constraint templates implemented in our system, which covers the most popular spreadsheet constraints we encountered.
+\caption{Overview of constraint templates implemented in \sname.
 The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
-Subgroups in normal font have to be vectors, while subgroups in \textbf{bold} may contain more than one vector.\label{table:constraints}}
+Subgroups in normal font have to be vectors, while subgroups in \textbf{bold} may contain more than one vector.
+Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.}\label{table:constraints}
   {\centering
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    $\ecalldiff{\sg_x}$
+    $\ecalldiff{\sg_x}^*$
       & $\discrete(\sg_x)$
@@
     %X = Y & & \\
-    $\ecperm{\sg_x}$
+    $\ecperm{\sg_x}^*$
       & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
@@
       \\ \hline
-    \ecfkey{\sg_{fk}}{\sg_{pk}} & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$ & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
+    $\ecfkey{\sg_{fk}}{\sg_{pk}}^*$ & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$ & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
     % TODO check out

### 2016-05-24 Paper/ECAI/tabular.tex
Comments formalization - dependencies

@@
 
-\paragraph{Dependencies}
-\sergey{question: the guidlines say caption for tables above the tables, but dsaa papers from 2014 and 2015 (I have checked three) put it below, that's confusing...}
+\subsubsection{Dependencies}
+\label{sec:form:dependencies}
 In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g.
@@
 
-One way to deal with such equivalences is to identify a \textit{canonical} constraint among the equivalent ones.
+\samuel{Should this not be part of approach / optimizations / redundancy?}One way to deal with such equivalences is to identify a \textit{canonical} constraint among the equivalent ones.
 For example, for the product constraint above, we could define a lexicographic ordering over the blocks and only consider the constraint with the smallest ordering.

### 2016-05-24 Paper/ECAI/tabular.tex
Updated appraoch

@@
 \newcommand{\learnconstraints}{\format{LearnConstraints}}
-\newcommand{\findassignment}{\format{FindSubassignments}}
+\newcommand{\findassignment}{\format{Subassignments}}
 \newcommand{\postprocess}{\format{pruneRedundant}}
@@
     \footnotesize
-    \State \textbf{Input:} $\groups$ -- groups, $s$ -- constraint template
+    \State \textbf{Input:} $\blocks$ -- blocks, $\temps$ -- constraint templates
     \State \textbf{Output:} $C$ -- learned constraints
-    \Procedure{LearnConstraints}{$\groups$, $s$}
+
+    \Procedure{LearnConstraints}{$\blocks$, $\temps$}
       \State $C \gets \emptyset$ %\Comment{The set of constraints}
+      \ForAll{$s~\mathbf{in}~\temps$}
         \State $n \gets$ number of arguments of template $s$
-    \State $A \gets \generategroups(s, \groups)$
+        \State $A \gets \generategroups(s, \blocks)$
         \ForAll{$(B_1, \dots, B_n) \in A$}
-    \ForAll{${(B_1',\dots,B_n')\in}~\findassignment(s, (B_1,\dots,B_n))$}
+        \State $A' \gets \findassignment(s, (B_1,\dots,B_n))$
+        \ForAll{$(B_1',\dots,B_n') \in A'$}
             \State $C \gets C \cup \{ s(B_1', \dots, B_n') \}$
+          \EndFor
         \EndFor
@@
 The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}.
-Given a constraint template~$s$, the algorithm first generates all combinations of superblocks (i.e. superblock assignments) that are consistent with the signature.
+For every constraint template~$s$, the algorithm first generates all combinations of superblocks (i.e. superblock assignments) that are consistent with the signature.
 In a second step it looks at every superblock assignment and finds subassignments that satisfy the signature and definition of~$s$.
@@
 To reason over the blocks, we automatically add to the constraint program a set of background facts that contain the properties of each block, namely its type, size, number of columns and rows, length, orientation, and table.
-\samuel{Review: We also add all previously found constraints of constraint templates that are implied by this template.}
 The actual constraints correspond to the requirements defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
@@
 \begin{example}
-  \samuel{Is this still necessary?}
   Consider the template $\ecsumprod{x}{y}{z}$, then the generated CSP will contain three variables $V_x, V_y, V_z$ corresponding to arguments $x, y$ and $z$ respectively.
@@
     \footnotesize
-    \Procedure{FindSubassignments}{$s$, $(B_1, \dots, B_n)$}
+    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
     \State $n \gets$ number of arguments of template $s$
@@
 
-Note that for some constraint definitions, there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
-Such duplicates are avoided by imposing an arbitrary but fixed group ordering on exchangeable blocks and discarding assignments not adhering to the ordering.
-
  %$G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
@@
 
-\samuel{MARKER-C}
 \subsection{Optimizations}
-This section discusses various essential design decisions and optimizations.
-The aim is to improve the algorithm in three areas: 1) speed; 2) redundant constraints; and 3) extensibility.
-
-\samuel{Start: Moved template ordering}
-\subsubsection{Template ordering}
-As discussed before, a constraint may imply another constraint.
+This section discusses various essential design decisions and optimizations aiming to improve the speed and extensibility of our system as well as to find less redundant constraints.
+
+\subsubsection{Template dependencies}
+As discussed in Section~\ref{sec:form:dependencies}, some constraints depend on other constraints.
 This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as \textit{LOOKUP} depending on \textit{FOREIGNKEY}.
+If a constraint template~$s_1$ includes template~$s_2$ in its signature, we say that~$s_1$ depends on~$s_2$.
 
 In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
-In our method, we use these dependencies/implications to define a dependency graph between the templates.
-We assume the signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
+Our approach uses the dependencies between templates to define a dependency graph.
+We assume that signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
 Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}. 
@@
   \includegraphics[width=0.8\linewidth]{figures/constraint_dependency.png}
-  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_1$ is implied by $s_2$ because the \CSignature of~$s_2$ depends on $s_1$ to hold on a specific subset of $s_2$ its groups.
-    The graph is extracted from the Signature column in Table \ref{table:constraints}.
+  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_2$ (the \CSignature of~$s_2$ includes $s_1$).
+    The graph is extracted from the Signature column in Table~\ref{table:constraints}.\samuel{Correct aggregates}
   }
@@
 
-The $\constrainttorder(\constraints,\dependencies)$ module in our method takes a set of {\template}s~(\constraints) and a DAG~(\dependencies) as input and produces a linear ordering of the templates such that the more general, implied constraint templates (the requirements) are ordered before the templates that depend on them.
-
-\paragraph{Example}
-As mentioned before, $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ implies $\ecfkey{G'_1}{H'_1}$ and hence the $\textit{LOOKUP}$ constraint template depends on $\textit{FOREIGNKEY}$ in Figure~\ref{fig:learning_order} and will be earlier in the ordering.
-
-% not helpful:
-%The \ecrank{v_y}{v_x} constraint does not have any dependencies, however in a restricted version without ties, groups assigned to $v_y$ would always be a permutation of numbers $1, 2, ..., \textit{length}(v_y)$.
-\samuel{End: Moved template ordering}
-
-We provide some more details of our implementation of the above algorithmic approach, to ease reproducibility.
+Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
+Concretely, the input templates~$\temps$ (see Algorithm~\ref{algo:tcl}) are reordered to agree with the partial ordering imposed by~\dependencies.
+
+The dependency graph represents a partial ordering and is used produce a linear ordering of the given templates such that more general templates occur before templates that depend on them (e.g.
+\textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
+Considering Algorithm~\ref{algo:tcl}, this step can be seen as preprocessing the input~\temps.
+Learning in this order allows templates to use constraints they depend on as these will have already been found.
+This is exploited in the \generategroups step to speedup the algorithm by reducing the search space.
+Let us illustrate this step with an example.
+
+\begin{example}
+  Consider $\ecfkey{B_{fk}}{B_{pk}}$ whose signature includes $\ecalldiff{B_{pk}}$ and assume that~$C$ are the constraints found for \textit{ALLDIFFERENT}.
+  Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
+  $P$~can be seen as a set of partial assignments in which the blocks for certain arguments have been fixed.
+\end{example}
+
+The \generategroups implementation now takes into account previously found constraints to obtain partial assignments~$P$ for a template.
+Instead of generating one CSP to find all assignments, a CSP is generated for every~$p \in P$ which finds all assignments completing~$p$.
+The procedure to generate CSPs remains the same, except that the partial assignment is used to fix the domains of the dependent variables.
+
+This approach has the additional benefit of ensuring that the (sub)blocks assigned to the dependent arguments already satisfy both the signature and definition of the original template.
+Therefore, the implementation of $\findassignment$ does not have to take these blocks into account.
 
 \subsubsection{Constraint solvers}
+\samuel{Tias, Sergey, have a look at his section, I'm torn about rewriting it, feeling it should either be reduced or expanded, especially the second part.}
 Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
@@
 
-\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the stand-alone Constraint Programming language MiniZinc \cite{minizinc} with the Gecode 4.4.0 solver which supports floats, as well as an embedded Python-based CP solver \cite{python_constraint}.
+\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the MiniZinc language \cite{minizinc} with the stand-alone Gecode 4.4.0 solver (which supports floats) and uses an embedded Python-based CP solver \cite{python_constraint}.
 The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints.
@@
 \subsubsection{Redundancy}
-There are only types of redundancy that we eliminate during search. The first is the removal of symmetric attribute assignments as explained in Section~\ref{sec:algo:subgr}.
-
-The second is that there be \textit{overlapping} subgroups that satisfy the definition, such as \tias{There is no example in the example figure?
-I was thinking of max/min, but it is over the entire vector always...}.
-In this case we chose to omit those subgroups that are fully contained within a larger subgroup since this can lead to over-fitting.
+We consider two types of redundancy that we aim to eliminate during the search.
+As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
+Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
+In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
+\samuel{Mention B1*B2 vs PRODUCT?}
+
+The second type of redundancy we aim to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
+For example, consider the constraint \ecsumc{B_r}{B[1:n]} where $B[k:n]$ consists of only zeros, then \ecsumc{B_r}{B[1:j]} will be true for all $k - 1 \leq j \leq n$.
+In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.
+In some cases the maximal subblock might falsely include irrelevant columns, in this case we suggest to put these irrelevant columns into a separate superblock.
 
@@
 The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
-All these constraints are available in popular spreadsheet software, except the first two constraints (alldifferent and permutation).
-These two constraints are \textit{structural} constraints, which we added so they can be used in the signature of other constraint templates; they are also popular constraints in constraint satisfaction~\cite{modelseeker}.
+Their inclusion was based on the analysis of the most popular Excel constraints as well as the spreadsheets we encountered.
+All these constraints are available in popular spreadsheet software, except for the structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}).
+These constraints added so they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
+
+Note the the template $\ecdiff{B_r}{B_1}{B_2}$ can be rewritten as $B_1 = B_r + B_2$.
+Our choice to use difference is a hardcoded preference based on the prevalence of this type of constraint in financial spreadsheet data.
 
@@
 
-Note that \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?}
-
-\tias{Not used, should be cut!?:}
-Generally \CSignature and / or the \CFunction can be relaxed to capture more constraints in the data, however, this also comes at a performance penalty.
-\samuel{More on this / example?} \tias{Yes, either explain in detail or cut}
+%Note that \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?}
 

### 2016-05-25 Paper/ECAI/tabular.tex
Various improvements

@@
   \begin{center}
-    \includegraphics[width=1\textwidth]{figures/Demo.png}
+    \includegraphics[width=1\textwidth]{figures/Demo2.png}
   \end{center}
@@
 
-Our algorithm requires tables to be specified, this can, however, be done in various ways.
-Tables are specified in a JSON configuration file which contains the table names, their coordinates within a CSV file and optionally the orientation of certain tables.
+Our algorithm requires tables to be specified.
+This information is provided to \sname in the form of a configuration file containing the name, coordinates within a spreadsheet and optionally the orientation of every table.
 The orientation can be used to indicate that a table should be interpreted as only rows or only columns.
 
-The generation of this JSON file can be done in many ways, ranging from manually specifying the tables to automatically detecting tables.
-Requiring the user to manually generate these files decreases the usability of our tool in a standalone setting, therefore, we offer two proof of concept approaches that aim to facilitate the specification of tables within a CSV.
+This information can be provided manually or generated automatically.
+We offer two proof of concept approaches that aim to facilitate the specification of tables.
 
@@
 The tool then generates the table definitions automatically.
-\\\\
-Aside from extending the automatic detection to be more applicable, one could combine the visual and automatic detection to let the user approve or correct detected table.
 
@@
 If a table has no fixed orientation, the algorithm will attempt to construct both row-blocks and column-blocks.
-In order to find row-blocks (column-blocks), the type of each row (column) is determined, i.e. whether it is numeric or textual.
-All adjacent rows (columns) that have the same type are merged to form a block. 
+In order to find blocks, the type of each vector is determined, i.e. whether it is numeric or textual.
+All adjacent vectors that have the same type are merged to form a block. 
 If any row (column) of the table contains both numeric and textual values, it is type inconsistent and the algorithm will not produce any row-blocks (column-blocks) for that table.
 For tables that have a fixed orientation only blocks of that orientation are built.
-
-While this step can be done algorithmically, there are cases where specifying finer-grained blocks manually may be beneficial.
-Firstly, blocks may group data from different contexts.
-For example, consider the type consistent block $B_3=\range{T_1}{\rangeall}{\rangeto{3}{8}}$ from the example in Figure~\ref{fig:main_example}: $\range{T_1}{\rangeall}{8}$ is contextually different from $\range{T_1}{\rangeall}{\rangeto{3}{7}}$, therefore, we could split~$B_3$ into two blocks.
-\samuel{Secondly, accidental blocks}
-Finer grained groups can reduce the amount of redundant constraints that are found.
 
@@
 %We will use constraint satisfaction technology to partly answer this challenge.
+\tias{No, separate selection of superblocks from generation of subblocks}
+\samuel{I do not fully agree, the property vs content aspect is still important, it is combined with superblock / subblock but this could be done differently}
 To answer this challenge we separate checking the signature (concerning properties of a block) of a constraint template from checking the definition (concerning the values in a block) of a constraint template.
@@
 \begin{algorithm}[t]
-  \begin{algorithmic}
+  \begin{algorithmic}[1]
     \footnotesize
@@
       \EndFor
-      \Return $C$
+      \State \Return $C$
     \EndProcedure
@@
 \begin{table}[htb]
-  \caption{Translation of signature requirements to superblock constraints}
+  \caption{Translation of signature requirements to superblock constraints.
+The requirements that do not need to be relaxed are: $\plength(B) = x$, $\por(B) = x$, $\ptable(B) = x$ and $\psize(B) \geq x$.}
   \label{tbl:translation}
   \begin{tabularx}{\linewidth}{lX}
-    \textbf{Constraint} & \textbf{Superblock constraint} \\ \hline \hline
-    $\plength(B) = x$ & $\plength(B) = x$ \\ \hline
-    $\por(B) = x$ & $\por(B) = x$ \\\hline
-    $\ptable(B) = x$ & $\ptable(B) = x$ \\ \hline
-    $\ptype(B) = x$ & $\mathit{supertype}(\ptype(B)) = \mathit{supertype}(x)$ \\ \hline
-    $\psize(B) \geq x$ & $\psize(B) \geq x$ \\ \hline
+    \textbf{Requirement} & \textbf{Superblock constraint} \\ \hline \hline
+    $\ptype(B) = x$ & $\mathit{basetype}(\ptype(B)) = \mathit{basetype}(x)$ \\ \hline
     $\psize(B) = x$ & $\psize(B) \geq x$ \\ \hline
@@
 \begin{example}
-  Consider the template $\ecsumprod{x}{y}{z}$, then the generated CSP will contain three variables $V_x, V_y, V_z$ corresponding to arguments $x, y$ and $z$ respectively.
-  Given blocks~\blocks, the domain of these variables are $D(V_x) = D(V_y) = D(V_z) = \{ 1, \dots, |\blocks| \}$.
+Br and Bx are numeric; rows(Bx) ⇤ 2; and columns(Bx ) = length(Br )
+
+  Consider the constraint template $\ecsumc{r}{x}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$r$ and~$x$.
+  Given blocks~\blocks, the domain of these variables are $D(V_r) = D(V_x) = \{ 1, \dots, |\blocks| \}$.
   Finally, the signature will be translated into constraints:
   \begin{align*}
-    & \numeric(V_x) \land \numeric(V_y) \land \numeric(V_z) \\
-    & (\por(V_x) = \mathit{row}) \implies (\pcols(V_x) = 1) \\
-    & (\por(V_x) = \mathit{column}) \implies (\prows(V_x) = 1) & \\
-    & \plength(V_y) \geq 2 \land \plength(V_z) \geq 2 \\
-    & \plength(V_y) = \plength(V_z)
+    & \numeric(V_r) \land \numeric(V_x) \land \prows(V_x) \geq 2 \\
+    & (\por(V_x) = \mathit{row}) \Rightarrow (\pcols(V_x) \geq \plength(V_r)) \\
+    & (\por(V_x) = \mathit{column}) \Rightarrow (\pcols(V_x) = \plength(V_r))
   \end{align*}
-  Every solution is a tuple of values for~$V_x$, $V_y$ and~$V_z$ corresponding directly with a superblock assignment over~\blocks.
+  Every solution is a tuple of values for~$V_r$ and~$V_x$ corresponding directly with a superblock assignment over~\blocks.
 \end{example}
@@
 In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.
-In some cases the maximal subblock might falsely include irrelevant columns, in this case we suggest to put these irrelevant columns into a separate superblock.
+
+In some cases a maximal subblock might falsely include irrelevant columns.
+By manually specifying more fine-grained blocks as input this behavior can often be avoided.
+Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$.
+If the target constraint is $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, the last vector of~$B_1$ can be split off into a separate block.
+Alternatively, the algorithm can output all valid constraints, e.g.
+$\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$, however, this might produce many redundant constraints and introduce constraints that accidentally cover too little data.
 

### 2016-05-26 Paper/ECAI/tabular.tex
Approach tweaks

@@
 Green borders indicate headerless tables; the blue elements are added to clarify some constraints and relations.
-\tias{TODO: headers not in green boxes} \sergey{show blocks?
-orientation only over blocks, not over tables?}\tias{Use specific constraint names, e.g.
+\tias{Use specific constraint names, e.g.
 'SUM-row' above Total in T1}}
@@
   \vspace{-10pt}
-  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).}
+  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).\samuel{Not the desired constraints ask : \tias{}}}
   \label{fig:sol_example}
@@
 \subsubsection{Generating superblock assignments}
-%The goal of this step is to prune out impossible groups from the beginning by looking at their properties rather than their actual content (\CSignature vs \CFunction).
 Given a constraint template~$s$ and the set of all blocks~\blocks, the goal is to find all combinations of blocks that satisfy the constraint signature.
@@
 assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.
+\samuel{Is this repetition?}
 While the previous step reasons about the \textit{properties} of blocks (type, size, length, ...), this step reasons about the actual \textit{content} of the vectors in the blocks, that is, the data.
@@
 We only consider continuous ranges as subblocks, so for a block of size $m$ there are $m*(m+1)/2$ possible subblocks to consider.
-Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheet.
+Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheets.
 
@@
     \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i \}$}
-      \If{$disjoint(\{B_1', \dots, B_n'\}) \land Test((B_1',\dots,B_n'))$}
+      \If{$disjoint(\{ B_1', \dots, B_n' \}) \land Test((B_1',\dots,B_n'))$}
         \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
@@
 In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
-\samuel{Mention B1*B2 vs PRODUCT?}
+Moreover, \sname currently uses product (\ecprod{B_R'}{B_1'}{B_2'}) instead of division (\eccalc{B_1'}{B_R' / B_2'}) and difference (\ecdiff{B_r}{B_1}{B_2}) instead of sum (\eccalc{B_1}{B_r + B_2}).
+These choices are hardcoded preferences based on the prevalence of difference and product in financial spreadsheet data, however, these constraints can be trivially rewritten if necessary.
 
@@
 These constraints added so they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
-
-Note the the template $\ecdiff{B_r}{B_1}{B_2}$ can be rewritten as $B_1 = B_r + B_2$.
-Our choice to use difference is a hardcoded preference based on the prevalence of this type of constraint in financial spreadsheet data.
 

### 2016-05-26 Code/src/core/constraint.py
Small code tweaks

@@ -273,6 +273,7 @@ class ConditionalAggregate(Constraint):
         source = ConstraintSource(variables, all_diff, {all_diff.x.name: "OK"})
         filters = [SameLength([self.o_key, self.result]), SameLength([self.f_key, self.values]),
                    SameTable([self.f_key, self.values]), Not(SameTable([self.f_key, self.o_key])),
+                   #SameTable([self.o_key, self.result]), #TODO think about this
                    NotPartial([self.o_key]), SameType([self.f_key, self.o_key]),
                    SameOrientation([self.o_key, self.result]), SameOrientation([self.f_key, self.values])]
         p_format = "{R} = " + name.upper() + "IF({FK}={OK}, {V})"

### 2016-05-26 Code/src/workflow.py

@@ -76,14 +76,16 @@ def main(csv_file, groups_file, verbose, silent=False, parse_silent=False, const
     ordered = order_constraints(supported)
 
     for constraint in ordered:
+        if verbose and not silent:
+            print(constraint.name, end=" ", flush=True)
         t_start = time.time()
         assignments = manager.find_assignments(constraint, groups, solutions)
         t_assign = time.time()
         solutions.add(constraint, manager.find_solutions(constraint, assignments, solutions))
         t_end = time.time()
-        f_string = "{name} [assignment time: {assign:.3f}, solving time: {solve:.3f}]"
         if verbose and not silent:
-            print(f_string.format(name=constraint.name, assign=t_assign - t_start, solve=t_end - t_assign))
+            f_string = "[assignment time: {assign:.3f}, solving time: {solve:.3f}]"
+            print(f_string.format(assign=t_assign - t_start, solve=t_end - t_assign))
         if len(solutions.get_solutions(constraint)) > 0 and not silent:
             print("\n".join(["\t" + constraint.to_string(s) for s in solutions.get_solutions(constraint)]))
         if (len(solutions.get_solutions(constraint)) > 0 or verbose) and not silent:

### 2016-05-26 Paper/ECAI/tabular.tex
reworked sec2 part about redundancy

@@
 \newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
+\newcommand{\ecdiv}[3]{\eccalc{#1}{{#2} / {#3}}}
 \newcommand{\ecdiff}[3]{\eccalc{#1}{#2 - #3}}
@@
 
-\tias{This should contain examples that recur later in the text}
-
-For example, for some constraints, the order of some of the arguments may not matter.
+Within the same constraint, there can be equivalent argument assignments if the order of some of the arguments does not matter.
 For example for the product constraint, $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecprod{\sg_r}{\sg_2}{\sg_1}$ so one can be considered redundant to the other.
 
-\samuel{Should this not be part of approach / optimizations / redundancy?}One way to deal with such equivalences is to identify a \textit{canonical} constraint among the equivalent ones.
-For example, for the product constraint above, we could define a lexicographic ordering over the blocks and only consider the constraint with the smallest ordering.
+Two different constraints may be logically equivalent due to their nature, e.g. for addition/subtraction and product/division: $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecdiv{\sg_1}{\sg_r}{\sg_2}$.
+
+When the data has rows or columns containing exactly the same data, then any constraint with such a vector in its argument assignment will also have an argument assignment with the other equivalent vectors.
 
 Because dealing with redundancy is often application-dependent, we explain in the next section our generic method for finding all constraints.
+We describe some optimisations that avoid obvious equivalences in Section~\ref{sec:opts} and will investigate the impact of redundancy in the experiments.
 
@@
 
-\subsection{Optimizations}
+\subsection{Optimizations} \label{sec:opts}
+\samuel{Should this not be part of approach / optimizations / redundancy?}One way to deal with such equivalences is to identify a \textit{canonical} constraint among the equivalent ones.
+For example, for the product constraint above, we could define a lexicographic ordering over the blocks and only consider the constraint with the smallest ordering.
+
 This section discusses various essential design decisions and optimizations aiming to improve the speed and extensibility of our system as well as to find less redundant constraints.

### 2016-05-26 Paper/ECAI/tabular.tex
Updated figure and removed todo's section 2

--- 2016-05-26 Paper/ECAI/tabular.tex
@@
 
-  \begin{subfigure}{.70\textwidth}
+  \begin{subfigure}{.65\textwidth}
   \begin{center}
@@
   \vspace{-10pt}
-  \caption{Example spreadsheet (black words and numbers only).
-Green borders indicate headerless tables; the blue elements are added to clarify some constraints and relations.
-\tias{Use specific constraint names, e.g.
-'SUM-row' above Total in T1}}
+  \caption{Example spreadsheet (black words and numbers only). Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.}
   \label{fig:main_example}
@@
 \hfill
-\begin{subfigure}{.30\textwidth}
+\begin{subfigure}{.35\textwidth}
   {\footnotesize
@@
 %
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}, \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}\\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^1, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^1 \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^1 , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^1 \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^1 \\
 %
@@
 	  % Table 4
-      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}} \\
+      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^2 \\
 %
@@
   \vspace{-10pt}
-  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).\samuel{Not the desired constraints ask : \tias{}}}
+  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).
+  $^1$(Spurious constraints) $^2$(Redundant constraints not originally present)}
   \label{fig:sol_example}
 \end{subfigure}
+  \caption{Running example}
 \end{figure*}
@@
 \item \textit{columns}: also orientation specific, for row-based blocks this is the length, for column-based blocks the size
-\item \tias{What is missing is 'discrete', not needed?}
 \end{itemize}
@@
 \begin{example}
-Consider headerless Table~1 in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e.
-they contain both numeric and textual data).
-The table can be partitioned into the following five column-oriented blocks:
-{\small
-\begin{align*}
-&B_1 = \range{T_1}{\rangeall}{1},
-&B_2 = \range{T_1}{\rangeall}{2},\\
-&B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
-&B_4 = \range{T_1}{\rangeall}{9},\\
-&B_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}.
-\end{align*}
-\tias{Indicate in figure? see also Sergey's comment in figure}
-}
+Consider the (headerless) table~$T_1$ in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
+However, the table can be partitioned into five column-oriented blocks $B_1, B_2, B_3, B_4, B_5$, as shown in the figure ($B_1 = \range{T_1}{\rangeall}{1}$, $B_2 = \range{T_1}{\rangeall}{2}$, $B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
+% {\small
+% B_1 = \range{T_1}{\rangeall}{1},
+% B_2 = \range{T_1}{\rangeall}{2},
+% B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
+% B_4 = \range{T_1}{\rangeall}{9},
+% B_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}
 \end{example}
@@
 \textbf{Block containment $\sqsubseteq$.} 
-A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (single orientation, contiguous, type consistent) and each of the vectors in $B'$ is also in $B$.
-For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=T[\rangeto{a}{b},:] \wedge B'=T[\rangeto{a'}{b'},:] \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
+A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (single orientation, contiguous, type consistent) with the same orientation and table, and each of the vectors in $B'$ is also in $B$.
+For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=\range{T}{\rangeto{a}{b}}{\rangeall} \wedge B'=\range{T}{\rangeto{a'}{b'}}{\rangeall} \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
 \end{definition}
-\sergey{do we need to mention explicitly that blocks have the same orientation and are in the same table?}
+
 We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.

### 2016-05-26 Paper/ECAI/tabular.tex
Applications

@@
 \subsection{Autocompletion}
-\sergey{I think the time discussion is irrelevant, should it be cut out completely?}
-
-Autocompletion can be seen as using knowledge derived from data at time~$t_0$ and new input data at time~$t_1 > t_0$ to predict data that the user has not yet written.
-Therefore, constraints are learned at time~$t_0$ and combined with data from a new row~$i$ (column~$j$) that the user has added to a table~$T$ by time~$t_1$.
-Constraints that do not conflict with the added data are considered \textit{viable}.
-Viable constraints that have enough input data to compute the result for a blank cell~$T[i,j]$ in the new row (column) are \textit{active}, which means they can be used to predict the outcome for that cell.
-\textit{Active} constraints can be used to autocomplete blank cells such that the user does not have to fill in the values manually.
-Once the user has stopped editing the new row~$i$ (column~$j$), the table~$T$ is updated to include~$i$ ($j$) and constraints are learned once again.
+Autocompletion can be seen as using knowledge derived from a snapshot of spreadsheet data to predict new values before they are written by the user.
+The snapshot~($D_0$) is used to learn constraints and these are combined with new data~($\Delta D$) that the user has added.
+Using~$\Delta D$ we can keep track of which vectors have changed and which vectors are now incomplete.
+A \textit{functional} constraint corresponds to a spreadsheet formula.
+Such a constraint, e.g. $\ecsumc{B_r}{\mathbf{B_x}}$, is able to compute a result~($B_r$) based on its input~($\mathbf{B_x}$).
+If all the input vectors / blocks have been filled and the result is incomplete, the constraint can predict the result value(s).
 
 Let us illustrate this on one of the spreadsheets from the gathered dataset.
-In Figure \ref{fig:autocompletion_example}, we see two tables the upper one is completed and in the lower one values in the last raw are not yet filled, then the system finds a satisfying constraint (\textit{SUMIF}) and based on it proposes filling values.
-
-\sergey{there should be a discussion on what to do if multiple constraints satisfy the data and propose predictions, just a sketch here for now}
-If multiple constraints satisfy the data and allow autocompletion, then there is a number of ways to handle it.
-Firstly, autocompletion can be disabled.
-Secondly, the most specific constrain can be applied.
-Thirdly, if one is not more specific than the other, one needs to introduce an autocompletion order, which would resolve the problem by picking only one constraint to suggest the values.
+Consider Figure \ref{fig:autocompletion_example} containing two tables.
+By learning constraints on the tables excluding the last row, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
+The new data $\Delta D$ consists of the new value \textit{West} in the last row.
+Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
+The \textit{SUMIF} constraints have enough input to fill the incomplete vectors and predict the values in the last row: $\range{T_2}{4}{2} = 1547$ and $\range{T_2}{4}{3} = 428128$.
+
+In case multiple constraints predict different values for the same cell, the algorithm will have to either prioritize constraints based on their specificity or a heuristic, not make a prediction, or present the choice to the user.
 
@@
 
-\sergey{Either we need to elaborate on the online detection or remove offline-online and simplify the rest, otherwise it is not clear what is going on; again I don't see any point in time here}
+\sergey{Either we need to elaborate on the online detection or remove offline-online and simplify the rest, otherwise it is not clear what is going on}\samuel{Is this still the case?}
 
@@
 The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
-However, instead of considering viable constraints, we look at \textit{conflict} constraints, i.e. constraints that were learned at time~$t_0$ but do not agree with the newly added data at time~$t_1$.
+However, we look at \textit{conflict} constraints, i.e. constraints that were learned on the snapshot data but do not agree with the newly added data~$\Delta D$.
 Such conflict constraints can indicate errors in the input.
-Since conflict constraints could also be accidental constraints that are correctly invalidated by the new data, the size of the data at time ~$t_0$ should be large enough.
+However, such conflicts can be wrongly caused by spurious constraints.
+Therefore, conflicts should only be detected if the snapshot data is large enough.
 
@@
 When run on this sheet, our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
-A possible approach to detect such missing or wrong values is to look at a sample of the table that excludes the erroneous row and compare the constraints learned on this sample with the constraints learned by adding the row containing the mistake.
+A possible approach to detect such missing or wrong values is to look at a sample of the table that excludes the erroneous row, e.g.
+trying to leave out each row, and compare the constraints learned on this sample with the constraints learned by adding the row containing the mistake.
 The sample needs to large enough such that we have enough confidence that the constraints are correct.
 For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
-If we define the sample to be our data at time~$t_0$ and the added row to be the additional input at~$t_1$, this task can be formulated as online error detection and dealt with in the same way.
+If we define the sample to be the snapshot data and the added row to be the additional input~$\Delta D$, this task can be formulated as online error detection and dealt with in the same way.
 %Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
@@
 
+\tias{write limitations: single cons, complete rows/cols only, (what else?) and deal with it as future work}.
 \bibliographystyle{plain}

### 2016-05-26 Paper/ECAI/tabular.tex
Changes formalization

@@
 \item \textit{type}: a block is type-consistent, so it has a type
-\item \textit{table}: the table that this block belongs to
+\item \textit{table}: the table that the block belongs to
 \item \textit{orientation}: either row-oriented or column-oriented
-\item \textit{size}: the size of a block is the number of vectors in it, that is, the number of rows or columns.
+\item \textit{size}: the number of vectors a block contains
 \item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
@@
 \textbf{Block containment $\sqsubseteq$.} 
-A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (single orientation, contiguous, type consistent) with the same orientation and table, and each of the vectors in $B'$ is also in $B$.
+A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (contiguous, type consistent) with the same orientation and table, and each of the vectors in $B'$ is also in $B$.
 For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=\range{T}{\rangeto{a}{b}}{\rangeall} \wedge B'=\range{T}{\rangeto{a'}{b'}}{\rangeall} \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
@@
 An example is that $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.
-
-\subsection{Constraint templates}
-The goal is to learn constraints over blocks in the data. The knowledge needed to learn a constraint is expressed through {\template}s.
-%
-A \template $s$ is a triple $s=\textit{(\CName, \CSignature, \CFunction)}$:
-%Let us elaborate on this:
-\begin{itemize}
-\item
-\textit{\CName}  specifies the syntactic form of the constraint $c(B_1, ...,B_n)$, that is, the name of the constraint $c$ together
-with $n$ abstract arguments $B_i$.
-Thus a constraint $c$ is viewed as a relation or predicate of arity $n$ in first order logic.
-Note that a function $B_r=f(B_1,...,B_n)$ can be represented with the $(n{+}1)$-ary predicate $c_f(B_r,B_1,...,B_n)$.
-Each argument will have to be instantiated with a block.
-% the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
-
-\item the \textit{\CSignature} defines the properties that the arguments of the predicate must satisfy.
-This can be any property of individual blocks, such as type, length, size and orientation, as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table, or have equal type or length.
-In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
-
-\item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
-Given an assignments of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks.
-%holds, in practice this will be a function that can be called when
-%the arguments are full instantiated to decide whether the constraint is satisfied or not.
-In logical and relational learning this is known as the background knowledge. %, which provides the definition of the predicate.
-\end{itemize}
-
-\begin{example}
-Several constraint templates are illustrated in Table \ref{table:constraints}. A non-trivial example is the constraint template for the row-based sum:
-
-\begin{itemize}
-  \item \CName: $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, where $\sg_r$ and $\mathbf{\sg_x}$ are the arguments;
-
-  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
-The two blocks have to be numeric.
-This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$.
-While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint;
-
-  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
-\end{itemize}
-\end{example}
-
-It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
-From a FOL perspective, the name of the constraint (e.g.
-\textit{SUM}) is just the predicate name, and the arguments $B_r$ and $B_x$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
-This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
-For example for $\ecrank{B_r}{B_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
-
-%When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
-%From a FOL perspective, the name of the constraint ($\mathit{RANK}$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
-%This also holds in our setting: when we write
-%      $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
-%we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ (provided~$T$ has an 8th column).
-%So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
-
-\newcommand{\sigc}{\ensuremath{\format{Sig}_c}}
-\newcommand{\defc}{\ensuremath{\format{Def}_c}}
-%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
-With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
-We say that a signature (definition) of a constraint template with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
-Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would write a Prolog-like clause: $c(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
-Under this interpretation, the term constraint and constraint template can be used interchangeably.
-
-\begin{definition}
-A \textbf{valid argument assignment} of a constraint $c$ is a tuple $(B_1, ..., B_n)$ such that $c(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
-\end{definition}
-
-%\samuel{\tias{Of the right (minimum) dimension??
-Note that this is also sloppy because the ':8' is not considered part of the variable...} True, I changed the wording a bit but personally I would leave it at this now, you can always say you look at $T[:,i:j]$ iff $cols(T) >= max(i,j)$}.
-%Alternatively, and perhaps more naturally, we can interpret the argument as the value it takes in a particular table $T$ such as $T_1$ in Figure 1, that is,
-%${\range{T_1}{\rangeall}{8}} =  [2,1,3,4]$.
-%Then we can also talk about constraints being satisfied.
-Basically, a constraint $c(A_1, ..., A_n)$ is {\em satisfied}  in a set of tables ${\cal T}$ if and only if the constraint holds on the values that the $A_i$ take in the set of tables ${\cal T}$.
-\samuel{\tias{I find this confusing: a group as we defined it is defined on one table, not on multiple tables.
-we should separate the table variable from the 'range' variables then...} waiting for Luc's opinion here}
-%Figure 3 provides a list of constraints that are satisfied by the tables in Figure 1.
 
@@
       & $\sg_{r}[i] = \sum_{i = 1}^{\plength(\sg_{1})} \sg_{1}[i] \times \sg_{2}[i]$.
-      \\
-
   \end{tabularx}}
 \end{table*}
+
+\subsection{Constraint templates}
+The goal is to learn constraints over blocks in the data. The knowledge needed to learn a constraint is expressed through {\template}s.
+%
+A \template $s$ is a triple $s=\textit{(\CName, \CSignature, \CFunction)}$:
+%Let us elaborate on this:
+\begin{itemize}
+\item
+\textit{\CName}  specifies the syntactic form of the constraint $s(B_1, ...,B_n)$, that is, the name of the template together
+with $n$ abstract arguments $B_i$.
+Thus a constraint is viewed as a relation or predicate of arity $n$ in first order logic.
+Note that a function $B_r=f(B_1,...,B_n)$ can be represented with the $(n{+}1)$-ary predicate $s_f(B_r,B_1,...,B_n)$.
+Each argument will have to be instantiated with a block.
+% the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
+
+\item the \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
+This can concern any property of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
+In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
+
+\item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
+Given an assignments of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks.
+%holds, in practice this will be a function that can be called when
+%the arguments are full instantiated to decide whether the constraint is satisfied or not.
+In logical and relational learning this is known as the background knowledge. %, which provides the definition of the predicate.
+\end{itemize}
+
+\begin{example}
+Several constraint templates are illustrated in Table \ref{table:constraints}. A non-trivial example is the constraint template for the row-based sum:
+
+\begin{itemize}
+  \item \CName: $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, where $\sg_r$ and $\mathbf{\sg_x}$ are the arguments;
+
+  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
+The two blocks have to be numeric.
+This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$.
+While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint;
+
+  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
+\end{itemize}
+\end{example}
+
+It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
+From a FOL perspective, the name of the constraint (e.g.
+\textit{SUM}) is just the predicate name, and the arguments $B_r$ and $B_x$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
+This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
+For example for $\ecrank{B_r}{B_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
+
+%When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+%From a FOL perspective, the name of the constraint ($\mathit{RANK}$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
+%This also holds in our setting: when we write
+%      $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
+%we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ (provided~$T$ has an 8th column).
+%So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
+
+\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
+\newcommand{\defc}{\ensuremath{\format{Def}_s}}
+%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
+With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
+We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
+Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would write a Prolog-like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
+Under this interpretation, the term constraint and constraint template can be used interchangeably.
+
+\begin{definition}
+A \textbf{valid argument assignment} of a constraint $s$ is a tuple $(B_1, ..., B_n)$ such that $s(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
+\end{definition}
 
@@
 In the context of spreadsheets, the variables would be (blocks of) cells, and one would be given the actual constraints and functions with the goal of finding the values in the cells.
-The inverse problem is, given only an instantiation of the cells, find the constraints that are satisfied in the spreadsheet.
+The inverse problem is, given only an instantiation of the cells, to find the constraints that are satisfied in the spreadsheet.
 
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $c(B'_1, ..., B'_n)$ where $c$ is a constraint with a corresponding template in ${\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the constraint $c$.
+{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the template~$s$.
 \end{definition}

### 2016-05-26 Paper/ECAI/tabular.tex
rework approach up to Algorithm subsection 1

@@
 The aim of our method is to detect constraints between rows and columns of tabular data.
-Recall that a valid argument assignment for a constraint is an assignment of each argument of the constraint to a block, such that the constraint, that is, its signature and definition, is satisfied.
-\samuel{In general: should it not be: assign blocks to arguments instead of arguments to blocks?}
-\tias{I here use constraint and constraint template interchangeably}
+Recall that a valid argument assignment for a constraint is an assignment of a block to each of the arguments of the constraint, such that the signature and definition of the constraint template is satisfied.
 
@@
 \item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
-\item Generate for each constraint all valid argument assignments in two steps:
+\item Generate for each constraint template all valid argument assignments in two steps:
 \begin{enumerate}
@@
 In principle, one could use a generate-and-test approach by generating all possible blocks from the superblock and testing each combination of blocks for each of the arguments of the constraints.
-However, this ...
-\textit{blows up} ...
-
-Hence, we divide this step into two parts: in the first part, we will not reason over individual blocks, but rather over all possible combinations of superblocks as candidates for the arguments.
-\tias{Example To develop in more detail and clarity:}
-Consider the sum constraints, the row-blocks of T1 are ...
-and column-blocks are ...
-instead of considering all possible blocks in each super-block, we reason at level of superblocks, for example block (with ID) can not be in a row-sum and neither can (block with Items sold total, Max items sold) because its size (1 and 2) is smaller than length of blocks (4).
-The only possible combination of superblocks is (1st Quarter ..
-Rank) for both arguments.
-
-In the second part, we start from such a superblock combination and generate and test all possible block assignments to arguments.
-As we only have to consider the blocks contained in each superblock, this is typically much easier.
-For the above example, \tias{to develop with more clarity} each of the vectors in the block will be considered as candidate for the left-hand side, and one could enumerate all subblocks for the right-hand side and verify for each of the rows in the vector.
-In practice, this can be optimized further.
+However, each superblock of size $m$ has $m*(m+1)/2$ contiguous subblocks, meaning that a constraint with $n$ arguments would have to check $O(n^{m^2})$ combinations of blocks.
+
+Instead, we divide this step into two parts: in the first part, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks.
+% and what combination of superblocks may lead to a valid argument assignment.\\
+Consider Table 1 and 2 in Figure~\ref{fig:main_example} and the $SUM_{col}$ constraint template.
+Instead of considering all possible subblocks of T1 and T2, we reason over the properties of the superblocks first, $B2$ and $B4$ are not numeric so can be immediately discarded.
+Looking at the sizes, $B1$ contains only one column and $B5$ 2, however, the rows in $T2$ have length 4 so we only need to consider superblocks of size at least 4, which only $B3$ satisfies.
+
+In the second part, we start from an atom $c(B_1, \ldots ,B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
+As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier.
+For the above example, each of the vectors of $B6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g.
+at least size $4$) and test the definition for each of the rows.
+%In practice, this can be optimized further.
 \\\\
-We now describe in more detail how the headerless tables are extracted and how the superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3).
+We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3).
 We then describe a number of optimizations that we perform to speed up step 3 and 4 (Section 4.4).
@@
 Furthermore, headers may be missing and their use often requires incorporating language-specific vocabulary, e.g.
-in English.
+English.
 
@@
   \includegraphics[width=1\linewidth]{figures/tabletool.png}
-  \caption{This figure shows the table selection tool of \sname.
-A user selects rectangular ranges of cells that correspond to (headerless) tables.
-If the data has an unambiguous orientation, this information can be provided as well.}
+  \caption{This figure shows a simple table selection tool. A user selects rectangular ranges of cells that correspond to (headerless) tables, an explicit orientation can also be provided.}
   \label{fig:visual_tool}
@@
 
-Our algorithm requires tables to be specified.
-This information is provided to \sname in the form of a configuration file containing the name, coordinates within a spreadsheet and optionally the orientation of every table.
-The orientation can be used to indicate that a table should be interpreted as only rows or only columns.
-
-This information can be provided manually or generated automatically.
-We offer two proof of concept approaches that aim to facilitate the specification of tables.
-
-\subsubsection{Automatic detection}
-As mentioned above, automatic detection of headers and tables is a complex task.
-Therefore, we define a restricted setting in which the task becomes easy enough to be automated.
-We assume that 1) tables are rectangular blocks of data not containing any empty cells; and 2) tables are separated by at least one empty cell from other tables.
-Additionally, if the first row contains only textual values it treated as a header, otherwise, if the first column consists of only textual data is will be the header.
+Instead, we assume tables are specified by means of their coordinates within a spreadsheet and optionally a fixed orientation of each table.
+The orientation can be used to indicate that a table should be interpreted as having only rows or only columns.
+
+We developed two simple prototypes to help with the specification of the tables:
+%This information can be provided manually or generated automatically.
+
+\paragraph{Automatic detection}
+%As mentioned above, automatic detection of headers and tables is a complex task.
+%Therefore, we define a restricted setting in which the task becomes easy enough to be automated.
+Under the following two assumptions, the table detection task becomes easy enough to be automated; 1) tables are rectangular blocks of data not containing any empty cells; and 2) tables are separated by at least one empty cell from other tables.
+The sheet is then processed row by row, splitting each row into ranges of non-empty cells and appending them to adjacent ranges in the previous row.
+Headers can be detected, for example, by checking if the first row or column contains only textual values.
 If a header row (column) is detected, it is removed from the table and the orientation of the table is fixed to columns (rows), otherwise, we assume there is no header and the orientation is not fixed.
 
-The tool will now detect ranges of non-empty cells in every row and greedily merge these ranges across columns to obtain rectangular tables.
-If the first row of a table contains 
-This tool removes the need for human intervention, however, it is rather limited and can only be used in simple settings.
-
-\subsubsection{Visual}
-Since the specification of tables is usually easy for humans, we propose a second approach which aims to decrease the complexity of specifying tables by allowing users to indicate tables using a visual tool (Figure~\ref{fig:visual_tool}).
+\paragraph{Visual}
+The above assumptions do not hold for many tables.
+However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (Figure~\ref{fig:visual_tool}).
 Users select tables excluding the header and optionally specify an orientation.
-The tool then generates the table definitions automatically.
+The tool then generates the specification of the tables automatically.
 
@@
 %A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) blocks.
-The goal of the block detection step is to partition tables into type-consistent (all-numeric or all-textual) blocks.
-We generate blocks automatically, partitioning tables into maximal type-consistent blocks.
-First, the spreadsheet data is preprocessed so that currency values and percentual values are transformed into their corresponding numeric (i.e.
+The goal of the block detection step is to partition tables into maximally type-consistent (all-numeric or all-textual) blocks.
+%We generate blocks automatically, partitioning tables into maximal type-consistent blocks.
+First, we preprocess the spreadsheet data so that currency values and percentual values are transformed into their corresponding numeric (i.e.
 integer or floating point) representation (e.g.
 $2.75 \$$ as $2.75$ or $85\%$ as~$0.85$).
-Afterwards, tables are partitioned into maximal type-consistent blocks.
-If a table has no fixed orientation, the algorithm will attempt to construct both row-blocks and column-blocks.
-In order to find blocks, the type of each vector is determined, i.e. whether it is numeric or textual.
-All adjacent vectors that have the same type are merged to form a block. 
-If any row (column) of the table contains both numeric and textual values, it is type inconsistent and the algorithm will not produce any row-blocks (column-blocks) for that table.
-For tables that have a fixed orientation only blocks of that orientation are built.
-
-\newcommand{\temps}{\ensuremath{S}}
+Then, each table is partitioned into maximal type-consistent blocks.
+
+To find row-blocks, each row is treated as a vector and must be type-consistent; similarly for column-blocks.
+Then, adjacent vectors that are of the same type are merged to obtain the maximally type-consistent superblocks.
 
@@
 \label{sec:algo}
-Our method assumes that tables and blocks are given and solves the \tcl~problem by checking for each template: what combination of given (super)blocks may satisfy the signature and which specific (sub)blocks satisfy both the signature and the definition.
-
-The main challenge we aim to address with our approach is to avoid having to check each possible combination of blocks for each constraint template as well as developing an \textit{extensible} method in which new constraint templates can be plugged in with relative ease.
-%We will use constraint satisfaction technology to partly answer this challenge.
-\tias{No, separate selection of superblocks from generation of subblocks}
-\samuel{I do not fully agree, the property vs content aspect is still important, it is combined with superblock / subblock but this could be done differently}
-To answer this challenge we separate checking the signature (concerning properties of a block) of a constraint template from checking the definition (concerning the values in a block) of a constraint template.
-Furthermore, we use constraint satisfaction technology to jointly search for and test the signature over all combinations of superblocks.
-
+Our method assumes that constraint templates and maximal blocks are given and solves the \tcl~problem by checking for each template $s$: what combination of maximal blocks can satisfy the signature (\textit{superblock assignment}) and which specific \textit{subblock assignments} satisfy both the signature and the definition.
+The pseudo-code of this approach is shown in Algorithm~\ref{algo:tcl}.
+%For every constraint template~$s$, the algorithm first generates all combinations of superblocks (i.e. superblock assignments) that are consistent with the signature.
+%In a second step it looks at every superblock assignment and finds subassignments that satisfy the signature and definition of~$s$.
+
+This separation of checking the \textit{properties} of superblock assignments from checking the \textit{actual data} in the subblock assignment controls the exponential blow-up of combinations to test.
+Furthermore, we will use constraint satisfaction technology in the first step to efficiently enumerate superblock assignments that are compatible with the signature.
+%These are the two key properties of our approach.
+
+\newcommand{\temps}{\ensuremath{S}}
 \begin{algorithm}[t]
@@
     \footnotesize
-    \State \textbf{Input:} $\blocks$ -- blocks, $\temps$ -- constraint templates
+    \State \textbf{Input:} $\temps$ -- constraint templates, $\blocks$ -- maximal blocks
     \State \textbf{Output:} $C$ -- learned constraints
@@
       \ForAll{$s~\mathbf{in}~\temps$}
-        \State $n \gets$ number of arguments of template $s$
+        %\State $n \gets$ number of arguments of template $s$
         \State $A \gets \generategroups(s, \blocks)$
@@
         \ForAll{$(B_1',\dots,B_n') \in A'$}
-            \State $C \gets C \cup \{ s(B_1', \dots, B_n') \}$
+            \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
           \EndFor
@@
 
-The pseudo-code of the main method is shown in Algorithm~\ref{algo:tcl}.
-For every constraint template~$s$, the algorithm first generates all combinations of superblocks (i.e. superblock assignments) that are consistent with the signature.
-In a second step it looks at every superblock assignment and finds subassignments that satisfy the signature and definition of~$s$.
-
 \subsubsection{Generating superblock assignments}
-Given a constraint template~$s$ and the set of all blocks~\blocks, the goal is to find all combinations of blocks that satisfy the constraint signature.
-Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.
-
-However, the properties defined in the signature, such as those related to the type or size of a block, can be used to avoid considering certain groups for certain arguments.
-Furthermore, the choice of one argument can influence the possible candidates for the other arguments, for example, if they must have the same length.
-Instead of writing specialized code for each of these cases, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.
+% Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.
+
+Given a constraint template~$s$ and the set of all maximal blocks~\blocks, the goal is to find all combinations of maximal blocks that are consistent with the constraint signature.
+An argument assignment $(B_1, ...
+B_n)$ is \textit{consistent} with the signature of template $s$ if for each block $B_i$ there exists at least one subblock $B_i' \sqsubseteq B_i$ that satisfies the signature.
+
+Furthermore, the choice of one argument can influence the possible candidates blocks for the other arguments, for example, if they must have the same length.
+Instead of writing specialized code to generate and test the superblock assignments, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.
 
@@
 In our case, we define one variable $V_i$ for each argument of a constraint template.
-Each variable can be assigned to each of the blocks in \blocks, so the domain consists of $|\{\blocks\}|$ identifiers.
+Each variable can be assigned each of the maximal blocks in \blocks, so the domain of the variables consists of $|\{\blocks\}|$ block identifiers.
 
@@
 
-To reason over the blocks, we automatically add to the constraint program a set of background facts that contain the properties of each block, namely its type, size, number of columns and rows, length, orientation, and table.
-The actual constraints correspond to the requirements defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
-
+To reason over the blocks, we add to the constraint program a set of background facts that contain the properties of each block, namely its type, table, orientation, size, length and number of rows and columns.
+%The actual constraints correspond to the requirements defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
+%
 % \tias{More detailed description of the properties commented out, feel free to put in}
@@
 %\end{description}
-
-\begin{table}[htb]
+%
+\begin{table}[t]
   \caption{Translation of signature requirements to superblock constraints.
@@
 \end{table}
-
-In order for our approach to be complete, the signature constraints can not always be enforced directly on a block.
-If a block matches the signature it is a valid candidate, however, if it does not there might exist subblocks that still satisfy the signature.
-\begin{example}
-  Consider, for example, the constraint template \ecseries{x} and a superblock candidate~$B$ for~$x$ containing both integer and floating point values.
-  While the signature of the template requires~$x$ to be integer we cannot discard~$B$ since there might be subblocks~$B' \sqsubseteq B$ that only contain integer values and satisfy the signature.
-\end{example}
-Table~\ref{tbl:translation} shows the how constraints from the signature are converted to constraints on superblocks.
+%
+The actual constraints must enforce that a superblock assignment is \textit{consistent} with the requirements defined in the signature.
+%
+%In order for our approach to be complete, the signature constraints can not always be enforced directly on a block.
+%If a block matches the signature it is a valid candidate, however, if it does not there might exist subblocks that still satisfy the signature.
+%\begin{example}
+%  Consider, for example, the constraint template \ecseries{x} and a superblock candidate~$B$ for~$x$ containing both integer and floating point values.
+%  While the signature of the template requires~$x$ to be integer we cannot discard~$B$ since there might be subblocks~$B' \sqsubseteq B$ that only contain integer values and satisfy the signature.
+%\end{example}
+
+Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints happens:
 Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
-Typing requirements need to be relaxed to check only for supertypes (numeric and textual).
+Typing requirements need to be relaxed to check only for base types (numeric and textual).
 Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since subgroups might contains fewer vectors.
@@
 The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all satisfying solutions.
-These solutions will correspond with the valid superblock assignments for constraint template~$s$.
+These solutions correspond to the valid superblock assignments for constraint template~$s$.
 
 \begin{example}
-Br and Bx are numeric; rows(Bx) ⇤ 2; and columns(Bx ) = length(Br )
-
-  Consider the constraint template $\ecsumc{r}{x}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$r$ and~$x$.
-  Given blocks~\blocks, the domain of these variables are $D(V_r) = D(V_x) = \{ 1, \dots, |\blocks| \}$.
-  Finally, the signature will be translated into constraints:
+%Br and Bx are numeric; rows(Bx) ⇤ 2; and columns(Bx ) = length(Br )
+
+  Consider the constraint template $\ecsumc{\sg_r}{\mathbf{\sg_x}}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$\sg_r$ and~$\mathbf{\sg_x}$.
+  Given maximal blocks~\blocks, the domain of these variables are $D(V_r) = D(V_x) = \{ 1, \dots, |\blocks| \}$.
+  Finally, the signature of $\ecsumc{\sg_r}{\mathbf{\sg_x}}$ (see Table~\ref{table:constraints}) will be translated into constraints:
   \begin{align*}
@@
   \end{align*}
-  Every solution is a tuple of values for~$V_r$ and~$V_x$ corresponding directly with a superblock assignment over~\blocks.
+  %Every solution is a tuple of values for~$V_r$ and~$V_x$ corresponding directly with a superblock assignment over~\blocks.
 \end{example}

### 2016-05-26 Paper/ECAI/tabular.tex
algorithm, subblocks

@@
 
-\subsubsection{Constraints over satisfying subgroups}
+\subsubsection{Generating subblock assignments}
 \label{sec:algo:subgr}
-In the previous step superblock assignments are generated for a constraint template $s$ with $n$ arguments.
-In this step, given such a superblock assignment $(B_1, \dots, B_n)$, the goal is to discover valid \textit{subassignments}, i.e.
+Given a superblock assignment $(B_1, \dots, B_n)$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e.
 assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.
-\samuel{Is this repetition?}
-While the previous step reasons about the \textit{properties} of blocks (type, size, length, ...), this step reasons about the actual \textit{content} of the vectors in the blocks, that is, the data.
-
+
+%While the previous step reasoned about the \textit{properties} of blocks, in this step we will also reason about the actual \textit{content} of the blocks, that is, the data.
 \begin{example}
@@
 
-The search for subassignments is represented as the generic procedure $\findassignment$.
-However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
-One possible method is to reformulate the problem as a CSP, similar to the approach we employed in the last section.
-Often, however, a simple generate-and-test approach will suffice.
-In contrast to the previous step, the number of subblocks to generate here are typically manageable: $O(n*m^2)$ with $m$ the size of the largest block and $n$ the number of arguments of the template.
-We only consider continuous ranges as subblocks, so for a block of size $m$ there are $m*(m+1)/2$ possible subblocks to consider.
-Furthermore, few constraint solvers support floating point numbers, which are prevalent in spreadsheets.
-
-For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
-Given a superblock assignment, all subassignments are generated taking into account the required size, columns or rows.
-The subassignments are sequentially tested using a template-specific procedure \textit{Test} and valid subassignments are returned.
-The \textit{Test} procedure verifies if the definition is satisfied and, if necessary, checks signature constraints that have not yet been enforced (e.g. subtype constraints).
+In this step, we could also formulate the problem as a CSP.
+%However, the search space is much smaller, e.g.
+$O(nm^2)$ with $n$ the number of arguments of the template and $m$ the size of the largest block.
+However, few CSP solvers support floating point numbers, which are prevalent in spreadsheets.
+Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
+This is inefficient, as the signature of the definition may already not be satisfied, or the first element in the data.
+
+Hence, a simple generate-and-test will usually suffice.
+%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
+%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
+Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
+The disjoint subassignments are then sequentially tested using a template-specific procedure \textit{Test} and valid subassignments are returned.
+The \textit{Test} procedure verifies the exact signature (e.g. subtypes will not have been checked yet) and checks whether the definition is satisfied for all data in the blocks.
 
@@
     \State $A_{sub} \gets \emptyset$
-    \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i \}$}
-      \If{$disjoint(\{ B_1', \dots, B_n' \}) \land Test((B_1',\dots,B_n'))$}
+    \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i  \wedge \nexists i,j,i\neq j: B'_i \sqsubseteq B'_j \}$}
+      \If{$Test((B_1',\dots,B_n'))$}
         \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
@@
 \subsection{Optimizations} \label{sec:opts}
-\samuel{Should this not be part of approach / optimizations / redundancy?}One way to deal with such equivalences is to identify a \textit{canonical} constraint among the equivalent ones.
-For example, for the product constraint above, we could define a lexicographic ordering over the blocks and only consider the constraint with the smallest ordering.
-
-This section discusses various essential design decisions and optimizations aiming to improve the speed and extensibility of our system as well as to find less redundant constraints.
+This section discusses various design decisions and optimizations aimed at improving the speed and extensibility of our system as well as to eliminate some obvious redundant constraints.
 

### 2016-05-26 Paper/ECAI/tabular.tex
some more tweaks to optimizations part

@@
 
-\subsubsection{Constraint solvers}
-\samuel{Tias, Sergey, have a look at his section, I'm torn about rewriting it, feeling it should either be reduced or expanded, especially the second part.}
-Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
-Constraint solvers can be categorized as being stand-alone solvers, meaning that they have a text-based modeling language in which the CSP must be formulated, and \textit{embedded}, meaning that they have a programmatic API to formulate the CSP.
-
-\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the MiniZinc language \cite{minizinc} with the stand-alone Gecode 4.4.0 solver (which supports floats) and uses an embedded Python-based CP solver \cite{python_constraint}.
-The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints.
+%\subsubsection{Constraint solvers}
+%\samuel{Tias, Sergey, have a look at his section, I'm torn about rewriting it, feeling it should either be reduced or expanded, especially the second part.}
+%Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
+%Constraint solvers can be categorized as being stand-alone solvers, meaning that they have a text-based modeling language in which the CSP must be formulated, and \textit{embedded}, meaning that they have a programmatic API to formulate the CSP.
+%
+%\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the MiniZinc language \cite{minizinc} with the stand-alone Gecode 4.4.0 solver (which supports floats) and uses an embedded Python-based CP solver \cite{python_constraint}.
+%The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints.
 The embedded solver is fastest as it does not have the overhead of file-writing and calling an external program, and is used in the experiments below.
@@
 In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
-Moreover, \sname currently uses product (\ecprod{B_R'}{B_1'}{B_2'}) instead of division (\eccalc{B_1'}{B_R' / B_2'}) and difference (\ecdiff{B_r}{B_1}{B_2}) instead of sum (\eccalc{B_1}{B_r + B_2}).
-These choices are hardcoded preferences based on the prevalence of difference and product in financial spreadsheet data, however, these constraints can be trivially rewritten if necessary.
-
-The second type of redundancy we aim to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
+Moreover, \sname currently contains only product (\ecprod{B_R'}{B_1'}{B_2'}) and not division (\eccalc{B_1'}{B_R' / B_2'}) as well as difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
+These choices are based on the prevalence of product and difference in financial spreadsheet data, however, division and sum could easily be added to, as can other constraints.
+
+A last type of redundancy we aim to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
 For example, consider the constraint \ecsumc{B_r}{B[1:n]} where $B[k:n]$ consists of only zeros, then \ecsumc{B_r}{B[1:j]} will be true for all $k - 1 \leq j \leq n$.
@@
 
-In some cases a maximal subblock might falsely include irrelevant columns.
-By manually specifying more fine-grained blocks as input this behavior can often be avoided.
+As a result, in some cases a maximal subblock might falsely include irrelevant columns.
 Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$.
-If the target constraint is $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, the last vector of~$B_1$ can be split off into a separate block.
-Alternatively, the algorithm can output all valid constraints, e.g.
+Should, the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
+
+Alternatively, we can output all valid argument assignments instead of only maximal subblocks, e.g.
 $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$, however, this might produce many redundant constraints and introduce constraints that accidentally cover too little data.

### 2016-05-26 Paper/ECAI/tabular.tex
Related work comments

--- 2016-05-26 Paper/ECAI/tabular.tex
@@
 For instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
-Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and also that ModelSeeker’s constraints are the typical ones from the constraint programming community, which are aimed at combinatorial optimization and scheduling rather than what is typically used in spreadsheets.
+Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and also that ModelSeeker’s constraints are the typical ones from the constraint programming community\tias{which is all over integers, we also over strings and real-valued, AND, not one vector but we have many indep.
+blocks}, which are aimed at combinatorial optimization and scheduling rather than what is typically used in spreadsheets.
 Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints using techniques inspired on Mitchell’s version spaces to process examples, some variants also asking informative questions to the user.
-As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
+As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints\tias{it looks only at pairwise constraints, so an important factor is how fast the algorithm can converget to the entire constraint set.
+Redundant and spur also issue for these methods}.
 As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.
@@
 At that point FlashFill would generate a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', would also apply it to automatically generate values for the other politicians (rows) in the table.
-While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
+While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'' \tias{only strings?}, FlashFill learns correct programs from very few examples.
 Flashfill has been extended to e.g.
-FlashExtract \cite{flashextract} to produce a set of normalized database tables starting from naively filled out spreadsheets.
+FlashExtract \cite{flashextract} to produce a set of normalized database tables \tias{can be more precise, does it add formulae?
+or this is real DB database, e.g a schema and PK/FK constraints?} starting from naively filled out spreadsheets.
 However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.

### 2016-05-26 Paper/ECAI/tabular.tex
Started reworking evaluation

@@
 %
-      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
+%      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
 %
@@
   \vspace{-10pt}
-  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16) and \textit{FOREIGNKEY} (5).
+  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
   $^1$(Spurious constraints) $^2$(Redundant constraints not originally present)}
@@
 In this section we experimentally validate our approach.
-We studied various questions, most notably with what accuracy our algorithm can find essential constraints.
+We studied various questions, most notably with what accuracy our algorithm can find constraints.
 
@@
 
+There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
+
 In this section we focus on \textit{functional} constraints that could be used in spreadsheets, ignoring constraints such as all-different or foreign-key.
-
 All experiments were run on a Macbook Pro, Intel Core i7 2.3 GHz with 16GB RAM.
@@
 Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
-This example combines several smaller examples that were used in an exercise session to teach Excel into one spreadsheet.
-Figure~\ref{fig:sol_example} shows the constraints that we expect to find.
+This example combines several smaller exercise session examples into one spreadsheet.
 
 \subsubsection{Results}
-\sname takes a few seconds to find 18 constraints, including all of the 12 solutions described in Figure~\ref{fig:sol_example}.
-Of the 6 remaining (redundant) constraints, 5 are $\mathit{rank}$ constraints that are true by accident, such as: \begin{align*}
-  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}
-\end{align*}
-The other redundant constraint is an additional $\mathit{lookup}$ that holds because the two vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).
-
-For this example our primary goal of finding all constraints is achieved.
-The implementation also returns a number of redundant constraints ($33.33\%$ of the total).
+\sname takes a few seconds to find the 18 constraints in Figure~\ref{fig:sol_example}.
+These include 5 spurious $\mathit{RANK}$ constraints, such as $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
+Moreover, there is 1 \textit{LOOKUP} constraint that was is present in the original spreadsheet (looking up ID based on Salesperson).
+
+For this example our primary goal of finding all original constraints is achieved.
+The implementation also returns 6 additional constraints ($33.33\%$ of the total).
 This ratio is, as we will show, rather high compared to other spreadsheets.

### 2016-05-27 Paper/ECAI/tabular.tex
Changed dependecy graph + caption and WIP evaluation

@@
 \usepackage{tabularx}
-\usepackage{subcaption}
+\usepackage[margin=0.1cm]{subcaption}
 \usepackage[inline]{enumitem}
@@
   \vspace{-10pt}
-  \caption{Example spreadsheet (black words and numbers only). Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.}
+  \caption{
+    Example spreadsheet (black words and numbers only).
+    Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.
+    This example was created by combining several smaller Excel sheets originating from an exercise session.
+  }
   \label{fig:main_example}
@@
   \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
-  $^1$(Spurious constraints) $^2$(Redundant constraints not originally present)}
+  $^1$(Spurious constraints) $^2$(Not present in original source)}
   \label{fig:sol_example}
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency.png}
-  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_2$ (the \CSignature of~$s_2$ includes $s_1$).
-    The graph is extracted from the Signature column in Table~\ref{table:constraints}.\samuel{Correct aggregates}
+  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency2.pdf}
+  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (the \CSignature of~$s_2$ includes $s_1$).
+    The graph is extracted from the Signature column in Table~\ref{table:constraints}.
   }
@@
 
-In this section we focus on \textit{functional} constraints that could be used in spreadsheets, ignoring constraints such as all-different or foreign-key.
-All experiments were run on a Macbook Pro, Intel Core i7 2.3 GHz with 16GB RAM.
+In this section we focus on \textit{functional} constraints that could be used in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
+All experiments were run using Python 3.5.1 (custom Anaconda environment) on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
 
@@
 Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
-This example combines several smaller exercise session examples into one spreadsheet.
 
 \subsubsection{Results}
-\sname takes a few seconds to find the 18 constraints in Figure~\ref{fig:sol_example}.
+\sname takes a few seconds to find the constraints in Figure~\ref{fig:sol_example} (excluding structural constraints).
 These include 5 spurious $\mathit{RANK}$ constraints, such as $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
-Moreover, there is 1 \textit{LOOKUP} constraint that was is present in the original spreadsheet (looking up ID based on Salesperson).
-
-For this example our primary goal of finding all original constraints is achieved.
-The implementation also returns 6 additional constraints ($33.33\%$ of the total).
-This ratio is, as we will show, rather high compared to other spreadsheets.
-However, this example contains many short vectors which increases the chance for constraints to be true by accident.
+Moreover, there is 1 \textit{LOOKUP} constraint that was not present in the original spreadsheet (looking up ID based on Salesperson).
+
+For this example our primary goal of finding all original constraints is achieved, the recall is $1.0$.
+The implementation also returns 6 additional constraints, therefore, our precision is $12/18 = 0.67$, which is rather low compared to other spreadsheets in the benchmark.
+The low precision can be explained by the many short vectors which increases the chance for constraints to be true by accident.
 
@@
 
-All spreadsheets have been converted to CSV with no manual intervention unless noted.
+All spreadsheets have been converted to CSV files.
 Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).
 
+\samuel{Revisit}
 For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
@@
 
-We examine three main questions concerning the accuracy, redundancy and efficiency of our approach, our main focus being accuracy.
+We examine three main questions concerning the accuracy (recall), redundancy (precision) and efficiency of our approach, our main focus being accuracy.
 
 \subsubsection*{Q1. How accurate is the approach?}
-Fortunately, our system is currently able to find $100\%$ of the implemented constraints (90) on the benchmark suite.
-There are only three essential constraints that have not been implemented yet, therefore our system currently identifies $96.77\%$ of the essential constraints.
+Fortunately, for the implemented constraint templates, our system achieves a recall of $1.0$, i.e. it is able to find all of the supported constraints (90) on the benchmark suite.
+There are only three templates that have not been implemented yet, so w.r.t. the ground truth the total recall of our system is $0.97$.
 
@@
 
-Redundant constraints can be reduced in different ways.
+Redundant constraints can be reduced in different ways.samuel
 Duplications should be detected in a post-processing step that uses entailment or heuristics to filter constraints.

### 2016-05-27 Paper/ECAI/tabular.tex
Started incorporating comments from Angelika

@@
 In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
-We see multiple promising applications for this technique, e.g.
-in rediscovering constraints, auto-completion and error checking.
+We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
 Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
@@
 The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
-However, large and complex sheets (possibly with multiple tables and relations between different them) can be hard to handle.
+However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
 Many end-users lack the understanding of the underlying structures and dependencies in such sheets and the data they contain.
-Especially when spreadsheets have been exported from
-other software such as Enterprise Resource Planning (ERP) systems; in this case, %, it is hard to understand the data as in this case,
-often a comma-separated values (CSV) format is used meaning that all formula's are lost, including inter-sheet formulas and relations.
+This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
+In this case, often a comma-separated values (CSV) format is used meaning that all formulas are lost, including inter-sheet formulas and relations.
 %This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
-Even in manually created spreadsheets, it can be challenging to be consistent and correct with formula's across big spreadsheets.
+Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big spreadsheets.
 For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain some mistakes in formulae.
 
-In this paper,  we investigate whether learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
+In this paper,  we investigate whether learning techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
 This is a new and unconventional machine learning problem.
@@
 
-%the field of Inductive Logic Programming has long...
-\tias{TODO for Sergey or Luc}.
-Additionally, the field of Constraint Programming has recently started investigating ways in which Constraint Satisfaction Problems can automatically be inferred from known solutions.
-Here, methods ranging from generate-and-test~\cite{todo:modelseeker}, concept learning~\cite{todo:conacq} and Inductive Logic Programming~\cite{todo:lallouet} have been used with encouraging results.
-
-%Another line of work from which we draw inspiration is that of \tias{TODO for ?, flashfill et al}.
-We will consider the case where all data is given at once.
-However, as in the use case of flashfill, one could also use this technique to derive formula's interactively while a user is entering values in a cell \tias{TODO SAMUEL: proofread/refine last statement}.
-
-The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
+\samuel{Move this up a paragraph}
+The question that we answer in this paper is: is it possible to discover or reconstruct constraints (relations, functions) in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
 %Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
@@
 In line with this definition, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
-We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independent of their orientation.
+We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independently of their orientation.
 
@@
 
-\item the \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
+\item \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
 This can concern any property of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
-In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
+In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
 
@@
 It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
+\samuel{Is terms the right word? Shouldn't it be arguments, too?}
 From a FOL perspective, the name of the constraint (e.g.
@@
 
-\paragraph{Redundancies}
+\subsubsection{Redundancies}
+\label{sec:form:redundancies}
 Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}.
@@
 
-%
 %More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever

### 2016-05-27 Paper/ECAI/tabular.tex
improved the references

--- 2016-05-27 Paper/ECAI/tabular.tex
@@
 approaches, it focuses on data in spreadsheets and on constraints that can be column or row-wise.
-Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,Llalouet}.
+Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
 Still several ideas from logical and relational learning

### 2016-05-27 Paper/ECAI/tabular.tex
Further comments Angelika

@@
 \begin{example}
-Consider the (headerless) table~$T_1$ in Figure~\ref{fig:main_example}, its rows are not type consistent (i.e. they contain both numeric and textual data).
+Consider the (headerless) table~$T_1$ in Figure~\ref{fig:main_example}.
+Its rows are not type consistent (i.e. they contain both numeric and textual data).
 However, the table can be partitioned into five column-oriented blocks $B_1, B_2, B_3, B_4, B_5$, as shown in the figure ($B_1 = \range{T_1}{\rangeall}{1}$, $B_2 = \range{T_1}{\rangeall}{2}$, $B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
@@
 % and what combination of superblocks may lead to a valid argument assignment.\\
-Consider Table 1 and 2 in Figure~\ref{fig:main_example} and the $SUM_{col}$ constraint template.
-Instead of considering all possible subblocks of T1 and T2, we reason over the properties of the superblocks first, $B2$ and $B4$ are not numeric so can be immediately discarded.
-Looking at the sizes, $B1$ contains only one column and $B5$ 2, however, the rows in $T2$ have length 4 so we only need to consider superblocks of size at least 4, which only $B3$ satisfies.
+Consider Table 1 and 2 in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{col}$ constraint template.
+Instead of considering all possible subblocks of T1 and T2, we first reason over the properties of the superblocks.
+For example, $B2$ and $B4$ are not numeric so they can be immediately discarded.
+\samuel{Iff Br is from T2?}
+Looking at the sizes, $B1$ and~$B2$ contain only one and two columns respectively, however, the rows in $T2$ have length 4 so we only need to consider superblocks with at least 4 columns, which only $B3$ satisfies.
 
@@
 \\\\
+\samuel{Section references not correct}
 We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3).
@@
 Typing requirements need to be relaxed to check only for base types (numeric and textual).
-Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since subgroups might contains fewer vectors.
-Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the group they are applied to.
-Subgroups of row-oriented (column-oriented) group will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
-
-The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all satisfying solutions.
+Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with two many vectors contain subblocks of the required size.
+Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
+Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
+
+The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all solutions.
 These solutions correspond to the valid superblock assignments for constraint template~$s$.
@@
 Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
+\samuel{Tias, there is something wrong here:}
 This is inefficient, as the signature of the definition may already not be satisfied, or the first element in the data.
@@
 
+\samuel{Angelika: trim this section for repeated information}
 Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
@@
 
-The dependency graph represents a partial ordering and is used produce a linear ordering of the given templates such that more general templates occur before templates that depend on them (e.g.
+The dependency graph represents a partial ordering and is used to produce a linear ordering of the given templates such that more general templates occur before templates that depend on them (e.g.
 \textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
@@
 Moreover, \sname currently contains only product (\ecprod{B_R'}{B_1'}{B_2'}) and not division (\eccalc{B_1'}{B_R' / B_2'}) as well as difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
-These choices are based on the prevalence of product and difference in financial spreadsheet data, however, division and sum could easily be added to, as can other constraints.
+These choices are based on the prevalence of product and difference in financial spreadsheet data, however, division and sum as well as other constraints could easily be added.
 
@@
 The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
-Their inclusion was based on the analysis of the most popular Excel constraints as well as the spreadsheets we encountered.
+Their inclusion was based on the analysis of the most popular Excel constraints \samuel{How do we know?} as well as the spreadsheets we encountered.
 All these constraints are available in popular spreadsheet software, except for the structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}).
-These constraints added so they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
+These constraints were added so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
 
@@
 In this section we experimentally validate our approach.
-We studied various questions, most notably with what accuracy our algorithm can find constraints.
+We studied various questions \samuel{Mention questions here?}, most notably with what accuracy our algorithm can find constraints.
 
@@
 
-There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
+There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel \samuel{Including running example?}, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
 
@@
 \subsection{Benchmark}
-There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
-The case study is also included.
-
 All spreadsheets have been converted to CSV files.
-Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).
-
-\samuel{Revisit}
+Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases blocks where also provided manually).
+
+\samuel{Revisit [what does "implemented" mean here? what's the difference between "implemented" and "algorithm", esp given that we are talking experiments? ]}
 For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
@@
 \begin{description}
-  \item[Implemented] Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
-  \item[Essential] Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
+  \item[Implemented] \samuel{Dupl} Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
+  \item[Essential] \samuel{Dupl} Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
   \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)
@@
 \subsubsection*{Q1. How accurate is the approach?}
+\samuel{Sounds like we were surprised}
 Fortunately, for the implemented constraint templates, our system achieves a recall of $1.0$, i.e. it is able to find all of the supported constraints (90) on the benchmark suite.

### 2016-05-27 Paper/ECAI/tabular.tex
Copied remaining comments Angelika

@@
 This is a new and unconventional machine learning problem.
+\samuel{A: be more specific about which headers, there's a lot to see in that figure - 
+use T1 / T2 to avoid confusion with regular table references
+ - 
+why is T2 oriented differently from the others?
+ - 
+what is "the corresponding column"?
+ - 
+I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now, also in terms of formulations ("the row-wise sum of the previous columns", hu?
+do you mean that in each row, total is the sum of the values for quarters?)}
 Consider the example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of spreadsheet operations such as \textit{average} or \textit{sum}.
@@
 %, since people tend to use spreadsheets for financial and accounting computations.
+\samuel{A: at this point I have an idea what the abstract means by "function" constraints, but wonder what the "relation" ones are}
 One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
@@
 In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
-
+\\\\
+\samuel{A: how do you get from the signature and the definition to these predicates? I guess this is what I'm missing on the formal side above… this needs to be made explicit}
 \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
-Given an assignments of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks.
+Given an assignment of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks.
 %holds, in practice this will be a function that can be called when
@@
 With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
+\samuel{A: how do you get from the signature and the definition to these predicates? I guess this is what I'm missing on the formal side above… this needs to be made explicit}
 We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
@@
 %
+\samuel{A: so you expect the user to define blocks? can this be automated?
+(also, later on, your solution seems to start from the tables rather than the blocks, why not match the two?)}
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\

### 2016-05-27 Paper/ECAI/tabular.tex
flashextract reworked

--- 2016-05-27 Paper/ECAI/tabular.tex
@@
 In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
-We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
+We see multiple promising applications for this technique, e.g.
+in rediscovering constraints, auto-completion and error checking.
 Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
@@
 The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
-However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
+However, large and complex sheets (possibly with multiple tables and relations between different them) can be hard to handle.
 Many end-users lack the understanding of the underlying structures and dependencies in such sheets and the data they contain.
-This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
-In this case, often a comma-separated values (CSV) format is used meaning that all formulas are lost, including inter-sheet formulas and relations.
+Especially when spreadsheets have been exported from
+other software such as Enterprise Resource Planning (ERP) systems; in this case, %, it is hard to understand the data as in this case,
+often a comma-separated values (CSV) format is used meaning that all formula's are lost, including inter-sheet formulas and relations.
 %This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
-Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big spreadsheets.
+Even in manually created spreadsheets, it can be challenging to be consistent and correct with formula's across big spreadsheets.
 For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain some mistakes in formulae.
 
-In this paper,  we investigate whether learning techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
+In this paper,  we investigate whether learning techniques can be used to infer constraints (formula's and other relations) from raw spreadsheet data.
 This is a new and unconventional machine learning problem.
@@
 
-\samuel{Move this up a paragraph}
-The question that we answer in this paper is: is it possible to discover or reconstruct constraints (relations, functions) in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
+%the field of Inductive Logic Programming has long...
+\tias{TODO for Sergey or Luc}.
+Additionally, the field of Constraint Programming has recently started investigating ways in which Constraint Satisfaction Problems can automatically be inferred from known solutions.
+Here, methods ranging from generate-and-test~\cite{todo:modelseeker}, concept learning~\cite{todo:conacq} and Inductive Logic Programming~\cite{todo:lallouet} have been used with encouraging results.
+
+%Another line of work from which we draw inspiration is that of \tias{TODO for ?, flashfill et al}.
+We will consider the case where all data is given at once.
+However, as in the use case of flashfill, one could also use this technique to derive formula's interactively while a user is entering values in a cell \tias{TODO SAMUEL: proofread/refine last statement}.
+
+The question that we answer in this paper is: is it possible to discover or reconstruct structural constraints (relations, functions) in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose method and system, named \sname (pronounced as the word ``tackle''), for discovering row-wise and column-wise constraints.
 %Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
@@
 In line with this definition, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
-We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independently of their orientation.
+We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independent of their orientation.
 
@@
 
-\item \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
+\item the \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
 This can concern any property of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
-In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
+In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the constraint its arguments are well-formed.
 
@@
 It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
-\samuel{Is terms the right word? Shouldn't it be arguments, too?}
 From a FOL perspective, the name of the constraint (e.g.
@@
 
-\subsubsection{Redundancies}
-\label{sec:form:redundancies}
+\paragraph{Redundancies}
 Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}.
@@
 
+%
 %More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
@@
 Therefore, conflicts should only be detected if the snapshot data is large enough.
+An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.
 
@@
 Flashfill has been extended to e.g.
-FlashExtract \cite{flashextract} to produce a set of normalized database tables \tias{can be more precise, does it add formulae?
-or this is real DB database, e.g a schema and PK/FK constraints?} starting from naively filled out spreadsheets.
+FlashExtract \cite{flashextract} to produce a set of tables starting from a document (a txt, spreadsheet or HTML file) with a set of examples marked for extraction.
 However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.

### 2016-05-27 Paper/ECAI/tabular.tex
Transferred all comments

--- 2016-05-27 Paper/ECAI/tabular.tex
@@
 
-%%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
-                  % Do NOT use for camera-ready version of paper.
-
 \begin{document}
 
-\title{Learning constraints in tabular data}
+\title{Learning constraints in spreadsheets and tabular data}
 % see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
@@
 Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
-However, modeling, maintaining and discovering formulas in tabular data and between spreadsheets can be time consuming and error-prone.
+However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
 In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
@@
 
-The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where
-the discovery of clausal constraints has been studied in an inductive logic programming setting \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
+The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
 and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
@@
 very few examples.  Our approach contrasts with these in that it focusses
-on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional CSPs.
+on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional constraint satisfaction problems (CSPs).
 It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
@@
 Our goal is to automatically discover constraints (functions and relations) between the rows and the columns of tables in a spreadsheet.
-This is applicable not just to data from spreadsheets, but any data in tabular form, hence the name.
+This is applicable not just to data from spreadsheets, but to any data in tabular form, hence the name.
 
@@
 
-We first introduce some terminology and the concept of \template, after which we define the problem and make some additional considerations.
+We first introduce some terminology and the concept of a \template, after which we define the problem and make some additional considerations.
 
@@
 \item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
+\\\\
+\luc{these are the vectors along \dots along}
 \item \textit{rows}: this is orientation specific, for row-based blocks this is the size, for column-based blocks the length
@@
   \end{tabularx}}
+
+  \luc{there are some fancy constraints here.
+it might be good to say something about them ...
+like an example of one of the more complex ones in the text ...
+maybe state what conditional means, what aggregate ...
+}
 \end{table*}
@@
 %
-A \template $s$ is a triple $s=\textit{(\CName, \CSignature, \CFunction)}$:
+A \template $s$ is a triple $t = \textit{(\CName, \CSignature, \CFunction)}$:
 %Let us elaborate on this:
@@
 
-  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
+  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$. \luc{example where it holds}
 \end{itemize}
@@
 We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
-Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would write a Prolog-like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
+Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming and logical relational learning, we would define the predicate~$s$ using a Prolog like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
 Under this interpretation, the term constraint and constraint template can be used interchangeably.
@@
 \begin{definition}
-A \textbf{valid argument assignment} of a constraint $s$ is a tuple $(B_1, ..., B_n)$ such that $s(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
+A \textbf{valid argument assignment} of a constraint $s$ is a tuple of blocks $(B_1, ..., B_n)$ such that $s(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
 \end{definition}
@@
 
-We define the {\bf Tabular Constraint Learning Problem} as follows:
+We define the inverse problem, that is the {\bf Tabular Constraint Learning Problem}, as follows:
 %
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$ and $(B'_1, ..., B'_n)$ is a valid argument assignment of the template~$s$.
+{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$, $\forall i: B_i' \sqsubseteq B_i \in {\cal B}$ and $(B'_1, ..., B'_n)$ is a satisfied argument assignment of the template~$s$.
 \end{definition}
@@
 This is because constraints may be logically equivalent or may be implied by other constraints.
+\luc{if you have permutation, alldiff is redundant.}
 
@@
 
-Because dealing with redundancy is often application-dependent, we explain in the next section our generic method for finding all constraints.
-We describe some optimisations that avoid obvious equivalences in Section~\ref{sec:opts} and will investigate the impact of redundancy in the experiments.
+Dealing with redundancy is often application-dependent.
+Therefore, in Section~\ref{sec:approach} we will first explain our generic method for finding all constraints, before describing some optimizations that avoid obvious equivalences (Section~\ref{sec:opts}).
+Later, we will investigate the impact of redundancy in the experiments.
 
@@
 \begin{enumerate}
-\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal \textit{superblock} and the $B_i$ are compatible with the signature.
-\item For each generated atom $c(B_1, \ldots ,B_n)$, find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
+\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal superblock and the $B_i$ are \textit{compatible} (defined below) with the signature. 
+\item For each generated atom $c(B_1, \ldots ,B_n)$ in 3a), find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
 \end{enumerate}
 \end{enumerate}
+\luc{introduce the notion of a maximal superblock earlier ???}
 
@@
 
-Instead, we divide this step into two parts: in the first part, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks.
+Instead, we divide this step into two parts: in step 3a, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks.
 % and what combination of superblocks may lead to a valid argument assignment.\\
-Consider Table 1 and 2 in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{col}$ constraint template.
-Instead of considering all possible subblocks of T1 and T2, we first reason over the properties of the superblocks.
-For example, $B2$ and $B4$ are not numeric so they can be immediately discarded.
+Consider tables $T_1$ and $T_2$ in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{col}$ constraint template.
+Instead of considering all possible subblocks of $T_1$ and $T_2$, we first reason over the properties of the superblocks.
+For example, $B_2$ and $B_4$ are not numeric so they can be immediately discarded.
 \samuel{Iff Br is from T2?}
-Looking at the sizes, $B1$ and~$B2$ contain only one and two columns respectively, however, the rows in $T2$ have length 4 so we only need to consider superblocks with at least 4 columns, which only $B3$ satisfies.
-
-In the second part, we start from an atom $c(B_1, \ldots ,B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
+Looking at the sizes, $B_1$ and~$B_2$ contain only one and two columns respectively, however, the rows in $T_2$ have length 4 so we only need to consider superblocks with at least 4 columns, which only $B_3$ satisfies.
+\luc{so the generated atom in step 3a is \dots }
+
+In step 3b, we start from an atom $c(B_1, \ldots, B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
 As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier.
-For the above example, each of the vectors of $B6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g.
-at least size $4$) and test the definition for each of the rows.
+For the above example, each of the vectors of $B_6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g.
+at least size $4$) and test whether for each row the definition is satisfied.
 %In practice, this can be optimized further.
@@
 We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3).
-We then describe a number of optimizations that we perform to speed up step 3 and 4 (Section 4.4).
+We then describe a number of optimizations that we perform to speed up steps 3 and 4 (Section 4.4).
 
@@
 This separation of checking the \textit{properties} of superblock assignments from checking the \textit{actual data} in the subblock assignment controls the exponential blow-up of combinations to test.
-Furthermore, we will use constraint satisfaction technology in the first step to efficiently enumerate superblock assignments that are compatible with the signature.
+Furthermore, we use constraint satisfaction technology in the first step to efficiently enumerate superblock assignments that are compatible with the signature.
 %These are the two key properties of our approach.
@@
         \ForAll{$(B_1',\dots,B_n') \in A'$}
+            \Comment \luc{if cs(B; …( satisfied/?}
             \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
@@
 
-\subsubsection{Generating superblock assignments}
+\subsubsection{Generating superblock assignments (Step 3a)}
 % Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.
@@
 
-Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints happens:
+Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints works:
 Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
@@
 
-\subsubsection{Generating subblock assignments}
+\subsubsection{Generating subblock assignments (Step 3b)}
 \label{sec:algo:subgr}
@@
   This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition: $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$, where $B[i][j]$ indicates the $i$th value of the $j$th vector in $B$.
-  In Figure~\ref{fig:main_example} there is exactly one such satisfying subassignment: $(T_1[:7], T_1[:,3:6])$.
+  In Figure~\ref{fig:main_example} there is exactly one such satisfying subassignment: $(T_1[:7], T_1[:,3:6])$. \luc{why [j][i] and not [j,i] ?}
 \end{example}
@@
 \samuel{Tias, there is something wrong here:}
-This is inefficient, as the signature of the definition may already not be satisfied, or the first element in the data.
+This is inefficient, as already the signature of the definition may not be satisfied, or the first element in the data.
 
@@
 %For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
+\luc{test = definition, why use another term ?}
 Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
@@
     \State $A_{sub} \gets \emptyset$
-    \ForAll{$(B_1', \dots, B_n') \in \{ (B_1', \dots, B_n') | \forall i: B_i' \sqsubseteq B_i  \wedge \nexists i,j,i\neq j: B'_i \sqsubseteq B'_j \}$}
+    \ForAll{$(B_1', \dots, B_n') \textbf{ where } \forall i: B_i' \sqsubseteq B_i  \land \textit{disjoint}_{j \neq i}(B_i', B_j')$}
       \If{$Test((B_1',\dots,B_n'))$}
+      \Comment \luc{Test = def is satisfied}
         \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
@@
   Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
-  $P$~can be seen as a set of partial assignments in which the blocks for certain arguments have been fixed.
+  $P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed. \samuel{is it clear now?}
 \end{example}
@@
 Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$.
-Should, the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
+Should the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
 
@@
 
-The implementation is illustrated using a case study on the spreadsheet corresponding with the previously introduced example (figure~\ref{fig:main_example}).
+The implementation is illustrated using a case study on the spreadsheet corresponding to the previously introduced example (Figure~\ref{fig:main_example}).
 In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 30 (\samuel{check number}) spreadsheets that we assembled from various sources.
@@
 
-\samuel{Revisit [what does "implemented" mean here? what's the difference between "implemented" and "algorithm", esp given that we are talking experiments? ]}
+\samuel{Revisit [what does "implemented" mean here? what's the difference between "implemented" and "algorithm", esp given that we are talking experiments? ] }
+\luc{what is not implemented, strange \dots in a paper !!!}
 For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
@@
 The constraints we considered to be redundant are either duplications (results that can be calculated in different ways, one of which seems superior) or constraints that hold by accident.
+\luc{be more precise here !!! Q1 / Q2 }
 
@@
 
-Redundant constraints can be reduced in different ways.samuel
-Duplications should be detected in a post-processing step that uses entailment or heuristics to filter constraints.
+Redundant constraints can be reduced in different ways.
+Duplications should be identified in a post-processing step that uses entailment or heuristics to filter constraints.
 Accidental constraints may be detected through heuristics, however, they can be also be removed by adding additional data.
@@
 
+\samuel{Single paragraph / section}
 \paragraph{Dependencies}
-In order for the algorithm to run efficiently it is crucial to use dependencies and find constraint incrementally whenever possible.
+In order for the algorithm to run efficiently it is crucial to use dependencies and find constraints incrementally whenever possible.
 This avoids some of the explosion of combinations for constraints that have many arguments.

### 2016-05-27 Paper/ECAI/tabular.tex
polished and fixed the conclusions

--- 2016-05-27 Paper/ECAI/tabular.tex
@@
   \label{fig:autocompletion_example}
-
 \end{subfigure}
-
 \end{figure*}
-
-\sergey{Either we need to elaborate on the online detection or remove offline-online and simplify the rest, otherwise it is not clear what is going on}\samuel{Is this still the case?}
 
@@
 trying to leave out each row, and compare the constraints learned on this sample with the constraints learned by adding the row containing the mistake.
-The sample needs to large enough such that we have enough confidence that the constraints are correct.
+The sample needs to be large enough such that we have enough confidence that the constraints are correct.
 For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
@@
 blocks}, which are aimed at combinatorial optimization and scheduling rather than what is typically used in spreadsheets.
-Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints using techniques inspired on Mitchell’s version spaces to process examples, some variants also asking informative questions to the user.
+Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples, some variants also asking informative questions to the user.
 As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints\tias{it looks only at pairwise constraints, so an important factor is how fast the algorithm can converget to the entire constraint set.
@@
 At that point FlashFill would generate a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', would also apply it to automatically generate values for the other politicians (rows) in the table.
-While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'' \tias{only strings?}, FlashFill learns correct programs from very few examples.
+While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
 Flashfill has been extended to e.g.
@@
 Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
-However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify “hard” constraints.
+However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify ``hard'' constraints.
 
@@
 
-\paragraph{Future work}
-There are multiple directions for future work on this topic.
-Integrating the system in an interactive setting would offer users the possibility to easily receive and provide feedback.
-Moreover, the system can be adapted to deal with additional types produce less redundant constraints through the use of (heuristic) filtering or post-processing.
-
-Currently, the system only learns single constraints, however, extending our approach to nested constraints would allow more expressive concepts to be learned.
-This shift would bring the approach in line with programming by example.
+For the future development of the system there are multiple directions. Integrating the system in an interactive setting would offer users the possibility to easily receive and provide feedback.
+Moreover, the system can be adapted to deal with additional types and produce less redundant constraints through the use of (heuristic) filtering or post-processing.
+
+Currently, the system only learns single constraints and extending our approach to nested constraints would allow more expressive concepts to be learned.
+Also, \sname only works on complete rows and columns, allowing partial vectors is an important future direction for developing applications based on the system.
+%This shift would bring the approach in line with programming by example.
 

### 2016-05-27 Paper/ECAI/tabular.tex
Fixes

--- 2016-05-27 Paper/ECAI/tabular.tex
@@
 \item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
-\\\\
-\luc{these are the vectors along \dots along}
-\item \textit{rows}: this is orientation specific, for row-based blocks this is the size, for column-based blocks the length
-\item \textit{columns}: also orientation specific, for row-based blocks this is the length, for column-based blocks the size
+\item \textit{rows}: the number of rows in the block
+(in row-oriented blocks this is equivalent to the size)
+\item \textit{columns}: the number of columns in the block (in row-oriented blocks this is equivalent to the length)
 \end{itemize}
@@
 This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$.
-While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint;
-
-  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$. \luc{example where it holds}
+While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint; \samuel{check}
+
+  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$. 
+  For example, $B_r = [12, 7]$ and $\mathbf{\sg_x}$ consists of columns: $[5, 15]$, $[7, -8]$.
 \end{itemize}
@@
 This is because constraints may be logically equivalent or may be implied by other constraints.
+
 \luc{if you have permutation, alldiff is redundant.}
@@
 \item Extract headerless tables from tabular data
-\item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
+\item Partition the tables into contiguous, type-consistent \textit{superblocks}
 \item Generate for each constraint template all valid argument assignments in two steps:
 \begin{enumerate}
-\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal superblock and the $B_i$ are \textit{compatible} (defined below) with the signature. 
-\item For each generated atom $c(B_1, \ldots ,B_n)$ in 3a), find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
+\item For each constraint template~$s$, generate all assignments $(B_1, \ldots ,B_n)$ where each $B_i$ is a superblock and the $B_i$ are \textit{compatible} (defined below) with the signature of~$s$. 
+\item For each generated assignment $(B_1, \ldots ,B_n)$ in 3a), find all valid constraints $s(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of~$s$ are satisfied.
 \end{enumerate}
@@
 
-In step 3b, we start from an atom $c(B_1, \ldots, B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
+In step 3b, we start from an assignment $(B_1, \ldots, B_n)$ with superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
 As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier.
@@
 %In practice, this can be optimized further.
-\\\\
-\samuel{Section references not correct}
-We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3).
-We then describe a number of optimizations that we perform to speed up steps 3 and 4 (Section 4.4).
+
+We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step 3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step 3b, Section~\ref{sec:algo:subgr}).
+In Section~\ref{sec:opts} we describe a number of optimizations designed to speed up steps the algorithm.
 
@@
         \ForAll{$(B_1',\dots,B_n') \in A'$}
-            \Comment \luc{if cs(B; …( satisfied/?}
             \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
@@
 \subsubsection{Generating superblock assignments (Step 3a)}
+\label{sec:algo:super}
+\
 % Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.
@@
 Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
-The disjoint subassignments are then sequentially tested using a template-specific procedure \textit{Test} and valid subassignments are returned.
-The \textit{Test} procedure verifies the exact signature (e.g. subtypes will not have been checked yet) and checks whether the definition is satisfied for all data in the blocks.
+For every (disjoint) subassignment the exact signature (e.g. subtypes will not have been checked yet) and definition will be tested and all satisfying subassignments are returned.
 
@@
     \ForAll{$(B_1', \dots, B_n') \textbf{ where } \forall i: B_i' \sqsubseteq B_i  \land \textit{disjoint}_{j \neq i}(B_i', B_j')$}
-      \If{$Test((B_1',\dots,B_n'))$}
-      \Comment \luc{Test = def is satisfied}
+      \If{$\mathit{Sig}_s(B_1',\dots,B_n') \land \mathit{Def}_s(B_1',\dots,B_n')$}
         \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$

### 2016-05-28 Paper/ECAI/tabular.tex
WIP Comments

@@
 This is a new and unconventional machine learning problem.
-\samuel{A: be more specific about which headers, there's a lot to see in that figure - 
-use T1 / T2 to avoid confusion with regular table references
- - 
-why is T2 oriented differently from the others?
- - 
-what is "the corresponding column"?
- - 
-I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now, also in terms of formulations ("the row-wise sum of the previous columns", hu?
-do you mean that in each row, total is the sum of the values for quarters?)}
-Consider the example in Figure~\ref{fig:main_example}, where the header's names already suggest the usage of spreadsheet operations such as \textit{average} or \textit{sum}.
-Looking at the first row in Table~2 and the data in Table~1, it is clear that the values in this row were computed by summing the values in the corresponding column of Table~1.
-Examining Table~1 also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by taking the row-wise sum of the previous four columns.
+\samuel{A: I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
+Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
+Looking at the first row in Table~$T_2$ and the data in Table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
+Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
 This provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns.
 Here everything is mixed.
-The data is relational on the one hand, since we have multiple tables with relationships between them.
+The data is relational on the one hand, since we have multiple tables with relationships between them (e.g.
+foreign-keys).
 And on the other hand, the data is most often mixed textual and numeric.
 %, since people tend to use spreadsheets for financial and accounting computations.
-\samuel{A: at this point I have an idea what the abstract means by "function" constraints, but wonder what the "relation" ones are}
 One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
@@
 It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
-A more detailed discussion on related work is contained in Section~\ref{sec:related_work}.
 
@@
 Section \ref{sec:applications} shows how our system can be used for applications.
-Section \ref{sec:related_work} presents an overview of the related work.
+Section \ref{sec:related_work} presents a more detailed discussion of the related work.
 Section \ref{sec:conclusions} provides conclusions.
@@
 \\\\
-\samuel{A: how do you get from the signature and the definition to these predicates? I guess this is what I'm missing on the formal side above… this needs to be made explicit}
+\samuel{A: I'm a bit confused by the fact that you seem to be mixing the formal way of writing these triples with textual explanation, both in the example in the text and in the table; I'd prefer to see the formal thing and then get the explanation, as that's what I'd need to use your approach}
 \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
@@
 It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
-\samuel{Is terms the right word? Shouldn't it be arguments, too?}
-From a FOL perspective, the name of the constraint (e.g.
-\textit{SUM}) is just the predicate name, and the arguments $B_r$ and $B_x$ are the terms, which can be seen as either uninstantiated variables or as concrete values.
+From a FOL perspective, a constraint of the form $\ecrank{B_r}{B_x}$ can be seen as a predicate $\textit{RANK}(B_r, B_x)$ where \textit{RANK} is the name of the predicate and its arguments~$B_r$ and~$B_x$ are terms, which can be seen as either uninstantiated variables or as concrete values.
 This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
-For example for $\ecrank{B_r}{B_x}$ and the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
+For example, for the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
 
@@
 %
-\samuel{A: so you expect the user to define blocks? can this be automated?
-(also, later on, your solution seems to start from the tables rather than the blocks, why not match the two?)}
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf Given} a set of instantiated blocks ${\cal B}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$, $\forall i: B_i' \sqsubseteq B_i \in {\cal B}$ and $(B'_1, ..., B'_n)$ is a satisfied argument assignment of the template~$s$.
+{\bf Given} a set of instantiated blocks~${\cal B}$ over tables~${\cal T}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$, $\forall i: B_i' \sqsubseteq B_i \in {\cal B}$ and $(B'_1, ..., B'_n)$ is a satisfied argument assignment of the template~$s$.
 \end{definition}
@@
 
+Blocks and tables are expected to be given, in Section~\ref{sec:approach} we will discuss how these can be extracted from a spreadsheet.
 Figure~\ref{fig:sol_example} shows the solution to the tabular constraint learning problem when applied on the blocks of Figure~\ref{fig:main_example} and constraint templates listed in Table~\ref{table:constraints}.
@@
 
-Instead, we divide this step into two parts: in step 3a, we will not reason over individual (sub)blocks and their data, but rather over the properties of the maximal superblocks.
+Instead, we divide this step into two parts: in step 3a, we will not reason over individual (sub)blocks and their data, but rather over the properties of the superblocks.
 % and what combination of superblocks may lead to a valid argument assignment.\\
-Consider tables $T_1$ and $T_2$ in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{col}$ constraint template.
-Instead of considering all possible subblocks of $T_1$ and $T_2$, we first reason over the properties of the superblocks.
+Consider table~$T_1$ in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{row}$ constraint template.
+Instead of considering all possible subblocks of $T_1$, we first reason over the properties of the superblocks.
 For example, $B_2$ and $B_4$ are not numeric so they can be immediately discarded.
-\samuel{Iff Br is from T2?}
-Looking at the sizes, $B_1$ and~$B_2$ contain only one and two columns respectively, however, the rows in $T_2$ have length 4 so we only need to consider superblocks with at least 4 columns, which only $B_3$ satisfies.
-\luc{so the generated atom in step 3a is \dots }
+Superblocks~$B_1, B_3$ and $B_5$ are valid candidates for $B_r$, however, since they all have length~4, $\mathbf{B_x}$ needs at least 4~columns which is only satisfied by $B_3$.
+So the assignments generated in step 3a would be $(B_1, B_3), (B_3, B_3), (B_5, B_3)$.
 
@@
 Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
-\samuel{Tias, there is something wrong here:}
-This is inefficient, as already the signature of the definition may not be satisfied, or the first element in the data.
+This is inefficient, as already the blocks may not satisfy the signature or the first elements in the data may not satisfy the definition.
 
@@
 %For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
-\luc{test = definition, why use another term ?}
 Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
@@
 \subsubsection{Template dependencies}
-As discussed in Section~\ref{sec:form:dependencies}, some constraints depend on other constraints.
-This is apparent when looking at the template signatures in Table~\ref{table:constraints}, where one can see that some signatures depend on other constraints, such as \textit{LOOKUP} depending on \textit{FOREIGNKEY}.
-If a constraint template~$s_1$ includes template~$s_2$ in its signature, we say that~$s_1$ depends on~$s_2$.
+As discussed in Section~\ref{sec:form:dependencies}, some constraint templates depend on others by including templates they depend on in their signature (see Table~\ref{table:constraints}).
 
@@
 We assume that signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
-Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signature definitions in Table~\ref{table:constraints}. 
+Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signatures in Table~\ref{table:constraints}. 
 Constraint templates that have no dependencies are omitted.
@@
 Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
-Concretely, the input templates~$\temps$ (see Algorithm~\ref{algo:tcl}) are reordered to agree with the partial ordering imposed by~\dependencies.
-
-The dependency graph represents a partial ordering and is used to produce a linear ordering of the given templates such that more general templates occur before templates that depend on them (e.g.
+Concretely, the input templates~$\temps$ (see Algorithm~\ref{algo:tcl}) are reordered to agree with the partial ordering imposed by~\dependencies (e.g.
 \textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
 Considering Algorithm~\ref{algo:tcl}, this step can be seen as preprocessing the input~\temps.
-Learning in this order allows templates to use constraints they depend on as these will have already been found.
+Learning in this order allows templates to use constraints they depend on, as these constraints will have already been learned.
 This is exploited in the \generategroups step to speedup the algorithm by reducing the search space.
-Let us illustrate this step with an example.
 
@@
 The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
-Their inclusion was based on the analysis of the most popular Excel constraints \samuel{How do we know?} as well as the spreadsheets we encountered.
-All these constraints are available in popular spreadsheet software, except for the structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}).
-These constraints were added so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
+We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
+We also added three structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
 
@@
 \subsection{Case study}
-\tias{Move explanation of origin to first time it is introduced}
 Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.

### 2016-05-28 Paper/ECAI/tabular.tex
Rewrite table definitions

@@
     Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.
+    Most tables only contain type-consistent columns, however, the data in table~$T_2$ can be used as both column- and row vectors.
     This example was created by combining several smaller Excel sheets originating from an exercise session.
@@
 \begin{table*}
-\caption{Overview of constraint templates implemented in \sname.
-The templates marked with $\dagger$ are aggregate constraint templates, the table shows the version for sum but the implementation also supports max, min, average, product and count.
-Subgroups in normal font have to be vectors, while subgroups in \textbf{bold} may contain more than one vector.
-Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.}\label{table:constraints}
+\caption{
+  Overview of constraint templates implemented in \sname.
+  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows the version for sum but \sname also supports max, min, average, product and count.
+  Subgroups in normal font have to be vectors, whereas subgroups in \textbf{bold} may contain more than one vector.
+  Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.
+  $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
+}\label{table:constraints}
   {\centering
@@
       & $\discrete(\sg_x)$
+
       & All values in $\sg_x$ are different: $\sg_x[i] \neq \sg_x[j]$ if $i \neq j$
@@
     %X = Y & & \\
+
     $\ecperm{\sg_x}^*$
       & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
+
       & The values in $\sg_{x}$ are a permutation of the numbers $1$ through $\plength(\sg_{x})$.
       \\ \hline
+
     \ecseries{\sg_x}
       & $\integer(\sg_{x})$ and $\ecperm{\sg_{x}}$
+
       & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
       \\ \hline
-    $\ecfkey{\sg_{fk}}{\sg_{pk}}^*$ & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$ & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
-    % TODO check out
+
+    $\ecfkey{\sg_{fk}}{\sg_{pk}}^*$
+      & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$
+
+      & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
+
     \eclookup{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
       & $\sg_{fk}$ and $\sg_{pk}$ are both $\discrete$; arguments $\{\sg_{fk}, \sg_{r}\}$ and $\{\sg_{pk}, \sg_{val}\}$ within the same set have the same \plength, \ptable and \por; $\sg_{r}$ and~$\sg_{val}$ have the same type; and \ecfkey{\sg_{fk}}{\sg_{pk}}.
+
       & $\sg_r[i]$ is the same value as looking up~$\sg_{fk}[i]$ in~$\sg_{pk}$  and returning the corresponding value in~$\sg_{val}$.
       \\ \hline
+
     \eclookupfuzzy{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
       & \textit{Same as lookup}
+
       & $\sg_r[i]$ is the same value as looking up the last item in~$\sg_{pk}$ smaller than~$\sg_{fk}[i]$ and returning the corresponding value in~$\sg_{val}$.
       \\ \hline
+
     \eclookupprod{\sg_r}{\sg_1}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
       & Arguments $\{\sg_{r}, \sg_{1}, \sg_{fk}\}$ are $\numeric$, arguments $\{\sg_{pk}, \sg_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sg_{fk}}{\sg_{pk}}.
+
       & $\sg_{r}[i]$ is the obtained by multiplying $\sg_{1}[i]$ with $\eclookupf{\sg_{fk}}{\sg_{pk}}{\sg_{val}}{}[i]$.
       \\ \hline
+
     \ecprod{\sg_r}{\sg_1}{\sg_2}
       & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$
+
       & $\sg_{r}[i] = \sg_{1}[i] \times \sg_{2}[i]$.
       \\ \hline
+
     \ecdiff{\sg_r}{\sg_1}{\sg_2}
       & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
+
       & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
       \\ \hline
+
     \ecproj{\sg_r}{\mathbf{\sg_x}}
-      & Arguments $\{\sg_{r}, \mathbf{\sg_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{\sg_x}$ contains at least~2 vectors; and $\sg_r = \mathit{SUM}(\mathbf{\sg_x}, \por(\mathbf{\sg_x}))$
+      & Arguments $\{\sg_{r}, \mathbf{\sg_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{\sg_x}$ contains at least~2 vectors; and $\sg_r = \mathit{SUM}_{\por(\mathbf{\sg_x})}(\mathbf{\sg_x})$
+
       & At every position~$i$ in $1$ through $\plength(\sg_{r})$ there is exactly one vector~$v$ in $\mathbf{\sg_x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sg_{r}[i]$.
       \\ \hline
+
     \ecrank{\sg_r}{\sg_x}
       & $\integer(\sg_{r})$; $\numeric(\sg_{x})$; and $\plength(\sg_{r}) = \plength(\sg_{x})$
+
       & The values in $\sg_{r}$ represent the rank (from largest to smallest) of the values in $\sg_{x}$ (including ties)
       \\ \hline
+
     \ectotal{\sg_r}{\sg_{pos}}{\sg_{neg}}
       & Arguments $\{\sg_{r}, \sg_{pos}, \sg_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
-      & The values in $\sg_{r}$ are a running total, each value $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
+
+      & $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
       \\ \hline
+
     $\ecsumr{\sg_r}{\mathbf{\sg_x}}^\dagger$
       & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\pcols(\mathbf{\sg_x}) \geq 2$; and $\prows(\mathbf{\sg_x}) = \plength(\sg_r)$
-      & Each value in $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
+
+      & $B_r[i] = \sum ($row~$i$ of $\mathbf{B_x})$
       \\ \hline
+
     $\ecsumc{\sg_r}{\mathbf{\sg_x}}^\dagger$
       & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\prows(\mathbf{\sg_x}) \geq 2$; and $\pcols(\mathbf{\sg_x}) = \plength(\sg_r)$
-      & Each value in $\sg_{r}$ is obtained by summing over the corresponding column in $\mathbf{\sg_x}$.
+
+      & $B_r[i] = \sum ($column~$i$ of $\mathbf{B_x})$
       \\ \hline
+
     $\ecsumif{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}^\dagger$
       & $\sg_{fk}, \sg_{pk}$ are $\discrete$; $\sg_{r}, \sg_{val}$ are $\numeric$; within the sets $\{\sg_{val}, \sg_{fk}\}$ and $\{\sg_{pk}, \sg_{r}\}$ arguments have the same $\plength$ and $\por$; $\sg_{fk}$ and $\sg_{val}$ have the same $\ptable$; $\sg_{fk}$ and $\sg_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sg_{pk}}
-      & The value for $\sg_{r}[i]$ is obtained by summing all values $\sg_{val}[j]$ where $\sg_{fk}[j] = \sg_{pk}[i]$
-      \\ \hline
+
+      & 
+      \[
+        B_r[i] = \sum_{j=1}^{\plength(B_{\textit{val}})} \begin{cases}
+          \sg_{val}[j] & \text{if } \sg_{fk}[j] = \sg_{pk}[i] \\
+          0 & \text{otherwise}
+        \end{cases}
+      \] \\ \hline
+
     \ecsumprod{\sg_r}{\sg_1}{\sg_2}
       & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are all $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
-      & $\sg_{r}[i] = \sum_{i = 1}^{\plength(\sg_{1})} \sg_{1}[i] \times \sg_{2}[i]$.
+
+      & $\sg_{r}[1] = \sum_{j = 1}^{\plength(\sg_{1})} \sg_{1}[j] \times \sg_{2}[j]$.
   \end{tabularx}}

### 2016-05-28 Paper/ECAI/tabular.tex
WIP working on comments

@@
   $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
-}\label{table:constraints}
+\luc{there are some fancy constraints here.
+it might be good to say something about them ...
+like an example of one of the more complex ones in the text ...
+maybe state what conditional means, what aggregate ...
+}
+}
+\label{table:constraints}
   {\centering
@@
 
-  \luc{there are some fancy constraints here.
-it might be good to say something about them ...
-like an example of one of the more complex ones in the text ...
-maybe state what conditional means, what aggregate ...
-}
 \end{table*}
@@
   \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
-The two blocks have to be numeric.
+  Both blocks have to be numeric.
   This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$.
-While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint; \samuel{check}
+  Moreover, we add to the bias that the number of columns to sum over is at least~$2$.
 
@@
 Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}.
-This is because constraints may be logically equivalent or may be implied by other constraints.
-
-\luc{if you have permutation, alldiff is redundant.}
+This is because constraints may be logically equivalent or may be implied by other constraints, e.g.
+if you know $\ecperm{B_x}$, $\ecalldiff{B_x}$ is redundant.
 
@@
 \end{enumerate}
-\luc{introduce the notion of a maximal superblock earlier ???}
 
@@
 
-\samuel{Angelika: trim this section for repeated information}
 Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
@@
 Aside from finding errors, the system can be extended to deal with noise and the ability to learn soft constraints.
-Soft constraints are the (potentially weighted) constraints that hold only on some of the data.
+Soft constraints are (potentially weighted) constraints that hold only on some of the data.
 This would extend the approach to new application domains as well as provide more native error detection.
-
-\tias{write limitations: single cons, complete rows/cols only, (what else?) and deal with it as future work}.
 \bibliographystyle{plain}

### 2016-05-29 Paper/ECAI/tabular.tex
Processing comments

@@
 This is a new and unconventional machine learning problem.
-\samuel{A: I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
 Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
@@
     This example was created by combining several smaller Excel sheets originating from an exercise session.
+    \samuel{A: I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
   }
@@
   $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
-\luc{there are some fancy constraints here.
+{\small\luc{there are some fancy constraints here.
 it might be good to say something about them ...
@@
 maybe state what conditional means, what aggregate ...
-}
+}}
 }

### 2016-05-29 Paper/ECAI/tabular.tex
Tweaks on table

@@
 \newcommand{\sg}{B}
+\newcommand{\sbl}[1]{\ensuremath{B_{\textit{#1}}}}
 
@@
 
-      & All values in $\sg_x$ are different: $\sg_x[i] \neq \sg_x[j]$ if $i \neq j$
+      & All values in $\sg_x$ are different. $i \neq j$: $\sg_x[i] \neq \sg_x[j]$
       \\ \hline
@@
     $\ecfkey{\sg_{fk}}{\sg_{pk}}^*$
-      & $\sg_{fk}$ and~$\sg_{pk}$ are both $\discrete$; $\ptype(\sg_{fk}) = \ptype(\sg_{pk})$; $\ptable(\sg_{fk}) \neq \ptable(\sg_{pk})$; and $\ecalldiff{\sg_{pk}}$
+      & $\sg_{\mathit{fk}}$ and~$\sg_{pk}$ are $\discrete$; they must have different $\ptable$s but the same $\ptype$; and $\ecalldiff{\sbl{pk}}$
 
@@
 
-    \eclookup{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
-      & $\sg_{fk}$ and $\sg_{pk}$ are both $\discrete$; arguments $\{\sg_{fk}, \sg_{r}\}$ and $\{\sg_{pk}, \sg_{val}\}$ within the same set have the same \plength, \ptable and \por; $\sg_{r}$ and~$\sg_{val}$ have the same type; and \ecfkey{\sg_{fk}}{\sg_{pk}}.
-
-      & $\sg_r[i]$ is the same value as looking up~$\sg_{fk}[i]$ in~$\sg_{pk}$  and returning the corresponding value in~$\sg_{val}$.
+    \eclookup{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
+      & $\sbl{fk}$ and $\sbl{pk}$ are $\discrete$; arguments $\{\sbl{fk}, \sbl{r}\}$ and $\{\sbl{pk}, \sbl{val}\}$ within the same set have the same \plength, \ptable and \por; $\sbl{r}$ and~$\sbl{val}$ have the same type; and \ecfkey{\sbl{fk}}{\sbl{pk}}.
+
+      & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] = \sbl{fk}[i]$
       \\ \hline
 
-    \eclookupfuzzy{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+    \eclookupfuzzy{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
       & \textit{Same as lookup}
 
-      & $\sg_r[i]$ is the same value as looking up the last item in~$\sg_{pk}$ smaller than~$\sg_{fk}[i]$ and returning the corresponding value in~$\sg_{val}$.
+      & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] \leq \sbl{fk}[i]$, $j$~maximal
       \\ \hline
 
-    \eclookupprod{\sg_r}{\sg_1}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}
+    \eclookupprod{\sbl{r}}{\sbl{1}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
       & Arguments $\{\sg_{r}, \sg_{1}, \sg_{fk}\}$ are $\numeric$, arguments $\{\sg_{pk}, \sg_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sg_{fk}}{\sg_{pk}}.
 
-      & $\sg_{r}[i]$ is the obtained by multiplying $\sg_{1}[i]$ with $\eclookupf{\sg_{fk}}{\sg_{pk}}{\sg_{val}}{}[i]$.
+      & $\sbl{r}[i] = \sbl{1}[i] \times \eclookupf{\sbl{fk}}{\sbl{pk}}{\sbl{val}}{}[i]$.
       \\ \hline
@@
     \ecsumprod{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are all $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
+      & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
 

### 2016-05-29 Paper/ECAI/tabular.tex
Improved evaluation

@@
 
+\newcommand{\benchsize}{??}
+
 \section{Evaluation}\label{sec:evaluation}
@@
 In this section we experimentally validate our approach.
-We studied various questions \samuel{Mention questions here?}, most notably with what accuracy our algorithm can find constraints.
+We studied three main questions concerning the accuracy (\textit{Q1}), precision (\textit{Q2}) and speed (\textit{Q3}) of our algorithm.
 
@@
 Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
-
-\subsubsection{Results}
 \sname takes a few seconds to find the constraints in Figure~\ref{fig:sol_example} (excluding structural constraints).
@@
 All spreadsheets have been converted to CSV files.
-Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases blocks where also provided manually).
+Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases with $\mathit{None}$ values, blocks where also provided manually).
+\samuel{chec}
 
@@
 \luc{what is not implemented, strange \dots in a paper !!!}
-For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expected to be discovered.
-We consider three types of constraints:
-\begin{description}
-  \item[Implemented] \samuel{Dupl} Constraint that have been implemented and are expected to be found (all constraints in Table~\ref{table:constraints})
-  \item[Essential] \samuel{Dupl} Constraint that have or have not been implemented but could be found using our algorithm (e.g. fuzzy conditional sum)
-  \item[Non-trivial] Constraints currently outside of the scope of this system (e.g. generic nested mathematical or logical formulas or n-ary constraints)
-\end{description}
-
-We examine three main questions concerning the accuracy (recall), redundancy (precision) and efficiency of our approach, our main focus being accuracy.
-
-\subsubsection*{Q1. How accurate is the approach?}
-\samuel{Sounds like we were surprised}
-Fortunately, for the implemented constraint templates, our system achieves a recall of $1.0$, i.e. it is able to find all of the supported constraints (90) on the benchmark suite.
-There are only three templates that have not been implemented yet, so w.r.t. the ground truth the total recall of our system is $0.97$.
-
-\subsubsection*{Q2. How many redundant constraints are discovered?}
-Our primary focus was, as mentioned before, accuracy, therefore we sometimes traded off less redundancy for more accuracy.
-The motivation is that solutions can still be pruned using entailment or heuristics afterwards.
-The constraints we considered to be redundant are either duplications (results that can be calculated in different ways, one of which seems superior) or constraints that hold by accident.
-\luc{be more precise here !!! Q1 / Q2 }
-
-Across all spreadsheets our implementation finds 121 constraints, 28 ($23.14\%$) of which are redundant.
-However, the average redundancy per spreadsheet is only $8.33\%$.
-Examining the redundant constraints reveals that many (12) of them are duplications occurring in a single spreadsheet.
-Many of the remaining constraints are duplications stemming from the overlapping role of difference and sum.
-Accidental constraints are limited to the 5 rank constraints that were discussed in the case study.
-
-Redundant constraints can be reduced in different ways.
-Duplications should be identified in a post-processing step that uses entailment or heuristics to filter constraints.
-Accidental constraints may be detected through heuristics, however, they can be also be removed by adding additional data.
+For every spreadsheet a ground-truth has been provided manually, specifying the \textbf{relevant} constraints that are expected to be (re-)discovered.
+Not all relevant constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
+Constraints that are currently outside of the scope of this system, e.g. generic nested mathematical or logical formulas, are ignored.
+
+\subsubsection*{Q1. How accurate \sname?}
+Our implementation achieves a recall of~$0.97$ for relevant constraints, i.e. it finds~90 of~93 relevant constraints in the benchmark suite.
+The recall for supported constraints is $1.0$ (90 of 90).
+
+\subsubsection*{Q2. How precise is \sname?}
+Our implementation finds 121 constraints across all spreadsheets, 28 of which are not present in the ground-truth.
+Therefore, the calculated precision is \samuel{Check numbers}.
+Notable, the average precision per spreadsheet is \samuel{math}.
+This gap is explained by the large number of additional constraints produced by few spreadsheets \samuel{More info?}.
+
+Examining the additional constraints more closely reveals that only \samuel{x} of the constraints are spurious, i.e. they hold by accident and can be disproved by adding additional data.
+Most constraints are \textit{redundant}, such as duplications or implications of other constraints.
+Many redundant constraints are the result of bad spreadsheet design, i.e. copying data in multiple places, and the overlapping role of difference and the sum aggregate.
+
+We consider precision (i.e. avoiding redundant constraints) a secondary goal, focusing primarily on accuracy.
+The rationale is that the precision can be increased by post-processing the constraints using entailment or heuristics.
 
@@
 Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
-For the 32 spreadsheets in the benchmark our implementation ran in ${\runtotal}s \pm {\runtotalstd}s$.
+For the \benchsize~spreadsheets in the benchmark our implementation ran in ${\runtotal}s \pm {\runtotalstd}s$.
 The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
@@
 \samuel{Single paragraph / section}
-\paragraph{Dependencies}
+\paragraph*{Dependencies}
 In order for the algorithm to run efficiently it is crucial to use dependencies and find constraints incrementally whenever possible.

### 2016-05-30 Paper/ECAI/tabular.tex
Several corrections

@@
 \newcommand{\sbl}[1]{\ensuremath{B_{\textit{#1}}}}
+\newcommand{\bsbl}[1]{\ensuremath{\mathbf{B_{\textit{#1}}}}}
 
@@
 
+% Moved up
+The question that we answer in this paper is: is it possible to discover or reconstruct constraints (relations, functions) in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
+It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given.
+%, and discovers column- and row-wise constraints.
+We demonstrate the utility of our approach in an experimental evaluation.
+Moreover, we sketch additional application scenario's such as autocompletion and error detection.
+
 The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
@@
 It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
-
-\samuel{Move this up a paragraph}
-The question that we answer in this paper is: is it possible to discover or reconstruct constraints (relations, functions) in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
-%Furthermore, the algorithm accepts declarative specifications of the constraints and predicates of interest (the "bias"). \tias{revise declarative claim?}
-It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given.
-%, and discovers column- and row-wise constraints.
-We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch additional application scenario's such as autocompletion and error detection.
 \\\\
@@
     This example was created by combining several smaller Excel sheets originating from an exercise session.
-    \samuel{A: I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
+    % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
   }
@@
 %
-An example is that $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.
+An example of block containment is $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.
 
@@
   $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
-{\small\luc{there are some fancy constraints here.
-it might be good to say something about them ...
-like an example of one of the more complex ones in the text ...
-maybe state what conditional means, what aggregate ...
-}}
+  {\small\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
 }
@@
 
-      & $B_r[i] = \sum ($row~$i$ of $\mathbf{B_x})$
+      & $B_r[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$
       \\ \hline
@@
 
-      & $B_r[i] = \sum ($column~$i$ of $\mathbf{B_x})$
+      & $B_r[i] = \sum_{j = 1}^{\prows(\bsbl{x})} \mathit{column}(i, \bsbl{x})[j]$
       \\ \hline
@@
 
-      & 
-      \[
-        B_r[i] = \sum_{j=1}^{\plength(B_{\textit{val}})} \begin{cases}
+      & \[ B_r[i] = \sum_{j=1}^{\plength(B_{\textit{val}})} \begin{cases}
           \sg_{val}[j] & \text{if } \sg_{fk}[j] = \sg_{pk}[i] \\
@@
 \end{table*}
+
+\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
+\newcommand{\defc}{\ensuremath{\format{Def}_s}}
 
@@
 In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
-\\\\
-\samuel{A: I'm a bit confused by the fact that you seem to be mixing the formal way of writing these triples with textual explanation, both in the example in the text and in the table; I'd prefer to see the formal thing and then get the explanation, as that's what I'd need to use your approach}
+We can capture this bias for a template~$s$ using a predicate $\sigc$.
 \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
@@
 %the arguments are full instantiated to decide whether the constraint is satisfied or not.
-In logical and relational learning this is known as the background knowledge. %, which provides the definition of the predicate.
+In logical and relational learning this is known as the background knowledge.
+We introduce the predicate \defc~to capture this background knowledge for a template~$s$.
 \end{itemize}
@@
 \begin{example}
-Several constraint templates are illustrated in Table \ref{table:constraints}. A non-trivial example is the constraint template for the row-based sum:
+The constraint templates implemented in \sname are defined in Table~\ref{table:constraints}.
+A non-trivial example is, for example, the constraint template for the row-based sum:
 
@@
 
-\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
-\newcommand{\defc}{\ensuremath{\format{Def}_s}}
-%\tias{TODO: conceptually $c(A_1, ..., A_n) :- Sig_c(A_1, ..., A_n) \wedge Def_c(A_1, ..., A_n)$ where '$c$' is here both the constraint name and the template}.
 With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
-\samuel{A: how do you get from the signature and the definition to these predicates? I guess this is what I'm missing on the formal side above… this needs to be made explicit}
 We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
@@
   \includegraphics[width=1\linewidth]{figures/tabletool.png}
-  \caption{This figure shows a simple table selection tool. A user selects rectangular ranges of cells that correspond to (headerless) tables, an explicit orientation can also be provided.}
+  \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding (headerless) tables, an explicit orientation can also be provided.}
   \label{fig:visual_tool}
@@
   Note that this assignment does not satisfy the signature yet, $\sg_r$ contains more than one vector and $\psize(\mathbf{\sg_x}) \neq \plength(B_r)$.
-  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition: $\forall i, \sg_r[0][i] = \sum_j \mathbf{\sg_x}[j][i]$, where $B[i][j]$ indicates the $i$th value of the $j$th vector in $B$.
-  In Figure~\ref{fig:main_example} there is exactly one such satisfying subassignment: $(T_1[:7], T_1[:,3:6])$. \luc{why [j][i] and not [j,i] ?}
+  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition:
+  $\forall i, \sbl{r}[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$.
+  In Figure~\ref{fig:main_example} there is exactly one such subassignment: $(\range{T_1}{\rangeall}{7}, \range{T_1}{\rangeall}{\rangeto{3}{6}})$.
 \end{example}
@@
   Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
-  $P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed. \samuel{is it clear now?}
+  $P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed.
 \end{example}
@@
 A last type of redundancy we aim to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
-For example, consider the constraint \ecsumc{B_r}{B[1:n]} where $B[k:n]$ consists of only zeros, then \ecsumc{B_r}{B[1:j]} will be true for all $k - 1 \leq j \leq n$.
+For example, consider the constraint \ecsumc{B_r}{B[1{:}n]} where $B[k{:}n]$ consists of only zeros, then \ecsumc{B_r}{B[1{:}j]} will be true for all $k - 1 \leq j \leq n$.
 In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.
@@
 As a result, in some cases a maximal subblock might falsely include irrelevant columns.
-Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$.
-Should the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
+Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$.
+Should the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
 
 Alternatively, we can output all valid argument assignments instead of only maximal subblocks, e.g.
-$\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1:3])}$, however, this might produce many redundant constraints and introduce constraints that accidentally cover too little data.
+$\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$, however, this might produce many redundant constraints and introduce constraints that accidentally cover too little data.
 
@@
 
-\subsubsection*{Q1. How accurate \sname?}
+\subsubsection*{Q1. How accurate is \sname?}
 Our implementation achieves a recall of~$0.97$ for relevant constraints, i.e. it finds~90 of~93 relevant constraints in the benchmark suite.
@@
 
-\subsubsection*{Q3. How fast is the algorithm?}
+\subsubsection*{Q3. How fast is \sname?}
 Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
@@
 \begin{table}
+  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
   \centering
@@
     & \textbf{Total} & \textbf{Average per spreadsheet} \\
-    \textbf{Constraints} & $93$ & $2.91$ \\
-    \textbf{Accuracy} & $96.77\%$ & $94.27\%$ \\
-    \textbf{Redundant} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
+    \textbf{Relevant constraints} & $93$ & $2.91$ \\
+    \textbf{Recall} & $96.77\%$ & $94.27\%$ \\
+    \textbf{Precision} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
     \textbf{Speed (s)} & ${\runtotal}s \pm {\runtotalstd}s$ & ${\runfile}s \pm {\runfilestd}s$
   \end{tabular}
-  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
 \end{table}
@@
 
-\samuel{Single paragraph / section}
-\paragraph*{Dependencies}
+We explained various optimizations, most notably the exploitation of dependencies.
+Using the dependencies to find constraints incrementally cuts down large parts of the search trees, especially for constraints that have many arguments.
+Essentially, one obtains a more specialized search without much effort.
+
 In order for the algorithm to run efficiently it is crucial to use dependencies and find constraints incrementally whenever possible.

### 2016-05-30 Paper/ECAI/tabular.tex
Shortening

@@
 It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
-very few examples.  Our approach contrasts with these in that it focusses
+very few examples.
+Our approach contrasts with these in that it focuses
 on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional constraint satisfaction problems (CSPs).
-It contrasts with Flashfill in that it can learn from numeric data too, as well as general constraints in addition to functions.
-\\\\
+It contrasts with Flashfill in that it can learn from numeric data, too, as well as general constraints in addition to functions.
+
 This paper is organized as follows.
@@
     Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.
-    Most tables only contain type-consistent columns, however, the data in table~$T_2$ can be used as both column- and row vectors.
-    This example was created by combining several smaller Excel sheets originating from an exercise session.
+    Most tables only contain type-consistent columns, however, table~$T_2$ also contains type-consistent row.
+    This example was created by combining several Excel sheets from several exercise sessions.
     % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
@@
 
-A row or a column is \textbf{type-consistent} if all cells in that row or column are of the same base type, that is, numeric or textual.
-We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$, and similarly $T[{:},a]$ for the $a$-th column.
-For example in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_5[:,1] = ['$Diana Coolen$', 5]$.
-The latter is not type-consistent while the former is.
+A row or a column is \textbf{type-consistent} if all cells in that row or column are of the same base type, i.e. numeric or textual.
+We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$ and similarly $T[{:},a]$ for the $a$-th column.
+For example, in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_5[:,1] = ['$Diana Coolen$', 5]$.
+$T_5[:,1]$ is not type-consistent while $T_1[1,:]$ is.
 %We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.
@@
 \item \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
-This can concern any property of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
+This can concern properties of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
 In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
@@
 \begin{itemize}
-  \item \CName: $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, where $\sg_r$ and $\mathbf{\sg_x}$ are the arguments;
+  \item \CName: $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, for arguments $\sg_r$ and $\mathbf{\sg_x}$.
 
@@
 Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$.
-Should the target constraint be $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
+Should the target constraint be $\eccalc{B_2}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
 
@@
 All spreadsheets have been converted to CSV files.
-Definitions of tables were added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases with $\mathit{None}$ values, blocks where also provided manually).
-\samuel{chec}
+Definitions of tables are given (in some harder cases with $\mathit{None}$ values, blocks are defined manually). \samuel{chec}
 
@@
 \begin{table}
-  \caption{Overview evaluation on essential constraints of the collected Spreadsheet dataset}
+  \caption{Overview evaluation on benchmark}
   \centering
   \begin{tabular}{lll}
-    & \textbf{Total} & \textbf{Average per spreadsheet} \\
+    & \textbf{Total} & \textbf{Spreadsheet (avg)} \\
     \textbf{Relevant constraints} & $93$ & $2.91$ \\

### 2016-05-30 Paper/ECAI/tabular.tex
h!
Line spacing
worked with refs
some tweaking on related work

@@
 \usepackage[dvipsnames]{xcolor}
+\newcommand{\PADSEP}{-7pt}
 \newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
@@
 
-\begin{table*}
+\begin{table*}[!h]
 \caption{
@@
       & All values in $\sg_x$ are different. $i \neq j$: $\sg_x[i] \neq \sg_x[j]$
-      \\ \hline
-    %X = Y & & \\
+      \\[\PADSEP] \hline
 
@@
       & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
 
-      & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\ \hline
+      & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\[\PADSEP] \hline
 
@@
       & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] = \sbl{fk}[i]$
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] \leq \sbl{fk}[i]$, $j$~maximal
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & $\sbl{r}[i] = \sbl{1}[i] \times \eclookupf{\sbl{fk}}{\sbl{pk}}{\sbl{val}}{}[i]$.
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & $\sg_{r}[i] = \sg_{1}[i] \times \sg_{2}[i]$.
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & At every position~$i$ in $1$ through $\plength(\sg_{r})$ there is exactly one vector~$v$ in $\mathbf{\sg_x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sg_{r}[i]$.
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & The values in $\sg_{r}$ represent the rank (from largest to smallest) of the values in $\sg_{x}$ (including ties)
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & $B_r[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
       & $B_r[i] = \sum_{j = 1}^{\prows(\bsbl{x})} \mathit{column}(i, \bsbl{x})[j]$
-      \\ \hline
+      \\[\PADSEP] \hline
 
@@
         \end{cases}
-      \] \\ \hline
+      \] \\[\PADSEP] \hline
 
@@
 \sname combines ideas from several existing
-approaches from multiple different fields of research.
+approaches and different fields of research.
 
@@
 constraints in multiple relations (or tables). However, unlike typical logical and relational learning
-approaches, it focuses on data in spreadsheets and on constraints that can be column or row-wise.
+approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
 Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
-Still several ideas from logical and relational learning
+Nevertheless, several ideas from logical and relational learning
 proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
 Also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha}
-although this line of research has focused on more specialized techniques for specific types of constraints. Furthermore, the constraints used
-in spreadsheets are often quite different than those in databases.
-
-Second, there exist several algorithms in the constraint programming community that induce constraint (programs)
-from one or more examples and questions. Two well-known lines of research include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
+although that line of research has focused on more specialized techniques for specific types of constraints. Furthermore, the constraints and formulas used
+in spreadsheets are often different from those in databases.
+
+Second, there exist a number of algorithms in the constraint programming community that induce constraint (programs)
+from one or more examples and questions. Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
 The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
-To this aim, it cleverly enumerates different tables from the vector.
-For instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
-Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and also that ModelSeeker’s constraints are the typical ones from the constraint programming community\tias{which is all over integers, we also over strings and real-valued, AND, not one vector but we have many indep.
-blocks}, which are aimed at combinatorial optimization and scheduling rather than what is typically used in spreadsheets.
-Quacq \cite{Quacq} and its predecessor Conacq \cite{Conacq} acquire a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples, some variants also asking informative questions to the user.
-As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints\tias{it looks only at pairwise constraints, so an important factor is how fast the algorithm can converget to the entire constraint set.
-Redundant and spur also issue for these methods}.
+To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
+Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and that ModelSeeker's constraints are the typical ones from the constraint programming community, aimed at combinatorial optimization and scheduling, and restricted to integer values; furthermore we can consider multiple tables and sheets.
+Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user whether a partical constraint is satisfied or not.
+Each of the examples/queries is processed for each constraint, and hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
+for Sudoku).
+The focus is again on integer variables and typical combinatorial optimisation constraints.
+%As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
 As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.
@@
 Third, our work borrows from the seminal work of Gulwani et al.
-\cite{flashfill} on program synthesis for tabular data that is incorporated in Microsoft’s Excel.
+\cite{flashfill} on program synthesis for tabular data, which is incorporated in Microsoft’s Excel.
 In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians.
 The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''.
-At that point FlashFill would generate a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', would also apply it to automatically generate values for the other politicians (rows) in the table.
+At that point FlashFill generates a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', and also applies it to the other politicians (rows) in the table.
 While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
 Flashfill has been extended to e.g.
-FlashExtract \cite{flashextract} to produce a set of tables starting from a document (a txt, spreadsheet or HTML file) with a set of examples marked for extraction.
+FlashExtract \cite{flashextract} to produce a set of tables starting from an unstructured document (a txt or HTML file) with a set of examples marked for extraction.
 However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.
@@
 This would extend the approach to new application domains as well as provide more native error detection.
+
 \bibliographystyle{plain}
-\bibliography{references}
+{\footnotesize
+  \bibliography{references}}
 \end{document}

### 2016-05-30 Code/src/accuracy.py
Added experiments and updated code

@@ -8,13 +8,53 @@ import workflow
 from core.constraint import *
 from engine.util import local
 
-files = ["bmi", "age_department_sumif", "average_ablebits", "columnwise-sum-rows", "examples", "expenses",
-         "external_revenue", "fbi_offenses_corr",
-         "financial_result", "fruits", "help_lookup", "household", "inventory", "multsum",
-         "paper_supply", "price_weight", "repair", "rides", "sales_blanks", "school", "score",
-         "shares", "sum_if_double_condition", "sum_table_from_guide", "sumif_blanks", "sumif_example",
-         "sumif_games_toys", "sumif_region", "sumif_uk", "sumproduct", "week_2_busn_store_2", "sumif_diskettes",
-         "pattern_mining"]
+exercises = [
+    "baltimore",
+    "bedrijfsresultaten",
+    "belgium",
+    "bmi",
+    "expenses",
+    "price_weight",
+    "rides",
+    "score",
+    "shares",
+]
+
+tutorials = [
+    "sumif_region",
+    "age_department_sumif",
+    "average_ablebits",
+    "fruits",
+    "help_lookup",
+    "household",
+    "inventory",
+    "multsum",
+    "paper_supply",
+    "repair",
+    "sales_blanks",
+    "school",
+    "sum_if_double_condition",
+    "sum_table_from_guide",
+    "sumif_blanks",
+    "sumif_diskettes",
+    "sumif_example",
+    "sumif_games_toys",
+    "sumif_uk",
+    "sumproduct",
+    "week_2_busn_store_2",
+]
+
+real = [
+    # glance - missing
+    "external_revenue",
+    "financial_result",
+    "fbi_offenses_corr",
+    "pattern_mining",
+]
+
+# files = exercises + tutorials + real
+files = real
+
 
 def is_excel_constraint(c: Constraint):
     return isinstance(c, Aggregate) or isinstance(c, ConditionalAggregate) or isinstance(c, Series) \

### 2016-05-30 Code/src/core/assignment.py

@@ -58,7 +58,7 @@ class Source:
 
             for variable in self.variables:
                 candidates = [assignment[variable.name]] if variable.name in assignment else groups
-                domain = list([g for g in candidates if g.dtype in variable.types])
+                domain = list([g for g in candidates if any(st in variable.types for st in g.vector_types)])
                 if len(domain) == 0:
                     return variable.name in assignment, []
                 problem.addVariable(variable.name, domain)

### 2016-05-30 Code/src/core/constraint.py

@@ -130,7 +130,7 @@ class Aggregate(Constraint):
         self._orientation = orientation
         self._operation = operation
         self.min_size = 2
-        self.min_vectors = 3 if operation == Operation.PRODUCT else 2
+        self.min_vectors = 3 if operation == Operation.PRODUCT or operation == Operation.SUM else 2
         size = Group.columns if orientation == Orientation.VERTICAL else Group.rows
         or_string = "col" if orientation == Orientation.VERTICAL else "row"
         op_string = operation.name
@@ -273,7 +273,7 @@ class ConditionalAggregate(Constraint):
         source = ConstraintSource(variables, all_diff, {all_diff.x.name: "OK"})
         filters = [SameLength([self.o_key, self.result]), SameLength([self.f_key, self.values]),
                    SameTable([self.f_key, self.values]), Not(SameTable([self.f_key, self.o_key])),
-                   #SameTable([self.o_key, self.result]), #TODO think about this
+                   SameTable([self.o_key, self.result]), #TODO think about this
                    NotPartial([self.o_key]), SameType([self.f_key, self.o_key]),
                    SameOrientation([self.o_key, self.result]), SameOrientation([self.f_key, self.values])]
         p_format = "{R} = " + name.upper() + "IF({FK}={OK}, {V})"

### 2016-05-30 Code/src/core/group.py

@@ -8,6 +8,19 @@ class GType(Enum):
     float = 1
     string = 2
 
+    @staticmethod
+    def super(gtype):
+        if gtype == GType.int:
+            return GType.float
+        else:
+            return gtype
+
+    @staticmethod
+    def max(gtype_set):
+        if any([gt == GType.string for gt in gtype_set]) and not all([gt == GType.string for gt in gtype_set]):
+            raise Exception("Inconsistent types")
+        return GType(max([gt.value for gt in gtype_set]))
+
 
 # --- Orientation ---
 
@@ -118,11 +131,12 @@ class Table:
 # --- Group ---
 
 class Group:
-    def __init__(self, table, bounds, row, data, dtype):
+    def __init__(self, table, bounds, row, data, gtype_set):
         self._table = table
         self._bounds = bounds
         self._row = row
-        self._dtype = dtype
+        self._vector_types = gtype_set
+        self._dtype = GType.max(gtype_set)
         self._data = data
         from core.constraint import blank_filter
         self._is_partial = not numpy.all(numpy.vectorize(blank_filter(self._data)[1])(self._data))
@@ -147,6 +161,10 @@ class Group:
     @property
     def dtype(self):
         return self._dtype
+
+    @property
+    def vector_types(self):
+        return self._vector_types
 
     @property
     def data(self):
@@ -200,8 +218,11 @@ class Group:
     def subgroup(self, bounds):
         if bounds not in self._subgroups:
             sub_bounds = Bounds([1, self.rows(), 1, self.columns()]).combine(bounds)
+            r1, r2, c1, c2 = sub_bounds.bounds
             sub_data = sub_bounds.subset(self.data)
-            group = Group(self.table, self.bounds.combine(bounds), self.row, sub_data, self.dtype)
+            indices = [r1, r2] if self.row else [c1, c2]
+            vector_types = [self._vector_types[i - 1] for i in range(indices[0], indices[1] + 1)]
+            group = Group(self.table, self.bounds.combine(bounds), self.row, sub_data, vector_types)
             self._subgroups[bounds] = group
             return group
         return self._subgroups[bounds]

### 2016-05-30 Code/src/parse/parser.py

@@ -141,9 +141,13 @@ def create_group(bounds_list, table: Table):
 
     data = bounds.subset(table.data)
     types = np.vectorize(detect_type)(data)
-    g_type = infer_type(types.flatten())
+    if row:
+        gtype_set = [infer_type(row) for row in types]
+    else:
+        gtype_set = [infer_type(col) for col in types.T]
+    g_type = GType.max(gtype_set)
     cast_f = np.vectorize(lambda t, v: cast(g_type, t, v), otypes=[np.object if g_type is GType.string else np.float])
-    return Group(table, bounds, row, cast_f(types, data), g_type)
+    return Group(table, bounds, row, cast_f(types, data), gtype_set)
 
 
 def detect_tables(type_data):
@@ -212,7 +216,7 @@ def detect_groups(type_data, tables):
         start = 0
         for col in range(cols):
             if not same[col]:
-                groups.append(create_group([":", start + 1, col + 1, ":"], table))
+                groups.append(create_group([":", start + 1, col + 1], table))
                 start = col + 1
 
     for t in tables:

### 2016-05-30 Code/src/engine/internal.py
Speedup precentual difference

@@ -305,9 +305,10 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             def is_diff(r, o1, o2):
                 if equal_v(o2, 0).any():
                     return False
-                calculated = (o1 - o2) / o2
-                equal_t = numpy.vectorize(lambda x, y: equal(x, y, True))
-                return equal_t(r, calculated).all()
+                for i in range(0, len(r)):
+                    if not equal(r[i], (o1[i] - o2[i]) / o2[i]):
+                        return False
+                return True
 
             keys = [c.result, c.first, c.second]
             return self._generate_test_vectors(assignments, keys, is_diff)

### 2016-05-30 Code/src/engine/internal.py
Minor speedup perc. diff

@@ -303,10 +303,8 @@ class InternalSolvingStrategy(DictSolvingStrategy):
 
         def percent_diff(c: PercentualDiff, assignments, solutions):
             def is_diff(r, o1, o2):
-                if equal_v(o2, 0).any():
-                    return False
                 for i in range(0, len(r)):
-                    if not equal(r[i], (o1[i] - o2[i]) / o2[i]):
+                    if o2[i] == 0 or not equal(r[i], (o1[i] - o2[i]) / o2[i]):
                         return False
                 return True
 

### 2016-05-30 Paper/ECAI/tabular.tex
Small fix evaluation
iteration of abstract and intro

@@
 However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
-In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
+In this work, we investigate the automatic discovery of constraints (functions and relations) in raw tabular data.
 We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
@@
 Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
-Constraint learning techniques are used to identify predicates (constraints) that hold between blocks of rows and columns of the tables.
+Constraint learning techniques are used to identify the predicates that hold between blocks of rows and columns of the tables.
 %New types of constraints can be easily added to the system by specifying them in a declarative language.
@@
 %This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
-Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big spreadsheets.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain some mistakes in formulae.
+Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
 
@@
 Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
-This provides a flavor of how this problem is different from standard data mining settings, where the data is just in rows and variables are in columns.
+This provides a flavor of how this problem is different from standard mining and learning settings, where the data is just in rows and variables are in columns.
 Here everything is mixed.
@@
 %, since people tend to use spreadsheets for financial and accounting computations.
-One can see that understanding and learning constraints in tabular data is hence a new and challenging problem for machine learning.
+Hence understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
 
 % Moved up
-The question that we answer in this paper is: is it possible to discover or reconstruct constraints (relations, functions) in flat tabular spreadsheet data?
+The question that we answer in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
 To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
-It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, as it reasons on that raw tabular data directly with no example constraint instantiations given.
+It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
 %, and discovers column- and row-wise constraints.
@@
 Our approach contrasts with these in that it focuses
-on learning both column- and row-constraints, as does Modelseeker~\cite{modelseeker} which is restricted to traditional constraint satisfaction problems (CSPs).
-It contrasts with Flashfill in that it can learn from numeric data, too, as well as general constraints in addition to functions.
+on learning both column- and row-constraints on integer, floating point as well as textual data.
 
@@
 Not all relevant constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
-Constraints that are currently outside of the scope of this system, e.g. generic nested mathematical or logical formulas, are ignored.
+Constraints that are currently outside of the scope of this system, e.g. generic nested mathematical or logical formulas and date-related constraints, are ignored.
 

### 2016-05-30 Code/src/engine/internal.py
Further speed imporvements (+ small bug fix)

@@ -290,13 +290,19 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             keys = [c.result, c.first, c.second]
 
             def is_product(r, o1, o2):
-                return numpy.vectorize(equal)(r, numpy.vectorize(Operation.PRODUCT.func)(o1, o2)).all()
+                for i in range(0, len(r)):
+                    if not equal(r[i], Operation.PRODUCT.func(o1[i], o2[i])):
+                        return False
+                return True
 
             return self._generate_test_vectors(assignments, keys, is_product, lambda r, o1, o2: ordered(o1, o2))
 
         def diff(c: Diff, assignments, solutions):
             def is_diff(r, o1, o2):
-                return numpy.vectorize(equal)(r, o1 - o2).all()
+                for i in range(0, len(r)):
+                    if not equal(r[i], o1[i] - o2[i]):
+                        return False
+                return True
 
             keys = [c.result, c.first, c.second]
             return self._generate_test_vectors(assignments, keys, is_diff, lambda r, _, o2: o2 < r)
@@ -304,7 +310,7 @@ class InternalSolvingStrategy(DictSolvingStrategy):
         def percent_diff(c: PercentualDiff, assignments, solutions):
             def is_diff(r, o1, o2):
                 for i in range(0, len(r)):
-                    if o2[i] == 0 or not equal(r[i], (o1[i] - o2[i]) / o2[i]):
+                    if o2[i] == 0 or not equal(r[i], (o1[i] - o2[i]) / o2[i], True):
                         return False
                 return True
 
@@ -352,7 +358,12 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             return solutions
 
         def equality(c: Equal, assignments, _):
-            test = lambda x, y: equal_v(x, y).all()
+            def test(x, y):
+                for i in range(0, len(x)):
+                    if not equal(x[i], y[i]):
+                        return False
+                return True
+
             return self._generate_test_vectors(assignments, [c.first, c.second], test, ordered)
 
         def equal_group(c: EqualGroup, assignments, solutions: Solutions):

### 2016-05-30 Paper/ECAI/tabular.tex
formalization and first part of method

@@
       &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}\\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^1, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^1 \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^1 , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^1 \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^1 \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^* \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^* , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^* \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^* \\
 %
@@
 	  % Table 4
-      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^2 \\
+      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^* \\
 %
@@
     \end{align*}}
-  \vspace{-10pt}
-  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
-  $^1$(Spurious constraints) $^2$(Not present in original source)}
+  \vspace{-18pt}
+  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).\\
+  $^*$: Not present in original document}
   \label{fig:sol_example}
@@
 \section{Formalization}\label{sec:formalization}
-Our goal is to automatically discover constraints (functions and relations) between the rows and the columns of tables in a spreadsheet.
+Our goal is to automatically discover constraints between the rows and the columns of tables in a spreadsheet.
 This is applicable not just to data from spreadsheets, but to any data in tabular form, hence the name.
@@
 \begin{itemize}
-\item \textit{type}: a block is type-consistent, so it has a type
+\item \textit{type}: a block is type-consistent, so it has one type
 \item \textit{table}: the table that the block belongs to
@@
   Overview of constraint templates implemented in \sname.
-  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows the version for sum but \sname also supports max, min, average, product and count.
-  Subgroups in normal font have to be vectors, whereas subgroups in \textbf{bold} may contain more than one vector.
+  Subgroups in normal font have to be vectors, whereas subgroups in \textbf{bold} may contain more than one vector;
+  $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
   Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.
-  $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
-  {\small\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
+  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows $SUM$ but \sname also supports $MAX$, $MIN$, $AVERAGE$, $PRODUCT$ and $COUNT$.
+  %{\small\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
 }
@@
   \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
-  For example, $B_r = [12, 7]$ and $\mathbf{\sg_x}$ consists of columns: $[5, 15]$, $[7, -8]$.
+  %For example, $B_r = [12, 7]$ and $\mathbf{\sg_x}$ consists of columns: $[5, 15]$, $[7, -8]$.
 \end{itemize}
+$SUM$ is an aggregate constraint, and a similar constraint is available for aggregate operators $MIN$, $MAX$, $AVERAGE$, $PRODUCT$ and $COUNT$.
+The \textit{conditional} aggregate variant $SUMIF$ only ranges over those cells that satisfy a condition on a related cell in that table, e.g.
+$\range{T_1}{\rangeall}{10}$ and the data in $T_5$ for matching salesperson names.
 \end{example}
@@
 We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
-Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming and logical relational learning, we would define the predicate~$s$ using a Prolog like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
+Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would define the predicate~$s$ using a Prolog like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
 Under this interpretation, the term constraint and constraint template can be used interchangeably.
@@
 
-Blocks and tables are expected to be given, in Section~\ref{sec:approach} we will discuss how these can be extracted from a spreadsheet.
+The input is a set of blocks, and in Section~\ref{sec:approach} we will discuss how these can be extracted from a spreadsheet.
 Figure~\ref{fig:sol_example} shows the solution to the tabular constraint learning problem when applied on the blocks of Figure~\ref{fig:main_example} and constraint templates listed in Table~\ref{table:constraints}.
@@
 For example for the product constraint, $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecprod{\sg_r}{\sg_2}{\sg_1}$ so one can be considered redundant to the other.
-
-Two different constraints may be logically equivalent due to their nature, e.g. for addition/subtraction and product/division: $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecdiv{\sg_1}{\sg_r}{\sg_2}$.
-
-When the data has rows or columns containing exactly the same data, then any constraint with such a vector in its argument assignment will also have an argument assignment with the other equivalent vectors.
+Two different constraints can also be logically equivalent, due to their nature e.g. for addition/subtraction and product/division: $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecdiv{\sg_1}{\sg_r}{\sg_2}$.
+
+Finally, when the data has rows or columns that contain \textit{exactly} the same values, then any constraint with such a vector in its argument assignment will also have an argument assignment with the other equivalent vectors.
 
 Dealing with redundancy is often application-dependent.
-Therefore, in Section~\ref{sec:approach} we will first explain our generic method for finding all constraints, before describing some optimizations that avoid obvious equivalences (Section~\ref{sec:opts}).
+Therefore, in Section~\ref{sec:approach} we first explain our generic method for finding all constraints, before describing some optimizations that avoid obvious equivalences (Section~\ref{sec:opts}).
 Later, we will investigate the impact of redundancy in the experiments.
@@
 As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier.
-For the above example, each of the vectors of $B_6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g.
+For the above example and $(B_5, B_3)$, each of the vectors of $B_5$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks of $B_3$ for the right-hand side and verify the signature (e.g.
 at least size $4$) and test whether for each row the definition is satisfied.
@@
 We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step 3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step 3b, Section~\ref{sec:algo:subgr}).
-In Section~\ref{sec:opts} we describe a number of optimizations designed to speed up steps the algorithm.
+In Section~\ref{sec:opts} we describe a number of optimizations designed to speed up the algorithm.
 
@@
 
-To find row-blocks, each row is treated as a vector and must be type-consistent; similarly for column-blocks.
+To find row-blocks, each row is treated as a vector and must be type-consistent; similarly for columns and column-blocks.
 Then, adjacent vectors that are of the same type are merged to obtain the maximally type-consistent superblocks.

### 2016-05-30 tacle/main.py
added core table selection files and flask

@@ -1,9 +1,9 @@
-from flask import Flask
+from flask import Flask, render_template
 app = Flask(__name__)
 
 @app.route("/")
 def hello():
-    return "Hello World!"
+    return render_template('table.html')
 
 if __name__ == "__main__":
     app.run()

### 2016-05-30 tacle/run.sh

@@ -1,0 +1,4 @@
+#flask run --host=0.0.0.0 -- > to make it visible to the outside world
+export FLASK_DEBUG=1 #set to false once release to the world
+export FLASK_APP=main.py
+flask run

### 2016-05-30 tacle/static/select-compiled.js

@@ -1,0 +1,250 @@
+var Orientation = {
+    row: "Row",
+    col: "Column",
+    none: "None"
+};
+
+function exportTables(tables) {
+    var converted = [];
+    for (var i = 0; i < tables.length; i++) {
+        converted.push({
+            "Name": "T" + (i + 1),
+            "Bounds": tables[i].bounds,
+            "Orientation": Orientation[tables[i].orientation]
+        });
+    }
+    return JSON.stringify({ "Tables": converted });
+}
+
+var textFile = null;
+
+function makeTextFile(text) {
+    var data = new Blob([text], { type: 'text/plain' });
+
+    // If we are replacing a previously generated file we need to
+    // manually revoke the object URL to avoid memory leaks.
+    if (textFile !== null) {
+        window.URL.revokeObjectURL(textFile);
+    }
+
+    textFile = window.URL.createObjectURL(data);
+
+    // returns a URL you can use as a href
+    return textFile;
+}
+
+var App = React.createClass({
+    displayName: "App",
+
+    getInitialState: function () {
+        return { data: [], start: null, end: null, orientation: "none", tables: [] };
+    },
+    update: function (state, callback) {
+        this.setState(state, callback);
+    },
+    setOrientation: function (orientation) {
+        this.setState({ orientation: orientation });
+    },
+    handleInput: function (event) {
+        this.setState({ rawData: event.target.value });
+    },
+    import: function (event) {
+        var file = event.target.files[0];
+        var textType = /text.*/;
+        if (file.type.match(textType)) {
+            var reader = new FileReader();
+
+            reader.onload = function () {
+                var result = Papa.parse(reader.result);
+                this.setState({ data: result.data });
+            }.bind(this);
+
+            reader.readAsText(file);
+        }
+    },
+    render: function () {
+        var or = function (or) {
+            return function () {
+                this.setOrientation(or);
+            }.bind(this);
+        }.bind(this);
+        return React.createElement(
+            "div",
+            { className: "full" },
+            React.createElement(
+                "div",
+                { className: "menu" },
+                React.createElement(ClearButton, { state: this.state, setState: this.update }),
+                Object.keys(Orientation).map(function (key) {
+                    return React.createElement(
+                        "label",
+                        { key: key },
+                        React.createElement("input", { type: "radio", value: key, name: "or", onChange: or(key),
+                            checked: this.state.orientation == key }),
+                        Orientation[key]
+                    );
[…]

### 2016-05-30 Paper/ECAI/tabular.tex
more of approach and applications

@@
       \State $C \gets \emptyset$ %\Comment{The set of constraints}
-      \ForAll{$s~\mathbf{in}~\temps$}
+      \ForAll{$s~\mathbf{in}~\temps$} \label{algo:tcl:for}
         %\State $n \gets$ number of arguments of template $s$
-        \State $A \gets \generategroups(s, \blocks)$
+        \State $A \gets \generategroups(s, \blocks)$  \label{algo:tcl:super}
         \ForAll{$(B_1, \dots, B_n) \in A$}
-        \State $A' \gets \findassignment(s, (B_1,\dots,B_n))$
+        \State{$A' \gets \findassignment(s, (B_1,\dots,B_n))$}
         \ForAll{$(B_1',\dots,B_n') \in A'$}
@@
 \begin{table}[t]
-  \caption{Translation of signature requirements to superblock constraints.
-The requirements that do not need to be relaxed are: $\plength(B) = x$, $\por(B) = x$, $\ptable(B) = x$ and $\psize(B) \geq x$.}
+  \caption{Translation of signature requirements to superblock constraints. The following need not be relaxed: $\plength(B) = x$, $\por(B) = x$, $\ptable(B) = x$ and $\psize(B) \geq x$.}
   \label{tbl:translation}
@@
 
-  Consider the constraint template $\ecsumc{\sg_r}{\mathbf{\sg_x}}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$\sg_r$ and~$\mathbf{\sg_x}$.
+  Consider the constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$\sg_r$ and~$\mathbf{\sg_x}$.
   Given maximal blocks~\blocks, the domain of these variables are $D(V_r) = D(V_x) = \{ 1, \dots, |\blocks| \}$.
-  Finally, the signature of $\ecsumc{\sg_r}{\mathbf{\sg_x}}$ (see Table~\ref{table:constraints}) will be translated into constraints:
+  Finally, the signature of $\ecsumr{\sg_r}{\mathbf{\sg_x}}$ (see Table~\ref{table:constraints}) will be translated into constraints:
   \begin{align*}
-    & \numeric(V_r) \land \numeric(V_x) \land \prows(V_x) \geq 2 \\
-    & (\por(V_x) = \mathit{row}) \Rightarrow (\pcols(V_x) \geq \plength(V_r)) \\
-    & (\por(V_x) = \mathit{column}) \Rightarrow (\pcols(V_x) = \plength(V_r))
+    & \numeric(V_r) \land \numeric(V_x) \land \pcols(V_x) \geq 2 \\
+    & (\por(V_x) = \mathit{column}) \Rightarrow (\prows(V_x) \geq \plength(V_r)) \\
+    & (\por(V_x) = \mathit{row}) \Rightarrow (\prows(V_x) = \plength(V_r))
   \end{align*}
@@
   An example superblock assignment from Figure~\ref{fig:main_example} is $(B_r, \mathbf{\sg_x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
-  Note that this assignment does not satisfy the signature yet, $\sg_r$ contains more than one vector and $\psize(\mathbf{\sg_x}) \neq \plength(B_r)$.
-  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that \textit{do} satisfy the signature and test whether they satisfy the definition:
-  $\forall i, \sbl{r}[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$.
+  Note that this assignment does not satisfy the signature yet because $\sg_r$ contains more than one vector.
+  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that do satisfy the signature and test whether they satisfy the definition:
+  $\forall i, \sbl{r}[i] = \sum_{j = 1}^{length(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$.
   In Figure~\ref{fig:main_example} there is exactly one such subassignment: $(\range{T_1}{\rangeall}{7}, \range{T_1}{\rangeall}{\rangeto{3}{6}})$.
@@
 Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
-This is inefficient, as already the blocks may not satisfy the signature or the first elements in the data may not satisfy the definition.
-
-Hence, a simple generate-and-test will usually suffice.
+This is inefficient, as already such blocks may not satisfy the signature or the first elements in the data may not satisfy the definition.
+
+On the other hand, a simple generate-and-test will usually suffice.
 %However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{figures/constraint_dependency2.pdf}
-  \caption{Dependency graph; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (the \CSignature of~$s_2$ includes $s_1$).
-    The graph is extracted from the Signature column in Table~\ref{table:constraints}.
+  \includegraphics[width=0.65\linewidth]{figures/constraint_dependency2.pdf}
+  \caption{Dependency graph of Table~\ref{table:constraints}; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (its signature includes $s_1$).
   }
@@
 Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
-Concretely, the input templates~$\temps$ (see Algorithm~\ref{algo:tcl}) are reordered to agree with the partial ordering imposed by~\dependencies (e.g.
+Concretely, in Algorithm~\ref{algo:tcl} line \ref{algo:tcl:for} the input templates~$\temps$ are handled following an ordering that agrees with the partial ordering imposed by~\dependencies (e.g.
 \textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
-Considering Algorithm~\ref{algo:tcl}, this step can be seen as preprocessing the input~\temps.
-Learning in this order allows templates to use constraints they depend on, as these constraints will have already been learned.
-This is exploited in the \generategroups step to speedup the algorithm by reducing the search space.
+When using this ordering, templates later in the ordering can use the learned constraints of the templates they depend on.
+This is made possible by adding, on line \ref{algo:tcl:super}, the set of learned constraints $C$ to the argument of \generategroups.
+
+This is used to speed up the \generategroups step.
+For constraint templates where some arguments are also part of a constraint that it depends on, it is not needed to search for superblocks and subblocks from scratch.
+Instead, one can start from all and only the actual argument assignments of the depending constraint, and only search matching superblocks for the remaining arguments.
 
 \begin{example}
-  Consider $\ecfkey{B_{fk}}{B_{pk}}$ whose signature includes $\ecalldiff{B_{pk}}$ and assume that~$C$ are the constraints found for \textit{ALLDIFFERENT}.
-  Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
-  $P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed.
+  Consider $\ecfkey{B_{fk}}{B_{pk}}$, which states that every value in $B_{fk}$ also exists in $B_{pk}$; its signature includes $\ecalldiff{B_{pk}}$.
+There are 16 \textit{ALLDIFFERENT} constraints to be found in Figure~\ref{fig:main_example}, hence, the \generategroups only needs to check which superblocks for $B_{pk}$ are compatible (different table, same type) with these 16 assignments to $B_{fk}$.
 \end{example}
 
-The \generategroups implementation now takes into account previously found constraints to obtain partial assignments~$P$ for a template.
-Instead of generating one CSP to find all assignments, a CSP is generated for every~$p \in P$ which finds all assignments completing~$p$.
-The procedure to generate CSPs remains the same, except that the partial assignment is used to fix the domains of the dependent variables.
-
-This approach has the additional benefit of ensuring that the (sub)blocks assigned to the dependent arguments already satisfy both the signature and definition of the original template.
-Therefore, the implementation of $\findassignment$ does not have to take these blocks into account.
+%Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
+%$P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed.
+Instead of generating one CSP to find all assignments, a CSP is generated for every known assignment of the depending constraint (and variables involved), which then searches for all assignments completing this partial assignment.
+The procedure to generate CSPs remains the same otherwise. %, except that the known subassignment(s) are added as extra blocks that can only be assigned to their corresponding variable.
+
+%This approach has the additional benefit of ensuring that the (sub)blocks assigned to the dependent arguments already satisfy both the signature and definition of the template it depends on.
+%Therefore, the implementation of $\findassignment$ does not have to take these blocks into account.
 
@@
 \subsubsection{Redundancy}
-We consider two types of redundancy that we aim to eliminate during the search.
+We consider two types of redundancy that we aim to eliminate during search.
 As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
@@
 In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
-Moreover, \sname currently contains only product (\ecprod{B_R'}{B_1'}{B_2'}) and not division (\eccalc{B_1'}{B_R' / B_2'}) as well as difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
-These choices are based on the prevalence of product and difference in financial spreadsheet data, however, division and sum as well as other constraints could easily be added.
-
-A last type of redundancy we aim to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
+Moreover, among the semantically equivalent constraints product (\ecprod{B_R'}{B_1'}{B_2'}) and division (\eccalc{B_1'}{B_R' / B_2'}) we only added division, as well as only difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
+However, division and sum could easily be added, as could it be added explicitly in post-processing based on the matching product and difference constraints.
+
+A last type of redundancy we chose to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
 For example, consider the constraint \ecsumc{B_r}{B[1{:}n]} where $B[k{:}n]$ consists of only zeros, then \ecsumc{B_r}{B[1{:}j]} will be true for all $k - 1 \leq j \leq n$.
@@
 Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$.
-Should the target constraint be $\eccalc{B_2}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach can find it.
+Should the target constraint be $\eccalc{B_2}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach could find it.
 
 Alternatively, we can output all valid argument assignments instead of only maximal subblocks, e.g.
-$\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$, however, this might produce many redundant constraints and introduce constraints that accidentally cover too little data.
+$\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$, however, this might produce many redundant constraints including constraints that accidentally cover too little data.
 
@@
 We also added three structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
+
+\tias{This would be a good place to say that for aggregates, you do SUBASSIGNMENTS with specialised code...}
 
@@
 
-We explained various optimizations, most notably the exploitation of dependencies.
-Using the dependencies to find constraints incrementally cuts down large parts of the search trees, especially for constraints that have many arguments.
-Essentially, one obtains a more specialized search without much effort.
+%We explained various optimizations, most notably the exploitation of dependencies.
+%Using the dependencies to find constraints incrementally cuts down large parts of the search trees, especially for constraints that have many arguments.
+%Essentially, one obtains a more specialized search without much effort.
 
@@
 Autocompletion can be seen as using knowledge derived from a snapshot of spreadsheet data to predict new values before they are written by the user.
-The snapshot~($D_0$) is used to learn constraints and these are combined with new data~($\Delta D$) that the user has added.
-Using~$\Delta D$ we can keep track of which vectors have changed and which vectors are now incomplete.
-A \textit{functional} constraint corresponds to a spreadsheet formula.
-Such a constraint, e.g. $\ecsumc{B_r}{\mathbf{B_x}}$, is able to compute a result~($B_r$) based on its input~($\mathbf{B_x}$).
-If all the input vectors / blocks have been filled and the result is incomplete, the constraint can predict the result value(s).
-
-Let us illustrate this on one of the spreadsheets from the gathered dataset.
+A snapshot~($D_0$) is used to learn constraints and these are combined with new data~($\Delta D$) that the user has added.
+Using~$\Delta D$ we can keep track of which vectors have changed and which vectors are now \textit{incomplete}.
+\textit{Functional} constraints, such as $\ecsumc{B_r}{\mathbf{B_x}}$, are able to compute a result~($B_r$) based on its input~($\mathbf{B_x}$).
+If all vectors in the input blocks have changed and the output vector is incomplete, the constraint can compute the resulting value(s).
+
 Consider Figure \ref{fig:autocompletion_example} containing two tables.
-By learning constraints on the tables excluding the last row, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
-The new data $\Delta D$ consists of the new value \textit{West} in the last row.
+Learning on a snapshot that does not contain line 17, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
+The new data $\Delta D$ consists of the new value \textit{West} on line 17.
 Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
@@
 
-In case multiple constraints predict different values for the same cell, the algorithm will have to either prioritize constraints based on their specificity or a heuristic, not make a prediction, or present the choice to the user.
+%In case multiple constraints predict different values for the same cell, the algorithm will have to either prioritize constraints based on their specificity or a heuristic, not make a prediction, or present the choice to the user.
 
@@
 The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
-However, we look at \textit{conflict} constraints, i.e. constraints that were learned on the snapshot data but do not agree with the newly added data~$\Delta D$.
+Here, we look at \textit{conflict} constraints, i.e. constraints that were learned on the snapshot data but do not agree with the newly added data~$\Delta D$.
 Such conflict constraints can indicate errors in the input.
 However, such conflicts can be wrongly caused by spurious constraints.
-Therefore, conflicts should only be detected if the snapshot data is large enough.
-An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.
+Therefore, conflicts should perhaps only be detected if the snapshot data is large enough.
+%An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.
 
 The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
-Consider, for example, a use case concerning crime statistics provided by the FBI.
-An extract of the spreadsheet is depicted in Figure~\ref{fig:fbi}.
-When run on this sheet, our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
-A possible approach to detect such missing or wrong values is to look at a sample of the table that excludes the erroneous row, e.g.
-trying to leave out each row, and compare the constraints learned on this sample with the constraints learned by adding the row containing the mistake.
-The sample needs to be large enough such that we have enough confidence that the constraints are correct.
+Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
+Our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
+A possible approach to detect such missing or wrong values is to look at a sample of the table, e.g.
+trying to leave out each row, and compare the constraints learned on this sample with the constraints learned on the whole table.
+%The sample needs to be large enough such that we have enough confidence that the constraints are correct.
 For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
-If we define the sample to be the snapshot data and the added row to be the additional input~$\Delta D$, this task can be formulated as online error detection and dealt with in the same way.
+If we define the sample to be the snapshot data and the other row(s) to be the additional input~$\Delta D$, this task can be formulated as online error detection and dealt with in the same way.
 %Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.

### 2016-05-31 Code/src/accuracy.py
Added size measuring code

@@ -6,6 +6,7 @@ from pandas import json
 import print_truth
 import workflow
 from core.constraint import *
+from core.group import Bounds
 from engine.util import local
 
 exercises = [
@@ -206,5 +207,18 @@ def calc_ratio(counters: List[CategoryCounter], ex_f1, ex_f2, per_file):
         return s1 / s2, s1, s2
 
 
+def measure_size():
+    for name in files:
+        groups_file = local("data/groups/{}.txt".format(name))
+        with open(groups_file) as file:
+            json_data = json.load(file)
+            table_count = len(json_data["Tables"])
+            cell_count = 0
+            for table in json_data["Tables"]:
+                bounds = Bounds(table["Bounds"])
+                cell_count += bounds.columns() * bounds.rows()
+            print("{}, {}, {}".format(name, table_count, cell_count))
+
+
 if __name__ == '__main__':
-    main()
+    measure_size()

### 2016-05-31 Code/src/workflow.py

@@ -30,6 +30,7 @@ constraint_list = [
 constraint_list += Aggregate.instances()
 constraint_list += ConditionalAggregate.instances()
 
+
 def order_constraints(constraints: List[Constraint]):
     ordered = []
     found = set()

### 2016-05-31 Paper/ECAI/tabular.tex

@@
 Our goal is to automatically identify constraints in a spreadsheet.
-We have presented and evaluated our approach, implemented as the system \sname, that is able to learn a large set of constraints from CSV data.
-Moreover, the amount of redundant constraints found by the system is limited despite the lack of more sophisticated post-processing steps.
-
-For the future development of the system there are multiple directions. Integrating the system in an interactive setting would offer users the possibility to easily receive and provide feedback.
-Moreover, the system can be adapted to deal with additional types and produce less redundant constraints through the use of (heuristic) filtering or post-processing.
-
-Currently, the system only learns single constraints and extending our approach to nested constraints would allow more expressive concepts to be learned.
-Also, \sname only works on complete rows and columns, allowing partial vectors is an important future direction for developing applications based on the system.
-%This shift would bring the approach in line with programming by example.
-
-Aside from finding errors, the system can be extended to deal with noise and the ability to learn soft constraints.
-Soft constraints are (potentially weighted) constraints that hold only on some of the data.
-This would extend the approach to new application domains as well as provide more native error detection.
+We have presented and evaluated our approach, implemented as the system \sname, that is able to learn many different constraints in tabular data.
+The resulting method has high accuracy, produces limited number of redundant constraints and is sufficiently efficient for normal as well as interactive use.
+
+The approach is designed to find constraints that hold over entire columns and rows.
+In future work we plan to extend this to learn arbitrary nested constraints (e.g.
+$B_r = (B_1 + B_2) / B_3$), as well as constraints over only a subset of the vectors.
+This may give rise to more redundant and spurious constraints, which was not problematic up to this point.
+
+One other line of work, which can also mitigate redundant and spurious constraints, is to integrate the approach an interactive setting, where users can receive and provide feedback; more similar to an active learning setting.
+Other ways to handle redundant constraints include the (heuristic) filtering or post-processing of the constraint set.
+
+A final direction is that the current approach assumes noise-free data (but see the part on error correction); in inductive logic programming there has been much work on how to handle noisy data, while in constraint satisfaction a popular alternative is to consider \textit{soft} constraints that can be violated (at a cost).
+This could extend the approach to new application domains, beyond traditional spreadsheets.
 

### 2016-06-07 Paper/ECAI/tabular.tex
Updated LaTeX

@@
     Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.
-    Most tables only contain type-consistent columns, however, table~$T_2$ also contains type-consistent row.
+    Most tables only contain type-consistent columns, however, table~$T_2$ also contains type-consistent rows.
+    Due to the location of the header of~$T_2$ we can infer that it consists of row data.
     This example was created by combining several Excel sheets from several exercise sessions.
@@
 %
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}\\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^* , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^* \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^*, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
       &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^* \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^* , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^* \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^* \\
 %
@@
 %
-      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
-%
-      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{5}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{2}} \\
+      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}} \\
+%
+      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}} \\
 %
@@
 	  % Table 4
-      &~\eclookup{\range{T_{4}}{\rangeall}{2}}{\range{T_{4}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^* \\
-%
-      &~\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
-%
-	  % Table 6
-      &~\ecsumc{\range{T_{6}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-      &~\range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
+      &~\ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+      &~\range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
+%
+    % Table 5
+      &~\eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^* \\
+%
+      &~\eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}
     \end{align*}}
@@
   \includegraphics[width=1\linewidth]{figures/tabletool.png}
-  \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding (headerless) tables, an explicit orientation can also be provided.}
+  \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding to (headerless) tables.
+  An explicit orientation can also be provided.}
   \label{fig:visual_tool}
@@
 In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
-Moreover, among the semantically equivalent constraints product (\ecprod{B_R'}{B_1'}{B_2'}) and division (\eccalc{B_1'}{B_R' / B_2'}) we only added division, as well as only difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
+Moreover, among the semantically equivalent constraints product (\ecprod{B_R'}{B_1'}{B_2'}) and division (\eccalc{B_1'}{B_R' / B_2'}) we only added product, as well as only difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
 However, division and sum could easily be added, as could it be added explicitly in post-processing based on the matching product and difference constraints.
@@
 
-\tias{This would be a good place to say that for aggregates, you do SUBASSIGNMENTS with specialised code...}
+For \textit{FOREIGNKEY}, \textit{LOOKUP}, aggregates and \textit{PROJECT} we use custom, optimized methods to find subassignments.
+Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
+While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
 
@@
 \section{Evaluation}\label{sec:evaluation}
-\sergey{we need to add a summary of the dataset, avg number of constraints, cells, rows, columns}
 In this section we experimentally validate our approach.
-We studied three main questions concerning the accuracy (\textit{Q1}), precision (\textit{Q2}) and speed (\textit{Q3}) of our algorithm.
+We studied three main questions concerning the recall (\textit{Q1}), precision (\textit{Q2}) and speed (\textit{Q3}) of our algorithm.
 
 The implementation is illustrated using a case study on the spreadsheet corresponding to the previously introduced example (Figure~\ref{fig:main_example}).
-In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 30 (\samuel{check number}) spreadsheets that we assembled from various sources.
-
-There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel \samuel{Including running example?}, spreadsheets from tutorials online and publicly available spreadsheets such as crime statistics or financial reports that demonstrate more real world usage\footnote{(attached to to the submitted paper)}.
-
-In this section we focus on \textit{functional} constraints that could be used in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
-All experiments were run using Python 3.5.1 (custom Anaconda environment) on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
+In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 34 spreadsheets that we assembled from various sources.
+
+\subsection{Method}
+\label{sec:evalualtion:method}
+In order to obtain a benchmark of spreadsheets we looked at three sources:
+1)~spreadsheets from an exercise session for teaching Excel; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports\footnote{(attached to to the submitted paper)}.
+All spreadsheets were converted to CSV and the table definitions are fixed.
+Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
+
+For every spreadsheet we specified a ground-truth of constraints, i.e. the \textbf{intended} constraints that are expected to be (re-)discovered.
+These were determined using the formulas (if present), context and headers in the original spreadsheets.
+We encoded all relevant constraints using our representation (in canonical form).
+Not all relevant constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
+Constraints that are currently outside of the scope of this system are, for example, generic nested mathematical or logical formulas.
+
+We focus on \textit{functional} constraints that could be used in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
+We included detected equalities as relevant constraints (in 5~spreadsheets).
+All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
 
@@
 
-For this example our primary goal of finding all original constraints is achieved, the recall is $1.0$.
-The implementation also returns 6 additional constraints, therefore, our precision is $12/18 = 0.67$, which is rather low compared to other spreadsheets in the benchmark.
+For this example our primary goal of finding all original constraints is achieved, the recall (i.e. $\frac{\text{relevant discovered}}{\text{relevant}}$) is~$1.0$.
+The implementation also returns 6 additional constraints, therefore, our precision (i.e.
+$\frac{\text{relevant discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is rather low compared to, for example, the spreadsheets in the benchmark.
 The low precision can be explained by the many short vectors which increases the chance for constraints to be true by accident.
@@
 \subsection{Benchmark}
-All spreadsheets have been converted to CSV files.
-Definitions of tables are given (in some harder cases with $\mathit{None}$ values, blocks are defined manually). \samuel{chec}
-
-\samuel{Revisit [what does "implemented" mean here? what's the difference between "implemented" and "algorithm", esp given that we are talking experiments? ] }
-\luc{what is not implemented, strange \dots in a paper !!!}
-For every spreadsheet a ground-truth has been provided manually, specifying the \textbf{relevant} constraints that are expected to be (re-)discovered.
-Not all relevant constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
-Constraints that are currently outside of the scope of this system, e.g. generic nested mathematical or logical formulas and date-related constraints, are ignored.
-
-\subsubsection*{Q1. How accurate is \sname?}
-Our implementation achieves a recall of~$0.97$ for relevant constraints, i.e. it finds~90 of~93 relevant constraints in the benchmark suite.
-The recall for supported constraints is $1.0$ (90 of 90).
+Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
+Notice that for the data category the average number of tables and cells is higher while the number of relevant constraints is lower compared to the other categories.
+
+{\setlength{\tabcolsep}{0.33em}
+\begin{table}[]
+  \centering
+  \caption{Summary of properties and results per spreadsheet category}
+  \label{tbl:category_overview}
+  \begin{tabularx}{\linewidth}{lccccccc}
+ & \multicolumn{2}{c}{\textbf{Exercises} (9)} & \multicolumn{2}{c}{\textbf{Tutorials} (21)} & \multicolumn{2}{c}{\textbf{Data} (4)} \\
+ & Overall & Sheet avg & Overall & Sheet avg & Overall & Sheet avg \\ \hline
+    \textbf{Tables} & 19 & 2.11 & 48 & 2.29 & 4 & 1 \\ \hline
+    \textbf{Cells} & 1231 & 137 & 1889 & 90 & 2320 & 580 \\ \hline
+    \textbf{\begin{tabular}[c]{@{}l@{}}Relevant\\[-4pt] Constraints\end{tabular}} & 34 & 3.78 & 52 & 2.48 & 6 & 1.50 \\ \hline \hline
+\textbf{Recall} & 0.85 & 0.83 & 0.88 & 0.87 & 1.00 & 1.00 \\ \hline
+\textbf{\begin{tabular}[c]{@{}l@{}}Recall\\[-4pt] Supported\end{tabular}} & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ \hline
+\textbf{Precision} & 0.97 & 0.98 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
+\textbf{Speed (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
+\end{tabularx}
+\end{table}}
+
+\subsubsection*{Q1. How many relevant constraints are found by \sname?}
+The overall recall achieved by \sname is~$0.88~(81/92)$.
+Similarly, the overall and average recall per category is high, as shown in Table~\ref{tbl:category_overview}.
+The recall for supported constraints is $1.00$, meaning that all supported constraints are found.
 
 \subsubsection*{Q2. How precise is \sname?}
-Our implementation finds 121 constraints across all spreadsheets, 28 of which are not present in the ground-truth.
-Therefore, the calculated precision is \samuel{Check numbers}.
-Notable, the average precision per spreadsheet is \samuel{math}.
-This gap is explained by the large number of additional constraints produced by few spreadsheets \samuel{More info?}.
-
-Examining the additional constraints more closely reveals that only \samuel{x} of the constraints are spurious, i.e. they hold by accident and can be disproved by adding additional data.
-Most constraints are \textit{redundant}, such as duplications or implications of other constraints.
-Many redundant constraints are the result of bad spreadsheet design, i.e. copying data in multiple places, and the overlapping role of difference and the sum aggregate.
-
-We consider precision (i.e. avoiding redundant constraints) a secondary goal, focusing primarily on accuracy.
-The rationale is that the precision can be increased by post-processing the constraints using entailment or heuristics.
+Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
+However, the overall precision per category reveals that the precision is much higher for the exercises and data categories ($0.97$ and $1.00$) and much lower for the tutorials category ($0.70$).
+Moreover, the average precision per spreadsheet is, for all categories, over $0.90$.
+Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
+A qualitative analysis of these spreadsheets shows that their low precision is caused by duplicate data or multiple ways to calculate the same results.
+For these spreadsheets the precision can be increased by post-processing the constraints using entailment (to detect implications) and heuristics (e.g. taking into account the alignment of results).
 
 \subsubsection*{Q3. How fast is \sname?}
-Concerning the speed of the algorithm we also prioritized accuracy when a trade-off had to be made.
-For the \benchsize~spreadsheets in the benchmark our implementation ran in ${\runtotal}s \pm {\runtotalstd}s$.
-The execution times vary widely though between spreadsheets, only four spreadsheets taking more than $0.2s$.
-Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
+The average runtime across all spreadsheets is $0.13s$.
+Across categories the average running times fluctuate in the range between $0.09s$ to $0.22s$.
+While most spreadsheets can be analyzed very quickly, running times can go up to around $1s$ for others.
+
+Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates. \samuel{More details}
 The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
 For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).
-
-\begin{table}
-  \caption{Overview evaluation on benchmark}
-  \centering
-  \begin{tabular}{lll}
-    & \textbf{Total} & \textbf{Spreadsheet (avg)} \\
-    \textbf{Relevant constraints} & $93$ & $2.91$ \\
-    \textbf{Recall} & $96.77\%$ & $94.27\%$ \\
-    \textbf{Precision} & $23.14\%~(28)$ & $8.33\%~(0.88)$ \\
-    \textbf{Speed (s)} & ${\runtotal}s \pm {\runtotalstd}s$ & ${\runfile}s \pm {\runfilestd}s$
-  \end{tabular}
-\end{table}
-\samuel{Expand summary}
-\sergey{would it make sense to specify the size of a spreadsheet in the number of non-zero cells?}
 
@@
 Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=1\linewidth]{figures/comparison.pdf}
+  \caption{Overview of the average properties and performance of spreadsheets within every benchmark category relative to the maximum across all spreadsheets.}
+  \label{fig:comparison}
+\end{figure}
+
+\subsection{Testing the limits}
+On many spreadsheets \sname achieves good results.
+However, factors such as incompatible formatting, bad spreadsheet design and rounding issues can lead to decreased recall, many additional constraints and long running times due to an exponential blow-up of candidate constraints.
+For example, we tested a spreadsheet (``exps\_tias'') with 6~tables, 531~cells and many duplicate columns across tables.
+While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when ignoring equalities.
+This shows the need for a (heuristic) post-processing step to increase precision and further optimizations to decrease running times.
 

### 2016-06-07 Code/src/accuracy.py
Improved Code

@@ -7,6 +7,7 @@ import print_truth
 import workflow
 from core.constraint import *
 from core.group import Bounds
+from draw import Plotter
 from engine.util import local
 
 exercises = [
@@ -51,17 +52,20 @@ real = [
     "financial_result",
     "fbi_offenses_corr",
     "pattern_mining",
+    "exps_tias"
 ]
 
-files = exercises + tutorials + real
+files = ["exps_tias"]  # exercises + tutorials + real
+
+
+def excel_constraints():
+    return Aggregate.instances() + ConditionalAggregate.instances() +\
+        [Series(), Rank(), Lookup(), FuzzyLookup(), RunningTotal(), Product(), Diff(), SumProduct(),
+         ForeignProduct(), Equal(), PercentualDiff(), Projection()]
 
 
 def is_excel_constraint(c: Constraint):
-    return isinstance(c, Aggregate) or isinstance(c, ConditionalAggregate) or isinstance(c, Series) \
-        or isinstance(c, Rank) or isinstance(c, Lookup) or isinstance(c, FuzzyLookup) or isinstance(c, RunningTotal) \
-        or isinstance(c, Product) or isinstance(c, Diff) or isinstance(c, SumProduct) \
-        or isinstance(c, ForeignProduct) or isinstance(c, Equal) or isinstance(c, PercentualDiff) \
-        or isinstance(c, Projection)
+    return c in excel_constraints()
 
 
 constraint_map = {c.name: c for c in workflow.constraint_list}
@@ -87,7 +91,7 @@ def main():
         if not os.path.isfile(csv_file):
             raise Exception("File not found: {}".format(csv_file))
         groups_file = local("data/groups/{}.txt".format(name))
-        silent = True
+        silent = False  # TODO True
         if not os.path.isfile(groups_file):
             with open(groups_file, "w+") as f:
                 print("{\n\t\"Tables\":\n\t\t[\n\n\t\t]\n}", file=f, flush=True)
@@ -207,18 +211,5 @@ def calc_ratio(counters: List[CategoryCounter], ex_f1, ex_f2, per_file):
         return s1 / s2, s1, s2
 
 
-def measure_size():
-    for name in files:
-        groups_file = local("data/groups/{}.txt".format(name))
-        with open(groups_file) as file:
-            json_data = json.load(file)
-            table_count = len(json_data["Tables"])
-            cell_count = 0
-            for table in json_data["Tables"]:
[…]

### 2016-06-07 Code/src/core/constraint.py

@@ -239,7 +239,7 @@ class Lookup(Constraint):
                    SameLength([self.f_key, self.f_value]), SameLength([self.o_key, self.o_value]),
                    SameTable([self.f_key, self.f_value]), SameTable([self.o_key, self.o_value]),
                    SameOrientation([self.f_key, self.f_value]), SameOrientation([self.o_key, self.o_value])]
-        super().__init__("lookup", "{FV} = LOOKUP({FK}, {OK}, {OV})", source, filters)
+        super().__init__("lookup", "{FV} = LOOKUP({FK}, {OK}, {OV})", source, filters, {Equal()})
 
 
 class FuzzyLookup(Constraint):
@@ -255,7 +255,7 @@ class FuzzyLookup(Constraint):
                    SameLength([self.f_key, self.f_value]), SameLength([self.o_key, self.o_value]),
                    SameTable([self.f_key, self.f_value]), SameTable([self.o_key, self.o_value]),
                    SameOrientation([self.f_key, self.f_value]), SameOrientation([self.o_key, self.o_value])]
-        super().__init__("fuzzy-lookup", "{FV} = FUZZY-LOOKUP({FK}, {OK}, {OV})", source, filters)
+        super().__init__("fuzzy-lookup", "{FV} = FUZZY-LOOKUP({FK}, {OK}, {OV})", source, filters, {Equal()})
 
 
 class ConditionalAggregate(Constraint):
@@ -273,11 +273,11 @@ class ConditionalAggregate(Constraint):
         source = ConstraintSource(variables, all_diff, {all_diff.x.name: "OK"})
         filters = [SameLength([self.o_key, self.result]), SameLength([self.f_key, self.values]),
                    SameTable([self.f_key, self.values]), Not(SameTable([self.f_key, self.o_key])),
-                   SameTable([self.o_key, self.result]), #TODO think about this
+                   SameTable([self.o_key, self.result]),  # TODO think about this
                    NotPartial([self.o_key]), SameType([self.f_key, self.o_key]),
                    SameOrientation([self.o_key, self.result]), SameOrientation([self.f_key, self.values])]
         p_format = "{R} = " + name.upper() + "IF({FK}={OK}, {V})"
-        super().__init__("{}-if".format(name.lower()), p_format, source, filters, depends_on={ForeignKey()})
+        super().__init__("{}-if".format(name.lower()), p_format, source, filters, depends_on={Lookup()})
 
     @property
     def operation(self) -> Operation:
[…]

### 2016-06-07 Code/src/draw.py

@@ -2,8 +2,92 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 
+class Plotter:
+    def __init__(self, output=None, labels=None, upper=None):
+        super().__init__()
+        self._categories = []
+        self._output = output
+        self._labels = labels
+        self._upper = upper
+
+    def add_category(self, name, data, color, errors=None, e_color=None):
+        self._categories.append((name, data, color, errors, e_color))
+
+    def plot(self):
+        cat_count = len(self._categories)
+        data_length = len(self._categories[0][1])
+        ind = np.array(range(data_length))
+        width = 0.1
+
+        fig = plt.figure()
+        ax = fig.add_subplot(111)
+        fig.set_size_inches(8, 2)
+        bars = []
+        for i in range(cat_count):
+            _, data, color, e, ec = self._categories[i]
+            if e is not None:
+                bars.append(plt.bar(ind + i * width, data, width=width, color=color, yerr=e, ecolor=ec))
+            else:
+                bars.append(plt.bar(ind + i * width, data, width=width, color=color))
+        plt.xticks(ind + (cat_count - 1) * width)
+        labels = list(label.replace(" ", "\n") for label in self._labels) \
+            if self._labels is not None else range(1, data_length + 1)
+        ax.set_xticklabels(labels)
+        plt.ylabel("Spreadsheet average in percent")
+        plt.yticks([0, 0.25, 0.5, 0.75, 1])
+        ax.set_yticklabels(["0%", "25%", "50%", "75%", "100%"])
[…]

### 2016-06-07 Code/src/engine/internal.py

@@ -2,6 +2,8 @@ import itertools
 
 import math
 
+import sys
+
 from core.constraint import *
 from core.group import Group
 from core.solutions import Solutions
@@ -140,7 +142,8 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                     if not pk_v.overlaps_with(pv_v):
                         pk_dict = dict(zip(pk_v.get_vector(1), pv_v.get_vector(1)))
                         for fk_v, fv_v in itertools.product(fk, fv):
-                            if not any(g1.overlaps_with(g2) for g1 in [pk_v, pv_v] for g2 in [fk_v, fv_v]) \
+                            if not (found_equal(pk_v, fk_v, solutions) and found_equal(pv_v, fv_v, solutions))\
+                                    and not any(g1.overlaps_with(g2) for g1 in [pk_v, pv_v] for g2 in [fk_v, fv_v]) \
                                     and not fk_v.overlaps_with(fv_v) \
                                     and is_lookup(pk_dict, fk_v.get_vector(1), fv_v.get_vector(1)):
                                 result = {c.o_key: pk_v, c.o_value: pv_v, c.f_key: fk_v, c.f_value: fv_v}
@@ -158,7 +161,8 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                         return i - 1
                 return len(collection) - 1
 
-            def is_lookup(ok, ov, fk, fv):
+            def test_equal(ok_v, ov_v, fk_v, fv_v):
+                ok, ov, fk, fv = (vec.get_vector(1) for vec in (ok_v, ov_v, fk_v, fv_v))
                 exact = True
                 for i in range(len(fk)):
                     index = find_fuzzy(fk[i], ok)
@@ -166,9 +170,9 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                         exact = False
                     if index is None or not equal(ov[index], fv[i]):
                         return False
-                return not exact
+                return not exact and not found_equal(ok_v, fk_v, solutions) and not found_equal(ov_v, fv_v, solutions)
 
-            return self._generate_test_vectors(assignments, keys, is_lookup)
+            return self._generate_test_vectors(assignments, keys, lambda *args: True, test_equal)
 
         def conditional_aggregate(c: ConditionalAggregate, assignments, solutions: Solutions):
             keys = [c.o_key, c.result, c.f_key, c.values]
[…]

### 2016-06-07 Code/src/experiment.py

@@ -1,0 +1,165 @@
+import os
+
+import time
+
+from pandas import json
+
+import workflow
+from core.constraint import *
+from core.group import Bounds
+from engine.util import local
+
+
+class ConstraintCounter:
+    def __init__(self, solutions):
+        self._relevant = set()
+        self._tracking = set()
+        self._additional = set()
+        for constraint in solutions:
+            for solution in solutions[constraint]:
+                self._tracking.add((constraint, frozenset(solution.items())))
+                self._relevant.add((constraint, frozenset(solution.items())))
+
+    @property
+    def relevant(self):
+        return self._relevant
+
+    @property
+    def additional(self):
+        return self._additional
+
+    @property
+    def missed(self):
+        return self._tracking
+
+    def register(self, constraint: Constraint, solution):
+        s_strings = frozenset({name: str(group) for name, group in solution.items()}.items())
+        if (constraint.name, s_strings) in self._relevant:
+            self._tracking.remove((constraint.name, s_strings))
+        elif is_excel_constraint(constraint):
+            self._additional.add((constraint.name, s_strings))
+
+    def count(self, found=None, relevant=None, supported=False):
+        relevant_list = self._relevant
+        if supported is True:
+            relevant_list = list(r for r in relevant_list if r in [c.name for c in excel_constraints()])
+        if relevant is True and found is True:
+                return len(relevant_list) - len(self._tracking)
+        elif relevant is True and found is False:
+                return len(self._tracking)
+        elif relevant is True and found is None:
+                return len(relevant_list)
+        elif relevant is False and found is True:
+                return len(self._additional)
+        elif relevant is None and found is True:
+                return len(relevant_list) - len(self._tracking) + len(self._additional)
+        raise Exception("Unknown")
+
+
+class Experiment:
+    def __init__(self, name):
+        super().__init__()
+        self._name = name
+        self._run = False
+        self._running_times = []
+        self._cells = None
+        self._tables = None
+        self._vectors = None
+        self._counter = None
+
+    @property
+    def name(self):
+        return self._name
+
+    @property
+    def cells(self):
+        self.run()
+        return self._cells
+
+    @property
+    def tables(self):
+        self.run()
+        return self._tables
+
+    @property
+    def vectors(self):
+        return self._vectors
+
+    @property
+    def counter(self):
+        self.run()
+        return self._counter
+
+    def running_times(self, amount=10):
[…]

### 2016-06-07 Code/src/playground.py

@@ -1,0 +1,77 @@
+import numpy
+from pandas import json
+
+from draw import Plotter
+from engine.util import local
+from experiment import Experiment, ConstraintCounter
+
+
+def main():
+    with open(local("data/data.txt")) as data_file:
+        data_json = json.load(data_file)
+        categories = ["Exercises", "Tutorials", "Data"]
+        log = {}
+        for category in data_json:
+            name = category["Category"]
+            files = category["Files"]
+            log[name] = []
+            for file in files:
+                experiment = Experiment(file)
+                print(file)
+                for accident in experiment.counter.additional:
+                    print("\tADDITIONAL:\t{}".format(accident))
+                log[name].append(experiment)
+
+        info = [
+            ("Tables", lambda e: e.tables, "grey", lambda x: str(x)),
+            #("Vectors", lambda e: e.vectors, "cyan", lambda x: str(x)),
+            ("Cells", lambda e: e.cells, "black", lambda x: str(x)),
+            ("Relevant Constraints", lambda e: e.counter.count(relevant=True), "magenta", lambda x: str(x)),
+            ("Recall", lambda e: recall(e.counter), "green", lambda x: "{:.2f}".format(x)),
+            ("Recall Supported", lambda e: recall(e.counter, True), "lightgreen", lambda x: "{:.2f}".format(x)),
+            ("Precision", lambda e: precision(e.counter), "yellow", lambda x: "{:.2f}".format(x)),
+            ("Runtime", lambda e: numpy.average(e.running_times(10)), "red", lambda x: "{:.2f}s".format(x))
+        ]
+
+        maximal = {name: max(max(f(e) for e in experiments) for experiments in log.values()) for name, f, _, _ in info}
+        colors = dict(zip(categories, ["red", "yellow", "black", "green"]))
+
+        print("\n" * 2)
+        max_label = list(t[3](maximal[t[0]]) for t in info)
+        global_plot = Plotter("comparison", list(t[0] for t in info), upper=max_label)
+        for category, experiments in [(cat, log[cat]) for cat in categories]:
+            print(category)
+            plotter = Plotter("benchmark_" + category)
+            averages = []
+            errors = []
+            for name, f, color, pretty in info:
+                plotter.add_category(name, list(f(e) / maximal[name] for e in experiments), color)
+                result_string = "\tAverage {name}: {average} - total: {total}"
+                average = numpy.average(list(f(e) for e in experiments))
+                averages.append(average / maximal[name])
[…]

### 2016-06-07 Paper/ECAI/tabular.tex
shortened the paper

@@
 
-Consider Figure \ref{fig:autocompletion_example} containing two tables.
-Learning on a snapshot that does not contain line 17, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
-The new data $\Delta D$ consists of the new value \textit{West} on line 17.
-Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
-The \textit{SUMIF} constraints have enough input to fill the incomplete vectors and predict the values in the last row: $\range{T_2}{4}{2} = 1547$ and $\range{T_2}{4}{3} = 428128$.
+% Consider Figure \ref{fig:autocompletion_example} containing two tables.
+% Learning on a snapshot that does not contain line 17, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
+% The new data $\Delta D$ consists of the new value \textit{West} on line 17.
+% Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
+% The \textit{SUMIF} constraints have enough input to fill the incomplete vectors and predict the values in the last row: $\range{T_2}{4}{2} = 1547$ and $\range{T_2}{4}{3} = 428128$.
 
@@
 \subsection{Error detection}
-\begin{figure*}[thb]
-\begin{subfigure}{.70\textwidth}
+\begin{figure}[thb]
+%\begin{subfigure}{.70\textwidth}
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/fbi_figure_highlighted.png}
+    \includegraphics[width=\linewidth]{figures/fbi_figure_highlighted.png}
   \end{center}
@@
   \label{fig:fbi}
-\end{subfigure}
-\hfill
-\begin{subfigure}{.29\textwidth}
-\begin{center}
-    \includegraphics[width=\textwidth]{figures/autocompletion_example.png}
-  \end{center}
-  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
-  \label{fig:autocompletion_example}
-\end{subfigure}
-\end{figure*}
+%\end{subfigure}
+% \hfill
+% \begin{subfigure}{.29\textwidth}
+% \begin{center}
+%     \includegraphics[width=\textwidth]{figures/autocompletion_example.png}
+%   \end{center}
+%   \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
+%   \label{fig:autocompletion_example}
+% \end{subfigure}
+\end{figure}
 

### 2016-06-07 Paper/ECAI/tabular.tex
removing and moving the figures from and back

--- 2016-06-07 Paper/ECAI/tabular.tex
@@
 
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{figures/tabletool.png}
-  \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding to (headerless) tables.
-  An explicit orientation can also be provided.}
-  \label{fig:visual_tool}
-\end{figure}
+% \begin{figure}[t]
+%   \centering
+%   \includegraphics[width=1\linewidth]{figures/tabletool.png}
+%   \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding to (headerless) tables.
+%   An explicit orientation can also be provided.}
+%   \label{fig:visual_tool}
+% \end{figure}
 
@@
 The above assumptions do not hold for many tables.
-However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (Figure~\ref{fig:visual_tool}).
+However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool.% (Figure~\ref{fig:visual_tool})
 Users select tables excluding the header and optionally specify an orientation.
@@
 \subsection{Error detection}
-\begin{figure*}[thb]
-\begin{subfigure}{.70\textwidth}
+\begin{figure}[thb]
+% %\begin{subfigure}{.70\textwidth}
+%   \begin{center}
+%     \includegraphics[width=\width]{figures/fbi_figure_highlighted.png}
+%   \end{center}
+%   \caption{Real world tabular constraint reconstruction: FBI crime statistics}
+%   \label{fig:fbi}
+%\end{subfigure}
+% \hfill
+% \begin{subfigure}{.29\textwidth}
 \begin{center}
-    \includegraphics[width=\textwidth]{figures/fbi_figure_highlighted.png}
-  \end{center}
-  \caption{Real world tabular constraint reconstruction: FBI crime statistics}
-  \label{fig:fbi}
-\end{subfigure}
-\hfill
-\begin{subfigure}{.29\textwidth}
-\begin{center}
-    \includegraphics[width=\textwidth]{figures/autocompletion_example.png}
+    \includegraphics[width=0.60\linewidth]{figures/autocompletion_example.png}
   \end{center}
@@
   \label{fig:autocompletion_example}
-\end{subfigure}
-\end{figure*}
+%\end{subfigure}
+\end{figure}
 
@@
 The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
-Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
-Our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
-A possible approach to detect such missing or wrong values is to look at a sample of the table, e.g.
+% Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
+% Our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
+% A possible approach to detect such missing or wrong values is to look at a sample of the table, e.g.
 trying to leave out each row, and compare the constraints learned on this sample with the constraints learned on the whole table.

### 2016-06-08 Paper/ECAI/tabular.tex
Fixes

--- 2016-06-07 Paper/ECAI/tabular.tex
@@
 \sname takes a few seconds to find the constraints in Figure~\ref{fig:sol_example} (excluding structural constraints).
-These include 5 spurious $\mathit{RANK}$ constraints, such as $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
+These include 5 spurious $\mathit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
 Moreover, there is 1 \textit{LOOKUP} constraint that was not present in the original spreadsheet (looking up ID based on Salesperson).
@@
 While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when ignoring equalities.
-This shows the need for a (heuristic) post-processing step to increase precision and further optimizations to decrease running times.
+By using a (heuristic) post-processing step the precision for this example could be increased.
+The running time can be improved by further optimizing the implementation, especially for conditional aggregates.
 

### 2016-06-09 Paper/ECAI/tabular.tex
Updated paper and figures

@@
     Example spreadsheet (black words and numbers only).
-    Green background indicate headerless tables, dark borders indicate maximal type-consistent blocks.
-    Most tables only contain type-consistent columns, however, table~$T_2$ also contains type-consistent rows.
-    Due to the location of the header of~$T_2$ we can infer that it consists of row data.
+    Green background indicates headerless tables, dark borders indicate maximal type-consistent blocks.
+    Most tables only contain type-consistent columns while table~$T_2$ also contains type-consistent rows.
+    However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
     This example was created by combining several Excel sheets from several exercise sessions.
@@
 
-The implementation is illustrated using a case study on the spreadsheet corresponding to the previously introduced example (Figure~\ref{fig:main_example}).
+First, our implementation is illustrated using a case study on the spreadsheet corresponding to the example introduced in Figure~\ref{fig:main_example}.
 In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 34 spreadsheets that we assembled from various sources.
@@
 
-Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates. \samuel{More details}
-The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
-For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).
+% Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
+% The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
+% For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 Consider Figure \ref{fig:autocompletion_example} containing two tables.
-Learning on a snapshot that does not contain line 17, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
-The new data $\Delta D$ consists of the new value \textit{West} on line 17.
+Learning on a snapshot that does not contain the row below $T_2$, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
+The new data $\Delta D$ consists of the new value \textit{West} in the row below $T_2$.
 Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
@@
 \begin{center}
-    \includegraphics[width=0.60\linewidth]{figures/autocompletion_example.png}
+    \includegraphics[width=0.65\linewidth]{figures/Learning.png}
   \end{center}
-  \caption{\sergey{Samuel, Tias have a look at it, should be corrected and extended, like with groups indications and formulas?}}
+  \caption{Autocompletion works by learning constraint on a snapshot and combining those constraints with new data $\Delta D$.
+  The figure shows how, given a new value \textit{West}, we can suggest values for the other cells in that row.}
   \label{fig:autocompletion_example}
@@
 We consider two cases of error detection.
-The first setting is the \textit{online} detection of errors which is similar to the autocompletion setting.
-Here, we look at \textit{conflict} constraints, i.e. constraints that were learned on the snapshot data but do not agree with the newly added data~$\Delta D$.
-Such conflict constraints can indicate errors in the input.
+The first setting is the \textit{online} detection of errors and is similar to the autocompletion setting.
+We look at \textit{conflict} constraints, i.e. constraints that were learned on the snapshot data but do not agree with the newly added data~$\Delta D$.
+Conflict constraints can indicate errors in the input.
 However, such conflicts can be wrongly caused by spurious constraints.
-Therefore, conflicts should perhaps only be detected if the snapshot data is large enough.
+Hence, conflicts should perhaps only be detected if the snapshot data is large enough.
 %An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.

### 2016-06-09 Paper/ECAI/tabular.tex
Improved formatting

@@
 %
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^* , \ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^* \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^*, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*, \ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^* \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^*\\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^* \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^*\\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*\\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^* \\
 %
@@
   \vspace{-18pt}
-  \caption{Constraints present in the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).\\
-  $^*$: Not present in original document}
+  \caption{Constraints learned for the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
+  Constraints marked with $*$ were not present in the original document}
   \label{fig:sol_example}

### 2016-06-11 Paper/ECAI/tabular.tex
Final touches
Version before proofread

@@
 
-  \begin{subfigure}{.65\textwidth}
+  \begin{subfigure}{.66\textwidth}
   \begin{center}
@@
 \hfill
-\begin{subfigure}{.35\textwidth}
+\begin{subfigure}{.34\textwidth}
   {\footnotesize
@@
   \vspace{-18pt}
-  \caption{Constraints learned for the running example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
+  \caption{Constraints learned for the example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
   Constraints marked with $*$ were not present in the original document}
@@
   Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.
-  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows $SUM$ but \sname also supports $MAX$, $MIN$, $AVERAGE$, $PRODUCT$ and $COUNT$.
+  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows \textit{SUM} but \sname also supports \textit{MAX}, \textit{MIN}, \textit{AVERAGE}, \textit{PRODUCT} and \textit{COUNT}.
   %{\small\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
@@
       & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
+      \\[\PADSEP] \hline
+
+    $\eccalc{\sg_r}{\frac{\sg_1 - \sg_2}{\sg_2}}$
+      & \textit{Same as \ecdiff{\sg_r}{\sg_1}{\sg_2}}
+
+      & $\sg_{r}[i] = (\sg_{1}[i] - \sg_{2}[i]) / \sg_{2}[i]$.
       \\[\PADSEP] \hline
@@
 \subsection{Optimizations} \label{sec:opts}
-This section discusses various design decisions and optimizations aimed at improving the speed and extensibility of our system as well as to eliminate some obvious redundant constraints.
+This section discusses various design decisions and optimizations aimed at improving the speed and extensibility of our system as well as to eliminate redundant constraints.
 
@@
 
-As a result, in some cases a maximal subblock might falsely include irrelevant columns.
+In some cases a maximal subblock might falsely include irrelevant columns.
 Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$.
-Should the target constraint be $\eccalc{B_2}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block before the above approach could find it.
-
-Alternatively, we can output all valid argument assignments instead of only maximal subblocks, e.g.
+Should the target constraint be $\eccalc{B_2}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block.
+
+Alternatively, we can output all valid assignments instead of only maximal subblocks, e.g.
 $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$, however, this might produce many redundant constraints including constraints that accidentally cover too little data.
@@
 These were determined using the formulas (if present), context and headers in the original spreadsheets.
-We encoded all relevant constraints using our representation (in canonical form).
-Not all relevant constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
-Constraints that are currently outside of the scope of this system are, for example, generic nested mathematical or logical formulas.
-
-We focus on \textit{functional} constraints that could be used in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
-We included detected equalities as relevant constraints (in 5~spreadsheets).
+We encoded all intended constraints using our representation (in canonical form).
+Not all intended constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
+Constraints that are currently outside of the scope of this system are, for example, nested mathematical or logical formulas.
+
+We focus on \textit{functional} constraints that could be used as formulas in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
+Moreover, we included detected equalities as intended constraints (in 5~spreadsheets).
 All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
@@
 
-For this example our primary goal of finding all original constraints is achieved, the recall (i.e. $\frac{\text{relevant discovered}}{\text{relevant}}$) is~$1.0$.
+For this example our primary goal of finding all original constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{intended}}$) is~$1.0$.
 The implementation also returns 6 additional constraints, therefore, our precision (i.e.
-$\frac{\text{relevant discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is rather low compared to, for example, the spreadsheets in the benchmark.
+$\frac{\text{intended discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is rather low compared to, for example, the spreadsheets in the benchmark.
 The low precision can be explained by the many short vectors which increases the chance for constraints to be true by accident.
@@
 Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
-Notice that for the data category the average number of tables and cells is higher while the number of relevant constraints is lower compared to the other categories.
+Notice that for the data category the average number of tables and cells is higher while the number of intended constraints is lower compared to the other categories.
 
@@
     \textbf{Cells} & 1231 & 137 & 1889 & 90 & 2320 & 580 \\ \hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}Relevant\\[-4pt] Constraints\end{tabular}} & 34 & 3.78 & 52 & 2.48 & 6 & 1.50 \\ \hline \hline
+    \textbf{\begin{tabular}[c]{@{}l@{}}Intended\\[-4.5pt] Constraints\end{tabular}} & 34 & 3.78 & 52 & 2.48 & 6 & 1.50 \\ \hline \hline
 \textbf{Recall} & 0.85 & 0.83 & 0.88 & 0.87 & 1.00 & 1.00 \\ \hline
-\textbf{\begin{tabular}[c]{@{}l@{}}Recall\\[-4pt] Supported\end{tabular}} & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ \hline
+\textbf{\begin{tabular}[c]{@{}l@{}}Recall\\[-4.5pt] Supported\end{tabular}} & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ \hline
 \textbf{Precision} & 0.97 & 0.98 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
@@
 
-\subsubsection*{Q1. How many relevant constraints are found by \sname?}
+\subsubsection*{Q1. How many intended constraints are found by \sname?}
 The overall recall achieved by \sname is~$0.88~(81/92)$.
@@
 Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
-However, the overall precision per category reveals that the precision is much higher for the exercises and data categories ($0.97$ and $1.00$) and much lower for the tutorials category ($0.70$).
-Moreover, the average precision per spreadsheet is, for all categories, over $0.90$.
+While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.97$ and $1.00$) than for the tutorials category ($0.70$).
 Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
@@
 
-In order for the algorithm to run efficiently it is crucial to use dependencies and find constraints incrementally whenever possible.
-This avoids some of the explosion of combinations for constraints that have many arguments.
-For example, using foreign keys as a base constraint to find conditional aggregates reduces the running time for the case study from about 3 seconds to below 1 second.
-Unfortunately, this assumption is sometimes too strong, when users are interested in aggregates for only some of the keys that are present in the data.
+To illustrate that using dependencies to find constraints incrementally can increase the efficiency, we ran the benchmarks using foreign keys as a base constraint to find conditional aggregates.
+The total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data) showing the strong effect of this dependency.
+However, the assumption that conditional aggregates are defined for all keys in the foreign key column is too strong in practice.
+
+Of the 32 constraint templates that are supported (including equality), the slowest by far is fuzzy lookup using about 36\% of the running time.
+It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates ($3.5\% - 4\%$).
 
@@
   \includegraphics[width=1\linewidth]{figures/comparison.pdf}
-  \caption{Overview of the average properties and performance of spreadsheets within every benchmark category relative to the maximum across all spreadsheets.}
+  \caption{Overview of the average properties and performance of spreadsheets within every benchmark category relative to the maximum across all spreadsheets.
+  Green bars indicate standard deviation within the category.}
   \label{fig:comparison}
@@
 \subsection{Testing the limits}
-On many spreadsheets \sname achieves good results.
+On many spreadsheets that adhere to the bias of tables, continuous blocks and vectors \sname achieves good results.
 However, factors such as incompatible formatting, bad spreadsheet design and rounding issues can lead to decreased recall, many additional constraints and long running times due to an exponential blow-up of candidate constraints.
@@
 By using a (heuristic) post-processing step the precision for this example could be increased.
-The running time can be improved by further optimizing the implementation, especially for conditional aggregates.
+The running time can be improved by implementing additional optimizations.
 
@@
 We have presented and evaluated our approach, implemented as the system \sname, that is able to learn many different constraints in tabular data.
-The resulting method has high accuracy, produces limited number of redundant constraints and is sufficiently efficient for normal as well as interactive use.
+The resulting method has high recall, produces limited number of redundant constraints and is sufficiently efficient for normal and interactive use.
 

### 2016-06-12 Paper/ECAI/tabular.tex
Final read & edits [Samuel]

@@
 
-\newcommand{\sg}{B}
-\newcommand{\sbl}[1]{\ensuremath{B_{\textit{#1}}}}
-\newcommand{\bsbl}[1]{\ensuremath{\mathbf{B_{\textit{#1}}}}}
+\newcommand{\sbs}{B}
+\newcommand{\sbl}[1]{\ensuremath{\sbs_{\textit{#1}}}}
+\newcommand{\ssbl}[1]{\ensuremath{\sbs'_{\textit{#1}}}}
+\newcommand{\bsbl}[1]{\ensuremath{\mathbf{\sbs_{\textit{#1}}}}}
 
@@
 \begin{abstract}
-%Spreadsheets, comma separated value files and other tabular data representations are in wide use today.
-However, the correct use and maintenance of functions in and between the tabular data can be error-prone and overwhelming.
-In this work, we investigate the automatic discovery of constraints (functions and relations) from raw tabular data.
-Our method takes inspiration form inductive logic programming and constraint satisfiability.
-We represent common spreadsheet functions as predicates with arguments that must satisfy number of constraints over the arguments.
-This allows generic constraint satisfaction techniques to be used to find all such satisfying predicates/functions within the rows, columns and submatrices of the tables.
-We show the effectiveness and accuracy of this approach on a number of spreadsheets from varying sources.
 Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
@@
 Constraint learning techniques are used to identify the predicates that hold between blocks of rows and columns of the tables.
-%New types of constraints can be easily added to the system by specifying them in a declarative language.
 We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
@@
 Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
-Looking at the first row in Table~$T_2$ and the data in Table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
+Looking at the first row in table~$T_2$ and the data in table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
 Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
@@
 foreign-keys).
-And on the other hand, the data is most often mixed textual and numeric.
+On the other hand, the data is most often mixed, e.g.
+textual and numeric.
 %, since people tend to use spreadsheets for financial and accounting computations.
-Hence understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
+Therefore, understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
 
@@
 very few examples.
-Our approach contrasts with these in that it focuses
-on learning both column- and row-constraints on integer, floating point as well as textual data.
+Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.
 
 This paper is organized as follows.
-Section \ref{sec:formalization} introduces concepts relevant to our approach.
-Section \ref{sec:approach} presents the problem statement and the approach.
+Section \ref{sec:formalization} introduces concepts relevant to our approach and our problem statement.
+Section \ref{sec:approach} presents our approach.
 Section \ref{sec:evaluation} presents the evaluation of the approach.
@@
     However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
-    This example was created by combining several Excel sheets from several exercise sessions.
+    This example was created by combining several Excel sheets from several exercise sessions based on~\cite{excel_book}.
     % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
@@
   \vspace{-18pt}
-  \caption{Constraints learned for the example (left), except \textit{ALLDIFFERENT} (16), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
+  \caption{Constraints learned for the example (left), except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
   Constraints marked with $*$ were not present in the original document}
@@
 We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$ and similarly $T[{:},a]$ for the $a$-th column.
-For example, in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_5[:,1] = ['$Diana Coolen$', 5]$.
-$T_5[:,1]$ is not type-consistent while $T_1[1,:]$ is.
+For example, in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_3[:,1] = ['$Diana Coolen$', 5]$.
+$T_3[1, :]$ is not type-consistent while $T_1[:, 1]$ is.
 %We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.
@@
 \begin{definition}
-A \textbf{block} has to satisfy three conditions: 1) it contains only entire rows or entire columns of a single headerless table 2) it is contiguous and 3) it is type-consistent.
+A \textbf{block} has to satisfy three conditions: 1)~it contains only entire rows or entire columns of a single headerless table; 2)~it is contiguous; and 3)~it is type-consistent.
 The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type. 
-If it contains only rows we say it has \textit{row-orientation}, if only columns, \textit{column-orientation}. 
+If a block contains only rows we say it has \textit{row-orientation}, if only columns, \textit{column-orientation}. 
 \end{definition}
@@
 Its rows are not type consistent (i.e. they contain both numeric and textual data).
-However, the table can be partitioned into five column-oriented blocks $B_1, B_2, B_3, B_4, B_5$, as shown in the figure ($B_1 = \range{T_1}{\rangeall}{1}$, $B_2 = \range{T_1}{\rangeall}{2}$, $B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
-% {\small
-% B_1 = \range{T_1}{\rangeall}{1},
-% B_2 = \range{T_1}{\rangeall}{2},
-% B_3 = \range{T_1}{\rangeall}{\rangeto{3}{8}},
-% B_4 = \range{T_1}{\rangeall}{9},
-% B_5 = \range{T_1}{\rangeall}{\rangeto{10}{11}}
+However, the table can be partitioned into five column-oriented blocks $\sbl{1}, \sbl{2}, \sbl{3}, \sbl{4}, \sbl{5}$, as shown in the figure ($\sbl{1} = \range{T_1}{\rangeall}{1}$, $\sbl{2} = \range{T_1}{\rangeall}{2}$, $\sbl{3} = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
 \end{example}
@@
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    $\ecalldiff{\sg_x}^*$
-      & $\discrete(\sg_x)$
-
-      & All values in $\sg_x$ are different. $i \neq j$: $\sg_x[i] \neq \sg_x[j]$
+    $\ecalldiff{\sbl{x}}^*$
+      & $\discrete(\sbl{x})$
+
+      & All values in $\sbl{x}$ are different. $i \neq j$: $\sbl{x}[i] \neq \sbl{x}[j]$
       \\[\PADSEP] \hline
 
-    $\ecperm{\sg_x}^*$
-      & $\numeric(\sg_{x})$, $\ecalldiff{\sg_{x}}$
-
-      & The values in $\sg_{x}$ are a permutation of the numbers $1$ through $\plength(\sg_{x})$.
+    $\ecperm{\sbl{x}}^*$
+      & $\numeric(\sbl{x})$, $\ecalldiff{\sbl{x}}$
+
+      & The values in $\sbl{x}$ are a permutation of the numbers $1$ through $\plength(\sbl{x})$.
       \\ \hline
 
-    \ecseries{\sg_x}
-      & $\integer(\sg_{x})$ and $\ecperm{\sg_{x}}$
-
-      & $\sg_{x}[1] = 1$ and $\sg_{x}[i] = \sg_{x}[i - 1] + 1$.
+    \ecseries{\sbl{x}}
+      & $\integer(\sbl{x})$ and $\ecperm{\sbl{x}}$
+
+      & $\sbl{x}[1] = 1$ and $\sbl{x}[i] = \sbl{x}[i - 1] + 1$.
       \\[\PADSEP] \hline
 
-    $\ecfkey{\sg_{fk}}{\sg_{pk}}^*$
-      & $\sg_{\mathit{fk}}$ and~$\sg_{pk}$ are $\discrete$; they must have different $\ptable$s but the same $\ptype$; and $\ecalldiff{\sbl{pk}}$
-
-      & Every value in~$\sg_{fk}$ also exist in~$\sg_{pk}$ \\[\PADSEP] \hline
+    $\ecfkey{\sbl{fk}}{\sbl{pk}}^*$
+      & $\sbl{fk}$ and~$\sbl{pk}$ are $\discrete$; they must have different $\ptable$s but the same $\ptype$; and $\ecalldiff{\sbl{pk}}$
+
+      & Every value in~$\sbl{fk}$ also exist in~$\sbl{pk}$ \\[\PADSEP] \hline
 
@@
     \eclookupprod{\sbl{r}}{\sbl{1}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{fk}\}$ are $\numeric$, arguments $\{\sg_{pk}, \sg_{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sg_{fk}}{\sg_{pk}}.
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{fk}\}$ are $\numeric$, arguments $\{\sbl{pk}, \sbl{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sbl{fk}}{\sbl{pk}}.
 
@@
 
-    \ecprod{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$
-
-      & $\sg_{r}[i] = \sg_{1}[i] \times \sg_{2}[i]$.
+    \ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}}
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are all $\numeric$ and have the same $\plength$
+
+      & $\sbl{r}[i] = \sbl{1}[i] \times \sbl{2}[i]$.
       \\[\PADSEP] \hline
 
-    \ecdiff{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_{r}, \sg_{1}, \sg_{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
-
-      & $\sg_{r}[i] = \sg_{1}[i] - \sg_{2}[i]$.
+    \ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
+
+      & $\sbl{r}[i] = \sbl{1}[i] - \sbl{2}[i]$.
       \\[\PADSEP] \hline
 
-    $\eccalc{\sg_r}{\frac{\sg_1 - \sg_2}{\sg_2}}$
-      & \textit{Same as \ecdiff{\sg_r}{\sg_1}{\sg_2}}
-
-      & $\sg_{r}[i] = (\sg_{1}[i] - \sg_{2}[i]) / \sg_{2}[i]$.
+    $\eccalc{\sbl{r}}{\frac{\sbl{1} - \sbl{2}}{\sbl{2}}}$
+      & \textit{Same as \ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}}
+
+      & $\sbl{r}[i] = (\sbl{1}[i] - \sbl{2}[i]) / \sbl{2}[i]$.
       \\[\PADSEP] \hline
 
-    \ecproj{\sg_r}{\mathbf{\sg_x}}
-      & Arguments $\{\sg_{r}, \mathbf{\sg_x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\mathbf{\sg_x}$ contains at least~2 vectors; and $\sg_r = \mathit{SUM}_{\por(\mathbf{\sg_x})}(\mathbf{\sg_x})$
-
-      & At every position~$i$ in $1$ through $\plength(\sg_{r})$ there is exactly one vector~$v$ in $\mathbf{\sg_x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sg_{r}[i]$.
+    \ecproj{\sbl{r}}{\bsbl{x}}
+      & Arguments $\{\sbl{r}, \bsbl{x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\bsbl{x}$ contains at least~2 vectors; and $\sbl{r} = \textit{SUM}_{\por(\bsbl{x})}(\bsbl{x})$
+
+      & At every position~$i$ in $1$ through $\plength(\sbl{r})$ there is exactly one vector~$v$ in $\bsbl{x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sbl{r}[i]$.
       \\[\PADSEP] \hline
 
-    \ecrank{\sg_r}{\sg_x}
-      & $\integer(\sg_{r})$; $\numeric(\sg_{x})$; and $\plength(\sg_{r}) = \plength(\sg_{x})$
-
-      & The values in $\sg_{r}$ represent the rank (from largest to smallest) of the values in $\sg_{x}$ (including ties)
+    \ecrank{\sbl{r}}{\sbl{x}}
+      & $\integer(\sbl{r})$; $\numeric(\sbl{x})$; and $\plength(\sbl{r}) = \plength(\sbl{x})$
+
+      & The values in $\sbl{r}$ represent the rank (from largest to smallest) of the values in $\sbl{x}$ (including ties)
       \\[\PADSEP] \hline
 
-    \ectotal{\sg_r}{\sg_{pos}}{\sg_{neg}}
-      & Arguments $\{\sg_{r}, \sg_{pos}, \sg_{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
-
-      & $\sg_{r}[i] = \sg_{r}[i - 1] + \sg_{pos}[i] - \sg_{neg}[i]$.
+    \ectotal{\sbl{r}}{\sbl{pos}}{\sbl{neg}}
+      & Arguments $\{\sbl{r}, \sbl{pos}, \sbl{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+
+      & $\sbl{r}[i] = \sbl{r}[i - 1] + \sbl{pos}[i] - \sbl{neg}[i]$.
       \\[\PADSEP] \hline
 
-    $\ecsumr{\sg_r}{\mathbf{\sg_x}}^\dagger$
-      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\pcols(\mathbf{\sg_x}) \geq 2$; and $\prows(\mathbf{\sg_x}) = \plength(\sg_r)$
-
-      & $B_r[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$
+    $\ecsumr{\sbl{r}}{\bsbl{x}}^\dagger$
+      & $\sbl{r}$ and $\bsbl{x}$ are $\numeric$; $\pcols(\bsbl{x}) \geq 2$; and $\prows(\bsbl{x}) = \plength(\sbl{r})$
+
+      & $\sbl{r}[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \textit{row}(i, \bsbl{x})[j]$
       \\[\PADSEP] \hline
 
-    $\ecsumc{\sg_r}{\mathbf{\sg_x}}^\dagger$
-      & $\sg_r$ and $\mathbf{\sg_x}$ are $\numeric$; $\prows(\mathbf{\sg_x}) \geq 2$; and $\pcols(\mathbf{\sg_x}) = \plength(\sg_r)$
-
-      & $B_r[i] = \sum_{j = 1}^{\prows(\bsbl{x})} \mathit{column}(i, \bsbl{x})[j]$
+    $\ecsumc{\sbl{r}}{\bsbl{x}}^\dagger$
+      & $\sbl{r}$ and $\bsbl{x}$ are $\numeric$; $\prows(\bsbl{x}) \geq 2$; and $\pcols(\bsbl{x}) = \plength(\sbl{r})$
+
+      & $\sbl{r}[i] = \sum_{j = 1}^{\prows(\bsbl{x})} \textit{column}(i, \bsbl{x})[j]$
       \\[\PADSEP] \hline
 
-    $\ecsumif{\sg_r}{\sg_{fk}}{\sg_{pk}}{\sg_{val}}^\dagger$
-      & $\sg_{fk}, \sg_{pk}$ are $\discrete$; $\sg_{r}, \sg_{val}$ are $\numeric$; within the sets $\{\sg_{val}, \sg_{fk}\}$ and $\{\sg_{pk}, \sg_{r}\}$ arguments have the same $\plength$ and $\por$; $\sg_{fk}$ and $\sg_{val}$ have the same $\ptable$; $\sg_{fk}$ and $\sg_{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sg_{pk}}
-
-      & \[ B_r[i] = \sum_{j=1}^{\plength(B_{\textit{val}})} \begin{cases}
-          \sg_{val}[j] & \text{if } \sg_{fk}[j] = \sg_{pk}[i] \\
+    $\ecsumif{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}^\dagger$
+      & $\sbl{fk}, \sbl{pk}$ are $\discrete$; $\sbl{r}, \sbl{val}$ are $\numeric$; within the sets $\{\sbl{val}, \sbl{fk}\}$ and $\{\sbl{pk}, \sbl{r}\}$ arguments have the same $\plength$ and $\por$; $\sbl{fk}$ and $\sbl{val}$ have the same $\ptable$; $\sbl{fk}$ and $\sbl{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sbl{pk}}
+
+      & \[ \sbl{r}[i] = \sum_{j=1}^{\plength(\sbl{val})} \begin{cases}
+          \sbl{val}[j] & \text{if } \sbl{fk}[j] = \sbl{pk}[i] \\
           0 & \text{otherwise}
@@
 
-    \ecsumprod{\sg_r}{\sg_1}{\sg_2}
-      & Arguments $\{\sg_r, \sg_1, \sg_2\}$ are $\numeric$; $\plength(\sg_{1}) = \plength(\sg_{2}) \geq 2$; and $\prows(\sg_{r}) = \pcols(\sg_{r}) = 1$
-
-      & $\sg_{r}[1] = \sum_{j = 1}^{\plength(\sg_{1})} \sg_{1}[j] \times \sg_{2}[j]$.
+    \ecsumprod{\sbl{r}}{\sbl{1}}{\sbl{2}}
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are $\numeric$; $\plength(\sbl{1}) = \plength(\sbl{2}) \geq 2$; and $\prows(\sbl{r}) = \pcols(\sbl{r}) = 1$
+
+      & $\sbl{r}[1] = \sum_{j = 1}^{\plength(\sbl{1})} \sbl{1}[j] \times \sbl{2}[j]$.
   \end{tabularx}}
@@
 \item
-\textit{\CName}  specifies the syntactic form of the constraint $s(B_1, ...,B_n)$, that is, the name of the template together
-with $n$ abstract arguments $B_i$.
-Thus a constraint is viewed as a relation or predicate of arity $n$ in first order logic.
-Note that a function $B_r=f(B_1,...,B_n)$ can be represented with the $(n{+}1)$-ary predicate $s_f(B_r,B_1,...,B_n)$.
+\textit{\CName}  specifies the syntactic form of the constraint $s(\sbl{1}, ...,\sbl{n})$, that is, the name of the template together
+with $n$ abstract arguments $\sbl{i}$.
+Thus, a constraint is viewed as a relation or predicate of arity $n$ in first order logic.
+Note that a function $\sbl{r}=f(\sbl{1},...,\sbl{n})$ can be represented with the $(n{+}1)$-ary predicate $s_f(\sbl{r},\sbl{1},...,\sbl{n})$.
 Each argument will have to be instantiated with a block.
@@
 \begin{itemize}
-  \item \CName: $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, for arguments $\sg_r$ and $\mathbf{\sg_x}$.
-
-  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
+  \item \CName: $\ecsumr{\sbl{r}}{\bsbl{x}}$, for arguments $\sbl{r}$ and $\bsbl{x}$.
+
+  \item \CSignature: $\sbl{r}$ has to be a single vector ($\textit{size}=1$) while $\bsbl{x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
   Both blocks have to be numeric.
-  This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$.
+  This constraint is orientation-specific, so it requires that the number of rows in $\bsbl{x}$ equals the length of $\sbl{r}$.
   Moreover, we add to the bias that the number of columns to sum over is at least~$2$.
 
-  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$.
-  %For example, $B_r = [12, 7]$ and $\mathbf{\sg_x}$ consists of columns: $[5, 15]$, $[7, -8]$.
+  \item \CFunction: each value in the vector $\sbl{r}$ is obtained by summing over the corresponding row in $\bsbl{x}$.
+  %For example, $\sbl{r} = [12, 7]$ and $\bsbl{x}$ consists of columns: $[5, 15]$, $[7, -8]$.
 \end{itemize}
-$SUM$ is an aggregate constraint, and a similar constraint is available for aggregate operators $MIN$, $MAX$, $AVERAGE$, $PRODUCT$ and $COUNT$.
-The \textit{conditional} aggregate variant $SUMIF$ only ranges over those cells that satisfy a condition on a related cell in that table, e.g.
+\textit{SUM} is an aggregate constraint, and a similar constraint is available for aggregate operators \textit{MIN}, \textit{MAX}, \textit{AVERAGE}, \textit{PRODUCT} and \textit{COUNT}.
+The \textit{conditional} aggregate variant \textit{SUMIF} only ranges over those cells that satisfy a condition on a related cell in that table, e.g.
 $\range{T_1}{\rangeall}{10}$ and the data in $T_5$ for matching salesperson names.
@@
 It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
-From a FOL perspective, a constraint of the form $\ecrank{B_r}{B_x}$ can be seen as a predicate $\textit{RANK}(B_r, B_x)$ where \textit{RANK} is the name of the predicate and its arguments~$B_r$ and~$B_x$ are terms, which can be seen as either uninstantiated variables or as concrete values.
+From a FOL perspective, a constraint of the form $\ecrank{\sbl{r}}{\sbl{x}}$ can be seen as a predicate $\textit{RANK}(\sbl{r}, \sbl{x})$ where \textit{RANK} is the name of the predicate and its arguments~$\sbl{r}$ and~$\sbl{x}$ are terms, which can be seen as either uninstantiated variables or as concrete values.
 This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
-For example, for the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $B_r$ is the $8$th vector in $T_1$: $B_r = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $B_x$ is the $7$th vector: $B_x = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
+For example, for the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $\sbl{r}$ is the $8$th vector in $T_1$: $\sbl{r} = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $\sbl{x}$ is the $7$th vector: $\sbl{x} = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
 
@@
 With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
-We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(B_1, ..., B_n)$ if $\sigc(B_1, ..., B_n)$ (respectively $\defc(B_1, ..., B_n)$) is satisfied.
-Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would define the predicate~$s$ using a Prolog like clause: $s(B_1, ..., B_n) \leftarrow \sigc(B_1, ..., B_n) \wedge \defc(B_1, ..., B_n)$.
+We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(\sbl{1}, ..., \sbl{n})$ if $\sigc(\sbl{1}, ..., \sbl{n})$ (respectively $\defc(\sbl{1}, ..., \sbl{n})$) is satisfied.
+Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would define the predicate~$s$ using a Prolog like clause: $s(\sbl{1}, ..., \sbl{n}) \leftarrow \sigc(\sbl{1}, ..., \sbl{n}) \wedge \defc(\sbl{1}, ..., \sbl{n})$.
 Under this interpretation, the term constraint and constraint template can be used interchangeably.
@@
 \begin{definition}
-A \textbf{valid argument assignment} of a constraint $s$ is a tuple of blocks $(B_1, ..., B_n)$ such that $s(B_1, ..., B_n)$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(B_1, ..., B_n)$ to the arguments.
+A \textbf{valid argument assignment} of a constraint template~$s$ is a tuple of blocks $(\sbl{1}, ..., \sbl{n})$ such that $s(\sbl{1}, ..., \sbl{n})$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(\sbl{1}, ..., \sbl{n})$ to the arguments.
 \end{definition}
@@
 \begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf Given} a set of instantiated blocks~${\cal B}$ over tables~${\cal T}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(B'_1, ..., B'_n)$ where $s \in {\cal S}$, $\forall i: B_i' \sqsubseteq B_i \in {\cal B}$ and $(B'_1, ..., B'_n)$ is a satisfied argument assignment of the template~$s$.
+{\bf Given} a set of instantiated blocks~${\cal B}$ over tables~${\cal T}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(\ssbl{1}, ..., \ssbl{n})$ where $s \in {\cal S}$, $\forall i: \sbl{i}' \sqsubseteq \sbl{i} \in {\cal B}$ and $(\ssbl{1}, ..., \ssbl{n})$ is a satisfied argument assignment of the template~$s$.
 \end{definition}
@@
 \label{sec:form:dependencies}
-In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g.
-for \textit{PERMUTATION}.
-This expresses a dependency of the constraint on that other constraint.
-This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the depending constraint, and its definition of its own definition plus the definition of the depending constraint.
-In FOL, we can see that one constraint entails the other, for example if $\ecperm{\sg_x}$ holds for a block $\sg_x$, then $\ecalldiff{\sg_{x}}$ also holds.
-
-Apart from easing the specification of the signature and definition, in Section~\ref{sec:approach} we will see how such dependencies can be used to speed up the search for constraints.
+In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g. for \textit{PERMUTATION}.
+This expresses a dependency of the constraint on that other \textit{base} constraint.
+This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the base constraint, and its definition of its own definition plus the definition of the base constraint.
+In FOL, we can see that one constraint entails the other, for example if $\ecperm{\sbl{x}}$ holds for a block $\sbl{x}$, then $\ecalldiff{\sbl{x}}$ also holds.
+
+In Section~\ref{sec:approach} we will see how, apart from easing the specification of the signature and definition, such dependencies can be used to speed up the search for constraints.
 
@@
 This is because constraints may be logically equivalent or may be implied by other constraints, e.g.
-if you know $\ecperm{B_x}$, $\ecalldiff{B_x}$ is redundant.
+if you know $\ecperm{\sbl{x}}$, $\ecalldiff{\sbl{x}}$ is redundant.
 
 Within the same constraint, there can be equivalent argument assignments if the order of some of the arguments does not matter.
-For example for the product constraint, $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecprod{\sg_r}{\sg_2}{\sg_1}$ so one can be considered redundant to the other.
-Two different constraints can also be logically equivalent, due to their nature e.g. for addition/subtraction and product/division: $\ecprod{\sg_r}{\sg_1}{\sg_2} \equiv \ecdiv{\sg_1}{\sg_r}{\sg_2}$.
+For example for the product constraint, $\ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}} \equiv \ecprod{\sbl{r}}{\sbl{2}}{\sbl{1}}$ so one can be considered redundant to the other.
+Two different constraints can also be logically equivalent, due to their nature e.g.
+for addition / subtraction and product / division: $\ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}} \equiv \ecdiv{\sbl{1}}{\sbl{r}}{\sbl{2}}$.
 
@@
 \begin{enumerate}
-\item For each constraint template~$s$, generate all assignments $(B_1, \ldots ,B_n)$ where each $B_i$ is a superblock and the $B_i$ are \textit{compatible} (defined below) with the signature of~$s$. 
-\item For each generated assignment $(B_1, \ldots ,B_n)$ in 3a), find all valid constraints $s(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of~$s$ are satisfied.
+\item For each constraint template~$s$, generate all assignments $(\sbl{1}, \ldots ,\sbl{n})$ where each~$\sbl{i}$ is a superblock and the~$\sbl{i}$ are \textit{compatible} (defined below) with the signature of~$s$.
+\item For each generated assignment $(\sbl{1}, \ldots ,\sbl{n})$ in 3a), find all valid constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that for all $i$ holds $\ssbl{i} \sqsubseteq \sbl{i}$ and the signature and definition of~$s$ are satisfied.
 \end{enumerate}
@@
 % and what combination of superblocks may lead to a valid argument assignment.\\
-Consider table~$T_1$ in Figure~\ref{fig:main_example} and the $\mathit{SUM}_{row}$ constraint template.
+Consider table~$T_1$ in Figure~\ref{fig:main_example} and the $\textit{SUM}_{row}$ constraint template.
 Instead of considering all possible subblocks of $T_1$, we first reason over the properties of the superblocks.
-For example, $B_2$ and $B_4$ are not numeric so they can be immediately discarded.
-Superblocks~$B_1, B_3$ and $B_5$ are valid candidates for $B_r$, however, since they all have length~4, $\mathbf{B_x}$ needs at least 4~columns which is only satisfied by $B_3$.
-So the assignments generated in step 3a would be $(B_1, B_3), (B_3, B_3), (B_5, B_3)$.
-
-In step 3b, we start from an assignment $(B_1, \ldots, B_n)$ with superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
-As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier.
-For the above example and $(B_5, B_3)$, each of the vectors of $B_5$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks of $B_3$ for the right-hand side and verify the signature (e.g.
+For example, $\sbl{2}$ and $\sbl{4}$ are not numeric so they can be immediately discarded.
+Superblocks~$\sbl{1}, \sbl{3}$ and $\sbl{5}$ are valid candidates for~$\sbl{r}$, however, since they all have length~4, $\mathbf{\sbl{x}}$ needs at least 4~columns which is only satisfied by $\sbl{3}$.
+So the assignments generated in step 3a would be $(\sbl{1}, \sbl{3}), (\sbl{3}, \sbl{3}), (\sbl{5}, \sbl{3})$.
+
+In step 3b, we start from an assignment $(\sbl{1}, \ldots, \sbl{n})$ with superblocks $\sbl{i}$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
+As we only have to consider the blocks $\ssbl{i} \sqsubseteq \sbl{i}$ contained in each superblock, this is typically much easier.
+For the above example and $(\sbl{5}, \sbl{3})$, each of the vectors of $\sbl{5}$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks of $\sbl{3}$ for the right-hand side and verify the signature (e.g.
 at least size $4$) and test whether for each row the definition is satisfied.
@@
 
-We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step 3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step 3b, Section~\ref{sec:algo:subgr}).
-In Section~\ref{sec:opts} we describe a number of optimizations designed to speed up the algorithm.
+We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step~3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step~3b, Section~\ref{sec:algo:subgr}).
+In Section~\ref{sec:opts} we describe a number of optimizations designed to improve the algorithm.
 
@@
 The above assumptions do not hold for many tables.
-However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool.% (Figure~\ref{fig:visual_tool})
+However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool.
 Users select tables excluding the header and optionally specify an orientation.
@@
         \State $A \gets \generategroups(s, \blocks)$  \label{algo:tcl:super}
-        \ForAll{$(B_1, \dots, B_n) \in A$}
-        \State{$A' \gets \findassignment(s, (B_1,\dots,B_n))$}
-        \ForAll{$(B_1',\dots,B_n') \in A'$}
-            \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
+        \ForAll{$(\sbl{1}, \dots, \sbl{n}) \in A$}
+        \State{$A' \gets \findassignment(s, (\sbl{1},\dots,\sbl{n}))$}
+        \ForAll{$(\sbl{1}',\dots,\sbl{n}') \in A'$}
+            \State $C \gets C \cup \{ c_s(\sbl{1}', \dots, \sbl{n}') \}$
           \EndFor
@@
 
-Given a constraint template~$s$ and the set of all maximal blocks~\blocks, the goal is to find all combinations of maximal blocks that are consistent with the constraint signature.
-An argument assignment $(B_1, ...
-B_n)$ is \textit{consistent} with the signature of template $s$ if for each block $B_i$ there exists at least one subblock $B_i' \sqsubseteq B_i$ that satisfies the signature.
-
-Furthermore, the choice of one argument can influence the possible candidates blocks for the other arguments, for example, if they must have the same length.
+Given a constraint template~$s$ and the set of all (maximal) superblocks~\blocks, the goal is to find all combinations of superblocks that are compatible with the constraint signature.
+An argument assignment $(\sbl{1}, ...
+\sbl{n})$ is \textit{compatible} with the signature of template $s$ if for each block $\sbl{i}$ there exists at least one subblock $\sbl{i}' \sqsubseteq \sbl{i}$ that satisfies the signature.
+
+The choice of one argument can influence the possible candidate blocks for the other arguments, for example, if they must have the same length.
 Instead of writing specialized code to generate and test the superblock assignments, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.
@@
 To reason over the blocks, we add to the constraint program a set of background facts that contain the properties of each block, namely its type, table, orientation, size, length and number of rows and columns.
-%The actual constraints correspond to the requirements defined in the signature, such as whether arguments are of a certain type, have certain dependencies or have to be equal length or orientation, etc.
-%
-% \tias{More detailed description of the properties commented out, feel free to put in}
-%Translation:
-%\begin{description}
-%  \item[Type] Require group that has supertype (e.g. integer => require numeric group, because subgroup might be integer)
-%  \item[Length] Require group to have exact length (subgroups have the same length)
-%  \item[Vectors / columns / rows] Require group to have at least the amount of vectors, columns, rows required
-%  \item[Dependency] Fix dependent variables, for every assignment construct CSP that attempts to find values for the other variables
-%  \item[Same length / table / orientation] Require groups to gave the same length table ...
-%\end{description}
-%
+
 \begin{table}[t]
@@
 \end{table}
-%
-The actual constraints must enforce that a superblock assignment is \textit{consistent} with the requirements defined in the signature.
-%
-%In order for our approach to be complete, the signature constraints can not always be enforced directly on a block.
-%If a block matches the signature it is a valid candidate, however, if it does not there might exist subblocks that still satisfy the signature.
-%\begin{example}
-%  Consider, for example, the constraint template \ecseries{x} and a superblock candidate~$B$ for~$x$ containing both integer and floating point values.
-%  While the signature of the template requires~$x$ to be integer we cannot discard~$B$ since there might be subblocks~$B' \sqsubseteq B$ that only contain integer values and satisfy the signature.
-%\end{example}
+
+The actual constraints must enforce that a superblock assignment is \textit{compatible} with the requirements defined in the signature.
 
 Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints works:
-Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment($\sqsubseteq$).
+Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment~($\sqsubseteq$).
 Typing requirements need to be relaxed to check only for base types (numeric and textual).
-Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with two many vectors contain subblocks of the required size.
+Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with too many vectors contain subblocks of the required (smaller) size.
 Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
@@
 
-  Consider the constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$\sg_r$ and~$\mathbf{\sg_x}$.
+  Consider the constraint template $\ecsumr{\sbl{r}}{\bsbl{x}}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$\sbl{r}$ and~$\bsbl{x}$.
   Given maximal blocks~\blocks, the domain of these variables are $D(V_r) = D(V_x) = \{ 1, \dots, |\blocks| \}$.
-  Finally, the signature of $\ecsumr{\sg_r}{\mathbf{\sg_x}}$ (see Table~\ref{table:constraints}) will be translated into constraints:
+  Finally, the signature of $\ecsumr{\sbl{r}}{\bsbl{x}}$ (see Table~\ref{table:constraints}) will be translated into constraints:
   \begin{align*}
@@
 \label{sec:algo:subgr}
-Given a superblock assignment $(B_1, \dots, B_n)$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e.
-assignments of subblocks $(B'_1, \dots, B'_n)$ (for all $i$, $B'_i \subseteq B_i$) that satisfy both the signature and the definition of template~$s$.
+Given a superblock assignment $(\sbl{1}, \dots, \sbl{n})$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e.
+assignments of subblocks $(\ssbl{1}, \dots, \ssbl{n})$ (for all $i$, $\ssbl{i} \sqsubseteq \sbl{i}$) that satisfy both the signature and the definition of template~$s$.
 
@@
 \begin{example}
-  Consider the sum-over-rows constraint template $\ecsumr{\sg_r}{\mathbf{\sg_x}}$.
-  An example superblock assignment from Figure~\ref{fig:main_example} is $(B_r, \mathbf{\sg_x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
-  Note that this assignment does not satisfy the signature yet because $\sg_r$ contains more than one vector.
-  This step aims to generate subassignments $(B_r' \sqsubseteq B_r, \mathbf{B_x}' \sqsubseteq \mathbf{B_x})$ that do satisfy the signature and test whether they satisfy the definition:
+  Consider the sum-over-rows constraint template $\ecsumr{\sbl{r}}{\bsbl{x}}$.
+  An example superblock assignment from Figure~\ref{fig:main_example} is $(\sbl{r}, \bsbl{x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
+  Note that this assignment does not satisfy the signature yet because $\sbl{r}$ contains more than one vector.
+  This step aims to generate subassignments $(\sbl{r}' \sqsubseteq \sbl{r}, \mathbf{\sbl{x}}' \sqsubseteq \mathbf{\sbl{x}})$ that do satisfy the signature and test whether they satisfy the definition:
   $\forall i, \sbl{r}[i] = \sum_{j = 1}^{length(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$.
@@
 
-On the other hand, a simple generate-and-test will usually suffice.
+On the other hand, a simple generate-and-test approach, as illustrated in Algorithm~\ref{algo:gen-test-subblocks}, will usually suffice.
 %However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
@@
     \footnotesize
-    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
+    \Procedure{Subassignments}{$s$, $(\sbl{1}, \dots, \sbl{n})$}
     \State $n \gets$ number of arguments of template $s$
     \State $A_{sub} \gets \emptyset$
-    \ForAll{$(B_1', \dots, B_n') \textbf{ where } \forall i: B_i' \sqsubseteq B_i  \land \textit{disjoint}_{j \neq i}(B_i', B_j')$}
-      \If{$\mathit{Sig}_s(B_1',\dots,B_n') \land \mathit{Def}_s(B_1',\dots,B_n')$}
-        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
+    \ForAll{$(\sbl{1}', \dots, \sbl{n}') \textbf{ where } \forall i: \sbl{i}' \sqsubseteq \sbl{i}  \land \textit{disjoint}_{j \neq i}(\sbl{i}', \sbl{j}')$}
+      \If{$\textit{Sig}_s(\sbl{1}',\dots,\sbl{n}') \land \textit{Def}_s(\sbl{1}',\dots,\sbl{n}')$}
+        \State $A_{sub} \gets A_{sub} \cup \{ (\sbl{1}', \dots, \sbl{n}') \}$
       \EndIf
@@
 \subsection{Optimizations} \label{sec:opts}
-This section discusses various design decisions and optimizations aimed at improving the speed and extensibility of our system as well as to eliminate redundant constraints.
+This section discusses various design decisions and optimizations aimed at improving the speed, the extensibility of our system and the elimination of redundant constraints.
 
@@
 
-In Inductive Logic Programming, one often exploits implications between constraints to structure the search space \cite{luc_book}.
+In Inductive Logic Programming, one often exploits implications between constraints to structure the search space~\cite{luc_book}.
 Our approach uses the dependencies between templates to define a dependency graph.
@@
 When using this ordering, templates later in the ordering can use the learned constraints of the templates they depend on.
-This is made possible by adding, on line \ref{algo:tcl:super}, the set of learned constraints $C$ to the argument of \generategroups.
-
-This is used to speed up the \generategroups step.
+To enable this reuse of constraints, the set of learned constraints $C$ is added to the argument of \generategroups on line \ref{algo:tcl:super}.
+
+The previously learned constraints are used to speed up the \generategroups step.
 For constraint templates where some arguments are also part of a constraint that it depends on, it is not needed to search for superblocks and subblocks from scratch.
-Instead, one can start from all and only the actual argument assignments of the depending constraint, and only search matching superblocks for the remaining arguments.
+Instead, one can start from all and only the actual argument assignments of the base constraint, and only search matching superblocks for the remaining arguments.
 
 \begin{example}
-  Consider $\ecfkey{B_{fk}}{B_{pk}}$, which states that every value in $B_{fk}$ also exists in $B_{pk}$; its signature includes $\ecalldiff{B_{pk}}$.
-There are 16 \textit{ALLDIFFERENT} constraints to be found in Figure~\ref{fig:main_example}, hence, the \generategroups only needs to check which superblocks for $B_{pk}$ are compatible (different table, same type) with these 16 assignments to $B_{fk}$.
+  Consider $\ecfkey{\sbl{fk}}{\sbl{pk}}$, which states that every value in $\sbl{fk}$ also exists in $\sbl{pk}$; its signature includes $\ecalldiff{\sbl{pk}}$.
+There are 18 \textit{ALLDIFFERENT} constraints to be found in Figure~\ref{fig:main_example}, hence, the \generategroups only needs to check which superblocks for $\sbl{fk}$ are compatible (different table, same type) with these 18 assignments to $\sbl{pk}$.
 \end{example}
 
-%Instead of generating assignments for~$B_{pk}$ using the given blocks~$\blocks$, we will only consider (sub)blocks~$P = \{ B' | \ecalldiff{B'} \in C \}$.
-%$P$~can be seen as a set of partial assignments in which the argument $B_{pk}$ has been fixed.
 Instead of generating one CSP to find all assignments, a CSP is generated for every known assignment of the depending constraint (and variables involved), which then searches for all assignments completing this partial assignment.
-The procedure to generate CSPs remains the same otherwise. %, except that the known subassignment(s) are added as extra blocks that can only be assigned to their corresponding variable.
-
-%This approach has the additional benefit of ensuring that the (sub)blocks assigned to the dependent arguments already satisfy both the signature and definition of the template it depends on.
-%Therefore, the implementation of $\findassignment$ does not have to take these blocks into account.
+The procedure to generate CSPs remains the same otherwise.
 
@@
 We consider two types of redundancy that we aim to eliminate during search.
-As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{B_R'}{B_1'}{B_2'} \Leftrightarrow \ecprod{B_R'}{B_2'}{B_1'}$.
+As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{\sbl{R}'}{\sbl{1}'}{\sbl{2}'} \Leftrightarrow \ecprod{\sbl{R}'}{\sbl{2}'}{\sbl{1}'}$.
 Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
 In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
-Moreover, among the semantically equivalent constraints product (\ecprod{B_R'}{B_1'}{B_2'}) and division (\eccalc{B_1'}{B_R' / B_2'}) we only added product, as well as only difference (\ecdiff{B_r}{B_1}{B_2}) and not sum (\eccalc{B_1}{B_r + B_2}).
-However, division and sum could easily be added, as could it be added explicitly in post-processing based on the matching product and difference constraints.
+Moreover, among the semantically equivalent constraints product (\ecprod{\sbl{R}'}{\sbl{1}'}{\sbl{2}'}) and division (\eccalc{\sbl{1}'}{\sbl{R}' / \sbl{2}'}) we only added product, as well as only difference (\ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}) and not sum (\eccalc{\sbl{1}}{\sbl{r} + \sbl{2}}).
+However, division and sum could easily be added, as they could be added explicitly in post-processing based on the matching product and difference constraints.
 
 A last type of redundancy we chose to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
-For example, consider the constraint \ecsumc{B_r}{B[1{:}n]} where $B[k{:}n]$ consists of only zeros, then \ecsumc{B_r}{B[1{:}j]} will be true for all $k - 1 \leq j \leq n$.
+For example, consider the constraint $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{n}}}$ where $\range{T}{\rangeall}{\rangeto{k}{n}}$ consists of only zeros, then $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{j}}}$ will be true for all $k - 1 \leq j \leq n$.
 In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.
@@
 In some cases a maximal subblock might falsely include irrelevant columns.
-Consider $B_1 = [[200, 300], [200, 150], [1, 2]]$ and $B_2 = [200, 300]$, then \sname will find the constraint $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$.
-Should the target constraint be $\eccalc{B_2}{\mathit{MAX}(B_1[1{:}2])}$, then the last vector of~$B_1$ must be split off into a separate block.
+Consider superblock $\bsbl{x} = \range{T}{\rangeall}{\rangeto{1}{3}} = [[200, 300], [200, 150], [1, 2]]$ and $\sbl{r} = [200, 300]$, then \sname will find the constraint $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$.
+Should the target constraint be $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$, then $\range{T}{\rangeall}{3}$ must be split off into a separate block.
 
 Alternatively, we can output all valid assignments instead of only maximal subblocks, e.g.
-$\eccalc{B_2[1]}{\mathit{MAX}(B_1[1])}$, $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}2])}$ and $\eccalc{B_2[1]}{\mathit{MAX}(B_1[1{:}3])}$, however, this might produce many redundant constraints including constraints that accidentally cover too little data.
+$\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{1})}$, $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$ and $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$, however, this might produce many redundant constraints including constraints that accidentally cover too little data.
 
@@
 In order to obtain a benchmark of spreadsheets we looked at three sources:
-1)~spreadsheets from an exercise session for teaching Excel; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports\footnote{(attached to to the submitted paper)}.
+1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports\footnote{(benchmark attached to to the submitted paper)}.
 All spreadsheets were converted to CSV and the table definitions are fixed.
@@
 \sname takes a few seconds to find the constraints in Figure~\ref{fig:sol_example} (excluding structural constraints).
-These include 5 spurious $\mathit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
+These include 5 spurious $\textit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
 Moreover, there is 1 \textit{LOOKUP} constraint that was not present in the original spreadsheet (looking up ID based on Salesperson).
 
-For this example our primary goal of finding all original constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{intended}}$) is~$1.0$.
+For this example our primary goal of finding all original constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{all intended}}$) is~$1.0$.
 The implementation also returns 6 additional constraints, therefore, our precision (i.e.
@@
 Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
-Notice that for the data category the average number of tables and cells is higher while the number of intended constraints is lower compared to the other categories.
+Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
 
@@
 To illustrate that using dependencies to find constraints incrementally can increase the efficiency, we ran the benchmarks using foreign keys as a base constraint to find conditional aggregates.
-The total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data) showing the strong effect of this dependency.
+The total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data), showing the strong effect of this dependency.
 However, the assumption that conditional aggregates are defined for all keys in the foreign key column is too strong in practice.
 
-Of the 32 constraint templates that are supported (including equality), the slowest by far is fuzzy lookup using about 36\% of the running time.
-It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates ($3.5\% - 4\%$).
+Of the 32 supported constraint templates (including equality), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
+It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates (each $3.5\% - 4\%$).
 
@@
 However, factors such as incompatible formatting, bad spreadsheet design and rounding issues can lead to decreased recall, many additional constraints and long running times due to an exponential blow-up of candidate constraints.
-For example, we tested a spreadsheet (``exps\_tias'') with 6~tables, 531~cells and many duplicate columns across tables.
+For example, we tested a spreadsheet (``exps'') with 6~tables, 531~cells and many duplicate columns across tables.
 While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when ignoring equalities.
@@
 Using~$\Delta D$ we can keep track of which vectors have changed and which vectors are now \textit{incomplete}.
-\textit{Functional} constraints, such as $\ecsumc{B_r}{\mathbf{B_x}}$, are able to compute a result~($B_r$) based on its input~($\mathbf{B_x}$).
+\textit{Functional} constraints, such as $\ecsumc{\sbl{r}}{\mathbf{\sbl{x}}}$, are able to compute a result~($\sbl{r}$) based on its input~($\mathbf{\sbl{x}}$).
 If all vectors in the input blocks have changed and the output vector is incomplete, the constraint can compute the resulting value(s).
@@
 For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
-If we define the sample to be the snapshot data and the other row(s) to be the additional input~$\Delta D$, this task can be formulated as online error detection and dealt with in the same way.
+This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g.
+leave out each row and consider it as $\Delta D_i$) to detect conflicts.
+
 %Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
@@
 In future work we plan to extend this to learn arbitrary nested constraints (e.g.
-$B_r = (B_1 + B_2) / B_3$), as well as constraints over only a subset of the vectors.
+$\sbl{r} = (\sbl{1} + \sbl{2}) / \sbl{3}$), as well as constraints over only a subset of the vectors.
 This may give rise to more redundant and spurious constraints, which was not problematic up to this point.
 
-One other line of work, which can also mitigate redundant and spurious constraints, is to integrate the approach an interactive setting, where users can receive and provide feedback; more similar to an active learning setting.
-Other ways to handle redundant constraints include the (heuristic) filtering or post-processing of the constraint set.
+Two promising ways to mitigate redundant and spurious constraints are heuristic filtering or post-processing of the constraint set and the integration of our approach in an interactive setting where users can receive and provide feedback.
+The latter is more similar to an active learning setting.
 

### 2016-06-12 Paper/ECAI/tabular.tex
Fixed row / column mistake
added a link to the github repository

@@
 \usepackage[inline]{enumitem}
+\usepackage{footmisc}
 
@@
 We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$ and similarly $T[{:},a]$ for the $a$-th column.
-For example, in Figure~\ref{fig:main_example}, $T_1[1,:] = [1,2,3,4]$ and $T_3[:,1] = ['$Diana Coolen$', 5]$.
+For example, in Figure~\ref{fig:main_example}, $T_1[:,1] = [1,2,3,4]$ and $T_3[1,:] = ['$Diana Coolen$', 5]$.
 $T_3[1, :]$ is not type-consistent while $T_1[:, 1]$ is.
@@
 The above assumptions do not hold for many tables.
-However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool.
+However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository\footnote{\url{https://github.com/for-DSAA2016-review/TaCLe}\label{github-link}}).
 Users select tables excluding the header and optionally specify an orientation.
@@
 In order to obtain a benchmark of spreadsheets we looked at three sources:
-1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports\footnote{(benchmark attached to to the submitted paper)}.
+1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports\footref{github-link}.
 All spreadsheets were converted to CSV and the table definitions are fixed.

### 2016-06-12 Paper/ECAI/tabular.tex
tweaks to exps, and a bit related work (and one word in abstract)

@@
 However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
-In this work, we investigate the automatic discovery of constraints (functions and relations) in raw tabular data.
+In this work, we investigate the automatic discovery of constraints (formula and relations) in raw tabular data.
 We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
@@
 These include 5 spurious $\textit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
-Moreover, there is 1 \textit{LOOKUP} constraint that was not present in the original spreadsheet (looking up ID based on Salesperson).
-
-For this example our primary goal of finding all original constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{all intended}}$) is~$1.0$.
+Moreover, there is 1 \textit{LOOKUP} constraint that was not intended in the original spreadsheet (looking up ID based on Salesperson) and which is symmetric to the intended one (Salesperson based on ID).
+
+For this example our primary goal of rediscovering all intended constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{all intended}}$) is~$1.0$.
 The implementation also returns 6 additional constraints, therefore, our precision (i.e.
-$\frac{\text{intended discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is rather low compared to, for example, the spreadsheets in the benchmark.
-The low precision can be explained by the many short vectors which increases the chance for constraints to be true by accident.
+$\frac{\text{intended discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is mostly due to the spurious RANK constraints.
+We believe this is due to the many short vectors which increases the chance for constraints to be true by accident.
 
@@
 \subsubsection*{Q1. How many intended constraints are found by \sname?}
-The overall recall achieved by \sname is~$0.88~(81/92)$.
-Similarly, the overall and average recall per category is high, as shown in Table~\ref{tbl:category_overview}.
-The recall for supported constraints is $1.00$, meaning that all supported constraints are found.
+The overall recall achieved by \sname is~$0.88~(81/92)$. The overall and average recall per category is also high, as shown in Table~\ref{tbl:category_overview}.
+The intended constraints not discovered correspond to constraints not supported by the system (supported recall is 1).
 
@@
 Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
-A qualitative analysis of these spreadsheets shows that their low precision is caused by duplicate data or multiple ways to calculate the same results.
-For these spreadsheets the precision can be increased by post-processing the constraints using entailment (to detect implications) and heuristics (e.g. taking into account the alignment of results).
+A qualitative analysis of these spreadsheets shows that their lower precision is caused by duplicate data or multiple ways to calculate the same vectors.
+For these spreadsheets the precision could be increased by post-processing the constraints to detect implications or heuristically selecting one among multiple ways to compute the same vector.
 
@@
 
-To illustrate that using dependencies to find constraints incrementally can increase the efficiency, we ran the benchmarks using foreign keys as a base constraint to find conditional aggregates.
-The total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data), showing the strong effect of this dependency.
-However, the assumption that conditional aggregates are defined for all keys in the foreign key column is too strong in practice.
-
 Of the 32 supported constraint templates (including equality), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
 It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates (each $3.5\% - 4\%$).
+
+To illustrate that using dependencies to find constraints incrementally can increase the efficiency, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF.
+This assumption is too strong in practice, not all keys have to be used in the conditions.
+However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data), showing the strong effect of the use of dependencies.
 
@@
 \subsection{Testing the limits}
-On many spreadsheets that adhere to the bias of tables, continuous blocks and vectors \sname achieves good results.
-However, factors such as incompatible formatting, bad spreadsheet design and rounding issues can lead to decreased recall, many additional constraints and long running times due to an exponential blow-up of candidate constraints.
+Our algorithm relies on a correct table and group detection step, and we noticed that exported CSV files can make this more difficult in case of rounding errors due to low-precision formatting as well as misformatting of strings so they are interpreted as numbers.
+Automatically generated sheets can also contain many duplicate (equal) rows or columns, which are a source of increased runtime and decreased recall for our algorithm.
 For example, we tested a spreadsheet (``exps'') with 6~tables, 531~cells and many duplicate columns across tables.
-While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when ignoring equalities.
-By using a (heuristic) post-processing step the precision for this example could be increased.
-The running time can be improved by implementing additional optimizations.
+While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when counting constraints with equal columns separately.
+%By using a (heuristic) post-processing step the precision for this example could be increased.
+%The running time can be improved by implementing additional optimizations.
 
@@
 \subsection{Autocompletion}
-Autocompletion can be seen as using knowledge derived from a snapshot of spreadsheet data to predict new values before they are written by the user.
+Autocompletion can be seen as using knowledge derived from the current data to predict new values before they are written by the user.
 A snapshot~($D_0$) is used to learn constraints and these are combined with new data~($\Delta D$) that the user has added.
 Using~$\Delta D$ we can keep track of which vectors have changed and which vectors are now \textit{incomplete}.
-\textit{Functional} constraints, such as $\ecsumc{\sbl{r}}{\mathbf{\sbl{x}}}$, are able to compute a result~($\sbl{r}$) based on its input~($\mathbf{\sbl{x}}$).
-If all vectors in the input blocks have changed and the output vector is incomplete, the constraint can compute the resulting value(s).
+\textit{Functional} constraints, such as $\ecsumc{\sbl{r}}{\mathbf{\sbl{x}}}$, are able to compute a result~($\sbl{r}$) based on its \textit{input}~($\mathbf{\sbl{x}}$).
+If all vectors in the input blocks have changed and the result vector is incomplete, the constraint can compute the resulting value(s).
 
@@
 \subsection{Error detection}
-\begin{figure}[thb]
+\begin{figure}[t]
 % %\begin{subfigure}{.70\textwidth}
@@
 The first setting is the \textit{online} detection of errors and is similar to the autocompletion setting.
-We look at \textit{conflict} constraints, i.e. constraints that were learned on the snapshot data but do not agree with the newly added data~$\Delta D$.
-Conflict constraints can indicate errors in the input.
-However, such conflicts can be wrongly caused by spurious constraints.
-Hence, conflicts should perhaps only be detected if the snapshot data is large enough.
+Using snapshots, we search for \textit{conflict} constraints, which are learned on a previous snapshot but do not agree with newly added data~$\Delta D$.
+Such conflicts can indicate (typing) errors in the data just entered. For example, in case 'Wes' instead of 'West' was added in Figure~\ref{fig:autocompletion_example}.
+%Conflict constraints can indicate errors in the input.
+%However, such conflicts can be wrongly caused by spurious constraints.
+%Hence, conflicts should perhaps only be detected if the snapshot data is large enough.
 %An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.
 
-The second type of error detection is \textit{offline} and attempts to detect errors in a spreadsheet.
+The second setting is \textit{offline} error detection in a static spreadsheet.
 % Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
@@
 For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
-This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g.
+Instead of snapshots, one can repeatedly search for conflict constraints between the original spreadsheet and a sheet with a row or column removed.
+Many schemes can be devised for choosing which row(s) or column(s) are removed.
+%This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g.
 leave out each row and consider it as $\Delta D_i$) to detect conflicts.
@@
 To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
-Key differences with our technique are that we assume the tables are given (a reasonable assumption if one starts from a spreadsheet), and that ModelSeeker's constraints are the typical ones from the constraint programming community, aimed at combinatorial optimization and scheduling, and restricted to integer values; furthermore we can consider multiple tables and sheets.
-Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user whether a partical constraint is satisfied or not.
-Each of the examples/queries is processed for each constraint, and hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
+Key differences are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and often typical for constraint programming community, e.g.
+aimed at combinatorial optimization and scheduling.
+Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user regarding satisfiability.
+Each of the examples/queries is processed for each constraint, hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
 for Sudoku).

### 2016-06-12 Paper/ECAI/tabular.tex
added github links

--- 2016-06-12 Paper/ECAI/tabular.tex
@@
 The above assumptions do not hold for many tables.
-However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository\footnote{\url{https://github.com/for-DSAA2016-review/TaCLe}\label{github-link}}).
+However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/for-DSAA2016-review/TaCLe}\label{github-link}}).
 Users select tables excluding the header and optionally specify an orientation.
@@
 In order to obtain a benchmark of spreadsheets we looked at three sources:
-1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports\footref{github-link}.
+1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports (available at the same repository\footref{github-link}).
 All spreadsheets were converted to CSV and the table definitions are fixed.

### 2016-06-12 Code/src/constraint_times.py
Source code changes

@@ -1,0 +1,44 @@
+times = dict()
+
+times["all-different"] = 0.00022602 + 0.00033402 + 0.00039411 + 0.00041890 + 0.00043607 + 0.00047183 + 0.00047803 + 0.00048614 + 0.00050712 + 0.00055885 + 0.00062203 + 0.00076103 + 0.00080395 + 0.00083303 + 0.00083590 + 0.00088382 + 0.00090885 + 0.00091696 + 0.00093389 + 0.00107718 + 0.00108600 + 0.00113893 + 0.00115895 + 0.00119996 + 0.00126481 + 0.00135684 + 0.00141001 + 0.00145292 + 0.00146699 + 0.00151515 + 0.00183797 + 0.00206208 + 0.00244999 + 0.00262499
+times["average (col)"] = 0.00008297 + 0.00008607 + 0.00008798 + 0.00012708 + 0.00013089 + 0.00013399 + 0.00013995 + 0.00013995 + 0.00014305 + 0.00018406 + 0.00019693 + 0.00021219 + 0.00024199 + 0.00024390 + 0.00025082 + 0.00027299 + 0.00027704 + 0.00032306 + 0.00036812 + 0.00044298 + 0.00044894 + 0.00045395 + 0.00051594 + 0.00054002 + 0.00054884 + 0.00067520 + 0.00088191 + 0.00093198 + 0.00194907 + 0.00210214 + 0.00500584 + 0.00995302 + 0.01466918 + 0.01785088
+times["average (row)"] = 0.00006986 + 0.00007486 + 0.00008488 + 0.00008702 + 0.00008988 + 0.00009298 + 0.00011206 + 0.00011301 + 0.00012898 + 0.00016999 + 0.00017905 + 0.00018620 + 0.00021005 + 0.00029707 + 0.00031209 + 0.00032616 + 0.00034308 + 0.00037193 + 0.00040889 + 0.00043511 + 0.00043917 + 0.00071716 + 0.00072098 + 0.00080800 + 0.00092697 + 0.00109506 + 0.00352812 + 0.00519800 + 0.01496005 + 0.02058411 + 0.02405596 + 0.02818799 + 0.03365207 + 0.03507900
[…]

### 2016-06-12 Code/src/core/learning.py

@@ -9,8 +9,8 @@ class LearningTask:
     def __init__(self, csv_file, groups_file, manager, constraints):
         self._manager = manager
         self._groups = list(get_groups_tables(csv_file, groups_file))
-        self._solutions = Solutions()
-        self._constraints = constraints
+        #self._solutions = Solutions()
+        self._constraints = ordered_constraints(constraints)
         self._times = dict()
 
     @property
@@ -20,10 +20,6 @@ class LearningTask:
     @property
     def manager(self):
         return self._manager
-
-    @property
-    def solutions(self):
-        return self._solutions
 
     def run(self):
         supported = []
@@ -47,12 +43,13 @@ class LearningTask:
             print_concise.nl()
 
         print_verbose = printing.get("detail", print_concise, on=False)
-        for constraint in self.ordered_constraints():
+        solutions = Solutions()
+        for constraint in self.constraints:
             print_verbose.write(constraint.name, end=" ")
             t_start = time.time()
-            assignments = self.manager.find_assignments(constraint, self._groups, self.solutions)
+            assignments = self.manager.find_assignments(constraint, self._groups, solutions)
             t_assign = time.time()
-            self.solutions.add(constraint, self.manager.find_solutions(constraint, assignments, self.solutions))
+            solutions.add(constraint, self.manager.find_solutions(constraint, assignments, solutions))
             t_end = time.time()
             assignment_time = t_assign - t_start
             solving_time = t_end - t_assign
@@ -60,13 +57,14 @@ class LearningTask:
             if print_verbose.on():
                 print_verbose.form("[assignment time: {assign:.3f}, solving time: {solve:.3f}]",
                                    assign=assignment_time, solve=solving_time)
-            if len(self.solutions.get_solutions(constraint)) > 0:
+            if len(solutions.get_solutions(constraint)) > 0:
                 print_concise.write(lambda: "\n".join(["\t" + constraint.to_string(s)
-                                                       for s in self.solutions.get_solutions(constraint)]))
-            if len(self.solutions.get_solutions(constraint)) > 0 or print_verbose.on():
+                                                       for s in solutions.get_solutions(constraint)]))
+            if len(solutions.get_solutions(constraint)) > 0 or print_verbose.on():
                 print_concise.nl()
 
         print_verbose.write(lambda: "Total: {0:.3f}".format(self.total_time()))
+        return solutions
 
     def total_time(self):
         return sum(self.time(c) for c in self.constraints)
@@ -75,13 +73,14 @@ class LearningTask:
[…]

### 2016-06-12 Code/src/engine/internal.py

@@ -82,7 +82,8 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             return self._generate_test_vectors(assignments, [c.x], test_list)
 
         def all_different(c: AllDifferent, assignments, solutions):
-            return self._generate_test_vectors(assignments, [c.x], lambda v: len(set(v)) == len(v))
+            return self._generate_test_vectors(assignments, [c.x], lambda v: len(set(v)) == len(v)
+                                               , lambda g: g.is_textual() or g.is_integer())
 
         def permutation(c: Permutation, assignments, solutions):
             def test_set(vector):

### 2016-06-12 Code/src/experiment.py

@@ -98,11 +98,11 @@ class Experiment:
 
     def run(self):
         if not self._run:
-            t_before = time.time()
             csv_file, config_file = self._get_csv_file(), self._get_config_file()
-            solutions = workflow.main(csv_file, config_file, False, silent=True, parse_silent=True)
-            self._running_times.append(time.time() - t_before)
-
+            learning_task = workflow.task(csv_file, config_file)
+            solutions = learning_task.run()
+#
+            self._running_times.append(learning_task.total_time())
             self.count_constraints(solutions)
             self.measure_size(config_file)
         self._run = True

### 2016-06-12 Code/src/parse/parser.py

@@ -5,10 +5,14 @@ import numpy as np
 import pandas as pd
 import re
 
+import time
+
 from core.group import Bounds, Table, Group, Orientation, GType
 
 
 # TODO Single vector => try both orientations
+from util import printing
+
 
 def parse(filename):
     with open(filename, "r") as file:
@@ -84,20 +88,21 @@ def infer_type(types):
     return GType.float
 
 
-def get_groups_tables(csv_file, groups_file=None, silent=False):
+def get_groups_tables(csv_file, groups_file=None):
+    parse_printer = printing.get(__name__, on=False)
+    t_start = time.time()
     data = parse(csv_file)
     type_data = np.vectorize(detect_type)(data)
     if groups_file is None:
         t = list(detect_tables(type_data))
         t = [(b, Table("T{}".format(i + 1), Bounds(b).subset(data), o)) for i, (b, o) in enumerate(t)]
-        if not silent:
-            print("PARSE: Detected tables: {}".format(", ".join(["{} = [{}:{}, {}:{}]".format(table.name, *bounds)
-                                                                 for bounds, table in t])))
+
+        if parse_printer.on():
+            tables = ["{} = [{}:{}, {}:{}]".format(table.name, *bounds) for bounds, table in t]
+            parse_printer.form("PARSE: Detected tables: {}", ", ".join(tables))
+
         groups = detect_groups(type_data, t)
-        if not silent:
-            print("PARSE: Detected groups: {}".format(", ".join(str(g) for g in groups)))
-            print()
-        return groups
+        parse_printer.form("PARSE: Detected groups: {}", ", ".join(str(g) for g in groups))
     else:
         table_dict = {}
         t = []
@@ -123,10 +128,9 @@ def get_groups_tables(csv_file, groups_file=None, silent=False):
                     groups.append(create_group(group_description["Bounds"], table))
             else:
                 groups = detect_groups(type_data, t)
-                if not silent:
-                    print("PARSE: Detected groups: {}".format(", ".join(str(g) for g in groups)))
-                    print()
-        return groups
+                parse_printer.form("PARSE: Detected groups: {}", ", ".join(str(g) for g in groups))
+    parse_printer.form("PARSE: Parsing took {}s", time.time() - t_start)
+    return groups
 
 
 def create_group(bounds_list, table: Table):

### 2016-06-12 Code/src/playground.py

@@ -4,56 +4,59 @@ from pandas import json
 from draw import Plotter
 from engine.util import local
 from experiment import Experiment, ConstraintCounter
+from util import printing
 
 
 def main():
+    info = [
+        ("Tables", lambda e: e.tables, "grey", lambda x: str(x)),
+        # ("Vectors", lambda e: e.vectors, "cyan", lambda x: str(x)),
+        ("Cells", lambda e: e.cells, "black", lambda x: str(x)),
+        ("Intended Constraints", lambda e: e.counter.count(relevant=True), "magenta", lambda x: str(x)),
+        ("Recall", lambda e: recall(e.counter), "green", lambda x: "{:.2f}".format(x)),
+        ("Recall Supported", lambda e: recall(e.counter, True), "lightgreen", lambda x: "{:.2f}".format(x)),
+        ("Precision", lambda e: precision(e.counter), "yellow", lambda x: "{:.2f}".format(x)),
+        ("Runtime", lambda e: numpy.average(e.running_times(1)), "red", lambda x: "{:.2f}s".format(x))
+    ]
+
     with open(local("data/data.txt")) as data_file:
         data_json = json.load(data_file)
         categories = ["Exercises", "Tutorials", "Data"]
-        log = {}
-        for category in data_json:
-            name = category["Category"]
-            files = category["Files"]
-            log[name] = []
+        stats = {}
+        for category in categories:
+            files = data_json[category]
+            stats[category] = []
             for file in files:
                 experiment = Experiment(file)
                 print(file)
                 for accident in experiment.counter.additional:
                     print("\tADDITIONAL:\t{}".format(accident))
-                log[name].append(experiment)
+                entry = {t[0]: t[1](experiment) for t in info}
+                entry["relevant.found"] = experiment.counter.count(relevant=True, found=True)
+                entry["extra.found"] = experiment.counter.count(relevant=False, found=True)
+                stats[category].append(entry)
 
-        info = [
-            ("Tables", lambda e: e.tables, "grey", lambda x: str(x)),
-            #("Vectors", lambda e: e.vectors, "cyan", lambda x: str(x)),
-            ("Cells", lambda e: e.cells, "black", lambda x: str(x)),
-            ("Relevant Constraints", lambda e: e.counter.count(relevant=True), "magenta", lambda x: str(x)),
-            ("Recall", lambda e: recall(e.counter), "green", lambda x: "{:.2f}".format(x)),
[…]

### 2016-06-12 Code/src/workflow.py

@@ -2,6 +2,7 @@ import argparse
 import time
 
 from core.constraint import *
+from core.learning import LearningTask
 from core.solutions import Solutions
 from core.strategy import StrategyManager
 from engine.idp import IdpAssignmentStrategy
@@ -9,26 +10,29 @@ from engine.internal import InternalCSPStrategy, InternalSolvingStrategy
 from engine.minizinc import MinizincAssignmentStrategy, MinizincSolvingStrategy
 from parse.parser import get_groups_tables
 
-constraint_list = [
-    # EqualGroup(),
-    Permutation(),
-    Series(),
-    AllDifferent(),
-    Projection(),
-    Rank(),
-    ForeignKey(),
-    Lookup(),
-    FuzzyLookup(),
-    RunningTotal(),
-    ForeignProduct(),
-    Product(),
-    Diff(),
-    PercentualDiff(),
-    SumProduct(),
-    Equal(),
-]
-constraint_list += Aggregate.instances()
-constraint_list += ConditionalAggregate.instances()
+
+def get_constraint_list():
+    constraint_list = [
+        # EqualGroup(),
+        Permutation(),
+        Series(),
+        AllDifferent(),
+        Projection(),
+        Rank(),
+        ForeignKey(),
+        Lookup(),
+        FuzzyLookup(),
+        RunningTotal(),
+        ForeignProduct(),
+        Product(),
+        Diff(),
+        PercentualDiff(),
+        SumProduct(),
+        Equal(),
+    ]
+    constraint_list += Aggregate.instances()
+    constraint_list += ConditionalAggregate.instances()
+    return constraint_list
 
 
 def order_constraints(constraints: List[Constraint]):
@@ -49,9 +53,16 @@ def order_constraints(constraints: List[Constraint]):
     return ordered
 
 
-def main(csv_file, groups_file, verbose, silent=False, parse_silent=False, constraints=constraint_list):
+def task(csv_file, groups_file, constraints=None):
+    if constraints is None:
+        constraints = get_constraint_list()
+    return LearningTask(csv_file, groups_file, get_manager(), constraints)
+
+
+def main(csv_file, groups_file, verbose, silent=False, constraints=None):
+    print("I AM BEING RUN")
     manager = get_manager()
-    groups = list(get_groups_tables(csv_file, groups_file, silent=parse_silent))
+    groups = list(get_groups_tables(csv_file, groups_file))
 
     solutions = Solutions()
[…]

### 2016-08-29 Paper/ECAI/tabular.tex
formula -> formulas
switched to ecml pkdd format; it got messed up a bit

@@
-\documentclass{IEEEtran}
-
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{latexsym}
-\usepackage{xspace}
-\usepackage{hyperref}
-\usepackage{amssymb}
-\usepackage{algorithm}
-\usepackage[noend]{algpseudocode}
-\usepackage[numbers]{natbib}
-\usepackage{notoccite}
-\usepackage{fixltx2e}
-\usepackage{framed}
-\usepackage{amsmath}
-\usepackage{tabularx}
-\usepackage[margin=0.1cm]{subcaption}
-\usepackage[inline]{enumitem}
-\usepackage{footmisc}
-
-\usepackage[dvipsnames]{xcolor}
-\newcommand{\PADSEP}{-7pt}
-\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
-\newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
-\newcommand{\tias}[1]{\textcolor{blue}{{\sc Tias:} #1}\xspace}
-\newcommand{\luc}[1]{\textcolor{red}{{\sc Luc:} #1}\xspace}
-
-\newcommand{\constraints}{\ensuremath{\mathcal{T}}\xspace}
-\newcommand{\format}[1]{\textit{#1}\xspace}
-\newcommand{\generategroups}{\format{SuperblockAssignments}}
-\newcommand{\extractgroups}{\format{extractGroups}}
-\newcommand{\extracttables}{\format{extractTables}}
-\newcommand{\learnconstraints}{\format{LearnConstraints}}
-\newcommand{\findassignment}{\format{Subassignments}}
-\newcommand{\postprocess}{\format{pruneRedundant}}
-\newcommand{\constrainttorder}{\format{templateOrder}}
-\newcommand{\template}{\format{constraint template}}
-\newcommand{\sname}{\format{TaCLe}}
-
-\newcommand{\CName}{Syntax\xspace}
-\newcommand{\CSignature}{Signature\xspace}
-\newcommand{\CFunction}{Definition\xspace}
-\newcommand{\dependencies}{\ensuremath{\mathcal{D}}\xspace}
-\newcommand{\groups}{\ensuremath{\mathcal{B}}\xspace}
-\newcommand{\blocks}{\ensuremath{\mathcal{B}}\xspace}
-
-\newcommand{\range}[3]{\ensuremath{#1[#2,#3]}}
-\newcommand{\rangeto}[2]{#1{:}#2}
-\newcommand{\rangeall}{:}
-
-\newcommand{\eccalc}[2]{\ensuremath{#1 = #2}}
-\newcommand{\ecrank}[2]{\eccalc{#1}{\textit{RANK}(#2)}}
-\newcommand{\ecfkey}[2]{\ensuremath{\textit{FOREIGNKEY}(#1,#2)}}
-\newcommand{\ecalldiff}[1]{\ensuremath{\textit{ALLDIFFERENT}(#1)}}
-\newcommand{\eclookupf}[4]{\ensuremath{\textit{LOOKUP}_{\textit{#4}}(#1, #2, #3)}}
-\newcommand{\eclookup}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{}}}
-\newcommand{\eclookupprod}[5]{\eccalc{#1}{#2 \times \eclookupf{#3}{#4}{#5}{}}}
-\newcommand{\eclookupfuzzy}[4]{\eccalc{#1}{\eclookupf{#2}{#3}{#4}{fuzzy}}}
-\newcommand{\ecperm}[1]{\ensuremath{\textit{PERMUTATION}(#1)}}
-\newcommand{\ecseries}[1]{\ensuremath{\textit{SERIES}(#1)}}
-\newcommand{\ecprod}[3]{\eccalc{#1}{#2 \times #3}}
-\newcommand{\ecdiv}[3]{\eccalc{#1}{{#2} / {#3}}}
-\newcommand{\ecdiff}[3]{\eccalc{#1}{#2 - #3}}
-\newcommand{\ectotal}[3]{\eccalc{#1}{\textit{PREV}(#1) + #2 - #3}}
-\newcommand{\ecproj}[2]{\eccalc{#1}{\textit{PROJECT}(#2)}}
-\newcommand{\ecaggc}[3]{\eccalc{#2}{\textit{#1\textsubscript{col}}(#3)}}
-\newcommand{\ecaggr}[3]{\eccalc{#2}{\textit{#1\textsubscript{row}}(#3)}}
-\newcommand{\ecsumc}[2]{\eccalc{#1}{\textit{SUM\textsubscript{col}}(#2)}}
-\newcommand{\ecsumr}[2]{\eccalc{#1}{\textit{SUM\textsubscript{row}}(#2)}}
-\newcommand{\ecaggif}[5]{\eccalc{#2}{\textit{#1IF}(#3, #4, #5)}}
-\newcommand{\ecsumif}[4]{\eccalc{#1}{\textit{SUMIF}(#2, #3, #4)}}
-\newcommand{\ecsumprod}[3]{\eccalc{#1}{\textit{SUMPRODUCT}(#2, #3)}}
-
-\newcommand{\numeric}{\format{numeric}}
-\newcommand{\textual}{\format{textual}}
-\newcommand{\integer}{\format{integer}}
-\newcommand{\discrete}{\format{discrete}}
-\newcommand{\plength}{\format{length}}
-\newcommand{\psize}{\format{size}}
-\newcommand{\ptype}{\format{type}}
-\newcommand{\ptable}{\format{table}}
-\newcommand{\por}{\format{orientation}}
-\newcommand{\prows}{\format{rows}}
-\newcommand{\pcols}{\format{columns}}
-\newcommand{\nat}{\mathcal{N}}
-
-\newcommand{\sbs}{B}
-\newcommand{\sbl}[1]{\ensuremath{\sbs_{\textit{#1}}}}
-\newcommand{\ssbl}[1]{\ensuremath{\sbs'_{\textit{#1}}}}
-\newcommand{\bsbl}[1]{\ensuremath{\mathbf{\sbs_{\textit{#1}}}}}
-
-\renewcommand{\arraystretch}{1.5}
-
-\usepackage{amsthm} % incompatible with ACM
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-\newtheorem{example}{Example}
-%\newdef{definition}{Definition} % ACM specific
+\input{journal_header}
+\input{packages}
+\input{macros}
 
@@
 
-%\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
+\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
 
@@
 However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
-In this work, we investigate the automatic discovery of constraints (formula and relations) in raw tabular data.
+In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data.
 We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.

### 2016-08-29 Paper/ECAI/tabular.tex
extended the related work section

@@
 The focus is again on integer variables and typical combinatorial optimisation constraints.
+\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
+It also connects with the approach of inducing constraints and functions by generalizing given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
+The key difference here that we start not from existing formulas but from the data directly.
 %As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
@@
 
+\sergey{added database dependency discovery; have a look at it} Fourth, our work is motivated by the functional dependency discover in databases \cite{flach_dependency_discovery}.
+Most of the discovery techniques rely on the database schema that does not directly apply in our case.
+Some techniques work directly on the data \cite{tane_dependency_discovery} and, contrary to our work, are focused on pure functional dependencies that are not typical constraints of a spreadsheet that might hold over multiple tables and on a mixture of columns and rows.
+
 Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
@@
 \bibliographystyle{plain}
-{\footnotesize
+{
   \bibliography{references}}

### 2016-08-29 Code/src/engine/internal.py
Removed recursion max-range

@@ -20,15 +20,18 @@ class MaxRange:
         self._find(start, end, size, limit, last)
 
     def _find(self, start, end, size, limit, last):
-        if last - start < size:
-            pass
-        elif end - start < size or end <= limit:
-            self._find(start + 1, last, size, max(start + size, limit), last)
-        else:
-            if self._test(start, end):
-                self._find(start + 1, last, size, end, last)
+        while True:
+            if last - start < size:
+                return
+            elif end - start < size or end <= limit:
+                start += 1
+                end = last
+                limit = max(start + size, limit)
             else:
-                self._find(start, end - 1, size, limit, last)
+                if self._test(start, end):
+                    start += 1
+                else:
+                    end -= 1
 
 
 # TODO Sum IF with Nones

### 2016-08-29 Code/src/core/constraint.py
Remember hash and added ordered

@@ -204,6 +204,16 @@ class AllDifferent(Constraint):
         super().__init__("all-different", "ALLDIFFERENT({X})", Source(variables), filters)
 
 
+class Ordered(Constraint):
+    x = Variable("X", types=numeric)
+
+    def __init__(self):
+        variables = [self.x]
+        source = Source(variables)
+        filters = [NotPartial(variables)]
+        super().__init__("ordered", "ORDERED({X})", source, filters)
+
+
 class Rank(Constraint):
     x = Variable("X", vector=True, types=numeric)
     y = Variable("Y", vector=True, types=integer)
@@ -250,7 +260,7 @@ class FuzzyLookup(Constraint):
 
     def __init__(self):
         variables = [self.o_key, self.o_value, self.f_key, self.f_value]
-        source = Source(variables)
+        source = ConstraintSource(variables, Ordered(), {Ordered.x.name: self.o_key.name})
         filters = [SameType([self.o_value, self.f_value]), NotPartial(variables),
                    SameLength([self.f_key, self.f_value]), SameLength([self.o_key, self.o_value]),
                    SameTable([self.f_key, self.f_value]), SameTable([self.o_key, self.o_value]),

### 2016-08-29 Code/src/core/group.py

@@ -141,6 +141,7 @@ class Group:
         from core.constraint import blank_filter
         self._is_partial = not numpy.all(numpy.vectorize(blank_filter(self._data)[1])(self._data))
         self._subgroups = dict()
+        self._hash = None
 
     @property
     def is_partial(self):
@@ -247,7 +248,9 @@ class Group:
         return self.subgroup(Bounds(l))
 
     def __hash__(self):
-        return hash((self.table, self.row, self.bounds))
+        if self._hash is None:
+            self._hash = hash((self.table, self.row, self.bounds))
+        return self._hash
 
     def __ne__(self, other):
         return not self == other

### 2016-08-29 Code/src/engine/internal.py

@@ -60,6 +60,7 @@ class InternalCSPStrategy(AssignmentStrategy):
         self.add_constraint(Diff())
         self.add_constraint(PercentualDiff())
         self.add_constraint(SumProduct())
+        self.add_constraint(Ordered())
 
     def add_constraint(self, constraint: Constraint):
         self._constraints.add(constraint)
@@ -207,6 +208,7 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                 vectors = {g: g.get_vector(1) for g in [ok, r, fk, v]}
 
                 for g in [ok, r, v]:
+                    # FIXME avoid all call
                     if g not in partial_cache:
                         partial_cache[g] = all(numpy.vectorize(blank_filter(vectors[g])[1])(vectors[g]))
                     if not partial_cache[g]:
@@ -434,8 +436,12 @@ class InternalSolvingStrategy(DictSolvingStrategy):
 
                 # Equal vectors, update cache:
                 if y_v in equal_map:
-                    raise RuntimeError("Violated assumption")
-                equal_map[y_v] = x_v
+                    found = equal_map[y_v]
+                    minimal, maximal = (x_v, found) if x_v < found else (found, x_v)
+                    equal_map[y_v] = minimal
+                    equal_map[maximal] = minimal
+                else:
+                    equal_map[y_v] = x_v
                 return True
 
             return self._generate_test_vectors(assignments, [c.first, c.second], lambda *args: True, test)
@@ -455,6 +461,15 @@ class InternalSolvingStrategy(DictSolvingStrategy):
                 max_range = MaxRange(test)
                 max_range.find(0, x.vectors(), 2)
             return result
+
+        def ordered_constraint(c: Ordered, assignments, solutions):
+            def test_ordering(x):
+                for i in range(1, len(x)):
+                    if x[i] <= x[i - 1]:
+                        return False
+                return True
+
+            return self._generate_test_vectors(assignments, [c.x], test_ordering)
 
         self.add_strategy(Equal(), equality)
         self.add_strategy(EqualGroup(), equal_group)
@@ -476,9 +491,11 @@ class InternalSolvingStrategy(DictSolvingStrategy):
         self.add_strategy(Diff(), diff)
         self.add_strategy(PercentualDiff(), percent_diff)
         self.add_strategy(SumProduct(), sum_product)
+        self.add_strategy(Ordered(), ordered_constraint)
 
     @staticmethod
[…]

### 2016-08-29 Code/src/workflow.py

@@ -29,6 +29,7 @@ def get_constraint_list():
         Diff(),
         PercentualDiff(),
         SumProduct(),
+        Ordered(),
     ]
     constraint_list += Aggregate.instances()
     constraint_list += ConditionalAggregate.instances()

### 2016-08-30 Paper/ECAI/tabular.tex
fixed the running example figure

@@
 
-\begin{figure*}[thb]
-
-  \begin{subfigure}{.66\textwidth}
+\begin{figure}[thb]
+
+  \begin{subfigure}{1\textwidth}
   \begin{center}
@@
     However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
-    This example was created by combining several Excel sheets from several exercise sessions based on~\cite{excel_book}.
+    This example was created by combining several Excel sheets from several exercise sessions based on the book MS Excel 2010~\cite{excel_book}.
     % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
@@
 \hfill
-\begin{subfigure}{.34\textwidth}
-  {\footnotesize
+\begin{subfigure}{1\textwidth}
+  {\scriptsize
     \begin{align*}
       % Table 1
-      &~\ecseries{\range{T_{1}}{\rangeall}{1}} \\
 %
@@
 %
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^*\\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^* \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^*\\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*\\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^* \\
-%
-      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}} \\
-%
-      &~\ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}} \\
+      &~\ecseries{\range{T_{1}}{\rangeall}{1}}                                  && \ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^*     && \ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}}  \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^*     && \ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^*    && \ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}       && \ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*     && \range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^*    && \eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^*\\
+%
+      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}}   && \eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
+%
+      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}}   && \ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}}
 %
 	  % Table 2
-      &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-      &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
-%
-      &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%     &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+%     &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+%     &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
+%
+%     &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
 %
 	  % Table 4
-      &~\ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-      &~\range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
+%     &~\ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+%     &~\range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
 %
     % Table 5
-      &~\eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^* \\
-%
-      &~\eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}
+%     &~\eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^* \\
+%
+%     &~\eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}
     \end{align*}}
-  \vspace{-18pt}
+  \vspace{-10pt}
   \caption{Constraints learned for the example (left), except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
@@
   \caption{Running example}
-\end{figure*}
+\end{figure}
 
@@
 The above assumptions do not hold for many tables.
-However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/for-DSAA2016-review/TaCLe}\label{github-link}}).
+However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/SergeyParamonov/TaCLe}\label{github-link}}).
 Users select tables excluding the header and optionally specify an orientation.

### 2016-08-31 Paper/ECAI/tabular.tex
rewritten related work section a bit

@@
 \begin{abstract}
+  \sergey{Samuel! Add a reference to python-constraint somewhere, including the references list at the end}\\
+  \sergey{Tias, Samuel, Luc, should we even mention miniZinc or ASP here, or is it not relevant?}
 Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
@@
 
-\begin{table*}[!h]
+\begin{table}[htb]
 \caption{
@@
   Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows \textit{SUM} but \sname also supports \textit{MAX}, \textit{MIN}, \textit{AVERAGE}, \textit{PRODUCT} and \textit{COUNT}.
-  %{\small\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
+  %{\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
 }
 \label{table:constraints}
-  {\centering
-  \begin{tabularx}{\textwidth}{l X X}
+\resizebox{1.0\columnwidth}{!}{%
+  {\normalsize \centering
+  \begin{tabularx}{1.70\textwidth}{l X X}
     \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
@@
 
-      & \[ \sbl{r}[i] = \sum_{j=1}^{\plength(\sbl{val})} \begin{cases}
+      & \[\sbl{r}[i] = \sum_{j=1}^{\plength(\sbl{val})} \begin{cases}
           \sbl{val}[j] & \text{if } \sbl{fk}[j] = \sbl{pk}[i] \\
@@
       & $\sbl{r}[1] = \sum_{j = 1}^{\plength(\sbl{1})} \sbl{1}[j] \times \sbl{2}[j]$.
-  \end{tabularx}}
-
-\end{table*}
+    \end{tabularx}
+  } % centering small
+} % resizebox
+
+\end{table}
 
@@
 
-\begin{algorithm}[tbh]
+\begin{algorithm}[t]
   \begin{algorithmic}
@@
   \centering
-  \includegraphics[width=0.65\linewidth]{figures/constraint_dependency2.pdf}
+  \includegraphics[width=0.50\linewidth]{figures/constraint_dependency2.pdf}
   \caption{Dependency graph of Table~\ref{table:constraints}; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (its signature includes $s_1$).
@@
 \begin{center}
-    \includegraphics[width=0.65\linewidth]{figures/Learning.png}
+    \includegraphics[width=1.0\linewidth]{figures/Learning.png}
   \end{center}
@@
 approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
-Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
+Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
 Nevertheless, several ideas from logical and relational learning
 proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
-Also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha}
-although that line of research has focused on more specialized techniques for specific types of constraints. Furthermore, the constraints and formulas used
-in spreadsheets are often different from those in databases.
+\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
+It also connects with the logical generalization of given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
+The key difference here that we start not from existing formulas but from the data directly.
 
@@
 The focus is again on integer variables and typical combinatorial optimisation constraints.
-\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
-It also connects with the approach of inducing constraints and functions by generalizing given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
-The key difference here that we start not from existing formulas but from the data directly.
 %As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
@@
 
-\sergey{added database dependency discovery; have a look at it} Fourth, our work is motivated by the functional dependency discover in databases \cite{flach_dependency_discovery}.
-Most of the discovery techniques rely on the database schema that does not directly apply in our case.
-Some techniques work directly on the data \cite{tane_dependency_discovery} and, contrary to our work, are focused on pure functional dependencies that are not typical constraints of a spreadsheet that might hold over multiple tables and on a mixture of columns and rows.
+\sergey{extended database dependency discovery section; have a look at it} Fourth, also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha} although that line of research has focused on more specialized techniques for specific types of constraints.
+Many of of the discovery techniques rely on the database schema that does not directly apply in our case \cite{flach_dependency_discovery}.
+Some work directly on the data \cite{tane_dependency_discovery} and focus on classical functional dependencies between attributes, while the constraints and formulas used in spreadsheets are often different from those in databases.
 

### 2016-08-31 Paper/ECAI/tabular.tex
fixed one table witdh

@@
   \label{tbl:category_overview}
-  \begin{tabularx}{\linewidth}{lccccccc}
+  \begin{tabular}{lccccccc}
  & \multicolumn{2}{c}{\textbf{Exercises} (9)} & \multicolumn{2}{c}{\textbf{Tutorials} (21)} & \multicolumn{2}{c}{\textbf{Data} (4)} \\
@@
 \textbf{Speed (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
-\end{tabularx}
+\end{tabular}
 \end{table}}

### 2016-08-31 Paper/ECAI/tabular.tex
moved related work

@@
   \sergey{Samuel! Add a reference to python-constraint somewhere, including the references list at the end}\\
-  \sergey{Tias, Samuel, Luc, should we even mention miniZinc or ASP here, or is it not relevant?}
 Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
@@
 Section \ref{sec:formalization} introduces concepts relevant to our approach and our problem statement.
+Section \ref{sec:related_work} presents a more detailed discussion of the related work.
 Section \ref{sec:approach} presents our approach.
@@
 Section \ref{sec:applications} shows how our system can be used for applications.
-Section \ref{sec:related_work} presents a more detailed discussion of the related work.
 Section \ref{sec:conclusions} provides conclusions.
@@
 %}
+\section{Related Work}\label{sec:related_work}
+\sname combines ideas from several existing
+approaches and different fields of research.
+
+First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
+constraints in multiple relations (or tables). However, unlike typical logical and relational learning
+approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
+Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
+Nevertheless, several ideas from logical and relational learning
+proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
+\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
+It also connects with the logical generalization of given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
+The key difference here that we start not from existing formulas but from the data directly.
+
+Second, there exist a number of algorithms in the constraint programming community that induce constraint (programs)
+from one or more examples and questions. Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
+The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
+To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
+Key differences are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and often typical for constraint programming community, e.g.
+aimed at combinatorial optimization and scheduling.
+Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user regarding satisfiability.
+Each of the examples/queries is processed for each constraint, hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
+for Sudoku).
+The focus is again on integer variables and typical combinatorial optimisation constraints.
+%As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
+As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.
+
+Third, our work borrows from the seminal work of Gulwani et al.
+\cite{flashfill} on program synthesis for tabular data, which is incorporated in Microsoft’s Excel.
+In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians.
+The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''.
+At that point FlashFill generates a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', and also applies it to the other politicians (rows) in the table.
+While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
+Flashfill has been extended to e.g.
+FlashExtract \cite{flashextract} to produce a set of tables starting from an unstructured document (a txt or HTML file) with a set of examples marked for extraction.
+However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.
+In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.
+Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
+
+\sergey{extended database dependency discovery section; have a look at it} Fourth, also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha} although that line of research has focused on more specialized techniques for specific types of constraints.
+Many of of the discovery techniques rely on the database schema that does not directly apply in our case \cite{flach_dependency_discovery}.
+Some work directly on the data \cite{tane_dependency_discovery} and focus on classical functional dependencies between attributes, while the constraints and formulas used in spreadsheets are often different from those in databases.
+
+Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
+However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify ``hard'' constraints.
 
@@
 
-\section{Related Work}\label{sec:related_work}
-\sname combines ideas from several existing
-approaches and different fields of research.
-
-First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
-constraints in multiple relations (or tables). However, unlike typical logical and relational learning
-approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
-Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
-Nevertheless, several ideas from logical and relational learning
-proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
-\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
-It also connects with the logical generalization of given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
-The key difference here that we start not from existing formulas but from the data directly.
-
-Second, there exist a number of algorithms in the constraint programming community that induce constraint (programs)
-from one or more examples and questions. Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
-The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
-To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
-Key differences are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and often typical for constraint programming community, e.g.
-aimed at combinatorial optimization and scheduling.
-Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user regarding satisfiability.
-Each of the examples/queries is processed for each constraint, hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
-for Sudoku).
-The focus is again on integer variables and typical combinatorial optimisation constraints.
-%As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
-As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.
-
-Third, our work borrows from the seminal work of Gulwani et al.
-\cite{flashfill} on program synthesis for tabular data, which is incorporated in Microsoft’s Excel.
-In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians.
-The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''.
-At that point FlashFill generates a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', and also applies it to the other politicians (rows) in the table.
-While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
-Flashfill has been extended to e.g.
-FlashExtract \cite{flashextract} to produce a set of tables starting from an unstructured document (a txt or HTML file) with a set of examples marked for extraction.
-However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.
-In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.
-Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
-
-\sergey{extended database dependency discovery section; have a look at it} Fourth, also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha} although that line of research has focused on more specialized techniques for specific types of constraints.
-Many of of the discovery techniques rely on the database schema that does not directly apply in our case \cite{flach_dependency_discovery}.
-Some work directly on the data \cite{tane_dependency_discovery} and focus on classical functional dependencies between attributes, while the constraints and formulas used in spreadsheets are often different from those in databases.
-
-Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
-However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify ``hard'' constraints.
-
 \section{Conclusions}\label{sec:conclusions}

### 2016-08-31 Paper/ECAI/tabular.tex
added equation discovery

@@
 \section{Related Work}\label{sec:related_work}
-\sname combines ideas from several existing
-approaches and different fields of research.
+\sname combines ideas from several existing approaches and different fields of research.
 
@@
 The focus is again on integer variables and typical combinatorial optimisation constraints.
+\sergey{added equation discover here, check if that's the right place for it and if it is in fact good enough description} In a similar spirit the field of equation discovery \cite{equation_discovery} focuses, guided by the heuristics based on the structure of the physical equations and unit measures, on the discovery of physical formulas over numbers and times series, but not on spreadsheet constraints.
 %As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.

### 2016-09-02 Paper/ECAI/tabular.tex

--- 2016-08-31 Paper/ECAI/tabular.tex
@@
   \label{tbl:category_overview}
-  \begin{tabular}{lccccccc}
+  \begin{tabularx}{\linewidth}{lccccccc}
  & \multicolumn{2}{c}{\textbf{Exercises} (9)} & \multicolumn{2}{c}{\textbf{Tutorials} (21)} & \multicolumn{2}{c}{\textbf{Data} (4)} \\
@@
 \textbf{Speed (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
-\end{tabular}
+\end{tabularx}
 \end{table}}
@@
 approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
-Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
+Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
 Nevertheless, several ideas from logical and relational learning
 proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
-\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
-It also connects with the logical generalization of given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
-The key difference here that we start not from existing formulas but from the data directly.
+Also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha}
+although that line of research has focused on more specialized techniques for specific types of constraints. Furthermore, the constraints and formulas used
+in spreadsheets are often different from those in databases.
 
@@
 
-\sergey{extended database dependency discovery section; have a look at it} Fourth, also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha} although that line of research has focused on more specialized techniques for specific types of constraints.
-Many of of the discovery techniques rely on the database schema that does not directly apply in our case \cite{flach_dependency_discovery}.
-Some work directly on the data \cite{tane_dependency_discovery} and focus on classical functional dependencies between attributes, while the constraints and formulas used in spreadsheets are often different from those in databases.
+\sergey{added database dependency discovery; have a look at it} Fourth, our work is motivated by the functional dependency discover in databases \cite{flach_dependency_discovery}.
+Most of the discovery techniques rely on the database schema that does not directly apply in our case.
+Some techniques work directly on the data \cite{tane_dependency_discovery} and, contrary to our work, are focused on pure functional dependencies that are not typical constraints of a spreadsheet that might hold over multiple tables and on a mixture of columns and rows.
 

### 2016-09-07 Paper/ECAI/tabular.tex
add contributions, rework abstract

--- 2016-08-31 Paper/ECAI/tabular.tex
@@
 
-\title{Learning constraints in spreadsheets and tabular data}
+\title{Discovering constraints in spreadsheets and tabular data}
 % see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
 
-\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
+\author{} % \institute{----------------------} }
 
@@
 \begin{abstract}
-  \sergey{Samuel! Add a reference to python-constraint somewhere, including the references list at the end}\\
 Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
-However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
-In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data.
-We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
-Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
-Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
-Constraint learning techniques are used to identify the predicates that hold between blocks of rows and columns of the tables.
-We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
+However, writing, maintaining and identifying good formulas for tabular data and spreadsheets can be time consuming and error-prone.
+In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data in an unsupervised way.
+Common spreadsheet formulas are represented as predicates whose arguments must satisfy the inherent properties of the function.
+The challenge is to discover these without user information or training examples.
+% Knowledge discovery
+We propose a two-stage generate and test method where the first stage uses constraint solving techniques to efficiently reduce candidates in a generic way based on the function signature. 
+Our approach to this knowledge discovery task takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
+We show that we are able to accurately discover constraints in spreadsheets from various sources.
 \end{abstract}
@@
 
+Our contributions are as follows:
+\begin{itemize}
+\item we define the tabular constraint learning problem, where the goal is to learn constraints (functions and relations) that range over entire rows or columns.
+This is an unsupervised learning problem;
+\item we propose a 2-stage generate and test method that reasons for each candidate constraint over the maximally contiguous submatrices in the first stage, and only in the second stage looks at individual rows and columns and their content;
+\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all tuples of submatrices compatible with the candidate constraint its arguments, using backtracking search and forward checking;
+\item experiments on a range of publicly available spreadsheets show that the system is able to extract constraints with high precision and recall in an unsupervised manner. 
+\end{itemize}
+
 The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
@@
 This paper is organized as follows.
+Section \ref{sec:related_work} presents a detailed discussion of the related work.
 Section \ref{sec:formalization} introduces concepts relevant to our approach and our problem statement.
-Section \ref{sec:related_work} presents a more detailed discussion of the related work.
 Section \ref{sec:approach} presents our approach.
@@
 \paragraph{Visual}
-The above assumptions do not hold for many tables.
+\sergey{Luc, can you check if the supplementary material at github is understandable for people?
+Do we have to add anything there like write that code would be available after?} The above assumptions do not hold for many tables.
 However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/SergeyParamonov/TaCLe}\label{github-link}}).
@@
 All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
+\sergey{Samuel! Add a reference to python-constraint, including the references list at the end}\\
 

### 2016-09-07 Paper/ECAI/tabular.tex
Updated precision

@@
 \textbf{\begin{tabular}[c]{@{}l@{}}Recall\\[-4.5pt] Supported\end{tabular}} & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ \hline
-\textbf{Precision} & 0.97 & 0.98 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
+\textbf{Precision} & 0.94 & 0.96 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
 \textbf{Speed (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
@@
 Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
-While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.97$ and $1.00$) than for the tutorials category ($0.70$).
+While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).
 Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.

### 2016-09-07 Paper/ECAI/tabular.tex
WIP runtime analysis

@@
 
+\samuel{fail-fast?}
 Of the 32 supported constraint templates (including equality), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
@@
 %The running time can be improved by implementing additional optimizations.
+
+\subsection{Runtime analysis}
+There are multiple factors that impact the running time of \sname: the number of vectors, the size of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
+Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
+
+For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
+For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of size~$k$, the worst case complexity becomes $O(n * m * k)$.
+However, in practice, most subassignments will not satisfy the constraint template.
+Since our implementation tries to account for this by failing fast, the amortized time is constant in the size of the vectors~$k$, unless $k$~is very large or many constraints hold in the data.
+For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).
+
+The benchmark experiments show that the number of cells in a spreadsheet only has a limited effect on the runtime.
+Since the implementation is tailored to fail fast when testing constraints, the runtime is in practice
 

### 2016-09-07 Paper/ECAI/tabular.tex
intro rework, unstable state

--- 2016-09-07 Paper/ECAI/tabular.tex
@@
 \section{Introduction}
+% relevance of formulas and spreadsheets
 Millions of people across the world use spreadsheets every day.
@@
 However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
-Many end-users lack the understanding of the underlying structures and dependencies in such sheets and the data they contain.
+Many end-users lack the understanding of the underlying structures and relations in such sheets and the data they contain.
 This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
@@
 Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
-
-In this paper,  we investigate whether learning techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt''\tias{[ADD CITE TO THIS PAPER]} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
+
+% why learn them
+Such issues and mistakes could be overcome through the development of intelligent user-assisting tools.
+A core aspect of such tools would be the ability to automatically discover what constraints and formulae are present in a spreadsheet.
+This is the problem we study in this paper.
+We see multiple promising applications that can build on this, such as rediscovery constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
+
+% why difficult
+In this paper,  we investigate whether knowledge discovery techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
+From a machine learning point of view this is an unconventional problem because
+
 This is a new and unconventional machine learning problem.
@@
 Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
+
 This provides a flavor of how this problem is different from standard mining and learning settings, where the data is just in rows and variables are in columns.
@@
 Moreover, we sketch additional application scenario's such as autocompletion and error detection.
+
+% contribs
+Our contributions are as follows:
+\begin{itemize}
+\item we define the tabular constraint learning problem, where the goal is to learn constraints (functions and relations) that range over entire rows or columns.
+This is an unsupervised learning problem;
+\item we propose a 2-stage generate and test method that reasons for each candidate constraint over the maximally contiguous submatrices in the first stage, and only in the second stage looks at individual rows and columns and their content;
+\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all tuples of submatrices compatible with the candidate constraint its arguments, using backtracking search and forward checking;
+\item experiments on a range of publicly available spreadsheets show that the system is able to extract constraints with high precision and recall in an unsupervised manner. 
+\end{itemize}
 

### 2016-09-07 Paper/ECAI/tabular.tex
reworked intro

@@
 We see multiple promising applications that can build on this, such as rediscovery constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
+In spirit, this builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
 
 % why difficult
-In this paper,  we investigate whether knowledge discovery techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
-From a machine learning point of view this is an unconventional problem because
-
-This is a new and unconventional machine learning problem.
-Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
-Looking at the first row in table~$T_2$ and the data in table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
-Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
-
-This provides a flavor of how this problem is different from standard mining and learning settings, where the data is just in rows and variables are in columns.
-Here everything is mixed.
-The data is relational on the one hand, since we have multiple tables with relationships between them (e.g.
-foreign-keys).
-On the other hand, the data is most often mixed, e.g.
-textual and numeric.
-%, since people tend to use spreadsheets for financial and accounting computations.
-Therefore, understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
+In this paper,  we investigate whether knowledge discovery techniques can be used to learn constraints (formulas and other relations) in spreadsheet data in an unsupervised way.
+From a machine learning point of view this is an unconventional problem because the data is in tabular form, but the constraints we wish to learn can involve both rows and columns of the table.
+Being unsupervised there is no labeled information either, although for every possible function one can verify whether a certain input satisfies the definition.
+%Unlike in other knowledge discovery tasks, such as pattern mining~\cite{}, there are many different types of formula and constraints we wish to discover, each with their own signature (number and properties of arguments) and definition.
+From a data mining point of view, the data is relational on the one hand, since we can have multiple tables with relationships between them, but also of mixed textual and numeric types.
+More closely related is clausal discovery~\cite{claudien,lallouet}, learning CSP constraints \cite{Quacq,Conacq,modelseeker} and dependency discovery in databases \cite{savnik}.
+But an important difference remains that we want to learn constraints on both columns and rows over integer, floating point as well as textual data.
 
 % Moved up
-The question that we answer in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
-It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
-%, and discovers column- and row-wise constraints.
-We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch additional application scenario's such as autocompletion and error detection.
-
+The question that we ask in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
+The main challenge is the amount of combinations of rows and columns that need to be tried as input to the constraints.
+To answer this challenge we propose a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
+%
 % contribs
@@
 \begin{itemize}
-\item we define the tabular constraint learning problem, where the goal is to learn constraints (functions and relations) that range over entire rows or columns.
-This is an unsupervised learning problem;
-\item we propose a 2-stage generate and test method that reasons for each candidate constraint over the maximally contiguous submatrices in the first stage, and only in the second stage looks at individual rows and columns and their content;
-\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all tuples of submatrices compatible with the candidate constraint its arguments, using backtracking search and forward checking;
-\item experiments on a range of publicly available spreadsheets show that the system is able to extract constraints with high precision and recall in an unsupervised manner. 
+\item we define the tabular constraint learning problem, where the goal is to learn constraints (functions and relations) that range over entire rows or columns in an unsupervised way;
+\item we propose a 2-stage generate and test method that reasons for each candidate constraint over maximally contiguous blocks in the first stage, and only in the second stage looks at individual rows and columns and their content;
+\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximal blocks compatible with the candidate constraint's argument, using backtracking search and forward checking;
+\item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall. 
 \end{itemize}
 
-The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
-and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
-very few examples.
-Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.
+%It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
+%, and discovers column- and row-wise constraints.
+%We demonstrate the utility of our approach in an experimental evaluation.
+%Moreover, we sketch additional application scenario's such as autocompletion and error detection.
 
@@
 \sname combines ideas from several existing approaches and different fields of research.
+
+\tias{OLD, from intro: 
+The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
+and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
+Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.}
 

### 2016-09-07 Paper/ECAI/tabular.tex
forgot to take out a duplicate part in rel work

@@
 \sname combines ideas from several existing approaches and different fields of research.
-
-\tias{OLD, from intro: 
-The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
-and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
-Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.}
 

### 2016-09-07 Paper/ECAI/tabular.tex
added needed citations

@@
 Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt''\tias{[ADD CITE TO THIS PAPER]} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
 

### 2016-09-07 Code/src/speed_tests.py
Configure experiments individually

@@ -94,27 +94,32 @@ def clear_folder(path):
             print(e)
 
 
-def generate_experiments(default_power=3, min_power=0, max_power=6, block_power=5, gtype=GType.int):
-    # Build increments
+def generate_experiments():
+    # Configuration
+    gtype = GType.int
     default = 2 ** 3
-    increments = list(2 ** p for p in range(min_power, max_power + 1))
+    block_power = 5
+    number_power = 6
+    size_power = 12
+
+    increment = lambda p_min, p_max: list(2 ** p for p in range(p_min, p_max + 1))
 
     # Vary number of vectors
     cat_1 = []
-    for cols in increments:
+    for cols in increment(0, number_power):
         rows = default
         cat_1.append(SpeedTestId(cols, rows, [cols], [gtype] * cols))
 
     # Vary vector size
     cat_2 = []
-    for rows in increments:
+    for rows in increment(0, size_power):
         cols = default
         cat_2.append(SpeedTestId(cols, rows, [cols], [gtype] * cols))
 
     # Vary block sizes
     cat_3 = []
     vectors = 2 ** block_power
-    for block_size in reversed(increments):
+    for block_size in reversed(increment(0, block_power)):
         if block_size <= vectors:
             cols = vectors
             rows = default
@@ -136,7 +141,7 @@ def setup_experiments(experiments):
 def main():
     # ID: cols, rows, blocks, int | str | float
 
-    categories = generate_experiments(default_power=3, max_power=5)
+    categories = generate_experiments()
     experiments = set(itertools.chain(*categories))
 
     runs = 1

### 2016-09-07 Paper/ECAI/tabular.tex
Changes

@@
 
-\title{Discovering constraints in spreadsheets and tabular data}
+\title{Learning constraints in spreadsheets and tabular data}
 % see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
 
-\author{} % \institute{----------------------} }
+\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
 
@@
 \begin{abstract}
+  \sergey{Samuel! Add a reference to python-constraint somewhere, including the references list at the end}\\
 Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
-However, writing, maintaining and identifying good formulas for tabular data and spreadsheets can be time consuming and error-prone.
-In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data in an unsupervised way.
-Common spreadsheet formulas are represented as predicates whose arguments must satisfy the inherent properties of the function.
-The challenge is to discover these without user information or training examples.
-% Knowledge discovery
-We propose a two-stage generate and test method where the first stage uses constraint solving techniques to efficiently reduce candidates in a generic way based on the function signature. 
-Our approach to this knowledge discovery task takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
-We show that we are able to accurately discover constraints in spreadsheets from various sources.
+However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
+In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data.
+We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
+Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
+Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
+Constraint learning techniques are used to identify the predicates that hold between blocks of rows and columns of the tables.
+We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
 \end{abstract}
@@
 \section{Introduction}
-% relevance of formulas and spreadsheets
 Millions of people across the world use spreadsheets every day.
@@
 However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
-Many end-users lack the understanding of the underlying structures and relations in such sheets and the data they contain.
+Many end-users lack the understanding of the underlying structures and dependencies in such sheets and the data they contain.
 This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
@@
 Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
-
-% why learn them
-Such issues and mistakes could be overcome through the development of intelligent user-assisting tools.
-A core aspect of such tools would be the ability to automatically discover what constraints and formulae are present in a spreadsheet.
-This is the problem we study in this paper.
-We see multiple promising applications that can build on this, such as rediscovery constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
-In spirit, this builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
-
-% why difficult
-In this paper,  we investigate whether knowledge discovery techniques can be used to learn constraints (formulas and other relations) in spreadsheet data in an unsupervised way.
-From a machine learning point of view this is an unconventional problem because the data is in tabular form, but the constraints we wish to learn can involve both rows and columns of the table.
-Being unsupervised there is no labeled information either, although for every possible function one can verify whether a certain input satisfies the definition.
-%Unlike in other knowledge discovery tasks, such as pattern mining~\cite{}, there are many different types of formula and constraints we wish to discover, each with their own signature (number and properties of arguments) and definition.
-From a data mining point of view, the data is relational on the one hand, since we can have multiple tables with relationships between them, but also of mixed textual and numeric types.
-More closely related is clausal discovery~\cite{claudien,lallouet}, learning CSP constraints \cite{Quacq,Conacq,modelseeker} and dependency discovery in databases \cite{savnik}.
-But an important difference remains that we want to learn constraints on both columns and rows over integer, floating point as well as textual data.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
+
+In this paper,  we investigate whether learning techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
+This is a new and unconventional machine learning problem.
+Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
+Looking at the first row in table~$T_2$ and the data in table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
+Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
+This provides a flavor of how this problem is different from standard mining and learning settings, where the data is just in rows and variables are in columns.
+Here everything is mixed.
+The data is relational on the one hand, since we have multiple tables with relationships between them (e.g.
+foreign-keys).
+On the other hand, the data is most often mixed, e.g.
+textual and numeric.
+%, since people tend to use spreadsheets for financial and accounting computations.
+Therefore, understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
 
 % Moved up
-The question that we ask in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
-The main challenge is the amount of combinations of rows and columns that need to be tried as input to the constraints.
-To answer this challenge we propose a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
-%
-% contribs
-Our contributions are as follows:
-\begin{itemize}
-\item we define the tabular constraint learning problem, where the goal is to learn constraints (functions and relations) that range over entire rows or columns in an unsupervised way;
-\item we propose a 2-stage generate and test method that reasons for each candidate constraint over maximally contiguous blocks in the first stage, and only in the second stage looks at individual rows and columns and their content;
-\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximal blocks compatible with the candidate constraint's argument, using backtracking search and forward checking;
-\item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall. 
-\end{itemize}
-
-%It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
+The question that we answer in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
+It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
 %, and discovers column- and row-wise constraints.
-%We demonstrate the utility of our approach in an experimental evaluation.
-%Moreover, we sketch additional application scenario's such as autocompletion and error detection.
+We demonstrate the utility of our approach in an experimental evaluation.
+Moreover, we sketch additional application scenario's such as autocompletion and error detection.
+
+The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
+and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
+very few examples.
+Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.
 
@@
 All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
-\sergey{Samuel! Add a reference to python-constraint, including the references list at the end}\\
 
@@
 %The running time can be improved by implementing additional optimizations.
+
+\subsection{Runtime analysis}
+There are multiple factors that impact the running time of \sname: the number of vectors, the size of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
+Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
+
+For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
+For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of size~$k$, the worst case complexity becomes $O(n * m * k)$.
+However, in practice, most subassignments will not satisfy the constraint template.
+Since our implementation tries to account for this by failing fast, the amortized time is constant in the size of the vectors~$k$, unless $k$~is very large or many constraints hold in the data.
+For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).
+
+The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
+Synthetic experiments with random integers and strings show that increasing the size of the vectors has a neglegible impact on the performance.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{journal_header}
+\input{packages}
+\input{macros}
+
+\begin{document}
+
+\title{Learning constraints in spreadsheets and tabular data}
+% see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
+
+\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
+
+\maketitle
+
+\begin{abstract}
+  \sergey{Samuel! Add a reference to python-constraint somewhere, including the references list at the end}\\
+  \sergey{Tias, Samuel, Luc, should we even mention miniZinc or ASP here, or is it not relevant?}
+Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
+However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
+In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data.
+We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
+Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
+Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
+Constraint learning techniques are used to identify the predicates that hold between blocks of rows and columns of the tables.
+We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
+\end{abstract}
+
+\section{Introduction}
+Millions of people across the world use spreadsheets every day.
+The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
+However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
+Many end-users lack the understanding of the underlying structures and dependencies in such sheets and the data they contain.
+This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
+In this case, often a comma-separated values (CSV) format is used meaning that all formulas are lost, including inter-sheet formulas and relations.
+%This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
+Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
+
+In this paper,  we investigate whether learning techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
+This is a new and unconventional machine learning problem.
+Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
+Looking at the first row in table~$T_2$ and the data in table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
+Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
+This provides a flavor of how this problem is different from standard mining and learning settings, where the data is just in rows and variables are in columns.
+Here everything is mixed.
+The data is relational on the one hand, since we have multiple tables with relationships between them (e.g.
+foreign-keys).
+On the other hand, the data is most often mixed, e.g.
+textual and numeric.
+%, since people tend to use spreadsheets for financial and accounting computations.
+Therefore, understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
+
+% Moved up
+The question that we answer in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
+To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
+It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
+%, and discovers column- and row-wise constraints.
+We demonstrate the utility of our approach in an experimental evaluation.
+Moreover, we sketch additional application scenario's such as autocompletion and error detection.
+
+The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
+and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
+It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
+very few examples.
+Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.
+
+This paper is organized as follows.
+Section \ref{sec:formalization} introduces concepts relevant to our approach and our problem statement.
+Section \ref{sec:approach} presents our approach.
+Section \ref{sec:evaluation} presents the evaluation of the approach.
+Section \ref{sec:applications} shows how our system can be used for applications.
+Section \ref{sec:related_work} presents a more detailed discussion of the related work.
+Section \ref{sec:conclusions} provides conclusions.
+
+%Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
+Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
+
+% {\it
+% \textbf{Motivation}:
+% \begin{itemize}
+%   \item USED -- File generated from model, model got lost, need to reconstruct
+%   \item Constraint programming is hard - is Excel hard?
+%   \item Avoid manual analysis, provide selection of constraints
+%   \item SOMEWHAT USED --Error checking
+%   \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
+% \end{itemize}
+
+% \textbf{Novelty:}
+% \begin{itemize}
+%   \item USED -- Unsupervised setting (contrary to flashfill, etc)
+%   \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
+%   \item USED -- Data format (2D) -- data is no longer in rows like a classic ML or DM settings
+%   \item USED -- Declarative, general / modular, stacking of constraint problems
+% \end{itemize}
+%}
+
+\begin{figure}[thb]
+
+  \begin{subfigure}{1\textwidth}
+  \begin{center}
+    \includegraphics[width=1\textwidth]{figures/Demo2.png}
+  \end{center}
+  \vspace{-10pt}
+  \caption{
+    Example spreadsheet (black words and numbers only).
+    Green background indicates headerless tables, dark borders indicate maximal type-consistent blocks.
+    Most tables only contain type-consistent columns while table~$T_2$ also contains type-consistent rows.
+    However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
+    This example was created by combining several Excel sheets from several exercise sessions based on the book MS Excel 2010~\cite{excel_book}.
+    % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
+  }
+  \label{fig:main_example}
+\end{subfigure}
+\hfill
+\begin{subfigure}{1\textwidth}
+  {\scriptsize
+    \begin{align*}
+      % Table 1
+%
+%      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
+%
+      &~\ecseries{\range{T_{1}}{\rangeall}{1}}                                  && \ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^*     && \ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}}  \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^*     && \ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^*    && \ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}       && \ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*     && \range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
+      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^*    && \eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^*\\
+%
+      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}}   && \eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
+%
+      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}}   && \ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}}
+%
+    % Table 2
+%     &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+%     &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+%     &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
+%
+%     &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
+%
+    % Table 4
+%     &~\ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
+%
+%     &~\range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
+%
+    % Table 5
+%     &~\eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^* \\
+%
+%     &~\eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}
+    \end{align*}}
+  \vspace{-10pt}
+  \caption{Constraints learned for the example (left), except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
+  Constraints marked with $*$ were not present in the original document}
+  \label{fig:sol_example}
+\end{subfigure}
+  \caption{Running example}
+\end{figure}
+
+\section{Formalization}\label{sec:formalization}
+Our goal is to automatically discover constraints between the rows and the columns of tables in a spreadsheet.
+This is applicable not just to data from spreadsheets, but to any data in tabular form, hence the name.
+
+%An example spreadsheet with tables is given in Figure~\ref{fig:main_example}, we will use this as running example throughout the paper.
+
+We first introduce some terminology and the concept of a \template, after which we define the problem and make some additional considerations.
+
+\subsection{Terminology}
+Spreadsheets and tabular data may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
+Note that a table can contain a header; however, we wish to reason over entire rows and columns of data, and hence we will consider \textbf{headerless tables} only.
+
+Formally, a (headerless) table is an $n \times m$ matrix. Each entry is called a \textit{cell}.
+A cell has a {\bf type}, which can be numeric or textual.
+We further distinguish numeric types in subtypes: integer and float.
+We also consider \textit{None} as a special type when a cell is empty; \textit{None} is a subtype of all other types.
+
+A row or a column is \textbf{type-consistent} if all cells in that row or column are of the same base type, i.e. numeric or textual.
+We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$ and similarly $T[{:},a]$ for the $a$-th column.
+For example, in Figure~\ref{fig:main_example}, $T_1[:,1] = [1,2,3,4]$ and $T_3[1,:] = ['$Diana Coolen$', 5]$.
+$T_3[1, :]$ is not type-consistent while $T_1[:, 1]$ is.
+%We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.
+
+The most important concept is that of a \textbf{block}. 
+\begin{definition}
+A \textbf{block} has to satisfy three conditions: 1)~it contains only entire rows or entire columns of a single headerless table; 2)~it is contiguous; and 3)~it is type-consistent.
+The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type. 
+If a block contains only rows we say it has \textit{row-orientation}, if only columns, \textit{column-orientation}. 
+\end{definition}
+
+In line with this definition, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
+We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independently of their orientation.
+
+A block has the following properties:
+\begin{itemize}
+\item \textit{type}: a block is type-consistent, so it has one type
+\item \textit{table}: the table that the block belongs to
+\item \textit{orientation}: either row-oriented or column-oriented
+\item \textit{size}: the number of vectors a block contains
+\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
+\item \textit{rows}: the number of rows in the block
+(in row-oriented blocks this is equivalent to the size)
+\item \textit{columns}: the number of columns in the block (in row-oriented blocks this is equivalent to the length)
+\end{itemize}
+
+%We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
+
+%For notational convenience, we will refer to vectors in {\rm roman} and to groups in {\bf bold}. {\bf CHECK}
+
+\begin{example}
+Consider the (headerless) table~$T_1$ in Figure~\ref{fig:main_example}.
+Its rows are not type consistent (i.e. they contain both numeric and textual data).
+However, the table can be partitioned into five column-oriented blocks $\sbl{1}, \sbl{2}, \sbl{3}, \sbl{4}, \sbl{5}$, as shown in the figure ($\sbl{1} = \range{T_1}{\rangeall}{1}$, $\sbl{2} = \range{T_1}{\rangeall}{2}$, $\sbl{3} = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
+\end{example}
+
+\begin{definition}
+\textbf{Block containment $\sqsubseteq$.} 
+A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (contiguous, type consistent) with the same orientation and table, and each of the vectors in $B'$ is also in $B$.
+For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=\range{T}{\rangeto{a}{b}}{\rangeall} \wedge B'=\range{T}{\rangeto{a'}{b'}}{\rangeall} \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
+\end{definition}
+
+We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
+%
+An example of block containment is $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.
+
+\begin{table}[htb]
+\caption{
+  Overview of constraint templates implemented in \sname.
+  Subgroups in normal font have to be vectors, whereas subgroups in \textbf{bold} may contain more than one vector;
+  $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
+  Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.
+  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows \textit{SUM} but \sname also supports \textit{MAX}, \textit{MIN}, \textit{AVERAGE}, \textit{PRODUCT} and \textit{COUNT}.
+  %{\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
+}
+\label{table:constraints}
+\resizebox{1.0\columnwidth}{!}{%
+  {\normalsize \centering
+  \begin{tabularx}{1.70\textwidth}{l X X}
+    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
+    $\ecalldiff{\sbl{x}}^*$
+      & $\discrete(\sbl{x})$
+
+      & All values in $\sbl{x}$ are different. $i \neq j$: $\sbl{x}[i] \neq \sbl{x}[j]$
+      \\[\PADSEP] \hline
+
+    $\ecperm{\sbl{x}}^*$
+      & $\numeric(\sbl{x})$, $\ecalldiff{\sbl{x}}$
+
+      & The values in $\sbl{x}$ are a permutation of the numbers $1$ through $\plength(\sbl{x})$.
+      \\ \hline
+
+    \ecseries{\sbl{x}}
+      & $\integer(\sbl{x})$ and $\ecperm{\sbl{x}}$
+
+      & $\sbl{x}[1] = 1$ and $\sbl{x}[i] = \sbl{x}[i - 1] + 1$.
+      \\[\PADSEP] \hline
+
+    $\ecfkey{\sbl{fk}}{\sbl{pk}}^*$
+      & $\sbl{fk}$ and~$\sbl{pk}$ are $\discrete$; they must have different $\ptable$s but the same $\ptype$; and $\ecalldiff{\sbl{pk}}$
+
+      & Every value in~$\sbl{fk}$ also exist in~$\sbl{pk}$ \\[\PADSEP] \hline
+
+    \eclookup{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
+      & $\sbl{fk}$ and $\sbl{pk}$ are $\discrete$; arguments $\{\sbl{fk}, \sbl{r}\}$ and $\{\sbl{pk}, \sbl{val}\}$ within the same set have the same \plength, \ptable and \por; $\sbl{r}$ and~$\sbl{val}$ have the same type; and \ecfkey{\sbl{fk}}{\sbl{pk}}.
+
+      & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] = \sbl{fk}[i]$
+      \\[\PADSEP] \hline
+
+    \eclookupfuzzy{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
+      & \textit{Same as lookup}
+
+      & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] \leq \sbl{fk}[i]$, $j$~maximal
+      \\[\PADSEP] \hline
+
+    \eclookupprod{\sbl{r}}{\sbl{1}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{fk}\}$ are $\numeric$, arguments $\{\sbl{pk}, \sbl{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sbl{fk}}{\sbl{pk}}.
+
+      & $\sbl{r}[i] = \sbl{1}[i] \times \eclookupf{\sbl{fk}}{\sbl{pk}}{\sbl{val}}{}[i]$.
+      \\[\PADSEP] \hline
+
+    \ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}}
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are all $\numeric$ and have the same $\plength$
+
+      & $\sbl{r}[i] = \sbl{1}[i] \times \sbl{2}[i]$.
+      \\[\PADSEP] \hline
+
+    \ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
+
+      & $\sbl{r}[i] = \sbl{1}[i] - \sbl{2}[i]$.
+      \\[\PADSEP] \hline
+
+    $\eccalc{\sbl{r}}{\frac{\sbl{1} - \sbl{2}}{\sbl{2}}}$
+      & \textit{Same as \ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}}
+
+      & $\sbl{r}[i] = (\sbl{1}[i] - \sbl{2}[i]) / \sbl{2}[i]$.
+      \\[\PADSEP] \hline
+
+    \ecproj{\sbl{r}}{\bsbl{x}}
+      & Arguments $\{\sbl{r}, \bsbl{x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\bsbl{x}$ contains at least~2 vectors; and $\sbl{r} = \textit{SUM}_{\por(\bsbl{x})}(\bsbl{x})$
+
+      & At every position~$i$ in $1$ through $\plength(\sbl{r})$ there is exactly one vector~$v$ in $\bsbl{x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sbl{r}[i]$.
+      \\[\PADSEP] \hline
+
+    \ecrank{\sbl{r}}{\sbl{x}}
+      & $\integer(\sbl{r})$; $\numeric(\sbl{x})$; and $\plength(\sbl{r}) = \plength(\sbl{x})$
+
+      & The values in $\sbl{r}$ represent the rank (from largest to smallest) of the values in $\sbl{x}$ (including ties)
+      \\[\PADSEP] \hline
+
+    \ectotal{\sbl{r}}{\sbl{pos}}{\sbl{neg}}
+      & Arguments $\{\sbl{r}, \sbl{pos}, \sbl{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
+
+      & $\sbl{r}[i] = \sbl{r}[i - 1] + \sbl{pos}[i] - \sbl{neg}[i]$.
+      \\[\PADSEP] \hline
+
+    $\ecsumr{\sbl{r}}{\bsbl{x}}^\dagger$
+      & $\sbl{r}$ and $\bsbl{x}$ are $\numeric$; $\pcols(\bsbl{x}) \geq 2$; and $\prows(\bsbl{x}) = \plength(\sbl{r})$
+
+      & $\sbl{r}[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \textit{row}(i, \bsbl{x})[j]$
+      \\[\PADSEP] \hline
+
+    $\ecsumc{\sbl{r}}{\bsbl{x}}^\dagger$
+      & $\sbl{r}$ and $\bsbl{x}$ are $\numeric$; $\prows(\bsbl{x}) \geq 2$; and $\pcols(\bsbl{x}) = \plength(\sbl{r})$
+
+      & $\sbl{r}[i] = \sum_{j = 1}^{\prows(\bsbl{x})} \textit{column}(i, \bsbl{x})[j]$
+      \\[\PADSEP] \hline
+
+    $\ecsumif{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}^\dagger$
+      & $\sbl{fk}, \sbl{pk}$ are $\discrete$; $\sbl{r}, \sbl{val}$ are $\numeric$; within the sets $\{\sbl{val}, \sbl{fk}\}$ and $\{\sbl{pk}, \sbl{r}\}$ arguments have the same $\plength$ and $\por$; $\sbl{fk}$ and $\sbl{val}$ have the same $\ptable$; $\sbl{fk}$ and $\sbl{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sbl{pk}}
+
+      & \[\sbl{r}[i] = \sum_{j=1}^{\plength(\sbl{val})} \begin{cases}
+          \sbl{val}[j] & \text{if } \sbl{fk}[j] = \sbl{pk}[i] \\
+          0 & \text{otherwise}
+        \end{cases}
+      \] \\[\PADSEP] \hline
+
+    \ecsumprod{\sbl{r}}{\sbl{1}}{\sbl{2}}
+      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are $\numeric$; $\plength(\sbl{1}) = \plength(\sbl{2}) \geq 2$; and $\prows(\sbl{r}) = \pcols(\sbl{r}) = 1$
+
+      & $\sbl{r}[1] = \sum_{j = 1}^{\plength(\sbl{1})} \sbl{1}[j] \times \sbl{2}[j]$.
+    \end{tabularx}
+  } % centering small
+} % resizebox
+
+\end{table}
+
+\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
+\newcommand{\defc}{\ensuremath{\format{Def}_s}}
+
+\subsection{Constraint templates}
+The goal is to learn constraints over blocks in the data. The knowledge needed to learn a constraint is expressed through {\template}s.
+%
+A \template $s$ is a triple $t = \textit{(\CName, \CSignature, \CFunction)}$:
+%Let us elaborate on this:
+\begin{itemize}
+\item
+\textit{\CName}  specifies the syntactic form of the constraint $s(\sbl{1}, ...,\sbl{n})$, that is, the name of the template together
+with $n$ abstract arguments $\sbl{i}$.
+Thus, a constraint is viewed as a relation or predicate of arity $n$ in first order logic.
+Note that a function $\sbl{r}=f(\sbl{1},...,\sbl{n})$ can be represented with the $(n{+}1)$-ary predicate $s_f(\sbl{r},\sbl{1},...,\sbl{n})$.
+Each argument will have to be instantiated with a block.
+% the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
+
+\item \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
+This can concern properties of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
+In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
+We can capture this bias for a template~$s$ using a predicate $\sigc$.
+\item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
+Given an assignment of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks.
+%holds, in practice this will be a function that can be called when
+%the arguments are full instantiated to decide whether the constraint is satisfied or not.
+In logical and relational learning this is known as the background knowledge.
+We introduce the predicate \defc~to capture this background knowledge for a template~$s$.
+\end{itemize}
+
+\begin{example}
+The constraint templates implemented in \sname are defined in Table~\ref{table:constraints}.
+A non-trivial example is, for example, the constraint template for the row-based sum:
+
+\begin{itemize}
+  \item \CName: $\ecsumr{\sbl{r}}{\bsbl{x}}$, for arguments $\sbl{r}$ and $\bsbl{x}$.
+
+  \item \CSignature: $\sbl{r}$ has to be a single vector ($\textit{size}=1$) while $\bsbl{x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
+  Both blocks have to be numeric.
+  This constraint is orientation-specific, so it requires that the number of rows in $\bsbl{x}$ equals the length of $\sbl{r}$.
+  Moreover, we add to the bias that the number of columns to sum over is at least~$2$.
+
+  \item \CFunction: each value in the vector $\sbl{r}$ is obtained by summing over the corresponding row in $\bsbl{x}$.
+  %For example, $\sbl{r} = [12, 7]$ and $\bsbl{x}$ consists of columns: $[5, 15]$, $[7, -8]$.
+\end{itemize}
+\textit{SUM} is an aggregate constraint, and a similar constraint is available for aggregate operators \textit{MIN}, \textit{MAX}, \textit{AVERAGE}, \textit{PRODUCT} and \textit{COUNT}.
+The \textit{conditional} aggregate variant \textit{SUMIF} only ranges over those cells that satisfy a condition on a related cell in that table, e.g.
+$\range{T_1}{\rangeall}{10}$ and the data in $T_5$ for matching salesperson names.
+\end{example}
+
+It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
+From a FOL perspective, a constraint of the form $\ecrank{\sbl{r}}{\sbl{x}}$ can be seen as a predicate $\textit{RANK}(\sbl{r}, \sbl{x})$ where \textit{RANK} is the name of the predicate and its arguments~$\sbl{r}$ and~$\sbl{x}$ are terms, which can be seen as either uninstantiated variables or as concrete values.
+This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
+For example, for the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $\sbl{r}$ is the $8$th vector in $T_1$: $\sbl{r} = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $\sbl{x}$ is the $7$th vector: $\sbl{x} = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
+
+%When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
+%From a FOL perspective, the name of the constraint ($\mathit{RANK}$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
+%This also holds in our setting: when we write
+%      $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
+%we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ (provided~$T$ has an 8th column).
+%So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
+
+With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
+We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(\sbl{1}, ..., \sbl{n})$ if $\sigc(\sbl{1}, ..., \sbl{n})$ (respectively $\defc(\sbl{1}, ..., \sbl{n})$) is satisfied.
+Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would define the predicate~$s$ using a Prolog like clause: $s(\sbl{1}, ..., \sbl{n}) \leftarrow \sigc(\sbl{1}, ..., \sbl{n}) \wedge \defc(\sbl{1}, ..., \sbl{n})$.
+Under this interpretation, the term constraint and constraint template can be used interchangeably.
+
+\begin{definition}
+A \textbf{valid argument assignment} of a constraint template~$s$ is a tuple of blocks $(\sbl{1}, ..., \sbl{n})$ such that $s(\sbl{1}, ..., \sbl{n})$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(\sbl{1}, ..., \sbl{n})$ to the arguments.
+\end{definition}
+
+\subsection{Problem Definition}\label{sec:problem_statement}
+The problem of learning constraints from tabular data can be seen as an inverse {\em constraint satisfaction problem} (CSP).
+In a CSP one is given a set of constraints over variables that must all be satisfied, and the goal is to find an instantiation of all the variables that satisfies these constraints.
+In the context of spreadsheets, the variables would be (blocks of) cells, and one would be given the actual constraints and functions with the goal of finding the values in the cells.
+The inverse problem is, given only an instantiation of the cells, to find the constraints that are satisfied in the spreadsheet.
+
+We define the inverse problem, that is the {\bf Tabular Constraint Learning Problem}, as follows:
+%
+\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
+{\bf Given} a set of instantiated blocks~${\cal B}$ over tables~${\cal T}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(\ssbl{1}, ..., \ssbl{n})$ where $s \in {\cal S}$, $\forall i: \sbl{i}' \sqsubseteq \sbl{i} \in {\cal B}$ and $(\ssbl{1}, ..., \ssbl{n})$ is a satisfied argument assignment of the template~$s$.
+\end{definition}
+% Let us now define what it means for a constraint from \template to hold in general.
+Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
+% \sergey{moved from formalization END}
+
+%Here we formalize the statement in terms of \template and group assignments as follows:
+%   \begin{tabular}{ll}
+%     \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
+%     \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
+%     \textbf{Find:}&  all constraints $C$ over \groups for each template $t$ in \constraints \\
+%   \end{tabular}
+
+The input is a set of blocks, and in Section~\ref{sec:approach} we will discuss how these can be extracted from a spreadsheet.
+Figure~\ref{fig:sol_example} shows the solution to the tabular constraint learning problem when applied on the blocks of Figure~\ref{fig:main_example} and constraint templates listed in Table~\ref{table:constraints}.
+
+\subsection{Other considerations}
+
+\subsubsection{Dependencies}
+\label{sec:form:dependencies}
+In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g. for \textit{PERMUTATION}.
+This expresses a dependency of the constraint on that other \textit{base} constraint.
+This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the base constraint, and its definition of its own definition plus the definition of the base constraint.
+In FOL, we can see that one constraint entails the other, for example if $\ecperm{\sbl{x}}$ holds for a block $\sbl{x}$, then $\ecalldiff{\sbl{x}}$ also holds.
+
+In Section~\ref{sec:approach} we will see how, apart from easing the specification of the signature and definition, such dependencies can be used to speed up the search for constraints.
+
+\subsubsection{Redundancies}
+\label{sec:form:redundancies}
+Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}.
+This is because constraints may be logically equivalent or may be implied by other constraints, e.g.
+if you know $\ecperm{\sbl{x}}$, $\ecalldiff{\sbl{x}}$ is redundant.
+
+Within the same constraint, there can be equivalent argument assignments if the order of some of the arguments does not matter.
+For example for the product constraint, $\ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}} \equiv \ecprod{\sbl{r}}{\sbl{2}}{\sbl{1}}$ so one can be considered redundant to the other.
+Two different constraints can also be logically equivalent, due to their nature e.g.
+for addition / subtraction and product / division: $\ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}} \equiv \ecdiv{\sbl{1}}{\sbl{r}}{\sbl{2}}$.
+
+Finally, when the data has rows or columns that contain \textit{exactly} the same values, then any constraint with such a vector in its argument assignment will also have an argument assignment with the other equivalent vectors.
+
+Dealing with redundancy is often application-dependent.
+Therefore, in Section~\ref{sec:approach} we first explain our generic method for finding all constraints, before describing some optimizations that avoid obvious equivalences (Section~\ref{sec:opts}).
+Later, we will investigate the impact of redundancy in the experiments.
+
+%More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
+%the  following statement holds: $$\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m).$$
+%
+%We shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
+%To simplify notation, and in line with logic programming, we shall often make the quantification implicit and simply write this as rules $ c_1(G_1, ...
+G_n) \rightarrow c_2(G'_1, ...
+G'_m)$.
+We say that the first constraint is more specific than the second, and the second is more general.
+%
+%
+%% where the second one ranges over a subsets of the arguments of the first, iff whenever the first constraint is true, the second must also be true.
+More formally, $\forall G'_1, ..., G'_n$ s.t.
+$\{G''_1, ..., G''_m\} \subseteq \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
+G'_n) \rightarrow c_2(G''_1, ..., G''_m)$ll.
+%
+%\paragraph{Example}
+%Consider the following implication:
+%$$\ecperm{G} \rightarrow \ecalldiff{G}$$
+%(it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G$ is a vector).
+%In this case the arguments are identical in both constraints.
+%Given that the problem is to find the set of {\em all} constraints, and that the more general constraint is entailed
+%by the more specific one, the $\ecdiff{G}$ constraint will be {\em redundant}.
+%So, for most applications, it will be sufficient to output only the most specific one.
+%
+%A more involved example is
+%$$\eclookup{G'_2}{G'_1}{H'_1}{H'_2} \rightarrow \ecfkey{G'_1}{H'_1}$$
+%which states
+%that a look-up constraint implies a foreign-key constraint.
+In the look-up constraint $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ the values in the vector $G'_2$ are determined by looking up the corresponding value in vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
+In Figure~\ref{fig:main_example} we have $\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}$.
+This implies a foreign-key constraint $\ecfkey{G'_1}{H'_1}$ where each of the values in vector $G'_1$ must appear in $H'_1$ for the look-up to work.
+
+%\tias{I am not defining a new problem for 'most specific canonical tabular constraint learning' because we loosely showed how implications and equivalences would be expressed, but not how to express canonicity and we did not claim that most specific is always best.}
+%\tias{Note also that our method finds ALL constraints; it uses specificity to faster traverse the search space, but the resulting set of constraints found would be the same, whether or not using the specificity (see 4.2.1).}
+
+%Let us introduce the version of tabular constraint learning by modifying Definition \ref{def:problem_statement} that takes into account the specificity relation between constraint templates.
+%
+%\begin{definition} \textit{Condensed Tabular Constraint Learning}\label{def:condensed_problem_statement}\\
+%  Given a set of instantiated tables~${\cal T}$ (all cells have values), a set of {\template}s~${\cal S}$, a set of groups~${\cal G}$ over the tables~${\cal T}$, and the set of all specifity rules~\dependencies: find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, the (sub)groups $G_1', ...
+, G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$ and there is no $s'$ in ${\cal C}$ such that $s'(G_h',\dots,G_k')$ holds (for $1 \leq h \leq k \leq n$) and  $s'$ is more specific than $s$ in \dependencies.
+%\end{definition}
+
+% \begin{minipage}[c]{14em}
+%   \vspace{5pt}
+%   \begin{tabular}{ll}
+%     \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
+%     \vspace{-4pt}
+%     &\\
+%     \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
+%     & a constraint specificity DAG \dependencies \\
+%     \textbf{Find:}& all the most specific wrt \dependencies constraints $C$ over \groups\\
+%     & for each $t$ in \constraints \\
+%   \end{tabular}
+%   \vspace{6pt}
+% \end{minipage}
+
+%This version explicitly takes into account the specificity of the constraints and invalidates some of the solutions that are entailed by the tabular other solutions.
+
+\newcommand{\tcl}{Tabular Constraint Learning}
+%\newcommand{\ctl}{Condensed Tabular Constraint Learning}
+\section{Approach to Tabular Constraint Learning}\label{sec:approach}
+The aim of our method is to detect constraints between rows and columns of tabular data.
+Recall that a valid argument assignment for a constraint is an assignment of a block to each of the arguments of the constraint, such that the signature and definition of the constraint template is satisfied.
+
+Our proposed methodology contains the following steps:
+\begin{enumerate}
+\item Extract headerless tables from tabular data
+\item Partition the tables into contiguous, type-consistent \textit{superblocks}
+\item Generate for each constraint template all valid argument assignments in two steps:
+\begin{enumerate}
+\item For each constraint template~$s$, generate all assignments $(\sbl{1}, \ldots ,\sbl{n})$ where each~$\sbl{i}$ is a superblock and the~$\sbl{i}$ are \textit{compatible} (defined below) with the signature of~$s$.
+\item For each generated assignment $(\sbl{1}, \ldots ,\sbl{n})$ in 3a), find all valid constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that for all $i$ holds $\ssbl{i} \sqsubseteq \sbl{i}$ and the signature and definition of~$s$ are satisfied.
+\end{enumerate}
+\end{enumerate}
+
+The core of our method is step 3.
+In principle, one could use a generate-and-test approach by generating all possible blocks from the superblock and testing each combination of blocks for each of the arguments of the constraints.
+However, each superblock of size $m$ has $m*(m+1)/2$ contiguous subblocks, meaning that a constraint with $n$ arguments would have to check $O(n^{m^2})$ combinations of blocks.
+
+Instead, we divide this step into two parts: in step 3a, we will not reason over individual (sub)blocks and their data, but rather over the properties of the superblocks.
+% and what combination of superblocks may lead to a valid argument assignment.\\
+Consider table~$T_1$ in Figure~\ref{fig:main_example} and the $\textit{SUM}_{row}$ constraint template.
+Instead of considering all possible subblocks of $T_1$, we first reason over the properties of the superblocks.
+For example, $\sbl{2}$ and $\sbl{4}$ are not numeric so they can be immediately discarded.
+Superblocks~$\sbl{1}, \sbl{3}$ and $\sbl{5}$ are valid candidates for~$\sbl{r}$, however, since they all have length~4, $\mathbf{\sbl{x}}$ needs at least 4~columns which is only satisfied by $\sbl{3}$.
+So the assignments generated in step 3a would be $(\sbl{1}, \sbl{3}), (\sbl{3}, \sbl{3}), (\sbl{5}, \sbl{3})$.
+
+In step 3b, we start from an assignment $(\sbl{1}, \ldots, \sbl{n})$ with superblocks $\sbl{i}$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
+As we only have to consider the blocks $\ssbl{i} \sqsubseteq \sbl{i}$ contained in each superblock, this is typically much easier.
+For the above example and $(\sbl{5}, \sbl{3})$, each of the vectors of $\sbl{5}$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks of $\sbl{3}$ for the right-hand side and verify the signature (e.g.
+at least size $4$) and test whether for each row the definition is satisfied.
+%In practice, this can be optimized further.
+
+We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step~3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step~3b, Section~\ref{sec:algo:subgr}).
+In Section~\ref{sec:opts} we describe a number of optimizations designed to improve the algorithm.
+
+\subsection{Table extraction}
+\label{sec:table_extraction}
+Many spreadsheets contain headers that provide hints at where the table(s) in the sheet are and what the meaning of the rows and columns is.
+However, detecting tables and headers is a complex problem on its own~\cite{header}.
+Furthermore, headers may be missing and their use often requires incorporating language-specific vocabulary, e.g.
+English.
+
+% \begin{figure}[t]
+%   \centering
+%   \includegraphics[width=1\linewidth]{figures/tabletool.png}
+%   \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding to (headerless) tables.
+%   An explicit orientation can also be provided.}
+%   \label{fig:visual_tool}
+% \end{figure}
+
+Instead, we assume tables are specified by means of their coordinates within a spreadsheet and optionally a fixed orientation of each table.
+The orientation can be used to indicate that a table should be interpreted as having only rows or only columns.
+
+We developed two simple prototypes to help with the specification of the tables:
+%This information can be provided manually or generated automatically.
+
+\paragraph{Automatic detection}
+%As mentioned above, automatic detection of headers and tables is a complex task.
+%Therefore, we define a restricted setting in which the task becomes easy enough to be automated.
+Under the following two assumptions, the table detection task becomes easy enough to be automated; 1) tables are rectangular blocks of data not containing any empty cells; and 2) tables are separated by at least one empty cell from other tables.
+The sheet is then processed row by row, splitting each row into ranges of non-empty cells and appending them to adjacent ranges in the previous row.
+Headers can be detected, for example, by checking if the first row or column contains only textual values.
+If a header row (column) is detected, it is removed from the table and the orientation of the table is fixed to columns (rows), otherwise, we assume there is no header and the orientation is not fixed.
+
+\paragraph{Visual}
+The above assumptions do not hold for many tables.
+However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/SergeyParamonov/TaCLe}\label{github-link}}).
+Users select tables excluding the header and optionally specify an orientation.
+The tool then generates the specification of the tables automatically.
+
+\subsection{Block detection} \label{sec:make_groups}
+%A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) blocks.
+The goal of the block detection step is to partition tables into maximally type-consistent (all-numeric or all-textual) blocks.
+%We generate blocks automatically, partitioning tables into maximal type-consistent blocks.
+First, we preprocess the spreadsheet data so that currency values and percentual values are transformed into their corresponding numeric (i.e.
+integer or floating point) representation (e.g.
+$2.75 \$$ as $2.75$ or $85\%$ as~$0.85$).
+Then, each table is partitioned into maximal type-consistent blocks.
+
+To find row-blocks, each row is treated as a vector and must be type-consistent; similarly for columns and column-blocks.
+Then, adjacent vectors that are of the same type are merged to obtain the maximally type-consistent superblocks.
+
+\subsection{Algorithm}
+\label{sec:algo}
+Our method assumes that constraint templates and maximal blocks are given and solves the \tcl~problem by checking for each template $s$: what combination of maximal blocks can satisfy the signature (\textit{superblock assignment}) and which specific \textit{subblock assignments} satisfy both the signature and the definition.
+The pseudo-code of this approach is shown in Algorithm~\ref{algo:tcl}.
+%For every constraint template~$s$, the algorithm first generates all combinations of superblocks (i.e. superblock assignments) that are consistent with the signature.
+%In a second step it looks at every superblock assignment and finds subassignments that satisfy the signature and definition of~$s$.
+
+This separation of checking the \textit{properties} of superblock assignments from checking the \textit{actual data} in the subblock assignment controls the exponential blow-up of combinations to test.
+Furthermore, we use constraint satisfaction technology in the first step to efficiently enumerate superblock assignments that are compatible with the signature.
+%These are the two key properties of our approach.
+
+\newcommand{\temps}{\ensuremath{S}}
+\begin{algorithm}[t]
+  \begin{algorithmic}[1]
+    \footnotesize
+    \State \textbf{Input:} $\temps$ -- constraint templates, $\blocks$ -- maximal blocks
+    \State \textbf{Output:} $C$ -- learned constraints
+
+    \Procedure{LearnConstraints}{$\blocks$, $\temps$}
+      \State $C \gets \emptyset$ %\Comment{The set of constraints}
+      \ForAll{$s~\mathbf{in}~\temps$} \label{algo:tcl:for}
+        %\State $n \gets$ number of arguments of template $s$
+        \State $A \gets \generategroups(s, \blocks)$  \label{algo:tcl:super}
+        \ForAll{$(\sbl{1}, \dots, \sbl{n}) \in A$}
+        \State{$A' \gets \findassignment(s, (\sbl{1},\dots,\sbl{n}))$}
+        \ForAll{$(\sbl{1}',\dots,\sbl{n}') \in A'$}
+            \State $C \gets C \cup \{ c_s(\sbl{1}', \dots, \sbl{n}') \}$
+          \EndFor
+        \EndFor
+      \EndFor
+      \State \Return $C$
+    \EndProcedure
+\end{algorithmic}
+\caption{Learn tabular constraints}
+\label{algo:tcl}
+\end{algorithm}
+
+\subsubsection{Generating superblock assignments (Step 3a)}
+\label{sec:algo:super}
+\
+% Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.
+
+Given a constraint template~$s$ and the set of all (maximal) superblocks~\blocks, the goal is to find all combinations of superblocks that are compatible with the constraint signature.
+An argument assignment $(\sbl{1}, ...
+\sbl{n})$ is \textit{compatible} with the signature of template $s$ if for each block $\sbl{i}$ there exists at least one subblock $\sbl{i}' \sqsubseteq \sbl{i}$ that satisfies the signature.
+
+The choice of one argument can influence the possible candidate blocks for the other arguments, for example, if they must have the same length.
+Instead of writing specialized code to generate and test the superblock assignments, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.
+
+A Constraint Satisfaction Problem (CSP) is a triple $(V,D,C)$ where $V$ is a set of \textit{variables}, $D$ the domain of possible values each variable can take and $C$ the set of constraints over $V$ that must be satisfied.
+In our case, we define one variable $V_i$ for each argument of a constraint template.
+Each variable can be assigned each of the maximal blocks in \blocks, so the domain of the variables consists of $|\{\blocks\}|$ block identifiers.
+
+%Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
+%For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that a group~$G \in \groups$ is selected if there is at least one subgroup~$G' \subseteq G$ that satisfies the \CSignature.
+
+To reason over the blocks, we add to the constraint program a set of background facts that contain the properties of each block, namely its type, table, orientation, size, length and number of rows and columns.
+
+\begin{table}[t]
+  \caption{Translation of signature requirements to superblock constraints. The following need not be relaxed: $\plength(B) = x$, $\por(B) = x$, $\ptable(B) = x$ and $\psize(B) \geq x$.}
+  \label{tbl:translation}
+  \begin{tabularx}{\linewidth}{lX}
+    \textbf{Requirement} & \textbf{Superblock constraint} \\ \hline \hline
+    $\ptype(B) = x$ & $\mathit{basetype}(\ptype(B)) = \mathit{basetype}(x)$ \\ \hline
+    $\psize(B) = x$ & $\psize(B) \geq x$ \\ \hline
+    $\pcols(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \pcols(B) \geq x$ \\ 
+    & $\mathit{if}~\por(B) = \mathit{row}: \pcols(B) = x$ \\ \hline
+    $\prows(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \prows(B) = x$ \\ 
+    & $\mathit{if}~\por(B) = \mathit{row}: \prows(B) \geq x$
+  \end{tabularx}
+\end{table}
+
+The actual constraints must enforce that a superblock assignment is \textit{compatible} with the requirements defined in the signature.
+
+Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints works:
+Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment~($\sqsubseteq$).
+Typing requirements need to be relaxed to check only for base types (numeric and textual).
+Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with too many vectors contain subblocks of the required (smaller) size.
+Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
+Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
+
+The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all solutions.
+These solutions correspond to the valid superblock assignments for constraint template~$s$.
+
+\begin{example}
+%Br and Bx are numeric; rows(Bx) ⇤ 2; and columns(Bx ) = length(Br )
+
+  Consider the constraint template $\ecsumr{\sbl{r}}{\bsbl{x}}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$\sbl{r}$ and~$\bsbl{x}$.
+  Given maximal blocks~\blocks, the domain of these variables are $D(V_r) = D(V_x) = \{ 1, \dots, |\blocks| \}$.
+  Finally, the signature of $\ecsumr{\sbl{r}}{\bsbl{x}}$ (see Table~\ref{table:constraints}) will be translated into constraints:
+  \begin{align*}
+    & \numeric(V_r) \land \numeric(V_x) \land \pcols(V_x) \geq 2 \\
+    & (\por(V_x) = \mathit{column}) \Rightarrow (\prows(V_x) \geq \plength(V_r)) \\
+    & (\por(V_x) = \mathit{row}) \Rightarrow (\prows(V_x) = \plength(V_r))
+  \end{align*}
+  %Every solution is a tuple of values for~$V_r$ and~$V_x$ corresponding directly with a superblock assignment over~\blocks.
+\end{example}
+
+\subsubsection{Generating subblock assignments (Step 3b)}
+\label{sec:algo:subgr}
+Given a superblock assignment $(\sbl{1}, \dots, \sbl{n})$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e.
+assignments of subblocks $(\ssbl{1}, \dots, \ssbl{n})$ (for all $i$, $\ssbl{i} \sqsubseteq \sbl{i}$) that satisfy both the signature and the definition of template~$s$.
+
+%While the previous step reasoned about the \textit{properties} of blocks, in this step we will also reason about the actual \textit{content} of the blocks, that is, the data.
+\begin{example}
+  Consider the sum-over-rows constraint template $\ecsumr{\sbl{r}}{\bsbl{x}}$.
+  An example superblock assignment from Figure~\ref{fig:main_example} is $(\sbl{r}, \bsbl{x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
+  Note that this assignment does not satisfy the signature yet because $\sbl{r}$ contains more than one vector.
+  This step aims to generate subassignments $(\sbl{r}' \sqsubseteq \sbl{r}, \mathbf{\sbl{x}}' \sqsubseteq \mathbf{\sbl{x}})$ that do satisfy the signature and test whether they satisfy the definition:
+  $\forall i, \sbl{r}[i] = \sum_{j = 1}^{length(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$.
+  In Figure~\ref{fig:main_example} there is exactly one such subassignment: $(\range{T_1}{\rangeall}{7}, \range{T_1}{\rangeall}{\rangeto{3}{6}})$.
+\end{example}
+
+In this step, we could also formulate the problem as a CSP.
+%However, the search space is much smaller, e.g.
+$O(nm^2)$ with $n$ the number of arguments of the template and $m$ the size of the largest block.
+However, few CSP solvers support floating point numbers, which are prevalent in spreadsheets.
+Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
+This is inefficient, as already such blocks may not satisfy the signature or the first elements in the data may not satisfy the definition.
+
+On the other hand, a simple generate-and-test approach, as illustrated in Algorithm~\ref{algo:gen-test-subblocks}, will usually suffice.
+%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
+%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
+Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
+For every (disjoint) subassignment the exact signature (e.g. subtypes will not have been checked yet) and definition will be tested and all satisfying subassignments are returned.
+
+\begin{algorithm}[t]
+  \begin{algorithmic}
+    \footnotesize
+    \Procedure{Subassignments}{$s$, $(\sbl{1}, \dots, \sbl{n})$}
+    \State $n \gets$ number of arguments of template $s$
+    \State $A_{sub} \gets \emptyset$
+    \ForAll{$(\sbl{1}', \dots, \sbl{n}') \textbf{ where } \forall i: \sbl{i}' \sqsubseteq \sbl{i}  \land \textit{disjoint}_{j \neq i}(\sbl{i}', \sbl{j}')$}
+      \If{$\textit{Sig}_s(\sbl{1}',\dots,\sbl{n}') \land \textit{Def}_s(\sbl{1}',\dots,\sbl{n}')$}
+        \State $A_{sub} \gets A_{sub} \cup \{ (\sbl{1}', \dots, \sbl{n}') \}$
+      \EndIf
+    \EndFor
+\Return $A_{sub}$
+\EndProcedure
+\end{algorithmic}
+\caption{Generate-and-test for $\findassignment$}
+\label{algo:gen-test-subblocks}
+\end{algorithm}
+
+% \begin{verbatim}
+% VERY PSEUDOcode:
+% something recursive that handles vectors/
+% subgroups differently like this?
+% check(Groups, X, Test):
+%   if |Groups| == 0:
+%     forall i in ?: vector elements...
+%       if not Test(X):
+%         return []
+%     return X
+%   // else
+%   all = []
+%   Group = Groups.pop(0)
+%   if vector(Group):
+%     forall Vec in Group:
+%       all += check(Groups, X + Vec, Test)
+%   else:
+%     forall Subgroup in Group:
+%       all += check(Groups, X + Subgroup, Test)
+%   return all
+% \end{verbatim}
+
+ %$G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
+\samuel{\tias{How are they stopped in the code?
+should be part of the pseud-code because not part of $Test(X)$?} will make it part of pseudo-code}
+
+% \paragraph{Example 2, not used (should be replaced by pseudo-code above?)}
+% Sum Product
+
+% Implementation: finds subgroups based on assignments\\
+% Method $Test(G_R, G_{O_1}, G_{O_2})$\\
+% $If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
+% $If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
+% \\\\
+% Method $FindSubgroups(Assignments)$\\
+% $Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g.
+vectors)?}
+
+\subsection{Optimizations} \label{sec:opts}
+This section discusses various design decisions and optimizations aimed at improving the speed, the extensibility of our system and the elimination of redundant constraints.
+
+\subsubsection{Template dependencies}
+As discussed in Section~\ref{sec:form:dependencies}, some constraint templates depend on others by including templates they depend on in their signature (see Table~\ref{table:constraints}).
+
+In Inductive Logic Programming, one often exploits implications between constraints to structure the search space~\cite{luc_book}.
+Our approach uses the dependencies between templates to define a dependency graph.
+We assume that signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
+Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signatures in Table~\ref{table:constraints}. 
+Constraint templates that have no dependencies are omitted.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.50\linewidth]{figures/constraint_dependency2.pdf}
+  \caption{Dependency graph of Table~\ref{table:constraints}; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (its signature includes $s_1$).
+  }
+  \label{fig:learning_order}
+\end{figure}
+
+Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
+Concretely, in Algorithm~\ref{algo:tcl} line \ref{algo:tcl:for} the input templates~$\temps$ are handled following an ordering that agrees with the partial ordering imposed by~\dependencies (e.g.
+\textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
+When using this ordering, templates later in the ordering can use the learned constraints of the templates they depend on.
+To enable this reuse of constraints, the set of learned constraints $C$ is added to the argument of \generategroups on line \ref{algo:tcl:super}.
+
+The previously learned constraints are used to speed up the \generategroups step.
+For constraint templates where some arguments are also part of a constraint that it depends on, it is not needed to search for superblocks and subblocks from scratch.
+Instead, one can start from all and only the actual argument assignments of the base constraint, and only search matching superblocks for the remaining arguments.
+
+\begin{example}
+  Consider $\ecfkey{\sbl{fk}}{\sbl{pk}}$, which states that every value in $\sbl{fk}$ also exists in $\sbl{pk}$; its signature includes $\ecalldiff{\sbl{pk}}$.
+There are 18 \textit{ALLDIFFERENT} constraints to be found in Figure~\ref{fig:main_example}, hence, the \generategroups only needs to check which superblocks for $\sbl{fk}$ are compatible (different table, same type) with these 18 assignments to $\sbl{pk}$.
+\end{example}
+
+Instead of generating one CSP to find all assignments, a CSP is generated for every known assignment of the depending constraint (and variables involved), which then searches for all assignments completing this partial assignment.
+The procedure to generate CSPs remains the same otherwise.
+
+%\subsubsection{Constraint solvers}
+%\samuel{Tias, Sergey, have a look at his section, I'm torn about rewriting it, feeling it should either be reduced or expanded, especially the second part.}
+%Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
+%Constraint solvers can be categorized as being stand-alone solvers, meaning that they have a text-based modeling language in which the CSP must be formulated, and \textit{embedded}, meaning that they have a programmatic API to formulate the CSP.
+%
+%\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the MiniZinc language \cite{minizinc} with the stand-alone Gecode 4.4.0 solver (which supports floats) and uses an embedded Python-based CP solver \cite{python_constraint}.
+%The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints.
+The embedded solver is fastest as it does not have the overhead of file-writing and calling an external program, and is used in the experiments below.
+
+\subsubsection{Redundancy}
+We consider two types of redundancy that we aim to eliminate during search.
+As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{\sbl{R}'}{\sbl{1}'}{\sbl{2}'} \Leftrightarrow \ecprod{\sbl{R}'}{\sbl{2}'}{\sbl{1}'}$.
+Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
+In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
+Moreover, among the semantically equivalent constraints product (\ecprod{\sbl{R}'}{\sbl{1}'}{\sbl{2}'}) and division (\eccalc{\sbl{1}'}{\sbl{R}' / \sbl{2}'}) we only added product, as well as only difference (\ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}) and not sum (\eccalc{\sbl{1}}{\sbl{r} + \sbl{2}}).
+However, division and sum could easily be added, as they could be added explicitly in post-processing based on the matching product and difference constraints.
+
+A last type of redundancy we chose to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
+For example, consider the constraint $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{n}}}$ where $\range{T}{\rangeall}{\rangeto{k}{n}}$ consists of only zeros, then $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{j}}}$ will be true for all $k - 1 \leq j \leq n$.
+In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.
+
+In some cases a maximal subblock might falsely include irrelevant columns.
+Consider superblock $\bsbl{x} = \range{T}{\rangeall}{\rangeto{1}{3}} = [[200, 300], [200, 150], [1, 2]]$ and $\sbl{r} = [200, 300]$, then \sname will find the constraint $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$.
+Should the target constraint be $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$, then $\range{T}{\rangeall}{3}$ must be split off into a separate block.
+
+Alternatively, we can output all valid assignments instead of only maximal subblocks, e.g.
+$\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{1})}$, $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$ and $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$, however, this might produce many redundant constraints including constraints that accidentally cover too little data.
+
+\subsubsection{Constraints}
+The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
+We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
+We also added three structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
+
+For \textit{FOREIGNKEY}, \textit{LOOKUP}, aggregates and \textit{PROJECT} we use custom, optimized methods to find subassignments.
+Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
+While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
+
+%Their dependencies are reflected in their \CSignature and shown in Figure~\ref{fig:learning_order}.
+%These constraints have been selected based on the most-used Excel functions as well as their occurrence in various example spreadsheets we observed.
+
+%Note that \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?}
+
+% Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}
+
+% \sergey{under question now, might worth deleting} \\
+% \sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}
+
+% \begin{equation}
+%   \forall i{:}~ v_r[i] \leq v[i] < v_r[i] + 0.5
+%   \label{eq:floor_extension}
+% \end{equation}
+
+% \subsection{Workflow}
+% \begin{algorithm}[thb]
+%   \begin{algorithmic}
+%     \footnotesize
+%     \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
+%     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
+%     \If{$T$ is \textbf{not} provided}
+%       \State $T \gets \extracttables(D)$
+%     \EndIf
+%     \If{$G$ is \textbf{not} provided}
+%       \State $G \gets \extractgroups(D, T)$
+%     \EndIf
+%     \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
+%     \State \Return $S$
+% \end{algorithmic}
+% \caption{Workflow}
+% \label{algo:workflow}
+% \end{algorithm}
+
+% \textbf{Approach}
+% \begin{itemize}
+%   \item Notation
+%   \item Algorithm (select constraints, find assignments, find solutions)
+% \end{itemize}
+
+% \samuel{Move next to subgroup satisfaction search}
+% \section{Declarative modeling}
+% \sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
+
+\newcommand{\runtotal}{16.12}
+\newcommand{\runtotalstd}{0.62}
+
+\newcommand{\runfile}{0.50}
+\newcommand{\runfilestd}{0.02}
+
+\newcommand{\benchsize}{??}
+
+\section{Evaluation}\label{sec:evaluation}
+In this section we experimentally validate our approach.
+We studied three main questions concerning the recall (\textit{Q1}), precision (\textit{Q2}) and speed (\textit{Q3}) of our algorithm.
+
+First, our implementation is illustrated using a case study on the spreadsheet corresponding to the example introduced in Figure~\ref{fig:main_example}.
+In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 34 spreadsheets that we assembled from various sources.
+
+\subsection{Method}
+\label{sec:evalualtion:method}
+In order to obtain a benchmark of spreadsheets we looked at three sources:
+1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports (available at the same repository\footref{github-link}).
+All spreadsheets were converted to CSV and the table definitions are fixed.
+Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
+
+For every spreadsheet we specified a ground-truth of constraints, i.e. the \textbf{intended} constraints that are expected to be (re-)discovered.
+These were determined using the formulas (if present), context and headers in the original spreadsheets.
+We encoded all intended constraints using our representation (in canonical form).
+Not all intended constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
+Constraints that are currently outside of the scope of this system are, for example, nested mathematical or logical formulas.
+
+We focus on \textit{functional} constraints that could be used as formulas in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
+Moreover, we included detected equalities as intended constraints (in 5~spreadsheets).
+All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
+
+\subsection{Case study}
+Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
+\sname takes a few seconds to find the constraints in Figure~\ref{fig:sol_example} (excluding structural constraints).
+These include 5 spurious $\textit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
+Moreover, there is 1 \textit{LOOKUP} constraint that was not intended in the original spreadsheet (looking up ID based on Salesperson) and which is symmetric to the intended one (Salesperson based on ID).
+
+For this example our primary goal of rediscovering all intended constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{all intended}}$) is~$1.0$.
+The implementation also returns 6 additional constraints, therefore, our precision (i.e.
+$\frac{\text{intended discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is mostly due to the spurious RANK constraints.
+We believe this is due to the many short vectors which increases the chance for constraints to be true by accident.
+
+\subsection{Benchmark}
+Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
+Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
+
+{\setlength{\tabcolsep}{0.33em}
+\begin{table}[]
+  \centering
+  \caption{Summary of properties and results per spreadsheet category}
+  \label{tbl:category_overview}
+  \begin{tabularx}{\linewidth}{lccccccc}
+ & \multicolumn{2}{c}{\textbf{Exercises} (9)} & \multicolumn{2}{c}{\textbf{Tutorials} (21)} & \multicolumn{2}{c}{\textbf{Data} (4)} \\
+ & Overall & Sheet avg & Overall & Sheet avg & Overall & Sheet avg \\ \hline
+    \textbf{Tables} & 19 & 2.11 & 48 & 2.29 & 4 & 1 \\ \hline
+    \textbf{Cells} & 1231 & 137 & 1889 & 90 & 2320 & 580 \\ \hline
+    \textbf{\begin{tabular}[c]{@{}l@{}}Intended\\[-4.5pt] Constraints\end{tabular}} & 34 & 3.78 & 52 & 2.48 & 6 & 1.50 \\ \hline \hline
+\textbf{Recall} & 0.85 & 0.83 & 0.88 & 0.87 & 1.00 & 1.00 \\ \hline
+\textbf{\begin{tabular}[c]{@{}l@{}}Recall\\[-4.5pt] Supported\end{tabular}} & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ \hline
+\textbf{Precision} & 0.97 & 0.98 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
+\textbf{Speed (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
+\end{tabularx}
+\end{table}}
+
+\subsubsection*{Q1. How many intended constraints are found by \sname?}
+The overall recall achieved by \sname is~$0.88~(81/92)$. The overall and average recall per category is also high, as shown in Table~\ref{tbl:category_overview}.
+The intended constraints not discovered correspond to constraints not supported by the system (supported recall is 1).
+
+\subsubsection*{Q2. How precise is \sname?}
+Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
+While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.97$ and $1.00$) than for the tutorials category ($0.70$).
+Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
+A qualitative analysis of these spreadsheets shows that their lower precision is caused by duplicate data or multiple ways to calculate the same vectors.
+For these spreadsheets the precision could be increased by post-processing the constraints to detect implications or heuristically selecting one among multiple ways to compute the same vector.
+
+\subsubsection*{Q3. How fast is \sname?}
+The average runtime across all spreadsheets is $0.13s$.
+Across categories the average running times fluctuate in the range between $0.09s$ to $0.22s$.
+While most spreadsheets can be analyzed very quickly, running times can go up to around $1s$ for others.
+
+% Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
+% The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
+% For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).
+
+%We explained various optimizations, most notably the exploitation of dependencies.
+%Using the dependencies to find constraints incrementally cuts down large parts of the search trees, especially for constraints that have many arguments.
+%Essentially, one obtains a more specialized search without much effort.
+
+Of the 32 supported constraint templates (including equality), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
+It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates (each $3.5\% - 4\%$).
+
+To illustrate that using dependencies to find constraints incrementally can increase the efficiency, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF.
+This assumption is too strong in practice, not all keys have to be used in the conditions.
+However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data), showing the strong effect of the use of dependencies.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{figures/comparison.pdf}
+  \caption{Overview of the average properties and performance of spreadsheets within every benchmark category relative to the maximum across all spreadsheets.
+  Green bars indicate standard deviation within the category.}
+  \label{fig:comparison}
+\end{figure}
+
+\subsection{Testing the limits}
+Our algorithm relies on a correct table and group detection step, and we noticed that exported CSV files can make this more difficult in case of rounding errors due to low-precision formatting as well as misformatting of strings so they are interpreted as numbers.
+Automatically generated sheets can also contain many duplicate (equal) rows or columns, which are a source of increased runtime and decreased recall for our algorithm.
+For example, we tested a spreadsheet (``exps'') with 6~tables, 531~cells and many duplicate columns across tables.
+While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when counting constraints with equal columns separately.
+%By using a (heuristic) post-processing step the precision for this example could be increased.
+%The running time can be improved by implementing additional optimizations.
+
+\section{Applications}\label{sec:applications}
+In this section, we illustrate how various motivating applications could be realized using our system.
+
+\subsection{Autocompletion}
+Autocompletion can be seen as using knowledge derived from the current data to predict new values before they are written by the user.
+A snapshot~($D_0$) is used to learn constraints and these are combined with new data~($\Delta D$) that the user has added.
+Using~$\Delta D$ we can keep track of which vectors have changed and which vectors are now \textit{incomplete}.
+\textit{Functional} constraints, such as $\ecsumc{\sbl{r}}{\mathbf{\sbl{x}}}$, are able to compute a result~($\sbl{r}$) based on its \textit{input}~($\mathbf{\sbl{x}}$).
+If all vectors in the input blocks have changed and the result vector is incomplete, the constraint can compute the resulting value(s).
+
+Consider Figure \ref{fig:autocompletion_example} containing two tables.
+Learning on a snapshot that does not contain the row below $T_2$, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
+The new data $\Delta D$ consists of the new value \textit{West} in the row below $T_2$.
+Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
+The \textit{SUMIF} constraints have enough input to fill the incomplete vectors and predict the values in the last row: $\range{T_2}{4}{2} = 1547$ and $\range{T_2}{4}{3} = 428128$.
+
+%In case multiple constraints predict different values for the same cell, the algorithm will have to either prioritize constraints based on their specificity or a heuristic, not make a prediction, or present the choice to the user.
+
+\subsection{Error detection}
+\begin{figure}[t]
+% %\begin{subfigure}{.70\textwidth}
+%   \begin{center}
+%     \includegraphics[width=\width]{figures/fbi_figure_highlighted.png}
+%   \end{center}
+%   \caption{Real world tabular constraint reconstruction: FBI crime statistics}
+%   \label{fig:fbi}
+%\end{subfigure}
+% \hfill
+% \begin{subfigure}{.29\textwidth}
+\begin{center}
+    \includegraphics[width=1.0\linewidth]{figures/Learning.png}
+  \end{center}
+  \caption{Autocompletion works by learning constraint on a snapshot and combining those constraints with new data $\Delta D$.
+  The figure shows how, given a new value \textit{West}, we can suggest values for the other cells in that row.}
+  \label{fig:autocompletion_example}
+%\end{subfigure}
+\end{figure}
+
+We consider two cases of error detection.
+The first setting is the \textit{online} detection of errors and is similar to the autocompletion setting.
+Using snapshots, we search for \textit{conflict} constraints, which are learned on a previous snapshot but do not agree with newly added data~$\Delta D$.
+Such conflicts can indicate (typing) errors in the data just entered. For example, in case 'Wes' instead of 'West' was added in Figure~\ref{fig:autocompletion_example}.
+%Conflict constraints can indicate errors in the input.
+%However, such conflicts can be wrongly caused by spurious constraints.
+%Hence, conflicts should perhaps only be detected if the snapshot data is large enough.
+%An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.
+
+The second setting is \textit{offline} error detection in a static spreadsheet.
+% Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
+% Our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
+% A possible approach to detect such missing or wrong values is to look at a sample of the table, e.g.
+trying to leave out each row, and compare the constraints learned on this sample with the constraints learned on the whole table.
+  %The sample needs to be large enough such that we have enough confidence that the constraints are correct.
+For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
+Instead of snapshots, one can repeatedly search for conflict constraints between the original spreadsheet and a sheet with a row or column removed.
+Many schemes can be devised for choosing which row(s) or column(s) are removed.
+%This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g.
+leave out each row and consider it as $\Delta D_i$) to detect conflicts.
+
+%Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
+%In the limited case of detecting single errors one can leave out every row (column), consider the remaining data to be the sample and
+%and the left out row (column) to be new input.
+
+\section{Related Work}\label{sec:related_work}
+\sname combines ideas from several existing
+approaches and different fields of research.
+
+First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
+constraints in multiple relations (or tables). However, unlike typical logical and relational learning
+approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
+Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
+Nevertheless, several ideas from logical and relational learning
+proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
+Also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha}
+although that line of research has focused on more specialized techniques for specific types of constraints. Furthermore, the constraints and formulas used
+in spreadsheets are often different from those in databases.
+
+Second, there exist a number of algorithms in the constraint programming community that induce constraint (programs)
+from one or more examples and questions. Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
+The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
+To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
+Key differences are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and often typical for constraint programming community, e.g.
+aimed at combinatorial optimization and scheduling.
+Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user regarding satisfiability.
+Each of the examples/queries is processed for each constraint, hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
+for Sudoku).
+The focus is again on integer variables and typical combinatorial optimisation constraints.
+\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
+It also connects with the approach of inducing constraints and functions by generalizing given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
+The key difference here that we start not from existing formulas but from the data directly.
+%As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
+As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.
+
+Third, our work borrows from the seminal work of Gulwani et al.
+\cite{flashfill} on program synthesis for tabular data, which is incorporated in Microsoft’s Excel.
+In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians.
+The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''.
+At that point FlashFill generates a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', and also applies it to the other politicians (rows) in the table.
+While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
+Flashfill has been extended to e.g.
+FlashExtract \cite{flashextract} to produce a set of tables starting from an unstructured document (a txt or HTML file) with a set of examples marked for extraction.
+However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.
+In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.
+Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
+
+\sergey{added database dependency discovery; have a look at it} Fourth, our work is motivated by the functional dependency discover in databases \cite{flach_dependency_discovery}.
+Most of the discovery techniques rely on the database schema that does not directly apply in our case.
+Some techniques work directly on the data \cite{tane_dependency_discovery} and, contrary to our work, are focused on pure functional dependencies that are not typical constraints of a spreadsheet that might hold over multiple tables and on a mixture of columns and rows.
+
+Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
+However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify ``hard'' constraints.
+
+\section{Conclusions}\label{sec:conclusions}
+
+% points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
+Our goal is to automatically identify constraints in a spreadsheet.
+We have presented and evaluated our approach, implemented as the system \sname, that is able to learn many different constraints in tabular data.
+The resulting method has high recall, produces limited number of redundant constraints and is sufficiently efficient for normal and interactive use.
+
+The approach is designed to find constraints that hold over entire columns and rows.
+In future work we plan to extend this to learn arbitrary nested constraints (e.g.
+$\sbl{r} = (\sbl{1} + \sbl{2}) / \sbl{3}$), as well as constraints over only a subset of the vectors.
+This may give rise to more redundant and spurious constraints, which was not problematic up to this point.
+
+Two promising ways to mitigate redundant and spurious constraints are heuristic filtering or post-processing of the constraint set and the integration of our approach in an interactive setting where users can receive and provide feedback.
+The latter is more similar to an active learning setting.
+
+A final direction is that the current approach assumes noise-free data (but see the part on error correction); in inductive logic programming there has been much work on how to handle noisy data, while in constraint satisfaction a popular alternative is to consider \textit{soft} constraints that can be violated (at a cost).
+This could extend the approach to new application domains, beyond traditional spreadsheets.
+
+\bibliographystyle{plain}
+{
+  \bibliography{references}}
+\end{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-09-08 Code/src/speed_tests.py
Added tables to speed test

@@ -12,28 +12,30 @@ from core.group import GType
 
 
 class SpeedTestId:
-    def __init__(self, cols, rows, blocks, types):
+    def __init__(self, cols, rows, blocks, types, tables):
         self.cols = cols
         self.rows = rows
         self.blocks = blocks
         self.types = types
+        self.tables = tables
 
     def __hash__(self, *args, **kwargs):
-        return hash((self.cols, self.rows, tuple(self.blocks), tuple(self.types)))
+        return hash((self.cols, self.rows, tuple(self.blocks), tuple(self.types), self.tables))
 
     def __eq__(self, other):
         if not isinstance(other, SpeedTestId):
             return False
         return self.cols == other.cols and self.rows == other.rows and self.blocks == other.blocks \
-            and self.types == other.types
+            and self.types == other.types and self.tables == other.tables
 
     def __repr__(self):
-        return "c{} r{} b{}".format(self.cols, self.rows, self.blocks)
+        return "c{} r{} b{} t{}".format(self.cols, self.rows, self.blocks, self.tables)
 
     @property
     def name(self):
         assert(all(b == self.blocks[0] for b in self.blocks))
-        return "c{}_r{}_b{}".format(self.cols, self.rows, "{}x{}".format(len(self.blocks), self.blocks[0]))
+        blocks = "{}x{}".format(len(self.blocks), self.blocks[0])
+        return "c{}_r{}_b{}_t{}".format(self.cols, self.rows, blocks, self.tables)
 
 
 def generate_random(test_id: SpeedTestId):
@@ -49,7 +51,8 @@ def generate_random(test_id: SpeedTestId):
         else:
             raise RuntimeError("Unexpected column type {}".format(column_type))
 
-    return list(list(str(column_data[col][row]) for col in range(test_id.cols)) for row in range(test_id.rows))
+    return list(list(str(column_data[col][row]) for col in range(test_id.cols) for _ in range(test_id.tables))
+                for row in range(test_id.rows))
 
 
 def generate_file(path, data_function, test_id: SpeedTestId, prefix):
@@ -68,17 +71,20 @@ def generate_file(path, data_function, test_id: SpeedTestId, prefix):
 
 def generate_blocks(test_id: SpeedTestId):
     content = dict()
-
     # Generate tables
-    bounds = [1, test_id.rows, 1, test_id.cols]
[…]

### 2016-09-08 Paper/ECAI/tabular.tex

@@
 Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
-
-In this paper,  we investigate whether learning techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
-This is a new and unconventional machine learning problem.
-Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
-Looking at the first row in table~$T_2$ and the data in table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
-Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
-This provides a flavor of how this problem is different from standard mining and learning settings, where the data is just in rows and variables are in columns.
-Here everything is mixed.
-The data is relational on the one hand, since we have multiple tables with relationships between them (e.g.
-foreign-keys).
-On the other hand, the data is most often mixed, e.g.
-textual and numeric.
-%, since people tend to use spreadsheets for financial and accounting computations.
-Therefore, understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
+
+% why learn them
+Such issues and mistakes could be overcome through the development of intelligent user-assisting tools.
+A core aspect of such tools would be the ability to automatically discover what constraints and formulae are present in a spreadsheet.
+This is the problem we study in this paper.
+We see multiple promising applications that can build on this, such as rediscovery constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
+In spirit, this builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
+
+% why difficult
+In this paper,  we investigate whether knowledge discovery techniques can be used to learn constraints (formulas and other relations) in spreadsheet data in an unsupervised way.
+From a machine learning point of view this is an unconventional problem because the data is in tabular form, but the constraints we wish to learn can involve both rows and columns of the table.
+Being unsupervised there is no labeled information either, although for every possible function one can verify whether a certain input satisfies the definition.
+%Unlike in other knowledge discovery tasks, such as pattern mining~\cite{}, there are many different types of formula and constraints we wish to discover, each with their own signature (number and properties of arguments) and definition.
+From a data mining point of view, the data is relational on the one hand, since we can have multiple tables with relationships between them, but also of mixed textual and numeric types.
+More closely related is clausal discovery~\cite{claudien,lallouet}, learning CSP constraints \cite{Quacq,Conacq,modelseeker} and dependency discovery in databases \cite{savnik}.
+But an important difference remains that we want to learn constraints on both columns and rows over integer, floating point as well as textual data.
 
 % Moved up
-The question that we answer in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
-It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
+The question that we ask in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
+The main challenge is the amount of combinations of rows and columns that need to be tried as input to the constraints.
+To answer this challenge we propose a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
+%
+% contribs
+Our contributions are as follows:
+\begin{itemize}
+\item we define the tabular constraint learning problem, where the goal is to learn constraints (functions and relations) that range over entire rows or columns in an unsupervised way;
+\item we propose a 2-stage generate and test method that reasons for each candidate constraint over maximally contiguous blocks in the first stage, and only in the second stage looks at individual rows and columns and their content;
+\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximal blocks compatible with the candidate constraint's argument, using backtracking search and forward checking;
+\item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall. 
+\end{itemize}
+
+%It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
 %, and discovers column- and row-wise constraints.
-We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch additional application scenario's such as autocompletion and error detection.
-
-The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
-and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
-very few examples.
-Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.
+%We demonstrate the utility of our approach in an experimental evaluation.
+%Moreover, we sketch additional application scenario's such as autocompletion and error detection.
 

### 2016-09-08 Paper/ECAI/tabular.tex
Expanded runtime analysis

@@
 However, in practice, most subassignments will not satisfy the constraint template.
-Since our implementation tries to account for this by failing fast, the amortized time is constant in the size of the vectors~$k$, unless $k$~is very large or many constraints hold in the data.
+Since our implementation tries to account for this by failing fast, the amortized time is constant in the size of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
 For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).
@@
 The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
-Synthetic experiments with random integers and strings show that increasing the size of the vectors has a neglegible impact on the performance.
+Synthetic experiments with random integers and strings show that increasing the size of the vectors has a negligible impact on the performance for most templates.
+
+Aggregate constraints, e.g. sum or max, have an argument that takes subblocks of varying sizes.
+Therefore, the size of the blocks in the spreadsheet influences the runtime of these constraints.
+For a block of size~$m$ there are $\frac{m(m + 1)}{2} = O(m^2)$ subblocks, thus, if a spreadsheet has smaller sized blocks, the search for aggregate constraints will be faster.
+
+Synthetic experiments with random integers confirm that splitting a constant number of vectors into smaller blocks improves the performance of the aggregate constraint search.
+
+These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes very small.
+Since these constraints have single vector arguments the block size should not have a strong effect.
+However, in practice, for very small blocks to the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
+Since this last phase is more adept at generating these combinations, shifting this workload to the assignment phase slows down the algorithm.
 

### 2016-09-08 Paper/ECAI/tabular.tex
Finished first version runtime analysis

@@
 Since these constraints have single vector arguments the block size should not have a strong effect.
-However, in practice, for very small blocks to the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
-Since this last phase is more adept at generating these combinations, shifting this workload to the assignment phase slows down the algorithm.
+However, in practice, for very small blocks the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
+Since the implementation of that last phase is more adept at generating these combinations, shifting this workload to the assignment phase slows down the algorithm.
 

### 2016-09-08 Paper/ECAI/tabular.tex
Review related work

@@
 The focus is again on integer variables and typical combinatorial optimisation constraints.
-\sergey{added equation discover here, check if that's the right place for it and if it is in fact good enough description} In a similar spirit the field of equation discovery \cite{equation_discovery} focuses, guided by the heuristics based on the structure of the physical equations and unit measures, on the discovery of physical formulas over numbers and times series, but not on spreadsheet constraints.
+\sergey{added equation discover here, check if that's the right place for it and if it is in fact good enough description} In a similar spirit the field of equation discovery \cite{equation_discovery} focuses, guided by  heuristics based on the structure of physical equations and unit measures, on the discovery of physical formulas over numbers and times series.
+However, they focus specifically on physical data, while we consider numbers in general and different types of spreadsheet constraints.
 %As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
@@
 Many of of the discovery techniques rely on the database schema that does not directly apply in our case \cite{flach_dependency_discovery}.
-Some work directly on the data \cite{tane_dependency_discovery} and focus on classical functional dependencies between attributes, while the constraints and formulas used in spreadsheets are often different from those in databases.
+Some techniques work directly on the data \cite{tane_dependency_discovery} and focus on classical functional dependencies between attributes, while the constraints and formulas used in spreadsheets are often different from those in databases.
 

### 2016-09-08 Paper/ECAI/tabular.tex
added markers for luc to check

@@
 
-\title{Discovering constraints in spreadsheets and tabular data}
+\title{Learning \wat \sergey{here should be learning and not discovering, right?} constraints in spreadsheets and tabular data}
 % see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
@@
 In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data in an unsupervised way.
-Common spreadsheet formulas are represented as predicates whose arguments must satisfy the inherent properties of the function.
+Common spreadsheet formulas are represented as predicates whose arguments must satisfy the inherent properties of the function \wat \sergey{predicates but properties of a function?
+sounds fine?}.
 The challenge is to discover these without user information or training examples.
@@
 We propose a two-stage generate and test method where the first stage uses constraint solving techniques to efficiently reduce candidates in a generic way based on the function signature. 
-Our approach to this knowledge discovery task takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
+Our approach to this knowledge discovery \wat task takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
 We show that we are able to accurately discover constraints in spreadsheets from various sources.
@@
 % why difficult
-In this paper,  we investigate whether knowledge discovery techniques can be used to learn constraints (formulas and other relations) in spreadsheet data in an unsupervised way.
+In this paper,  we investigate whether knowledge discovery \wat techniques can be used to learn constraints (formulas and other relations) in spreadsheet data in an unsupervised way.
 From a machine learning point of view this is an unconventional problem because the data is in tabular form, but the constraints we wish to learn can involve both rows and columns of the table.

### 2016-09-08 Paper/ECAI/tabular.tex
uncommitted tweaks

@@
 First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
-constraints in multiple relations (or tables). However, unlike typical logical and relational learning
+constraints in multiple relations (or tables). Unlike typical logical and relational learning
 approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
-Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
+Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of a conjunction of literals as in clausal discovery \cite{claudien,lallouet}.
 Nevertheless, several ideas from logical and relational learning

### 2016-09-08 Paper/ECAI/tabular.tex
Added explanation for number of vectors

--- 2016-09-08 Paper/ECAI/tabular.tex
@@
 For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small). 
-
 The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
@@
 
+On the contrary, the number of vectors in a spreadsheet have a significant impact on the running time.
+Many constraints take single vector arguments, therefore, the number of subassignments to be tested depends on the number of vectors in the spreadsheets and the arity of the constraint.
+For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
+As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a large arity.
+This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors is increased.
+However, in our benchmark experiments there will be various factors that limit the rapid increase in running time because the assignment step can disqualify entire blocks that do not match the signature requirements, i.e.
+blocks of vectors with different lengths for the product constraint.
+
 Aggregate constraints, e.g. sum or max, have an argument that takes subblocks of varying sizes.
-Therefore, the size of the blocks in the spreadsheet influences the runtime of these constraints.
-For a block of size~$m$ there are $\frac{m(m + 1)}{2} = O(m^2)$ subblocks, thus, if a spreadsheet has smaller sized blocks, the search for aggregate constraints will be faster.
-
+Therefore, the size of the blocks in the spreadsheet also influences the runtime of these constraints.
+For a block of size~$m$ there are $\frac{m(m + 1)}{2}$ or $O(m^2)$ subblocks, thus, if a spreadsheet has more but smaller sized blocks, the search for aggregate constraints will be faster.
+Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$.
 Synthetic experiments with random integers confirm that splitting a constant number of vectors into smaller blocks improves the performance of the aggregate constraint search.

### 2016-09-08 Paper/ECAI/tabular.tex
applied the comments of Luc

--- 2016-09-08 Paper/ECAI/tabular.tex
@@
 
-\title{Learning \wat \sergey{here should be learning and not discovering, right?} constraints in spreadsheets and tabular data}
+\title{Learning constraints in spreadsheets and tabular data}
 % see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
@@
 In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data in an unsupervised way.
-Common spreadsheet formulas are represented as predicates whose arguments must satisfy the inherent properties of the function \wat \sergey{predicates but properties of a function?
-sounds fine?}.
-The challenge is to discover these without user information or training examples.
-% Knowledge discovery
+Common spreadsheet formulas are represented using predicates and expressions whose arguments must satisfy the inherent properties of the function.
+The challenge is to discover these without user information or labeled examples.
 We propose a two-stage generate and test method where the first stage uses constraint solving techniques to efficiently reduce candidates in a generic way based on the function signature. 
-Our approach to this knowledge discovery \wat task takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
+Our approach to this knowledge discovery task takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
 We show that we are able to accurately discover constraints in spreadsheets from various sources.
@@
 This is the problem we study in this paper.
-We see multiple promising applications that can build on this, such as rediscovery constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
-In spirit, this builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
+We discuss multiple promising applications that can build on this, such as rediscovery constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
+Our techniques build on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
 
 % why difficult
-In this paper,  we investigate whether knowledge discovery \wat techniques can be used to learn constraints (formulas and other relations) in spreadsheet data in an unsupervised way.
+In this paper,  we investigate whether machine learning and knowledge discovery techniques can be used to learn constraints (formulas and other relations) in spreadsheet data in an unsupervised way.
 From a machine learning point of view this is an unconventional problem because the data is in tabular form, but the constraints we wish to learn can involve both rows and columns of the table.
@@
 The question that we ask in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
-The main challenge is the amount of combinations of rows and columns that need to be tried as input to the constraints.
+The main challenge is the number of combinations of rows and columns that need to be tried as input to the constraints.
 To answer this challenge we propose a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
@@
 First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
-constraints in multiple relations (or tables). Unlike typical logical and relational learning
+constraints in multiple relations (or tables). However, unlike typical logical and relational learning
 approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
-Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of a conjunction of literals as in clausal discovery \cite{claudien,lallouet}.
+Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
 Nevertheless, several ideas from logical and relational learning
@@
 \paragraph{Visual}
-\sergey{Luc, can you check if the supplementary material at github is understandable for people?
-Do we have to add anything there like write that code would be available after?} The above assumptions do not hold for many tables.
+The above assumptions do not hold for many tables.
 However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/SergeyParamonov/TaCLe}\label{github-link}}).
@@
 \end{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\input{journal_header}
-\input{packages}
-\input{macros}
-
-\begin{document}
-
-\title{Learning constraints in spreadsheets and tabular data}
-% see http://www.acm.org/binaries/content/assets/publications/article-templates/sig-alternate-sample.tex
-
-\author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3 \institute{----------------------} }
-
-\maketitle
-
-\begin{abstract}
-  \sergey{Samuel! Add a reference to python-constraint somewhere, including the references list at the end}\\
-  \sergey{Tias, Samuel, Luc, should we even mention miniZinc or ASP here, or is it not relevant?}
-Spreadsheets, CSV (comma separated value) files and other tabular data representations are in wide use today.
-However, modeling, maintaining and discovering formulas in tabular data and spreadsheets can be time consuming and error-prone.
-In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data.
-We see multiple promising applications for this technique, e.g. in rediscovering constraints, auto-completion and error checking.
-Our method takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
-Common spreadsheet functions are represented as predicates whose arguments must satisfy a number of constraints.
-Constraint learning techniques are used to identify the predicates that hold between blocks of rows and columns of the tables.
-We show that our approach is able to accurately discover constraints in spreadsheets from various sources.
-\end{abstract}
-
-\section{Introduction}
-Millions of people across the world use spreadsheets every day.
-The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
-However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
-Many end-users lack the understanding of the underlying structures and dependencies in such sheets and the data they contain.
-This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
-In this case, often a comma-separated values (CSV) format is used meaning that all formulas are lost, including inter-sheet formulas and relations.
-%This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
-Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
-
-In this paper,  we investigate whether learning techniques can be used to infer constraints (formulas and other relations) from raw spreadsheet data.
-This is a new and unconventional machine learning problem.
-Consider the example in Figure~\ref{fig:main_example}, where header names such as \textbf{Total} and \textbf{Rank} already suggest the usage of spreadsheet operations such as \textit{SUM} or \textit{RANK}.
-Looking at the first row in table~$T_2$ and the data in table~$T_1$, it is clear that the values in this row were computed by summing the values in the corresponding columns (\textbf{1st Quarter} through \textbf{Total}) of $T_1$.
-Examining $T_1$ also shows that computations are not only performed column-wise but row-wise as well: the cells in the column \textbf{Total} are obtained by summing the rows of the previous four columns.
-This provides a flavor of how this problem is different from standard mining and learning settings, where the data is just in rows and variables are in columns.
-Here everything is mixed.
-The data is relational on the one hand, since we have multiple tables with relationships between them (e.g.
-foreign-keys).
-On the other hand, the data is most often mixed, e.g.
-textual and numeric.
-%, since people tend to use spreadsheets for financial and accounting computations.
-Therefore, understanding and learning constraints in tabular data is a novel and challenging problem for machine learning.
-
-% Moved up
-The question that we answer in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
-To answer this question we contribute a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
-It operates directly on (headerless) tables of a spreadsheet in an unsupervised setting, that is, it reasons on raw tabular data directly with no example constraint instantiations given.
-%, and discovers column- and row-wise constraints.
-We demonstrate the utility of our approach in an experimental evaluation.
-Moreover, we sketch additional application scenario's such as autocompletion and error detection.
-
-The approach that we introduce borrows techniques from logical and relational learning \cite{luc_book}, where the discovery of clausal constraints has been studied in inductive logic programming \cite{claudien,lallouet}; and from constraint learning,  where several approaches to learning sets of constraints have been developed in a constraint programming setting \cite{Quacq,Conacq,modelseeker};
-and from mining in databases, where inductive techniques have been applied to discover constraints (such as functional and multi-valued dependencies) in relational databases \cite{savnik}.
-It also builds on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from
-very few examples.
-Our approach contrasts with these in that it focuses on learning both column- and row-constraints across tables on integer, floating point as well as textual data.
-
-This paper is organized as follows.
-Section \ref{sec:formalization} introduces concepts relevant to our approach and our problem statement.
-Section \ref{sec:approach} presents our approach.
-Section \ref{sec:evaluation} presents the evaluation of the approach.
-Section \ref{sec:applications} shows how our system can be used for applications.
-Section \ref{sec:related_work} presents a more detailed discussion of the related work.
-Section \ref{sec:conclusions} provides conclusions.
-
-%Due to the complexity of numeric computations in spreadsheets, people often fail to grasp the properties of the data, which leads to the so called Spreadsheet risk\footnote{\url{https://en.wikipedia.org/wiki/Spreadsheet\#Spreadsheet_risk}} and even caused major flaw in the famous economics papers \cite{flaw_excel} and billion-losses in the financial industry \cite{spreadsheet_risk_loss}.
-Tabular constraint learning provides a potential cure by indicating learned constraint violations and suggest possible fixes.
-
-% {\it
-% \textbf{Motivation}:
-% \begin{itemize}
-%   \item USED -- File generated from model, model got lost, need to reconstruct
-%   \item Constraint programming is hard - is Excel hard?
-%   \item Avoid manual analysis, provide selection of constraints
-%   \item SOMEWHAT USED --Error checking
-%   \item Completion, gain speed and insights (Complicated constraints, also complicated to verify, too much output)
-% \end{itemize}
-
-% \textbf{Novelty:}
-% \begin{itemize}
-%   \item USED -- Unsupervised setting (contrary to flashfill, etc)
-%   \item Numeric, different constraints (contrary to single textual function solution in flashfill, etc)
-%   \item USED -- Data format (2D) -- data is no longer in rows like a classic ML or DM settings
-%   \item USED -- Declarative, general / modular, stacking of constraint problems
-% \end{itemize}
-%}
-
-\begin{figure}[thb]
-
-  \begin{subfigure}{1\textwidth}
-  \begin{center}
-    \includegraphics[width=1\textwidth]{figures/Demo2.png}
-  \end{center}
-  \vspace{-10pt}
-  \caption{
-    Example spreadsheet (black words and numbers only).
-    Green background indicates headerless tables, dark borders indicate maximal type-consistent blocks.
-    Most tables only contain type-consistent columns while table~$T_2$ also contains type-consistent rows.
-    However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
-    This example was created by combining several Excel sheets from several exercise sessions based on the book MS Excel 2010~\cite{excel_book}.
-    % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
-  }
-  \label{fig:main_example}
-\end{subfigure}
-\hfill
-\begin{subfigure}{1\textwidth}
-  {\scriptsize
-    \begin{align*}
-      % Table 1
-%
-%      &~\ecperm{\range{T_{1}}{\rangeall}{1}}, \ecperm{\range{T_{1}}{\rangeall}{8}} \\
-%
-      &~\ecseries{\range{T_{1}}{\rangeall}{1}}                                  && \ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{5}}^*     && \ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}}  \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{6}}^*     && \ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{10}}^*    && \ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{7}}       && \ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{3}}^*     && \range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
-      &~\ecrank{\range{T_{1}}{\rangeall}{8}}{\range{T_{1}}{\rangeall}{4}}^*    && \eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^*\\
-%
-      &~\ecsumr{\range{T_{1}}{\rangeall}{7}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}}   && \eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}} \\
-%
-      &~\ecaggif{\textit{SUM}}{\range{T_{1}}{\rangeall}{10}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}}   && \ecaggif{\textit{MAX}}{\range{T_{1}}{\rangeall}{11}}{\range{T_{3}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{3}}{\rangeall}{2}}
-%
-    % Table 2
-%     &~\ecsumc{\range{T_{2}}{1}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-%     &~\ecaggc{\textit{AVERAGE}}{\range{T_{2}}{2}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-%     &~\ecaggc{\textit{MAX}}{\range{T_{2}}{3}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}},  \\
-%
-%     &~\ecaggc{\textit{MIN}}{\range{T_{2}}{4}{\rangeall}}{\range{T_{1}}{\rangeall}{\rangeto{3}{7}}} \\
-%
-    % Table 4
-%     &~\ecsumc{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{\rangeto{3}{6}}} \\
-%
-%     &~\range{T_{4}}{\rangeall}{4} = PREV(\range{T_{4}}{\rangeall}{4}) + \range{T_{4}}{\rangeall}{2} - \range{T_{4}}{\rangeall}{3} \\
-%
-    % Table 5
-%     &~\eclookup{\range{T_{5}}{\rangeall}{2}}{\range{T_{5}}{\rangeall}{3}}{\range{T_{1}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}^* \\
-%
-%     &~\eclookup{\range{T_{5}}{\rangeall}{3}}{\range{T_{5}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}
-    \end{align*}}
-  \vspace{-10pt}
-  \caption{Constraints learned for the example (left), except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
-  Constraints marked with $*$ were not present in the original document}
-  \label{fig:sol_example}
-\end{subfigure}
-  \caption{Running example}
-\end{figure}
-
-\section{Formalization}\label{sec:formalization}
-Our goal is to automatically discover constraints between the rows and the columns of tables in a spreadsheet.
-This is applicable not just to data from spreadsheets, but to any data in tabular form, hence the name.
-
-%An example spreadsheet with tables is given in Figure~\ref{fig:main_example}, we will use this as running example throughout the paper.
-
-We first introduce some terminology and the concept of a \template, after which we define the problem and make some additional considerations.
-
-\subsection{Terminology}
-Spreadsheets and tabular data may conceptually consist of multiple tables, such as in Figure~\ref{fig:main_example}.
-Note that a table can contain a header; however, we wish to reason over entire rows and columns of data, and hence we will consider \textbf{headerless tables} only.
-
-Formally, a (headerless) table is an $n \times m$ matrix. Each entry is called a \textit{cell}.
-A cell has a {\bf type}, which can be numeric or textual.
-We further distinguish numeric types in subtypes: integer and float.
-We also consider \textit{None} as a special type when a cell is empty; \textit{None} is a subtype of all other types.
-
-A row or a column is \textbf{type-consistent} if all cells in that row or column are of the same base type, i.e. numeric or textual.
-We will use notation $T[a,{:}]$ to refer to the $a$-th row of table $T$ and similarly $T[{:},a]$ for the $a$-th column.
-For example, in Figure~\ref{fig:main_example}, $T_1[:,1] = [1,2,3,4]$ and $T_3[1,:] = ['$Diana Coolen$', 5]$.
-$T_3[1, :]$ is not type-consistent while $T_1[:, 1]$ is.
-%We will write \textbf{vector} to refer to a single row or column when its orientation does not matter.
-
-The most important concept is that of a \textbf{block}. 
-\begin{definition}
-A \textbf{block} has to satisfy three conditions: 1)~it contains only entire rows or entire columns of a single headerless table; 2)~it is contiguous; and 3)~it is type-consistent.
-The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type. 
-If a block contains only rows we say it has \textit{row-orientation}, if only columns, \textit{column-orientation}. 
-\end{definition}
-
-In line with this definition, we can use the following notation to refer to blocks: $B = T[\rangeto{a}{b},:]$ for a row-oriented block containing rows $a$ to $b$ in table $T$; and similarly $B = T[{:},\rangeto{a}{b}]$ for a column-oriented block.
-We will refer to the \textit{vectors} of a block when we wish to refer to its rows/columns independently of their orientation.
-
-A block has the following properties:
-\begin{itemize}
-\item \textit{type}: a block is type-consistent, so it has one type
-\item \textit{table}: the table that the block belongs to
-\item \textit{orientation}: either row-oriented or column-oriented
-\item \textit{size}: the number of vectors a block contains
-\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
-\item \textit{rows}: the number of rows in the block
-(in row-oriented blocks this is equivalent to the size)
-\item \textit{columns}: the number of columns in the block (in row-oriented blocks this is equivalent to the length)
-\end{itemize}
-
-%We call a group $G$ \textit{numeric} (\textit{textual}, etc), written as \textit{numeric(G)}, if all its vectors contain numeric (textual, etc) elements.
-
-%For notational convenience, we will refer to vectors in {\rm roman} and to groups in {\bf bold}. {\bf CHECK}
-
-\begin{example}
-Consider the (headerless) table~$T_1$ in Figure~\ref{fig:main_example}.
-Its rows are not type consistent (i.e. they contain both numeric and textual data).
-However, the table can be partitioned into five column-oriented blocks $\sbl{1}, \sbl{2}, \sbl{3}, \sbl{4}, \sbl{5}$, as shown in the figure ($\sbl{1} = \range{T_1}{\rangeall}{1}$, $\sbl{2} = \range{T_1}{\rangeall}{2}$, $\sbl{3} = \range{T_1}{\rangeall}{\rangeto{3}{8}}$, \dots).
-\end{example}
-
-\begin{definition}
-\textbf{Block containment $\sqsubseteq$.} 
-A block $B'$ is contained in a block $B$, $B' \sqsubseteq B$, iff both are valid blocks (contiguous, type consistent) with the same orientation and table, and each of the vectors in $B'$ is also in $B$.
-For row-oriented blocks: $B' \sqsubseteq B \Leftrightarrow B=\range{T}{\rangeto{a}{b}}{\rangeall} \wedge B'=\range{T}{\rangeto{a'}{b'}}{\rangeall} \wedge a \leq a' \wedge b' \leq b$ and similarly for column-oriented blocks.
-\end{definition}
-
-We will sometimes write that $B'$ is a \textit{subblock} of $B$ or that $B$ is a \textit{superblock} of $B'$.
-%
-An example of block containment is $T_1[:,\rangeto{3}{6}] \sqsubseteq T_1[:,\rangeto{3}{8}] $, which contains the sales numbers of all employees for the four quarters.
-
-\begin{table}[htb]
-\caption{
-  Overview of constraint templates implemented in \sname.
-  Subgroups in normal font have to be vectors, whereas subgroups in \textbf{bold} may contain more than one vector;
-  $\discrete(x)$ is a shortcut for: $\textual(x) \lor \numeric(x)$.
-  Templates marked with $*$ are \textit{structural}, i.e. they are not functional and cannot be implemented in most spreadsheet software.
-  Templates marked with $\dagger$ are \textit{aggregate} templates, the table shows \textit{SUM} but \sname also supports \textit{MAX}, \textit{MIN}, \textit{AVERAGE}, \textit{PRODUCT} and \textit{COUNT}.
-  %{\luc{maybe say s.th. about some fancy constraints: example of a more complex one in the text (what does conditional mean, aggregate ... )}}
-}
-\label{table:constraints}
-\resizebox{1.0\columnwidth}{!}{%
-  {\normalsize \centering
-  \begin{tabularx}{1.70\textwidth}{l X X}
-    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ \hline \hline
-    $\ecalldiff{\sbl{x}}^*$
-      & $\discrete(\sbl{x})$
-
-      & All values in $\sbl{x}$ are different. $i \neq j$: $\sbl{x}[i] \neq \sbl{x}[j]$
-      \\[\PADSEP] \hline
-
-    $\ecperm{\sbl{x}}^*$
-      & $\numeric(\sbl{x})$, $\ecalldiff{\sbl{x}}$
-
-      & The values in $\sbl{x}$ are a permutation of the numbers $1$ through $\plength(\sbl{x})$.
-      \\ \hline
-
-    \ecseries{\sbl{x}}
-      & $\integer(\sbl{x})$ and $\ecperm{\sbl{x}}$
-
-      & $\sbl{x}[1] = 1$ and $\sbl{x}[i] = \sbl{x}[i - 1] + 1$.
-      \\[\PADSEP] \hline
-
-    $\ecfkey{\sbl{fk}}{\sbl{pk}}^*$
-      & $\sbl{fk}$ and~$\sbl{pk}$ are $\discrete$; they must have different $\ptable$s but the same $\ptype$; and $\ecalldiff{\sbl{pk}}$
-
-      & Every value in~$\sbl{fk}$ also exist in~$\sbl{pk}$ \\[\PADSEP] \hline
-
-    \eclookup{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
-      & $\sbl{fk}$ and $\sbl{pk}$ are $\discrete$; arguments $\{\sbl{fk}, \sbl{r}\}$ and $\{\sbl{pk}, \sbl{val}\}$ within the same set have the same \plength, \ptable and \por; $\sbl{r}$ and~$\sbl{val}$ have the same type; and \ecfkey{\sbl{fk}}{\sbl{pk}}.
-
-      & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] = \sbl{fk}[i]$
-      \\[\PADSEP] \hline
-
-    \eclookupfuzzy{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
-      & \textit{Same as lookup}
-
-      & $\sbl{r}[i] = \sbl{val}[j]$ where $\sbl{pk}[j] \leq \sbl{fk}[i]$, $j$~maximal
-      \\[\PADSEP] \hline
-
-    \eclookupprod{\sbl{r}}{\sbl{1}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
-      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{fk}\}$ are $\numeric$, arguments $\{\sbl{pk}, \sbl{val}\}$ are $\discrete$ and within both sets all arguments have the same \plength, \ptable and \por; also \ecfkey{\sbl{fk}}{\sbl{pk}}.
-
-      & $\sbl{r}[i] = \sbl{1}[i] \times \eclookupf{\sbl{fk}}{\sbl{pk}}{\sbl{val}}{}[i]$.
-      \\[\PADSEP] \hline
-
-    \ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}}
-      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are all $\numeric$ and have the same $\plength$
-
-      & $\sbl{r}[i] = \sbl{1}[i] \times \sbl{2}[i]$.
-      \\[\PADSEP] \hline
-
-    \ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}
-      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are all $\numeric$ and have the same $\plength$ and $ \por$
-
-      & $\sbl{r}[i] = \sbl{1}[i] - \sbl{2}[i]$.
-      \\[\PADSEP] \hline
-
-    $\eccalc{\sbl{r}}{\frac{\sbl{1} - \sbl{2}}{\sbl{2}}}$
-      & \textit{Same as \ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}}
-
-      & $\sbl{r}[i] = (\sbl{1}[i] - \sbl{2}[i]) / \sbl{2}[i]$.
-      \\[\PADSEP] \hline
-
-    \ecproj{\sbl{r}}{\bsbl{x}}
-      & Arguments $\{\sbl{r}, \bsbl{x}\}$ all have the same $\plength$, $\por$, $\ptable$ and $\ptype$; $\bsbl{x}$ contains at least~2 vectors; and $\sbl{r} = \textit{SUM}_{\por(\bsbl{x})}(\bsbl{x})$
-
-      & At every position~$i$ in $1$ through $\plength(\sbl{r})$ there is exactly one vector~$v$ in $\bsbl{x}$ such that $v[i]$ is a non-blank value, then $v[i] = \sbl{r}[i]$.
-      \\[\PADSEP] \hline
-
-    \ecrank{\sbl{r}}{\sbl{x}}
-      & $\integer(\sbl{r})$; $\numeric(\sbl{x})$; and $\plength(\sbl{r}) = \plength(\sbl{x})$
-
-      & The values in $\sbl{r}$ represent the rank (from largest to smallest) of the values in $\sbl{x}$ (including ties)
-      \\[\PADSEP] \hline
-
-    \ectotal{\sbl{r}}{\sbl{pos}}{\sbl{neg}}
-      & Arguments $\{\sbl{r}, \sbl{pos}, \sbl{neg}\}$ are all $\numeric$ and all have the same $\plength$, which is at least $2$
-
-      & $\sbl{r}[i] = \sbl{r}[i - 1] + \sbl{pos}[i] - \sbl{neg}[i]$.
-      \\[\PADSEP] \hline
-
-    $\ecsumr{\sbl{r}}{\bsbl{x}}^\dagger$
-      & $\sbl{r}$ and $\bsbl{x}$ are $\numeric$; $\pcols(\bsbl{x}) \geq 2$; and $\prows(\bsbl{x}) = \plength(\sbl{r})$
-
-      & $\sbl{r}[i] = \sum_{j = 1}^{\pcols(\bsbl{x})} \textit{row}(i, \bsbl{x})[j]$
-      \\[\PADSEP] \hline
-
-    $\ecsumc{\sbl{r}}{\bsbl{x}}^\dagger$
-      & $\sbl{r}$ and $\bsbl{x}$ are $\numeric$; $\prows(\bsbl{x}) \geq 2$; and $\pcols(\bsbl{x}) = \plength(\sbl{r})$
-
-      & $\sbl{r}[i] = \sum_{j = 1}^{\prows(\bsbl{x})} \textit{column}(i, \bsbl{x})[j]$
-      \\[\PADSEP] \hline
-
-    $\ecsumif{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}^\dagger$
-      & $\sbl{fk}, \sbl{pk}$ are $\discrete$; $\sbl{r}, \sbl{val}$ are $\numeric$; within the sets $\{\sbl{val}, \sbl{fk}\}$ and $\{\sbl{pk}, \sbl{r}\}$ arguments have the same $\plength$ and $\por$; $\sbl{fk}$ and $\sbl{val}$ have the same $\ptable$; $\sbl{fk}$ and $\sbl{pk}$ must have different $\ptable$s but the same $\ptype$; and \ecalldiff{\sbl{pk}}
-
-      & \[\sbl{r}[i] = \sum_{j=1}^{\plength(\sbl{val})} \begin{cases}
-          \sbl{val}[j] & \text{if } \sbl{fk}[j] = \sbl{pk}[i] \\
-          0 & \text{otherwise}
-        \end{cases}
-      \] \\[\PADSEP] \hline
-
-    \ecsumprod{\sbl{r}}{\sbl{1}}{\sbl{2}}
-      & Arguments $\{\sbl{r}, \sbl{1}, \sbl{2}\}$ are $\numeric$; $\plength(\sbl{1}) = \plength(\sbl{2}) \geq 2$; and $\prows(\sbl{r}) = \pcols(\sbl{r}) = 1$
-
-      & $\sbl{r}[1] = \sum_{j = 1}^{\plength(\sbl{1})} \sbl{1}[j] \times \sbl{2}[j]$.
-    \end{tabularx}
-  } % centering small
-} % resizebox
-
-\end{table}
-
-\newcommand{\sigc}{\ensuremath{\format{Sig}_s}}
-\newcommand{\defc}{\ensuremath{\format{Def}_s}}
-
-\subsection{Constraint templates}
-The goal is to learn constraints over blocks in the data. The knowledge needed to learn a constraint is expressed through {\template}s.
-%
-A \template $s$ is a triple $t = \textit{(\CName, \CSignature, \CFunction)}$:
-%Let us elaborate on this:
-\begin{itemize}
-\item
-\textit{\CName}  specifies the syntactic form of the constraint $s(\sbl{1}, ...,\sbl{n})$, that is, the name of the template together
-with $n$ abstract arguments $\sbl{i}$.
-Thus, a constraint is viewed as a relation or predicate of arity $n$ in first order logic.
-Note that a function $\sbl{r}=f(\sbl{1},...,\sbl{n})$ can be represented with the $(n{+}1)$-ary predicate $s_f(\sbl{r},\sbl{1},...,\sbl{n})$.
-Each argument will have to be instantiated with a block.
-% the list of its variables $v_1,\dots,v_n$.  In ILP terminology, it is known as vocabulary.
-
-\item \textit{\CSignature} defines the requirements that the arguments of the predicate must satisfy.
-This can concern properties of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
-In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
-We can capture this bias for a template~$s$ using a predicate $\sigc$.
-\item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
-Given an assignment of blocks to its arguments, it can be used to verify whether the constraint is satisfied or not by the actual data present in the blocks.
-%holds, in practice this will be a function that can be called when
-%the arguments are full instantiated to decide whether the constraint is satisfied or not.
-In logical and relational learning this is known as the background knowledge.
-We introduce the predicate \defc~to capture this background knowledge for a template~$s$.
-\end{itemize}
-
-\begin{example}
-The constraint templates implemented in \sname are defined in Table~\ref{table:constraints}.
-A non-trivial example is, for example, the constraint template for the row-based sum:
-
-\begin{itemize}
-  \item \CName: $\ecsumr{\sbl{r}}{\bsbl{x}}$, for arguments $\sbl{r}$ and $\bsbl{x}$.
-
-  \item \CSignature: $\sbl{r}$ has to be a single vector ($\textit{size}=1$) while $\bsbl{x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
-  Both blocks have to be numeric.
-  This constraint is orientation-specific, so it requires that the number of rows in $\bsbl{x}$ equals the length of $\sbl{r}$.
-  Moreover, we add to the bias that the number of columns to sum over is at least~$2$.
-
-  \item \CFunction: each value in the vector $\sbl{r}$ is obtained by summing over the corresponding row in $\bsbl{x}$.
-  %For example, $\sbl{r} = [12, 7]$ and $\bsbl{x}$ consists of columns: $[5, 15]$, $[7, -8]$.
-\end{itemize}
-\textit{SUM} is an aggregate constraint, and a similar constraint is available for aggregate operators \textit{MIN}, \textit{MAX}, \textit{AVERAGE}, \textit{PRODUCT} and \textit{COUNT}.
-The \textit{conditional} aggregate variant \textit{SUMIF} only ranges over those cells that satisfy a condition on a related cell in that table, e.g.
-$\range{T_1}{\rangeall}{10}$ and the data in $T_5$ for matching salesperson names.
-\end{example}
-
-It is helpful to see the analogy of constraint templates with first order logic (FOL) and constraint satisfaction.
-From a FOL perspective, a constraint of the form $\ecrank{\sbl{r}}{\sbl{x}}$ can be seen as a predicate $\textit{RANK}(\sbl{r}, \sbl{x})$ where \textit{RANK} is the name of the predicate and its arguments~$\sbl{r}$ and~$\sbl{x}$ are terms, which can be seen as either uninstantiated variables or as concrete values.
-This also holds in our setting, where an instantiation of a variable corresponds to a concrete block.
-For example, for the spreadsheet in Figure~\ref{fig:main_example}, when we write $\ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{7}}$, then the value of $\sbl{r}$ is the $8$th vector in $T_1$: $\sbl{r} = \range{T_1}{\rangeall}{8} = [2,1,3,4]$ and the value of $\sbl{x}$ is the $7$th vector: $\sbl{x} = \range{T_1}{\rangeall}{7} = [1514, 1551, 691, 392]$.
-
-%When looking into constraints like \textit{$B$ = RANK($A$)}, it is helpful to see the analogy with both first order logic (FOL) and with constraint satisfaction problems.
-%From a FOL perspective, the name of the constraint ($\mathit{RANK}$) is just the predicate, and the arguments $B$ and $A$ are the terms, which can be seen as either uninstantiated variables or as values (concrete groups with values).
-%This also holds in our setting: when we write
-%      $\ecrank{\range{T}{\rangeall}{8}}{\range{T}{\rangeall}{7}}$,
-%we can interpret the argument ${\range{T}{\rangeall}{8}}$ as a group variable that would apply to any table $T$ (provided~$T$ has an 8th column).
-%So, $T$ is viewed as table variable here, with the effect that  ${\range{T}{\rangeall}{8}}$ is a group variable.
-
-With this interpretation, we can speak about the signature and definition of a constraint template being \textit{satisfied}.
-We say that a signature (definition) of a constraint template~$s$ with $n$ arguments is satisfied by the blocks $(\sbl{1}, ..., \sbl{n})$ if $\sigc(\sbl{1}, ..., \sbl{n})$ (respectively $\defc(\sbl{1}, ..., \sbl{n})$) is satisfied.
-Likewise, the template is satisfied if both the signature and definition are satisfied; in logic programming, we would define the predicate~$s$ using a Prolog like clause: $s(\sbl{1}, ..., \sbl{n}) \leftarrow \sigc(\sbl{1}, ..., \sbl{n}) \wedge \defc(\sbl{1}, ..., \sbl{n})$.
-Under this interpretation, the term constraint and constraint template can be used interchangeably.
-
-\begin{definition}
-A \textbf{valid argument assignment} of a constraint template~$s$ is a tuple of blocks $(\sbl{1}, ..., \sbl{n})$ such that $s(\sbl{1}, ..., \sbl{n})$ is satisfied, that is, both the signature and the definition of the corresponding constraint template are satisfied by the assignment of $(\sbl{1}, ..., \sbl{n})$ to the arguments.
-\end{definition}
-
-\subsection{Problem Definition}\label{sec:problem_statement}
-The problem of learning constraints from tabular data can be seen as an inverse {\em constraint satisfaction problem} (CSP).
-In a CSP one is given a set of constraints over variables that must all be satisfied, and the goal is to find an instantiation of all the variables that satisfies these constraints.
-In the context of spreadsheets, the variables would be (blocks of) cells, and one would be given the actual constraints and functions with the goal of finding the values in the cells.
-The inverse problem is, given only an instantiation of the cells, to find the constraints that are satisfied in the spreadsheet.
-
-We define the inverse problem, that is the {\bf Tabular Constraint Learning Problem}, as follows:
-%
-\begin{definition} \textit{Tabular Constraint Learning.}\label{def:problem_statement}\\
-{\bf Given} a set of instantiated blocks~${\cal B}$ over tables~${\cal T}$ and a set of {\template}s ${\cal S}$: {\bf find} all constraints $s(\ssbl{1}, ..., \ssbl{n})$ where $s \in {\cal S}$, $\forall i: \sbl{i}' \sqsubseteq \sbl{i} \in {\cal B}$ and $(\ssbl{1}, ..., \ssbl{n})$ is a satisfied argument assignment of the template~$s$.
-\end{definition}
-% Let us now define what it means for a constraint from \template to hold in general.
-Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
-% \sergey{moved from formalization END}
-
-%Here we formalize the statement in terms of \template and group assignments as follows:
-%   \begin{tabular}{ll}
-%     \multicolumn{2}{l}{{\textbf{Tabular Constraint Learning Problem}}}\\
-%     \textbf{Given:}& the set of all groups $\groups$ and of \template $\constraints$\\
-%     \textbf{Find:}&  all constraints $C$ over \groups for each template $t$ in \constraints \\
-%   \end{tabular}
-
-The input is a set of blocks, and in Section~\ref{sec:approach} we will discuss how these can be extracted from a spreadsheet.
-Figure~\ref{fig:sol_example} shows the solution to the tabular constraint learning problem when applied on the blocks of Figure~\ref{fig:main_example} and constraint templates listed in Table~\ref{table:constraints}.
-
-\subsection{Other considerations}
-
-\subsubsection{Dependencies}
-\label{sec:form:dependencies}
-In Table~\ref{table:constraints} one can see that for some constraints we used the predicate of another constraint in its signature, e.g. for \textit{PERMUTATION}.
-This expresses a dependency of the constraint on that other \textit{base} constraint.
-This can be interpreted as follows: the signature of the constraint consists of its own signature plus the signature of the base constraint, and its definition of its own definition plus the definition of the base constraint.
-In FOL, we can see that one constraint entails the other, for example if $\ecperm{\sbl{x}}$ holds for a block $\sbl{x}$, then $\ecalldiff{\sbl{x}}$ also holds.
-
-In Section~\ref{sec:approach} we will see how, apart from easing the specification of the signature and definition, such dependencies can be used to speed up the search for constraints.
-
-\subsubsection{Redundancies}
-\label{sec:form:redundancies}
-Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}.
-This is because constraints may be logically equivalent or may be implied by other constraints, e.g.
-if you know $\ecperm{\sbl{x}}$, $\ecalldiff{\sbl{x}}$ is redundant.
-
-Within the same constraint, there can be equivalent argument assignments if the order of some of the arguments does not matter.
-For example for the product constraint, $\ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}} \equiv \ecprod{\sbl{r}}{\sbl{2}}{\sbl{1}}$ so one can be considered redundant to the other.
-Two different constraints can also be logically equivalent, due to their nature e.g.
-for addition / subtraction and product / division: $\ecprod{\sbl{r}}{\sbl{1}}{\sbl{2}} \equiv \ecdiv{\sbl{1}}{\sbl{r}}{\sbl{2}}$.
-
-Finally, when the data has rows or columns that contain \textit{exactly} the same values, then any constraint with such a vector in its argument assignment will also have an argument assignment with the other equivalent vectors.
-
-Dealing with redundancy is often application-dependent.
-Therefore, in Section~\ref{sec:approach} we first explain our generic method for finding all constraints, before describing some optimizations that avoid obvious equivalences (Section~\ref{sec:opts}).
-Later, we will investigate the impact of redundancy in the experiments.
-
-%More formally, a constraint $c_1(G_1, ... G_n)$ \textit{implies} another constraint~$c_2(G'_1, ..., G'_m)$ whenever
-%the  following statement holds: $$\forall G_i, G'_j : c_1(G_1, ... G_n) \rightarrow c_2(G'_1, ... G'_m).$$
-%
-%We shall only consider {\em range-restricted} implications, that is, we shall require that $\{G'_1, ..., G'_m\} \subseteq \{G_1, ..., G_n\}$.
-%To simplify notation, and in line with logic programming, we shall often make the quantification implicit and simply write this as rules $ c_1(G_1, ...
-G_n) \rightarrow c_2(G'_1, ...
-G'_m)$.
-We say that the first constraint is more specific than the second, and the second is more general.
-%
-%
-%% where the second one ranges over a subsets of the arguments of the first, iff whenever the first constraint is true, the second must also be true.
-More formally, $\forall G'_1, ..., G'_n$ s.t.
-$\{G''_1, ..., G''_m\} \subseteq \{G'_1, ..., G'_n\},\, c_1(G'_1, ...
-G'_n) \rightarrow c_2(G''_1, ..., G''_m)$ll.
-%
-%\paragraph{Example}
-%Consider the following implication:
-%$$\ecperm{G} \rightarrow \ecalldiff{G}$$
-%(it can be seen in Table~\ref{table:constraints} that both constraints only hold if $G$ is a vector).
-%In this case the arguments are identical in both constraints.
-%Given that the problem is to find the set of {\em all} constraints, and that the more general constraint is entailed
-%by the more specific one, the $\ecdiff{G}$ constraint will be {\em redundant}.
-%So, for most applications, it will be sufficient to output only the most specific one.
-%
-%A more involved example is
-%$$\eclookup{G'_2}{G'_1}{H'_1}{H'_2} \rightarrow \ecfkey{G'_1}{H'_1}$$
-%which states
-%that a look-up constraint implies a foreign-key constraint.
-In the look-up constraint $\eclookup{G'_2}{G'_1}{H'_1}{H'_2}$ the values in the vector $G'_2$ are determined by looking up the corresponding value in vector $G'_1$ in vector $H'_1$ and returning the corresponding value in vector $H'_2$.
-In Figure~\ref{fig:main_example} we have $\eclookup{\range{T_{4}}{\rangeall}{3}}{\range{T_{4}}{\rangeall}{2}}{\range{T_{1}}{\rangeall}{1}}{\range{T_{1}}{\rangeall}{2}}$.
-This implies a foreign-key constraint $\ecfkey{G'_1}{H'_1}$ where each of the values in vector $G'_1$ must appear in $H'_1$ for the look-up to work.
-
-%\tias{I am not defining a new problem for 'most specific canonical tabular constraint learning' because we loosely showed how implications and equivalences would be expressed, but not how to express canonicity and we did not claim that most specific is always best.}
-%\tias{Note also that our method finds ALL constraints; it uses specificity to faster traverse the search space, but the resulting set of constraints found would be the same, whether or not using the specificity (see 4.2.1).}
-
-%Let us introduce the version of tabular constraint learning by modifying Definition \ref{def:problem_statement} that takes into account the specificity relation between constraint templates.
-%
-%\begin{definition} \textit{Condensed Tabular Constraint Learning}\label{def:condensed_problem_statement}\\
-%  Given a set of instantiated tables~${\cal T}$ (all cells have values), a set of {\template}s~${\cal S}$, a set of groups~${\cal G}$ over the tables~${\cal T}$, and the set of all specifity rules~\dependencies: find the set of all constraints $s(G_1', ..., G_n')$ that are satisfied  in ${\cal T}$, for which all $G_i' \subseteq G_i \in {\cal G}$, the (sub)groups $G_1', ...
-, G_n'$ satisfy the signature of the constraint template~$s \in {\cal C}$ and there is no $s'$ in ${\cal C}$ such that $s'(G_h',\dots,G_k')$ holds (for $1 \leq h \leq k \leq n$) and  $s'$ is more specific than $s$ in \dependencies.
-%\end{definition}
-
-% \begin{minipage}[c]{14em}
-%   \vspace{5pt}
-%   \begin{tabular}{ll}
-%     \multicolumn{2}{l}{{\textbf{Condensed Tabular Constraint Learning Problem}}}\\
-%     \vspace{-4pt}
-%     &\\
-%     \textbf{Given:}& the set of all groups $\groups$ and of \template's $\constraints$,\\
-%     & a constraint specificity DAG \dependencies \\
-%     \textbf{Find:}& all the most specific wrt \dependencies constraints $C$ over \groups\\
-%     & for each $t$ in \constraints \\
-%   \end{tabular}
-%   \vspace{6pt}
-% \end{minipage}
-
-%This version explicitly takes into account the specificity of the constraints and invalidates some of the solutions that are entailed by the tabular other solutions.
-
-\newcommand{\tcl}{Tabular Constraint Learning}
-%\newcommand{\ctl}{Condensed Tabular Constraint Learning}
-\section{Approach to Tabular Constraint Learning}\label{sec:approach}
-The aim of our method is to detect constraints between rows and columns of tabular data.
-Recall that a valid argument assignment for a constraint is an assignment of a block to each of the arguments of the constraint, such that the signature and definition of the constraint template is satisfied.
-
-Our proposed methodology contains the following steps:
-\begin{enumerate}
-\item Extract headerless tables from tabular data
-\item Partition the tables into contiguous, type-consistent \textit{superblocks}
-\item Generate for each constraint template all valid argument assignments in two steps:
-\begin{enumerate}
-\item For each constraint template~$s$, generate all assignments $(\sbl{1}, \ldots ,\sbl{n})$ where each~$\sbl{i}$ is a superblock and the~$\sbl{i}$ are \textit{compatible} (defined below) with the signature of~$s$.
-\item For each generated assignment $(\sbl{1}, \ldots ,\sbl{n})$ in 3a), find all valid constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that for all $i$ holds $\ssbl{i} \sqsubseteq \sbl{i}$ and the signature and definition of~$s$ are satisfied.
-\end{enumerate}
-\end{enumerate}
-
-The core of our method is step 3.
-In principle, one could use a generate-and-test approach by generating all possible blocks from the superblock and testing each combination of blocks for each of the arguments of the constraints.
-However, each superblock of size $m$ has $m*(m+1)/2$ contiguous subblocks, meaning that a constraint with $n$ arguments would have to check $O(n^{m^2})$ combinations of blocks.
-
-Instead, we divide this step into two parts: in step 3a, we will not reason over individual (sub)blocks and their data, but rather over the properties of the superblocks.
-% and what combination of superblocks may lead to a valid argument assignment.\\
-Consider table~$T_1$ in Figure~\ref{fig:main_example} and the $\textit{SUM}_{row}$ constraint template.
-Instead of considering all possible subblocks of $T_1$, we first reason over the properties of the superblocks.
-For example, $\sbl{2}$ and $\sbl{4}$ are not numeric so they can be immediately discarded.
-Superblocks~$\sbl{1}, \sbl{3}$ and $\sbl{5}$ are valid candidates for~$\sbl{r}$, however, since they all have length~4, $\mathbf{\sbl{x}}$ needs at least 4~columns which is only satisfied by $\sbl{3}$.
-So the assignments generated in step 3a would be $(\sbl{1}, \sbl{3}), (\sbl{3}, \sbl{3}), (\sbl{5}, \sbl{3})$.
-
-In step 3b, we start from an assignment $(\sbl{1}, \ldots, \sbl{n})$ with superblocks $\sbl{i}$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
-As we only have to consider the blocks $\ssbl{i} \sqsubseteq \sbl{i}$ contained in each superblock, this is typically much easier.
-For the above example and $(\sbl{5}, \sbl{3})$, each of the vectors of $\sbl{5}$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks of $\sbl{3}$ for the right-hand side and verify the signature (e.g.
-at least size $4$) and test whether for each row the definition is satisfied.
-%In practice, this can be optimized further.
-
-We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step~3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step~3b, Section~\ref{sec:algo:subgr}).
-In Section~\ref{sec:opts} we describe a number of optimizations designed to improve the algorithm.
-
-\subsection{Table extraction}
-\label{sec:table_extraction}
-Many spreadsheets contain headers that provide hints at where the table(s) in the sheet are and what the meaning of the rows and columns is.
-However, detecting tables and headers is a complex problem on its own~\cite{header}.
-Furthermore, headers may be missing and their use often requires incorporating language-specific vocabulary, e.g.
-English.
-
-% \begin{figure}[t]
-%   \centering
-%   \includegraphics[width=1\linewidth]{figures/tabletool.png}
-%   \caption{This simple table selection tool allows users to select rectangular ranges of cells corresponding to (headerless) tables.
-%   An explicit orientation can also be provided.}
-%   \label{fig:visual_tool}
-% \end{figure}
-
-Instead, we assume tables are specified by means of their coordinates within a spreadsheet and optionally a fixed orientation of each table.
-The orientation can be used to indicate that a table should be interpreted as having only rows or only columns.
-
-We developed two simple prototypes to help with the specification of the tables:
-%This information can be provided manually or generated automatically.
-
-\paragraph{Automatic detection}
-%As mentioned above, automatic detection of headers and tables is a complex task.
-%Therefore, we define a restricted setting in which the task becomes easy enough to be automated.
-Under the following two assumptions, the table detection task becomes easy enough to be automated; 1) tables are rectangular blocks of data not containing any empty cells; and 2) tables are separated by at least one empty cell from other tables.
-The sheet is then processed row by row, splitting each row into ranges of non-empty cells and appending them to adjacent ranges in the previous row.
-Headers can be detected, for example, by checking if the first row or column contains only textual values.
-If a header row (column) is detected, it is removed from the table and the orientation of the table is fixed to columns (rows), otherwise, we assume there is no header and the orientation is not fixed.
-
-\paragraph{Visual}
-The above assumptions do not hold for many tables.
-However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/SergeyParamonov/TaCLe}\label{github-link}}).
-Users select tables excluding the header and optionally specify an orientation.
-The tool then generates the specification of the tables automatically.
-
-\subsection{Block detection} \label{sec:make_groups}
-%A spreadsheet may contain multiple tables, and a table in turn will contain multiple (type-consistent) blocks.
-The goal of the block detection step is to partition tables into maximally type-consistent (all-numeric or all-textual) blocks.
-%We generate blocks automatically, partitioning tables into maximal type-consistent blocks.
-First, we preprocess the spreadsheet data so that currency values and percentual values are transformed into their corresponding numeric (i.e.
-integer or floating point) representation (e.g.
-$2.75 \$$ as $2.75$ or $85\%$ as~$0.85$).
-Then, each table is partitioned into maximal type-consistent blocks.
-
-To find row-blocks, each row is treated as a vector and must be type-consistent; similarly for columns and column-blocks.
-Then, adjacent vectors that are of the same type are merged to obtain the maximally type-consistent superblocks.
-
-\subsection{Algorithm}
-\label{sec:algo}
-Our method assumes that constraint templates and maximal blocks are given and solves the \tcl~problem by checking for each template $s$: what combination of maximal blocks can satisfy the signature (\textit{superblock assignment}) and which specific \textit{subblock assignments} satisfy both the signature and the definition.
-The pseudo-code of this approach is shown in Algorithm~\ref{algo:tcl}.
-%For every constraint template~$s$, the algorithm first generates all combinations of superblocks (i.e. superblock assignments) that are consistent with the signature.
-%In a second step it looks at every superblock assignment and finds subassignments that satisfy the signature and definition of~$s$.
-
-This separation of checking the \textit{properties} of superblock assignments from checking the \textit{actual data} in the subblock assignment controls the exponential blow-up of combinations to test.
-Furthermore, we use constraint satisfaction technology in the first step to efficiently enumerate superblock assignments that are compatible with the signature.
-%These are the two key properties of our approach.
-
-\newcommand{\temps}{\ensuremath{S}}
-\begin{algorithm}[t]
-  \begin{algorithmic}[1]
-    \footnotesize
-    \State \textbf{Input:} $\temps$ -- constraint templates, $\blocks$ -- maximal blocks
-    \State \textbf{Output:} $C$ -- learned constraints
-
-    \Procedure{LearnConstraints}{$\blocks$, $\temps$}
-      \State $C \gets \emptyset$ %\Comment{The set of constraints}
-      \ForAll{$s~\mathbf{in}~\temps$} \label{algo:tcl:for}
-        %\State $n \gets$ number of arguments of template $s$
-        \State $A \gets \generategroups(s, \blocks)$  \label{algo:tcl:super}
-        \ForAll{$(\sbl{1}, \dots, \sbl{n}) \in A$}
-        \State{$A' \gets \findassignment(s, (\sbl{1},\dots,\sbl{n}))$}
-        \ForAll{$(\sbl{1}',\dots,\sbl{n}') \in A'$}
-            \State $C \gets C \cup \{ c_s(\sbl{1}', \dots, \sbl{n}') \}$
-          \EndFor
-        \EndFor
-      \EndFor
-      \State \Return $C$
-    \EndProcedure
-\end{algorithmic}
-\caption{Learn tabular constraints}
-\label{algo:tcl}
-\end{algorithm}
-
-\subsubsection{Generating superblock assignments (Step 3a)}
-\label{sec:algo:super}
-\
-% Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.
-
-Given a constraint template~$s$ and the set of all (maximal) superblocks~\blocks, the goal is to find all combinations of superblocks that are compatible with the constraint signature.
-An argument assignment $(\sbl{1}, ...
-\sbl{n})$ is \textit{compatible} with the signature of template $s$ if for each block $\sbl{i}$ there exists at least one subblock $\sbl{i}' \sqsubseteq \sbl{i}$ that satisfies the signature.
-
-The choice of one argument can influence the possible candidate blocks for the other arguments, for example, if they must have the same length.
-Instead of writing specialized code to generate and test the superblock assignments, we make use of the built-in reasoning mechanisms of constraint satisfaction solvers.
-
-A Constraint Satisfaction Problem (CSP) is a triple $(V,D,C)$ where $V$ is a set of \textit{variables}, $D$ the domain of possible values each variable can take and $C$ the set of constraints over $V$ that must be satisfied.
-In our case, we define one variable $V_i$ for each argument of a constraint template.
-Each variable can be assigned each of the maximal blocks in \blocks, so the domain of the variables consists of $|\{\blocks\}|$ block identifiers.
-
-%Finding candidate groups for a \template~$t$ can be seen as a Constraint Satisfaction Problem (CSP) dependent on the \CSignature of~$t$.
-%For every argument in~$t$ we consider all groups~\groups (or all candidates from a more general constraint) and prune these candidate groups using their types and other requirements imposed by the templates \CSignature, such that a group~$G \in \groups$ is selected if there is at least one subgroup~$G' \subseteq G$ that satisfies the \CSignature.
-
-To reason over the blocks, we add to the constraint program a set of background facts that contain the properties of each block, namely its type, table, orientation, size, length and number of rows and columns.
-
-\begin{table}[t]
-  \caption{Translation of signature requirements to superblock constraints. The following need not be relaxed: $\plength(B) = x$, $\por(B) = x$, $\ptable(B) = x$ and $\psize(B) \geq x$.}
-  \label{tbl:translation}
-  \begin{tabularx}{\linewidth}{lX}
-    \textbf{Requirement} & \textbf{Superblock constraint} \\ \hline \hline
-    $\ptype(B) = x$ & $\mathit{basetype}(\ptype(B)) = \mathit{basetype}(x)$ \\ \hline
-    $\psize(B) = x$ & $\psize(B) \geq x$ \\ \hline
-    $\pcols(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \pcols(B) \geq x$ \\ 
-    & $\mathit{if}~\por(B) = \mathit{row}: \pcols(B) = x$ \\ \hline
-    $\prows(B) = x$ & $\mathit{if}~\por(B) = \mathit{column}: \prows(B) = x$ \\ 
-    & $\mathit{if}~\por(B) = \mathit{row}: \prows(B) \geq x$
-  \end{tabularx}
-\end{table}
-
-The actual constraints must enforce that a superblock assignment is \textit{compatible} with the requirements defined in the signature.
-
-Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints works:
-Requirements on the lengths, orientations and tables of blocks can be directly enforced since they are invariant under block containment~($\sqsubseteq$).
-Typing requirements need to be relaxed to check only for base types (numeric and textual).
-Minimum sizes can be directly enforced, but exact size requirements are relaxed to minimum sizes, since blocks with too many vectors contain subblocks of the required (smaller) size.
-Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
-Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
-
-The $\generategroups(\textit{s,\blocks})$ method will use these conversion rules to construct a CSP and query a solver for all solutions.
-These solutions correspond to the valid superblock assignments for constraint template~$s$.
-
-\begin{example}
-%Br and Bx are numeric; rows(Bx) ⇤ 2; and columns(Bx ) = length(Br )
-
-  Consider the constraint template $\ecsumr{\sbl{r}}{\bsbl{x}}$, then the generated CSP will contain two variables $V_r, V_x$ corresponding to arguments~$\sbl{r}$ and~$\bsbl{x}$.
-  Given maximal blocks~\blocks, the domain of these variables are $D(V_r) = D(V_x) = \{ 1, \dots, |\blocks| \}$.
-  Finally, the signature of $\ecsumr{\sbl{r}}{\bsbl{x}}$ (see Table~\ref{table:constraints}) will be translated into constraints:
-  \begin{align*}
-    & \numeric(V_r) \land \numeric(V_x) \land \pcols(V_x) \geq 2 \\
-    & (\por(V_x) = \mathit{column}) \Rightarrow (\prows(V_x) \geq \plength(V_r)) \\
-    & (\por(V_x) = \mathit{row}) \Rightarrow (\prows(V_x) = \plength(V_r))
-  \end{align*}
-  %Every solution is a tuple of values for~$V_r$ and~$V_x$ corresponding directly with a superblock assignment over~\blocks.
-\end{example}
-
-\subsubsection{Generating subblock assignments (Step 3b)}
-\label{sec:algo:subgr}
-Given a superblock assignment $(\sbl{1}, \dots, \sbl{n})$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e.
-assignments of subblocks $(\ssbl{1}, \dots, \ssbl{n})$ (for all $i$, $\ssbl{i} \sqsubseteq \sbl{i}$) that satisfy both the signature and the definition of template~$s$.
-
-%While the previous step reasoned about the \textit{properties} of blocks, in this step we will also reason about the actual \textit{content} of the blocks, that is, the data.
-\begin{example}
-  Consider the sum-over-rows constraint template $\ecsumr{\sbl{r}}{\bsbl{x}}$.
-  An example superblock assignment from Figure~\ref{fig:main_example} is $(\sbl{r}, \bsbl{x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
-  Note that this assignment does not satisfy the signature yet because $\sbl{r}$ contains more than one vector.
-  This step aims to generate subassignments $(\sbl{r}' \sqsubseteq \sbl{r}, \mathbf{\sbl{x}}' \sqsubseteq \mathbf{\sbl{x}})$ that do satisfy the signature and test whether they satisfy the definition:
-  $\forall i, \sbl{r}[i] = \sum_{j = 1}^{length(\bsbl{x})} \mathit{row}(i, \bsbl{x})[j]$.
-  In Figure~\ref{fig:main_example} there is exactly one such subassignment: $(\range{T_1}{\rangeall}{7}, \range{T_1}{\rangeall}{\rangeto{3}{6}})$.
-\end{example}
-
-In this step, we could also formulate the problem as a CSP.
-%However, the search space is much smaller, e.g.
-$O(nm^2)$ with $n$ the number of arguments of the template and $m$ the size of the largest block.
-However, few CSP solvers support floating point numbers, which are prevalent in spreadsheets.
-Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
-This is inefficient, as already such blocks may not satisfy the signature or the first elements in the data may not satisfy the definition.
-
-On the other hand, a simple generate-and-test approach, as illustrated in Algorithm~\ref{algo:gen-test-subblocks}, will usually suffice.
-%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
-%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
-Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
-For every (disjoint) subassignment the exact signature (e.g. subtypes will not have been checked yet) and definition will be tested and all satisfying subassignments are returned.
-
-\begin{algorithm}[t]
-  \begin{algorithmic}
-    \footnotesize
-    \Procedure{Subassignments}{$s$, $(\sbl{1}, \dots, \sbl{n})$}
-    \State $n \gets$ number of arguments of template $s$
-    \State $A_{sub} \gets \emptyset$
-    \ForAll{$(\sbl{1}', \dots, \sbl{n}') \textbf{ where } \forall i: \sbl{i}' \sqsubseteq \sbl{i}  \land \textit{disjoint}_{j \neq i}(\sbl{i}', \sbl{j}')$}
-      \If{$\textit{Sig}_s(\sbl{1}',\dots,\sbl{n}') \land \textit{Def}_s(\sbl{1}',\dots,\sbl{n}')$}
-        \State $A_{sub} \gets A_{sub} \cup \{ (\sbl{1}', \dots, \sbl{n}') \}$
-      \EndIf
-    \EndFor
-\Return $A_{sub}$
-\EndProcedure
-\end{algorithmic}
-\caption{Generate-and-test for $\findassignment$}
-\label{algo:gen-test-subblocks}
-\end{algorithm}
-
-% \begin{verbatim}
-% VERY PSEUDOcode:
-% something recursive that handles vectors/
-% subgroups differently like this?
-% check(Groups, X, Test):
-%   if |Groups| == 0:
-%     forall i in ?: vector elements...
-%       if not Test(X):
-%         return []
-%     return X
-%   // else
-%   all = []
-%   Group = Groups.pop(0)
-%   if vector(Group):
-%     forall Vec in Group:
-%       all += check(Groups, X + Vec, Test)
-%   else:
-%     forall Subgroup in Group:
-%       all += check(Groups, X + Subgroup, Test)
-%   return all
-% \end{verbatim}
-
- %$G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-\samuel{\tias{How are they stopped in the code?
-should be part of the pseud-code because not part of $Test(X)$?} will make it part of pseudo-code}
-
-% \paragraph{Example 2, not used (should be replaced by pseudo-code above?)}
-% Sum Product
-
-% Implementation: finds subgroups based on assignments\\
-% Method $Test(G_R, G_{O_1}, G_{O_2})$\\
-% $If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
-% $If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
-% \\\\
-% Method $FindSubgroups(Assignments)$\\
-% $Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g.
-vectors)?}
-
-\subsection{Optimizations} \label{sec:opts}
-This section discusses various design decisions and optimizations aimed at improving the speed, the extensibility of our system and the elimination of redundant constraints.
-
-\subsubsection{Template dependencies}
-As discussed in Section~\ref{sec:form:dependencies}, some constraint templates depend on others by including templates they depend on in their signature (see Table~\ref{table:constraints}).
-
-In Inductive Logic Programming, one often exploits implications between constraints to structure the search space~\cite{luc_book}.
-Our approach uses the dependencies between templates to define a dependency graph.
-We assume that signatures do not contain equivalences or loops, and hence the resulting graph is a directed acyclic graph (DAG).
-Figure~\ref{fig:learning_order} shows the dependency graph extracted from the signatures in Table~\ref{table:constraints}. 
-Constraint templates that have no dependencies are omitted.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=0.50\linewidth]{figures/constraint_dependency2.pdf}
-  \caption{Dependency graph of Table~\ref{table:constraints}; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (its signature includes $s_1$).
-  }
-  \label{fig:learning_order}
-\end{figure}
-
-Using the dependency graph~(\dependencies) we can reorder the templates such that a template occurs after any template it depends on.
-Concretely, in Algorithm~\ref{algo:tcl} line \ref{algo:tcl:for} the input templates~$\temps$ are handled following an ordering that agrees with the partial ordering imposed by~\dependencies (e.g.
-\textit{ALLDIFFERENT} before \textit{FOREIGNKEY}).
-When using this ordering, templates later in the ordering can use the learned constraints of the templates they depend on.
-To enable this reuse of constraints, the set of learned constraints $C$ is added to the argument of \generategroups on line \ref{algo:tcl:super}.
-
-The previously learned constraints are used to speed up the \generategroups step.
-For constraint templates where some arguments are also part of a constraint that it depends on, it is not needed to search for superblocks and subblocks from scratch.
-Instead, one can start from all and only the actual argument assignments of the base constraint, and only search matching superblocks for the remaining arguments.
-
-\begin{example}
-  Consider $\ecfkey{\sbl{fk}}{\sbl{pk}}$, which states that every value in $\sbl{fk}$ also exists in $\sbl{pk}$; its signature includes $\ecalldiff{\sbl{pk}}$.
-There are 18 \textit{ALLDIFFERENT} constraints to be found in Figure~\ref{fig:main_example}, hence, the \generategroups only needs to check which superblocks for $\sbl{fk}$ are compatible (different table, same type) with these 18 assignments to $\sbl{pk}$.
-\end{example}
-
-Instead of generating one CSP to find all assignments, a CSP is generated for every known assignment of the depending constraint (and variables involved), which then searches for all assignments completing this partial assignment.
-The procedure to generate CSPs remains the same otherwise.
-
-%\subsubsection{Constraint solvers}
-%\samuel{Tias, Sergey, have a look at his section, I'm torn about rewriting it, feeling it should either be reduced or expanded, especially the second part.}
-%Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
-%Constraint solvers can be categorized as being stand-alone solvers, meaning that they have a text-based modeling language in which the CSP must be formulated, and \textit{embedded}, meaning that they have a programmatic API to formulate the CSP.
-%
-%\sname supports the ASP \cite{whaisasp} formalism, using the stand-alone Clingo 4.5.4 \cite{clingo} solver, the MiniZinc language \cite{minizinc} with the stand-alone Gecode 4.4.0 solver (which supports floats) and uses an embedded Python-based CP solver \cite{python_constraint}.
-%The constraint signature that must be satisfied to generate the groups can be specified in each of the tree formalisms, to ease prototyping and extendability towards new constraints.
-The embedded solver is fastest as it does not have the overhead of file-writing and calling an external program, and is used in the experiments below.
-
-\subsubsection{Redundancy}
-We consider two types of redundancy that we aim to eliminate during search.
-As noted in Section~\ref{sec:form:dependencies}, for some constraint templates there are \textit{symmetries} over the arguments that lead to trivially equivalent solutions, for example, $\ecprod{\sbl{R}'}{\sbl{1}'}{\sbl{2}'} \Leftrightarrow \ecprod{\sbl{R}'}{\sbl{2}'}{\sbl{1}'}$.
-Such duplicates are avoided by defining a \textit{canonical} form for these constraints.
-In practice, we define an arbitrary but fixed block ordering and require those blocks that are interchangeable to adhere to this ordering.
-Moreover, among the semantically equivalent constraints product (\ecprod{\sbl{R}'}{\sbl{1}'}{\sbl{2}'}) and division (\eccalc{\sbl{1}'}{\sbl{R}' / \sbl{2}'}) we only added product, as well as only difference (\ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}) and not sum (\eccalc{\sbl{1}}{\sbl{r} + \sbl{2}}).
-However, division and sum could easily be added, as they could be added explicitly in post-processing based on the matching product and difference constraints.
-
-A last type of redundancy we chose to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
-For example, consider the constraint $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{n}}}$ where $\range{T}{\rangeall}{\rangeto{k}{n}}$ consists of only zeros, then $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{j}}}$ will be true for all $k - 1 \leq j \leq n$.
-In \sname we have, therefore, chosen to only consider \textit{maximal} subblocks.
-
-In some cases a maximal subblock might falsely include irrelevant columns.
-Consider superblock $\bsbl{x} = \range{T}{\rangeall}{\rangeto{1}{3}} = [[200, 300], [200, 150], [1, 2]]$ and $\sbl{r} = [200, 300]$, then \sname will find the constraint $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$.
-Should the target constraint be $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$, then $\range{T}{\rangeall}{3}$ must be split off into a separate block.
-
-Alternatively, we can output all valid assignments instead of only maximal subblocks, e.g.
-$\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{1})}$, $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$ and $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$, however, this might produce many redundant constraints including constraints that accidentally cover too little data.
-
-\subsubsection{Constraints}
-The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
-We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
-We also added three structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
-
-For \textit{FOREIGNKEY}, \textit{LOOKUP}, aggregates and \textit{PROJECT} we use custom, optimized methods to find subassignments.
-Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
-While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
-
-%Their dependencies are reflected in their \CSignature and shown in Figure~\ref{fig:learning_order}.
-%These constraints have been selected based on the most-used Excel functions as well as their occurrence in various example spreadsheets we observed.
-
-%Note that \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?}
-
-% Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}
-
-% \sergey{under question now, might worth deleting} \\
-% \sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}
-
-% \begin{equation}
-%   \forall i{:}~ v_r[i] \leq v[i] < v_r[i] + 0.5
-%   \label{eq:floor_extension}
-% \end{equation}
-
-% \subsection{Workflow}
-% \begin{algorithm}[thb]
-%   \begin{algorithmic}
-%     \footnotesize
-%     \State \textbf{Input:} $D$ -- dataset, \constraints -- constraints, \dependencies -- dependencies \\(optional: tables $T$, groups $G$)
-%     \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
-%     \If{$T$ is \textbf{not} provided}
-%       \State $T \gets \extracttables(D)$
-%     \EndIf
-%     \If{$G$ is \textbf{not} provided}
-%       \State $G \gets \extractgroups(D, T)$
-%     \EndIf
-%     \State $S \gets \learnconstraints(G,\constraints,\dependencies)$
-%     \State \Return $S$
-% \end{algorithmic}
-% \caption{Workflow}
-% \label{algo:workflow}
-% \end{algorithm}
-
-% \textbf{Approach}
-% \begin{itemize}
-%   \item Notation
-%   \item Algorithm (select constraints, find assignments, find solutions)
-% \end{itemize}
-
-% \samuel{Move next to subgroup satisfaction search}
-% \section{Declarative modeling}
-% \sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
-
-\newcommand{\runtotal}{16.12}
-\newcommand{\runtotalstd}{0.62}
-
-\newcommand{\runfile}{0.50}
-\newcommand{\runfilestd}{0.02}
-
-\newcommand{\benchsize}{??}
-
-\section{Evaluation}\label{sec:evaluation}
-In this section we experimentally validate our approach.
-We studied three main questions concerning the recall (\textit{Q1}), precision (\textit{Q2}) and speed (\textit{Q3}) of our algorithm.
-
-First, our implementation is illustrated using a case study on the spreadsheet corresponding to the example introduced in Figure~\ref{fig:main_example}.
-In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 34 spreadsheets that we assembled from various sources.
-
-\subsection{Method}
-\label{sec:evalualtion:method}
-In order to obtain a benchmark of spreadsheets we looked at three sources:
-1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports (available at the same repository\footref{github-link}).
-All spreadsheets were converted to CSV and the table definitions are fixed.
-Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
-
-For every spreadsheet we specified a ground-truth of constraints, i.e. the \textbf{intended} constraints that are expected to be (re-)discovered.
-These were determined using the formulas (if present), context and headers in the original spreadsheets.
-We encoded all intended constraints using our representation (in canonical form).
-Not all intended constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
-Constraints that are currently outside of the scope of this system are, for example, nested mathematical or logical formulas.
-
-We focus on \textit{functional} constraints that could be used as formulas in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
-Moreover, we included detected equalities as intended constraints (in 5~spreadsheets).
-All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
-
-\subsection{Case study}
-Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
-\sname takes a few seconds to find the constraints in Figure~\ref{fig:sol_example} (excluding structural constraints).
-These include 5 spurious $\textit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
-Moreover, there is 1 \textit{LOOKUP} constraint that was not intended in the original spreadsheet (looking up ID based on Salesperson) and which is symmetric to the intended one (Salesperson based on ID).
-
-For this example our primary goal of rediscovering all intended constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{all intended}}$) is~$1.0$.
-The implementation also returns 6 additional constraints, therefore, our precision (i.e.
-$\frac{\text{intended discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is mostly due to the spurious RANK constraints.
-We believe this is due to the many short vectors which increases the chance for constraints to be true by accident.
-
-\subsection{Benchmark}
-Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
-Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
-
-{\setlength{\tabcolsep}{0.33em}
-\begin{table}[]
-  \centering
-  \caption{Summary of properties and results per spreadsheet category}
-  \label{tbl:category_overview}
-  \begin{tabularx}{\linewidth}{lccccccc}
- & \multicolumn{2}{c}{\textbf{Exercises} (9)} & \multicolumn{2}{c}{\textbf{Tutorials} (21)} & \multicolumn{2}{c}{\textbf{Data} (4)} \\
- & Overall & Sheet avg & Overall & Sheet avg & Overall & Sheet avg \\ \hline
-    \textbf{Tables} & 19 & 2.11 & 48 & 2.29 & 4 & 1 \\ \hline
-    \textbf{Cells} & 1231 & 137 & 1889 & 90 & 2320 & 580 \\ \hline
-    \textbf{\begin{tabular}[c]{@{}l@{}}Intended\\[-4.5pt] Constraints\end{tabular}} & 34 & 3.78 & 52 & 2.48 & 6 & 1.50 \\ \hline \hline
-\textbf{Recall} & 0.85 & 0.83 & 0.88 & 0.87 & 1.00 & 1.00 \\ \hline
-\textbf{\begin{tabular}[c]{@{}l@{}}Recall\\[-4.5pt] Supported\end{tabular}} & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ \hline
-\textbf{Precision} & 0.97 & 0.98 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
-\textbf{Speed (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
-\end{tabularx}
-\end{table}}
-
-\subsubsection*{Q1. How many intended constraints are found by \sname?}
-The overall recall achieved by \sname is~$0.88~(81/92)$. The overall and average recall per category is also high, as shown in Table~\ref{tbl:category_overview}.
-The intended constraints not discovered correspond to constraints not supported by the system (supported recall is 1).
-
-\subsubsection*{Q2. How precise is \sname?}
-Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
-While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.97$ and $1.00$) than for the tutorials category ($0.70$).
-Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
-A qualitative analysis of these spreadsheets shows that their lower precision is caused by duplicate data or multiple ways to calculate the same vectors.
-For these spreadsheets the precision could be increased by post-processing the constraints to detect implications or heuristically selecting one among multiple ways to compute the same vector.
-
-\subsubsection*{Q3. How fast is \sname?}
-The average runtime across all spreadsheets is $0.13s$.
-Across categories the average running times fluctuate in the range between $0.09s$ to $0.22s$.
-While most spreadsheets can be analyzed very quickly, running times can go up to around $1s$ for others.
-
-% Most of the execution time in these cases goes towards searching either aggregate constraints or conditional aggregates.
-% The search for aggregates will be slow on spreadsheets containing larger groups of numeric data.
-% For conditional aggregates the number of candidate primary keys (all-different) and numeric vectors determines the running time (e.g. the case study example).
-
-%We explained various optimizations, most notably the exploitation of dependencies.
-%Using the dependencies to find constraints incrementally cuts down large parts of the search trees, especially for constraints that have many arguments.
-%Essentially, one obtains a more specialized search without much effort.
-
-Of the 32 supported constraint templates (including equality), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
-It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates (each $3.5\% - 4\%$).
-
-To illustrate that using dependencies to find constraints incrementally can increase the efficiency, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF.
-This assumption is too strong in practice, not all keys have to be used in the conditions.
-However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data), showing the strong effect of the use of dependencies.
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{figures/comparison.pdf}
-  \caption{Overview of the average properties and performance of spreadsheets within every benchmark category relative to the maximum across all spreadsheets.
-  Green bars indicate standard deviation within the category.}
-  \label{fig:comparison}
-\end{figure}
-
-\subsection{Testing the limits}
-Our algorithm relies on a correct table and group detection step, and we noticed that exported CSV files can make this more difficult in case of rounding errors due to low-precision formatting as well as misformatting of strings so they are interpreted as numbers.
-Automatically generated sheets can also contain many duplicate (equal) rows or columns, which are a source of increased runtime and decreased recall for our algorithm.
-For example, we tested a spreadsheet (``exps'') with 6~tables, 531~cells and many duplicate columns across tables.
-While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when counting constraints with equal columns separately.
-%By using a (heuristic) post-processing step the precision for this example could be increased.
-%The running time can be improved by implementing additional optimizations.
-
-\section{Applications}\label{sec:applications}
-In this section, we illustrate how various motivating applications could be realized using our system.
-
-\subsection{Autocompletion}
-Autocompletion can be seen as using knowledge derived from the current data to predict new values before they are written by the user.
-A snapshot~($D_0$) is used to learn constraints and these are combined with new data~($\Delta D$) that the user has added.
-Using~$\Delta D$ we can keep track of which vectors have changed and which vectors are now \textit{incomplete}.
-\textit{Functional} constraints, such as $\ecsumc{\sbl{r}}{\mathbf{\sbl{x}}}$, are able to compute a result~($\sbl{r}$) based on its \textit{input}~($\mathbf{\sbl{x}}$).
-If all vectors in the input blocks have changed and the result vector is incomplete, the constraint can compute the resulting value(s).
-
-Consider Figure \ref{fig:autocompletion_example} containing two tables.
-Learning on a snapshot that does not contain the row below $T_2$, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
-The new data $\Delta D$ consists of the new value \textit{West} in the row below $T_2$.
-Therefore, vector~$\range{T_2}{:}{1}$ is changed and vectors~$\range{T_2}{:}{2}$, $\range{T_2}{:}{3}$ have become incomplete.
-The \textit{SUMIF} constraints have enough input to fill the incomplete vectors and predict the values in the last row: $\range{T_2}{4}{2} = 1547$ and $\range{T_2}{4}{3} = 428128$.
-
-%In case multiple constraints predict different values for the same cell, the algorithm will have to either prioritize constraints based on their specificity or a heuristic, not make a prediction, or present the choice to the user.
-
-\subsection{Error detection}
-\begin{figure}[t]
-% %\begin{subfigure}{.70\textwidth}
-%   \begin{center}
-%     \includegraphics[width=\width]{figures/fbi_figure_highlighted.png}
-%   \end{center}
-%   \caption{Real world tabular constraint reconstruction: FBI crime statistics}
-%   \label{fig:fbi}
-%\end{subfigure}
-% \hfill
-% \begin{subfigure}{.29\textwidth}
-\begin{center}
-    \includegraphics[width=1.0\linewidth]{figures/Learning.png}
-  \end{center}
-  \caption{Autocompletion works by learning constraint on a snapshot and combining those constraints with new data $\Delta D$.
-  The figure shows how, given a new value \textit{West}, we can suggest values for the other cells in that row.}
-  \label{fig:autocompletion_example}
-%\end{subfigure}
-\end{figure}
-
-We consider two cases of error detection.
-The first setting is the \textit{online} detection of errors and is similar to the autocompletion setting.
-Using snapshots, we search for \textit{conflict} constraints, which are learned on a previous snapshot but do not agree with newly added data~$\Delta D$.
-Such conflicts can indicate (typing) errors in the data just entered. For example, in case 'Wes' instead of 'West' was added in Figure~\ref{fig:autocompletion_example}.
-%Conflict constraints can indicate errors in the input.
-%However, such conflicts can be wrongly caused by spurious constraints.
-%Hence, conflicts should perhaps only be detected if the snapshot data is large enough.
-%An exact specification of what large enough means (i.e., how to measure the confidence in the constraint violation) is a separate application-dependent research question and will be investigated in the future work.
-
-The second setting is \textit{offline} error detection in a static spreadsheet.
-% Consider a use case concerning crime statistics provided by the FBI in Figure~\ref{fig:fbi}.
-% Our system is able to detect the second sum constraint, however, a missing value prevents the first sum constraint from being learned.
-% A possible approach to detect such missing or wrong values is to look at a sample of the table, e.g.
-trying to leave out each row, and compare the constraints learned on this sample with the constraints learned on the whole table.
-  %The sample needs to be large enough such that we have enough confidence that the constraints are correct.
-For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
-Instead of snapshots, one can repeatedly search for conflict constraints between the original spreadsheet and a sheet with a row or column removed.
-Many schemes can be devised for choosing which row(s) or column(s) are removed.
-%This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g.
-leave out each row and consider it as $\Delta D_i$) to detect conflicts.
-
-%Of course, generating large enough samples and rows (columns) to be added can be done in multiple ways.
-%In the limited case of detecting single errors one can leave out every row (column), consider the remaining data to be the sample and
-%and the left out row (column) to be new input.
-
-\section{Related Work}\label{sec:related_work}
-\sname combines ideas from several existing
-approaches and different fields of research.
-
-First, it borrows techniques from logical and relational learning \cite{luc_book}, as it finds
-constraints in multiple relations (or tables). However, unlike typical logical and relational learning
-approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
-Furthermore, it derives a set of simple “atomic” constraints rather than a set of rules that each consist of multiple literals as in clausal discovery \cite{claudien,lallouet}.
-Nevertheless, several ideas from logical and relational learning
-proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
-Also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha}
-although that line of research has focused on more specialized techniques for specific types of constraints. Furthermore, the constraints and formulas used
-in spreadsheets are often different from those in databases.
-
-Second, there exist a number of algorithms in the constraint programming community that induce constraint (programs)
-from one or more examples and questions. Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
-The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
-To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
-Key differences are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and often typical for constraint programming community, e.g.
-aimed at combinatorial optimization and scheduling.
-Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user regarding satisfiability.
-Each of the examples/queries is processed for each constraint, hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
-for Sudoku).
-The focus is again on integer variables and typical combinatorial optimisation constraints.
-\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
-It also connects with the approach of inducing constraints and functions by generalizing given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
-The key difference here that we start not from existing formulas but from the data directly.
-%As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
-As Modelseeker, these systems focus on the types of constraints in combinatorial optimization but unlike ModelSeeker and our approach, they do not assume the variables are organized in tabular form.
-
-Third, our work borrows from the seminal work of Gulwani et al.
-\cite{flashfill} on program synthesis for tabular data, which is incorporated in Microsoft’s Excel.
-In FlashFill, the end-user might work on a table containing the surnames (column A) and names (column B) of politicians.
-The first row might contain the values A1 = ``Blair'' and B1 = ``Tony'', and the user might then enter in cell C1 the value ``T.B.''.
-At that point FlashFill generates a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', and also applies it to the other politicians (rows) in the table.
-While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
-Flashfill has been extended to e.g.
-FlashExtract \cite{flashextract} to produce a set of tables starting from an unstructured document (a txt or HTML file) with a set of examples marked for extraction.
-However, unlike our approach, Flashfill requires the user to identify explicitly the desired inputs/outputs from the function as well as to provide explicitly some positive and sometimes also negative examples.
-In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.
-Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
-
-\sergey{added database dependency discovery; have a look at it} Fourth, our work is motivated by the functional dependency discover in databases \cite{flach_dependency_discovery}.
-Most of the discovery techniques rely on the database schema that does not directly apply in our case.
-Some techniques work directly on the data \cite{tane_dependency_discovery} and, contrary to our work, are focused on pure functional dependencies that are not typical constraints of a spreadsheet that might hold over multiple tables and on a mixture of columns and rows.
-
-Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
-However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify ``hard'' constraints.
-
-\section{Conclusions}\label{sec:conclusions}
-
-% points: 1) can be a base for an actual plugin or a function in excel 2) novel problem and challenge for systems and constraint solvers 3) can be part of the complicated pipeline togehter with the header detection
-Our goal is to automatically identify constraints in a spreadsheet.
-We have presented and evaluated our approach, implemented as the system \sname, that is able to learn many different constraints in tabular data.
-The resulting method has high recall, produces limited number of redundant constraints and is sufficiently efficient for normal and interactive use.
-
-The approach is designed to find constraints that hold over entire columns and rows.
-In future work we plan to extend this to learn arbitrary nested constraints (e.g.
-$\sbl{r} = (\sbl{1} + \sbl{2}) / \sbl{3}$), as well as constraints over only a subset of the vectors.
-This may give rise to more redundant and spurious constraints, which was not problematic up to this point.
-
-Two promising ways to mitigate redundant and spurious constraints are heuristic filtering or post-processing of the constraint set and the integration of our approach in an interactive setting where users can receive and provide feedback.
-The latter is more similar to an active learning setting.
-
-A final direction is that the current approach assumes noise-free data (but see the part on error correction); in inductive logic programming there has been much work on how to handle noisy data, while in constraint satisfaction a popular alternative is to consider \textit{soft} constraints that can be violated (at a cost).
-This could extend the approach to new application domains, beyond traditional spreadsheets.
-
-\bibliographystyle{plain}
-{
-  \bibliography{references}}
-\end{document}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-09-14 Code/src/runtime_rendering.py
Better rendering

@@ -1,40 +1,70 @@
 import matplotlib.pyplot as plt
-
-number_vectors = [("Runtime",
-                   [0.013, 0.028, 0.171, 2.475, 37.709, 637.579])]
-vector_size = [("Runtime",
-                [2.738, 2.727, 2.475, 2.475, 2.373, 1.603, 0.899, 0.375, 0.408, 0.479, 0.589, 0.837, 1.253]),
-               ("Runtime without RANK",
-                [2.735, 2.724, 2.471, 2.47, 2.367, 1.595, 0.883, 0.348, 0.358, 0.373, 0.385, 0.446, 0.5])]
-number_blocks = [("Runtime aggregate constraints",
-                  [1.66, 0.783, 0.291, 0.167, 0.046]),
-                 ("Runtime non-aggregate constraints",
-                  [36.051, 37.729, 37.974, 48.377, 90.721])]
+import numpy
 
 
-def x(length):
-    return [2 ** power for power in range(length)]
-
-
-def plot(data):
-    fig = plt.figure()
-    ax = fig.add_subplot(111)
-    # fig.set_size_inches(8, 2)  TODO Check out
-
+class ScatterData:
     colors = ["black", "green"]
-    plots = []
-    for i in range(len(data)):
-        title, times = data[i]
-        plots.append(ax.scatter(x(len(times)), times, color=colors[i]))
-        ax.plot(x(len(times)), times, color=colors[i])
+    markers = ["o", "v"]
 
-    ax.legend(plots, (title for title, _ in data), loc="lower left")
+    def __init__(self, title, x_data):
+        self.title = title
+        self.x = x_data
+        self.data = []
+        self.limits = None, None
 
-    ax.set_yscale('log')
-    ax.set_xscale('log')
+    def add_data(self, name, data):
+        self.data.append((name, data))
+        return self
 
-    plt.show()
+    @property
+    def size(self):
+        return len(self.data)
 
-plot(number_vectors)
-plot(vector_size)
-plot(number_blocks)
+    def x_lim(self, limits):
+        self.limits = limits, self.limits[1]
+
+    def y_lim(self, limits):
+        self.limits = self.limits[0], limits
+
+    def render(self, ax, lines=True, log_x=True, log_y=True):
+        plots = []
+        for i in range(self.size):
+            title, times = self.data[i]
+            x_data = self.x(times) if callable(self.x) else self.x
[…]

### 2016-09-14 Code/src/speed_tests.py
Plotting + parsing

@@ -1,6 +1,8 @@
 import json
 import random
 import string
+from collections import OrderedDict
+
 import numpy as np
 import time
 
@@ -8,7 +10,9 @@ import workflow
 import os
 import itertools
 
+from core.constraint import Rank, Aggregate
 from core.group import GType
+from runtime_rendering import ScatterData
 
 
 class TableSpec:
@@ -42,6 +46,23 @@ class TableSpec:
         self.block_types.append((size, b_type))
         return self
 
+    @staticmethod
+    def parse(input_string):
+        _, row_string, blocks_string = input_string.split("-")
+        rows = int(row_string[1:])
+        type_string = blocks_string[0]
+        if type_string == "i":
+            b_type = GType.int
+        elif type_string == "s":
+            b_type = GType.string
+        else:
+            raise RuntimeError("Unexpected type: {}".format(type_string))
+        block_count, block_size = (int(s) for s in blocks_string[1:].split("x"))
+        table = TableSpec(rows)
+        for _ in range(block_count):
+            table.add_block(block_size, b_type)
+        return table
+
 
 class SpeedTestId:
     def __init__(self, tables):
@@ -61,6 +82,16 @@ class SpeedTestId:
     @property
     def name(self):
         return "_".join("T-{}".format(table.name) for table in self.tables)
+
+    @staticmethod
+    def parse(input_string):
+        table_parts = input_string.split("_")
+        return SpeedTestId(list(TableSpec.parse(table_part) for table_part in table_parts))
+
+
+def get_constraints():
+    from core.learning import ordered_constraints
+    return ordered_constraints(workflow.get_constraint_list())
 
 
 def generate_random(test_id: SpeedTestId):
@@ -170,7 +201,7 @@ def generate_experiments():
                 table.add_block(block_size, gtype)
             cat_3.append(SpeedTestId([table, TableSpec(rows).add_block(min_table, gtype)]))
 
-    return [cat_3]
+    return [cat_1, cat_2, cat_3]
 
 
 def setup_experiments(experiments):
@@ -183,49 +214,131 @@ def setup_experiments(experiments):
     return paths
 
 
-def print_table(categories, tasks):
-    print()
-    print("Test", "Total", *[c for c in tasks[categories[0][0]][0].constraints], sep="\t")
+def tasks_to_table(categories, tasks):
+    tables = []
     for category in categories:
+        table = OrderedDict()
         for test_id in category:
-            row = [test_id.name]
             results = tasks[test_id]
[…]

### 2016-09-14 Code/src/speed_tests.py
Named categories

@@ -201,7 +201,7 @@ def generate_experiments():
                 table.add_block(block_size, gtype)
             cat_3.append(SpeedTestId([table, TableSpec(rows).add_block(min_table, gtype)]))
 
-    return [cat_1, cat_2, cat_3]
+    return {"cat1": cat_1, "cat2": cat_2, "cat3": cat_3}
 
 
 def setup_experiments(experiments):
@@ -215,8 +215,8 @@ def setup_experiments(experiments):
 
 
 def tasks_to_table(categories, tasks):
-    tables = []
-    for category in categories:
+    tables = OrderedDict()
+    for c_name, category in categories.items():
         table = OrderedDict()
         for test_id in category:
             results = tasks[test_id]
@@ -225,13 +225,13 @@ def tasks_to_table(categories, tasks):
             for c in results[0].constraints:
                 numbers[c] = list(task.time(c) for task in results)
             table[test_id] = numbers
-        tables.append(table)
+        tables[c_name] = table
     return tables
 
 
 def runs_to_table(categories, runs):
-    tables = []
-    for category in categories:
+    tables = OrderedDict()
+    for c_name, category in categories.items():
         table = OrderedDict()
         for test_id in category:
             results = runs[test_id]
@@ -240,7 +240,7 @@ def runs_to_table(categories, runs):
             for c, values in zip(get_constraints(), results[1:]):
                 series[c] = values
             table[test_id] = series
-        tables.append(table)
+        tables[c_name] = table
     return tables
 
 
@@ -260,7 +260,8 @@ def import_log_file(log_file):
 
 
 def print_tables(tables):
-    for table in tables:
+    for c_name, table in tables.items():
+        print(c_name)
         print("Test", "Total", *get_constraints(), sep="\t")
         for test_id in table:
             row = [test_id.name]
@@ -271,45 +272,58 @@ def print_tables(tables):
 
 
 def draw_figures(tables):
-    scatter_1 = ScatterData("Number of vectors", list(test_id.tables[0].cols for test_id in tables[0]))
-    scatter_1.add_data("All constraints", list(np.average(series["total"]) for series in tables[0].values()))
-    scatter_1.x_lim((10**0, 10**4))
-    scatter_1.y_lim((10**-2, 10**2))
+    scatters = []
 
-    scatter_2 = ScatterData("Vector size", list(test_id.tables[0].rows for test_id in tables[1]))
[…]

### 2016-09-14 Paper/ECAI/tabular.tex

@@
 proved to be useful, such as the use of a canonical form, a refinement graph, and pruning for redundancy.
-\sergey{that's where I added a new reference, needs to be checked and might need to be moved in the right place}.
 It also connects with the logical generalization of given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
@@
 The focus is again on integer variables and typical combinatorial optimisation constraints.
-\sergey{added equation discover here, check if that's the right place for it and if it is in fact good enough description} In a similar spirit the field of equation discovery \cite{equation_discovery} focuses, guided by  heuristics based on the structure of physical equations and unit measures, on the discovery of physical formulas over numbers and times series.
+In a similar spirit the field of equation discovery \cite{equation_discovery} focuses, guided by  heuristics based on the structure of physical equations and unit measures, on the discovery of physical formulas over numbers and times series.
 However, they focus specifically on physical data, while we consider numbers in general and different types of spreadsheet constraints.
@@
 
-\sergey{extended database dependency discovery section; have a look at it} Fourth, also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha} although that line of research has focused on more specialized techniques for specific types of constraints.
+Fourth, also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha} although that line of research has focused on more specialized techniques for specific types of constraints.
 Many of of the discovery techniques rely on the database schema that does not directly apply in our case \cite{flach_dependency_discovery}.
@@
 Let $t$ be a \template, $S$ be a mapping from variables to the subgroups of \groups, and $C$ be the constraint $t^S$, then $C$ \textit{holds} iff all constraints in $\text{\CSignature}^S$ and $\text{\CFunction}^S$ of $C$ hold.
-% \sergey{moved from formalization END}
 
@@
 
-% \sergey{under question now, might worth deleting} \\
-% \sergey{example new template: vr = floor(v), float(v), int(vr), len(v)=len(vr), v,vr are single vectors?}
-
 % \begin{equation}
@@
 % \section{Declarative modeling}
-% \sergey{We need to fit ASP, Minizinc and all that here, I mean it is supposed to be an important point after all}
 

### 2016-09-16 Paper/ECAI/tabular.tex
Included plots in runtime analysis

@@
 \subsection{Runtime analysis}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{figures/scatter_plots.pdf}
+  \caption{Loglog plots of the experiments with synthetic (random) data.
+In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plots~$c$ and~$d$ the number of blocks is increased.}
+  \label{fig:runtime_analysis}
+\end{figure}
+
 There are multiple factors that impact the running time of \sname: the number of vectors, the size of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
 Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
+
+In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the size of the vectors; and 3)~the size of the blocks.
+This analysis is supported by a set of experiments we have run on synthetic spreadsheets.
+These spreadsheets contains random integers\footnote{For strings and floats the results are similar but the running times are typically much lower.} and are made up of two tables.
+For one table the number of vectors ($2$) and block size ($2$) is fixed, while for the other table these properties are varied.
+Moreover, the size of vectors (for both tables) can be adapted, while the default is~$8$.
+The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}.
 
@@
 The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
-Synthetic experiments with random integers and strings show that increasing the size of the vectors has a negligible impact on the performance for most templates.
+As shown in Figure~\ref{fig:runtime_analysis}b, the synthetic experiments confirm that increasing the size of the vectors has a negligible impact on the performance for most template.
+Most of the increase in running time can be attributed to the RANK constraint which is less suited for failing fast.
 
@@
 For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
-As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a large arity.
-This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors is increased.
+As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
+This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors is increased (Figure~\ref{fig:runtime_analysis}a)
 However, in our benchmark experiments there will be various factors that limit the rapid increase in running time because the assignment step can disqualify entire blocks that do not match the signature requirements, i.e.
@@
 Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$.
-Synthetic experiments with random integers confirm that splitting a constant number of vectors into smaller blocks improves the performance of the aggregate constraint search.
-
-These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes very small.
-Since these constraints have single vector arguments the block size should not have a strong effect.
+The synthetic experiments confirm that splitting a constant number of vectors into more, smaller blocks improves the performance of the aggregate constraint search (Figure~\ref{fig:runtime_analysis}c).
+
+These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes very small ($1$ or $2$).
+Since these constraints have single vector arguments, the block size should not have a strong effect.
 However, in practice, for very small blocks the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
-Since the implementation of that last phase is more adept at generating these combinations, shifting this workload to the assignment phase slows down the algorithm.
+Since the implementation of that last phase is more adept at generating these combinations and does not need to check for constraints of all these combinations, shifting this workload to the assignment phase slows down the algorithm.
 

### 2016-09-16 Paper/ECAI/tabular.tex
Various edits

@@
 
-\author{} % \institute{----------------------} }
+\author{Samuel Kolb \and Sergey Paramonov \and Thias Guns \and Luc De Raedt} % \institute{KULeuven}
 
@@
 This is the problem we study in this paper.
-We discuss multiple promising applications that can build on this, such as rediscovery constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
+We discuss multiple promising applications that can build on this, such as rediscovery of constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
 Our techniques build on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
@@
 It also connects with the logical generalization of given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
-The key difference here that we start not from existing formulas but from the data directly.
+The key difference here is that we start not from existing formulas but from the data directly.
 
@@
   \vspace{-10pt}
-  \caption{Constraints learned for the example (left), except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
+  \caption{Constraints learned for the example (top), except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
   Constraints marked with $*$ were not present in the original document}
@@
 The \textit{conditional} aggregate variant \textit{SUMIF} only ranges over those cells that satisfy a condition on a related cell in that table, e.g.
-$\range{T_1}{\rangeall}{10}$ and the data in $T_5$ for matching salesperson names.
+for the spreadsheet in Figure~\ref{fig:main_example}: $\range{T_1}{\rangeall}{10}$ and the data in $T_5$ for matching salesperson names.
 \end{example}
@@
 
-\begin{table}[t]
+\begin{table}[tbh]
   \caption{Translation of signature requirements to superblock constraints. The following need not be relaxed: $\plength(B) = x$, $\por(B) = x$, $\ptable(B) = x$ and $\psize(B) \geq x$.}
@@
 \label{sec:algo:subgr}
+
 Given a superblock assignment $(\sbl{1}, \dots, \sbl{n})$ from the previous step, the goal is to discover valid \textit{subassignments}, i.e.
@@
 
-On the other hand, a simple generate-and-test approach, as illustrated in Algorithm~\ref{algo:gen-test-subblocks}, will usually suffice.
-%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
-%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
-Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
-For every (disjoint) subassignment the exact signature (e.g. subtypes will not have been checked yet) and definition will be tested and all satisfying subassignments are returned.
-
-\begin{algorithm}[t]
+\begin{algorithm}[tbh]
   \begin{algorithmic}
@@
 \end{algorithm}
+
+On the other hand, a simple generate-and-test approach, as illustrated in Algorithm~\ref{algo:gen-test-subblocks}, will usually suffice.
+%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
+%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
+Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
+For every (disjoint) subassignment the exact signature (e.g. subtypes will not have been checked yet) and definition will be tested and all satisfying subassignments are returned.
 
@@
 While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
+Based on the assumptions that most subassignments will not satisfy a constraint template, the implementations are geared to fail-fast when possible.
 
@@
 $\frac{\text{intended discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is mostly due to the spurious RANK constraints.
-We believe this is due to the many short vectors which increases the chance for constraints to be true by accident.
+This spreadsheet contains many short vectors which increases the chance for constraints to be true by accident.
 
@@
 
-\samuel{fail-fast?}
 Of the 32 supported constraint templates (including equality), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
@@
 
-There are multiple factors that impact the running time of \sname: the number of vectors, the size of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
+There are multiple factors that impact the running time of \sname: the number of vectors, the length of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
 Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
 
-In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the size of the vectors; and 3)~the size of the blocks.
+In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the length of the vectors; and 3)~the size of the blocks.
 This analysis is supported by a set of experiments we have run on synthetic spreadsheets.
@@
 For one table the number of vectors ($2$) and block size ($2$) is fixed, while for the other table these properties are varied.
-Moreover, the size of vectors (for both tables) can be adapted, while the default is~$8$.
+Moreover, the length of the vectors (for both tables) can be adapted, while the default is~$8$.
 The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}.
@@
 For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
-For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of size~$k$, the worst case complexity becomes $O(n * m * k)$.
+For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of length~$k$, the worst case complexity becomes $O(n * m * k)$.
 However, in practice, most subassignments will not satisfy the constraint template.
-Since our implementation tries to account for this by failing fast, the amortized time is constant in the size of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
+Since our implementation tries to account for this by failing fast, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
 For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small). 
 The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
-As shown in Figure~\ref{fig:runtime_analysis}b, the synthetic experiments confirm that increasing the size of the vectors has a negligible impact on the performance for most template.
+As shown in Figure~\ref{fig:runtime_analysis}b, the synthetic experiments confirm that increasing the length of the vectors has a negligible impact on the performance for most template.
 Most of the increase in running time can be attributed to the RANK constraint which is less suited for failing fast.
@@
 As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
-This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors is increased (Figure~\ref{fig:runtime_analysis}a)
-However, in our benchmark experiments there will be various factors that limit the rapid increase in running time because the assignment step can disqualify entire blocks that do not match the signature requirements, i.e.
-blocks of vectors with different lengths for the product constraint.
+This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors is increased (Figure~\ref{fig:runtime_analysis}a).
+In practice, in our benchmark experiments, the assignment step of our algorithm is able to limit this rapid increase in running time for larger numbers of vectors.
+The superblock assignment phase is able to disqualify entire blocks that do not match the signature requirements, i.e. blocks of vectors with different lengths for the product constraint.
 
@@
 
-These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes very small ($1$ or $2$).
+These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes small.
 Since these constraints have single vector arguments, the block size should not have a strong effect.
-However, in practice, for very small blocks the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
-Since the implementation of that last phase is more adept at generating these combinations and does not need to check for constraints of all these combinations, shifting this workload to the assignment phase slows down the algorithm.
+However, in practice, when there are many small blocks, the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
+Since the implementation of that last phase is more adept at generating these combinations and does not need to check for constraints on all these combinations, shifting this workload to the assignment phase slows down the algorithm.
 

### 2016-09-16 Paper/ECAI/tabular.tex
Minor fixes

@@
 
-\author{Samuel Kolb \and Sergey Paramonov \and Thias Guns \and Luc De Raedt} % \institute{KULeuven}
+\author{Samuel Kolb \and Sergey Paramonov \and Tias Guns \and Luc De Raedt}
+\institute{KULeuven}
 
@@
 All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
-\sergey{Samuel! Add a reference to python-constraint, including the references list at the end}\\
+The constraint solver that is used by the superblock assignment phase is python-constraint\cite{python_constraint}.
 

### 2016-09-16 Code/src/speed_tests.py
Subfigure titles

@@ -276,7 +276,7 @@ def draw_figures(tables):
 
     if "cat1" in tables:
         table = tables["cat1"]
-        scatter_1 = ScatterData("Number of vectors", list(test_id.tables[0].cols for test_id in table))
+        scatter_1 = ScatterData("(a) Number of vectors", list(test_id.tables[0].cols for test_id in table))
         scatter_1.add_data("All constraints", list(np.average(series["total"]) for series in table.values()))
         scatter_1.x_lim((10**0, 10**4))
         scatter_1.y_lim((10**-2, 10**2))
@@ -284,7 +284,7 @@ def draw_figures(tables):
 
     if "cat2" in tables:
         table = tables["cat2"]
-        scatter_2 = ScatterData("Vector size", list(test_id.tables[0].rows for test_id in table))
+        scatter_2 = ScatterData("(b) Vector size", list(test_id.tables[0].rows for test_id in table))
         total_run_times = list(np.average(series["total"]) for series in table.values())
         rank_run_times = list(np.average(series[Rank()]) for series in table.values())
         no_rank_run_times = list(total - rank for total, rank in zip(total_run_times, rank_run_times))
@@ -298,7 +298,7 @@ def draw_figures(tables):
         table = tables["cat3"]
 
         number_of_blocks = list(len(test_id.tables[0].block_types) for test_id in table)
-        scatter_3 = ScatterData("Block size (aggregates)", number_of_blocks)
+        scatter_3 = ScatterData("(c) Number of blocks (aggregates)", number_of_blocks)
         total_times = list(np.average(series["total"]) for series in table.values())
         aggregates = list(sum(np.average(series[c]) for c in Aggregate.instances()) for series in table.values())
         scatter_3.add_data("All constraints", total_times)
@@ -307,7 +307,7 @@ def draw_figures(tables):
         scatter_3.y_lim((10**-2, 10**2))
         scatters.append(scatter_3)
 
-        scatter_4 = ScatterData("Block size (non-aggregates)", number_of_blocks)
+        scatter_4 = ScatterData("(d) Number of blocks (non-aggregates)", number_of_blocks)
         non_aggregates = list(total - aggregate for total, aggregate in zip(total_times, aggregates))
         scatter_4.add_data("All constraints", total_times)
         scatter_4.add_data("Vector constraints", non_aggregates)

### 2016-09-16 Paper/ECAI/tabular.tex
pass over abstract, intro and rel.work

@@
 However, writing, maintaining and identifying good formulas for tabular data and spreadsheets can be time consuming and error-prone.
-In this work, we investigate the automatic discovery of constraints (formulas and relations) in raw tabular data in an unsupervised way.
-Common spreadsheet formulas are represented using predicates and expressions whose arguments must satisfy the inherent properties of the function.
-The challenge is to discover these without user information or labeled examples.
-We propose a two-stage generate and test method where the first stage uses constraint solving techniques to efficiently reduce candidates in a generic way based on the function signature. 
-Our approach to this knowledge discovery task takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
+In this work, we investigate the automatic learning of constraints (formulas and relations) in raw tabular data in an unsupervised way.
+We represent common spreadsheet formulas and relations through predicates and expressions whose arguments must satisfy the inherent properties of the formula/relation.
+The challenge is to automatically infer the set of constraints that is present in the data, without user information or labeled examples.
+We propose a two-stage generate and test method where the first stage uses constraint solving techniques to efficiently reduce the number of candidates in a generic way, based on the predicate signatures.
+Our approach takes inspiration from inductive logic programming, constraint learning and constraint satisfaction.
 We show that we are able to accurately discover constraints in spreadsheets from various sources.
@@
 %This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
-Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across big tables.
+Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across all tables.
 For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
@@
 This is the problem we study in this paper.
-We discuss multiple promising applications that can build on this, such as rediscovery of constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
-Our techniques build on work on program synthesis, in particular, on Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
+There are multiple applications that could build on this, such as rediscovery of constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
 
@@
 But an important difference remains that we want to learn constraints on both columns and rows over integer, floating point as well as textual data.
+Our inspiration comes from work on program synthesis, in particular Flashfill \cite{flashfill}, where the definition of a function (over textual cells only) is learned in spreadsheet data from very few examples.
 
 % Moved up
-The question that we ask in this paper is: is it possible to discover or reconstruct constraints in flat tabular spreadsheet data?
-The main challenge is the number of combinations of rows and columns that need to be tried as input to the constraints.
+The question that we ask in this paper is: is it possible to learn the set of constraints present in a spreadsheet, from just flat tabular spreadsheet data?
+The main challenge is the number of possible constraints and combinations of rows and columns that need to be tried as input to the constraints.
 To answer this challenge we propose a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
@@
 \begin{itemize}
-\item we define the tabular constraint learning problem, where the goal is to learn constraints (functions and relations) that range over entire rows or columns in an unsupervised way;
-\item we propose a 2-stage generate and test method that reasons for each candidate constraint over maximally contiguous blocks in the first stage, and only in the second stage looks at individual rows and columns and their content;
-\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximal blocks compatible with the candidate constraint's argument, using backtracking search and forward checking;
+\item we define the tabular constraint learning problem, where the goal is to constraints that range over entire rows or columns in an unsupervised way;
+\item we propose an effective 2-stage generate and test method where the first stage reasons only over properties of contiguous blocks of rows/columns, and the second stage continues to investigate individual rows and columns and their content;
+\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximally contiguous blocks compatible with the candidate constraint's argument; %, using backtracking search and forward checking;
 \item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall. 
@@
 The key difference here is that we start not from existing formulas but from the data directly.
-
-Second, there exist a number of algorithms in the constraint programming community that induce constraint (programs)
-from one or more examples and questions. Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
+\tias{I disagree, we start from the formula and then match it on the data, rewrite}
+
+Second, there exist a number of algorithms in the constraint programming community that induce constraints from one or more examples and questions.
+Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
 The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
@@
 However, they focus specifically on physical data, while we consider numbers in general and different types of spreadsheet constraints.
+\tias{Last sentence too vague, what is specific about physical data?
+btw, it is supervised, e.g.
+input-output and we are general?
+Also, it is 'one' conjunction of functions, e.g.
+one equation, or not?
+Please rewrite more precisely}
 %As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
@@
 At that point FlashFill generates a program that produces the output ``T.B.'' for the inputs ``Blair'' and ``Tony'', and also applies it to the other politicians (rows) in the table.
-While it might need an extra example to deal with names such as ``Van Rompuy, Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
+While it might need an extra example to deal with names such as ``Van Rompuy'', ``Herman'' and determine whether it should output ``H.V.'' or ``H.V.R.'', FlashFill learns correct programs from very few examples.
 Flashfill has been extended to e.g.
@@
 Fourth, also related to this line of research is the work on deriving constraints in databases such as functional and multi-valued dependencies \cite{savnik, Mannila-Raiha} although that line of research has focused on more specialized techniques for specific types of constraints.
-Many of of the discovery techniques rely on the database schema that does not directly apply in our case \cite{flach_dependency_discovery}.
-Some techniques work directly on the data \cite{tane_dependency_discovery} and focus on classical functional dependencies between attributes, while the constraints and formulas used in spreadsheets are often different from those in databases.
-
-Finally, there are the works on  BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
+Many of of the discovery techniques rely on the database schema~\cite{flach_dependency_discovery}, which is not available in our case.
+Techniques that work directly on the data have been investigated for the discovery of functional dependencies between attributes~\cite{tane_dependency_discovery}.
+The constraints and formulas we wish to learn go beyond that, including arithmetic, conditional arithmetic and \textit{fuzzy} lookups, for example.
+
+Finally, worth mentioning is BayesDB \cite{BayesDB} and Tabular \cite{tabular}, two recent probabilistic modeling languages that have been specifically designed for dealing with relational data in tabular form and that can, as FlashFill, also automatically fill out missing values in a table.
 However, unlike the other mentioned approaches, it is based on an underlying probabilistic graphical model that performs probabilistic inference rather than identify ``hard'' constraints.
@@
     Green background indicates headerless tables, dark borders indicate maximal type-consistent blocks.
-    Most tables only contain type-consistent columns while table~$T_2$ also contains type-consistent rows.
-    However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
-    This example was created by combining several Excel sheets from several exercise sessions based on the book MS Excel 2010~\cite{excel_book}.
+    Most tables only contain type-consistent columns while table~$T_2$ also contains type-consistent rows (though its header indicates that it consists of row data).
+%    However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
+    This example combines several Excel sheets based on exercise sessions in the book ``MS Excel 2010''~\cite{excel_book}.
     % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
@@
   \vspace{-10pt}
-  \caption{Constraints learned for the example (top), except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5).
-  Constraints marked with $*$ were not present in the original document}
+  \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5) constraints.
+  Constraints marked with $*$ were not present in the original sheets.}
   \label{fig:sol_example}
@@
 \section{Formalization}\label{sec:formalization}
-Our goal is to automatically discover constraints between the rows and the columns of tables in a spreadsheet.
+Our goal is to automatically learn the constraints present between rows and/or columns in a spreadsheet.
 This is applicable not just to data from spreadsheets, but to any data in tabular form, hence the name.

### 2016-09-16 Paper/ECAI/tabular.tex
experiments reworked

@@
   \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5) constraints.
-  Constraints marked with $*$ were not present in the original sheets.}
+  Constraints marked with $*$ were not present in the original spreadsheets.}
   \label{fig:sol_example}
@@
 
-\subsubsection{Constraints}
+\subsubsection{Constraints} \label{sec:which_cons}
 The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
@@
 In this section we experimentally validate our approach.
-We studied three main questions concerning the recall (\textit{Q1}), precision (\textit{Q2}) and speed (\textit{Q3}) of our algorithm.
-
-First, our implementation is illustrated using a case study on the spreadsheet corresponding to the example introduced in Figure~\ref{fig:main_example}.
-In order to quantify the results and generalize our findings we also evaluate our algorithm on a benchmark of 34 spreadsheets that we assembled from various sources.
-
-\subsection{Method}
+We first explain the experimental setup and illustrate it on the example in Figure~\ref{fig:main_example}, we then investigate the effectiveness of our method on generated spreadsheets, after which we evaluate our method on spreadsheets from various sources.
+
+\paragraph{Experimental setup}
 \label{sec:evalualtion:method}
+All spreadsheets used in the experiments are in CSV format, and for each we have ran the table extraction tool once (using a combination of the rather primitive automatic detection + visual verification that the tables are meaningful (e.g.
+merge split-up tables because of None values, or remove headers of all-textual tables).
+\\ % split of weird sentence
+Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
+\tias{What does this mean, 'group definitons'??
+Is this 'block detection' (Sec 4.2), if so, how do we explain this process??}
+
+All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints that are expected to be (re-)discovered.
+These were determined by us using the formulas of the original sheet (if present), the headers in the sheet as well as general context if need.
+\tias{last part too vague!!
+Give example of where 'context' was needed/used.
+If nowhere, cut that part of sentence}.
+In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas.
+We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
+All constraints are stored in their canonical form.
+5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors.
+
+Section~\ref{sec:which_cons} discusses the 32 constraints currently supported by \sname \tias{It does, right?
+Should also contain '32' number}.
+Some of the intended constraints are currently outside of the scope of \sname however, more specifically nested mathematical or nested logical formulas.
+We denote by \textbf{supported} constraints the subset of intended constraints that the system can find in theory.
+
+We will use recall and precision to measure how well our tool is performing.
+Recall is the fraction of intended constraints actually discovered by the system: $\frac{\text{intended discovered}}{\text{all intended}}$, while precision is the fraction of constraints found by the system that are indeed intended: $\frac{\text{intended discovered}}{\text{all discovered}}$.
+
+All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
+The constraint solver that is used by the superblock assignment phase (Step 3a) is python-constraint\cite{python_constraint}.
+
+\subsection{Results on the running example}
+For the example presented in Figure~\ref{fig:main_example},
+\sname takes a few seconds to find the constraints listed in Figure~\ref{fig:sol_example}.
+These include 5 spurious $\textit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
+Moreover, there is 1 \textit{LOOKUP} constraint that was not intended in the original spreadsheet (looking up ID based on Salesperson) and which is symmetric to the intended one (Salesperson based on ID).
+
+Hence, for this example we achieve a recall of 100\% (all intended constraints are discovered) and a precision of $12/18 = 67\%$.
+All intended constraints are supported.
+The recall is as good as we hoped, but the precision is rather low.
+This is mostly due to the spurious RANK constraints.
+We note that the tables are quite short, which increases the chance of a constraint like 'RANK' to be true by chance; we expect less spurious constraints on larger datasets.
+
+We now investigate three main questions that influence the quality of the solutions found by a constraint learning method:
+\begin{itemize}
+\item Q1: it may fail to find intended constraints if it takes too much time to find them; what are the factors that influence the runtime of the method most?
+\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
+\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
+\end{itemize}
+We first investigate Q1 on generated spreadsheets, and Q2 and Q3 on a collection of real spreadsheets.
+
+\subsection{Effectiveness on generated data}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{figures/scatter_plots.pdf}
+  \caption{Loglog plots of the experiments with synthetic (random) data.
+In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plots~$c$ and~$d$ the number of blocks is increased.}
+  \label{fig:runtime_analysis}
+\end{figure}
+
+There are multiple factors that can impact the running time of \sname: the number of vectors, the length of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
+Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
+
+In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the length of the vectors; and 3)~the size of the blocks.
+\tias{Why only these and not the other?
+Motivated/explain, or also measure their effect ; )}
+
+Recall that our method uses a 2-phase approach where in the first phase, the superblock assignments are generated (Section ...) and in the second phase, the subblock assignments are generated (Section ...).
+
+\paragraph{Spreadsheet generation}
+We do this analysis on a set of synthetically generated spreadsheets.
+The spreadsheets contains random integers\footnote{For strings and floats the results are similar but the running times are typically much lower.} and are made up of two tables.
+For one table the number of vectors and block size is fixed to two, while for the other table these properties are varied.
+Moreover, the length of the vectors (for both tables) can be adapted, while the default is~$8$.
+%The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}. % no such general pointer, specific pointers only
+
+\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}
+For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
+For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of length~$k$, the worst case complexity becomes $O(n * m * k)$.
+However, in practice, most subassignments will not satisfy the constraint template.
+Since our implementation tries to account for this by failing fast\tias{explain more precisely what you mean with 'failing fast'}, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
+For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).
+
+\paragraph{Effect of length of vectors}
+%The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
+As can be seen in Figure~\ref{fig:runtime_analysis}b\tias{This is the first figure discussed so must be a}, increasing the length of the vectors has a negligible impact on the performance for most template.
+Most of the increase in running time can be attributed to the RANK constraint which is less suited for failing fast \tias{Too vague, 'less suited', be precise}; the runtime for detecting all constraints except RANK is shown in green to highlight this.
+
+\paragraph{Effect of number of vectors}
+On the contrary, the number of vectors in a spreadsheet can have a significant impact on the running time.
+Many constraints take single vector arguments, therefore, the number of subassignments \tias{unclear, this is about phase 2, the number to test, or the actual testing of each?
+explain and relate to Sec 4.3.2 what aspect you are discussing here} to be tested depends on the number of vectors in the spreadsheets and the arity of the constraint.
+For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
+\tias{Better move to 4.3.2?
+or have there too anyway, no new facts in 'experiment' section}
+As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
+This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors \tias{more details about equal sized: is this a strict requirement for this to happen?
+I assume this is worst case and does not happen that much in practice?
+(the next sentences half hint at that, be more explicit)} is increased (Figure~\ref{fig:runtime_analysis}a).
+In practice, in our benchmark experiments, the assignment step of our algorithm is able to limit this rapid increase in running time for larger numbers of vectors \tias{How?
+explain, if you explain in the next sentence then start it with 'This is because' or alike, reads better}.
+The superblock assignment phase is able to disqualify entire blocks that do not match the signature requirements, i.e. blocks of vectors with different lengths for the product constraint.
+
+Aggregate constraints, e.g. sum or max, have an argument that takes subblocks of varying sizes.
+Therefore, the size of the blocks \tias{in what phase? the max. contigiuous? be more precise} in the spreadsheet also influences the runtime of these constraints.
+For a block of size~$m$ there are $\frac{m(m + 1)}{2}$ or $O(m^2)$ subblocks, thus, if a spreadsheet has more but smaller sized blocks, the search for aggregate constraints will be faster.
+Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$.
+The \tias{explain what THE experiment is, and how you set it up} synthetic experiments confirm that splitting a constant number of vectors into more, smaller blocks improves the performance of the aggregate constraint search (Figure~\ref{fig:runtime_analysis}c, green line).
+
+These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes small.
+Since these constraints have single vector arguments, the block size should not have a strong effect.
+However, in practice, when there are many small blocks, the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
+Since the implementation of that last phase is more adept at generating these combinations and does not need to check for constraints on all these combinations, shifting this workload to the assignment \tias{more precise: to superblock?} phase slows down the algorithm.
+\tias{Point to figure to support this (actually, move that line into 4c??); also, if this is true, you are basically saying that the CP-based first phase is worse then the second phase?
+That kind-of demotivates the idea of using CP in first phase...
+What do you really want to share with the reader?}
+
+\subsection{Effectiveness on real spreadsheets}
 In order to obtain a benchmark of spreadsheets we looked at three sources:
 1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports (available at the same repository\footref{github-link}).
-All spreadsheets were converted to CSV and the table definitions are fixed.
-Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
-
-For every spreadsheet we specified a ground-truth of constraints, i.e. the \textbf{intended} constraints that are expected to be (re-)discovered.
-These were determined using the formulas (if present), context and headers in the original spreadsheets.
-We encoded all intended constraints using our representation (in canonical form).
-Not all intended constraints are currently supported by \sname, therefore, we sometimes consider the subset of \textbf{supported} constraints.
-Constraints that are currently outside of the scope of this system are, for example, nested mathematical or logical formulas.
-
-We focus on \textit{functional} constraints that could be used as formulas in spreadsheets, ignoring the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
-Moreover, we included detected equalities as intended constraints (in 5~spreadsheets).
-All experiments were run using Python 3.5.1 on a Macbook Pro, Intel Core i7 2.3~GHz with 16GB RAM.
-The constraint solver that is used by the superblock assignment phase is python-constraint\cite{python_constraint}.
-
-\subsection{Case study}
-Let us illustrate our implementation using the example presented in Figure~\ref{fig:main_example}.
-\sname takes a few seconds to find the constraints in Figure~\ref{fig:sol_example} (excluding structural constraints).
-These include 5 spurious $\textit{RANK}$ constraints, e.g. $\ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5}}$, that are true by accident.
-Moreover, there is 1 \textit{LOOKUP} constraint that was not intended in the original spreadsheet (looking up ID based on Salesperson) and which is symmetric to the intended one (Salesperson based on ID).
-
-For this example our primary goal of rediscovering all intended constraints is achieved, the recall (i.e. $\frac{\text{intended discovered}}{\text{all intended}}$) is~$1.0$.
-The implementation also returns 6 additional constraints, therefore, our precision (i.e.
-$\frac{\text{intended discovered}}{\text{all discovered}}$) is $12/18 = 0.67$, which is mostly due to the spurious RANK constraints.
-This spreadsheet contains many short vectors which increases the chance for constraints to be true by accident.
-
-\subsection{Benchmark}
+
 Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
 Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
+\tias{Weird here because purpose unclear, point to this when discussing some result for which this is important}
 
@@
 \textbf{Precision} & 0.94 & 0.96 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
-\textbf{Speed (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
+\textbf{Runtime (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
 \end{tabular}
@@
 
-\subsubsection*{Q1. How many intended constraints are found by \sname?}
-The overall recall achieved by \sname is~$0.88~(81/92)$. The overall and average recall per category is also high, as shown in Table~\ref{tbl:category_overview}.
-The intended constraints not discovered correspond to constraints not supported by the system (supported recall is 1).
-
-\subsubsection*{Q2. How precise is \sname?}
-Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
-While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).
-Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
-A qualitative analysis of these spreadsheets shows that their lower precision is caused by duplicate data or multiple ways to calculate the same vectors.
-For these spreadsheets the precision could be increased by post-processing the constraints to detect implications or heuristically selecting one among multiple ways to compute the same vector.
-
-\subsubsection*{Q3. How fast is \sname?}
+\paragraph{Q1: Effectiveness}
+Our first questions, Q1, is about the inability to find intended constraints due to excessive runtime.
+As Table~\ref{tbl:category_overview} (last row) shows, the runtime of our method on these spreadsheets is always fast.
 The average runtime across all spreadsheets is $0.13s$.
 Across categories the average running times fluctuate in the range between $0.09s$ to $0.22s$.
-While most spreadsheets can be analyzed very quickly, running times can go up to around $1s$ for others.
+While most spreadsheets can be analyzed very quickly, running times can go up to around $1s$ for others. ... \tias{Explain by relating to previous section of synthetic data}
 
@@
 
-Of the 32 supported constraint templates (including equality), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
+Looking at the runtime per constraint template individually (including equality\tias{what does this mean?}), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
 It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates (each $3.5\% - 4\%$).
 
-To illustrate that using dependencies to find constraints incrementally can increase the efficiency, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF.
-This assumption is too strong in practice, not all keys have to be used in the conditions.
-However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data), showing the strong effect of the use of dependencies.
+Our method uses dependencies to increase efficiency by finding constraints incrementally.
+To illustrate the effect this can have, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF.
+This is a stronger assumption, and a too strong in practice as not all keys have to be used in the conditions.
+However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data).
+This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.
 
@@
 
-\subsection{Testing the limits}
-Our algorithm relies on a correct table and group detection step, and we noticed that exported CSV files can make this more difficult in case of rounding errors due to low-precision formatting as well as misformatting of strings so they are interpreted as numbers.
-Automatically generated sheets can also contain many duplicate (equal) rows or columns, which are a source of increased runtime and decreased recall for our algorithm.
+\paragraph{Q2: Intended and supported constraints}
+%\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
+The overall recall achieved by \sname is~$0.88~(81/92)$. The overall and average recall per category is also high, as shown in Table~\ref{tbl:category_overview}.
+The intended constraints not discovered correspond to constraints not supported by the system (supported recall is 1).
+\tias{Rewrite and go into more detail about all supported found, some intended not support, which ones, what category, just some more details}
+
+\paragraph{Q3: precision}
+%\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
+The final question, Q3, is about the amount and type of non-intended constraints found.
+Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
+While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).
+
+Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
+This is caused by duplicate data or multiple ways to calculate the same vectors.
+\tias{More details or examples; we copy intended constraints for equivalent/duplicate vectors, so not an issue?}
+For these spreadsheets the precision could be increased by post-processing the constraints to detect implications or heuristically selecting one among multiple ways to compute the same vector.
+
+%\subsection{Testing the limits}
+%Our algorithm relies on a correct table and group detection step, and we noticed that exported CSV files can make this more difficult in case of rounding errors due to low-precision formatting as well as misformatting of strings so they are interpreted as numbers.
+%Automatically generated sheets can also contain many duplicate (equal) rows or columns, which are a source of increased runtime and decreased recall for our algorithm.
 For example, we tested a spreadsheet (``exps'') with 6~tables, 531~cells and many duplicate columns across tables.
-While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when counting constraints with equal columns separately.
-%By using a (heuristic) post-processing step the precision for this example could be increased.
-%The running time can be improved by implementing additional optimizations.
-
-\subsection{Runtime analysis}
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{figures/scatter_plots.pdf}
-  \caption{Loglog plots of the experiments with synthetic (random) data.
-In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plots~$c$ and~$d$ the number of blocks is increased.}
-  \label{fig:runtime_analysis}
-\end{figure}
-
-There are multiple factors that impact the running time of \sname: the number of vectors, the length of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
-Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
-
-In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the length of the vectors; and 3)~the size of the blocks.
-This analysis is supported by a set of experiments we have run on synthetic spreadsheets.
-These spreadsheets contains random integers\footnote{For strings and floats the results are similar but the running times are typically much lower.} and are made up of two tables.
-For one table the number of vectors ($2$) and block size ($2$) is fixed, while for the other table these properties are varied.
-Moreover, the length of the vectors (for both tables) can be adapted, while the default is~$8$.
-The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}.
-
-For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
-For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of length~$k$, the worst case complexity becomes $O(n * m * k)$.
-However, in practice, most subassignments will not satisfy the constraint template.
-Since our implementation tries to account for this by failing fast, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
-For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small). 
-The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
-As shown in Figure~\ref{fig:runtime_analysis}b, the synthetic experiments confirm that increasing the length of the vectors has a negligible impact on the performance for most template.
-Most of the increase in running time can be attributed to the RANK constraint which is less suited for failing fast.
-
-On the contrary, the number of vectors in a spreadsheet have a significant impact on the running time.
-Many constraints take single vector arguments, therefore, the number of subassignments to be tested depends on the number of vectors in the spreadsheets and the arity of the constraint.
-For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
-As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
-This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors is increased (Figure~\ref{fig:runtime_analysis}a).
-In practice, in our benchmark experiments, the assignment step of our algorithm is able to limit this rapid increase in running time for larger numbers of vectors.
-The superblock assignment phase is able to disqualify entire blocks that do not match the signature requirements, i.e. blocks of vectors with different lengths for the product constraint.
-
-Aggregate constraints, e.g. sum or max, have an argument that takes subblocks of varying sizes.
-Therefore, the size of the blocks in the spreadsheet also influences the runtime of these constraints.
-For a block of size~$m$ there are $\frac{m(m + 1)}{2}$ or $O(m^2)$ subblocks, thus, if a spreadsheet has more but smaller sized blocks, the search for aggregate constraints will be faster.
-Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$.
-The synthetic experiments confirm that splitting a constant number of vectors into more, smaller blocks improves the performance of the aggregate constraint search (Figure~\ref{fig:runtime_analysis}c).
-
-These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes small.
-Since these constraints have single vector arguments, the block size should not have a strong effect.
-However, in practice, when there are many small blocks, the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
-Since the implementation of that last phase is more adept at generating these combinations and does not need to check for constraints on all these combinations, shifting this workload to the assignment phase slows down the algorithm.
+%While we obtained a recall of~$1.0$, the algorithm ran for $14.30s$ and the overall precision was~$0.52$ and only~$0.11$ when counting constraints with equal columns separately.
+%%By using a (heuristic) post-processing step the precision for this example could be increased.
+%%The running time can be improved by implementing additional optimizations.
 

### 2016-09-20 Paper/ECAI/tabular.tex
rewritten

@@
 It also connects with the logical generalization of given formulas in a spreadsheet \cite{Isakowitz}, where one abstracts present formulas by introducing parameters in the place of particular cell references.
-The key difference here is that we start not from existing formulas but from the data directly.
-\tias{I disagree, we start from the formula and then match it on the data, rewrite}
+The key difference here is that we discover constraints (formulas) in the data, we do not have or use any prior information on the formulas in a spreadsheet.
 
@@
 The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
-To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, … and Modelseeker would then look for both column, row  and even diagonal constraints.
+To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, $\dots$ and Modelseeker would then look for both column, row  and even diagonal constraints.
 Key differences are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and often typical for constraint programming community, e.g.
@@
 In a similar spirit the field of equation discovery \cite{equation_discovery} focuses, guided by  heuristics based on the structure of physical equations and unit measures, on the discovery of physical formulas over numbers and times series.
-However, they focus specifically on physical data, while we consider numbers in general and different types of spreadsheet constraints.
+However, they focus specifically on noisy physical data and formulas, e.g.
+a single variable can take values from a series of measurements to get a regression parameter estimate or they apply specific data transformation such as numerical calculation of time and partial derivatives, i.e.
+the method works with approximations; the equation structures are heavily biased towards parameter estimation, while we focus on discovery of different types of spreadsheet constraints which is an exact constraint satisfaction problem.
+\sergey{rewritten, Samuel, does it sound fine now?}
 \tias{Last sentence too vague, what is specific about physical data?

### 2016-09-23 Paper/ECAI/tabular.tex
shared first author
Rewriting evaluation

@@
 \subsubsection{Constraints} \label{sec:which_cons}
-The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
+The \samuel{count} constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
 We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
@@
 merge split-up tables because of None values, or remove headers of all-textual tables).
+\samuel{Should we not say just with the visual tool?}
  \\ % split of weird sentence
-Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
+Blocks are detected automatically using our block detection algorithm (Section~\ref{sec:make_groups}) unless a spreadsheet contains ambiguous \textit{None} values.
+For those ambiguous cases the blocks are specified manually. \samuel{I have yet to make this more precise}
+
+% Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
 \tias{What does this mean, 'group definitons'??
@@
 
-All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints that are expected to be (re-)discovered.
-These were determined by us using the formulas of the original sheet (if present), the headers in the sheet as well as general context if need.
-\tias{last part too vague!!
-Give example of where 'context' was needed/used.
-If nowhere, cut that part of sentence}.
+All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints, that are expected to be (re-)discovered.
+These were determined by us using either the formulas of the original sheet (if present), or inferring them based on, for example, the header names in the sheet.
 In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas.
@@
 Section~\ref{sec:which_cons} discusses the 32 constraints currently supported by \sname \tias{It does, right?
-Should also contain '32' number}.
+Should also contain '32' number}\sam{I will recount at the end}.
 Some of the intended constraints are currently outside of the scope of \sname however, more specifically nested mathematical or nested logical formulas.
@@
 This is mostly due to the spurious RANK constraints.
-We note that the tables are quite short, which increases the chance of a constraint like 'RANK' to be true by chance; we expect less spurious constraints on larger datasets.
+We note that the tables are quite short, which increases the chance of a constraint like \textit{RANK} to be true by chance; we expect less spurious constraints on larger datasets.
 
@@
 
-There are multiple factors that can impact the running time of \sname: the number of vectors, the length of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
-Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
-
 In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the length of the vectors; and 3)~the size of the blocks.
-\tias{Why only these and not the other?
-Motivated/explain, or also measure their effect ; )}
-
-Recall that our method uses a 2-phase approach where in the first phase, the superblock assignments are generated (Section ...) and in the second phase, the subblock assignments are generated (Section ...).
+This analysis is accomplished by testing our system on generated spreadsheets.
+
+Recall that our method uses a 2-phase approach where in the first phase, the superblock assignments are generated (Section~\ref{sec:algo:super} and in the second phase, the subblock assignments are generated (Section~\ref{sec:algo:subgr}).
 
 \paragraph{Spreadsheet generation}
-We do this analysis on a set of synthetically generated spreadsheets.
-The spreadsheets contains random integers\footnote{For strings and floats the results are similar but the running times are typically much lower.} and are made up of two tables.
+For our analysis, we perform multiple experiments on synthetically generated spreadsheets.
+
+Many constraints require either numeric or discrete data, therefore, the type of cells also has an impact on the running time.
+Since integers fulfill both of those typing restrictions, the generated spreadsheets contain only integer data.
+
+One of the assumptions guiding the implementation is that most of the possible constraints are false.
+To simulate this setting, the generated spreadsheets contain random data.
+There will be few constraints that hold on these spreadsheets, however, random data is likely to produce vectors that fulfill the \textit{ALLDIFFERENT} constraint, which is a requirement for many other constraint templates.
+
+In order to also analyse constraints that require their arguments to be picked from different tables, the spreadsheets are made up of two tables.
 For one table the number of vectors and block size is fixed to two, while for the other table these properties are varied.
-Moreover, the length of the vectors (for both tables) can be adapted, while the default is~$8$.
+The number of vectors and the block size in the experiments are only reported for the adaptable table.
+The length of the vectors can also be adapted and applies to both tables.
+Unless otherwise mentioned, the adaptable table consists of one block containing all its vectors and by default the number of vectors and vector length are both~$8$.
 %The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}. % no such general pointer, specific pointers only
 
-\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}
+\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}\sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
 For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
@@
 
+\paragraph{Effect of number of vectors}
+Figure~\ref{fig:runtime_analysis}a shows the results of our experiments, in which \sname was run on generated spreadsheets containing an increasing number of (equally sized) vectors.
+As shown in the figure, the number of vectors in a spreadsheet can have a significant impact on the running time.
+
+This behavior is to be expected when not many superblock assignments can be discarded in phase~1, for example, when all vectors have the same type and size.
+In this case, for a constraint template that requires $n$~single vector arguments and a spreadsheet with $v$~vectors, up to $\frac{v!}{(v - n)!}$ subassignments are generated throughout the algorithm.
+
+Typically, we expect the number of generated subassignments to be much smaller in real world spreadsheets and, therefore, the running time to be less sensitive to the number of vectors.
+This is because in phase~1 entire blocks can be disqualified that do not match the signature requirements, e.g. blocks of vectors with different lengths for the product constraint.
+% \tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}\samuel{?}
+\samuel{=> benchmark}
+
+% Many constraints take single vector arguments, therefore, the total number of subassignments that are generated in the second phase of the algorithm depends on the number of vectors in the spreadsheet and the arity of the constraint.
+
+% \tias{unclear, this is about phase 2, the number to test, or the actual testing of each?
+explain and relate to Sec 4.3.2 what aspect you are discussing here} to be tested depends on the number of vectors in the spreadsheets and the arity of the constraint.
+
+% For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
+\tias{Better move to 4.3.2?
+or have there too anyway, no new facts in 'experiment' section}
+% As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
+% This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors \tias{more details about equal sized: is this a strict requirement for this to happen?
+I assume this is worst case and does not happen that much in practice?
+(the next sentences half hint at that, be more explicit)} is increased (Figure~\ref{fig:runtime_analysis}a).
+
 \paragraph{Effect of length of vectors}
 %The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
-As can be seen in Figure~\ref{fig:runtime_analysis}b\tias{This is the first figure discussed so must be a}, increasing the length of the vectors has a negligible impact on the performance for most template.
-Most of the increase in running time can be attributed to the RANK constraint which is less suited for failing fast \tias{Too vague, 'less suited', be precise}; the runtime for detecting all constraints except RANK is shown in green to highlight this.
-
-\paragraph{Effect of number of vectors}
-On the contrary, the number of vectors in a spreadsheet can have a significant impact on the running time.
-Many constraints take single vector arguments, therefore, the number of subassignments \tias{unclear, this is about phase 2, the number to test, or the actual testing of each?
-explain and relate to Sec 4.3.2 what aspect you are discussing here} to be tested depends on the number of vectors in the spreadsheets and the arity of the constraint.
-For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
-\tias{Better move to 4.3.2?
-or have there too anyway, no new facts in 'experiment' section}
-As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
-This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors \tias{more details about equal sized: is this a strict requirement for this to happen?
-I assume this is worst case and does not happen that much in practice?
-(the next sentences half hint at that, be more explicit)} is increased (Figure~\ref{fig:runtime_analysis}a).
-In practice, in our benchmark experiments, the assignment step of our algorithm is able to limit this rapid increase in running time for larger numbers of vectors \tias{How?
-explain, if you explain in the next sentence then start it with 'This is because' or alike, reads better}.
-The superblock assignment phase is able to disqualify entire blocks that do not match the signature requirements, i.e. blocks of vectors with different lengths for the product constraint.
+To measure the effect that the length of the vectors has on the running time, we ran our system on spreadsheets in which the length of the vectors was gradually increased.
+Figure~\ref{fig:runtime_analysis}b summarizes our results and shows that the length of the vectors has a small impact on the performance for most templates.
+
+As discussed in \samuel{???}, our implementation attempts to detect as fast as possible if a subassignments does not satisfy a constraint template.
+Our experiments satisfy the assumption that the large majority of possible subassignments do not satisfy most constraints.
+Therefore, the results confirm that our implementation is able to severely limit the impact of the length of vectors.
+
+Most of the increase in running time can be attributed to the \textit{RANK} constraint.
+The implementation is not always able to detect subassignments that do not satisfy \textit{RANK} quickly and often has to fall back to analyzing an entire vector.
+
+\paragraph{Effect of block size}
+To measure the effect of changing the block size, the number of vectors is set to~$32$ and the vectors are split into varying number of blocks (up to ~$32$).
+Figure~\ref{fig:runtime_analysis}c shows the running times of \sname for various settings.
+As the blocks become smaller, the total running time first decreases before increasing again.
+Looking at the running times for aggregate constraints and non-aggregate constraints  separately reveals that they are affected differently by the block size.
+For aggregate constraints (\samuel{xxx line}) the running time decreases with the block size, while for non-aggregate constraints (\samuel{yyy line}) the running time first stays more or less constant but then increases more strongly as the size of the blocks becomes very small.
 
 Aggregate constraints, e.g. sum or max, have an argument that takes subblocks of varying sizes.
-Therefore, the size of the blocks \tias{in what phase? the max. contigiuous? be more precise} in the spreadsheet also influences the runtime of these constraints.
-For a block of size~$m$ there are $\frac{m(m + 1)}{2}$ or $O(m^2)$ subblocks, thus, if a spreadsheet has more but smaller sized blocks, the search for aggregate constraints will be faster.
-Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$.
-The \tias{explain what THE experiment is, and how you set it up} synthetic experiments confirm that splitting a constant number of vectors into more, smaller blocks improves the performance of the aggregate constraint search (Figure~\ref{fig:runtime_analysis}c, green line).
-
-These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes small.
-Since these constraints have single vector arguments, the block size should not have a strong effect.
-However, in practice, when there are many small blocks, the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
-Since the implementation of that last phase is more adept at generating these combinations and does not need to check for constraints on all these combinations, shifting this workload to the assignment \tias{more precise: to superblock?} phase slows down the algorithm.
-\tias{Point to figure to support this (actually, move that line into 4c??); also, if this is true, you are basically saying that the CP-based first phase is worse then the second phase?
+Contrary to non aggregate constraints, which only have single vector argument, superblock assignments from phase~1 are not split into single vectors subassignments in phase~2.
+Instead, for an argument with variable size, the system tries to find any subblock of the given superblock that satisfies the constraint.
+Since, for a superblock of size~$m$, the number of subblocks is $O(m^2)$ \samuel{ref?}, the search for aggregate constraints will be faster if there are more but smaller blocks\footnote{Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$}.
+
+Non-aggregate constraints have only single vector arguments.
+Therefore, the total number of subassignments generated by the algorithm is not influenced by the block size.
+By splitting a set of vectors into many blocks, however, there will be a larger number of superblocks in phase~1 and, therefore, a larger number of superblock assignments have to be generated.
+Recall that in phase~1 the algorithm additionally tests block level properties for every superblock assignment.
+In the generated spreadsheets these tests can only disqualify few or no superblock assignments, therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.
+
+Splitting unrelated vectors into smaller subblocks can decrease the running time for some spreadsheets by reducing the workload for aggregate constraints.
+However, the two phase approach is also well suited to operate on large blocks and eliminate superblock assignments that are not compatible with constraint template signatures.
+\samuel{Does this conclusion sound sound?}
+
+% \tias{Point to figure to support this (actually, move that line into 4c??); also, if this is true, you are basically saying that the CP-based first phase is worse then the second phase?
 That kind-of demotivates the idea of using CP in first phase...
@@
 \subsection{Effectiveness on real spreadsheets}
-In order to obtain a benchmark of spreadsheets we looked at three sources:
+In order to test our approach on real data we assembled a benchmark of spreadsheets from three sources:
 1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports (available at the same repository\footref{github-link}).
@@
 Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
-Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
+% Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
 \tias{Weird here because purpose unclear, point to this when discussing some result for which this is important}
@@
 The average runtime across all spreadsheets is $0.13s$.
-Across categories the average running times fluctuate in the range between $0.09s$ to $0.22s$.
-While most spreadsheets can be analyzed very quickly, running times can go up to around $1s$ for others. ... \tias{Explain by relating to previous section of synthetic data}
+Across categories the average running times fluctuate in the range between $0.08s$ to $0.20s$.
+While most spreadsheets can be processed very quickly, running times can go up to around $0.85s$ for others.
+In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
+Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets∞m.
 
@@
 
-Looking at the runtime per constraint template individually (including equality\tias{what does this mean?}), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
-It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates (each $3.5\% - 4\%$).
+Looking at the runtime per constraint template individually, the slowest by far is fuzzy lookup using $\pm 21\%$ of the running time.
+It is followed by product ($\pm 7\%$), relative difference ($\pm 6\%$), running total ($\pm 5\%$), difference ($\pm 5\%$) and various row-aggregates (each $5\% - 3\%$).
 
@@
 However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data).
-This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.
+This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.\sam{Check numbers}
 

### 2016-09-23 Paper/ECAI/tabular.tex
Edits evaluation and constraints section (fail fast)

@@
 
-\author{Samuel Kolb \and Sergey Paramonov \and Tias Guns \and Luc De Raedt}
-\institute{KULeuven}
+\author{Samuel Kolb* \and Sergey Paramonov* \and Tias Guns \and Luc De Raedt}
+\institute{KULeuven\\ {*} Shared first author}
 
@@
 While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
-Based on the assumptions that most subassignments will not satisfy a constraint template, the implementations are geared to fail-fast when possible.
+
+Based on the assumptions that most subassignments will not satisfy a constraint template, the implementations are geared to \textit{fail-fast} when possible.
+Specifically, the tests try to prove that a subassignment does not satisfy a constraint template as fast as possible.
+For example, it can be often shown that a constraint does not hold by only looking first entries of every subblock (e.g.
+if $\sbl{r}[0] \neq \sbl{1}[0] \times \sbl{2}[0]$ then $\sbl{r} = \sbl{1} \times \sbl{2}$ does not hold).
+As a result, the time to check if a subassignment satisfies a constraint templates with a fail-fast implementation requires constant time\footnote{If the number of entries to look at is bounded by a constant, the vector lengths are smaller than the number of subassignments to test ($a$) and the time to test subassignments that satisfy the constraint is $O(a)$, then the time to test a subassignment is amortized $O(1)$} in most cases and does not depend on the length of the vectors.
+
+%\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}\sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
+%For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
+%For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of length~$k$, the worst case complexity becomes $O(n * m * k)$.
+%However, in practice, most subassignments will not satisfy the constraint template.
+%Since our implementation tries to account for this by failing fast\tias{explain more precisely what you mean with 'failing fast'}, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
+%For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).
 
@@
 5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors.
+\samuel{They are usually not copied}
 
@@
 %The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}. % no such general pointer, specific pointers only
-
-\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}\sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
-For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
-For a constraint with arity~$n$, and a subassignment consisting of $n$~blocks with $m$~vectors of length~$k$, the worst case complexity becomes $O(n * m * k)$.
-However, in practice, most subassignments will not satisfy the constraint template.
-Since our implementation tries to account for this by failing fast\tias{explain more precisely what you mean with 'failing fast'}, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
-For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).
 
@@
 %\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
-The overall recall achieved by \sname is~$0.88~(81/92)$. The overall and average recall per category is also high, as shown in Table~\ref{tbl:category_overview}.
-The intended constraints not discovered correspond to constraints not supported by the system (supported recall is 1).
-\tias{Rewrite and go into more detail about all supported found, some intended not support, which ones, what category, just some more details}
-
-\paragraph{Q3: precision}
+The overall recall achieved by \sname is~$0.88~(81/92)$.
+When looking at the recall per category, \sname is able to consistently obtain high values for both overall and average recall, as shown in Table~\ref{tbl:category_overview}.
+The intended constraints not discovered correspond to constraints that are not supported by the system (supported recall is 1).
+Therefore, \sname is able to find all constraints for supported constraint templates, which confirms that our bias (signature) is not too restrictive.
+
+The intended constraints that are not discovered correspond with nested constraints (e.g.
+$R = \frac{W}{(L/100)^2}$), variants of supported constraints (e.g.
+series that do not start at 1) or more complex versions of supported constraints (e.g.
+conditional aggregates that automatically infer the condition).
+
+\paragraph{Q3: Precision}
 %\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
 The final question, Q3, is about the amount and type of non-intended constraints found.
-Across all spreadsheets \sname achieves a precision of~$0.79~(81/102)$.
-While the average precision per spreadsheet is over $0.90$ for all categories, the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).
-
-Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints.
-This is caused by duplicate data or multiple ways to calculate the same vectors.
-\tias{More details or examples; we copy intended constraints for equivalent/duplicate vectors, so not an issue?}
-For these spreadsheets the precision could be increased by post-processing the constraints to detect implications or heuristically selecting one among multiple ways to compute the same vector.
+Across all spreadsheets \sname achieves a precision of~$0.79~(81/103)$.
+While the average precision per spreadsheet is high, over $0.90$ in every category, however the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).
+
+Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints, e.g.
+an spreadsheet computes aggregates on its inventory but copies the data column for every aggregate.
+Since every aggregate can be calculated based on any of the (equal) columns, many of the resulting constraints will be marked as unintended.
+Usually, unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data.
+A way to increase the precision is by post-processing the constraints to detect implications and heuristically selecting one among multiple ways to compute the same vector.
 

### 2016-09-23 Paper/ECAI/tabular.tex
Updated plots, plot references (and block detection)

@@
 \samuel{Should we not say just with the visual tool?}
- \\ % split of weird sentence
-Blocks are detected automatically using our block detection algorithm (Section~\ref{sec:make_groups}) unless a spreadsheet contains ambiguous \textit{None} values.
-For those ambiguous cases the blocks are specified manually. \samuel{I have yet to make this more precise}
+Blocks are detected automatically using our block detection algorithm (Section~\ref{sec:make_groups}) except for spreadsheets that contain \textit{None} values or (partially) numeric headers, in which case the blocks are specified manually.
 
@@
   \caption{Loglog plots of the experiments with synthetic (random) data.
-In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plots~$c$ and~$d$ the number of blocks is increased.}
+In plot~$a$ the number of vectors is increased; in plot~$b$ the size of the vectors is increased; and in plot~$c$ the number of blocks is increased (the block size is decreased).}
   \label{fig:runtime_analysis}
@@
 % \tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}\samuel{?}
-\samuel{=> benchmark}
 
@@
 
-As discussed in \samuel{???}, our implementation attempts to detect as fast as possible if a subassignments does not satisfy a constraint template.
+As discussed in Section~\ref{sec:which_cons}, our implementation attempts to detect as fast as possible if a subassignments does not satisfy a constraint template.
 Our experiments satisfy the assumption that the large majority of possible subassignments do not satisfy most constraints.
@@
 Looking at the running times for aggregate constraints and non-aggregate constraints  separately reveals that they are affected differently by the block size.
-For aggregate constraints (\samuel{xxx line}) the running time decreases with the block size, while for non-aggregate constraints (\samuel{yyy line}) the running time first stays more or less constant but then increases more strongly as the size of the blocks becomes very small.
+For aggregate constraints (green line) the running time decreases with the block size, while for non-aggregate constraints (red line) the running time first stays more or less constant but then increases more strongly as the size of the blocks becomes small.
 
@@
 In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
-Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets∞m.
+Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets.
 

### 2016-09-23 Paper/ECAI/tabular.tex
abstract, intro and rel. work pass

--- 2016-09-23 Paper/ECAI/tabular.tex
@@
 We represent common spreadsheet formulas and relations through predicates and expressions whose arguments must satisfy the inherent properties of the formula/relation.
-The challenge is to automatically infer the set of constraints that is present in the data, without user information or labeled examples.
+The challenge is to automatically infer the set of constraints that is present in the data, without labeled examples or user feedback.
 We propose a two-stage generate and test method where the first stage uses constraint solving techniques to efficiently reduce the number of candidates in a generic way, based on the predicate signatures.
@@
 \begin{itemize}
-\item we define the tabular constraint learning problem, where the goal is to constraints that range over entire rows or columns in an unsupervised way;
+\item we define the tabular constraint learning problem, where the goal is to find constraints that range over entire rows or columns in an unsupervised way;
 \item we propose an effective 2-stage generate and test method where the first stage reasons only over properties of contiguous blocks of rows/columns, and the second stage continues to investigate individual rows and columns and their content;
@@
 This paper is organized as follows.
-Section \ref{sec:related_work} presents a detailed discussion of the related work.
-Section \ref{sec:formalization} introduces concepts relevant to our approach and our problem statement.
-Section \ref{sec:approach} presents our approach.
-Section \ref{sec:evaluation} presents the evaluation of the approach.
-Section \ref{sec:applications} shows how our system can be used for applications.
-Section \ref{sec:conclusions} provides conclusions.
+Section \ref{sec:related_work} discusses related work.
+Section \ref{sec:formalization} introduces the relevant concepts and defines the problem formally while Section \ref{sec:approach} presents our approach.
+This is evaluated in Section \ref{sec:evaluation} after which we show two applications of how our system could be used in Section \ref{sec:applications}, and we conclude in Section \ref{sec:conclusions}.
 
@@
 constraints in multiple relations (or tables). Unlike typical logical and relational learning
-approaches, it focuses on data in spreadsheets and constraints on columns or rows in the sheets.
+approaches, it focuses on data in spreadsheets and constraints involving columns and rows in the sheets.
 Furthermore, it derives a set of simple ``atomic'' constraints rather than a set of rules that each consist of a conjunction of literals as in clausal discovery \cite{claudien,lallouet}.
@@
 Two well-known approaches include ModelSeeker \cite{modelseeker} and Quacq \cite{Quacq}.
-The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints from a large constraint library that hold in the vector.
+The former starts from a single example in the form of a vector of values and then exhaustively looks for the most specific constraints that hold given a large constraint library.
 To this aim, it cleverly enumerates different de-vectorized tables; for instance, if the initial vector was of dimension $1 \times 20$, it would consider rearrangements of size $2 \times 10$, $4 \times 5$, $5 \times 4$, $\dots$ and Modelseeker would then look for both column, row  and even diagonal constraints.
-Key differences are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and often typical for constraint programming community, e.g.
+Key differences with our work are that we assume given spreadsheets with \textit{multiple} tables; and that ModelSeeker's constraints are restricted to integer values and constraints that are typical for the constraint programming community, e.g.
 aimed at combinatorial optimization and scheduling.
-Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired on Mitchell's version spaces to process examples; its successor Quacq uses active learning to query the user regarding satisfiability.
+Conacq \cite{Conacq} acquires a conjunction of constraints using techniques inspired by Mitchell's version space algorithm to process examples; its successor Quacq uses active learning to query the user regarding satisfiability.
 Each of the examples/queries is processed for each constraint, hence an important issue is how fast the algorithm converges to the entire solution set (e.g.
@@
 However, they focus specifically on noisy physical data and formulas, e.g.
-a single variable can take values from a series of measurements to get a regression parameter estimate or they apply specific data transformation such as numerical calculation of time and partial derivatives, i.e.
-the method works with approximations; the equation structures are heavily biased towards parameter estimation, while we focus on discovery of different types of spreadsheet constraints which is an exact constraint satisfaction problem.
-\sergey{rewritten, Samuel, does it sound fine now?}
-\tias{Last sentence too vague, what is specific about physical data?
-btw, it is supervised, e.g.
-input-output and we are general?
-Also, it is 'one' conjunction of functions, e.g.
-one equation, or not?
-Please rewrite more precisely}
+a single variable can take values from a series of measurements to get a regression parameter estimate or they apply specific data transformation such as numerical calculation of time and partial derivatives.
+In other words, the method works with approximations and the equation structures are heavily biased towards parameter estimation, while our goal is to discover hard constraints that hold over a subset of the rows and columns.
 %As the conjunction for typical constraint programming problems (such as Sudoku) can be quite large, a key problem is to cope with redundant constraints.
@@
 %    However, due to the location of the header of~$T_2$ we can infer that it consists of row data.
-    This example combines several Excel sheets based on exercise sessions in the book ``MS Excel 2010''~\cite{excel_book}.
+    This example combines several Excel sheets based on exercises in the book ``MS Excel 2010''~\cite{excel_book}.
     % I guess you want to save space by having everything in one figure, but it might be useful to have a smaller figure on the first page just to explain this example in a bit more detail (or use a smaller example here); it's quite overwhelming now}
@@
   \vspace{-10pt}
-  \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5) constraints.
+  \caption{Constraints learned for the tables above, except 18 \textit{ALLDIFFERENT}, 2 \textit{PERMUTATION} and 5 \textit{FOREIGNKEY} constraints not shown.
   Constraints marked with $*$ were not present in the original spreadsheets.}
@@
 \subsubsection{Constraints} \label{sec:which_cons}
-The \samuel{count} constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
+The constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
 We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
@@
 merge split-up tables because of None values, or remove headers of all-textual tables).
-\samuel{Should we not say just with the visual tool?}
 \\ % split of weird sentence
-Blocks are detected automatically using our block detection algorithm (Section~\ref{sec:make_groups}) unless a spreadsheet contains ambiguous \textit{None} values.
-For those ambiguous cases the blocks are specified manually. \samuel{I have yet to make this more precise}
-
-% Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
+Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
 \tias{What does this mean, 'group definitons'??
@@
 
-All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints, that are expected to be (re-)discovered.
-These were determined by us using either the formulas of the original sheet (if present), or inferring them based on, for example, the header names in the sheet.
+All spreadsheets also have a set of ground-truth of constraints, we call these the \textbf{intended} constraints that are expected to be (re-)discovered.
+These were determined by us using the formulas of the original sheet (if present), the headers in the sheet as well as general context if need.
+\tias{last part too vague!!
+Give example of where 'context' was needed/used.
+If nowhere, cut that part of sentence}.
 In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas.
@@
 Section~\ref{sec:which_cons} discusses the 32 constraints currently supported by \sname \tias{It does, right?
-Should also contain '32' number}\sam{I will recount at the end}.
+Should also contain '32' number}.
 Some of the intended constraints are currently outside of the scope of \sname however, more specifically nested mathematical or nested logical formulas.
@@
 This is mostly due to the spurious RANK constraints.
-We note that the tables are quite short, which increases the chance of a constraint like \textit{RANK} to be true by chance; we expect less spurious constraints on larger datasets.
+We note that the tables are quite short, which increases the chance of a constraint like 'RANK' to be true by chance; we expect less spurious constraints on larger datasets.
 
@@
 
+There are multiple factors that can impact the running time of \sname: the number of vectors, the length of the vectors, the number of vectors per block, the type of cells, the number of tables and the number of constraints that are in the data.
+Moreover, the running time depends on the number of the constraint templates implemented in the system and their arity.
+
 In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the length of the vectors; and 3)~the size of the blocks.
-This analysis is accomplished by testing our system on generated spreadsheets.
-
-Recall that our method uses a 2-phase approach where in the first phase, the superblock assignments are generated (Section~\ref{sec:algo:super} and in the second phase, the subblock assignments are generated (Section~\ref{sec:algo:subgr}).
+\tias{Why only these and not the other?
+Motivated/explain, or also measure their effect ; )}
+
+Recall that our method uses a 2-phase approach where in the first phase, the superblock assignments are generated (Section ...) and in the second phase, the subblock assignments are generated (Section ...).
 
 \paragraph{Spreadsheet generation}
-For our analysis, we perform multiple experiments on synthetically generated spreadsheets.
-
-Many constraints require either numeric or discrete data, therefore, the type of cells also has an impact on the running time.
-Since integers fulfill both of those typing restrictions, the generated spreadsheets contain only integer data.
-
-One of the assumptions guiding the implementation is that most of the possible constraints are false.
-To simulate this setting, the generated spreadsheets contain random data.
-There will be few constraints that hold on these spreadsheets, however, random data is likely to produce vectors that fulfill the \textit{ALLDIFFERENT} constraint, which is a requirement for many other constraint templates.
-
-In order to also analyse constraints that require their arguments to be picked from different tables, the spreadsheets are made up of two tables.
+We do this analysis on a set of synthetically generated spreadsheets.
+The spreadsheets contains random integers\footnote{For strings and floats the results are similar but the running times are typically much lower.} and are made up of two tables.
 For one table the number of vectors and block size is fixed to two, while for the other table these properties are varied.
-The number of vectors and the block size in the experiments are only reported for the adaptable table.
-The length of the vectors can also be adapted and applies to both tables.
-Unless otherwise mentioned, the adaptable table consists of one block containing all its vectors and by default the number of vectors and vector length are both~$8$.
+Moreover, the length of the vectors (for both tables) can be adapted, while the default is~$8$.
 %The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}. % no such general pointer, specific pointers only
 
-\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}\sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
+\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}
 For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
@@
 
+\paragraph{Effect of length of vectors}
+%The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
+As can be seen in Figure~\ref{fig:runtime_analysis}b\tias{This is the first figure discussed so must be a}, increasing the length of the vectors has a negligible impact on the performance for most template.
+Most of the increase in running time can be attributed to the RANK constraint which is less suited for failing fast \tias{Too vague, 'less suited', be precise}; the runtime for detecting all constraints except RANK is shown in green to highlight this.
+
 \paragraph{Effect of number of vectors}
-Figure~\ref{fig:runtime_analysis}a shows the results of our experiments, in which \sname was run on generated spreadsheets containing an increasing number of (equally sized) vectors.
-As shown in the figure, the number of vectors in a spreadsheet can have a significant impact on the running time.
-
-This behavior is to be expected when not many superblock assignments can be discarded in phase~1, for example, when all vectors have the same type and size.
-In this case, for a constraint template that requires $n$~single vector arguments and a spreadsheet with $v$~vectors, up to $\frac{v!}{(v - n)!}$ subassignments are generated throughout the algorithm.
-
-Typically, we expect the number of generated subassignments to be much smaller in real world spreadsheets and, therefore, the running time to be less sensitive to the number of vectors.
-This is because in phase~1 entire blocks can be disqualified that do not match the signature requirements, e.g. blocks of vectors with different lengths for the product constraint.
-% \tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}\samuel{?}
-\samuel{=> benchmark}
-
-% Many constraints take single vector arguments, therefore, the total number of subassignments that are generated in the second phase of the algorithm depends on the number of vectors in the spreadsheet and the arity of the constraint.
-
-% \tias{unclear, this is about phase 2, the number to test, or the actual testing of each?
+On the contrary, the number of vectors in a spreadsheet can have a significant impact on the running time.
+Many constraints take single vector arguments, therefore, the number of subassignments \tias{unclear, this is about phase 2, the number to test, or the actual testing of each?
 explain and relate to Sec 4.3.2 what aspect you are discussing here} to be tested depends on the number of vectors in the spreadsheets and the arity of the constraint.
-
-% For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
+For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
 \tias{Better move to 4.3.2?
 or have there too anyway, no new facts in 'experiment' section}
-% As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
-% This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors \tias{more details about equal sized: is this a strict requirement for this to happen?
+As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
+This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors \tias{more details about equal sized: is this a strict requirement for this to happen?
 I assume this is worst case and does not happen that much in practice?
 (the next sentences half hint at that, be more explicit)} is increased (Figure~\ref{fig:runtime_analysis}a).
-
-\paragraph{Effect of length of vectors}
-%The benchmark experiments show that the number of cells in a spreadsheet has a very limited impact on the running time.
-To measure the effect that the length of the vectors has on the running time, we ran our system on spreadsheets in which the length of the vectors was gradually increased.
-Figure~\ref{fig:runtime_analysis}b summarizes our results and shows that the length of the vectors has a small impact on the performance for most templates.
-
-As discussed in \samuel{???}, our implementation attempts to detect as fast as possible if a subassignments does not satisfy a constraint template.
-Our experiments satisfy the assumption that the large majority of possible subassignments do not satisfy most constraints.
-Therefore, the results confirm that our implementation is able to severely limit the impact of the length of vectors.
-
-Most of the increase in running time can be attributed to the \textit{RANK} constraint.
-The implementation is not always able to detect subassignments that do not satisfy \textit{RANK} quickly and often has to fall back to analyzing an entire vector.
-
-\paragraph{Effect of block size}
-To measure the effect of changing the block size, the number of vectors is set to~$32$ and the vectors are split into varying number of blocks (up to ~$32$).
-Figure~\ref{fig:runtime_analysis}c shows the running times of \sname for various settings.
-As the blocks become smaller, the total running time first decreases before increasing again.
-Looking at the running times for aggregate constraints and non-aggregate constraints  separately reveals that they are affected differently by the block size.
-For aggregate constraints (\samuel{xxx line}) the running time decreases with the block size, while for non-aggregate constraints (\samuel{yyy line}) the running time first stays more or less constant but then increases more strongly as the size of the blocks becomes very small.
+In practice, in our benchmark experiments, the assignment step of our algorithm is able to limit this rapid increase in running time for larger numbers of vectors \tias{How?
+explain, if you explain in the next sentence then start it with 'This is because' or alike, reads better}.
+The superblock assignment phase is able to disqualify entire blocks that do not match the signature requirements, i.e. blocks of vectors with different lengths for the product constraint.
 
 Aggregate constraints, e.g. sum or max, have an argument that takes subblocks of varying sizes.
-Contrary to non aggregate constraints, which only have single vector argument, superblock assignments from phase~1 are not split into single vectors subassignments in phase~2.
-Instead, for an argument with variable size, the system tries to find any subblock of the given superblock that satisfies the constraint.
-Since, for a superblock of size~$m$, the number of subblocks is $O(m^2)$ \samuel{ref?}, the search for aggregate constraints will be faster if there are more but smaller blocks\footnote{Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$}.
-
-Non-aggregate constraints have only single vector arguments.
-Therefore, the total number of subassignments generated by the algorithm is not influenced by the block size.
-By splitting a set of vectors into many blocks, however, there will be a larger number of superblocks in phase~1 and, therefore, a larger number of superblock assignments have to be generated.
-Recall that in phase~1 the algorithm additionally tests block level properties for every superblock assignment.
-In the generated spreadsheets these tests can only disqualify few or no superblock assignments, therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.
-
-Splitting unrelated vectors into smaller subblocks can decrease the running time for some spreadsheets by reducing the workload for aggregate constraints.
-However, the two phase approach is also well suited to operate on large blocks and eliminate superblock assignments that are not compatible with constraint template signatures.
-\samuel{Does this conclusion sound sound?}
-
-% \tias{Point to figure to support this (actually, move that line into 4c??); also, if this is true, you are basically saying that the CP-based first phase is worse then the second phase?
+Therefore, the size of the blocks \tias{in what phase? the max. contigiuous? be more precise} in the spreadsheet also influences the runtime of these constraints.
+For a block of size~$m$ there are $\frac{m(m + 1)}{2}$ or $O(m^2)$ subblocks, thus, if a spreadsheet has more but smaller sized blocks, the search for aggregate constraints will be faster.
+Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$.
+The \tias{explain what THE experiment is, and how you set it up} synthetic experiments confirm that splitting a constant number of vectors into more, smaller blocks improves the performance of the aggregate constraint search (Figure~\ref{fig:runtime_analysis}c, green line).
+
+These experiments, however, also show that the running time for various non-aggregate constraints increases as the size of blocks becomes small.
+Since these constraints have single vector arguments, the block size should not have a strong effect.
+However, in practice, when there are many small blocks, the superblock assignment phase has to generate a very large number of combinations, while for larger blocks this work is done mostly during the subassignment generation phase.
+Since the implementation of that last phase is more adept at generating these combinations and does not need to check for constraints on all these combinations, shifting this workload to the assignment \tias{more precise: to superblock?} phase slows down the algorithm.
+\tias{Point to figure to support this (actually, move that line into 4c??); also, if this is true, you are basically saying that the CP-based first phase is worse then the second phase?
 That kind-of demotivates the idea of using CP in first phase...
@@
 \subsection{Effectiveness on real spreadsheets}
-In order to test our approach on real data we assembled a benchmark of spreadsheets from three sources:
+In order to obtain a benchmark of spreadsheets we looked at three sources:
 1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports (available at the same repository\footref{github-link}).
@@
 Table~\ref{tbl:category_overview} gives an overview of the spreadsheets in the different categories and summarizes the results of our experiments.
-% Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
+Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
 \tias{Weird here because purpose unclear, point to this when discussing some result for which this is important}
@@
 The average runtime across all spreadsheets is $0.13s$.
-Across categories the average running times fluctuate in the range between $0.08s$ to $0.20s$.
-While most spreadsheets can be processed very quickly, running times can go up to around $0.85s$ for others.
-In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
-Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets∞m.
+Across categories the average running times fluctuate in the range between $0.09s$ to $0.22s$.
+While most spreadsheets can be analyzed very quickly, running times can go up to around $1s$ for others. ... \tias{Explain by relating to previous section of synthetic data}
 
@@
 
-Looking at the runtime per constraint template individually, the slowest by far is fuzzy lookup using $\pm 21\%$ of the running time.
-It is followed by product ($\pm 7\%$), relative difference ($\pm 6\%$), running total ($\pm 5\%$), difference ($\pm 5\%$) and various row-aggregates (each $5\% - 3\%$).
+Looking at the runtime per constraint template individually (including equality\tias{what does this mean?}), the slowest by far is fuzzy lookup using $\pm 36\%$ of the running time.
+It is followed by product ($\pm 5\%$), relative difference ($\pm 5\%$) and various row-aggregates (each $3.5\% - 4\%$).
 
@@
 However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data).
-This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.\sam{Check numbers}
+This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.
 

### 2016-09-23 Paper/ECAI/tabular.tex
Added ordered constraints

--- 2016-09-23 Paper/ECAI/tabular.tex
@@
   \vspace{-10pt}
-  \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2) and \textit{FOREIGNKEY} (5) constraints.
+  \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2), \textit{FOREIGNKEY} (5) and \textit{ORDERED} (6) constraints.
   Constraints marked with $*$ were not present in the original spreadsheets.}
@@
 
+    \ecordered{\sbl{x}}
+      & $\numeric(\sbl{x})$
+
+      & $i < j$: $\sbl{x}[i] \leq \sbl{x}[j]$.
+      \\[\PADSEP] \hline
+
     \ecseries{\sbl{x}}
@@
     \eclookupfuzzy{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}}
-      & \textit{Same as lookup}
+      & \textit{Same as lookup}; except $\ecordered{\sbl{pk}}$ instead of \ecfkey{\sbl{fk}}{\sbl{pk}}
 
@@
   \centering
-  \includegraphics[width=0.50\linewidth]{figures/constraint_dependency2.pdf}
+  \includegraphics[width=0.60\linewidth]{figures/constraint_dependency.pdf}
   \caption{Dependency graph of Table~\ref{table:constraints}; an arrow from~$s_1$ to~$s_2$ indicates that $s_2$ depends on~$s_1$ (its signature includes $s_1$).
@@
 \subsubsection{Constraints} \label{sec:which_cons}
-The \samuel{count} constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
+The 33 constraints that are currently supported in our system are shown in Table~\ref{table:constraints}.
 We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
@@
 In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas.
-We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY} and \textit{PERMUTATION}.
+We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY}, \textit{PERMUTATION} and \textit{ORDERED}.
 All constraints are stored in their canonical form.
@@
 
-Section~\ref{sec:which_cons} discusses the 32 constraints currently supported by \sname \tias{It does, right?
-Should also contain '32' number}\sam{I will recount at the end}.
+Section~\ref{sec:which_cons} discusses the 33 constraints currently supported by \sname.
 Some of the intended constraints are currently outside of the scope of \sname however, more specifically nested mathematical or nested logical formulas.
@@
   \centering
-  \includegraphics[width=1\linewidth]{figures/scatter_plots.pdf}
+  \includegraphics[width=0.8\linewidth]{figures/scatter_plots.pdf}
   \caption{Loglog plots of the experiments with synthetic (random) data.
@@
 However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data).
-This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.\sam{Check numbers}
+This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.\sam{Check numbers} % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41
 
@@
 Therefore, \sname is able to find all constraints for supported constraint templates, which confirms that our bias (signature) is not too restrictive.
+
+% The intended constraints not discovered correspond to constraints that are not supported by the system (supported recall is 0.99), except for one intended \textit{SUMIF} constraint.
+% The supported recall value mostly confirms that our bias (signature) is not too restrictive and is able to find constraints for supported constraint templates.
+% By changing the signature of \textit{SUMIF} to allow \sbl{pk}
 

### 2016-09-23 Paper/ECAI/tabular.tex
Formatting

@@
 \item For each constraint template~$s$, generate all assignments $(\sbl{1}, \ldots ,\sbl{n})$ where each~$\sbl{i}$ is a superblock and the~$\sbl{i}$ are \textit{compatible} (defined below) with the signature of~$s$.
-\item For each generated assignment $(\sbl{1}, \ldots ,\sbl{n})$ in 3a), find all valid constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that for all $i$ holds $\ssbl{i} \sqsubseteq \sbl{i}$ and the signature and definition of~$s$ are satisfied.
+\item For each assignment $(\sbl{1}, \ldots ,\sbl{n})$ generated in 3a), find all constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that for all $i$ holds $\ssbl{i} \sqsubseteq \sbl{i}$ and the signature and definition of~$s$ are satisfied.
 \end{enumerate}

### 2016-09-23 Paper/ECAI/tabular.tex
section 3 and beginning of section 4

--- 2016-09-23 Paper/ECAI/tabular.tex
@@
 
-We first introduce some terminology and the concept of a \template, after which we define the problem and make some additional considerations.
+We first introduce some terminology and the concept of a \template, after which we define the problem and discuss some additional considerations.
 
@@
 \begin{definition}
-A \textbf{block} has to satisfy three conditions: 1)~it contains only entire rows or entire columns of a single headerless table; 2)~it is contiguous; and 3)~it is type-consistent.
+A \textbf{block} has to satisfy three conditions: 1)~it contains entire rows or entire columns of a single headerless table; 2)~it is contiguous; and 3)~it is type-consistent.
 The rows or columns have to be contiguous in the original table meaning that they must visually form a block in the table; and each of the rows/columns has to be of the same type. 
@@
 \begin{itemize}
-\item \textit{type}: a block is type-consistent, so it has one type
-\item \textit{table}: the table that the block belongs to
-\item \textit{orientation}: either row-oriented or column-oriented
-\item \textit{size}: the number of vectors a block contains
-\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
+\item \textit{type}: a block is type-consistent, so it has one type;
+\item \textit{table}: the table that the block belongs to;
+\item \textit{orientation}: either row-oriented or column-oriented;
+\item \textit{size}: the number of vectors a block contains;
+\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length;
 \item \textit{rows}: the number of rows in the block
-(in row-oriented blocks this is equivalent to the size)
-\item \textit{columns}: the number of columns in the block (in row-oriented blocks this is equivalent to the length)
+(in row-oriented blocks this is equivalent to the size);
+\item \textit{columns}: the number of columns in the block (in row-oriented blocks this is equivalent to the length).
 \end{itemize}
@@
 This can concern properties of individual blocks as well as relations between properties of arguments, for example that the corresponding blocks must belong to the same table or have equal length.
-In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies when the arguments of a constraint are well-formed.
-We can capture this bias for a template~$s$ using a predicate $\sigc$.
+In terms of logical and relational learning \cite{luc_book}, the \CSignature is known as the {\em bias} of the learner, it specifies (a conjunction of) the properties that the arguments must have for the constraint to be well-formed.
+We denote this bias for a template~$s$ using the predicate $\sigc$.
 \item \textit{\CFunction} is the actual definition of the constraint that specifies when the constraint holds.
@@
 
-  \item \CSignature: $\sbl{r}$ has to be a single vector ($\textit{size}=1$) while $\bsbl{x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font.
+  \item \CSignature: $\sbl{r}$ has to be a single vector ($\textit{size}=1$) while $\bsbl{x}$ can be a block ($\textit{size} \geq 1$), which can be derived from the use of a normal or \textbf{bold} font.
   Both blocks have to be numeric.
   This constraint is orientation-specific, so it requires that the number of rows in $\bsbl{x}$ equals the length of $\sbl{r}$.
-  Moreover, we add to the bias that the number of columns to sum over is at least~$2$.
+  Moreover, the bias specifies that the number of columns to sum over is at least~$2$ (to avoid $\bsbl{x}$ being a single column).
 
@@
 Therefore, in Section~\ref{sec:approach} we first explain our generic method for finding all constraints, before describing some optimizations that avoid obvious equivalences (Section~\ref{sec:opts}).
-Later, we will investigate the impact of redundancy in the experiments.
+In the experiments we will investigate the impact of redundancy further.
 
@@
 \item For each constraint template~$s$, generate all assignments $(\sbl{1}, \ldots ,\sbl{n})$ where each~$\sbl{i}$ is a superblock and the~$\sbl{i}$ are \textit{compatible} (defined below) with the signature of~$s$.
-\item For each generated assignment $(\sbl{1}, \ldots ,\sbl{n})$ in 3a), find all valid constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that for all $i$ holds $\ssbl{i} \sqsubseteq \sbl{i}$ and the signature and definition of~$s$ are satisfied.
+\item For each generated assignment $(\sbl{1}, \ldots ,\sbl{n})$, find all valid constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that $\ssbl{i} \sqsubseteq \sbl{i}$ holds for all $i$ and the signature and definition of~$s$ are satisfied.
 \end{enumerate}
@@
 Superblocks~$\sbl{1}, \sbl{3}$ and $\sbl{5}$ are valid candidates for~$\sbl{r}$, however, since they all have length~4, $\mathbf{\sbl{x}}$ needs at least 4~columns which is only satisfied by $\sbl{3}$.
-So the assignments generated in step 3a would be $(\sbl{1}, \sbl{3}), (\sbl{3}, \sbl{3}), (\sbl{5}, \sbl{3})$.
-
-In step 3b, we start from an assignment $(\sbl{1}, \ldots, \sbl{n})$ with superblocks $\sbl{i}$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks.
-As we only have to consider the blocks $\ssbl{i} \sqsubseteq \sbl{i}$ contained in each superblock, this is typically much easier.
-For the above example and $(\sbl{5}, \sbl{3})$, each of the vectors of $\sbl{5}$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks of $\sbl{3}$ for the right-hand side and verify the signature (e.g.
+So the assignments generated for $\textit{SUM}_{row}$ in step 3a would be $(\sbl{1}, \sbl{3}), (\sbl{3}, \sbl{3}), (\sbl{5}, \sbl{3})$.
+
+In step 3b, we start from an assignment $(\sbl{1}, \ldots, \sbl{n})$ with superblocks $\sbl{i}$ and generate and test all possible (sub)block assignments to the arguments, using the properties and actual data of the blocks.
+As we only have to consider the blocks $\ssbl{i} \sqsubseteq \sbl{i}$ contained in each superblock, this is typically much easier then the previous step.
+For $\textit{SUM}_{row}$ in the above example and assignment $(\sbl{5}, \sbl{3})$, each of the vectors of $\sbl{5}$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks of $\sbl{3}$ for the right-hand side and verify the signature (e.g.
 at least size $4$) and test whether for each row the definition is satisfied.
@@
 We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step~3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step~3b, Section~\ref{sec:algo:subgr}).
-In Section~\ref{sec:opts} we describe a number of optimizations designed to improve the algorithm.
+In Section~\ref{sec:opts} we describe a number of optimizations designed to improve the effectiveness algorithm.
 

### 2016-09-23 Paper/ECAI/tabular.tex
Rerun numbers

--- 2016-09-23 Paper/ECAI/tabular.tex
@@
     \textbf{\begin{tabular}[c]{@{}l@{}}Intended\\[-4.5pt] Constraints\end{tabular}} & 34 & 3.78 & 52 & 2.48 & 6 & 1.50 \\ \hline \hline
-\textbf{Recall} & 0.85 & 0.83 & 0.88 & 0.87 & 1.00 & 1.00 \\ \hline
+\textbf{Recall} & 0.88 & 0.85 & 0.88 & 0.87 & 1.00 & 1.00 \\ \hline
 \textbf{\begin{tabular}[c]{@{}l@{}}Recall\\[-4.5pt] Supported\end{tabular}} & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ \hline
 \textbf{Precision} & 0.94 & 0.96 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
-\textbf{Runtime (s)} & 1.62 & 0.18 & 1.76 & 0.08 & 0.81 & 0.20
+\textbf{Runtime (s)} & 0.93 & 0.10 & 0.93 & 0.04 & 0.79 & 0.20
 \end{tabular}
@@
 As Table~\ref{tbl:category_overview} (last row) shows, the runtime of our method on these spreadsheets is always fast.
-The average runtime across all spreadsheets is $0.13s$.
-Across categories the average running times fluctuate in the range between $0.08s$ to $0.20s$.
-While most spreadsheets can be processed very quickly, running times can go up to around $0.85s$ for others.
+The average runtime across all spreadsheets is $0.08s$.
+Across categories the average running times fluctuate in the range between $0.04s$ to $0.20s$.
+While most spreadsheets can be processed very quickly, running times can go up to around $0.65s$ for others.
 In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
@@
 Looking at the runtime per constraint template individually, the slowest by far is fuzzy lookup using $\pm 21\%$ of the running time.
-It is followed by product ($\pm 7\%$), relative difference ($\pm 6\%$), running total ($\pm 5\%$), difference ($\pm 5\%$) and various row-aggregates (each $5\% - 3\%$).
+It is followed by product ($\pm 7\%$), relative difference ($\pm 6\%$), running total ($\pm 5\%$), difference ($\pm 5\%$) and various row-aggregates ($\pm 5\%$).
 
 Our method uses dependencies to increase efficiency by finding constraints incrementally.
-To illustrate the effect this can have, we ran the benchmarks using FOREIGNKEY instead of ALLDIFFERENT as base constraint to find conditional aggregates such as SUMIF.
-This is a stronger assumption, and a too strong in practice as not all keys have to be used in the conditions.
-However, the total running times per category are decreased by 6\% (exercises), 30\% (tutorials) and~3\% (data).
-This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.\sam{Check numbers} % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41
+To illustrate the effect this can have, we ran the benchmarks without using \textit{ORDERED} as base constraint for fuzzy lookup.
+The total running times per category increase by 74\% (exercises), 84\% (tutorials) and~13\% (data).
+This shows that the use of dependencies can have a strong effect on running times by sharing the computation and pruning done for one constraint with another.
+ % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41
 
@@
 %\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
-The overall recall achieved by \sname is~$0.88~(81/92)$.
+The overall recall achieved by our system, \sname, is~$0.88~(81/92)$.
 When looking at the recall per category, \sname is able to consistently obtain high values for both overall and average recall, as shown in Table~\ref{tbl:category_overview}.

### 2016-09-23 Paper/ECAI/tabular.tex
pass over more of sec 4

--- 2016-09-23 Paper/ECAI/tabular.tex
@@
 
-Instead, we assume tables are specified by means of their coordinates within a spreadsheet and optionally a fixed orientation of each table.
+Instead, our algorithm will assume tables are specified by means of their coordinates within a spreadsheet and optionally a fixed orientation of each table.
 The orientation can be used to indicate that a table should be interpreted as having only rows or only columns.
@@
 The above assumptions do not hold for many tables.
-However, since the specification of tables is usually easy for humans, we propose a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/SergeyParamonov/TaCLe}\label{github-link}}).
+However, since the specification of tables is usually easy for humans, we employ a second approach which allows users to indicate tables using a visual tool (screenshots are available in the accompanying github repository with meta-data\footnote{\url{https://github.com/SergeyParamonov/TaCLe}\label{github-link}}).
 Users select tables excluding the header and optionally specify an orientation.
@@
 integer or floating point) representation (e.g.
-$2.75 \$$ as $2.75$ or $85\%$ as~$0.85$).
+$2.75 \$$ as $2.75$ and $85\%$ as~$0.85$).
 Then, each table is partitioned into maximal type-consistent blocks.
@@
 In our case, we define one variable $V_i$ for each argument of a constraint template.
-Each variable can be assigned each of the maximal blocks in \blocks, so the domain of the variables consists of $|\{\blocks\}|$ block identifiers.
+Each variable can be assigned to one of the superblocks in \blocks, so the domain of the variables consists of $|\blocks|$ block identifiers.
 
@@
 
-To reason over the blocks, we add to the constraint program a set of background facts that contain the properties of each block, namely its type, table, orientation, size, length and number of rows and columns.
-
-\begin{table}[tbh]
+To reason over the blocks, we add to the constraint program a set of background facts that contain the properties of each superblock, namely its type, table, orientation, size, length and number of rows and columns.
+
+\begin{table}[t]
   \caption{Translation of signature requirements to superblock constraints. The following need not be relaxed: $\plength(B) = x$, $\por(B) = x$, $\ptable(B) = x$ and $\psize(B) \geq x$.}
@@
 The actual constraints must enforce that a superblock assignment is \textit{compatible} with the requirements defined in the signature.
-
+%
 Table~\ref{tbl:translation} shows how the conversion from requirements of the signature to CSP constraints works:
@@
 Finally, restrictions on the number of rows or columns behave as length or size constraints based on the orientation of the block they are applied to.
-Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
+%Subgroups of row-oriented (column-oriented) blocks will always have the same number of columns (rows), however, the number of rows (columns) might decrease.
 
@@
 \begin{example}
-  Consider the sum-over-rows constraint template $\ecsumr{\sbl{r}}{\bsbl{x}}$.
-  An example superblock assignment from Figure~\ref{fig:main_example} is $(\sbl{r}, \bsbl{x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
+  Consider the sum-over-rows template $\ecsumr{\sbl{r}}{\bsbl{x}}$ again;
+  an example superblock assignment from Figure~\ref{fig:main_example} is $(\sbl{r}, \bsbl{x}) = (\range{T_1}{\rangeall}{\rangeto{3}{8}}, \range{T_1}{\rangeall}{\rangeto{3}{8}})$.
   Note that this assignment does not satisfy the signature yet because $\sbl{r}$ contains more than one vector.
@@
 Furthermore, in a CSP approach we would have to \textit{ground} out the definition for each of the corresponding elements in the vectors.
-This is inefficient, as already such blocks may not satisfy the signature or the first elements in the data may not satisfy the definition.
+This is inefficient, as such blocks may not even satisfy the signature or just the first element in the data may already not satisfy the definition.
 
@@
 For every (disjoint) subassignment the exact signature (e.g. subtypes will not have been checked yet) and definition will be tested and all satisfying subassignments are returned.
+
+\tias{todo me: check that more here}
 

### 2016-09-23 Paper/ECAI/tabular.tex
last changes in sec 4

@@
 \subsection{Optimizations} \label{sec:opts}
-This section discusses various design decisions and optimizations aimed at improving the speed, the extensibility of our system and the elimination of redundant constraints.
+This section discusses various design decisions and optimizations aimed at improving the extensibility and effectiveness of our method, as well as reducing the redundancy in the output.
 
@@
 
-Instead of generating one CSP to find all assignments, a CSP is generated for every known assignment of the depending constraint (and variables involved), which then searches for all assignments completing this partial assignment.
+In this case, instead of generating one CSP to find all assignments, a CSP is generated for every known assignment of the depending constraint (and variables involved), which then searches for all assignments completing this partial assignment.
 The procedure to generate CSPs remains the same otherwise.
@@
 Moreover, among the semantically equivalent constraints product (\ecprod{\sbl{R}'}{\sbl{1}'}{\sbl{2}'}) and division (\eccalc{\sbl{1}'}{\sbl{R}' / \sbl{2}'}) we only added product, as well as only difference (\ecdiff{\sbl{r}}{\sbl{1}}{\sbl{2}}) and not sum (\eccalc{\sbl{1}}{\sbl{r} + \sbl{2}}).
-However, division and sum could easily be added, as they could be added explicitly in post-processing based on the matching product and difference constraints.
-
-A last type of redundancy we chose to reduce is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
+However, division and sum could easily be added, and they could equivalently be added in post-processing based on the matching product and difference constraints.
+
+A last type of redundancy we chose to avoid during search is that there may be multiple \textit{overlapping} subblocks that satisfy the definition of a constraint template.
 For example, consider the constraint $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{n}}}$ where $\range{T}{\rangeall}{\rangeto{k}{n}}$ consists of only zeros, then $\ecsumc{\sbl{r}}{\range{T}{\rangeall}{\rangeto{1}{j}}}$ will be true for all $k - 1 \leq j \leq n$.
@@
 Consider superblock $\bsbl{x} = \range{T}{\rangeall}{\rangeto{1}{3}} = [[200, 300], [200, 150], [1, 2]]$ and $\sbl{r} = [200, 300]$, then \sname will find the constraint $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$.
-Should the target constraint be $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$, then $\range{T}{\rangeall}{3}$ must be split off into a separate block.
-
-Alternatively, we can output all valid assignments instead of only maximal subblocks, e.g.
+Should the target constraint be $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$, then \sname is only able to find exactly that constraint if $\range{T}{\rangeall}{3}$ is split off into a separate block in the block detection phase.
+
+%Alternatively, we can output all valid assignments instead of only maximal subblocks, e.g.
 $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{1})}$, $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{2}})}$ and $\eccalc{\sbl{r}}{\textit{MAX}(\range{T}{\rangeall}{\rangeto{1}{3}})}$, however, this might produce many redundant constraints including constraints that accidentally cover too little data.

### 2016-09-23 Paper/ECAI/tabular.tex

--- 2016-09-23 Paper/ECAI/tabular.tex
@@
   \vspace{-10pt}
-  \caption{Constraints learned for the tables above, except for \textit{ALLDIFFERENT} (18), \textit{PERMUTATION} (2), \textit{FOREIGNKEY} (5) and \textit{ORDERED} (6) constraints.
+  \caption{Constraints learned for the tables above, except 18 \textit{ALLDIFFERENT}, 2 \textit{PERMUTATION} and 5 \textit{FOREIGNKEY} and 6 \textit{ORDERED} constraints not shown.
   Constraints marked with $*$ were not present in the original spreadsheets.}
@@
 \item For each constraint template~$s$, generate all assignments $(\sbl{1}, \ldots ,\sbl{n})$ where each~$\sbl{i}$ is a superblock and the~$\sbl{i}$ are \textit{compatible} (defined below) with the signature of~$s$.
-\item For each assignment $(\sbl{1}, \ldots ,\sbl{n})$ generated in 3a), find all constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that for all $i$ holds $\ssbl{i} \sqsubseteq \sbl{i}$ and the signature and definition of~$s$ are satisfied.
+\item For each generated assignment $(\sbl{1}, \ldots ,\sbl{n})$, find all constraints $s(\ssbl{1}, \ldots, \ssbl{n})$ such that $\ssbl{i} \sqsubseteq \sbl{i}$ holds for all $i$ and the signature and definition of~$s$ are satisfied.
 \end{enumerate}

### 2016-09-23 Paper/ECAI/tabular.tex
more of sec4 and some of exp section

@@
 
-\tias{todo me: check that more here}
+Based on the assumptions that most subassignments will not satisfy the definition of the constraint template, the implementation of the definition check is geared to \textit{fail-fast} when possible.
+Specifically, the entries of every subblock are fetched one by one, and as soon as one does not hold failure is returned.
+In many cases, the first entry will already not satisfy the definition (e.g.
+if $\sbl{r}[0] \neq \sbl{1}[0] \times \sbl{2}[0]$ then $\sbl{r} = \sbl{1} \times \sbl{2}$ does not hold).
+As a result, many subassignments can be discared by looking at just one or a few entries, and considering all subassignments that will be checked, the runtime of the checks will typically not be influenced much by the length of the vectors.
+\tias{Moved the discussion from 4.4.3 to here, where it better fits. Plus, I find the O(1) claim too strong! O() is worst-case time complexity, not average case}
+%the time to check if a subassignment satisfies a constraint definition with a fail-fast implementation requires constant time\footnote{If the number of entries to look at is bounded by a constant, the vector lengths are smaller than the number of subassignments to test ($a$) and the time to test subassignments that satisfy the constraint is $O(a)$, then the time to test a subassignment is amortized $O(1)$} in most cases and does not depend on the length of the vectors.
 
@@
 We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
-We also added three structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
-
-For \textit{FOREIGNKEY}, \textit{LOOKUP}, aggregates and \textit{PROJECT} we use custom, optimized methods to find subassignments.
-Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
-While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
-
-Based on the assumptions that most subassignments will not satisfy a constraint template, the implementations are geared to \textit{fail-fast} when possible.
-Specifically, the tests try to prove that a subassignment does not satisfy a constraint template as fast as possible.
-For example, it can be often shown that a constraint does not hold by only looking first entries of every subblock (e.g.
-if $\sbl{r}[0] \neq \sbl{1}[0] \times \sbl{2}[0]$ then $\sbl{r} = \sbl{1} \times \sbl{2}$ does not hold).
-As a result, the time to check if a subassignment satisfies a constraint templates with a fail-fast implementation requires constant time\footnote{If the number of entries to look at is bounded by a constant, the vector lengths are smaller than the number of subassignments to test ($a$) and the time to test subassignments that satisfy the constraint is $O(a)$, then the time to test a subassignment is amortized $O(1)$} in most cases and does not depend on the length of the vectors.
+We also added four structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION}, \textit{ORDER} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
+
+\tias{Just below this comment I commented out the paragraph that FK,Lookup,proj have optimized methods becuase it does not properly explain what is used instead.
+That creates confusion and makes our method not reproducible.
+Leave out until it is rewritten with sufficient information.}
+%As an optimization, for \textit{FOREIGNKEY}, \textit{LOOKUP}, aggregates and \textit{PROJECT} we use custom methods to find subassignments.
+%Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
+\tias{Really, this is unacceptably vague.
+I have no clue what you mean with that, meaning that our research is not reproducable and hence not scientific}
+%While conditional aggregates use generate-and-test, they employ similar optimizations in their test method.
 
@@
 \label{sec:evalualtion:method}
-All spreadsheets used in the experiments are in CSV format, and for each we have ran the table extraction tool once (using a combination of the rather primitive automatic detection + visual verification that the tables are meaningful (e.g.
+All spreadsheets used in the experiments are in CSV format, and for each we have ran the table extraction tool once using the visual tool.
+Manual intervention of the simple automatic detection was needed only in the case of None values (to merge split-up tables cause by the None's) and to remove all-textual headers of all-textual tables.%(using a combination of the rather primitive automatic detection + visual verification that the tables are meaningful (e.g.
 merge split-up tables because of None values, or remove headers of all-textual tables).
-\samuel{Should we not say just with the visual tool?}
-Blocks are detected automatically using our block detection algorithm (Section~\ref{sec:make_groups}) except for spreadsheets that contain \textit{None} values or (partially) numeric headers, in which case the blocks are specified manually.
+%\samuel{Should we not say just with the visual tool?} \tias{Rewrote a bit, I like the amount of detail as it provides insight into the limitations}
+Blocks are then detected automatically using the block detection algorithm (Section~\ref{sec:make_groups}).
+Manual intervention was required for tables that contain \textit{None} values, as well as for (partially) numeric headers.
+\tias{This does not make sense to me: headers are removed before block detection; should this be part of manual intervention of table extraction tool??}
 
@@
 5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors.
-\samuel{They are usually not copied}
+\samuel{They are usually not copied} \tias{'usually' is not acceptable in a scientific paper.
+Why would you sometimes do it and sometimes not?
+Preferably always or never copy, and if not specify when and why exactly you deviate.}
 
@@
 All intended constraints are supported.
-The recall is as good as we hoped, but the precision is rather low.
+The recall is perfect, but the precision is rather low.
 This is mostly due to the spurious RANK constraints.
@@
 
-In this section the impact on the running time is analyzed of 1)~the number of vectors; 2)~the length of the vectors; and 3)~the size of the blocks.
+In this section we investigate the factors that influence the runtime of our method the most; we do this by investigating the impact on the running time of 1)~the number of vectors; 2)~the length of the vectors; and 3)~the size of the blocks.
 This analysis is accomplished by testing our system on generated spreadsheets.
 
-Recall that our method uses a 2-phase approach where in the first phase, the superblock assignments are generated (Section~\ref{sec:algo:super} and in the second phase, the subblock assignments are generated (Section~\ref{sec:algo:subgr}).
-
-\paragraph{Spreadsheet generation}
-For our analysis, we perform multiple experiments on synthetically generated spreadsheets.
-
-Many constraints require either numeric or discrete data, therefore, the type of cells also has an impact on the running time.
-Since integers fulfill both of those typing restrictions, the generated spreadsheets contain only integer data.
-
-One of the assumptions guiding the implementation is that most of the possible constraints are false.
-To simulate this setting, the generated spreadsheets contain random data.
-There will be few constraints that hold on these spreadsheets, however, random data is likely to produce vectors that fulfill the \textit{ALLDIFFERENT} constraint, which is a requirement for many other constraint templates.
-
-In order to also analyse constraints that require their arguments to be picked from different tables, the spreadsheets are made up of two tables.
-For one table the number of vectors and block size is fixed to two, while for the other table these properties are varied.
-The number of vectors and the block size in the experiments are only reported for the adaptable table.
-The length of the vectors can also be adapted and applies to both tables.
-Unless otherwise mentioned, the adaptable table consists of one block containing all its vectors and by default the number of vectors and vector length are both~$8$.
+\paragraph{Spreadsheet generation.}
+%For our analysis, we perform multiple experiments on synthetically generated spreadsheets.
+The generated sheets should trigger the checking of many different constraints, both in the first phase (superblock assignment) and in the second phase (subblock assignment), in order to clearly study the effect on the constraint checking.
+As many constraints require numeric or discrete data, we will use integers in the generated sheets as these are both numeric and discrete.
+
+We generate spreadsheets with random data.
+The result is that there are no intended constraints in the data.
+This is reasonable, as even in real data most constraints that are checked will not hold.
+Additionally, for vectors in random data the \textit{ALLDIFFERENT} constraint is likely to hold, which is a requirement for many other constraint templates.
+
+The generates spreadsheets contain two tables, as a number of constraints require that their arguments are picked from different tables.
+For one table the number of vectors and block size is fixed to two, while for the other table these properties are varied according to the experimental setup.
+The reported number of vectors and the block size in the experiments refer to those of the second table.
+The length of the vectors can also be adapted and applies to both tables (this is a required property of many constraints).
+Unless otherwise mentioned, the adaptable table consists of one block of $8$ vectors of length $8$ (and hence 2 vectors of length $8$ in the 'fixed' table).
 %The running times for various configurations are plotted in Figure~\ref{fig:runtime_analysis}. % no such general pointer, specific pointers only
+
+%Recall that our method uses a 2-phase approach where in the first phase, the superblock assignments are generated (Section~\ref{sec:algo:super} and in the second phase, the subblock assignments are generated (Section~\ref{sec:algo:subgr}).
 
@@
 
-This behavior is to be expected when not many superblock assignments can be discarded in phase~1, for example, when all vectors have the same type and size.
-In this case, for a constraint template that requires $n$~single vector arguments and a spreadsheet with $v$~vectors, up to $\frac{v!}{(v - n)!}$ subassignments are generated throughout the algorithm.
+This behavior is to be expected when few superblock assignments can be discarded in the first phase, for example, when all vectors have the same type and size as we artificially enforced in this experiment.
+In this case, for a constraint template that requires $n$~single vector arguments and a spreadsheet with $v$~vectors, up to $\frac{v!}{(v - n)!}$ subassignments are generated and checked by the algorithm.
 
 Typically, we expect the number of generated subassignments to be much smaller in real world spreadsheets and, therefore, the running time to be less sensitive to the number of vectors.
-This is because in phase~1 entire blocks can be disqualified that do not match the signature requirements, e.g. blocks of vectors with different lengths for the product constraint.
+This is because in the first phase entire blocks can be disqualified that do not match the signature requirements, e.g. blocks of vectors with different lengths. % for the product constraint.
 % \tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}\samuel{?}
@@
 Most of the increase in running time can be attributed to the \textit{RANK} constraint.
-The implementation is not always able to detect subassignments that do not satisfy \textit{RANK} quickly and often has to fall back to analyzing an entire vector.
+The implementation is often not able to quickly detect that subassignments do not satisfy \textit{RANK}, meaning that it will have to analyze many if not all entries in the vector.
 
 \paragraph{Effect of block size}
-To measure the effect of changing the block size, the number of vectors is set to~$32$ and the vectors are split into varying number of blocks (up to ~$32$).
+To measure the effect of changing the block size, the number of vectors is set to~$32$ and the vectors are explicitly split into varying number of blocks/tables (up to ~$32$).
 Figure~\ref{fig:runtime_analysis}c shows the running times of \sname for various settings.
@@
 
-Aggregate constraints, e.g. sum or max, have an argument that takes subblocks of varying sizes.
-Contrary to non aggregate constraints, which only have single vector argument, superblock assignments from phase~1 are not split into single vectors subassignments in phase~2.
-Instead, for an argument with variable size, the system tries to find any subblock of the given superblock that satisfies the constraint.
-Since, for a superblock of size~$m$, the number of subblocks is $O(m^2)$ \samuel{ref?}, the search for aggregate constraints will be faster if there are more but smaller blocks\footnote{Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$}.
+Aggregate constraints, e.g.
+sum or max, have an argument that allows subblocks of varying sizes while non-aggregates only allow single vector arguments.
+Hence, for aggregate constraints our method will try to find any subblock in the given superblock that satisfies the constraint.
+Since a superblock of size~$m$ has $m(m+1)/2$ contiguous subblocks, the search for aggregate constraints becomes faster if there are more but smaller blocks\footnote{Consider, for example, $v$~vectors distributed over $b$~blocks of size $v/b$, the number of subblock candidates then is $O(b * v^2 / b^2) = O(v^2 / b)$}.
 
 Non-aggregate constraints have only single vector arguments.
-Therefore, the total number of subassignments generated by the algorithm is not influenced by the block size.
-By splitting a set of vectors into many blocks, however, there will be a larger number of superblocks in phase~1 and, therefore, a larger number of superblock assignments have to be generated.
-Recall that in phase~1 the algorithm additionally tests block level properties for every superblock assignment.
-In the generated spreadsheets these tests can only disqualify few or no superblock assignments, therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.
-
-Splitting unrelated vectors into smaller subblocks can decrease the running time for some spreadsheets by reducing the workload for aggregate constraints.
-However, the two phase approach is also well suited to operate on large blocks and eliminate superblock assignments that are not compatible with constraint template signatures.
+%Therefore, the total number of subassignments generated by the algorithm is not influenced by the block size.
+By splitting a set of vectors into many blocks, however, there will be a larger number of superblocks given as input to the first phase. 
+%Recall that in phase~1 the algorithm additionally tests block level properties for every superblock assignment.
+In the generated spreadsheets, the first phase, which looks at properties of superblocks alone, can disqualify few or no superblock assignments, leading to some increase in runtime as more blocks are added.
+%therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.
+
+%\tias{I'm cutting this paragraph, in an unsupervised method you do not control the number of blocks and will hence not 'split unrelated vectors'; that would be supervised.}
+%Splitting unrelated vectors into smaller subblocks can decrease the running time for some spreadsheets by reducing the workload for aggregate constraints.
+%However, the two phase approach is also well suited to operate on large blocks and eliminate superblock assignments that are not compatible with constraint template signatures.
 \samuel{Does this conclusion sound sound?}
@@
 What do you really want to share with the reader?}
+
+$ $
+
+\noindent 
+These experiments with generated data have given us a better understanding of what effect the properties of vectors and blocks can have on runtime and hence on the ability of the algorithm to finish within reasonable time.
+We next investigate the behaviour of our method on real spreadsheets.
+
+\tias{I'm up to this point}
 

### 2016-09-24 Paper/ECAI/tabular.tex
slight retouch and some more comments by Tias

@@
 
-\tias{I'm up to this point}
-
 \subsection{Effectiveness on real spreadsheets}
@@
 In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
+\tias{Add something like 'on data generated to trigger many checks (equal sizes and lengths of many vectors)??
+You choose}
 Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets.
@@
  % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41
+\tias{Very nice!!!}
 
@@
 Therefore, \sname is able to find all constraints for supported constraint templates, which confirms that our bias (signature) is not too restrictive.
+\tias{'confirms'?
+We never write anywhere that we doubt this (and if we did I probably removed it : ///) Alternative could be something like "shows that our method is able to learn all constraints using the bias chosen (the formalism of signature and definition)"?}
 
@@
 conditional aggregates that automatically infer the condition).
+\tias{if you can clarify the latter, please do; if too difficult OK but 'automatically infer' sounds..
+unusual for Excel}
 
@@
 %\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
-The final question, Q3, is about the amount and type of non-intended constraints found.
+The final question, Q3, is about the amount and type of \textbf{non}-intended constraints found.
 Across all spreadsheets \sname achieves a precision of~$0.79~(81/103)$.
-While the average precision per spreadsheet is high, over $0.90$ in every category, however the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).
+While the average precision per spreadsheet is high, over $0.90$ in every category, we observe that the overall precision is much higher for the exercises and data categories ($0.94$ and $1.00$) than for the tutorials category ($0.70$).
 
 Examining the tutorial spreadsheets confirms that there are a few spreadsheets that have a high number of additional constraints, e.g.
-an spreadsheet computes aggregates on its inventory but copies the data column for every aggregate.
+one spreadsheet computes aggregates on inventory data, but copies the data column for every aggregate.
 Since every aggregate can be calculated based on any of the (equal) columns, many of the resulting constraints will be marked as unintended.
-Usually, unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data.
-A way to increase the precision is by post-processing the constraints to detect implications and heuristically selecting one among multiple ways to compute the same vector.
+\tias{rewrite 'many of the..' into 'our method will find all of these; only one in 'true' set, others drive down precision...'}
+
+Such unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data.
+One way to increase precision for this case is to post-process the constraints to detect equivalences and implications, and heuristically select one among multiple ways to compute the same vector.
 
@@
 \section{Applications}\label{sec:applications}
-In this section, we illustrate how various motivating applications could be realized using our system.
+In this section, we illustrate how various applications can build on our method for solving the tabular constraint learning problem.
 
@@
 Consider Figure \ref{fig:autocompletion_example} containing two tables.
-Learning on a snapshot that does not contain the row below $T_2$, we find constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
+Learning on a snapshot that does not contain the row below $T_2$, our method finds constraints $\ecsumif{\range{T_2}{\rangeall}{2}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{3}}$ and $\ecsumif{\range{T_2}{\rangeall}{3}}{\range{T_1}{\rangeall}{2}}{\range{T_2}{\rangeall}{1}}{\range{T_1}{\rangeall}{4}}$.
 The new data $\Delta D$ consists of the new value \textit{West} in the row below $T_2$.
@@
 The first setting is the \textit{online} detection of errors and is similar to the autocompletion setting.
-Using snapshots, we search for \textit{conflict} constraints, which are learned on a previous snapshot but do not agree with newly added data~$\Delta D$.
+Using snapshots, one can learn the constraints in the previous snapshot and search for \textit{conflict} constraints, which do not agree with newly added data~$\Delta D$.
 Such conflicts can indicate (typing) errors in the data just entered. For example, in case 'Wes' instead of 'West' was added in Figure~\ref{fig:autocompletion_example}.
@@
 Many schemes can be devised for choosing which row(s) or column(s) are removed.
+\tias{Expand 'one can repeatedly search' into the actual steps: take spreadsheet, remove row/column, detect constraints and then search for any constraints not present in the full data (can easily be optimized to avoid known constraints btw)}
 %This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g.
@@
 We have presented and evaluated our approach, implemented as the system \sname, that is able to learn many different constraints in tabular data.
-The resulting method has high recall, produces limited number of redundant constraints and is sufficiently efficient for normal and interactive use.
+The resulting method has high recall, produces a limited number of redundant constraints and is sufficiently efficient for normal and interactive use.
+\tias{Smth about extendible: more constraints can be added by specifying signature in terms of properties and implementing the subblock finding (or in less detail, can be added, e.g.
+all native Excel constraints...
+should this be true)}
 

### 2016-09-25 Paper/ECAI/tabular.tex
Updated and reworked comments

@@
 Based on the assumptions that most subassignments will not satisfy the definition of the constraint template, the implementation of the definition check is geared to \textit{fail-fast} when possible.
-Specifically, the entries of every subblock are fetched one by one, and as soon as one does not hold failure is returned.
-In many cases, the first entry will already not satisfy the definition (e.g.
-if $\sbl{r}[0] \neq \sbl{1}[0] \times \sbl{2}[0]$ then $\sbl{r} = \sbl{1} \times \sbl{2}$ does not hold).
+Specifically, for many constraints the entries of every subblock are fetched one by one, and as soon as one does not hold failure is returned.
+Often, the first entry will already not satisfy the definition (e.g. if $\sbl{r}[0] \neq \sbl{1}[0] \times \sbl{2}[0]$ then $\sbl{r} = \sbl{1} \times \sbl{2}$ does not hold).
 As a result, many subassignments can be discared by looking at just one or a few entries, and considering all subassignments that will be checked, the runtime of the checks will typically not be influenced much by the length of the vectors.
-\tias{Moved the discussion from 4.4.3 to here, where it better fits. Plus, I find the O(1) claim too strong! O() is worst-case time complexity, not average case}
+%tias{Moved the discussion from 4.4.3 to here, where it better fits. Plus, I find the O(1) claim too strong! O() is worst-case time complexity, not average case}
 %the time to check if a subassignment satisfies a constraint definition with a fail-fast implementation requires constant time\footnote{If the number of entries to look at is bounded by a constant, the vector lengths are smaller than the number of subassignments to test ($a$) and the time to test subassignments that satisfy the constraint is $O(a)$, then the time to test a subassignment is amortized $O(1)$} in most cases and does not depend on the length of the vectors.
@@
 We included most formulas that we encountered in tutorial spreadsheets including the popular \textit{SUM} and \textit{LOOKUP} constraints\footnote{\href{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}{https://support.office.com/en-us/article/Excel-functions-by-category-5F91F4E9-7B42-46D2-9BD1-63F26A86C0EB}}.
-We also added four structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION}, \textit{ORDER} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
-
-\tias{Just below this comment I commented out the paragraph that FK,Lookup,proj have optimized methods becuase it does not properly explain what is used instead.
+We also added four structural constraints (\textit{ALLDIFFERENT}, \textit{PERMUTATION}, \textit{ORDERED} and \textit{FOREIGNKEY}) so that they can be used in the signature of other constraint templates; they are also popular in constraint satisfaction~\cite{modelseeker}.
+
+For \textit{LOOKUP}, aggregates and \textit{PROJECT} we use optimized implementations to find subassignments instead of generic generate-and-test.
+
+Given a superblock assignment, the implementation for \textit{LOOKUP} first generates candidate vectors for \sbl{pk} and \sbl{val}, then generates a datastructure that allows it to quickly look up keys, afterwards it generates candidate vectors for the remaining arguments and, finally, checks which subassignments are valid.
+
+For aggregates and project, a custom implementation is used that tries to find maximal subblocks that satisfy the constraint templates.
+Instead of generating all possible subblocks, given a superblock assignment, these implementations start from a superblock and repeatedly generate and test smaller subblocks.
+Therefore, once a match is found, all smaller subblocks do not need to be considered anymore.
+
+Finally, for \textit{FOREIGNKEY} and conditional aggregates, generate-and-test is used, but in combination with caches that store intermediate results and a preparation step to populate some of the caches.
+For example, for conditional aggregates, superblock assignments are analyzed to find overlapping vectors.
+These results are then cached and used to quickly reject subassignments where there is no overlap between \sbl{pk} and \sbl{fk}.
+
+%tias{Just below this comment I commented out the paragraph that FK,Lookup,proj have optimized methods becuase it does not properly explain what is used instead.
 That creates confusion and makes our method not reproducible.
@@
 \label{sec:evalualtion:method}
-All spreadsheets used in the experiments are in CSV format, and for each we have ran the table extraction tool once using the visual tool.
-Manual intervention of the simple automatic detection was needed only in the case of None values (to merge split-up tables cause by the None's) and to remove all-textual headers of all-textual tables.%(using a combination of the rather primitive automatic detection + visual verification that the tables are meaningful (e.g.
+All spreadsheets used in the experiments are in CSV format and for each spreadsheet the tables are specified.
+The tables were obtained by running the table extraction tool (Section~\ref{sec:table_extraction}).
+Manual intervention of the simple automatic detection was needed only in the case of \textit{None} values (e.g.
+to merge split-up tables cause by the \textit{None}'s) and in cases where the headers are ambiguous (e.g.
+headers containing textual and numeric cells).
+In these cases the tables specifications were generated using our visual tool.
+
+%(using a combination of the rather primitive automatic detection + visual verification that the tables are meaningful (e.g.
 merge split-up tables because of None values, or remove headers of all-textual tables).
-%\samuel{Should we not say just with the visual tool?} \tias{Rewrote a bit, I like the amount of detail as it provides insight into the limitations}
+%{Should we not say just with the visual tool?} {Rewrote a bit, I like the amount of detail as it provides insight into the limitations}
 Blocks are then detected automatically using the block detection algorithm (Section~\ref{sec:make_groups}).
-Manual intervention was required for tables that contain \textit{None} values, as well as for (partially) numeric headers.
-\tias{This does not make sense to me: headers are removed before block detection; should this be part of manual intervention of table extraction tool??}
+Manual intervention was required for tables that contain \textit{None} values (e.g. if the type of empty cells cannot be inferred).
 
 % Group definitions are only specified if the spreadsheet contains ambiguous \textit{None} values.
-\tias{What does this mean, 'group definitons'??
+\{What does this mean, 'group definitons'??
 Is this 'block detection' (Sec 4.2), if so, how do we explain this process??}
@@
 In the following experiments, we focus on spreadsheets specifically and hence only include \textit{functional} constraints, constraints that can be expressed as formulas.
-We ignoring structural constraints such as \textit{ALLDIFFERENT}, \textit{FOREIGNKEY}, \textit{PERMUTATION} and \textit{ORDERED}.
+Therefore, we ignore the structural constraints \textit{ALLDIFFERENT}, \textit{FOREIGNKEY}, \textit{PERMUTATION} and \textit{ORDERED}.
 All constraints are stored in their canonical form.
-5 spreadsheets contained rows or columns that were entirely identical, and any intended constraint involving one such vector was also added for the equivalent vectors.
-\samuel{They are usually not copied} \tias{'usually' is not acceptable in a scientific paper.
+5 spreadsheets contained rows or columns that were entirely identical in which case results can be computed in multiple ways and there are multiple valid constraints.
+Therefore, the original formulas or the headers and positioning of the cells is used to infer the intended constraint.
+%and any intended constraint involving one such vector was also added for the equivalent vectors.
+\{They are usually not copied} \{'usually' is not acceptable in a scientific paper.
 Why would you sometimes do it and sometimes not?
@@
 \begin{itemize}
-\item Q1: it may fail to find intended constraints if it takes too much time to find them; what are the factors that influence the runtime of the method most?
-\item Q2: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
-\item Q3: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
+\item \textit{Q1}: it may fail to find intended constraints if it takes too much time to find them; what are the factors that influence the runtime of the method most?
+\item \textit{Q2}: it may fail to find intended constraints because it does not support such constraints; how does or method perform on real spreadsheets?
+\item \textit{Q3}: it may find non-intended constraints; how many and what type of non-intended constraints are found on real spreadsheets?
 \end{itemize}
-We first investigate Q1 on generated spreadsheets, and Q2 and Q3 on a collection of real spreadsheets.
+We first investigate \textit{Q1} on generated spreadsheets and investigate \textit{Q1}, \textit{Q2} and \textit{Q3} on a collection of real spreadsheets.
 
@@
 \textbf{Precision} & 0.94 & 0.96 & 0.70 & 0.91 & 1.00 & 1.00 \\ \hline
-\textbf{Runtime (s)} & 0.93 & 0.10 & 0.93 & 0.04 & 0.79 & 0.20
+\textbf{Runtime (s)} & 0.91 & 0.10 & 0.87 & 0.04 & 0.79 & 0.20
 \end{tabular}
@@
 Across categories the average running times fluctuate in the range between $0.04s$ to $0.20s$.
-While most spreadsheets can be processed very quickly, running times can go up to around $0.65s$ for others.
-In the previous section we analyzed the effect of various factors that can help explain runtime behavior.
-\tias{Add something like 'on data generated to trigger many checks (equal sizes and lengths of many vectors)??
-You choose}
-Our benchmark experiments show that our algorithm is able to perform well in practice, even for larger spreadsheets.
+While most spreadsheets can be processed very quickly, running times can go up to around $0.61s$ for others.
+In the previous section we analyzed the effect of various factors that can help explain runtime behavior on data generated to allow many candidate subassignments (e.g.
+using only vectors of the same type and length).
+Our benchmark experiments show that our algorithm is able to perform well on real spreadsheets, even for larger numbers of vectors.
 
@@
  % (0.83 + 1.03 + 0.99) = 2.85 / (1.34 + 1.46 + 0.46) = 3.3  / (0.99 + 0.98 + 0.44) = 2.41
-\tias{Very nice!!!}
+%tias{Very nice!!!}
 
@@
 The intended constraints not discovered correspond to constraints that are not supported by the system (supported recall is 1).
-Therefore, \sname is able to find all constraints for supported constraint templates, which confirms that our bias (signature) is not too restrictive.
-\tias{'confirms'?
+Therefore, \sname is able to find all constraints for supported constraint templates, which shows that our method is able to learn all constraints using the bias chosen (the formalism of signature and definition).
+%tias{'confirms'?
 We never write anywhere that we doubt this (and if we did I probably removed it : ///) Alternative could be something like "shows that our method is able to learn all constraints using the bias chosen (the formalism of signature and definition)"?}
@@
 series that do not start at 1) or more complex versions of supported constraints (e.g.
-conditional aggregates that automatically infer the condition).
-\tias{if you can clarify the latter, please do; if too difficult OK but 'automatically infer' sounds..
+conditional aggregates whose conditions are constant values).
+%tias{if you can clarify the latter, please do; if too difficult OK but 'automatically infer' sounds..
 unusual for Excel}
@@
 one spreadsheet computes aggregates on inventory data, but copies the data column for every aggregate.
-Since every aggregate can be calculated based on any of the (equal) columns, many of the resulting constraints will be marked as unintended.
-\tias{rewrite 'many of the..' into 'our method will find all of these; only one in 'true' set, others drive down precision...'}
-
-Such unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data.
+Since every aggregate can be calculated based on any of the (equal) columns, our method will find all of these constraints.
+However, for every aggregate only one constraint is \textit{intended}, the others will be considered as redundant and decrease the precision.
+
+%many of the resulting constraints will be marked as unintended. tias{rewrite 'many of the..' into 'our method will find all of these; only one in 'true' set, others drive down precision...'}
+
+Such redundant or unintended constraints are the result of multiple ways to calculate the same result, e.g. if there is duplicate data as in the example above.
 One way to increase precision for this case is to post-process the constraints to detect equivalences and implications, and heuristically select one among multiple ways to compute the same vector.
@@
 Using snapshots, one can learn the constraints in the previous snapshot and search for \textit{conflict} constraints, which do not agree with newly added data~$\Delta D$.
-Such conflicts can indicate (typing) errors in the data just entered. For example, in case 'Wes' instead of 'West' was added in Figure~\ref{fig:autocompletion_example}.
+Such conflicts can indicate (typing) errors in the data just entered.
+For example, in case $1572$ was added in Figure~\ref{fig:autocompletion_example} for \textit{Units per region} in the newly added row (\textit{West}).
+
 %Conflict constraints can indicate errors in the input.
@@
 For example, in Figure \ref{fig:fbi} all but one row satisfy the constraint, which is a strong indication that there is a constraint violation.
-Instead of snapshots, one can repeatedly search for conflict constraints between the original spreadsheet and a sheet with a row or column removed.
+Instead of snapshots, one can repeatedly remove one or multiple rows or columns from the spreadsheet, detect constraints ($S$) that hold in the modified data and search for constraints in~$S$ that do not hold on the original spreadsheet (conflict constraints).
 Many schemes can be devised for choosing which row(s) or column(s) are removed.
-\tias{Expand 'one can repeatedly search' into the actual steps: take spreadsheet, remove row/column, detect constraints and then search for any constraints not present in the full data (can easily be optimized to avoid known constraints btw)}
+
+% between the original spreadsheet and a sheet with a row or column removed.
+Many schemes can be devised for choosing which row(s) or column(s) are removed.
+tias{Expand 'one can repeatedly search' into the actual steps: take spreadsheet, remove row/column, detect constraints and then search for any constraints not present in the full data (can easily be optimized to avoid known constraints btw)}
 %This task can be formulated as online error detection by repeatedly partitioning the spreadsheet into a snapshot~$D_i$ and ``input'' $\Delta D_i$ (e.g.
@@
 The resulting method has high recall, produces a limited number of redundant constraints and is sufficiently efficient for normal and interactive use.
-\tias{Smth about extendible: more constraints can be added by specifying signature in terms of properties and implementing the subblock finding (or in less detail, can be added, e.g.
-all native Excel constraints...
-should this be true)}
+Our approach also allows new constraint templates to be easily added by specifying their signature in terms of properties and providing an implementation that finds subassignments satisfying the definition.
 

### 2016-09-25 Paper/ECAI/tabular.tex
Small changes

@@
 
-Given a superblock assignment, the implementation for \textit{LOOKUP} first generates candidate vectors for \sbl{pk} and \sbl{val}, then generates a datastructure that allows it to quickly look up keys, afterwards it generates candidate vectors for the remaining arguments and, finally, checks which subassignments are valid.
+Given a superblock assignment, the implementation for \eclookup{\sbl{r}}{\sbl{fk}}{\sbl{pk}}{\sbl{val}} first generates candidate vectors for \sbl{pk} and \sbl{val}, then populates a hash table that allows it to quickly look up keys, afterwards it generates candidate vectors for the remaining arguments and, finally, checks which subassignments are valid.
 
@@
 Finally, for \textit{FOREIGNKEY} and conditional aggregates, generate-and-test is used, but in combination with caches that store intermediate results and a preparation step to populate some of the caches.
-For example, for conditional aggregates, superblock assignments are analyzed to find overlapping vectors.
+For example, for conditional aggregates, superblock assignments are analyzed to find vectors with overlapping values.
 These results are then cached and used to quickly reject subassignments where there is no overlap between \sbl{pk} and \sbl{fk}.
@@
 This is because in the first phase entire blocks can be disqualified that do not match the signature requirements, e.g. blocks of vectors with different lengths. % for the product constraint.
-% \tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}\samuel{?}
+% tias{Better move to 4.3.2? or have there too anyway, no new facts in 'experiment' section}samuel{?}
 
@@
 
-% \tias{unclear, this is about phase 2, the number to test, or the actual testing of each?
+% tias{unclear, this is about phase 2, the number to test, or the actual testing of each?
 explain and relate to Sec 4.3.2 what aspect you are discussing here} to be tested depends on the number of vectors in the spreadsheets and the arity of the constraint.
@@
 % For a constraint template with arity~$n$ and $v$~vectors there are $n^v$ subassignments (or $\frac{n!}{(n - v)!}$ if the arguments have to be different).
-\tias{Better move to 4.3.2?
+tias{Better move to 4.3.2?
 or have there too anyway, no new facts in 'experiment' section}
@@
 In the generated spreadsheets, the first phase, which looks at properties of superblocks alone, can disqualify few or no superblock assignments, leading to some increase in runtime as more blocks are added.
-%therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.
-
-%\tias{I'm cutting this paragraph, in an unsupervised method you do not control the number of blocks and will hence not 'split unrelated vectors'; that would be supervised.}
+%therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.\sam{Not sure}
+
+%tias{I'm cutting this paragraph, in an unsupervised method you do not control the number of blocks and will hence not 'split unrelated vectors'; that would be supervised.}
 %Splitting unrelated vectors into smaller subblocks can decrease the running time for some spreadsheets by reducing the workload for aggregate constraints.
 %However, the two phase approach is also well suited to operate on large blocks and eliminate superblock assignments that are not compatible with constraint template signatures.
-\samuel{Does this conclusion sound sound?}
-
-% \tias{Point to figure to support this (actually, move that line into 4c??); also, if this is true, you are basically saying that the CP-based first phase is worse then the second phase?
+samuel{Does this conclusion sound sound?}
+
+% tias{Point to figure to support this (actually, move that line into 4c??); also, if this is true, you are basically saying that the CP-based first phase is worse then the second phase?
 That kind-of demotivates the idea of using CP in first phase...
@@
 % Notice that for the data category the average number of cells is higher while the number of intended constraints and tables is lower compared to the other categories.
-\tias{Weird here because purpose unclear, point to this when discussing some result for which this is important}
+tias{Weird here because purpose unclear, point to this when discussing some result for which this is important}
 

### 2016-09-25 Paper/ECAI/tabular.tex
Removed comments
Cosmetic fix
minor changes

@@
 \author{Samuel Kolb* \and Sergey Paramonov* \and Tias Guns \and Luc De Raedt}
-\institute{KULeuven\\ {*} Shared first author}
+\institute{KU Leuven\\ {*} Shared first author}
 
@@
 Often, the first entry will already not satisfy the definition (e.g. if $\sbl{r}[0] \neq \sbl{1}[0] \times \sbl{2}[0]$ then $\sbl{r} = \sbl{1} \times \sbl{2}$ does not hold).
-As a result, many subassignments can be discared by looking at just one or a few entries, and considering all subassignments that will be checked, the runtime of the checks will typically not be influenced much by the length of the vectors.
+As a result, many subassignments can be discarded by looking at just one or a few entries, and considering all subassignments that will be checked, the runtime of the checks will typically not be influenced much by the length of the vectors.
 %tias{Moved the discussion from 4.4.3 to here, where it better fits. Plus, I find the O(1) claim too strong! O() is worst-case time complexity, not average case}
@@
 \paragraph{Effect of number of vectors}
-Figure~\ref{fig:runtime_analysis}a shows the results of our experiments, in which \sname was run on generated spreadsheets containing an increasing number of (equally sized) vectors.
+Figure~\ref{fig:runtime_analysis} (a) shows the results of our experiments, in which \sname was run on generated spreadsheets containing an increasing number of (equally sized) vectors.
 As shown in the figure, the number of vectors in a spreadsheet can have a significant impact on the running time.
@@
 To measure the effect that the length of the vectors has on the running time, we ran our system on spreadsheets in which the length of the vectors was gradually increased.
-Figure~\ref{fig:runtime_analysis}b summarizes our results and shows that the length of the vectors has a small impact on the performance for most templates.
+Figure~\ref{fig:runtime_analysis} (b) summarizes our results and shows that the length of the vectors has a small impact on the performance for most templates.
 
@@
 To measure the effect of changing the block size, the number of vectors is set to~$32$ and the vectors are explicitly split into varying number of blocks/tables (up to ~$32$).
-Figure~\ref{fig:runtime_analysis}c shows the running times of \sname for various settings.
+Figure~\ref{fig:runtime_analysis} (c) shows the running times of \sname for various settings.
 As the blocks become smaller, the total running time first decreases before increasing again.

### 2016-11-23 Paper/response_letter/response_letter.tex
added response letter

\documentclass{article}
\usepackage{enumerate}
\usepackage[english]{babel}
\usepackage{times}
\usepackage{latexsym}    % to get LASY symbols
\usepackage{graphicx}    % to insert PostScript figures
\usepackage{xspace}
\usepackage[usenames,dvipsnames]{xcolor}

\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
\newcommand{\sam}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
\newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
\newcommand{\tias}[1]{\textcolor{blue}{{\sc Tias:} #1}\xspace}
\newcommand{\luc}[1]{\textcolor{red}{{\sc Luc:} #1}\xspace}
\newcommand{\added}[1]{\textcolor{red}{added:} {#1} \textcolor{red}{end}\xspace}

\newcommand{\scaleconstant}{1.0}
\bibliographystyle{plainnat}
\author{Samuel Kolb \and Sergey Paramonov \and Tias Guns \and Luc {De Raedt}}
\title{MLJ Revision Letter:\\ learning constraints in spreadsheets and tabular data}
\begin{document}
\maketitle

\paragraph{Competitor algorithms and related work:} \textit{The reviewers point out that you should compare with other methods. One reviewer stresses that Microsoft Excel provides functionality that alleviates part of the problem you address. Another reviewer suggests that you compare to a baseline that considers all constraints. You must compare to algorithms that solve the same problem (at least part of it).}

We have elaborated the differences between TaCLe and FlashFill in the related work section (highlighted in \textcolor{red}{red}), also indicating why FlashFill cannot be applied to the examples we consider. Let us summarize the key differences here \sergey{Samuel, you had a point on this, can you add it ?}:
\begin{itemize}
  \item FlashFill works only with a single table, while TaCLe can handle multiple tables
  \item FlashFill only searches for a string transformation program, while TaCLe works with textual and numeric constraints
  \item FlashFill requires marked input-output columns (and columns are treated as vectors), while TaCLe is unsupervised and works across columns and rows
  \item \sergey{here should be the point by Samuel} like they produce a single program with a single solution, while we enumerate multiple constraints that hold in the data
\end{itemize}

Also, we have compared TaCLe with a baseline: one where the graph dependency is disabled and another where we disabled Step 1 (candidate generation based on properties) and graph dependency. \sergey{Samuel, Tias -- can you expand and correct this one?}

\end{document}

### 2016-11-23 Paper/ECAI/tabular.tex
added response letter

@@
 In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.
-Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
+\added{Furthermore, FlashFill handles only single tables with textual data in them, while the examples we consider are multi-tables with mixed numeric and textual data.} Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
 
@@
 
-%\tias{I am not defining a new problem for 'most specific canonical tabular constraint learning' because we loosely showed how implications and equivalences would be expressed, but not how to express canonicity and we did not claim that most specific is always best.}
-%\tias{Note also that our method finds ALL constraints; it uses specificity to faster traverse the search space, but the resulting set of constraints found would be the same, whether or not using the specificity (see 4.2.1).}
+%tias{I am not defining a new problem for 'most specific canonical tabular constraint learning' because we loosely showed how implications and equivalences would be expressed, but not how to express canonicity and we did not claim that most specific is always best.}
+%tias{Note also that our method finds ALL constraints; it uses specificity to faster traverse the search space, but the resulting set of constraints found would be the same, whether or not using the specificity (see 4.2.1).}
 
@@
  %$G_1 = +(G_2, G_3) \leftrightarrow G_1 = +(G_3, G_2)$.
-\samuel{\tias{How are they stopped in the code?
+samuel{tias{How are they stopped in the code?
 should be part of the pseud-code because not part of $Test(X)$?} will make it part of pseudo-code}
@@
 % Method $Test(G_R, G_{O_1}, G_{O_2})$\\
-% $If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$\tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
-% $If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ \tias{does not show that reasons over content of vectors}
+% $If \lnot G_{O_1} < G_{O_2}: Reject~\#~Canonicity$tias{ill-defined, $G_{O_1}$ is a group and $<$ not defined on it (on vector identifiers rather than content?)}\\
+% $If data(G_R) = data(G_{O_1}) \bullet data(G_{O_2}): Accept$ tias{does not show that reasons over content of vectors}
 % \\\\
 % Method $FindSubgroups(Assignments)$\\
-% $Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ \tias{does not enforce size of subgroups (e.g.
+% $Return \{ (G_R', G_{O_1}', G_{O_2}') | \exists (G_R, G_{O_1}, G_{O_2}) \in Assignments \land G_R' \subseteq G_R \land G_{O_1}' \subseteq G_{O_1} \land G_{O_2}' \subseteq G_{O_2} \land Test(G_R', G_{O_1}', G_{O_2}') \}$ tias{does not enforce size of subgroups (e.g.
 vectors)?}
@@
 %\subsubsection{Constraint solvers}
-%\samuel{Tias, Sergey, have a look at his section, I'm torn about rewriting it, feeling it should either be reduced or expanded, especially the second part.}
+%samuel{Tias, Sergey, have a look at his section, I'm torn about rewriting it, feeling it should either be reduced or expanded, especially the second part.}
 %Generic constraint solving technology can be used for the candidate group generation as well as for enumerating the satisfying subgroups.
@@
 
-For aggregates and project, a custom implementation is used that tries to find maximal subblocks that satisfy the constraint templates.
+For aggregates (e.g. \textit{SUM}) and \textit{PROJECT}, a custom implementation is used that tries to find maximal subblocks that satisfy the constraint templates.
 Instead of generating all possible subblocks, given a superblock assignment, these implementations start from a superblock and repeatedly generate and test smaller subblocks.
@@
 %Instead of the more generic generate-and-test implementation, we use, for example, caching, maximal range-finding and partial tests.
-\tias{Really, this is unacceptably vague.
+tias{Really, this is unacceptably vague.
 I have no clue what you mean with that, meaning that our research is not reproducable and hence not scientific}
@@
 
-%\paragraph{Theoretical analysis of runtime, this does not belong in experiments}\tias{move to technical section}\sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
+%\paragraph{Theoretical analysis of runtime, this does not belong in experiments}tias{move to technical section}sam{Has to be moved yet to 'Constraints' section, explaining fail-fast}
 %For most constraint templates the time required to check whether a subassignment satisfies the template is linear in the size of the subassignment (the number of subblocks and their size).
@@
 %However, in practice, most subassignments will not satisfy the constraint template.
-%Since our implementation tries to account for this by failing fast\tias{explain more precisely what you mean with 'failing fast'}, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
+%Since our implementation tries to account for this by failing fasttias{explain more precisely what you mean with 'failing fast'}, the amortized time is constant in the length of the vectors~$k$, unless $k$~is very large\footnote{Specifically, if $k$ is much larger than the number of candidate subassignments} or many constraints hold in the data.
 %For templates requiring their arguments to be single vectors $m$~is~$1$, and the total complexity for checking a subassignment becomes~$O(n) \simeq O(1)$ (since $n$ is typically small).
@@
 
-%Note that \samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?}
-
-% Extending the system with a new constraint template amounts to \tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}
+%Note that samuel{Motivate why product, not sum (=diff) you only need one, hardcoded preference for display - diff not present - pruning relation?}
+
+% Extending the system with a new constraint template amounts to tias{TODO: samuel, explain how to define syntax; signature-for-groupgen; signature-for-def; definition}
 
@@
 
-% \samuel{Move next to subgroup satisfaction search}
+% samuel{Move next to subgroup satisfaction search}
 % \section{Declarative modeling}
@@
 % As the number of vectors in a spreadsheet increases, the running time will quickly be dominated by constraints with a larger arity.
-% This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors \tias{more details about equal sized: is this a strict requirement for this to happen?
+% This effect is noticeable in our synthetic experiments when the number of (equally sized) vectors tias{more details about equal sized: is this a strict requirement for this to happen?
 I assume this is worst case and does not happen that much in practice?
@@
 In the generated spreadsheets, the first phase, which looks at properties of superblocks alone, can disqualify few or no superblock assignments, leading to some increase in runtime as more blocks are added.
-%therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.\sam{Not sure}
+%therefore, it will be faster to group the same vectors into larger super blocks, generate fewer superblock assignments to test and split these into single vector subassignments in phase~2.
 

### 2016-11-23 Paper/response_letter/response_letter.tex
updating the letter and the paper

@@
 \maketitle
+\section{Key issues}
 
@@
 
+\section{Detailed comments}
+
+\paragraph{Sensitivity to the noise}
+
+\paragraph{How the data was obtained and its detailed overview}
+We have extended the description of the gathered collection of spreadsheets in the paper (Section 5.3), also we provided an extensive technical description of the dataset in the corresponding github repository: \url{https://github.com/sergeyparamonov/tacle}.
+\sergey{Samuel, can you make a detailed overview of the dataset in github https://github.com/sergeyparamonov/tacle}?
+
 \end{document}

### 2016-11-23 Paper/ECAI/tabular.tex
updating the letter and the paper

@@
 In contrast, our approach is unsupervised, although it would be interesting to study whether it could be improved through user interaction and through user examples.
-\added{Furthermore, FlashFill handles only single tables with textual data in them, while the examples we consider are multi-tables with mixed numeric and textual data.} Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
+\added{Furthermore, FlashFill handles only single tables with textual data to obtain a string transformation program, while the examples we consider are multi-tables with mixed numeric and textual data and multiple constraints in them.} Another interesting open issue (for all of these techniques with the exception of logical and relational learning) is how to deal with possibly incomplete or noisy data.
 
@@
 \subsection{Effectiveness on real spreadsheets}
+\added{
 In order to test our approach on real data we assembled a benchmark of spreadsheets from three sources:
-1)~spreadsheets from an exercise session for teaching Excel based on~\cite{excel_book}; 2)~spreadsheets from tutorials online; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics or financial reports (available at the same repository\footref{github-link}).
+  1)~spreadsheets from an exercise session for teaching Excel at the Brussels Business School based on the book~\cite{excel_book}, covering the most popular Excel formulas; 2)~spreadsheets from online tutorials on Excel formulas; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics (FBI:UCR) or financial reports (US BEA).
+The data and its extended description are available at the same repository\footref{github-link} with the links to all original publicly available spreadsheets used in the experiments (in the file \textit{links.txt}).
+}
+
+\sergey{Samuel, can you provide an extended technical description in github? Like straight in the README?}
 

### 2016-11-24 Code/src/core/assignment.py
Submission fixes and tweaks

@@ -3,6 +3,7 @@ from typing import List, Dict, Set
 import itertools
 from constraint import Problem
 from core.group import Group, Orientation
+from core.solutions import Solutions
 from parse.parser import GType
 
 
@@ -46,9 +47,9 @@ class Source:
         return self._variables
 
     def candidates(self, groups, solutions, filters):
-        return self._complete([{}], groups, filters)
+        return self._complete([{}], groups, filters, solutions)
 
-    def _complete(self, assignments, groups, filters):
+    def _complete(self, assignments, groups, filters, solutions):
         result = []
 
         # TODO type as constraint
@@ -67,17 +68,17 @@ class Source:
                 variables = list([v.name for v in f.variables])
 
                 def c_j(ff, vv):
-                    return lambda *args: ff.test({vv[i]: args[i] for i in range(len(args))})
+                    return lambda *args: ff.test({vv[i]: args[i] for i in range(len(args))}, solutions)
 
                 problem.addConstraint(c_j(f, variables), variables)
 
             return True, list(problem.getSolutions())
 
         for assignment in assignments:
-            resume, solutions = try_assignment()
+            resume, candidate_solutions = try_assignment()
             if not resume:
                 return []
-            result += solutions
+            result += candidate_solutions
         return result
 
     def depends_on(self) -> Set:
@@ -96,7 +97,7 @@ class ConstraintSource(Source):
 
     def candidates(self, groups, solutions, filters):
         assignments = [{self.dictionary[k]: v for k, v in s.items()} for s in solutions.get_solutions(self.constraint)]
-        return self._complete(assignments, groups, filters)
+        return self._complete(assignments, groups, filters, solutions)
 
     def depends_on(self):
         return {self.constraint}
@@ -110,7 +111,7 @@ class Filter:
     def variables(self):
         return self._variables
 
-    def test(self, assignment: Dict[str, Group]):
+    def test(self, assignment: Dict[str, Group], solutions):
         raise NotImplementedError()
 
     def test_same(self, assignment, f):
@@ -127,8 +128,8 @@ class Not(Filter):
         super().__init__(original_filter.variables)
         self._original_filter = original_filter
 
-    def test(self, assignment):
-        return not self._original_filter.test(assignment)
+    def test(self, assignment, solutions):
+        return not self._original_filter.test(assignment, solutions)
 
 
[…]

### 2016-11-24 Code/src/core/constraint.py

@@ -4,7 +4,7 @@ from typing import List, Dict
 import numpy
 
 from core.assignment import Source, Filter, Variable, SameLength, ConstraintSource, SameTable, \
-    SameOrientation, SameType, SizeFilter, Not, NotPartial, Partial
+    SameOrientation, SameType, SizeFilter, Not, NotPartial, Partial, SatisfiesConstraint
 from core.group import GType, Group, Orientation
 
 
@@ -136,7 +136,7 @@ class Aggregate(Constraint):
         op_string = operation.name
         variables = [self.x, self.y]
 
-        def test(_, a: Dict[str, Group]):
+        def test(_, a: Dict[str, Group], _solutions):
             x_group, y_group = [a[v.name] for v in variables]
             o_match = x_group.row == (orientation == Orientation.HORIZONTAL)
             if not o_match and x_group.vectors() < self.min_vectors:
@@ -260,6 +260,7 @@ class FuzzyLookup(Constraint):
 
     def __init__(self):
         variables = [self.o_key, self.o_value, self.f_key, self.f_value]
+        # source = Source(variables)
         source = ConstraintSource(variables, Ordered(), {Ordered.x.name: self.o_key.name})
         filters = [SameType([self.o_value, self.f_value]), NotPartial(variables),
                    SameLength([self.f_key, self.f_value]), SameLength([self.o_key, self.o_value]),
@@ -283,7 +284,7 @@ class ConditionalAggregate(Constraint):
         source = ConstraintSource(variables, all_diff, {all_diff.x.name: "OK"})
         filters = [SameLength([self.o_key, self.result]), SameLength([self.f_key, self.values]),
                    SameTable([self.f_key, self.values]), Not(SameTable([self.f_key, self.o_key])),
-                   SameTable([self.o_key, self.result]),  # TODO think about this
+                   # SameTable([self.o_key, self.result]),  # TODO think about this
                    NotPartial([self.o_key]), SameType([self.f_key, self.o_key]),
                    SameOrientation([self.o_key, self.result]), SameOrientation([self.f_key, self.values])]
         p_format = "{R} = " + name.upper() + "IF({FK}={OK}, {V})"

### 2016-11-24 Code/src/experiment.py

@@ -42,7 +42,7 @@ class ConstraintCounter:
     def count(self, found=None, relevant=None, supported=False):
         relevant_list = self._relevant
         if supported is True:
-            relevant_list = list(r for r in relevant_list if r in [c.name for c in excel_constraints()])
+            relevant_list = list(r for r, _ in relevant_list if r in [c.name for c in excel_constraints()])
         if relevant is True and found is True:
                 return len(relevant_list) - len(self._tracking)
         elif relevant is True and found is False:
@@ -101,6 +101,8 @@ class Experiment:
             csv_file, config_file = self._get_csv_file(), self._get_config_file()
             learning_task = workflow.task(csv_file, config_file)
             solutions = learning_task.run()
+            # print(*learning_task.constraints, sep="\t")
+            # print(*["{:.3f}".format(learning_task.time(c)) for c in learning_task.constraints], sep="\t")
 #
             self._running_times.append(learning_task.total_time())
             self.count_constraints(solutions)

### 2016-11-24 Code/src/speed_tests.py

@@ -103,7 +103,7 @@ def generate_random(test_id: SpeedTestId):
             for _ in range(size):
                 max_rows = max(max_rows, table.rows)
                 if b_type == GType.int:
-                    column_data.append(np.random.randint(1, 100, table.rows))
+                    column_data.append(np.random.randint(1, 1000, table.rows))
                 elif b_type == GType.string:
                     letter = lambda: random.choice(string.ascii_lowercase)
                     word = lambda: ''.join(letter() for _ in range(np.random.randint(5, 15)))
@@ -136,8 +136,10 @@ def generate_blocks(test_id: SpeedTestId):
     content = dict()
     # Generate tables
     tables = []
+    index = 1
     for i, table in enumerate(test_id.tables):
-        bounds = [1, table.rows, table.cols * i + 1, table.cols * (i + 1)]
+        bounds = [1, table.rows, index, index + table.cols - 1]
+        index += table.cols
         tables.append({"Name": "T{}".format(i + 1), "Bounds": bounds, "Orientation": "Column"})
     content["Tables"] = tables
 
@@ -278,7 +280,7 @@ def draw_figures(tables):
         table = tables["cat1"]
         scatter_1 = ScatterData("(a) Number of vectors", list(test_id.tables[0].cols for test_id in table))
         scatter_1.add_data("All constraints", list(np.average(series["total"]) for series in table.values()))
-        scatter_1.x_lim((10**0, 10**4))
+        scatter_1.x_lim((10**0, 10**2))
         scatter_1.y_lim((10**-2, 10**2))
         scatters.append(scatter_1)
 
@@ -301,19 +303,13 @@ def draw_figures(tables):
         scatter_3 = ScatterData("(c) Number of blocks (aggregates)", number_of_blocks)
         total_times = list(np.average(series["total"]) for series in table.values())
         aggregates = list(sum(np.average(series[c]) for c in Aggregate.instances()) for series in table.values())
+        non_aggregates = list(total - aggregate for total, aggregate in zip(total_times, aggregates))
         scatter_3.add_data("All constraints", total_times)
         scatter_3.add_data("Aggregate constraints", aggregates)
-        scatter_3.x_lim((10**0, 10**4))
+        scatter_3.add_data("Non-aggregate constraints", non_aggregates)
+        scatter_3.x_lim((10**0, 10**2))
         scatter_3.y_lim((10**-2, 10**2))
         scatters.append(scatter_3)
-
[…]

### 2016-11-24 Paper/ECAI/information_sheet.tex

\documentclass{article}
\usepackage{enumerate}
\usepackage[english]{babel}
\usepackage{times}
\usepackage{latexsym}    % to get LASY symbols
\usepackage{graphicx}    % to insert PostScript figures
\newcommand{\scaleconstant}{1.0}
\bibliographystyle{plainnat}
\author{Samuel Kolb \and Sergey Paramonov \and Tias Guns \and Luc {De Raedt}}
\title{MLJ contribution information sheet:\\ learning constraints in spreadsheets and tabular data}
\begin{document}
\maketitle

\paragraph{Question} \textit{What is the main claim of the paper? Why is this an important contribution to the machine learning/data mining literature?}

It is possible to learn constraints in flat tabular spreadsheet data in an unsupervised way. We propose a general-purpose method and system, named TaCLe (from: Tabular Constraint Learner) for discovering constraints spanning entire rows and columns. This is important because millions of people across the world use spreadsheets every day yet errors are not uncommon. This could be overcome through the development of intelligent user-assisting tools, of which the problem we study here is at the core.

Our contributions are as follows:
\begin{itemize}
  \item we define the tabular constraint learning problem, where the goal is to find constraints that range over entire rows or columns in an unsupervised way;
  \item we propose an effective 2-stage generate and test method where the first stage reasons only over properties of contiguous blocks of rows/columns, and the second stage continues to investigate individual rows and columns and their content;
  \item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximally contiguous blocks compatible with the candidate constraint's argument; 
  \item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall. 
\end{itemize}

\paragraph{Question} \textit{What is the evidence provided to support claims? Be precise.}

We have provided theoretical as well as experimental evidence. The main challenge is the number of possible constraints and combinations of rows and columns that need to be tried as input to the constraints. In Section 4 "Approach to Tabular Constraint Learning", we have presented theoretical analysis and a number of optimizations demonstrating how the algorithm reduces the search space and effectively navigates it using a two step procedure for candidate generation. In Section 5 "Evaluation", we have presented experimental evaluation of the TaCLe system on generated data demonstrating how different factors such as the length of vectors or the number of vectors affect the runtime, as well as evaluating it on a collection of Excel documents (with a link to a github repository with the data and screenshots of the system). The experiments demonstrate that the system has high precision and recall of learned constraint, while keeping runtime within reasonable bounds static and interactive use.

\paragraph{Question} \textit{Report 3-5 most closely related contributions in the past 7 years (authored by researchers outside the authors’ research group) and briefly state the relation of the submission to them.}

\begin{itemize}
  \item Flashfill [8] (POPL, 2011) and Flashextract [13] (PLDI, 2014) by Sumit Gulwani et al. The former learns a string transformation function from a very few examples in an Excel spreadsheet, the latter learns a function extracting particular parts from a textual file such as xls, html or txt, given a set of marked examples. Both of them work in the supervised manner learning a string transformation function from a document with a set of examples. In our case, we work with unlabeled data in an unsupervised manner, we can have multiple tables with relationships between them with mixed textual and numeric data, contrary to both systems where columns are variables and rows are data points, in our work columns and row can be variables and data points.
  \item Model-seeker [1]  (CP, 2012) by Nicolas Beldiceanu and Helmut Simonis. The Model seeker system is designed to find several types of numeric constraints that are common for constraint programming community such as scheduling or combinatorial optimization. We consider both numerical and textual constraint with a focus on spreadsheet formulas that are different from what typically studied in constraint programming literature.
  \item Toward a logical/physical theory of spreadsheet modeling (TOIS, 1995) by Tomas Isakowitz et al. [11]. They have studied generalization of the formulas present in an Excel document and the development of a logical model of spreadsheets. The key difference is that they generalize existing formulas, while we learn them directly from the data.
\end{itemize}

\paragraph{Question} \textit{Have you published parts of your paper before, for instance in a conference? If so, give details of your previous paper(s) and a precise statement detailing how your paper provides a significant contribution beyond the previous paper(s).}
No.

\paragraph{Question} \textit{Who are the most appropriate reviewers for the paper? Authors are required to suggest up to four candidate reviewers (especially if external to the Guest Editorial Board), including a brief motivation for each suggestion.}
\begin{itemize}
  \item Michele Sebag for the work in constraint programming and learning
  \item Arnaud Lallouet for the work in constraint programming, constraints learning and data mining
  \item Jos\'e Hern\'andez-Orallo for the work in inductive programming
\end{itemize}

\end{document}

### 2016-11-24 Code/src/core/constraint.py
Submission fixes and tweaks

@@ -174,7 +174,6 @@ class Aggregate(Constraint):
 
 # TODO Subset -> Fuzzy lookup
 
-
 class Permutation(Constraint):
     x = Variable("X", types=numeric)
 

### 2016-11-28 Paper/response_letter/response_letter.tex
adding comments
updated noise response

@@
 \usepackage{times}
+\usepackage{hyperref}
 \usepackage{latexsym}    % to get LASY symbols
@@
 \paragraph{Sensitivity to the noise}
+In the field of Inductive Logic Programming and Constraint Learning, there has been a line of work devoted to noise handling. Let us briefly overview the key methods applicable in our case:
+\begin{itemize}
+  \item \textbf{Noisy semantics}: a possible approach is to modify the semantics of constraints such as $\sum_{i=1}^n b_i = c$ into $\sum_{i=1}^n b_i = c + \epsilon$, where $\epsilon$ is the noise level that might be dependent on $b_i$'s, $n$ or $c$.
+  \item \textbf{Skip semantics}: another possibility is to allow skipping of elements in vectors, i.e.
+in the constraint  $\sum_{i=1}^n b_i = c$, we can allow $i$ to skip up to $k$ elements to match $c$, then we treat skipped values as noise and do not take them into account.
+  \item \textbf{Soft constraints}: a constraint satisfaction is determined by a real-value measure such as $\sum_{i=1}^n b_i = c + \sigma|c|$, where $\sigma=0$ denotes the perfect match and the other values denote deviations in absolute values of $c$, then our goal would be to select top-$k$ best matching constraints according to the designed measure.
+\end{itemize}
 
@@
 
+\sergey{Also, they wanted some real examples of formulas from the dataset in Section 6}
+
 \end{document}

### 2016-11-28 Code/src/get_constraint_stats.py

@@ -1,6 +1,7 @@
 import os
 from collections import defaultdict
 import json
+from workflow import get_constraint_list
 
 def get_category(filename, categories):
     for key, value in categories.items():
@@ -10,6 +11,7 @@ def get_category(filename, categories):
     raise Exception("that shouldn't have happened")
 
 def main():
+    supported_constraints = {x.name for x in get_constraint_list()}
     folder    = "data/truth/"
     filenames = set(os.listdir(folder)) - set(['examples.txt','fbi_offenses.txt',"columnwise-sum-rows.txt",'exps.txt'])
     stats_absolute              = defaultdict(int)
@@ -17,6 +19,8 @@ def main():
     number_of_files_in_category = defaultdict(int)
     all_constraints = set()
     all_categories  = set()
+    overall_counter = 0
+    constraints_by_category = defaultdict(int)
     with open("data/data.txt", "r") as catetoriesfile:
         categories = json.load(catetoriesfile)
     for filename in filenames:
@@ -26,9 +30,18 @@ def main():
             number_of_files_in_category[category] += 1
             raw         = json.load(afile)
             constraints = raw['Essential']
+          # if raw['Non-trivial']:
+          #     constraints = constraints.update(raw['Non-trivial'])
             for key, value in constraints.items():
-                print("key",key,"value",value)
-                stats_absolute[(category,key)] += 1
+                if key not in supported_constraints:
+                    continue
+#               print("key",key,"value",value)
+                stats_absolute[(category,key)] += len(value)
+                if len(value) > 1:
+                    print(value)
+                    print("len",len(value))
+                overall_counter += len(value)
+                constraints_by_category[category] += len(value)
                 stats_fraction_aux[(category,key)].add(filename)
                 all_constraints.add(key)
     stats_fraction = defaultdict(float)
@@ -38,20 +51,25 @@ def main():
    #print(stats_absolute)
    #print(stats_fraction)
    #print(number_of_files_in_category)
-    
-    print("constraint," + ",".join(all_categories))
+    all_constraints = sorted(all_constraints)
+    print("| constraint " + " | ".join(all_categories) + " |")
     for constraint in all_constraints:
-        print(constraint, end=",")
+        print("| "+constraint, end=" | ")
         for category in all_categories:
-            print("{:.2f}".format(stats_fraction[(category,constraint)]),end=",")
+            print("{:.2f}".format(stats_fraction[(category,constraint)]),end=" | ")
[…]

### 2016-11-28 Paper/response_letter/response_letter.tex
editing the response letter

@@
 \maketitle
+\sergey{just in case, the text is preliminary, feel free to cut, expand and rewrite, I think it should be 1 or 2 pages at the end, I hope}
 \section{Key issues}
+
+\paragraph{Goal and motivation}
+\sergey{should we even mention possible approach to compression here?}
+
+\paragraph{Approach}
+\sergey{here is the response on the unclear definitions and notation}
 
@@
 
+We have expanded the conclusions section to accommodate these proposals.
+
 \paragraph{How the data was obtained and its detailed overview}

### 2016-11-28 Paper/response_letter/response_letter.tex
making comments more precise

@@
 \begin{itemize}
-  \item \textbf{Noisy semantics}: a possible approach is to modify the semantics of constraints such as $\sum_{i=1}^n b_i = c$ into $\sum_{i=1}^n b_i = c + \epsilon$, where $\epsilon$ is the noise level that might be dependent on $b_i$'s, $n$ or $c$.
+  \item \textbf{Noisy semantics}: a possible approach is to modify the semantics of constraints such as $\sum_{i=1}^n b_i = c$ into $|\sum_{i=1}^n b_i - c| \leq \epsilon$, where $\epsilon$ is the noise level that might be dependent on $b_i$'s, $n$ or $c$.
   \item \textbf{Skip semantics}: another possibility is to allow skipping of elements in vectors, i.e.
 in the constraint  $\sum_{i=1}^n b_i = c$, we can allow $i$ to skip up to $k$ elements to match $c$, then we treat skipped values as noise and do not take them into account.
-  \item \textbf{Soft constraints}: a constraint satisfaction is determined by a real-value measure such as $\sum_{i=1}^n b_i = c + \sigma|c|$, where $\sigma=0$ denotes the perfect match and the other values denote deviations in absolute values of $c$, then our goal would be to select top-$k$ best matching constraints according to the designed measure.
+  \item \textbf{Soft constraints}: a constraint satisfaction is determined by a real-value measure such as $|\sum_{i=1}^n b_i - c | = \sigma|c|$, where $\sigma=0$ denotes the perfect match and the other values denote deviations in absolute values of $c$, then our goal would be to select top-$k$ best matching constraints according to the designed measure.
 \end{itemize}

### 2016-11-28 Paper/ECAI/tabular.tex
making comments more precise

@@
 
-A final direction is that the current approach assumes noise-free data (but see the part on error correction); in inductive logic programming there has been much work on how to handle noisy data, while in constraint satisfaction a popular alternative is to consider \textit{soft} constraints that can be violated (at a cost).
+A final direction is that the current approach assumes noise-free data (but see the part on error correction); in inductive logic programming there has been much work on how to handle noisy data such as allowing a tolerance level $\epsilon$ when a rule or a constraint $\sum_{i=1}^n b_i = c$ matches up to   $\sum_{i=1}^n b_i = c + \epsilon$, while in constraint satisfaction a popular alternative is to consider \textit{soft} constraints that can be violated (at a cost).
 This could extend the approach to new application domains, beyond traditional spreadsheets.

### 2016-11-28 Paper/ECAI/tabular.tex
added a note on the noise handling

--- 2016-11-23 Paper/ECAI/tabular.tex
@@
 
-A final direction is that the current approach assumes noise-free data (but see the part on error correction); in inductive logic programming there has been much work on how to handle noisy data, while in constraint satisfaction a popular alternative is to consider \textit{soft} constraints that can be violated (at a cost).
+\added{A final direction is that the current approach assumes noise-free data (but see the part on error correction); in inductive logic programming there has been much work on how to handle noisy data such as allowing a tolerance level $\epsilon$ when a rule or a constraint $\sum_{i=1}^n b_i = c$ matches up to $\epsilon$: $|\sum_{i=1}^n b_i - c| \leq \epsilon$, while in constraint satisfaction a popular alternative is to consider \textit{soft} constraints that can be violated at a cost, which is a real value and then the task is to find top-$k$ best matching constraints.
+}
 This could extend the approach to new application domains, beyond traditional spreadsheets.

### 2016-12-05 Code/src/core/assignment.py
Solved sum-product relaxation bug

@@ -68,7 +68,7 @@ class Source:
                 variables = list([v.name for v in f.variables])
 
                 def c_j(ff, vv):
-                    return lambda *args: ff.test({vv[i]: args[i] for i in range(len(args))}, solutions)
+                    return lambda *args: ff.test_relaxed({vv[i]: args[i] for i in range(len(args))}, solutions)
 
                 problem.addConstraint(c_j(f, variables), variables)
 
@@ -113,6 +113,12 @@ class Filter:
 
     def test(self, assignment: Dict[str, Group], solutions):
         raise NotImplementedError()
+
+    def test_relaxed(self, assignment, solutions):
+        return self.test(assignment, solutions)
+
+    def is_relaxed(self):
+        return False
 
     def test_same(self, assignment, f):
         groups = list([assignment[v.name] for v in self.variables])
@@ -202,6 +208,20 @@ class SizeFilter(Filter):
             and all(self._cols is None or op(g.columns(), self._cols) for g in groups) \
             and all(self._length is None or op(g.length(), self._length) for g in groups) \
             and all(self._vectors is None or op(g.vectors(), self._vectors) for g in groups)
+
+    def test_relaxed(self, assignment: Dict[str, Group], solutions):
+        if self._max_size:
+            # Max size can not be enforced at the super-block level
+            groups = list([assignment[v.name] for v in self.variables])
+            valid = all(self._rows is None or g.row_oriented() or g.rows() <= self._rows for g in groups) and all(
+                self._cols is None or not g.row_oriented() or g.columns() <= self._cols for g in groups) and all(
+                self._length is None or g.length() <= self._length for g in groups)
+            return valid
+        else:
+            return self.test(assignment, solutions)
+
+    def is_relaxed(self):
+        return self._max_size
 
 
 class OrientationFilter(Filter):

### 2016-12-05 Code/src/engine/internal.py

@@ -439,13 +439,13 @@ class InternalSolvingStrategy(DictSolvingStrategy):
             keys = [c.result, c.first, c.second]
 
             def is_sum_product(r_v, o1_v, o2_v):
-                if not ordered(o1_v, o2_v):
+                if not ordered(o1_v, o2_v) or r_v.rows() != 1 or r_v.columns() != 1:
                     return False
 
                 r, o1, o2 = to_data(r_v, o1_v, o2_v)
 
                 # TODO too many vector operations (easy)
-                return numpy.vectorize(equal)(r, numpy.sum(numpy.vectorize(Operation.PRODUCT.func)(o1, o2))).all()
+                return equal_v(r, numpy.sum(numpy.vectorize(Operation.PRODUCT.func)(o1, o2))).all()
 
             return self._generate_test_vectors(assignments, keys, is_sum_product)
 

### 2016-12-05 Code/src/playground.py

@@ -44,7 +44,7 @@ def main():
         print("\n" * 2)
         max_label = list(t[3](maximal[t[0]]) for t in info)
         global_plot = Plotter("comparison", list(t[0] for t in info), upper=max_label)
-        test = {"Exercises": (30, 2), "Tutorials": (46, 20), "Data": (6, 0)}
+        test = {"Exercises": (30, 2), "Tutorials": (46, 21), "Data": (6, 0)}
         for category, cat_stats in [(cat, stats[cat]) for cat in categories]:
             print(category)
             plotter = Plotter("benchmark_" + category)

### 2016-12-05 Paper/response_letter/response_letter.tex

@@
 \maketitle
+\sergey{We need a discussion section on: noise, (non)-contiguous blocks and combination of constraints} 
 \sergey{just in case, the text is preliminary, feel free to cut, expand and rewrite, I think it should be 1 or 2 pages at the end, I hope}
@@
 \paragraph{Goal and motivation}
-\sergey{should we even mention possible approach to compression here?}
+\sergey{Reviewer pointed out that (and we should thank him and use it, I think):}
+
+\textbf{Reviewer \#3:}
+Section 6.
+The reviewer would like to point out the important field of data cleaning (of dirty data, as it is called in the database community) and anomaly detection, beyond simple error detection.
+Because your method does not only detect errors, but can also show *why* it deems something as erronous your method falls in the desirable category of descriptive or explainable anomaly detection algorithms.\\
+\sergey{A sentence on compression, maybe?}\\
+\sergey{We need to connect better Sections 5 and 6 -- from the reviews}
 
@@
 You must compare to algorithms that solve the same problem (at least part of it).}
+
+\sergey{Tias, you also had a point here that flash fill is searching for a combination of string transformations, we search for atomic constraints and etc, could you integrate it here?
+and indicate what part should be added to the paper?}
 

### 2016-12-09 Paper/ECAI/tabular.tex
Reworked motivation

@@
 Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across all tables.
-For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.
+For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulas.
 
 % why learn them
+\added{
 Such issues and mistakes could be overcome through the development of intelligent user-assisting tools.
-A core aspect of such tools would be the ability to automatically discover what constraints and formulae are present in a spreadsheet.
+A core aspect of such tools would be the ability to automatically discover what constraints and formulas are present in a spreadsheet.
 This is the problem we study in this paper.
-There are multiple applications that could build on this, such as rediscovery of constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
+Concretely we consider five key applications that intelligent tools based on automatic constraint discovery may facilitate: auto-completion, error checking, formula suggestion, rich import and data compression.
+Auto-completion helps users quickly populate spreadsheets by suggesting values for unfilled cells based on constraints that have previously been detected within the spreadsheet.
+Error checking helps uncover erroneous values in spreadsheets, for example, while typing, by finding previously discovered constraint that are violated by new values.
+Both of these applications can facilitate the use of spreadsheets even for users who do not explicitly use formulas at all.
+Of course, these applications are also able to explain to more advanced users what constraints are responsible for suggesting or rejecting a value.
+Formula suggestion is a variant of auto-completion that, instead of suggesting a \emph{value}, provides a formula that computes a user provided value.
+This setting can help users to find formulas that perform a certain task or find the right syntax for a formula.
+We use rich import to denote the task of importing semi-structured data from, for example, CSV files and reconstruct constraints that were used, implicitly or explicitly, to generate the data.
+Rich import goes hand-in-hand with data compression, since a file can be compressed by identifying all values that are generated by a formula and storing that formula instead of the values generated by it.
+%There are multiple applications that could build on this, such as rediscovery of constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
+}
 
@@
 \item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall.
+\added{\item we describe on a high level how auto-completion and error checking can be realized using \sname}
 \end{itemize}
@@
 Section \ref{sec:formalization} introduces the relevant concepts and defines the problem formally while Section \ref{sec:approach} presents our approach.
-This is evaluated in Section \ref{sec:evaluation} after which we show two applications of how our system could be used in Section \ref{sec:applications}, and we conclude in Section \ref{sec:conclusions}.
+This is evaluated in Section \ref{sec:evaluation} after which \added{we show how to realize two of the motivating applications using our system} in Section \ref{sec:applications}, and we conclude in Section \ref{sec:conclusions}.
 

### 2016-12-12 Paper/response_letter/response_letter.tex
should we mention dataset as contribution?

@@
 \sergey{A sentence on compression, maybe?}\\
+\sergey{Can we claim as a contribution creation of the first dataset for tabular constraint learning? I guess it is a valid claim, it didn't exist before at all, right?}\\
 \sergey{We need to connect better Sections 5 and 6 -- from the reviews}
@@
 \paragraph{How the data was obtained and its detailed overview}
-We have extended the description of the gathered collection of spreadsheets in the paper (Section 5.3), also we provided an extensive technical description of the dataset in the corresponding github repository: \url{https://github.com/sergeyparamonov/tacle}.
+We have extended the description of the gathered collection of spreadsheets in the paper (Section 5.3), also we provided an extensive technical description of the dataset in the corresponding github repository: \url{https://github.com/sergeyparamonov/tacle} and added a table indicating present constraints by category into the appendix.
 \sergey{Samuel, can you make a detailed overview of the dataset in github https://github.com/sergeyparamonov/tacle}?
@@
 
+\sergey{Samuel?
+I guess, we have to say that it should be specified in the constraint itself, on what it depends, kind of given}~ \textbf{Reviwer \#2}: C.
+According to my understanding, the proposed method exploits possible dependencies and redundancies that may exist between the different constraints.
+However, it is unclear how such dependencies/redundancies are found.
+Are they assumed to be given as input?
+Are they detected by the algorithm?
+What would be the performance (runtime and precision/recall) of the algorithm if no dependencies/redundancies exist?
+
 \end{document}

### 2016-12-12 Paper/ECAI/tabular.tex
should we mention dataset as contribution?

@@
 1)~spreadsheets from an exercise session for teaching Excel at the Brussels Business School based on the book~\cite{excel_book}, covering the most popular Excel formulas; 2)~spreadsheets from online tutorials on Excel formulas; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics (FBI:UCR) or financial reports (US BEA).
-The data and its extended description are available at the same repository\footref{github-link} with the links to all original publicly available spreadsheets used in the experiments (in the file \textit{links.txt}).
+The data and its extended description are available in Appendix \ref{appendix:dataset_overview} in Table \ref{tab:constraint_by_category} and at the same repository\footref{github-link} with the links to all original publicly available spreadsheets used in the experiments (in the file \textit{links.txt}).
 }
 
-\sergey{Samuel, can you provide an extended technical description in github? Like straight in the README?}
+\sergey{Samuel, can you provide an extended technical description in github? Like straight in the README next to the tables?}
 
@@
   \bibliography{references}}
+\appendix
+\section{Spreadsheet Dataset Overview}
+\sergey{Samuel, we need an extended discussion here on how the data was obtained, I'll try a first sketch here}\\
+There are three main sources for spreadsheets we have collected.
+\begin{itemize}
+  \item 
+  \item 
+  \item 
+\end{itemize}
+
+\begin{table}
+  \centering
+  \rowcolors{2}{gray!25}{white}
+  \begin{tabular}{l | c c c }
+    \textbf{Constraint} & \textbf{Exercises} & \textbf{Data} & \textbf{Tutorials}  \\\hline
+    average (col) & 0 & 0 & 2 \\
+    average (row) & 0 & 1 & 0 \\
+    average-if & 1 & 0 & 0 \\
+    count (col) & 0 & 0 & 1 \\
+    count-if & 1 & 0 & 0 \\
+    difference & 2 & 1 & 0 \\
+    equal & 0 & 0 & 10 \\
+    foreign-product & 1 & 0 & 0 \\
+    fuzzy-lookup & 2 & 0 & 1 \\
+    lookup & 1 & 0 & 0 \\
+    max (col) & 0 & 0 & 2 \\
+    max-if & 1 & 0 & 0 \\
+    min (col) & 0 & 0 & 3 \\
+    min-if & 1 & 0 & 0 \\
+    percentual-diff & 5 & 1 & 0 \\
+    product & 2 & 0 & 3 \\
+    project & 1 & 0 & 0 \\
+    rank & 1 & 0 & 0 \\
+    series & 2 & 1 & 0 \\
+    sum (col) & 5 & 0 & 10 \\
+    sum (row) & 2 & 2 & 3 \\
+    sum-if & 2 & 0 & 9 \\
+    sum-product & 0 & 0 & 2 \\
+  \end{tabular}
+  \caption{Constraint occurrence by category}
+  \label{tab:constraint_by_category}
+\end{table}
+
 \end{document}

### 2016-12-12 Paper/ECAI/tabular.tex
updating appendix to put an overview of the dataset there

@@
 \section{Spreadsheet Dataset Overview}
+\label{appendix:dataset_overview}
 \sergey{Samuel, we need an extended discussion here on how the data was obtained, I'll try a first sketch here}\\
@@
 \begin{itemize}
-  \item 
-  \item 
-  \item 
+  \item After identification of popular Excel functions (MS Office web page has an overview of popular functions), we have searched for online tutorials about these functions and collected them under the category \textbf{Tutorials}.
+  \item We have collected the exercises under the category \textbf{Exercises} from the introductory Excel book \cite{excel_book} that focused on popular Excel functions.
+  \item We have collected under the category \textbf{Data} publicly available economic and crime reporting data from the U.S.
+Bureau of Economic Analysis (BEA) and U.S.
+FBI Uniform Crime Reporting (UCR) Program.
 \end{itemize}
+
+The overview of constraint distributions per category is presented in Table \ref{tab:constraint_by_category}.
 
@@
   \rowcolors{2}{gray!25}{white}
+  \caption{Constraint occurrence by category}
+  \label{tab:constraint_by_category}
   \begin{tabular}{l | c c c }
@@
   \end{tabular}
-  \caption{Constraint occurrence by category}
-  \label{tab:constraint_by_category}
 \end{table}

### 2016-12-13 Paper/ECAI/tabular.tex
updating revision letter

@@
 \appendix
-\section{Spreadsheet Dataset Overview}
-\label{appendix:dataset_overview}
+\section{Spreadsheet Dataset Overview} \label{appendix:dataset_overview}
 \sergey{Samuel, we need an extended discussion here on how the data was obtained, I'll try a first sketch here}\\
@@
 \begin{itemize}
-  \item After identification of popular Excel functions (MS Office web page has an overview of popular functions), we have searched for online tutorials about these functions and collected them under the category \textbf{Tutorials}.
+  \item After identification of popular Excel functions (the MS Office web page has an overview of popular functions), we have searched for online tutorials about these functions and collected them under the category \textbf{Tutorials} (a link to each webpage found and used is provided in the accompanying github repository).
   \item We have collected the exercises under the category \textbf{Exercises} from the introductory Excel book \cite{excel_book} that focused on popular Excel functions.
@@
 Bureau of Economic Analysis (BEA) and U.S.
-FBI Uniform Crime Reporting (UCR) Program.
+FBI Uniform Crime Reporting (UCR) Program (the links are also provided in the accompanying github repository).
 \end{itemize}

### 2016-12-19 Paper/ECAI/tabular.tex
WIP Autocompletion

--- 2016-12-12 Paper/ECAI/tabular.tex
@@
   1)~spreadsheets from an exercise session for teaching Excel at the Brussels Business School based on the book~\cite{excel_book}, covering the most popular Excel formulas; 2)~spreadsheets from online tutorials on Excel formulas; and 3)~publicly available \textit{data} spreadsheets such as real-world crime statistics (FBI:UCR) or financial reports (US BEA).
-The data and its extended description are available in Appendix \ref{appendix:dataset_overview} in Table \ref{tab:constraint_by_category} and at the same repository\footref{github-link} with the links to all original publicly available spreadsheets used in the experiments (in the file \textit{links.txt}).
+The data and its extended description are available at the same repository\footref{github-link} with the links to all original publicly available spreadsheets used in the experiments (in the file \textit{links.txt}).
 }
 
-\sergey{Samuel, can you provide an extended technical description in github? Like straight in the README next to the tables?}
+\sergey{Samuel, can you provide an extended technical description in github? Like straight in the README?}
 
@@
 \section{Applications}\label{sec:applications}
-In this section, we illustrate how various applications can build on our method for solving the tabular constraint learning problem.
-
-\subsection{Autocompletion}
-Autocompletion can be seen as using knowledge derived from the current data to predict new values before they are written by the user.
+\added{In this section, we illustrate how two of the motivating applications, auto-completion and error checking, can build on our method for solving the tabular constraint learning problem.}
+
+\subsection{Auto-completion}
+Auto-completion can be seen as using knowledge derived from the current data to predict new values before they are written by the user.
+\added{We can decompose this process into three phases: 1) learn constraints on a snapshot~($D_0$) of the spreadsheet, 2) new data~($\Delta D$) is added by the user, and 3) the system uses the learned constraints to predict missing values.
+}
 A snapshot~($D_0$) is used to learn constraints and these are combined with new data~($\Delta D$) that the user has added.
@@
   \bibliography{references}}
-\appendix
-\section{Spreadsheet Dataset Overview}
-\sergey{Samuel, we need an extended discussion here on how the data was obtained, I'll try a first sketch here}\\
-There are three main sources for spreadsheets we have collected.
-\begin{itemize}
-  \item 
-  \item 
-  \item 
-\end{itemize}
-
-\begin{table}
-  \centering
-  \rowcolors{2}{gray!25}{white}
-  \begin{tabular}{l | c c c }
-    \textbf{Constraint} & \textbf{Exercises} & \textbf{Data} & \textbf{Tutorials}  \\\hline
-    average (col) & 0 & 0 & 2 \\
-    average (row) & 0 & 1 & 0 \\
-    average-if & 1 & 0 & 0 \\
-    count (col) & 0 & 0 & 1 \\
-    count-if & 1 & 0 & 0 \\
-    difference & 2 & 1 & 0 \\
-    equal & 0 & 0 & 10 \\
-    foreign-product & 1 & 0 & 0 \\
-    fuzzy-lookup & 2 & 0 & 1 \\
-    lookup & 1 & 0 & 0 \\
-    max (col) & 0 & 0 & 2 \\
-    max-if & 1 & 0 & 0 \\
-    min (col) & 0 & 0 & 3 \\
-    min-if & 1 & 0 & 0 \\
-    percentual-diff & 5 & 1 & 0 \\
-    product & 2 & 0 & 3 \\
-    project & 1 & 0 & 0 \\
-    rank & 1 & 0 & 0 \\
-    series & 2 & 1 & 0 \\
-    sum (col) & 5 & 0 & 10 \\
-    sum (row) & 2 & 2 & 3 \\
-    sum-if & 2 & 0 & 9 \\
-    sum-product & 0 & 0 & 2 \\
-  \end{tabular}
-  \caption{Constraint occurrence by category}
-  \label{tab:constraint_by_category}
-\end{table}
-
 \end{document}

### 2016-12-19 Paper/response_letter/response_letter.tex

--- 2016-12-05 Paper/response_letter/response_letter.tex
@@
 \documentclass{article}
+\usepackage[usenames,dvipsnames,table]{xcolor}
 \usepackage{enumerate}
@@
 
-\paragraph{Sensitivity to the noise}
+\paragraph{Reviewer \#2, point B:How the data was obtained and its detailed overview}
+We have extended the description of the gathered collection of spreadsheets in the paper (Section 5.3), also we provided an extensive technical description of the dataset in the corresponding github repository: \url{https://github.com/SergeyParamonov/tacle} and added a table indicating present constraints by category into the Appendix together with description of spreadsheet sources.
+Let us present the overview here as well: there are three main sources for spreadsheets we have collected.
+\begin{itemize}
+  \item After identification of popular Excel functions (the MS Office web page has an overview of popular functions), we have searched for online tutorials about these functions and collected them under the category \textbf{Tutorials} (a link to each webpage found and used is provided in the accompanying github repository).
+  \item We have collected the exercises under the category \textbf{Exercises} from the introductory Excel book \cite{excel_book} that focused on popular Excel functions.
+  \item We have collected under the category \textbf{Data} publicly available economic and crime reporting data from the U.S.
+Bureau of Economic Analysis (BEA) and U.S.
+FBI Uniform Crime Reporting (UCR) Program (the links are also provided in the accompanying github repository).
+\end{itemize}
+
+The overview of constraint distributions per category is presented in Table \ref{tab:constraint_by_category}.
+\begin{table}
+  \centering
+  \rowcolors{2}{gray!25}{white}
+  \caption{Constraint occurrence by category}
+  \label{tab:constraint_by_category}
+  \begin{tabular}{l | c c c }
+    \textbf{Constraint} & \textbf{Exercises} & \textbf{Data} & \textbf{Tutorials}  \\\hline
+    average (col) & 0 & 0 & 2 \\
+    average (row) & 0 & 1 & 0 \\
+    average-if    & 1 & 0 & 0 \\
+    count (col) & 0 & 0 & 1 \\
+    count-if & 1 & 0 & 0 \\
+    difference & 2 & 1 & 0 \\
+    equal & 0 & 0 & 10 \\
+    foreign-product & 1 & 0 & 0 \\
+    fuzzy-lookup & 2 & 0 & 1 \\
+    lookup & 1 & 0 & 0 \\
+    max (col) & 0 & 0 & 2 \\
+    max-if & 1 & 0 & 0 \\
+    min (col) & 0 & 0 & 3 \\
+    min-if & 1 & 0 & 0 \\
+    percentual-diff & 5 & 1 & 0 \\
+    product & 2 & 0 & 3 \\
+    project & 1 & 0 & 0 \\
+    rank & 1 & 0 & 0 \\
+    series & 2 & 1 & 0 \\
+    sum (col) & 5 & 0 & 10 \\
+    sum (row) & 2 & 2 & 3 \\
+    sum-if & 2 & 0 & 9 \\
+    sum-product & 0 & 0 & 2 \\
+  \end{tabular}
+\end{table}
+
+\sergey{Samuel, can you have a look at what I wrote in the appendix and in the README and expand it?}
+
+\sergey{Also, they wanted some real examples of formulas from the dataset in Section 6}
+
+\textbf{Reviwer \#2, point C}:  \textit{According to my understanding, the proposed method exploits possible dependencies and redundancies that may exist between the different constraints.
+However, it is unclear how such dependencies/redundancies are found.
+Are they assumed to be given as input?
+Are they detected by the algorithm?
+What would be the performance (runtime and precision/recall) of the algorithm if no dependencies/redundancies exist?}
+
+\sergey{Samuel? I guess, we have to say that it should be specified in the constraint itself, on what it depends, kind of given}~
+
+Also, we have evaluated the influence on the runtime of the presence of the dependent graph by disabling it in the system and comparing the results on the datasets and synthetic data as presented in \sergey{Samuel, could you put a ref to where the baseline experiment is?}.
+There is no difference in the precision/recall, since the dependency graph only allows to avoid checking constraints that are guaranteed to fail, i.e., without the dependency graph the system would have to check the constraints that will fail to establish that indeed they do not hold.
+
+\paragraph{Review \#3, major drawback of the method: sensitivity to the noise}
 In the field of Inductive Logic Programming and Constraint Learning, there has been a line of work devoted to noise handling. Let us briefly overview the key methods applicable in our case:
@@
 
-\paragraph{How the data was obtained and its detailed overview}
-We have extended the description of the gathered collection of spreadsheets in the paper (Section 5.3), also we provided an extensive technical description of the dataset in the corresponding github repository: \url{https://github.com/sergeyparamonov/tacle}.
-\sergey{Samuel, can you make a detailed overview of the dataset in github https://github.com/sergeyparamonov/tacle}?
+\paragraph{Reviewer \#3, (non)-contiguous blocks} \textit{
+The requirement that blocks are contiguous makes sense, but also feels like a severe limitation. Please explain more clearly why this is (not) the case.}
+\sergey{Samuel, we have a short discussion that it breaks everything, right? Like you cannot prune, for example and also that skip semantics kind of imitates non-contiguous blocks}
 
-\sergey{Also, they wanted some real examples of formulas from the dataset in Section 6}
-
+\bibliography{references}
 \end{document}

### 2017-05-17 Code/src/engine/internal.py
Applied baseline bug-fix

@@ -585,7 +585,7 @@ def rank_data(a):
 def equal(x, y, scale=False):
     if x is None or y is None:
         return x is y
-    if isinstance(x, float) or isinstance(y, float):
+    if isinstance(x, (float, int)) and isinstance(y, (float, int)) and (isinstance(x, float) or isinstance(y, float)):
         delta = pow(10, -10)
         if scale:
             n_digits = min(precision_and_scale(x)[1], precision_and_scale(y)[1])

### 2017-05-17 Code/src/accuracy.py
Fix in accuracy file

@@ -68,7 +68,7 @@ def is_excel_constraint(c: Constraint):
     return c in excel_constraints()
 
 
-constraint_map = {c.name: c for c in workflow.constraint_list}
+constraint_map = {c.name: c for c in workflow.get_constraint_list()}
 
 
 def print_constraints(prefix, constraints):
